### 2018-01-10 tactictoe_jv.tex
first commit

\documentclass[runningheads,a4paper,draft]{svjour3}

%\usepackage{etex}
%packages indispensables
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{todonotes}
\newcommand{\todoi}[1]{\todo[inline]{#1}}

%\usepackage{graphicx}
%packages utiles
\usepackage{alltt} %program code
%\usepackage{enumerate}
%\usepackage{amssymb} %lettres mathématiques
%\let\oldemptyset\emptyset
%let\emptyset\varnothing
%\usepackage{amsthm}
\usepackage{amsfonts}
%\usepackage{bussproofs} %derivation 
%\usepackage{hyperref} %to write path.
%\usepackage{color} % colouring text
%\usepackage{tabularx} % table
%\usepackage{fancyvrb}
%\usepackage[toc,page]{appendix}
%\usepackage{cite}
\usepackage{booktabs}

\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{pgfplots}
%\usepackage{float}
%\restylefloat{table}
\usepackage{xcolor}
\usepackage{url}
\usepackage{subfigure}

\usepackage{xspace}
\usepackage{amsmath}
\DeclareMathOperator*{\argmin}{\arg\!\max}
\DeclareMathOperator*{\argmax}{\arg\!\max}

\def\holfour{\textsf{HOL4}\xspace}
\def\isabelle{\textsf{Isabelle}\xspace}
\def\hollight{\textsf{HOL Light}\xspace}
\def\mizar{\textsf{Mizar}\xspace}
\def\coq{\textsf{Coq}\xspace}
\def\ocaml{\textsf{OCaml}\xspace}
\def\vampire{\textsf{Vampire}\xspace}
\def\eprover{\textsf{E-prover}\xspace}
\def\zthree{\textsf{z3}\xspace}

\def\sml{\textsf{SML}\xspace}
\def\polyml{\textsf{Poly/ML}\xspace}
\def\holyhammer{\textsf{HOL(y)Hammer}\xspace}
\def\sledgehammer{\textsf{Sledgehammer}\xspace}
\def\mizAR{\textsf{MizAR}\xspace}

\def\metis{\textsf{Metis}\xspace}
\def\leancop{\textsf{leanCoP}\xspace}
\def\tactictoe{\textsf{TacticToe}\xspace}
\newcommand{\bq}[1]{\textbackslash"{#1}\textbackslash"}
\newcommand{\ra}[1]{\renewcommand{\arraystretch}{#1}}
%\theoremstyle{remark}
%\newtheorem{example}{Example} 
%\newtheorem{remark}{Remark} 
%\newtheorem{definition}{Definition} 
\clubpenalty=10000
\widowpenalty=10000

\usepackage{fancyvrb}
\usepackage{tikz} % graphics
% graphs
\usetikzlibrary{arrows,shapes}
\usetikzlibrary{trees,positioning,fit}
\tikzstyle{block} = [rectangle, draw,% fill=blue!20,
    text width=5em, text centered, rounded corners, minimum height=4em]
\tikzstyle{line}=[draw]
\tikzstyle{cloud} = [draw, ellipse,%fill=red!20,
    node distance=3cm,
    minimum height=2em]

% graphs
\usetikzlibrary{arrows,shapes,calc}
\usetikzlibrary{trees,positioning,fit}
\tikzstyle{arrow}=[draw,-to,thick]
\tikzstyle{bluearrow}=[draw,-to,thick,blue]
\tikzstyle{tcircle} = [circle, draw, fill=white]
\tikzstyle{tsquare} = [rectangle, draw, fill=white]
\tikzstyle{bcircle} = [circle, draw, blue, fill=blue]
\tikzstyle{gcircle} = [circle, draw, mygreen, fill=mygreen]

%\title{TacticToe: A theorem prover trained from formal proofs}
%\title{TacticToe: Smart guidance in tactical proof search}
%\title{TacticToe: Proof  by Tactic Selection}
\title{TacticToe: Learning the Game of Tactics}
%\title{TacticToe: Learning the Tactic Game}
%\title{TacticToe: Automatic tactic-level proof generation}

\author{\mbox{Thibault Gauthier} \and \mbox{Cezary Kaliszyk} \and \mbox{Josef 
Urban} \and \mbox{Ramana Kumar} \and \mbox{Michael Norrish}}

\authorrunning{Gauthier et al.}
%\titlerunning{TacticToe: Learning to Reason with HOL4 Tactics}

\institute{Cezary Kaliszyk \at
Department of Computer Science, University of Innsbruck,
%Technikerstr. 21a/2,
Innsbruck, Austria\\ \url{bla@bla}
\and
Karol Pak \at Institute of Informatics, University of Bialystok, Bialystok, Poland\\\url{bla@bla}}

%\institute{
%  University of Innsbruck\\
%  \email{\{thibault.gauthier,cezary.kaliszyk\}@uibk.ac.at}
%\and
%  Czech Technical University, Prague.\\
%  \email{josef.urban@gmail.com}\\
%}

\usepackage[final]{listings}
%\usepackage[scaled=.95]{couriers}
\lstdefinelanguage{SML}%
{columns=fullflexible,
  keywords={THEN,val,open,store_thm,fun,fn},%
  frame=none,
  sensitive=true,
  keywordstyle=\fontfamily{lmss}\footnotesize\selectfont,%
  basicstyle=\fontfamily{pcr}\selectfont,%
  stringstyle=\tt,
  morestring=[b]",
  literate={\ }{$\mkern6mu$}1%
   {=}{{\tt\raisebox{-.15mm}{=}}}1%
   {[}{{\tt\raisebox{-.15mm}{[}}}1%
   {]}{{\tt\raisebox{-.15mm}{]}}}1%
   {->}{{$\rightarrow$}}1%
   {=>}{{$\Rightarrow$}}1%
   {'a}{{$\alpha$}}1%
   {'b}{{$\beta$}}1%
   {ldots}{$\ldots$}1%
   {!}{$\forall$}1%
   {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
  }

\lstdefinelanguage{SMLSmall}%
{columns=fullflexible,
  keywords={THEN,val,open,store_thm,fun,fn},%
  frame=none,
  sensitive=true,
  keywordstyle=\fontfamily{lmss}\scriptsize\selectfont,%
  basicstyle=\fontfamily{pcr}\small\selectfont,%
  stringstyle=\tt,
  morestring=[b]",
  literate={\ }{$\mkern6mu$}1%
   {=}{{\tt\raisebox{-.15mm}{=}}}1%
   {[}{{\tt\raisebox{-.15mm}{[}}}1%
   {]}{{\tt\raisebox{-.15mm}{]}}}1%
   {->}{{$\rightarrow$}}1%
   {=>}{{$\Rightarrow$}}1%
   {'a}{{$\alpha$}}1%
   {'b}{{$\beta$}}1%
   {ldots}{$\ldots$}1%
   {!}{$\forall$}1%
   {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
  }
\lstset{language=SML}
\begin{document}
\maketitle

\begin{abstract}
Formalization of mathematical statements in interactive theorem provers (ITPs) 
relies nowadays on many methods to automate the construction of a proof.
Some of this method are general purpose and with enough time can prove any 
theorems. Yet, in a formalized proof script 
a mathematician take advantages of a combination of specialized rules, 
theory-based strategy and general ones. All of them are implemented as tactics 
in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
strategies. Sometimes, a strategy can be specialized for only one theory as in 
SMT solving. But it is difficult for an automated theorem prover to rely on 
rules tailored for a particular case. 
Indeed, this would require implementing many specific rules and more 
importantly figuring out in which cases they can be applied.

We bridge this gap by implementing our tactical prover TacticToe on top 
of the ITP HOL4. 
After learning in which situations tactics are effective from 
tactic-level proof and what kind of formulas are the results of effective 
steps, we implement a Monte Carlo tree search algorithm that explores the most 
promising tactic-level proof paths. Experiments on the \holfour standard 
library shows that TacticToe is able to re-prove 59\% of 7088 theorems in less 
than 10 seconds where as the ATP Eprover with auto-schedule
solves 29\%.

%on their first-order translations?

%
%Techniques combining machine learning with translation to automated
%reasoning have recently become an important component of formal proof 
%assistants. 
%Such ``hammer'' techniques complement traditional proof
%assistant automation as implemented by tactics and decision
%procedures. n this paper we present a unified proof assistant
%automation approach which attempts to automate the selection of
%appropriate tactics and tactic-sequences combined with an optimized
%small-scale hammering approach.
%
%We implement the technique as a tactic-level automation for HOL4:
%TacticToe. It implements a modified A*-algorithm directly in HOL4 that
%explores different tactic-level proof paths, guiding their selection
%by learning from a large number of previous tactic-level
%proofs. Unlike the existing hammer methods, TacticToe avoids
%translation to FOL, working directly on the HOL level. By combining tactic 
%prediction and premise selection, TacticToe is able to re-prove 39\% of 7902 
%HOL4 theorems in 
%%less than
%5 seconds whereas the best single HOL(y)Hammer strategy solves 32\% in the 
%same 
%amount of time.

\end{abstract}

\section{Introduction}
%\begin{example} Proof automatically generated by \tactictoe for the user given 
%goal
%\small
%\begin{center}
%
%
%\begin{BVerbatim}[commandchars=\\\{\}]
%Goal: ``\(\forall\)l. FOLDL (\(\lambda\)xs x. SNOC x xs) [] l = l``
%Proof:
%  SNOC_INDUCT_TAC THENL
%  [ REWRITE_TAC [APPEND_NIL, FOLDL],
%    ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC]
%      THEN CONV_TAC (DEPTH_CONV BETA_CONV)
%      THEN ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC] ]
%\end{BVerbatim}
%
%
%\end{center}
%
%\end{example}
%

\todo{These 2 paragraphs are copied}
Many of the state-of-the-art interactive theorem provers (ITPs) such as 
  \holfour~\cite{hol4}, \hollight~\cite{Harrison09hollight}, 
  \isabelle~\cite{isabelle} 
  and \coq~\cite{coq-book}
  provide high-level parameterizable tactics for constructing the
  proofs.  Such tactics typically analyze the current goal state and
  assumptions, apply nontrivial proof transformations, which get
  expanded into possibly many basic kernel-level inferences or significant 
  parts of the proof term.
%\marginpar{not Coq}
  In this
  work we develop a tactic-level automation procedure for the \holfour ITP
  which guides selection of the tactics by learning from previous
  proofs.  Instead of relying on translation to first-order automated
  theorem provers (ATPs) as done by the hammer 
  systems~\cite{hammers4qed,tgck-cpp15}, the technique
%\marginpar{too many cites?}
%BlanchetteGKKU16,holyhammer,mizAR40
  directly searches for sequences of tactic applications that lead to
  the ITP proof, thus avoiding the translation and
  proof-reconstruction phases needed by the hammers.

  To do this, we \emph{extract and record} tactic invocations from the ITP 
  proofs (Section~\ref{sec:recording}) and
  \emph{build efficient machine learning classifiers} based on such training
  examples (Section~\ref{s:prediction}).  The learned data serves as a 
  guidance for a \emph{Monte Carlo tree search} that explores the different 
  proof paths (Section~\ref{sec:proofsearch}). The result, if
  successful, is a certified human-level proof composed of \holfour
  tactics.  The system is evaluated on a large set of theorems originating
  from \holfour (Section~\ref{s:experiments}), and we show that the performance 
  of the single best \tactictoe strategy exceeds the performance of a hammer 
  system used with a single strategy and a single efficient external prover.
%Thanks to a very different proof-search procedure, \tactictoe also
%proves many theorems whose proofs are not found by the hammering
%approach.

\subsection{Contributions}
This project is an extension of the work described in \cite{tgckju-lpar17}. We 
explain here the additional contributions and their effect across our system:

\begin{itemize}
\item Proof recording at the tactic level is made more precise. Pattern 
matching constructions, opening of submodules are supported. The different 
recording steps (parsing, unfolding and replaying) are regrouped into a single 
\sml function. Intermediate files storing the result of each step are not 
necessary anymore. So, our recording is more complete and fully 
automatized.
\item Existing tactic orthogonalization on recorded data is made stronger by 
also including the internal ATP \metis in the orthogonalization 
competition. Because we manage a database of theorems for argument 
prediction, we apply here a pruning step to remove easy theorems.
\item Two types of arguments (list of theorems and terms) can be predicted 
independently. Combinatorial explosion in the number of tactics 
is prevented through the introduction of abstracted tactics. As desired, their 
instantiations is dependent on the target goal.
\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
algorithm. MCTS gives a more balanced feedback to guide proof search by 
comparing subtrees instead of leafs. (todo: A first successful evaluation 
function.)
\item The internal ATP \metis used during our proof search is complemented by 
calls to \holyhammer. For efficiency reasons, those calls are run in parallel 
and asynchronously.
\item First evaluation of \tactictoe and \holyhammer success rates relative to 
the length of the original proofs.
\item User interaction is enhanced by the minimization and prettification 
applied on the generated proof.
\end{itemize}

\section{Related Work}\todo{reorganise}
There are several essential components of our work that are comparable to 
previous approaches: tactic-level proof recording, tactic 
selection through machine learning techniques and automatic tactic-based proof 
search. Our work is also related to previous approaches that use machine 
learning to select premises for the ATP systems and guide ATP proof search 
internally.

For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15} 
can transform a packed tactical proof into a series of interactive tactic 
calls. Its principal application 
was so far refactoring the library and teaching common proof techniques to new 
ITP users. In our work, the splitting of a proof into a sequence of tactics is 
essential for the
tactic recording procedure, used to train our tactic prediction module.

The system \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14} 
groups related proofs thanks to its clustering 
algorithms. It allows \coq users to inspire themselves from similar proofs and 
notice 
duplicated proofs. Our predictions comes from a much more detailed description 
of the open goal.
However, we simply create a single label for each tactic call whereas each of 
its
arguments is treated independently in \textsf{ML4PG}. 
Our choice is motivated by the k-NN algorithm already used in
\holyhammer for the selection of theorems.
%, which can be easly adapted to the selection of tactics but would be improved 
%by a 
%version that can learn to predict tactics and their arguments independently.

\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to 
generate
proof scripts from previous \coq proof examples.
Its strength lies in its ability to produce likely sequences 
of tactics for solving domain specific goals. It operates by creating a model 
for common sequences of tactics for a specific library.
This means that in order to propose the following tactic, only the previously 
called tactics
are considered.
%intuition tells us that it should not be the main factor
Our algorithm, on the other hand, relies mainly on the characteristics of the 
current goal 
to decide
which tactics to apply next. In this way, our learning mechanism has to 
rediscover why each 
tactic was applied for the current subgoals. It may lack some useful bias for 
common sequences 
of tactics, but is more reactive to subtle changes. Indeed, it can be trained 
on a large library and only tactics relevant to the current subgoal will be 
selected. 
Concerning the proof search, \textsf{SEPIA}'s %simple 
breadth-first search is replaced by an A*-algorithm which allows for heuristic 
guidance in 
the exploration of the search tree.
Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
theorems) of the 
\coq library demonstrating that it globally outperforms individual \coq 
tactics. In contrast, we demonstrate the competitiveness of our system against 
the successful general-purpose hammers on the \holfour standard library (7902 
theorems).

Machine learning has also been used to advise the best library lemmas for new 
ITP goals.
This can be done either in an interactive way, when the user completes the 
proof based on the recommended lemmas, as in the \textsc{Mizar Proof 
Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma 
selection is handed over to the ATP component of a \emph{hammer} 
system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.

Internal learning-based selection of tactical steps inside an ITP is analogous 
to internal learning-based selection of clausal steps inside ATPs such as 
\textsc{MaLeCoP}~\cite{malecop} and \textsc{FEMaLeCoP}~\cite{femalecop}. These 
systems
use the naive Bayes classifier to  select clauses for the extension steps in
tableaux proof search based on many previous proofs. Satallax~\cite{Brown2012a} 
can guide its
search internally~\cite{mllax} using a command classifier, which can estimate 
the priority of the 11 kinds of
commands in the priority queue based on positive and negative examples.
There are several essential components of our work that are comparable to 
previous approaches: tactic-level proof recording, tactic 
selection through machine learning techniques and automatic tactic-based proof 
search. Our work is also related to previous approaches that use machine 
learning to select premises for the ATP systems and guide ATP proof search 
internally.

For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15} 
can transform a packed tactical proof into a series of interactive tactic 
calls. Its principal application 
was so far refactoring the library and teaching common proof techniques to new 
ITP users. In our work, the splitting of a proof into a sequence of tactics is 
essential for the
tactic recording procedure, used to train our tactic prediction mechanism.

The system \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14} 
groups related proofs thanks to its clustering 
algorithms. It allows \coq users to inspire themselves from similar proofs and 
notice 
duplicated proofs. Our predictions comes from a much more detailed description 
of the open goal.
However, we simply create a single label for each tactic call whereas each of 
its
arguments is treated independently in \textsf{ML4PG}. 
Our choice is motivated by the k-NN algorithm already used in
\holyhammer for the selection of theorems.
%, which can be easly adapted to the selection of tactics but would be improved 
%by a 
%version that can learn to predict tactics and their arguments independently.

\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to 
generate
proof scripts from previous \coq proof examples.
Its strength lies in its ability to produce likely sequences 
of tactics for solving domain specific goals. It operates by creating a model 
for common sequences of tactics for a specific library.
This means that in order to propose the following tactic, only the previously 
called tactics
are considered.
%intuition tells us that it should not be the main factor
Our algorithm, on the other hand, relies mainly on the characteristics of the 
current goal 
to decide
which tactics to apply next. In this way, our learning mechanism has to 
rediscover why each 
tactic was applied for the current subgoals. It may lack some useful bias for 
common sequences 
of tactics, but is more reactive to subtle changes. Indeed, it can be trained 
on a large library and only tactics relevant to the current subgoal will be 
selected. 
Concerning the proof search, \textsf{SEPIA}'s %simple 
breadth-first search is replaced by MCTS which allows for heuristic 
guidance in 
the exploration of the search tree.
Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
theorems) of the 
\coq library demonstrating that it globally outperforms individual \coq 
tactics. In contrast, we demonstrate the competitiveness of our system against 
the successful general-purpose hammers on the \holfour standard library.
A recent improvement on this system~\cite{} includes the possibility to replace 
theorems in tactics by ones belonging to the same cluster, independently of the 
current goal. We have the opposite approach, we forget about the original 
theorems in the tactic creating an more general abstract tactic
and only use the goal to guide predictions. Finding a way to combine those two 
approaches could be the key to stronger theorem predictions.

A similar effort for stronger automation is performed in \isabelle. Nagashima 
developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the user to 
program with regular tactics and combine them with tools such as 
\sledgehammer~\cite{sledgehammer10} or Nitpick~\cite{Nitpick10}. 
Selection of tactics from recorded proofs has yet to be implemented and now it 
is up to come up with a good meta-tactic.

%In \holfour, 
%a tactic is a just a \sml function of a special type and \holyhammer can 
%easily 
%be modified to be a tactic. So this meta-level is not 
%necessary, all programming can occur at the tactic level and even \tactictoe  
%is a tactic.

%* Theory Exploration (Moa Johanson+, ITP 2017)  I think that is more related 
%to conjeturing and do not think it fits here.

Machine learning has also been used to advise the best library lemmas for new 
ITP goals.
This can be done either in an interactive way, when the user completes the 
proof based on the recommended lemmas, as in the \textsc{Mizar Proof 
Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma 
selection is handed over to the ATP component of a \emph{hammer} 
system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.

Internal learning-based selection of tactical steps inside an ITP is analogous 
to internal learning-based selection of clausal steps inside ATPs such as 
\textsc{MaLeCoP}~\cite{malecop} and \textsc{FEMaLeCoP}~\cite{femalecop}. These 
systems
use the naive Bayes classifier to  select clauses for the extension steps in
tableaux proof search based on many previous proofs. Satallax~\cite{Brown2012a} 
can guide its
search internally~\cite{mllax} using a command classifier, which can estimate 
the priority of the 11 kinds of
commands in the priority queue based on positive and negative examples.

A first successful use of orthogonalization techniques for theorems appeared in
a work on lemma mining~\cite{ckju-jsc15} in \hollight. Theorems were mined from 
proofs recorded at the kernel level. Because the number of recorded lemmas was 
in millions, page ranking techniques was used as a filter to extract ``useful'' 
lemmas.

\section{Proof Recording}\label{sec:recording}
Recording proofs in an LCF-style proof assistant can be done at different levels.
In \holfour all existing approaches relied on modifying the kernel. This was used
either to export the primitive inference steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable for our
purpose of learning proving strategies at the intermediate tactic level. We therefore
discuss recording proofs in an LCF-style proof assistant, with the focus on \holfour
in this section.

% Because our aim is to 
% learn how to prove from existing examples, we record the proofs at 
% an intermediate tactic level closest to the one presented to 
% a user. Since human developers build almost all their proofs from tactics, 
% we believe this is an appropriate level to apply machine learning on.

%More information is contained in tactical scripts than 
%in top-level dependencies on which was based machine-learning guidance in 
%interactive theorem proving so far. Learning from atomic inferences would 
%require better predictions and abstractions of rules to handle the larger 
%search space.

%\subsection{Recording Tactic Calls}
%We first discuss parsing the proofs and identifying 
%individual tactic calls. We next show how tactics calls are recorded together 
%with their corresponding goals using a modified proof script.
%Finally, the recorded data is organized as feature vectors with tactics as 
%labels and characteristics of their associated goals as features. 
%These feature vectors constitute the training set for our selection algorithm.

Rather than to rely on the underlying programming language, we parse the
original theory file containing proof scripts into an
abstract syntax tree (AST). The AST is then transformed into a globalized AST
where every top-level value is made independent from
the context. \todo{some reason why this is needed? nice recording level} We 
split the proofs into a series of tactic units and wrap them 
with a recording function.
By replaying the proofs produced by the globalized AST, the wrapper can
record the globalized tactic string with its input goal as training examples in
a database.
The combined effect of these steps on a \holfour proof will be illustrated on Example~\ref{ex:running}.

\begin{example}\label{ex:running}(Running example)
\small
\begin{lstlisting}[language=SMLSmall,frame=tb]
open boolLib Tactic Prim_rec Rewrite
ldots
val LIST_INDUCT_TAC = INDUCT_THEN list_INDUCT ASSUME_TAC
ldots
val MAP_APPEND = store_thm("MAP_APPEND",
  ``!(f:'a->'b) l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f l2)``,
  STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND])
\end{lstlisting}
\end{example}

\subsection{Minimal \sml Code Trees}
\holfour proofs are organized in theory files.
Each theory file contains definitions and statements of theorems together with
their proofs. Those theories and the contained proofs
are written
in the functional programming language \sml. To automate
proof search strategies a user can implement programming language functions
that correspond to more involved proofs.
Functions that correspond to backward proofs, called tactics, are combined with each other with the help of 
tacticals (functions that takes
tactics as arguments). Such a combination of tactics will be called a \emph{proof script} when it
proves a theorem stated in a theory.

In order to apply machine learning to the existing proof scripts, we will need to
parse and modify them, which involves parsing and modification of \sml code.
Acquiring access to the string representation of a tactic (code for the tactic) 
allows us generating
proof scripts that can be inspected by a human.
As some \holfour theories are built in separate sessions, it is not possible
to transfer \sml values directly, but instead tactic code needs to transferred instead.
To do this, the transferred information will be saved externally.

Using the standard types $string$ and $list$ and their type constructors,
we define a minimal code tree by the following AST:
\begin{align*}
M &=_{def} string \\
I &=_{def} string\ \vert\ Module(M,I)\\
D &=_{def} Open(M)\ \vert\  Val(I,E) \\
E &=_{def} Code(I\ list)\ \vert\   Let (D,E)
\end{align*}

In this AST, the constructors $Let$, $Val$ and $Open$ correspond to
the reserved \sml tokens \texttt{let}, \texttt{val} and \texttt{open} 
respectively. $Module(M,I)$ refers to the \sml token \texttt{M.I} where 
\texttt{M} is a module (also called structure in \sml) and \texttt{I} is an 
identifier. $D$ is a declaration, $E$ an expression. Inside a 
declaration $val(I,E)$, $E$ is a definition for $I$.
To construct such an AST for a theory file, we rely on a custom parser. It transform a source file into a 
minimal code tree taking into account the scope of each constructor.
An additional reason to refrain from the use of the interpreter in our case, is that
the \polyml compiler needs to execute the code preceding an expression 
before parsing it, which is significantly slower than a custom parser.

\subsection{Globalization}
To transfer \sml values as code across theories in a conservative manner, we 
need to 
globalize \sml identifiers. To this end, we replace each identifier by some 
\sml code containing all the information to obtain the value of this identifier 
without knowing the context. The result is a standalone \sml code executable in 
any \holfour theories which will be interpreted in the same way in the absence 
of side effects. Two techniques are helpful to produce such code. The first one 
is to prefix global identifiers by the module (and sub-modules) in which they 
were declared. And the second one is to replace local identifiers 
by their definitions. Prefixing and 
unfolding are applied recursively in 
the definitions.

For a more formal explanation, we define a target AST build upon the original 
one, where each identifer $I$ is replace by its globalized expression $E'$:

\begin{align*}
M' &=_{def} string \\
I' &=_{def} string\ \vert\ Module(M',I')\\
D' & =_{def} Open(M')\ \vert\  Val (I',E') \\
E' &=_{def} Code(E'\ list)\ \vert\ Let (D',E') 
\end{align*}

Then, we express two mutually recursive functions $f_d$ and 
$f_e$ that globalize the original representation. The 
function $f_d$ modifies declaration and $f_e$ modifies expressions. Since a 
\sml program is a list of declarations, we will be able transform it by 
iteratively applying $f_d$. The stack 
$\mathbb{S}$ keep track of each previous declaration: the identifier that is 
declared and their globalized definition. To save a declaration to 
the stack, we use the $push$ 
function. Opening a module has the effect of pushing all identifiers with 
their associated prefixed version to the stack. The scope of an identifier is 
automatically handled by construction of the AST. To find the globalized 
version of an identifier $i$, we use the $lookup$ function on a stack $s$, 
which returns the globalization associated with $i$ appearing highest in $s$. 
We then build $f_e$ and $f_d$ in the following way:
\vspace{5mm}

\begin{math}
\begin{array}{lcl}
f_d:& \mathbb{S} \times D &\rightarrow\ \ \mathbb{S} \times D'\\
& (s, Open(m)) &\mapsto\ \ (push(m,s),\ Open(m)) \\
& (s, Val(i,e)) &\mapsto\ \ (push ((i,\hat{e}),s),\ Val(i, \hat{e})) 
\mbox{ where } \hat{e} = \pi_2^2 (f_e (s,e))
\end{array}
\end{math}

\vspace{5mm}

\begin{math}
\begin{array}{lcl}
f_e:& \mathbb{S} \times E &\rightarrow\ \ \mathbb{S} \times E'\\
& (s, List(l))  &\mapsto\ \ (s,\ lookup(s,l))\\
& (s, Let (d,e)) &\mapsto\ \ (\hat{s}, f_e (\hat{s}, e))
\mbox{ where } \hat{s} = \pi_1^2 (f_d (s,d))
\end{array}
\end{math}

\vspace{5mm}

%\begin{example} Example
%\small
%\begin{alltt}
%val MAP_APPEND = store_thm ("MAP_APPEND", 
%  --`!(f:'a->'b).!l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f 
%l2)`--,
%    STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]);
%\end{alltt}
%\end{example}  
%
%\begin{example} Running example (extracted proof)
%\small
%\begin{alltt}
%STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
%\end{alltt}
%\end{example}

%The goal of our program is to transform the original program is to associate to
%each expr token.

%\subsubsection{Structures}
%Structures:
%
%Globalizing functions:
%examples:
%
%let val x = 5;  end;
%substructres

Thanks to the simplicity of the minimal AST, we are able to present our 
recursive unfolding algorithm in a concise and formal manner. But for practical 
purpose, the minimal AST is not sufficient. So, we extend the scope of our 
algorithm by enumerating common \sml constructions and explain 
independently how we globalize them. We first start by mentioning additional 
details about module and value declarations.

\paragraph{Modules}
Declaring a module is made through the keyword 
\texttt{open}. We can fetch 
the values declared in a module \texttt{m} by first erasing all elements from 
\texttt{PolyML.NameSpace} and then calling \texttt{open m}. Additionally, it 
gives us the information of whether a value is a constructor or an exception. 
This will prove to be particularly useful to handle pattern matching 
construction correctly. During the opening of a module, nested modules can be 
declared and push to the stack. In this way, we can globalize them by prefixing 
them by their parent modules when they appear in the program.

\paragraph{Values}
The standard ML language allows to declare multiple values in a single $let$ 
construction. In effect, this list of declarations is equivalent to nested 
declarations.

Because we do not want to unfold proofs into other proofs, we 
make an exception 
for values of type theorem and do not replace them by their definition.
Instead, we fetch theorems from the 
database by looking for its formula in the database of theorems.
Because this database is reconstructed from files saved on disk, a call to 
the function \texttt{DB.fetch} works across theories.

\paragraph{Functions}
To globalize a function identifier, we include 
its declaration in the globalized expression. 
Given a declaration \texttt{fun f v = e}, the globalized 
expression for the identifier \texttt{f} is \texttt{let fun f v = e' in f 
end}, where \texttt{e'} is the globalized version of \texttt{e}. The function 
\texttt{f} and its argument \texttt{v} are binders. So, we prevent our 
algorithm from trying to globalize them during the construction of \texttt{e'}.
This approach is slightly more verbose than using an 
anonymous function construction. But it is prettier as 
we regroup declarations at the top of the proof expression before returning it
to the user (see Section~\ref{sec:proofdisplay}).

\paragraph{Pattern matching}
A pattern is an \sml expression that 
contains \sml 
constructors (or exceptions in the case of \texttt{handle}) and binding 
variables. Our algorithm can recognize cases where pattern matching appears and 
globalize the expression appropriately.  Let \texttt{p} be a pattern 
and 
\texttt{e} be abd expression, the following constructions can 
occur: \texttt{val p = e}, \texttt{case \_ of p => e}, \texttt{fun f p = e}, 
\texttt{fn p => e}, \texttt{handle p => e}. A binding identifiers in \texttt{p}
is a token that is not a constructor (or exception). To produce the globalized 
version of each construction, we globalize constructors (or exception) in the 
pattern \texttt{p} and we protect binding variables in \texttt{e} to prevent 
their modification. When a value identifier \texttt{x} is declared in a pattern 
\texttt{p}, the globalized expression \texttt{e'} for \texttt{x} becomes 
\texttt{let val p = e' in x end}.

\paragraph{Infixity}
In \sml, it is possible to declare infixity rules for a string $s$ that 
is not prefixed by a module. The infixity applies only when an identifier 
exactly match $s$. Therefore, an identifier \texttt{x} could have been treated 
as infix in an original expression \texttt{E} but when 
we substitute it locally by its 
globalized version \texttt{e'}, \texttt{e'} is now considered with prefix 
infixity in \texttt{E} which makes \texttt{E} an invalid expression. Therefore, 
because we 
want to avoid global changes to \texttt{E}, we encapsulate \texttt{e'} by two 
newly 
defined 
infix guards to reproduce 
the original behaviour.
In total there are 20 possible infix settings \texttt{x} can have in \sml.
Those infix guards are defined to have the 
same 
precedence level $\texttt{P}\in[0,9]$ and same associativity 
$\texttt{A}\in[l,r]$ (right or left) as \texttt{x}.
Definitions for the infix guards depend on their associativity:

\begin{definition}(Code of infix guards)
\begin{lstlisting}
fun a infixlX_open f = fn x => f (a,x)  
fun g infixlX_close b = g b

fun f infixrX_close b = fn x => f (x,b)
fun a infixrX_open g = g a
\end{lstlisting}
\end{definition}

\begin{proof}(Does anyone want a proof that this functions are correctly 
reproducing the original function?)
\end{proof}

The previously decribed transformation affects each identifier in a theory 
file. However, when we 
create the recording theory file, only proof scripts are replaced by their 
globalized version and the rest of the code is left unchanged. The effect on 
the proof script contained in the running example is shown in 
Example~\ref{ex:global}.

\begin{example} (Globalization)
\begin{lstlisting}
Tactic.STRIP_TAC 
  infixl0_open boolLib.THEN infixl0_close
Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC
  infixl0_open boolLib.THEN infixl0_close
Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
\end{lstlisting}
\end{example}

%\sml expression delimiters ``([,]){}='' are left 
%unchanged by the globalization process.
%
% pattern matching
%By distinguishing which values are constructors in our Pattern matching in 
%\sml 
%is allowed inside a value 
%
%
%infix operators
%Our goal is to create a globalized version of each generalizable token.
%
%In this description we are omitting handling of "fun", "case of | =>", "fn =>"
%The only thing to consider is that var appearing as arguments of a function.
%
%infix constructor: 
%  Infix constructors cannot be replaced by their globalized version without 
%  confusing the parser.
%
%  
%
%  We can globalize an infix operartor locally in the following way;
%
%Original code: some\_code1 infix some\_code2
%
%the infix code does not carry the information on how to parse it.
%
%            infix\_open (infix\_code) infix\_close.
%
%where infix\_open and infix\_closed are declared to be the same arity with the 
%same associactivity rule as infix.
%
%Depending if the operator is left associative or right associative we give two 
%different definitions for infix\_open and infix\_close.
%
%
%
%constructions with bounded variables: fun (case of) (
%
%proof scripts (arguments of prove, store\_thm, maybe\_thm)
%
%watch list:
%
%stateful tactics

\subsection{Recording tactic calls}

We now restrict the definition of a proof script to being the third 
argument of a call to \texttt{store\_thm}. And we precisely define the 
modification applied to the theory file. A modified theory file is identical to 
the original theory file except every proof scripts \texttt{p} is
replaced by \texttt{Split "p'"} where \texttt{Split} is the splitting function 
to be defined and \texttt{"p'"} is the string representation of the globalized 
version of the proof script 
\texttt{p}. 
.
\begin{example}(Proof script transformation)
\begin{alltt}
original declaration: val id = store_thm (name, term, p)
modified declaration: val id = store_thm (name, term, Split "p'")
\end{alltt}
\end{example}

The intended effect of the splitting function is to detect and wrap \sml 
expressions of type tactics. Its aim is to is encapsulate each tactic 
unit $t$ within a recording function $R$. This function $R$ is designed to 
observe what goals $t$ receives at run time. Since there can be maybe nested 
level of tactics, (i.e 
tactics that contain tactics as subexpressions), we choose one that is
close to what a human understand to be a tactic unit. 
Most tacticals are infix, and we do not want to consider combination of 
tactics to be a tactic unit. For our wrapper \texttt{Split}, a tactic unit is a 
subterm of 
type tactics that do not contain a infix operator as its top symbol. Thanks to
globalized infixity, we just have to verify that the expression 
is of type tactic and that an infix guard is not a top symbol of the expression.
To determine that, we run the \polyml compiler on \texttt{"p'"} to give us a 
detailed code tree of the expression. Each tactic unit is then substituted 
by a wrapped version to produce the final proof script.

\begin{example} (Wrap)
\begin{alltt}\small
R "Tactic.STRIP_TAC"
  infixl0_open boolLib.THEN infixl0_close
R "Prim_rec.INDUCT_THEN (DB.fetch \bq{list}\bq{INDUCT}) Tactic.ASSUME_TAC"
  infixl0_open boolLib.THEN infixl0_close
R "Rewrite.ASM_REWRITE_TAC 
     [DB.fetch \bq{list} \bq{MAP}, DB.fetch \bq{list} \bq{APPEND}]"
\end{alltt}
\end{example}

During its execution, the recording function \texttt{R} (as coded in 
Definition~\ref{ex:record}) 
saves the tactic 
string and the goal, verifies that the tactic string as the same effect as the 
original tactic and apply it to continue the proof. To perform this task, we 
convert the tactic string back into a tactic via the function 
\texttt{read\_tactic}. This function will be used whenever we want to execute a 
tactic string. So in the next sections, we will not make a difference between a 
tactic and its code and call both of them tactics.

\begin{example}\label{ex:record} (Code of the recording function)
\begin{lstlisting}
fun R tactic_string goal = 
  (save (tactic_string, goal); (read_tactic tactic\_string) goal)
\end{lstlisting}
\end{example}

\subsection{Storing Training Data}
We describe here the format preferred to store and load the 
data recorded by the function \texttt{save}. 
During saving, we add fields to the input pair $(tactic,goal)$ to record 
precomputed information in order to make the prediction faster. The final 
feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
The time $time$ taken by $tactic$ is not currently used by our 
predictor. The list of goals $goal list$ is the output produced by $tactic$ on 
$goal$. And the features $features$ of the goal $goal$ are extracted and hashed 
as described in Section~\ref{sec:features}.

\begin{remark}
To simplify the description of our machine learning algorithms in future 
sections, we consider a feature vector to be a pair $(tactic,goal)$ and only 
explicitly rely on the 
other fields when necessary.
\end{remark}

We now discuss the type of file we want to store our feature vectors on and
how we plan to write and read \sml values. We start by taking 
inspiration from an existing implementation in \holfour.

To avoid the computational cost of re-proving every theorems at the begining of 
each session, theorem values in \holfour are stored.
They are encoded them in to a string by calling \texttt{Term.write\_raw}.
Conversely, the function \texttt{Term.read\_raw} transforms the string back 
into the original term.
To have theorems binded to sml identifiers, it is necessary to create a 
\sml file where those binders are declared. And for simplicity sake, all theory 
data was fed into this \sml file. However, it is
more efficient to store theory data into a text file and make a custom 
parser which can be run faster than the \polyml compiler.
Thanks to this insight, we have 
modified the way in which \holfour theories are saved to include only \sml 
bindings in a \sml file. All other theory data is now stored separately 
into a text file. Michael Norrish and Ramana Kumar contributed to integrated 
these changes in the current distribution of \holfour. After the changes, 
\holfour theories load three times faster in average.

Our \tactictoe data management system do not need to create new \sml
bindings for the user. So we can store all \sml values into a text file which 
can be loaded rapidly.
Storing time, tactic string and features is easy since they are respectively 
floats, strings and integers. And because a goal is composed of terms, we can 
re-use the \holfour theory data reading and writing functions 
for the two other fields.

%The previous step demonstrated how to extract training data from proofs. We 
%now 
%need to store this data on disk so that this training data is re-usable in 
%other theories. We gathered information about a global string representation 
%of 
%a tactic, the time taken to apply this tactic and the input goal and output 
%list of goals. For printing the tactic string and the time, we just make use 
%of 
%printing functions for those types  and 
%\texttt{Real.toString}. 
%However for the terms that forms the goals, it is not advisable to use the 
%function 
%\texttt{term\_to\_string} as it may not contain all the information necessary 
%to 
%reconstruct the same term. Instead we re-use the functions for storing terms 
%of 
%theorems in theories \texttt{Term.write\_raw} and and \texttt{Term.read\_raw}, 
%constructing sharing tables for types and terms.
%Contrary to \holfour theories that are saved into a \sml file, our data is 
%stored as text files. This allow us a much faster loading time, since we are 
%never calling the PolyML compiler. 
%Already exists in \holfour for recording theorems.
%Mention better export for HOL4.
%
%Side effects:
%By minimizing the size of the sml files, we manage to multiply by 3 the speed 
%at which theories are loaded in \holfour. 
%Mention Ramana Kumar and Michael Norrish.

Finally, we mention some limitations that comes with 
the need to record tactics as a string compared to being able to store a \sml 
value. Access to the name of the tactic enhances machine 
learning ability and user interaction. Yet, because some tactics 
depends on the state of \holfour, for example the state of the global 
simpset, a tactic is re-parsed before each proof attempt to be up-to-date.
As a result, a tactic may have different effect on goals than suggested by the 
database which could compromise the prediction accuracy. In practice, this 
should rarely happen. Indeed, most tactics are stateless and an update on a 
stateful tactic is usually improving it.

%By recording also the sml value it would also possible to eliminate those 
%drawbacks as long as you stay in the same session. This is typically the case 
%for the \hollight theorem prover but not in \holfour where every theory is 
%build in a different session loading previous theorems from disks.

\section{Prediction}\label{s:prediction}

The learning-based selection of relevant lemmas significantly improves the 
automation for hammers~\cite{BlanchetteGKKU16}. Therefore we propose to adapt 
one of the strongest hammer lemma selection 
methods: the modified distance-weighted \emph{$k$ nearest-neighbour} 
classifier~\cite{ckju-pxtp13,DudaniS76}.
In \tactictoe, we predict four different objects: theorems, terms, tactics and 
list of goals. We derive from tactic selection a prior policy for 
our search algorithm. Lists of goals (products of tactics) evaluate the success 
of the policy and adapt it accordingly. Term and theorems predictions are 
necessary to find the suitable arguments for abstracted tactics (see Section 
\ref{sec:synthesis}).

%Premise selection usually only prunes the initial set of formulas given to the 
%ATPs, which then try to solve the pruned problems on their own. 
%the goal of predicting useful lemmas is to reduce the search space,
% of the external ATPs called by a hammer,

%Mention explictly all the predictors that will be using this distance.
%tactics:
%theorems:
%terms:
%goal list: (product of tactics)
\begin{itemize}
\item tactic:\\
Given a goal $g$, the classifier selects a list of previously solved goals $l$
ordered by their similarity to $g$. Since each solved goal is associated 
with a tactic in our database, the ordered list of goals induces a ordered  
list of associated tactics. And tactics are tried in this order during the 
proof search. Therefore, to order the tactics correctly, it is crucial to 
estimate the similarity between goals with mathematically relevant features.
\item term:\\

\item theorem:\\

\item list of goals:\\

\end{itemize}

%We will next discuss 
%the extraction 
%of features from goals and the
%actual prediction of tactics. 
%Both have been integrated in the \sml proof search.

\subsection{Features for Terms and Goals}\label{sec:features}

We start by extracting the syntactic features that have been successfully used 
in premise selection from terms:
\begin{itemize}
\item names of constants, including the logical operators,
\item type constructors present in the types of constants and variables,
\item subterms with all variables replaced by a single place holder $V$.
\item names of the variables
\end{itemize}

We found that the names of variables present in a term are particularly 
important for predicting tactics such as 
case splitting on a variable 
(\texttt{Cases\_on var}) or instantiation of a variable 
(\texttt{SPEC\_TAC var term}).

When a formula is represented as a goal, we also add a tag to each 
feature to remember whether it was extracted from the assumptions or the 
conclusion of a goal.
%We additionally extract the following features:

%\begin{itemize}
%\item the top-level logical structure with atoms substituted by a single place 
%holder $A$ and all its substructures.
%\end{itemize}

%Determining the presence of top-level logical 
%operators (i.e 
%implication) is essential to assess if a "logical" tactic should be applied. 
%For example, 
%the presence of an implication may lead to the application of the tactic 
%\texttt{DISCH\_TAC} that moves the precondition to the assumptions. Top-level 
%logical structure gives a more detailed view of the relationship between those 
%logical components.

\subsection{Feature distance}\label{sec:predictions}

In all proofs, we record each tactic invocation and link (associate) the 
currently open goal with the tactic's name in our database.
Given a new open goal $g$, the \emph{score of a tactic $T$ wrt. $g$} is defined 
to be the score (similarity) of $T$'s 
associated goal which is most similar to $g$.
The idea is that tactics with high scores are more likely to solve the open 
goal $g$, 
since they were able to solve similar goals before.

We estimate the similarity (or co-distance) between an open goal $g_o$ and a 
previously recorded goal $g_p$ using their respective feature sets $f_o$ and 
$f_p$.
The similarity function $sim$ computed by the k-NN algorithm is analogous 
to the
one used in the premise selection task~\cite{ckju-pxtp13}. The main idea 
is to find the features shared by the two goals and estimate the rarity of 
those features calculated via the TF-IDF~\cite{Jones72astatistical} heuristics.
In a second co-distance $sim_2$, we additionally take into account 
the
total number of 
features to reduce the seemingly unfair advantage of big feature sets in the 
first scoring function.
\[sim_1 (f_o, f_p) = {\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{\tau_1}}}\]
\[sim_2 (f_o, f_p) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
f_p}{\text{tfidf}(f)^{\tau_1}}}}
{ln (e + card(f_o) + card(f_p))}
\]

\begin{remark}
We use $\tau_1 = 6$ because it was found optimal in previous experiments.
\end{remark}

In addition to predicting tactics, we also used these similarity function to 
predict theorems and terms. Suprisingly, the first similarity function $sim_1$
is best to predict tactics and theorems. It is probably because there exists 
monotonocity heuristic where if a tactic can be effective on a large goal, it 
will be also effective on smaller goals that contain subterms of the larger 
goal. The same heuristic can be applied for theorems, when a large theorem 
contains more information than a theorem which is a subformula of the larger 
theorem. Nevertheless, we want to be able to predict variables (which are small 
terms) during term argument prediction. So in this case, weight down big terms, 
as implemented by the $sim_2$ function, should have a beneficial impact.
The distance $sim_2$ is also used to evaluate list of goals for proof node 
selection in the Monte Carlo tree search(MCTS) algorithm described in 
Section~\ref{sec:proofsearch}.

%\[tactic\_score_2 (f_o,f_p) = \frac{tactic\_score_1 (f_o,f_p)}{(1+ ln\ (1 + 
%card\ f_o))} \]

%Moreover, we would like to compare the distance of a recorded goal to 
%different goals opened at different moment of the search. That is why we 
%normalize the scores by dividing them by the similarity of the open goal with 
%itself. As a result, every score will lie in the interval $[0, 1]$ where 1 is 
%the highest degree of similarity (i.e. the shortest distance). We respectively 
%refer to those normalized scores later as $tactic\_norm\_score_1$ and 
%$tactic\_norm\_score_2$.

\subsection{Dependencies}\label{sec:dependencies}
We first rank feature vectors by their distance to the conjecture. But better 
ranked vectors only contains symbols from 
the conjecture and may not be suitable for later stages of the proof where the 
goals may have derived significantly from the conjecture. To solve this 
problem, we reorder the list so that descendants (see 
Definition~\ref{def:descendant}) of vectors inherits the similarity of their 
parent.

\begin{definition}(Descendant of a feature vector)\label{def:descendant}
Let $\mathfrak{F}$ be a set of recorded feature vectors. 
The set of all descendants $D_\infty$ of a feature vector $f$ is 
inductively defined by:

\begin{align*}
D_0 &=_{def} \lbrace f \rbrace \\
D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists 
(\hat{t},\hat{g}) \in D_n.\ g \in \hat{t}(\hat{g}) \rbrace  \\
D_\infty &=_{def} \bigcup_{i \in \mathbb{N}^*} D_i\\
\end{align*}
where $\hat{t}(\hat{g})$ denotes the resulting set of goals produced by the 
application 
of the tactic $\hat{t}$ to the goal $\hat{g}$.
\end{definition}

We augment the ordered predictions by inserting after each predicted vector 
its descendants and removing lower ranked duplicates.

This is algorithm is used to predict tactics in the preselection
phase of the algorithm. Similarly, dependencies between theorems are used to 
preselect theorems.

\subsection{Preselection}\label{sec:pretactic}
Since predicting efficiently is crucial during proof 
search, we preselect a number of tactics before the search.
We take the first 500 vectors of the augmented ordered list. Only members of 
this list can be predicted during proof search. When an 
open goal $g$ is created, the scores of those 500 preselected tactics are
recomputed which gives a preferential order of tactics to be tried on $g$.

%In practice, such preselection is necessary because our nearest neighbor 
%algorithm  is consuming more than half of the proof search time when there were
%more than 50,000 feature vectors to select from.

%\begin{remark}
%A preselection technique is also in place for the 
%theorem predictor and the evaluation predictor defined later in this paper, 
%because those predictors are also called during the proof search.
%\end{remark}

\section{Orthogonalization}

Different tactics may transform a single goal in the same way. Exploring such 
equivalent paths
is undesirable, as it leads to inefficiency in automated proof search.
To solve this problem, we do not directly assign a goal to the associated 
tactic, but organize a competition on the closest feature vectors. The number 
of selected feature vectors is a parameter that is set for each experiment.
The winner is the tactic appearing in the most feature vectors provided that it 
has the 
same effect as the original tactic. We associate this tactic with the features 
of the targeted goal instead of the original in our feature database. As a 
result, already successful tactics are preferred, and new tactics are 
considered only if they provide a different contribution. We also advantage 
tactics that have a large coverage. In the following, we give formal definition 
of the concepts of subsumption and coverage that are required for expressing 
the orthogonalization algorithm.

\begin{definition} (Coverage)\\ 
Let $\mathfrak{F}$ be a recorded set of feature vectors. We can define the 
coverage $C(t)$ of a tactic $t$ by the number of times this tactic appears in 
$\mathfrak{F}$. Expressing this in a formula, we get:
  \[C(t) =_{def} card\ \lbrace (x,g)\ |\ (x,g) \in \mathfrak{F} \wedge x = t
  \rbrace  \]
Intuitively, this notion estimates how general a tactic is by counting the 
number of different goals it is useful for.
\end{definition}

\begin{definition} (Goal subsumption)\\ 
A goal subsumption $S$ is a partial reflexive relation on a set of goals.\\
Given two goals $g_1$ and $g_2$, we note:
$g_1 \ge g_2 \Leftrightarrow_{def} (g_1,g_2) \in S$.\\
Assuming we have a way to test if a goal is a consequence of another, a 
possibly useful subsumption definition would be: 
$g_1 \ge g_2  \Leftrightarrow_{def} g_1\ \mbox{is a consequence of}\ g_2$\\
In our experiments, we take the minimal subsumption defined by $g_1 \ge g_2  
\Leftrightarrow_{def} g_1 =_{\alpha} g_2$ where $=_{\alpha}$ is 
$\alpha$-equivalence.
\end{definition}

We can naturally extend a goal subsumption to a set of goals in the following 
way.
\begin{definition} (Goal set subsumption)\\
A goal set subsumption is an partial relation on a set of goals.
Given two sets of goals $l_1$ and $l_2$, we can define it from goal 
subsumption $\ge$ by:
\[l_1 \ge l_2  \Leftrightarrow_{def} \forall g_2 \in l_2\ \exists g_1 \in l_1\ 
g_1 \ge g_2\]
\end{definition}

We now build a subsumption on the tactic level.
\begin{definition} (Tactic subsumption)\\  
Given two non-failing tactics $t_1$ and $t_2$ on $g$, a tactic $t_1$ subsumes a 
tactic $t_2$ on a goal $g$, noted $\ge_g$, when:

 \[t_1 \ge_g t_2 \Leftrightarrow_{def} t_1(g) \ge t_2(g)\]

If one of the tactic is failing then the couple $(t_1,t_2)$ is not an element 
of the 
relation.
Given the minimal subsumption on goals, a tactic $t_1$ subsumes another $t_2$ 
on a goal 
$g$ if and only if the resulting set of goals produced by $t_1$ is a subset 
of the output of $t_2$ on $g$. 
\end{definition}

We can now express the orthogonalization algorithm given in a single formula 
with parameters: $P_n$ for the tactic predictor where $n$ for the number of 
predicted
tactics, $S$ for the tactic subsumption relation and $C$ for the coverage of 
a tactic. The orthogonalization process $\mathbb{O}$ is given as input a 
recorded feature vector $(t,g)$ where $t$ is the tactic originally applied to 
the goal $g$. The winning tactic of the competition is given by the formula:

\[\mathbb{O}(t,g) = \argmax_{x\ \in\ P_n(g) \cup \lbrace t \rbrace} \lbrace 
C(x)\ 
|\ x \ge_g t\rbrace\]

After orthogonalization, because the orthogonalized tactic is supposedly 
better, we record the feature vector $(\mathbb{O}(t,g),g)$ instead of 
$(t,g)$. To sum up, the winning tactic is the one with the most coverage and 
that is able to replicate the effect of the original tactic (or do better if we 
woudl consider more complex goal subsumptions).

\section{Tactic Abstraction}\label{sec:synthesis}
One of the great weakness of the previous version of \tactictoe is that
it could not create its own tactics. Indeed, sometimes no previous tactic is 
adapted for a certain goal and creating a new tactic is necessary.
%Experiments in section, show how different degrees of tactic synthesis could 
%improve the success rate of \tactictoe.
We present a way to create tactics with different arguments
by abstracting them and re-instantiating them using a predictor for that kind 
of argument. In the spirit of the orthogonalization method, we will try to 
create tactics that are more general but have the same effect as the 
ungeneralized one.

\paragraph{Abstraction of Tactic Arguments}
The first step is to abstract arguments of tactics by replacing them by a 
place holder creating an tactic with an unspecified argument.
Suppose we have a recorded tactic $T$. Since $T$ is a \sml code tree, we can 
try to abstract any of the 
\sml subterms. Let $u$ be a \sml subterm of $T$, and $h$ a placeholder $u$, we 
can substitute $u$ by $h$ to create an abstracted tactic.
We note this intermediate tactic $T[h/u]$. By repeating the process, multiple 
arguments may be abstracted in the same tactic. Ultimately, the more 
abstractions are performed the more general a tactic is, as many tactics 
become instances of the same abstracted tactic. Yet, it becomes harder and 
harder to predict suitable arguments.

\paragraph{Instantiation of an Abstracted Tactic}
An abstracted tactic is not immediately applicable to a goal, since it contains
unspecified arguments. To apply an abstracted tactic $A$, we first need to 
instantiate that should placeholders $h$ inside $A$. Because it is difficult to 
build a general predictor that effective on each type of argument, we manually 
design different argument predictors for each type. Those predictors are 
usually given a goal has input and returns the best chosen argument for this 
goal(see Section~\ref{sec:arg_pred}). All in all, an abstracted tactics has 
dynamic arguments, it decides how to instantiate them depending on the given 
goal.

\paragraph{Selection of Abstracted Tactics}
As abstracted tactics are not used in human proofs, we need to find a way to 
predict them so that they can contribute to the \tactictoe proof search.
A straightforward idea is to duplicate each tactic during the proof search by 
creating an abstracted version of the tactic. However, this is computationally 
expensive when the tactic does not instantiate back to its original form, we 
risk doing unnecessary work as both version may have a similar effect.
Therefore, we need to decide beforehand which of the two is better.
In fact, we can re-use the orthogonalization module to do this task for us.
During the orthogonalization competition on goal $g$, before trying a tactic 
$T$, we now also try to apply its abstracted version $A$. If $A$ succeeds, it 
becomes a a label for $g$ is including in the database of tactic features and 
thus can be predicted during a proof search.

\begin{remark}
Forcing $A$ into the competition is necessary because it has never been seen 
before so it has no coverage. But after the coverage $c_A$ of $A$ in the 
database exceeds the coverage of $T$, $A$ will be selected before $T$ without 
the 
previously required forcing.
\end{remark}

\subsection{Argument Predictors}\label{sec:arg_pred}
We made argument predictors for two types of arguments: list 
of theorems and term. We choose this two types because they are the most common 
inside tactics and because perfect synthesis of those arguments would lead to 
the largest improvements (see Section~\ref{sec:perfect_exp}). Arguments with 
different types 
are currently never abstracted in tactics since we do 
not have yet a way to predict those arguments and instantiate those tactics.

\paragraph{List of Theorems}
Predicting theorems is made through the same system as premise selection
used for \holyhammer for general purpose automation which produces a list of 
theorems $l$. The original list of theorem $l'$ abstracted in the tactic is 
forgotten. For the abstracted tactic 
to replace the original one, the assumption is that the tactic does not break 
when adding arguments. If this condition is met, the abstracted tactic will 
always subsumes the original tactic when $l$ is a superset of $l'$. Such 
condition is called monotonicity of the 
tactic in its argument. Predicting such superset may not be necessary as the  
predicted theorems can sometimes be even more suitable than the original one 
because they depend directly on the current goal. One of the biggest advantage 
of predicting theorems independently is that tactics may now rely on theorems 
that were never part of a tactic.

\paragraph{Term}
Predicting terms is harder since we do not have a list of terms to choose from.
As an approximation for this list, we take the list of subterms $l$ contained 
in a goal $g$. Even then choosing the right subterms require us to look at the 
term $t$ the tactic argument was abstracted from. The term predictor then 
selects the subterm in $l$ that is closest to $t$. In an attempt to predict 
variables (which are terms in \holfour) at the place where they were previously 
variable, we use the distance $sim_2$ that penalizes large subterms.
Since we rely on the original term argument $t$, the kind of tactics produced 
are much less general that for abstracted list of terms and thus have smaller
coverage.

%In the orthogonalization settings, the abstracted forms compete alongside the 
%originals and are ranked according to their frequency. 
%Freq (abstract) = max (orig,abstr). And the abstracted from is always 
%preferred 
%to the original form of a tactic if both had the effect that was fought upon.
%
%
%Abstracted tactics then become part of the feature space and can be predicted 
%by the tactic predictors but are slightly slower because of the added overhead
%of predicting list of theorems for their unspecified arguments.
%
%We first detect if the tactics contains one or more theorem list. This 
%detection
%is done by looking if the first element of the list is of type thm.
%
%This method is difficult to generalize to any tokens?
%
%By forgetting what was the original argument, we lose an important piece of 
%information on what kind of arguments a tactic accepts. This is less problem 
%for
%a list of theorems because predicted theorems that are not usable by the 
%tactic 
%are usually ignored.
%
%What if you would have to predict a single theorem, or term, or function?
%
%There are two complementary guiding methods to predict them:
%
%- Dependent on the goal.
%- Similar to previous arguments for this tactic.
%
%For example one way to predict subterms, would be to take the subterm that is 
%the closest to predicted subterm that appears on the goal.
%
%For that we take slight variation of the previously defined distance to 
%eliminate large subterms:

%\subsubsection{Estimating the difficulty of a goal}
%A first idea is to guess what is the expected length of its shortest proof.
%
%From the recorded feature vectors, we can compute the length of the shortest 
%proof for each goal $g$ and can be defined recursively by:
%
%length :::= mathematical stuff
%
%Now to compute the difficulty of a goal. We can average the difficulty of the 
%$k$ nearest neighbor. 
%
%However, we believe that it is more important to try to guess if a goal is 
%likely to be provable or not.
%
%We augment our data by looking at all the produced list of goals
%that did not subsumes the original one during 
%orthogonalization. We had the list to our list of negative examples. The list 
%of positive examples are the goal list that are present in the feature vectors.
%
%
%We try to estimate if a goal was useful by taking the k nearest neighbor of 
%the 
%goal and see how many appeared in a previous proofs and how many are hard 
%negatives. The average give us a probability on how provable it is.
%We multiply this probability by a coefficient k.
%
%Using this heuristic to guide our proof search, we hope to reduce the amount 
%of 
%time that the search spends trying to prove unprovable goals.
%
%Formula for the heuristic.

%\begin{remark}
%In practice, each proof node contains a list of goals to be proven and we 
%estimate the difficulty of proving a conjunction of goals.
%\end{remark}

\section{Proof Search}\label{sec:proofsearch}
Despite the best efforts of the prediction algorithms, the selected tactic may 
not 
solve the current goal, proceed in the wrong direction
or even loop. For that reason, the prediction needs to be accompanied by a  
proof search mechanism that allows for backtracking and 
can choose which proof tree to extend next and in 
which direction.
Our search algorithm takes inspiration from advanced tree search algorithm such 
that

The first necessary adaptation of the algorithm stems
from the fact that a proof is in general not a
path but a tree. This means that our search space has two branching factors: 
the choice of a tactic, and the number of goals produced by each tactic.
The first one can be thought as a disjunction of tactics. Indeed, 
if there exists a tactic that solves a goal then it is solved. And 
conversely, the second branching factor is a conjunction of goals. All pending 
goals must be proven in order to prove the parent goal. For example, in the 
case of induction tactics the pending goals are the base case and the inductive 
case.

If we suppose that each tactic produces at most one goal, then the standard 
monte carlo tree search algorithm applies. So we will first discuss how our 
algorithm works in this case. And then explain how we extend it to support 
tactics producing 
multiple goals.

%Our search algorithm takes inspiration from the A*-algorithm~\cite{HartNR68} 
%which uses 
%a cost function and heuristics to estimate the total distance to the 
%destination and
%choose the shortest route.
%The proof is not finished when the current tactic solves its goal because it 
%often leaves new pending open goals along the current path.

\subsection{Diagonalized tree search}
(Move this section to Monte Carlo tree search and simplify it) 
(Don't mention diagonlazied tree search)

%%
%%Our first attempt was to create a breadth first search algorithm with 
%%preferences for the best predicted tactic for each goal. In 
%%Figure~\ref{fig:br1}, we try to prove the goal $g$ using list of tactics 
%%$t_1,t_2,t_3 \ldots$ ordered by the predictor. Those tactics produce 
%%respectively the goals $g_1,g_2,g_3 \ldots$. By choosing tactics
%%for each produced goal $g_i$, we can recursively construct the whole search 
%%tree. We observe that if we apply the breadth first search algorithm to this 
%%tree then the predictor preferences are of little use, as we will not explore 
%%deeper more promising branches. Another related inconvenience is that we may 
%%not ever be able to explore the tree at depth 2 since there maybe too many 
%%tactics to try at depth 1. One way to solve those this problem is to 
%%diagonalized our tree by adding dummy square nodes as presented in 
%%Figure~\ref{fig:diag_choice}. The new branching factor of this tree is 
%%constant 
%%equal to two. The exploration of lower ranked tactics are delayed by the 
%%diagonalization during breadth search. And after $2^n$ breadth-first steps, 
%%the 
%%most promising branch has been explored up to depth $n$.
%%
%%\begin{figure}{}
%%\begin{center}
%%\begin{tikzpicture}[node distance=2cm]
%%\node [tcircle] (0) {$g$};
%%\node [above of=0] (m0) {};
%%\node [tcircle, left of=m0, node distance=1cm] (2) {$g_2$};
%%\node [tcircle, left of=2] (1) {$g_1$};
%%\node [tcircle, right of=m0,node distance=1cm] (3) {$g_3$};
%%\node [right of=3] (inf) {};
%%
%%\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1);
%%\draw[-to,thick] (0) to node[xshift=-10](t2){$t_2$} (2);
%%\draw[-to,thick] (0) to node[xshift=-10](t3){$t_3$} (3);
%%\draw[-to,dotted,thick] (0) to (inf);
%%\end{tikzpicture}
%%%\end{center}
%%%\caption{Choice of a tactic}
%%%\label{fig:choice}
%%%\end{figure}
%%%\begin{figure}{}
%%%\begin{center}
%%\begin{tikzpicture}[node distance=1cm]
%%\node [tcircle] (0) {$g$};
%%\node [above of=0] (m0) {};
%%\node [tcircle, left of=m0] (1) {$g_1$};
%%\node [tsquare, right of=m0] (s1) {};
%%\node [above of=s1] (m1) {};
%%\node [tcircle, left of=m1] (2) {$g_2$};
%%\node [tsquare, right of=m1] (s2) {};
%%\node [above of=s2] (m2) {};
%%\node [tcircle, left of=m2] (3) {$g_3$};
%%\node [right of=m2] (s3) {};
%%
%%\draw[-to,thick] (0) to node[xshift=-5,yshift=-5](t1){$t_1$} (1);
%%\draw[-to,thick] (0) to (s1);
%%\draw[-to,thick] (s1) to node[xshift=-5,yshift=-5](t1){$t_2$} (2);
%%\draw[-to,thick] (s1) to (s2);
%%\draw[-to,thick] (s2) to node[xshift=-5,yshift=-5](t1){$t_3$} (3);
%%\draw[-to,dotted,thick] (s2) to (s3);
%%\end{tikzpicture}
%%\end{center}
%%\caption{Transformation of a search tree into a diagonalized search tree,
%%caused   by a change in the relation between a node and its 
%%children.}
%%\label{fig:diag_choice}
%%\end{figure}
%%
%%
%%In the particular case where tactics produces at most one goal, a proof is 
%%found in the search tree when a tactic $t$ does not produce any new goal. The 
%%proof of the root goal can be reconstructed by extracting the path leading 
%%to the final tactic $t$.
%%
%%
%%
%%\subsection{General case}
%%We now tackle the more general case where tactics $t_i$ may produces multiple 
%%goals $g_{i,1}$ $g_{i,2}$ $g_{i,n}$ from a parent goal $g_p$.
%%
%%All created node are kept in memory for later inspection but some of the 
%%nodes 
%%are deactivated which means they are removed from the proof search.
%%
%%Definition (deactivated node):
%%  A solved node or a node that has a solved ancestor.
%%
%%Then at first, the breadth first search operates normally with only the first 
%%goal $g_{i,1}$ is present in the search tree. 
%%And as soon as a proof is found for the goal $g_{i,1}$, the subtree with root 
%%$g_{i,1}$ is erased and replaced by a node containing $g_{i,2}$. The process 
%%continues with node with the lowest depth being treated first.
%%The process stops if the root goal is solved, the time limit is exceeded or 
%%there is no more tactic to be applied to an active node.

\subsection{Monte Carlo Tree Search}

We will reuse the prediction

%Search Algorithm AlphaGo Zero uses a much simpler variant of the asynchronous 
%policy and
%value MCTS algorithm (APV-MCTS) used in AlphaGo Fan and AlphaGo Lee.
%Each node s in the search tree contains edges (s, a) for all legal actions a ∈ 
%A(s). Each edge
%stores a set of statistics,
%{N (s, a), W (s, a), Q(s, a), P (s, a)},
%where N (s, a) is the visit count, W (s, a) is the total action-value, Q(s, a) 
%is the mean action-value,
%and P (s, a) is the prior probability of selecting that edge. Multiple 
%simulations are executed in
%parallel on separate search threads. The algorithm proceeds by iterating over 
%three phases (a–c in
%Figure 2), and then selects a move to play (d in Figure 2).
%25Select (Figure 2a). The selection phase is almost identical to AlphaGo Fan 
%12 ; we recapitulate
%here for completeness. The first in-tree phase of each simulation begins at 
%the root node of the
%search tree, s 0 , and finishes when the simulation reaches a leaf node s L at 
%time-step L. At each
%of these time-steps, t < L, an action is selected according to the statistics 
%in the search tree,
%
%a t = argmax Q(s t , a) + U (s t , a) , using a variant of the PUCT algorithm 
%24 ,
%a
%pP
%U (s, a) = c puct P (s, a)
%N (s, b)
%1 + N (s, a)
%b
%where c puct is a constant determining the level of exploration; this search 
%control strategy initially
%prefers actions with high prior probability and low visit count, but 
%asympotically prefers actions
%with high action-value.
%Expand and evaluate (Figure 2b). The leaf node s L is added to a queue for 
%neural network
%evaluation, (d i (p), v) = f θ (d i (s L )), where d i is a dihedral 
%reflection or rotation selected uniformly
%at random from i ∈ [1..8].
%Positions in the queue are evaluated by the neural network using a mini-batch 
%size of 8; the
%search thread is locked until evaluation completes. The leaf node is expanded 
%and each edge (s L , a)
%is initialised to {N (s L , a) = 0, W (s L , a) = 0, Q(s L , a) = 0, P (s L , 
%a) = p a }; the value v is then
%backed up.
%Backup (Figure 2c). The edge statistics are updated in a backward pass through 
%each step
%t ≤ L. The visit counts are incremented, N (s t , a t ) = N (s t , a t ) + 1, 
%and the action-value is updated
%to the mean value, W (s t , a t ) = W (s t , a t ) + v, Q(s t , a t ) =
%W (s t ,a t )
%.
%N (s t ,a t )
%We use virtual loss to ensure
%each thread evaluates different nodes 69 .

A similar trade-off between exploration and exploitation exists in algorithm 
like Monte Carlo. However, Monte Carlo tree search metrics are based on many 
simulations and do not reflect our exploration strategy that only

We present here a monte carlo tree search with no roll out which is essentially 
the same as the one use by AlphaGo Zero~\cite{} originally defined by \cite{} 
in \cite{}.

\[Value(a_i) = Exploitation(a_i) + c_{exploration} * Exploration(a_i)\]

\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{Policy(a_i)}\]

\[Policy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(a_i))}}\]

The exploitation function $Exploitation$ is the average of the evaluation of 
all nodes descendant of $a_i$ in the current search tree.

MCTS with prior policy
This modification of the search tree can be simulated by weighting each arrow 
by the distance shown by the diagonalized tree. To simulated the 
diagonalization in the original tree, the arrow $t_i$ will be given weight $i$.
Methods for exploring a tree with a large branching factor has been explored  
in Go~\cite{alphago} and

However for very large branching factor RTS Combinatorial Multi-armed Bandits
for Real-Time Strategy Games. The method of choice for solving those 
problems 
is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
to each branches

In the context of changing wiehgts, the diagonalized tree has the advantage of 
being a little faster to explore as the ordered of tactic is inscribed in the 
shape of the tree. if the prior shapes

\subsection{Optimization}

\begin{definition} (Productive tactic)
During a proof search, after a tactic $t$ is applied to a node with active goal 
$g$, 
$t$ is called 
productive if and only if all the following conditions are satisified:
\begin{itemize}
\item It does not fail. We note $l$ the list of goals returned by the tactic.
\item It does not loop. $l$ do not contain any goals that appears in its 
ancestor nodes.
\item It is not a parallel step. $l$ is not a superset of a list of goals 
produced from the same node with the same active goal.
\end{itemize}

The third point is a partial attempt to prevent confluent branches. The general 
case where two branches becomes parallel after $n$ steps is handle by a
tactic cache which memorizes tactic effects on goals. Similarly there is also a 
prediction cache, which memorize the tactic predictions for each goal. Those 
two caches allow a fast exploration of confluent branches but they remain 
separated in our search tree.
\end{definition}

\subsection{Evaluation}

We now concentrate on defining a reasonable evaluation function that also 
influences node selection in MCTS. Essentially, we estimate how 
interesting is the list of goals of a tactic to determine if we should continue 
searching in this direction. Theoretically, an interesting list of goal is one 
that has the shortest proof, and we could measure the likelyhood of being 
provable in a number of steps from previous proof search. However, this 
approach produces a lot of negatives. So to reduce this number, we limit 
ourselves to feature vectors tested during orthogonalization. We declare a list 
of goals l to be positive, noted $l \in Positive$, if it is has been 
produced by the winner of an orthogonalization competition. We extend the 
definition of a $distance$ to a list of goals by considering the union of the 
features of the goals. The set of the $r$ set of goals closest to a set of 
goal $l$ is noted $Neighbour(r,l)$. We can define the set of positives and 
negatives within a radius $r$ of a 
list of goals $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
Positive \rbrace$.

During our experiments, we evaluate output goals in three different manners 
with the following functions:

\begin{align*}
E_{none} (a_i) &= 0 \\
E_{productive} (a_i) &= 1\\
E_r (a_i) &= \frac{card\ Positive(r,a_i)}{r}\\
\end{align*}

For all evaluations functions, if a tactic is not productive then there is no 
output to evaluate, so the result is a failure and a reward of $0$ will be 
backpropagated.

A large difference between diagonalized search and our MCTS algorithm is in 
what is considered to be the width of the tree. In diagonalized search, 
non-productive tactics contribute to the extension of the width where as in 
MCTSn the width includes only the total number of children produced. The 
influence of non-productive on the proof exploration is recovered when we 
include an 
evaluation function that takes into account the productivity of tactics such as 
$E_{productive}$ or $E_r$.

%Obviously, we can consider the 
%trivial evaluation function where 
%a goal is given reward if and only if it is proven. Yet, to make this approach 
%effective would require to learn on a very large amount of relatively easy 
%problems. To lessen the requirement of an very large amount of data, we 
%propose 
%to use our goal distance to select the $k$ closest goals that were proven 
%before. The following formulas shows how we can combine positive examples and 
%negatives examples to evaluate an unseen 
%goals.
%
%%Hard Negative mining (cite Deep math)
%
%\[\mathfrak{N}_m=\lbrace g\ |\ (N(g) \wedge n(g) \geq m) \vee 
%(P(g) \wedge p(g) >  m)\rbrace\]
%\[\mathfrak{P}_m=\lbrace g\ |\ P(g) \wedge p(g) \leq m \rbrace\]
%
%
%\[E(m) = \frac{1}{m} \times \frac{card\ \mathfrak{P}_m}
%    {card\ \mathfrak{N}_m + card\ \mathfrak{P}_m} \]
%
%
%\[E = \argmax_{m \in \mathbb{N}} E(m)\]
%

%We need
%MCTS.
%We use supervised learning to estimate evaluation functions.
%
%Positives are given the number of intermediate steps needed to finish the 
%proof.
%And the distance for negatives is two times the number of intermediate steps 
%tried. If we only tried one more tactic on this goal we would solve it.

%We use K-nn on them with k = 10 and calculate an the following weighted mean.
%
%
%The heuristic component should represent the number of steps (or total time) 
%estimated required to 
%solve the goal.
%If a goal is not provable then it's the number of time the goal was visited 
%during the search + 1.

%
%
%Combining with general automation:
%metis.
%holyhammer. (with orthogonalized theorems).
%divide by the sum of the weight of all features.
%
%
%New astar heuristic:
%The astar algorithm was simplified and now is represented in the standard 
%additive form instead of the multiplicative one described in our previous 
%paper.
%
%We also tried to estimate the difficulty of a goal. How likely that a 
%conjunction of goals are provable.
%
%Heuristic from prediction score:
%
%Heuristic from positives:
%
%Heuristic from positives + hard negatives.
%
%Admissibility of the heuristic:
%  
%Connection with UCB, UCT and Monte Carlo.
%
%Why is it not applicable in our case?
%
%1) Monte carlo requires to explore every action at least once. 
%This could be fixed by predicting fewer tactics 10-20. But we do not have
%yet an automatic way to tell.
%
%2) Monte Carlo requires a evaluation of the provability of the goal.
%This is insight can be gain.
%
%\subsection{Node selection formula}
%Monte Carlo tree search is a tree search algorithm. It explores the proof tree
%by performing simulation from the root to a leaf of the tree by selecting each 
%node by looking at that includes exploration and exploitation. Intuitively, 
%the 
%trade-off between exploitation and exploration means that the algorithm will 
%exploit deeper more promises branches but leaving enough time for exoloring 
%less likely alternatives.
%
%
%
%Connection with previous tree search. The prior probabilities are decreasing
%exponential of 2s. $\sqrt (n) > 1/2^n$
%
%
%
%
%
%calls simulation. 
%
%
%
%\subsection{Description of the evaluation function}
%
%Non local heuristic for evaluation. Should consider also what is needed to be 
%proven.
%
%
%In order to evaluate how long it takes to prove a goal. We collect data from 
%previous proof searches. Those include negative and positive examples.
%Positives examples are associated with the minimal number of steps needed to 
%find their proofs. Negative examples carry along the largest proof search 
%unsuccessfully performed.
%
%Combining this measure in an single evaluation of the proof length is a 
%challenging task. We use the following formulas P1 for estimating the 
%probability of being provable in less than one step.
%
%Neg(i) = not provable in i steps.
%70 percent change of not being provable in two steps.
%50 percent chance of closing the proof in 1 step: 2
%1 percent chance of being provable in 1 steps.
%20 percent chance of being provable in 2 steps. 5*2 = value = 10
%100 percent chance of being provable in 5 steps. 5
%
%\[P_i = Neg(i) | Knn(10) \] 
%
%the total estimation is given by:
%\[E = (\sum_{i=1}^5 (1-P_i)) + 1\] 
%
%\subsection{Combining evaluation function on goals}
%
%Independence of each probabilities.
%Full correlation: $(P1 * P2 * P3) ^{1/3}$
%Min probabiity:

%
%\subsection{Advantage of Monte Carlo Tree Search over A*-search}
%- Take into account all estimation along the proof and not only the last leaf.
%- Easy to link conjunction estimation   together.
-

\section{ATP and Lemma Prediction Integration}

General-purpose proof automation mechanisms which combine proof translation to
ATPs with machine learning (``hammers'') have become quite successful in
enhancing the automation level in proof assistants~\cite{hammers4qed}.
As external automated reasoning techniques sometimes outperform the combined 
power of tactics, we would like to combine the \tactictoe search with 
\holyhammer for \holfour~\cite{tgck-cpp15}. The principle is that we will 
create a \holyhammer tactic that can be used during the proof search. Since 
external calls to \holyhammer are computationally expensive (5 seconds at a 
minimum), we will also 
experiment with a ``small hammer'' tactic comprised of: a faster premise 
selection algorithm from previously proven theorems $\mathfrak{T}$ and
a short call to the internal prover \metis~\cite{metis}.

It is best tactic available but we do not have any data about it since it does 
not appear in any proofs. So as a first approximation, we 
artificially give it the best prediction scores on any goal and the best 
coverage.

\paragraph{Effect on orthogonalization of tactics}
During orthogonalization, the general purpose tactic will be the first 
competitor and if it manages to prove the goal it will subsume the original 
tactic on this goal since the empty set of goals is a subset of any set. 
Therefore, it will replace the original tactic in the feature vector that is 
competed for. This is desirable since we would like to remove alternative 
proofs.

\paragraph{Orthogonalization of theorems}
A ``hammer'' tactic can be useful to remove parallel
theorems. We consider a theorem to be parallel if it can be proven 
from previous theorems using a general tactic with a timeout equal or less 
than the one it is given during the proof search. Parallel theorems are marked 
and removed from the set of theorems $\mathfrak{T}$ from 
which premise selection is performed.

\begin{remark}
The ``smaller hammer'' is chosen to remove parallel theorems and as the first 
try during orthogonalization of tactics. 
\end{remark}

\paragraph{Effect on the proof search}
First, before the proof search, we preselect 500 theorems 
for the whole proof search tree using a premise selection algorithm 
that takes into account dependencies between theorems.
During the proof search, when a new goal is created or a fresh pending goal is 
considered, the ``hammer'' tactic is always tried first because if it succeeds 
the node would be closed rapidly.

\paragraph{Parallelization of external calls}
It is possible to parallelize calls to external provers so that they happen 
asynchronously during the proof search. We let the prover runs \tactictoe 
continue its proof in parallel of a ``hammer'' tactic call. Additional care 
needs to be taken after each proof step to close proofs solved by the general 
tactic. When the number of thread reaches a limit, a ``hammer'' tactic call 
should
wait for a thread to be freed but the proof search goes on and the feedback of 
the ``hammer'' tactic is delayed.

\section{Experimental Evaluation}\label{s:experiments}

\subsection{Methodology} \todo{is copied}
The evaluation imitates the construction of the library: For each theorem only 
the previous
human proofs are known. These are used as the learning base for the predictions.
To achieve this scenario we re-prove all theorems during a modified build of 
\holfour.
As theorems are proved, they are recorded and included in the training examples.
For each theorem we first attempt to run the \tactictoe search with a time 
limit of 5 seconds,
before processing the original proof script.
In this way, the fairness of the
experiments is guaranteed by construction. 
Only previously declared \sml 
variables (essentially tactics, theorems and simpsets) are accessible. 
And for each theorem to be re-proven \tactictoe is only trained on previous 
proofs.

Although the training process in each strategy on its own is fair, the 
selection of the 
best strategy in Section~\ref{sec:sel_param} should also be considered as a 
learning 
process. To ensure the global fairness, the final experiments in 
Section~\ref{sec:exp_full} 
runs the best strategy on the full dataset which is about 10 times larger. The
performance is minimally better on this validation set.

\subsection{Tuning \tactictoe}

\subsubsection{Feature parameters}
The first experiment concerns the choice of the right kind of features and 
feature scoring 
mechanism. The results are presented in Table~\ref{tab:feature_param}.
We observe that the higher-order features and the feature of the top logical 
structure increase minimally the number of problems solved. The attempted 
length penalty on the total number of features is actually slightly harmful.
We did not turn on the top feature flag in the next experiments because they 
are slightly more expensive too compute.

%
%\begin{table}[t]
%\centering\ra{1.3}
%\small
%\begin{tabular}{llll}
%\toprule
% ID & Learning parameter & Solved & $U(D_1)$ \\
%\midrule
% $D_0$ & $codist_1$ (length penalty) & 172 (20.0\%) & 5 \\
% $D_1$ & $codist_0$ (default) & 179 (20.8\%) & 0 \\
% $D_2$ & no top features & 175 (20.3\%) & 18 \\
% $D_3$ & no higher-order features & 178 (20.7\%) & 8 \\
%\bottomrule
%\end{tabular}
%\caption{\label{tab:feature_param}Success rate of strategies with different 
%learning parameters on the training set.}
%\end{table}
%
%\subsubsection{Search parameters}
%
%\paragraph{tactic timeout}
%\begin{table}[t]
%\centering\ra{1.3}
%\small
%\begin{tabular}{llll}
%\toprule
% ID & Searching parameter & Solved & $U(D_{9})$ \\
%\midrule
% $D_1$ & $codist_0$ (default)      & 179 (20.8\%) & 10 \\
% $D_4$ & tactic timeout 0.004 sec & 175 (20.3\%) & 9 \\
% $D_5$ & tactic timeout 0.1 sec   & 178 (20.7\%) & 10 \\
%\bottomrule
%\end{tabular}
% \caption{\label{tab:search_param}Success rate of strategies with different 
% search parameters on the training set.}
%\end{table}

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{llllll}
\toprule
 $c_{policy}$ & Width & Ortho & Eval & $c_{exploration}$ & Solved \\
\midrule
 2/3 & All & No & $E_{none}$ & 1 & 120\\
 1/2 & All & No & $E_{none}$ & 1 & 125\\
% 1/2 & All & Yes & $E_{none}$ & 1 & 130 \\
 1/3 & All & No & $E_{none}$ & 1 & 117\\
% 1/2 & Productive & No & $E_{none}$ & 1 & 120\\
% 1/2 & Productive & No & $E_{productive}$ & 1 & 127\\
\bottomrule
\end{tabular}
\caption{\label{tab:cost_param} on a local machine}
\end{table}

%This experiments were performed without the square root in the policy 
%function.

\begin{table}[ht]
\centering\ra{1.3}
\small
\begin{tabular}{llllll}
\toprule
  $c_{policy}$ & Width & Ortho & Eval & $c_{exploration}$ & Solved \\
\midrule
 1/2 & All & No & $E_{none}$ & 1 & 122\\
 1/2 & All & Yes & $E_{none}$ & 1 & 128\\
 1/2 & Productive & Yes & $E_{none}$ & 1 & 128\\
 1/2 & Productive & Yes & $E_{productive}$ & 1 & 134\\ 
 &  & &  & 4 & 130\\
 1/2 & Productive & Yes & $E_1$ & 1 & 127\\ 
 & & & & 4 & 126\\ 
 1/2 & Productive & Yes & $E_{10}$ & 1 & 136\\ 
 & & & & 4 & 137\\ 
\bottomrule
\end{tabular}
\caption{\label{tab:learning} on the server}
\end{table}

We settled for the first parameters where an evaluation function based on 
machine learning was beneficial. It is highly likely than better parameters for 
the policy, radius of evaluation and exploration coefficient can be found with 
further investigation.

%\begin{table}[ht]
%\centering\ra{1.3}
%\small
%\begin{tabular}{ll}
%\toprule
% & Solved \\
%\midrule
% diag coeff: 0.5 & 120\\
% diag coeff: 1   & 125\\
% diag coeff: 1 + orthogonalization & 130 \\
% diag coeff: 2 &  117\\
% no evaluation & 120\\
% trivial evaluation E1 & 127\\
%\bottomrule
%\end{tabular}
%\caption{\label{tab:cost_param} on a local machine}
%\end{table}
%
%\begin{table}[ht]
%\centering\ra{1.3}
%\small
%\begin{tabular}{ll}
%\toprule
% & Solved \\
%\midrule
% none & 122\\
% ortho & 129\\
% no evaluation + ortho & 128\\
% trivial evaluation + ortho    & E1:134 E4:130\\
% evaluation radius: 1 + ortho  & E1:127 E4:126\\
% evaluation radius: 10 + ortho & E1:136 E4:137\\
%\bottomrule
%\end{tabular}
%\caption{\label{tab:learning} on the server}
%\end{table}

We preselect around each feature vectors, one hundred feature of but this takes 
time. What kills us is the time needed to select the neighbors among the 
maximum of 50,000 feature vectors but maybe we can hope to 
taking half the time of the proof search.

\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}

We investigate what are the limits of \tactictoe and what benefits additional 
knowledge could provide. To this end, we will modify the set of features to 
include features vectors that should not be known in a regular experiment. We  
include some vectors from the recorded human proof and perform the proof search 
after the recording instead of before in all other experiments. This experiment 
is not fair and should only be compared to 
experiments in the same settings.

For clarity, we will name the set of feature vectors that can be recorded from 
the current human proof $\mathbb{H}$ and the set of vectors recorded from 
previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
$\pi_1^1$ the projection on the 
first component returning the label of a feature vector. And we define the 
function $T$ that takes a tactic and return the list of its tokens.

We will progressively add bigger and bigger 
subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
search success rates.

Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we consider:
\begin{align*}
\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
   \rbrace \\
\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
\rbrace\\
\end{align*}

$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
prediction for those tactics.

\begin{table}[h]
\centering\ra{1.3}
\small
\begin{tabular}{lll}
\toprule
 & Additional features & Solved \\
\midrule
$\emptyset$ & none & 119 (43.6\%) \\ %NONEv2
$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
$\mathbb{H}$     & all & 239 (87.5\%) \\ %AFTERSMALLv2
% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
\bottomrule
\end{tabular}
\caption{\label{tab:featue_param} Success rates of with an artificially 
augmented dataset of features fetched from the human proof of the tested 
theorem}
\end{table}

And we observe in Table~\ref{tab:featue_param} a 
slight increase compared to the default \tactictoe. We then included all 
vectors that contains tactics that can be rebuild from tokens of
tactics in our database. 
Our search algorithm proved two proofs less that if it had perfect prediction 
so it means that our prediction algorithm (policy) is doing a good  job at 
least in this early settings where proofs are easy. 
Suprisingly, being able to recombine tokens of tactics from the database is 
barely sufficient to prove more than half of the theorems in this dataset. The 
Therefore, additionally including in tactics theorems and terms that did not 
appear in previous tactics is essential. Creating from scratch other arguments 
may not help as much. This is why we restricted ourselves to predicting 
arguments of theorem list type and the term type during tactic synthesis. 
Finally our proof search algorithm sometimes fails even if it knows all the 
tactic necessary. The two main reasons are that either a crucial tactic exceeds 
its timeout or that the human proof is too deep to be rediscovered (more than 
10 steps long).

\subsubsection{Tactic synthesis}
Here, we test the benefit of tactic synthesis compared to the old version.
In all those experiments orthogonalization is performed with radius 20 but not 
orthogonalization with respect to metis.

All experiments in this table are using orthogonalization.
\begin{table}[t]
\centering\ra{1.3}
\small
\begin{tabular}{ll}
\toprule
  & Solved \\
\midrule
% holyhammer & 79 \\
 none & 122\\
 ortho & 129\\
 ortho + term & 126\\    
 ortho + theorem list & 160\\
 ortho + metis & 162\\
 %ortho + metis + thmortho & 155\\
 %ortho + metis + stacortho & 160\\
 ortho + metis + theorem list & 175\\ 
\bottomrule
\end{tabular}
\caption{\label{tab:cfot_param}
Sucess rates of tactictoe with synthesis for different arguments.}
\end{table}
We want to know what the optimal number of predictions for tactics different 
from \metis since \metis has a special status in our algorithm is handled 
independently.

\subsubsection{Hammer parameters}

The third experiment, presented in Table~\ref{tab:lh_param},
evaluates the effect of integrating the ``small hammer'' in the \tactictoe 
search.
At a first glance, the increased success rate is significant for all tested 
parameters. 
Further analysis reveals that increasing the number of premises from 8 to 16 
with a 
timeout of 0.02 seconds is detrimental. The $D_{19}$ experiment  demonstrates 
that 0.1 
seconds is a better time limit for reasoning with 16 premises. And the $D_{18}$ 
experiment reveals the disadvantage of unnecessarily increasing the timeout of 
\metis. This reduces the time available for the rest of the proof search, which 
makes the success rate drop.

%\begin{table*}[t]
%\centering\ra{1.3}
%\small
%\begin{tabular}{llll}
%\toprule
% ID & ``small hammer'' parameter & Solved &  $U(D_{19})$ \\
%\midrule
% $D_9$ & $codist_5(0.8,0.8)$ (default: no small hammer) & 211 (24.5\%) & 19 \\
% $D_{16}$ & 8 premises + timeout 0.02 sec & 281 (32.7\%) & 21 \\ 
% $D_{17}$ & 16 premises + timeout 0.02 sec & 270 (31.4\%) & 18 \\
% $D_{18}$ & 8 premises + timeout 0.1 sec & 280 (32.6\%) & 11 \\
% $D_{19}$ & 16 premises + timeout 0.1 sec & 289 (33.6\%) & 0 \\
%\bottomrule
%\end{tabular}
% \caption{\label{tab:lh_param}Success rate of strategies with different 
% parameters of
% ``small hammer'' on the training set.}
%\end{table*}

The best strategy which does not rely on the ``small hammer'' approach $D_9$ 
will be called \tactictoe(NH) (for no ``small hammer'') in the remaining part 
of the paper, and the best strategy relying on the approach, $D_{19}$, will be 
referred to as \tactictoe(SH) (``small hammer'').

%\begin{table}[h!]
%\centering\ra{1.3}
%\small
%\begin{tabular}{lll}
%\toprule
% ID & Orthogonalization of theorems & Solved \\
%\midrule
% $NONE$           &                             & 118\\
% $METIS16$        & using metis                 & 143\\
% $METIS16_NEW$    & no theorems from boolTheory & 155\\
% $METIS16v2NOINT$ & no internal theorems        & 159\\
% $METIS16v2DIAG$  & diagonal exploration        & 166\\ 
% $METIS16v2INT$   & internal theorems           & 168\\
% $METIS16_30$     & 30 seconds timeout          & 162\\
%\bottomrule
%\end{tabular}
%\caption{\label{tab:cst_param}}
%\end{table}

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{ll}
\toprule
  & Solved \\
\midrule
   \eprover &  106 \\
   \tactictoe &  139 \\
   \tactictoe + \metis & 182 \\
   \tactictoe + asynchronous \eprover & 182 \\
%x tactictoe + async holyhammer 10 & 183 \\
   \tactictoe + \metis + asynchronous \eprover & 190 \\

\bottomrule
\end{tabular}
\caption{Experiments with a timeout of 60 seconds \label{tab:param}}
\end{table}

\subsection{Full-scale experiment}

Mention the conflicts.
Theorem orthogonalization is not turn on during instantiation of tacitcs except 
for the metis tactic.
Best parameters + comparison with holyhammer.

Until now, all our feature vectors were learned form human proofs. 
human proofs. We can also add to this tactics that were part of a proof found 
by our algorithm. To prevent duplication of effort, orthogonalization of those 
tactics is essential to have a beneficial effect (see Experiments).
Since recording and re-proving are intertwined, the 
additional data is available for the next proof search.
The hope is that the algorithm will improve faster by learning from its own 
discovered proofs than from the human proof 
scripts~\cite{DBLP:conf/cade/Urban07}. It is also the first step toward 
self-reinforcement learning.

\begin{table}[h!]
\centering\ra{1.3}
\small
\begin{tabular}{llll}
\toprule
  & & Small(273) & Big(7088) \\
\midrule
  & \eprover (auto-schedule + 128 premises) & 86 & 2069 (29.19\%)\\ 
%  & \tactictoe (without special care for metis) & 154 & 3868 (\\
  & \tactictoe (best parameters) & 187 & 4186 (59.06\%) \\
%  & \tactictoe (+ self-learning) & 187 (187) &  4201 (59.27\%) \\
\bottomrule
\end{tabular}
\caption{Experiments with a timeout of 10 seconds \label{tab:_param}}
\end{table}

It is learning to use \metis however it cannot automatically learn to give it 
more time making it still a bit weaker than the adhoc startegy.

Table~\ref{theories} compares the success rates of re-proving for different
\holfour theories. \tactictoe(SH) outperforms \tactictoe(NH) on every 
considered theory.
Even if a lot weaker due to the missing premise selection component, 
\tactictoe(NH)
is hugely better than \holyhammer(blistr) in 4 theories: \texttt{measure}, 
\texttt{list}, \texttt{sorting} and \texttt{finite\_map}. The main reason is 
that 
those theories are equipped with specialized tactics, performing complex 
transformation such as structural induction and \tactictoe(NH) can reuse them. 
Conversely, \holyhammer(blistr) is more suited to deal with dense theories such 
as 
\texttt{real} or \texttt{complex} where a lot of related theorems are available 
and 
most proofs are usually completed by rewriting tactics.

\begin{table}[b!]
\centering
\setlength{\tabcolsep}{3mm}
\begin{tabular}{@{}ccccc@{}}
\toprule
\phantom{ab} & {arith} & {real} & {compl} 
& {meas} \\
\midrule
\tactictoe & 74.6 & 67.3 & 77.3 & 28.2\\
\eprover & 50.8 & 67.6 & 63.4 & 10.8 \\
\midrule
\phantom{abc}  & {proba} & {list} & {sort} & {f\_map} \\
\midrule
\tactictoe & 37.3 & 71.4 & 58.4 & 73.6 \\
\eprover & 19.3 & 20.7 & 15.8 & 20.9 \\
\bottomrule
\end{tabular}
\caption{\label{theories}Percentage (\%) of re-proved theorems in the theories 
\texttt{arithmetic}, \texttt{real}, \texttt{complex}, \texttt{measure},  
\texttt{probability}, \texttt{list}, \texttt{sorting} and \texttt{finite\_map}. 
}
\end{table}

\subsection{Time and space complexity}

\pgfplotscreateplotcyclelist{my black}{
solid, mark repeat=100, mark phase=0, black!100\\
dashed, mark repeat=100, mark phase=0, black!100\\
solid, mark repeat=20, mark phase=10, black!100, mark=triangle*\\
solid, mark repeat=20, mark phase=10, black!75, mark=square*\\
solid, mark repeat=20, mark phase=10, black!50, mark=*\\
solid, mark repeat=20, mark phase=10, black!50, mark=o\\
}

\begin{figure}[p]
\centering           
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=south east, at={(0.9,0.1)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=10,
  ymin=0, ymax=4500,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=time, y=solved] {data/tt_ptime};
\addplot table[x=time, y=solved] {data/e_ptime};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\caption{Number of problem solved in less than x seconds.}
\end{figure}

\begin{figure}[p]
\centering           
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=10,
  ymin=0, ymax=100,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=size, y=solved] {data/tt_percent};
\addplot table[x=size, y=solved] {data/e_percent};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\caption{Percentage of problem solved with respect to the length of the 
original 
proof}
\end{figure}

\begin{table*}[t]
\centering\ra{1.3}
\small
\begin{tabular}{llllll}
\toprule
 ID & \multicolumn{2}{c}{nodes}   & \multicolumn{2}{c}{proof size} & 
 \multicolumn{1}{c}{time}\\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} 
 \cmidrule(lr){6-6}
     & average & max & average & max & average \\
\midrule
 \tactictoe & 94.66 & 421 & 3.34 & 39 & 0.66 \\
\bottomrule
\end{tabular}
 \caption{\label{tab:stats}search statistics}
\end{table*}

\section{Proof Presentation}\label{sec:proofdisplay}

\paragraph{Reconstruction}
When a proof search succeeds (there are no more pending goals at the root)
we need to reconstruct a \holfour human-style proof.
The saved nodes consist of a set of trees where each edge is a tactic and
the proof tree is the one starting at the root.
In order to obtain a single \holfour proof, we need to combine the tactics
gathered in the trees using tacticals.
By the design of the search, a single tactic combinator, \texttt{THENL}, is 
sufficient. It combines a tactic with a list of subsequent ones, in such a way 
that after the parent tactic is called, for each created goal a respective 
tactic from the list is called.
The proof tree is transformed into a final single proof script
 by the following recursive function $P$ taking a
tree node $t$ and returning a string:
\begin{equation*}
P(t) =
\begin{cases}
P(c) & \text{if $t$ is a root},\\
tac & \text{if $t$ is a leaf},\\
tac\ \texttt{THENL}\ [P(c_0),\ldots,P(c_n)] & \text{otherwise.}
\end{cases}
\end{equation*}
where $tac$ is the tactic that produced the node, $c$ is the
only successful child of the root and $c_0, \ldots, c_n$ are the 
children of the node produced by the successful tactic.

The readability of the created proof scripts is improved, by replacing
replacing  \texttt{THENL} by \texttt{THEN} when the list has length 1.
Further post-processing such as
removing unnecessary tactics and theorems has been developed and
improve the user experience greatly~\cite{DBLP:conf/sefm/Adams15}.

\paragraph{Minimizing length and time of the proof} 
First a weak minimization is possible when processing the final proof. If the 
tactic \texttt{A THEN B} appears in the proof and has the same effect as 
\texttt{B} then we can drop \texttt{A} from the proof.
Stronger minimization can be obtained by rerunning the prover with feature 
vectors containing only tactics 
from the proof. To obtain a shorter proof, the diagonalization coefficient is
reduced to encourage exploration in the width of the proof tree. Similarly by 
ponderating each tactic application by its time and setting the 
diagonalization  coefficient is set to a tenth of the timeout of a tactic, we 
can aim for faster proofs.

\paragraph{Minimizing tactic arguments}
One of the argument that may contain unnecessary information are lists.
We test if the minimized tactic has the same effect as the original one.
This is a generalization of the simplest method used for minimizing a list of 
theorems in ``hammers'' \cite{}. The fact that the lists are generally small 
make this approach feasible.

\paragraph{Prettification}
Without prettification, the returned tactic is barely readable as it contains 
information to guarantees that each \sml subterm is interpreted in the way in 
any context. Our approach here is to strip this information from the tactic and 
test if the tactic has the same effect as the original one to guarantee that we 
are not changing the proof. For efficiency, 
this is performed on each tactic step separately. We strip module prefixes, 
local declaration. If local declaration can't be 
stripped, we try to group all local declarations under a single "let" binding 
which improves readability and prevents duplication of this declarations.

%\paragraph{Step by step exploration.}
%Because a human may have better insight on which tactic to chose next, we also 
%provides an interactive mode, where the user is presented the best $n$ 
%predicted tactics for a goal and to be chosen from. The number $n$ is 
%parameter 
%that can be chosen by the user. Those tactics are curated so that only 
%productive tactics are returned. (define productive)

\section{Case study}
Investigating further the different qualities of \tactictoe, we study its 
generated proof scripts on an example in \texttt{list} theory (see 
Example~\ref{ex:casestudy}). 
The theorem to be proven states the equivalence 
between the fact that a number $n$ is greater than the length of a list 
$\mathit{ls}$ with the 
fact that dropping $n$ elements from this list returns an empty list.

The human proof proceeds by induction on $n$ followed by solving both goals 
using rewrite steps combined with an arithmetic decision procedure. 
Both \tactictoe proofs (NH and SH) follow the general idea of reasoning by 
induction but solve 
the base case and the inductive case in a different way.
The base case only needs rewriting using the global simpset in the 
\tactictoe(NH)
proof, which is simulated by a call to \metis in the (SH) proof.
The inductive case should require an arithmetic decision procedure as hinted by 
the human 
proof. This is achieved by rewriting using an arithmetic simpset in the second 
proof. In the first proof however, a rewriting step and case splitting step 
were used
to arrive at a point where \metis calls succeed.
The tactic proof produced by \tactictoe(NH) often looks better than the one 
discovered by \tactictoe(SH) in that it does not involve \metis calls with a 
large numbers of premises.

%In general, \tactictoe does not search for beauty, but sometimes it is 
%discovered by accident.

%\begin{remark}
%The additional difficulty produced by the necessary induction is such that the 
%theorem could not be proven by \holyhammer.
%\end{remark}

%SURJ_INJ_INV
%proof length: 7
%proof time: 0.092014
%Statistics
%  infstep : 500
%  nodes   : 132
%  maxdepth: 7
%Time: 9.574603
%  inferstep time: 7.287698
%  node_find time: 0.135171
%  node_crea time: 1.674807
%    pred time: 0.335816
%    thmpred time: 0.000018
%    mc time: 0.704148
%    inst time: 0.0

\newcommand{\wed}{\begin{math}\wedge\end{math}}

\begin{example}\label{ex:casestudy}(
Proof discovered by \tactictoe for theorem \texttt{SURJ\_INJ\_INV} In 
theory \texttt{pred\_set})
\small
\begin{center}
\begin{alltt}
    SRW_TAC [] [auto] 
  THEN 
    POP_ASSUM (MP_TAC o 
      CONV_RULE (DEPTH_CONV RIGHT_IMP_EXISTS_CONV 
                 THENC SKOLEM_CONV THENC REWRITE_CONV [EXISTS_DEF] 
                 THENC DEPTH_CONV BETA_CONV)) 
  THEN
    Q.SPEC_TAC (`@y. !x. x IN t ==> y x IN s \wed (f (y x) = x)`,`g`) 
  THEN
    SRW_TAC [] [INJ_DEF] 
  THEN
    Q.EXISTS_TAC `g` 
  THEN
    SRW_TAC [] [INJ_DEF] 
  THEN
    METIS_TAC [auto]

\end{alltt} 
\end{center}
\end{example}

%The term and the conversion rules were not synthesized but simply copied from 
%previous 
%proofs.

\section{Conclusion}\label{sec:concl}

\todo{is copied (modified a bit)}
We proposed a new proof assistant automation technique which combines 
tactic-based proof search with machine learning tactic prediction
Its implementation,
\tactictoe, achieves an overall performance of 59\% theorems on most of the 
\holfour standard library surpassing \eprover with auto-schedule. Its 
effectiveness is especially visible on 
theories which use inductive data structures, specialized decision procedures, 
and custom built simplification sets.
Thanks to the learning abilities of \tactictoe, the generated proof scripts 
usually reveal the high-level structure of the proof. %For these reasons, 
We therefore believe that predicting ITP tactics based on the current goal 
features is a very reasonable approach to automatically guiding proof search, 
and that accurate predictions can be obtained by learning from the knowledge 
available in today's large formal proof corpora. 
%And it seems to get us 
%closer to the way how humans learn to do theorem proving.

There is plenty of future work in the directions opened here.
To improve the quality of the predicted tactics, 
we would like to predict their arguments independently.
To be even more precise, the relation between the 
tactic arguments and their respective goals could be used.
Additionally, we could aim for a tighter combination with the ATP-based hammer
systems. This would perhaps make \tactictoe slower, but it might allow
finding proofs that are so far both beyond the ATPs and \tactictoe's
powers. The idea of reusing high-level blocks of reasoning and
then learning their selection could also be explored
further in various contexts. Larger frequent blocks of (instantiated) tactics
in ITPs as well as blocks of inference patterns in ATPs could be detected
automatically, their
usefulness in particular proof situations learned from the large corpora of
ITP and ATP proofs, and reused in high-level proof search.

%%%Improve tactic pred with term predictions. 
%%%just use a similarity on terms.
%%%should also take into account the things that are not similar.
%%%Variables should lead to predicting variables. Maybe treat variables 
%%%separately make a is_var.

\paragraph{Acknowledgments}\label{sect:acks}
This work has been supported by the
ERC Consolidator grant no.\ 649043 \textit{AI4REASON} and ERC starting
grant no.\ 714034 \textit{SMART}.

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}
% LocalWords: TacticToe tactictoe precedences ITP ITPs THENL

### 2018-01-11 tactictoe_jv.tex
figure for search tree

@@
 
-\todo{These 2 paragraphs are copied}
+\todo{These 2 paragraphs are copied and slightly edited}
 Many of the state-of-the-art interactive theorem provers (ITPs) such as 
@@
 \item The internal ATP \metis used during our proof search is complemented by 
-calls to \holyhammer. For efficiency reasons, those calls are run in parallel 
-and asynchronously.
-\item First evaluation of \tactictoe and \holyhammer success rates relative to 
+asynchronous calls to \eprover.
+\item First evaluation of \tactictoe and \eprover success rates relative to 
 the length of the original proofs.
@@
 
-\subsection{Dependencies}\label{sec:dependencies}
+\subsection{Preselecting feature vectors and their 
+dependencies}\label{sec:dependencies}
 We first rank feature vectors by their distance to the conjecture. But better 
@@
 We augment the ordered predictions by inserting after each predicted vector 
-its descendants and removing lower ranked duplicates.
-
-This is algorithm is used to predict tactics in the preselection
-phase of the algorithm. Similarly, dependencies between theorems are used to 
-preselect theorems.
-
-\subsection{Preselection}\label{sec:pretactic}
+its descendants and removing lower ranked duplicates. This contrast with 
+theorem predictions where we only select parents. This phase is only applied 
+during preselection, so is of no great consequence as long as we select enough 
+theorems and tactics.
+
 Since predicting efficiently is crucial during proof 
@@
 Let $\mathfrak{F}$ be a recorded set of feature vectors. We can define the 
-coverage $C(t)$ of a tactic $t$ by the number of times this tactic appears in 
+coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times 
+this tactic 
+appears in 
 $\mathfrak{F}$. Expressing this in a formula, we get:
-  \[C(t) =_{def} card\ \lbrace (x,g)\ |\ (x,g) \in \mathfrak{F} \wedge x = t
+  \[\textit{\mbox{Coverage}}(t) =_{def} card\ \lbrace (x,g)\ |\ (x,g) \in 
+  \mathfrak{F} 
+  \wedge x 
+  = t
   \rbrace  \]
@@
 \begin{definition} (Goal subsumption)\\ 
-A goal subsumption $S$ is a partial reflexive relation on a set of goals.\\
-Given two goals $g_1$ and $g_2$, we note:
-$g_1 \ge g_2 \Leftrightarrow_{def} (g_1,g_2) \in S$.\\
+A goal subsumption $S$ is a partial relation on a set of goals.\\
+Given two goals $g_1$ and $g_2$, we note $g_1 \ge g_2$ when  $(g_1,g_2) \in 
+S$.\\
 Assuming we have a way to test if a goal is a consequence of another, a 
 possibly useful subsumption definition would be: 
-$g_1 \ge g_2  \Leftrightarrow_{def} g_1\ \mbox{is a consequence of}\ g_2$\\
-In our experiments, we take the minimal subsumption defined by $g_1 \ge g_2  
-\Leftrightarrow_{def} g_1 =_{\alpha} g_2$ where $=_{\alpha}$ is 
-$\alpha$-equivalence.
+\[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is a consequence of } g_2\]
+In our experiments, we take a minimal subsumption defined by:
+\[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is $\alpha$-equivalent to }  
+ g_2\]
 \end{definition}
 
-We can naturally extend a goal subsumption to a set of goals in the following 
-way.
+We can naturally extend a goal subsumption to a set of goals.
 \begin{definition} (Goal set subsumption)\\
@@
 predicted
-tactics, $S$ for the tactic subsumption relation and $C$ for the coverage of 
+tactics, $S$ for the tactic subsumption relation and $\textit{\mbox{Coverage}}$ 
+for the coverage of 
 a tactic. The orthogonalization process $\mathbb{O}$ is given as input a 
@@
 \[\mathbb{O}(t,g) = \argmax_{x\ \in\ P_n(g) \cup \lbrace t \rbrace} \lbrace 
-C(x)\ 
+\textit{\mbox{Coverage}}(x)\ 
 |\ x \ge_g t\rbrace\]
@@
 $(t,g)$. To sum up, the winning tactic is the one with the most coverage and 
-that is able to replicate the effect of the original tactic (or do better if we 
-woudl consider more complex goal subsumptions).
+that is able to replicate the effect of the original tactic.
 
@@
 which direction.
-Our search algorithm takes inspiration from advanced tree search algorithm such 
-that
+Our search algorithm is an advanced tree search algorithm that relies on 
+evaluation and policy functions to guide the search.
 
@@
 
-\subsection{Diagonalized tree search}
-(Move this section to Monte Carlo tree search and simplify it) 
-(Don't mention diagonlazied tree search)
+%\subsection{Diagonalized tree search}
+%(Move this section to Monte Carlo tree search and simplify it) 
+%(Don't mention diagonlazied tree search)
 
@@
 
-\subsection{Monte Carlo Tree Search}
-
-We will reuse the prediction
-
 %Search Algorithm AlphaGo Zero uses a much simpler variant of the asynchronous 
@@
 
+\subsection{A proof tree and solved node}
+
+A tactical proof tree is a set of nodes containing a list of goals where each 
+goal $g$ of a node $n$ is linked to a node $m$ when the list of
+
+\begin{figure}{}
+\begin{center}
+\begin{tikzpicture}[node distance=0.7cm]
+\node [tcircle] (0) {$g_0^i$};
+\node [right of=0] (0r) {.\ .\ .};
+\node [tcircle,right of=0r] (0rr) {$g_0^n$};
+\node [left of=0] (0l) {.\ .\ .};
+\node [tcircle,left of=0l] (0ll) {$g_0^1$};
+\node [draw,fit=(0rr) (0ll)] {}; 
+\node [above of=0,node distance=2cm] (2) {.\ .\ .};
+\node [left of=2, tcircle] (2l) {$g_2^1$};
+\node [tcircle,right of=2] (2r) {$g_2^p$};
+\node [draw, fit=(2l) (2r)] (2b) {};
+
+\node [tcircle, left of=2l, node distance=2cm] (1r) {$g_1^q$};
+\node [left of=1r] (1) {.\ .\ .};
+\node [tcircle, left of=1] (1l) {$g_1^1$};
+\node [draw, fit=(1l) (1r)] (1b) {};
+
+\node [right of=2r,node distance=2cm] (3l) {$\phantom{g_3^p}$};
+\node [right of=3l] (3) {\phantom{.\ .\ .}};
+\node [right of=3] (3r) {$\phantom{g_3^p}$};
+\node [fit=(3l) (3r)] (3b) {};
+
+\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
+\draw[-to,thick] (0) to node[xshift=-10](t2){$t_2$} (2b);
+\draw[-to,dotted,thick] (0) to (3b);
+\end{tikzpicture}
+\end{center}
+\caption{Choice of a goal and a tactic}
+\label{fig:choice}
+\end{figure}
+
+Starting from the bottom node, we can construct descendant of this node by 
+choosing a goal $g_i^0$ on this node and choosing tactics $t1,t2,\ldots$ to 
+apply to this goal. Here the result produces new nodes each containing a list 
+of goals to be proven. If one of those list where empty, the goal $g_i^0$ would 
+be proved.
+
+A node is solved when every of its goal are solved.
+A goal is solved when every of its children node are solved.
+
+\subsection{Guiding the proof search}
+
 A similar trade-off between exploration and exploitation exists in algorithm 
@@
 
-The third point is a partial attempt to prevent confluent branches. The general 
+\end{definition}
+
+The third point of the definition is a partial attempt to prevent confluent 
+branches. The general 
 case where two branches becomes parallel after $n$ steps is handle by a
@@
 separated in our search tree.
-\end{definition}
 
@@
 \section{Experimental Evaluation}\label{s:experiments}
+
+Disclaimer:
+Comparing results from different tables is possible but not advisable since 
+conditions of the experiments may vary slightly because of various factors such 
+as overhead cost of additional features or the fact that they were run on 
+different machines on a local machine or on the server.
 
@@
 \end{tabular}
- \caption{\label{tab:stats}search statistics}
+ \caption{\label{tab:stats} old search statistics}
 \end{table*}
@@
 \section{Case study}
-Investigating further the different qualities of \tactictoe, we study its 
-generated proof scripts on an example in \texttt{list} theory (see 
-Example~\ref{ex:casestudy}). 
-The theorem to be proven states the equivalence 
-between the fact that a number $n$ is greater than the length of a list 
-$\mathit{ls}$ with the 
-fact that dropping $n$ elements from this list returns an empty list.
-
-The human proof proceeds by induction on $n$ followed by solving both goals 
-using rewrite steps combined with an arithmetic decision procedure. 
-Both \tactictoe proofs (NH and SH) follow the general idea of reasoning by 
-induction but solve 
-the base case and the inductive case in a different way.
-The base case only needs rewriting using the global simpset in the 
-\tactictoe(NH)
-proof, which is simulated by a call to \metis in the (SH) proof.
-The inductive case should require an arithmetic decision procedure as hinted by 
-the human 
-proof. This is achieved by rewriting using an arithmetic simpset in the second 
-proof. In the first proof however, a rewriting step and case splitting step 
-were used
-to arrive at a point where \metis calls succeed.
-The tactic proof produced by \tactictoe(NH) often looks better than the one 
-discovered by \tactictoe(SH) in that it does not involve \metis calls with a 
-large numbers of premises.
+%Investigating further the different qualities of \tactictoe, we study its 
+%generated proof scripts on an example in \texttt{list} theory (see 
+%Example~\ref{ex:casestudy}). 
+%The theorem to be proven states the equivalence 
+%between the fact that a number $n$ is greater than the length of a list 
+%$\mathit{ls}$ with the 
+%fact that dropping $n$ elements from this list returns an empty list.
+%
+%The human proof proceeds by induction on $n$ followed by solving both goals 
+%using rewrite steps combined with an arithmetic decision procedure. 
+%Both \tactictoe proofs (NH and SH) follow the general idea of reasoning by 
+%induction but solve 
+%the base case and the inductive case in a different way.
+%The base case only needs rewriting using the global simpset in the 
+%\tactictoe(NH)
+%proof, which is simulated by a call to \metis in the (SH) proof.
+%The inductive case should require an arithmetic decision procedure as hinted 
+%by 
+%the human 
+%proof. This is achieved by rewriting using an arithmetic simpset in the second 
+%proof. In the first proof however, a rewriting step and case splitting step 
+%were used
+%to arrive at a point where \metis calls succeed.
+%The tactic proof produced by \tactictoe(NH) often looks better than the one 
+%discovered by \tactictoe(SH) in that it does not involve \metis calls with a 
+%large numbers of premises.
 

### 2018-01-11 tactictoe_jv.tex
table of contents: for organizing

@@
 %
-
+\vspace{100mm}
+
+\tableofcontents
+
+\vspace{100mm}
 \todo{These 2 paragraphs are copied and slightly edited}
@@
 
-A first successful use of orthogonalization techniques for theorems appeared in
-a work on lemma mining~\cite{ckju-jsc15} in \hollight. Theorems were mined from 
-proofs recorded at the kernel level. Because the number of recorded lemmas was 
-in millions, page ranking techniques was used as a filter to extract ``useful'' 
-lemmas.
-
 \section{Proof Recording}\label{sec:recording}
@@
 
-\section{Tactic Abstraction}\label{sec:synthesis}
+\section{Abstraction}\label{sec:synthesis}
 One of the great weakness of the previous version of \tactictoe is that
@@
 
-The first necessary adaptation of the algorithm stems
-from the fact that a proof is in general not a
-path but a tree. This means that our search space has two branching factors: 
+(todo: missing link)
+A proof is in general not a path but a tree. This means that our search space 
+has two branching factors: 
 the choice of a tactic, and the number of goals produced by each tactic.
@@
 \node [above of=0,node distance=2cm] (2) {.\ .\ .};
-\node [left of=2, tcircle] (2l) {$g_2^1$};
-\node [tcircle,right of=2] (2r) {$g_2^p$};
+\node [left of=2, tcircle] (2l) {$g_j^1$};
+\node [tcircle,right of=2] (2r) {$g_j^p$};
 \node [draw, fit=(2l) (2r)] (2b) {};
 
-\node [tcircle, left of=2l, node distance=2cm] (1r) {$g_1^q$};
+\node [tcircle, left of=2l, node distance=3cm] (1r) {$g_1^q$};
 \node [left of=1r] (1) {.\ .\ .};
@@
 
-\node [right of=2r,node distance=2cm] (3l) {$\phantom{g_3^p}$};
+\node [right of=2r,node distance=3cm] (3l) {$\phantom{g_3^p}$};
 \node [right of=3l] (3) {\phantom{.\ .\ .}};
@@
 
+\node [fit=(1r) (2l)] (1m) {}; 
+\node [fit=(2l) (3r)] (2m) {};
+
 \draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
-\draw[-to,thick] (0) to node[xshift=-10](t2){$t_2$} (2b);
+\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
+\draw[-to,dotted,thick] (0) to (1m);
+\draw[-to,dotted,thick] (0) to (2m);
 \draw[-to,dotted,thick] (0) to (3b);
@@
 \end{center}
-\caption{Choice of a goal and a tactic}
+\caption{The two branching factors: choice of a goal $g_0^i$ and choice of a 
+tactic $t_j$.}
 \label{fig:choice}
@@
 
-\section{ATP and Lemma Prediction Integration}
+\section{ATP Integration}
 
@@
 minimum), we will also 
-experiment with a ``small hammer'' tactic comprised of: a faster premise 
+experiment with a small hammer tactic comprised of: a faster premise 
 selection algorithm from previously proven theorems $\mathfrak{T}$ and
-a short call to the internal prover \metis~\cite{metis}.
-
-It is best tactic available but we do not have any data about it since it does 
-not appear in any proofs. So as a first approximation, we 
-artificially give it the best prediction scores on any goal and the best 
-coverage.
-
-\paragraph{Effect on orthogonalization of tactics}
-During orthogonalization, the general purpose tactic will be the first 
-competitor and if it manages to prove the goal it will subsume the original 
-tactic on this goal since the empty set of goals is a subset of any set. 
-Therefore, it will replace the original tactic in the feature vector that is 
-competed for. This is desirable since we would like to remove alternative 
-proofs.
-
-\paragraph{Orthogonalization of theorems}
-A ``hammer'' tactic can be useful to remove parallel
-theorems. We consider a theorem to be parallel if it can be proven 
-from previous theorems using a general tactic with a timeout equal or less 
-than the one it is given during the proof search. Parallel theorems are marked 
-and removed from the set of theorems $\mathfrak{T}$ from 
-which premise selection is performed.
-
-\begin{remark}
-The ``smaller hammer'' is chosen to remove parallel theorems and as the first 
-try during orthogonalization of tactics. 
-\end{remark}
-
-\paragraph{Effect on the proof search}
-First, before the proof search, we preselect 500 theorems 
-for the whole proof search tree using a premise selection algorithm 
-that takes into account dependencies between theorems.
-During the proof search, when a new goal is created or a fresh pending goal is 
-considered, the ``hammer'' tactic is always tried first because if it succeeds 
-the node would be closed rapidly.
-
-\paragraph{Parallelization of external calls}
-It is possible to parallelize calls to external provers so that they happen 
-asynchronously during the proof search. We let the prover runs \tactictoe 
-continue its proof in parallel of a ``hammer'' tactic call. Additional care 
-needs to be taken after each proof step to close proofs solved by the general 
-tactic. When the number of thread reaches a limit, a ``hammer'' tactic call 
-should
-wait for a thread to be freed but the proof search goes on and the feedback of 
-the ``hammer'' tactic is delayed.
+a short call (0.1 seconds) to the internal prover \metis~\cite{metis} or a
+longer call to external prover \eprover.
+
+We construct this type of tactic manually but in the case of \metis it could be 
+rediscovered automatically as it is essentially a call to \metis where its 
+theorem list argument has been abstracted.
+
+Since we do not always have the data to learn on which goal this general 
+tactics are succesful, we assume as a first approximation that they are the 
+strongest all around. And we try them first on each node during the proof 
+search with the advantage that if it succeeds the node is closed rapidly in the 
+case of \metis.
+
+To integrate a general tactic that requires a higher time out such as \eprover, 
+we make use of asynchronous calls to prevent the ATP 
+from slowing down the search.  The number of parallel threads can be capped.
+And if tactictoe or an asynchronous thread proves a goal, all descendants of 
+these goal are deactivated terminating associated threads. This frees space for 
+more pending threads associated to newly created nodes to start.
+
+%needs to be taken after each proof step to close proofs solved by \eprover
+%When the number of thread reaches a limit, a ``hammer'' tactic call 
+%should
+%wait for a thread to be freed but the proof search goes on and the feedback of 
+%the ``hammer'' tactic is delayed. 
+%
+%
+%It is best tactic available but we do not have any data about it since it does 
+%not appear in any proofs. So , we 
+%artificially try it on
+
+%\paragraph{Effect on orthogonalization of tactics}
+%During orthogonalization, the general purpose tactic will be the first 
+%competitor and if it manages to prove the goal it will subsume the original 
+%tactic on this goal since the empty set of goals is a subset of any set. 
+%Therefore, it will replace the original tactic in the feature vector that is 
+%competed for. This is desirable since we would like to remove alternative 
+%proofs.
+
+%\paragraph{Orthogonalization of theorems}
+%A ``hammer'' tactic can be useful to remove parallel
+%theorems. We consider a theorem to be parallel if it can be proven 
+%from previous theorems using a general tactic with a timeout equal or less 
+%than the one it is given during the proof search. Parallel theorems are marked 
+%and removed from the set of theorems $\mathfrak{T}$ from 
+%which premise selection is performed.
+%
+%\begin{remark}
+%The ``smaller hammer'' is chosen to remove parallel theorems and as the first 
+%try during orthogonalization of tactics. 
+%\end{remark}
+
+%\paragraph{Effect on the proof search}
+%First, before the proof search, we preselect 500 theorems 
+%for the whole proof search tree using a premise selection algorithm 
+%that takes into account dependencies between theorems.
+%During the proof search, when a new goal is created or a fresh pending goal is 
+%considered, the ``hammer'' tactic is always tried first because if it succeeds 
+%the node would be closed rapidly.
+
+%\paragraph{Parallelization of external calls}
+%It is possible to parallelize calls to external provers so that they happen 
+%asynchronously during the proof search. We let the prover runs \tactictoe 
+%continue its proof in parallel of a ``hammer'' tactic call.
 
@@
 
-\subsubsection{Tactic synthesis}
-Here, we test the benefit of tactic synthesis compared to the old version.
+\subsubsection{Abstraction}
+Here, we test the benefit of different kind of tactic abstraction. 
 In all those experiments orthogonalization is performed with radius 20 but not 
@@
 
-\subsubsection{Hammer parameters}
-
-The third experiment, presented in Table~\ref{tab:lh_param},
-evaluates the effect of integrating the ``small hammer'' in the \tactictoe 
-search.
-At a first glance, the increased success rate is significant for all tested 
-parameters. 
-Further analysis reveals that increasing the number of premises from 8 to 16 
-with a 
-timeout of 0.02 seconds is detrimental. The $D_{19}$ experiment  demonstrates 
-that 0.1 
-seconds is a better time limit for reasoning with 16 premises. And the $D_{18}$ 
-experiment reveals the disadvantage of unnecessarily increasing the timeout of 
-\metis. This reduces the time available for the rest of the proof search, which 
-makes the success rate drop.
+\subsubsection{ATP integration}
+
+%The third experiment, presented in Table~\ref{tab:lh_param},
+%evaluates the effect of integrating the ``small hammer'' in the \tactictoe 
+%search.
+%At a first glance, the increased success rate is significant for all tested 
+%parameters. 
+%Further analysis reveals that increasing the number of premises from 8 to 16 
+%with a 
+%timeout of 0.02 seconds is detrimental. The $D_{19}$ experiment  demonstrates 
+%that 0.1 
+%seconds is a better time limit for reasoning with 16 premises. And the 
+%$D_{18}$ 
+%experiment reveals the disadvantage of unnecessarily increasing the timeout of 
+%\metis. This reduces the time available for the rest of the proof search, 
+%which 
+%makes the success rate drop.
+
+The best setting for \metis was to be in previous experiments to be 16
+theorems and
 
@@
 
+Analysing failed proof attempt can give us an idea of the behaviour of the 
+algorithm. In a proof search of 10 seconds, in average 500 (todo: real 
+estimation) tactics are applied.
+The productive tactics create an average of 100 (todo: real estimation) nodes. 
+This means that in this settings 1 out of 5 tactics are productive.
+
+Increasing the timeout to more than 10 seconds would probably be beneficial as 
+we can see that the slope of the curve is far for being flat in Fig~..
+
 \pgfplotscreateplotcyclelist{my black}{
@@
 dashed, mark repeat=100, mark phase=0, black!100\\
-solid, mark repeat=20, mark phase=10, black!100, mark=triangle*\\
-solid, mark repeat=20, mark phase=10, black!75, mark=square*\\
-solid, mark repeat=20, mark phase=10, black!50, mark=*\\
-solid, mark repeat=20, mark phase=10, black!50, mark=o\\
 }
 
-\begin{figure}[p]
+\begin{figure}[h]
 \centering           
@@
 
-\begin{figure}[p]
+(Todo: change next figure to an histogram)
+\begin{figure}[h]
 \centering           
@@
 
-\begin{table*}[t]
-\centering\ra{1.3}
-\small
-\begin{tabular}{llllll}
-\toprule
- ID & \multicolumn{2}{c}{nodes}   & \multicolumn{2}{c}{proof size} & 
- \multicolumn{1}{c}{time}\\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} 
- \cmidrule(lr){6-6}
-     & average & max & average & max & average \\
-\midrule
- \tactictoe & 94.66 & 421 & 3.34 & 39 & 0.66 \\
-\bottomrule
-\end{tabular}
- \caption{\label{tab:stats} old search statistics}
-\end{table*}
+%\begin{table*}[t]
+%\centering\ra{1.3}
+%\small
+%\begin{tabular}{llllll}
+%\toprule
+% ID & \multicolumn{2}{c}{nodes}   & \multicolumn{2}{c}{proof size} & 
+% \multicolumn{1}{c}{time}\\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} 
+% \cmidrule(lr){6-6}
+%     & average & max & average & max & average \\
+%\midrule
+% \tactictoe & 94.66 & 421 & 3.34 & 39 & 0.66 \\
+%\bottomrule
+%\end{tabular}
+% \caption{\label{tab:stats} old search statistics}
+%\end{table*}
 
@@
 
+Future work:
+  Promising techniques that could not be evaluated. Orthogonalization of 
+  theorems~\cite{ckju-jsc15}. Term predictions.
+
+Cite in future works.
+%A first successful use of orthogonalization techniques for theorems appeared in
+%a work on lemma mining~\cite{ckju-jsc15} in \hollight. Theorems were mined 
+%from 
+%proofs recorded at the kernel level. Because the number of recorded lemmas was 
+%in millions, page ranking techniques was used as a filter to extract 
+%``useful'' 
+%lemmas.
+
 \todo{is copied (modified a bit)}

### 2018-01-12 tactictoe_jv.tex
better affiliations

@@
 
-\institute{Cezary Kaliszyk \at
+\institute{Thibault Gauthier and Cezary Kaliszyk \at
 Department of Computer Science, University of Innsbruck,
 %Technikerstr. 21a/2,
-Innsbruck, Austria\\ \url{bla@bla}
+Innsbruck, Austria\\ \url{{thibault.gauthier,cezary.kaliszyk}@uibk.ac.at}
 \and
-Karol Pak \at Institute of Informatics, University of Bialystok, Bialystok, Poland\\\url{bla@bla}}
+Josef Urban \at Czech Technical University, Prague\\\url{josef.urban@gmail.com}
+\and Ramana Kumar and Michael Norrish \at Data61}
 

### 2018-01-12 tactictoe_jv.tex
reformulations in the abstract

@@
 
+\todo{JU: Still feels more like an intro than an abstract. Could be shorter and punchier.}
 \begin{abstract}
-Formalization of mathematical statements in interactive theorem provers (ITPs) 
-relies nowadays on many methods to automate the construction of a proof.
-Some of this method are general purpose and with enough time can prove any 
-theorems. Yet, in a formalized proof script 
-a mathematician take advantages of a combination of specialized rules, 
-theory-based strategy and general ones. All of them are implemented as tactics 
-in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
+Formalization of mathematics in interactive theorem provers (ITPs) 
+relies today on many proof automation methods.
+Some of these methods are general-purpose and with enough time they can prove any 
+theorem. Many formalized proof scripts however also
+take advantage of specialized rules and theory-based strategies. 
+In the HOL4 ITP system, practically all of them are implemented as \emph{tactics}.
+
+Automated theorem provers (ATPs) rely on parameterizable general 
 strategies. Sometimes, a strategy can be specialized for only one theory as in 
@@
 We bridge this gap by implementing our tactical prover TacticToe on top 
-of the ITP HOL4. 
-After learning in which situations tactics are effective from 
-tactic-level proof and what kind of formulas are the results of effective 
-steps, we implement a Monte Carlo tree search algorithm that explores the most 
-promising tactic-level proof paths. Experiments on the \holfour standard 
-library shows that TacticToe is able to re-prove 59\% of 7088 theorems in less 
-than 10 seconds where as the ATP Eprover with auto-schedule
-solves 29\%.
+of the HOL4 ITP. 
+TacticToe first learns from many proof scripts
+the relevance of tactics in particular proof situations.
+% are effective from 
+% tactic-level proofs and what kind of formulas are the results of effective 
+% steps, 
+This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths. 
+Experiments on the \holfour standard 
+library show that TacticToe is able to re-prove 59\% of the 7088 theorems in less 
+than 10 seconds compared to 29\% proved by the HolyHammer system using the E ATP in its auto-schedule mode.
 

### 2018-01-16 tactictoe_jv.tex
Adding TODOs to the paper

--- 2018-01-11 tactictoe_jv.tex
@@
 
-\begin{abstract}
+\begin{abstract}\todo{CK review + choose title}
 Formalization of mathematical statements in interactive theorem provers (ITPs) 
@@
 
-\section{Introduction}
+\section{Introduction}\todo{TG + CK review}
 %\begin{example} Proof automatically generated by \tactictoe for the user given 
@@
 \vspace{100mm}
-\todo{These 2 paragraphs are copied and slightly edited}
+\todo{TG : comparison with hammers and advantages for the user and over hammers}
 Many of the state-of-the-art interactive theorem provers (ITPs) such as
@@
 
+\todo{TG: rewrite the plan}
   To do this, we \emph{extract and record} tactic invocations from the ITP 
@@
 
+\todo{CK review}
 \subsection{Contributions}
@@
 recording steps (parsing, unfolding and replaying) are regrouped into a single 
-\sml function. Intermediate files storing the result of each step are not 
-necessary anymore. So, our recording is more complete and fully 
-automatized.
+\sml function. So, our recording is more complete and fully automatized.
 \item Existing tactic orthogonalization on recorded data is made stronger by 
@@
 \item Two types of arguments (list of theorems and terms) can be predicted 
-independently. Combinatorial explosion in the number of tactics 
+independently of the tactic. Combinatorial explosion in the number of tactics 
 is prevented through the introduction of abstracted tactics. As desired, their 
@@
 algorithm. MCTS gives a more balanced feedback to guide proof search by 
-comparing subtrees instead of leafs. (todo: A first successful evaluation 
-function.)
+comparing subtrees instead of leafs. The policy and evaluation are learned 
+through supervised learning.
 \item The internal ATP \metis used during our proof search is complemented by 
@@
 
-\section{Related Work}\todo{reorganise}
+\section{Related Work}\todo{TG reorganise + other motivation?}
 There are several essential components of our work that are comparable to 
@@
 \section{Proof Recording}\label{sec:recording}
+\todo{MN + RK: comment technical and hard-to read presentation of the unfolding
+mechanism and replace by a high level abstraction. Alternatively, make the 
+current version more accessible.}
+
 Recording proofs in an LCF-style proof assistant can be done at different levels.
@@
 original theory file containing proof scripts into an
-abstract syntax tree (AST). The AST is then transformed into a globalized AST
+abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
+human ==> better for the machine learning algorithm (learn the right thing)}
+
+ The AST is then transformed into a globalized AST
 where every top-level value is made independent from
-the context. \todo{some reason why this is needed? nice recording level} We 
+the context. \todo{some reason why this is needed? easy transfer between 
+theories} We 
 split the proofs into a series of tactic units and wrap them 
@@
 
-\section{Prediction}\label{s:prediction}
+\section{Prediction}\label{s:prediction}\todo{TG: complete term theorem and 
+list of goals specificities in terms of predictions and justify by their use 
+case}
 
@@
 
-\section{Orthogonalization}
+\section{Orthogonalization}\todo{MN + RK: review + fix typos}
 
@@
 
-\section{Abstraction}\label{sec:synthesis}
+\section{Abstraction}\label{sec:synthesis}\todo{MN + RK: review + fix typos}
 One of the great weakness of the previous version of \tactictoe is that
@@
 \section{Proof Search}\label{sec:proofsearch}
+\todo
+{
+TG:
+describe the prior policy, the evaluation, 
+the node selection algorithm,
+the node extension (+ case of solved goal) + backpropagation.
+add description of end of the algorithm.
+}
 Despite the best efforts of the prediction algorithms, the selected tactic may 
@@
 
-(todo: missing link)
+\todo{TG: add transition}
 A proof is in general not a path but a tree. This means that our search space 
@@
 
+\todo{TG: add general overview of the algorithm and describe the plan}
 If we suppose that each tactic produces at most one goal, then the standard 
@@
 
-%Search Algorithm AlphaGo Zero uses a much simpler variant of the asynchronous 
-%policy and
-%value MCTS algorithm (APV-MCTS) used in AlphaGo Fan and AlphaGo Lee.
-%Each node s in the search tree contains edges (s, a) for all legal actions a ∈ 
-%A(s). Each edge
-%stores a set of statistics,
-%{N (s, a), W (s, a), Q(s, a), P (s, a)},
-%where N (s, a) is the visit count, W (s, a) is the total action-value, Q(s, a) 
-%is the mean action-value,
-%and P (s, a) is the prior probability of selecting that edge. Multiple 
-%simulations are executed in
-%parallel on separate search threads. The algorithm proceeds by iterating over 
-%three phases (a–c in
-%Figure 2), and then selects a move to play (d in Figure 2).
-%25Select (Figure 2a). The selection phase is almost identical to AlphaGo Fan 
-%12 ; we recapitulate
-%here for completeness. The first in-tree phase of each simulation begins at 
-%the root node of the
-%search tree, s 0 , and finishes when the simulation reaches a leaf node s L at 
-%time-step L. At each
-%of these time-steps, t < L, an action is selected according to the statistics 
-%in the search tree,
-%
-%a t = argmax Q(s t , a) + U (s t , a) , using a variant of the PUCT algorithm 
-%24 ,
-%a
-%pP
-%U (s, a) = c puct P (s, a)
-%N (s, b)
-%1 + N (s, a)
-%b
-%where c puct is a constant determining the level of exploration; this search 
-%control strategy initially
-%prefers actions with high prior probability and low visit count, but 
-%asympotically prefers actions
-%with high action-value.
-%Expand and evaluate (Figure 2b). The leaf node s L is added to a queue for 
-%neural network
-%evaluation, (d i (p), v) = f θ (d i (s L )), where d i is a dihedral 
-%reflection or rotation selected uniformly
-%at random from i ∈ [1..8].
-%Positions in the queue are evaluated by the neural network using a mini-batch 
-%size of 8; the
-%search thread is locked until evaluation completes. The leaf node is expanded 
-%and each edge (s L , a)
-%is initialised to {N (s L , a) = 0, W (s L , a) = 0, Q(s L , a) = 0, P (s L , 
-%a) = p a }; the value v is then
-%backed up.
-%Backup (Figure 2c). The edge statistics are updated in a backward pass through 
-%each step
-%t ≤ L. The visit counts are incremented, N (s t , a t ) = N (s t , a t ) + 1, 
-%and the action-value is updated
-%to the mean value, W (s t , a t ) = W (s t , a t ) + v, Q(s t , a t ) =
-%W (s t ,a t )
-%.
-%N (s t ,a t )
-%We use virtual loss to ensure
-%each thread evaluates different nodes 69 .
-
 \subsection{A proof tree and solved node}
@@
 
-\subsection{Optimization}
+\subsection{Node extension}
+
+To prevent tactics from looping, we add timeout of 0.02 seconds to all tactics 
+(optimized from previous paper)
 
@@
 searching in this direction. Theoretically, an interesting list of goal is one 
-that has the shortest proof, and we could measure the likelyhood of being 
+that has the shortest proof, and we could measure the likelihood of being 
 provable in a number of steps from previous proof search. However, this 
-approach produces a lot of negatives. So to reduce this number, we limit 
+approach produces a lot of negatives\todo{TG: cite Hard Negative mining (cite 
+Deep math)}. So to reduce this number, we limit 
 ourselves to feature vectors tested during orthogonalization. We declare a list 
@@
 
-\section{ATP Integration}
+\section{ATP Integration}\todo{JU + CK: rework and review}
 
@@
 external calls to \holyhammer are computationally expensive (5 seconds at a 
-minimum), we will also 
+minimum), we  
 experiment with a small hammer tactic comprised of: a faster premise 
 selection algorithm from previously proven theorems $\mathfrak{T}$ and
-a short call (0.1 seconds) to the internal prover \metis~\cite{metis} or a
+a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
+internal prover \metis~\cite{metis} or a
 longer call to external prover \eprover.
@@
 And if tactictoe or an asynchronous thread proves a goal, all descendants of 
-these goal are deactivated terminating associated threads. This frees space for 
-more pending threads associated to newly created nodes to start.
+these goal are deactivated, terminating associated threads. This frees space 
+for more pending threads, associated to newly created nodes, to start.
 
@@
 
-\section{Experimental Evaluation}\label{s:experiments}
+\section{Experimental Evaluation}\label{s:experiments}\todo{TG: introduction + 
+consistent tables}
 
 Disclaimer:
-Comparing results from different tables is possible but not advisable since 
-conditions of the experiments may vary slightly because of various factors such 
+Conditions of the experiments may vary slightly between different tables 
+because of various factors such 
 as overhead cost of additional features or the fact that they were run on 
@@
 
-\subsection{Methodology} \todo{is copied}
+\subsection{Methodology} \todo{TG: is copied, update references of sections and
+new algorithm + mention when orthogonalization is done}
+
 The evaluation imitates the construction of the library: For each theorem only 
-the previous
-human proofs are known. These are used as the learning base for the predictions.
+the previous human proofs are known. These are used as the learning base for 
+the predictions.
 To achieve this scenario we re-prove all theorems during a modified build of 
 \holfour.
-As theorems are proved, they are recorded and included in the training examples.
+As theorems are proved, their tactical proofs and their statements are recorded 
+and included in the 
+training examples.
 For each theorem we first attempt to run the \tactictoe search with a time 
@@
 
-\subsubsection{Feature parameters}
+\subsubsection{Prediction parameters}\todo{TG: remove this section or mention
+that parameters were optimized in the previous version}
 The first experiment concerns the choice of the right kind of features and 
@@
 %\end{table}
+
+\subsubsection{MCTS parameters: policy and evaluation}
 
@@
 
-\section{Proof Presentation}\label{sec:proofdisplay}
+\section{Proof Presentation}\label{sec:proofdisplay}\todo{MN + RK: add 
+motivations and transitions + review}
 
@@
 
-\section{Case study}
+\section{Case study}\todo{TG: rerun experiment with prettification and 
+minimization + add another example. 
+  TG + RK: describe the two examples + find examples that did not appear during 
+  evaluation
+  RK: After implementation and small experiment is done (in about 1 week) 
+      Test tactictoe on new goals and manual exploration mode.
+      Give feedback.
+}
 %Investigating further the different qualities of \tactictoe, we study its 
@@
 
-\section{Conclusion}\label{sec:concl}
+\section{Conclusion}\label{sec:concl}\todo{TG + JU review}
 
 Future work:
-  Promising techniques that could not be evaluated. Orthogonalization of 
-  theorems~\cite{ckju-jsc15}. Term predictions.
-
-Cite in future works.
+  Promising techniques that could not be evaluated. 
+  + Orthogonalization of theorems~\cite{ckju-jsc15}. 
+  + Try harder on term predictions. \cite{latest version of sepia}.
+
 %A first successful use of orthogonalization techniques for theorems appeared in

### 2018-01-16 tactictoe_jv.tex

@@
 
-\begin{abstract}\todo{CK review + choose title}
-Formalization of mathematical statements in interactive theorem provers (ITPs) 
-relies nowadays on many methods to automate the construction of a proof.
-Some of this method are general purpose and with enough time can prove any 
-theorems. Yet, in a formalized proof script 
-a mathematician take advantages of a combination of specialized rules, 
-theory-based strategy and general ones. All of them are implemented as tactics 
-in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
+%
+%\begin{abstract}
+%Formalization of mathematical statements in interactive theorem provers (ITPs) 
+%relies nowadays on many methods to automate the construction of a proof.
+%Some of this method are general purpose and with enough time can prove any 
+%theorems. Yet, in a formalized proof script 
+%a mathematician take advantages of a combination of specialized rules, 
+%theory-based strategy and general ones. All of them are implemented as tactics 
+%in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
+%=======
+\todo{JU review: Still feels more like an intro than an abstract. Could be 
+shorter and punchier.}
+\begin{abstract}
+Formalization of mathematics in interactive theorem provers (ITPs) 
+relies today on many proof automation methods.
+Some of these methods are general-purpose and with enough time they can prove any 
+theorem. Many formalized proof scripts however also
+take advantage of specialized rules and theory-based strategies. 
+In the HOL4 ITP system, practically all of them are implemented as \emph{tactics}.
+
+Automated theorem provers (ATPs) rely on parameterizable general 
 strategies. Sometimes, a strategy can be specialized for only one theory as in 
@@
 
-\section{Introduction}\todo{TG + CK review}
+\section{Introduction}\todo{TG + JU + CK}
 %\begin{example} Proof automatically generated by \tactictoe for the user given 
@@
 
-\todo{CK review}
+\todo{JU or CK review}
 \subsection{Contributions}
@@
 
-\section{Conclusion}\label{sec:concl}\todo{TG + JU review}
+\section{Conclusion}\label{sec:concl}\todo{TG + JU or CK review}
 

### 2018-01-16 tactictoe_jv.tex
Proof presentation: update

@@
 \vspace{100mm}
-
 \tableofcontents
@@
 vectors containing only tactics 
-from the proof. To obtain a shorter proof, the diagonalization coefficient is
-reduced to encourage exploration in the width of the proof tree. Similarly by 
-ponderating each tactic application by its time and setting the 
-diagonalization  coefficient is set to a tenth of the timeout of a tactic, we 
-can aim for faster proofs.
+from the proof. To obtain a shorter proof, the exploration coefficient should be
+increased and the policy coefficient lowered to encourage exploration in the 
+width of the proof tree. Additionally, by giving a weaker evaluation to slow 
+tactics, we can aim for faster proofs.
 
@@
 One of the argument that may contain unnecessary information are lists.
-We test if the minimized tactic has the same effect as the original one.
+Let $t$ be a tactic applied to a goal $g$ containing a list $l$ as argument. 
+And let $t'$ be the tactic $t$ where one element $e$ of $l$ as be removed. If 
+$t$ and $t'$ have the same effect on $g$ then $t'$ can replace $t$ in the final 
+proof.
+
 This is a generalization of the simplest method used for minimizing a list of 
-theorems in ``hammers'' \cite{}. The fact that the lists are generally small 
+theorems in ``hammers'' \cite{}. The fact that lists are generally small
 make this approach feasible.

### 2018-01-16 tactictoe_jv.tex
Proof presentation: update

@@
 \section{Proof Presentation}\label{sec:proofdisplay}\todo{MN + RK: add 
-motivations and transitions + review}
+motivations and transitions}
+
+Motivation: 
++ minimal effort to integrate the tactictoe results in proof scripts.
+
+Side effect:
++ could be also used to simplify existing proof scripts.
 

### 2018-01-17 tactictoe_jv.tex
tactictoe: case study examples

@@
 {columns=fullflexible,
-  keywords={THEN,val,open,store_thm,fun,fn},%
+  keywords={THEN,THENL,val,open,store_thm,fun,fn},%
   frame=none,
@@
    {->}{{$\rightarrow$}}1%
-   {=>}{{$\Rightarrow$}}1%
+   {==>}{{$\Rightarrow$}}1%
+   {wedge}{{$\wedge$}}1%
+   {>=}{{$\geq$}}1%
    {'a}{{$\alpha$}}1%
@@
    {->}{{$\rightarrow$}}1%
-   {=>}{{$\Rightarrow$}}1%
+   {==>}{{$\Rightarrow$}}1%
+   {wedge}{{$\wedge$}}1%
+   {<=}{{$\leq$}}1%
+   {>=}{{$\geq$}}1%
    {'a}{{$\alpha$}}1%
@@
 %
-\vspace{100mm}
-\tableofcontents
+\vspace{100mm}\tableofcontents
 
@@
 
-\begin{table}[b!]
+\begin{table}[]
 \centering
@@
 
-(Todo: change next figure to an histogram)
 \begin{figure}[h]
@@
 
-\section{Case study}\todo{TG: rerun experiment with prettification and 
-minimization + add another example. 
-  TG + RK: describe the two examples + find examples that did not appear during 
-  evaluation
+\section{Case study}\todo{
+  TG + RK: describe the two examples.
   RK: After implementation and small experiment is done (in about 1 week) 
-      Test tactictoe on new goals and manual exploration mode.
+      Test tactictoe find new examples on new goals and manual exploration mode.
       Give feedback.
@@
 
+%The term and the conversion rules were not synthesized but simply copied from 
+%previous 
+%proofs.
+
+%SURJ_INJ_INV
+%proof length: 2
+%proof time: 0.050853
+%Statistics
+%  infstep : 31
+%  nodes   : 11
+%  maxdepth: 3
+%Time: 1.300789
+%  inferstep time: 1.071869
+%  node_find time: 0.018046
+%  node_crea time: 0.147604
+%    pred time: 0.04263
+%    thmpred time: 0.000005
+%    mc time: 0.071282
+%    inst time: 0.0
+%Proof found: 
+%Org tac number: 15
+%Replaying proof: 0.562829
+
 %\begin{remark}
@@
 
-%SURJ_INJ_INV
-%proof length: 7
-%proof time: 0.092014
+\newcommand{\wed}{\begin{math}\wedge\end{math}}
+
+Proofs discovered by \tactictoe discovered during evaluation.
+
+for theorem \texttt{SURJ\_INJ\_INV} In 
+theory \texttt{pred\_set})
+
+\begin{example}\label{ex:casestudy}()
+
+%val SURJ_INJ_INV = store_thm(
+%  "SURJ_INJ_INV",
+%  ``SURJ f s t ==> ?g. INJ g t s /\ !y. y IN t ==> (f (g y) = y)``,
+\vspace{5mm}
+\small
+
+Goal
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+!f s t. SURJ f s t ==> ?g. INJ g t s wedge !y. y IN t ==> (f (g y) = y)
+\end{lstlisting}
+
+Original \holfour proof script
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+REWRITE_TAC [IMAGE_SURJ] THEN
+  DISCH_TAC THEN Q.EXISTS_TAC `THE o LINV_OPT f s` THEN
+  BasicProvers.VAR_EQ_TAC THEN REPEAT STRIP_TAC
+  THENL [
+  irule INJ_COMPOSE THEN Q.EXISTS_TAC `IMAGE SOME s` THEN
+    REWRITE_TAC [INJ_LINV_OPT_IMAGE] THEN REWRITE_TAC [INJ_DEF, IN_IMAGE] THEN
+    REPEAT STRIP_TAC THEN REPEAT BasicProvers.VAR_EQ_TAC THEN
+    FULL_SIMP_TAC std_ss [THE_DEF],
+  ASM_REWRITE_TAC [LINV_OPT_def, o_THM, THE_DEF] THEN
+    RULE_ASSUM_TAC (Ho_Rewrite.REWRITE_RULE
+      [IN_IMAGE', GSYM SELECT_THM, BETA_THM]) THEN ASM_REWRITE_TAC [] ]
+\end{lstlisting}
+
+\vspace{5mm}
+
+\tactictoe proof script
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+SRW_TAC [] [SURJ_DEF, INJ_DEF] THEN METIS_TAC []
+\end{lstlisting}
+
+Goal
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+!l1 l2 n x. LENGTH l1 <= n ==> 
+  (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n - LENGTH l1) l2))
+\end{lstlisting}
+
+\vspace{5mm}
+
+Original \holfour proof script
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+  rw[] THEN simp[LIST_EQ_REWRITE] THEN Q.X_GEN_TAC `z` THEN
+  simp[EL_LUPDATE] THEN rw[] THEN simp[EL_APPEND2,EL_LUPDATE] THEN
+  fs[] THEN Cases_on `z < LENGTH l1` THEN
+  fs[] THEN simp[EL_APPEND1,EL_APPEND2,EL_LUPDATE]
+\end{lstlisting}
+
+%val LUPDATE_APPEND2 = Q.store_thm("LUPDATE_APPEND2",
+%   `!l1 l2 n x.
+%      LENGTH l1 <= n ==>
+%      (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n-LENGTH l1) l2))`,
+%);
+\vspace{5mm}
+
+\tactictoe proof script
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+Induct_on `l1` THENL [SRW_TAC [] [],
+  Cases_on `n` THENL [SRW_TAC [] [],
+    FULL_SIMP_TAC (srw_ss ()) [] THEN METIS_TAC [LUPDATE_def]]]
+\end{lstlisting}
+
+If a discovered proof is reasonable short, fast and understandable, they could 
+replace its original in the \holfour standard library.
+
+%
+%LUPDATE_APPEND2
+%proof length: 6
+%proof time: 0.017156
 %Statistics
-%  infstep : 500
-%  nodes   : 132
-%  maxdepth: 7
-%Time: 9.574603
-%  inferstep time: 7.287698
-%  node_find time: 0.135171
-%  node_crea time: 1.674807
-%    pred time: 0.335816
-%    thmpred time: 0.000018
-%    mc time: 0.704148
+%  infstep : 70
+%  nodes   : 27
+%  maxdepth: 4
+%Time: 3.02716
+%  inferstep time: 2.12842
+%  node_find time: 0.07146
+%  node_crea time: 0.508384
+%    pred time: 0.164047
+%    thmpred time: 0.000012
+%    mc time: 0.265314
 %    inst time: 0.0
-
-\newcommand{\wed}{\begin{math}\wedge\end{math}}
-
-\begin{example}\label{ex:casestudy}(
-Proof discovered by \tactictoe for theorem \texttt{SURJ\_INJ\_INV} In 
-theory \texttt{pred\_set})
-\small
-\begin{center}
-\begin{alltt}
-    SRW_TAC [] [auto] 
-  THEN 
-    POP_ASSUM (MP_TAC o 
-      CONV_RULE (DEPTH_CONV RIGHT_IMP_EXISTS_CONV 
-                 THENC SKOLEM_CONV THENC REWRITE_CONV [EXISTS_DEF] 
-                 THENC DEPTH_CONV BETA_CONV)) 
-  THEN
-    Q.SPEC_TAC (`@y. !x. x IN t ==> y x IN s \wed (f (y x) = x)`,`g`) 
-  THEN
-    SRW_TAC [] [INJ_DEF] 
-  THEN
-    Q.EXISTS_TAC `g` 
-  THEN
-    SRW_TAC [] [INJ_DEF] 
-  THEN
-    METIS_TAC [auto]
-
-\end{alltt} 
-\end{center}
+%Proof found: 
+%Org tac number: 10
+%Original proof time: 0.0
+%Replaying proof: 0.470118
+
 \end{example}
-
-%The term and the conversion rules were not synthesized but simply copied from 
-%previous 
-%proofs.
 

### 2018-01-17 tactictoe_jv.tex
AITP integration: wip

@@
 
-A large difference between diagonalized search and our MCTS algorithm is in 
-what is considered to be the width of the tree. In diagonalized search, 
-non-productive tactics contribute to the extension of the width where as in 
-MCTSn the width includes only the total number of children produced. The 
-influence of non-productive on the proof exploration is recovered when we 
-include an 
-evaluation function that takes into account the productivity of tactics such as 
-$E_{productive}$ or $E_r$.
+%A large difference between diagonalized search and our MCTS algorithm is in 
+%what is considered to be the width of the tree. In diagonalized search, 
+%non-productive tactics contribute to the extension of the width where as in 
+%MCTSn the width includes only the total number of children produced. The 
+%influence of non-productive on the proof exploration is recovered when we 
+%include an 
+%evaluation function that takes into account the productivity of tactics such 
+%as 
+%$E_{productive}$ or $E_r$.
 
@@
 
-\section{ATP Integration}\todo{JU + CK: rework and review}
-
+\section{ATP Integration}\todo{TG + JU + CK: rework and review}
 General-purpose proof automation mechanisms which combine proof translation to
@@
 power of tactics, we would like to combine the \tactictoe search with 
-\holyhammer for \holfour~\cite{tgck-cpp15}. The principle is that we will 
-create a \holyhammer tactic that can be used during the proof search. Since 
-external calls to \holyhammer are computationally expensive (5 seconds at a 
-minimum), we  
+stronger general purpose provers such as the one in \holyhammer for 
+\holfour~\cite{tgck-cpp15}.
+
+During our experiments,
+We chose to distinguish two these general purpose tactic \metis and \eprover.
+\metis is integrated in \holfour, so can be call with frequently with low 
+timeout.
+And the strongest external provers according to past experiments \cite{hh4h4}.
+Where as...
+external calls to \eprover are computationally expensive (5 seconds at a 
+minimum +  128 premises), we  
 experiment with a small hammer tactic comprised of: a faster premise 
-selection algorithm from previously proven theorems $\mathfrak{T}$ and
+selection algorithm and
 a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
@@
 
-We construct this type of tactic manually but in the case of \metis it could be 
-rediscovered automatically as it is essentially a call to \metis where its 
+\paragraph{Prediction of general purpose tactics}
+Since we do not always have the data to learn on which goal this general 
+tactics are successful, we assume as a first approximation that they are the 
+strongest all around.
+
+And we try them first on each node during the proof search with the advantage 
+that if it succeeds the node is closed rapidly.
+
+This general tactics requires a list of relevant lemmas.
+
+This forced behavior could often be rediscovered from the training data for the 
+internal provers \metis by using theorem abstraction. essentially a call to 
+\metis where its 
 theorem list argument has been abstracted.
 
-Since we do not always have the data to learn on which goal this general 
-tactics are succesful, we assume as a first approximation that they are the 
-strongest all around. And we try them first on each node during the proof 
-search with the advantage that if it succeeds the node is closed rapidly in the 
-case of \metis.
-
+\paragraph{Application of general purpose tactics}
+The slightly higher timeout of 0.1 seconds given to \metis does not slowdown 
+the proof too much but...
 To integrate a general tactic that requires a higher time out such as \eprover, 
-we make use of asynchronous calls to prevent the ATP 
-from slowing down the search.  The number of parallel threads can be capped.
-And if tactictoe or an asynchronous thread proves a goal, all descendants of 
-these goal are deactivated, terminating associated threads. This frees space 
-for more pending threads, associated to newly created nodes, to start.
+we make use of asynchronous calls via PolyML.fork to prevent the ATP 
+from slowing down the search.
+
+%The number of parallel threads can be capped.
+%And if tactictoe or an asynchronous thread proves a goal, all descendants of 
+%these goal are deactivated, terminating associated threads. This frees space 
+%for more pending threads, associated to newly created nodes, to start.
 

### 2018-01-19 tactictoe_jv.tex
adding examples of generated tactictoe scripts
proof search: wip

@@
 Experiments on the \holfour standard 
-library show that TacticToe is able to re-prove 59\% of the 7088 theorems in less 
-than 10 seconds compared to 29\% proved by the HolyHammer system using the E ATP in its auto-schedule mode.
+library show that TacticToe is able to re-prove 59\% of 7088 theorems in less 
+than 10 seconds. This is to be compared with the 29\% proved by the ATP Eprover 
+in auto-schedule mode with 128 lemmas predicted and translated by HolyHammer.
 
@@
 
-\subsection{A proof tree and solved node}
+\subsection{A proof tree with two branching factors}
 
@@
 apply to this goal. Here the result produces new nodes each containing a list 
-of goals to be proven. If one of those list where empty, the goal $g_i^0$ would 
-be proved.
-
-A node is solved when every of its goal are solved.
-A goal is solved when every of its children node are solved.
-
-\subsection{Guiding the proof search}
+of goals to be proven.
+
+Supervised learning introduction.
+\subsection{Prior tactic policy}
+
+\subsection{Prior node evaluation}
+
+\subsection{Node selection}
 
@@
 
-\end{definition}
-
 The third point of the definition is a partial attempt to prevent confluent 
@@
 separated in our search tree.
+
+\paragraph{Solving a goal and a node}
+Describe the case where a proof for a goal in a node is found.
+If one of those list where empty, the goal $g_i^0$ would 
+be proved.
+A node is solved when every of its goal are solved.
+
+\subsection{End of the proof search}
+The proof search stops when one of these 3 conditions is realized:
+\item It found a proof (i.e. the root node is solved). In this case, 
+the search returns a minimized and 
+prettified proof script (see Section~\ref{sec:proofdisplay}).
+\item It has saturated. 
+There is no more tactics to be applied to any open goal. (todo: define open 
+goal somewhere).
+
+\item It exceeded a given time limit.
+
+The second case happens very rarely (less than 10 times when evaluating 7088 
+theorems) and only at the beginning of the library when the learned dataset is 
+small. So in our experiments, we just give the number of proof found relative 
+to the number of proof attempts.
+
+\end{definition}
 
@@
 
+\paragraph{Dataset}
+
+All top-level theorems in the standard library are considered except:
+\begin{itemize}
+\item the 619 calls to store\_thm that contains a let binding, these are 
+typically 
+harder for \tactictoe as it may not have seen the particular local declaration 
+before and therefore would have to find an alternative proofs. We also do not 
+learn from these proofs because they were explictly local.
+\item the 1175 calls to save\_thm. These are typically mostly trivial and easy 
+theorems.
+The remaining 7088 top-level theorems are evaluated in the full scale 
+evaluation and the level of difficulty of these problem is assessed by a 
+comparison with \eprover.
+\end{itemize}
+
+Among these types of theorems, every tenth theorems of the first third of the 
+library is evaluated during tuning, which give us a total of 273 theorems to 
+train different parameters on.
+
+\paragraph{Strategy selection}
 Although the training process in each strategy on its own is fair, the 
@@
 
+Constant progress: 
+  From 45 percent on the same dataset in 5 seconds 
+       59 percent in december in 10 seconds
+  Probably 65-70 percent en of January in 60 seconds.
+
+The progress cannot be attributed only to the time increase but to testing and 
+debugging of parameters.
+
 \section{Proof Presentation}\label{sec:proofdisplay}\todo{MN + RK: add 
@@
 
+%Name: 218 SURJ_INJ_INV
+%Original proof time: 0.002033
+
 %SURJ_INJ_INV
@@
 
-Original \holfour proof script
+Original \holfour proof script (2.03 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 
-\tactictoe proof script
+\tactictoe proof script (50.85 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 
-Original \holfour proof script
+Original \holfour proof script (62.69 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 
+%Name: 224 LUPDATE_APPEND2
+%Original proof time:
+
 %val LUPDATE_APPEND2 = Q.store_thm("LUPDATE_APPEND2",
@@
 
-\tactictoe proof script
+\tactictoe proof script (17.16 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 
-If a discovered proof is reasonable short, fast and understandable, they could 
-replace its original in the \holfour standard library.
+If a discovered proof is reasonable short, fast, maintainable  and 
+understandable, they could 
+replace its original in the \holfour standard library. In the first case the 
+tactictoe proof is slower but could still be considered becaues it is much more 
+readable. In the second case, it is faster and more readable.
 
@@
   + Orthogonalization of theorems~\cite{ckju-jsc15}. 
-  + Try harder on term predictions. \cite{latest version of sepia}.
+  + Try harder and smarted on term predictions. \cite{latest version of sepia}. 
+  \cite{conjecturing}
+
+More training than just orthogonalization of tactics.
 
@@
 Its implementation,
-\tactictoe, achieves an overall performance of 59\% theorems on most of the 
-\holfour standard library surpassing \eprover with auto-schedule. Its 
+\tactictoe, achieves an overall performance of 59\% theorems on 7088 theorems  
+of the \holfour standard library surpassing \eprover with 
+auto-schedule. Its 
 effectiveness is especially visible on 

### 2018-01-20 tactictoe_jv.tex
search: wip

@@
 
+Since there is a bijection between tactics of the tree search 
+and the node they 
+produced, any function define on tactics of the tree search can be map to a 
+function defined on nodes.
+
 Supervised learning introduction.
@@
 
+The estimation through the distance produced by the tactic selectors are hard 
+to translate into a prior tactic policy, so we rely only on the order of 
+tactics to estimate the probability with which a tactic should be tried.
+Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
+distance to a goal $g$. Let $c_policy$ be a constant heuristic that estimates 
+how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
+tactics $t_{i+1},\ldots,t_500$. We make the strong assumption that $c_policy$ 
+is constant across the search. We can now calculate the prior policy by:
+
+\[PriorPolicy(t_{i}) = (1 - c_policy)^{i} * c_policy\]
+
+The first term of the multiplication is the probability that none of the better 
+ranked tactic is selected.
+
+In fact, these probabilities are never used directly and only serves as a 
+guidance to our deterministic node selection algorithm over many simulations.
+
 \subsection{Prior node evaluation}
-
-\subsection{Node selection}
-
-A similar trade-off between exploration and exploitation exists in algorithm 
-like Monte Carlo. However, Monte Carlo tree search metrics are based on many 
-simulations and do not reflect our exploration strategy that only
-
-We present here a monte carlo tree search with no roll out which is essentially 
-the same as the one use by AlphaGo Zero~\cite{} originally defined by \cite{} 
-in \cite{}.
-
-\[Value(a_i) = Exploitation(a_i) + c_{exploration} * Exploration(a_i)\]
-
-\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{Policy(a_i)}\]
-
-\[Policy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(a_i))}}\]
-
-The exploitation function $Exploitation$ is the average of the evaluation of 
-all nodes descendant of $a_i$ in the current search tree.
-
-MCTS with prior policy
-This modification of the search tree can be simulated by weighting each arrow 
-by the distance shown by the diagonalized tree. To simulated the 
-diagonalization in the original tree, the arrow $t_i$ will be given weight $i$.
-Methods for exploring a tree with a large branching factor has been explored  
-in Go~\cite{alphago} and
-
-However for very large branching factor RTS Combinatorial Multi-armed Bandits
-for Real-Time Strategy Games. The method of choice for solving those 
-problems 
-is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
-to each branches
-
-In the context of changing wiehgts, the diagonalized tree has the advantage of 
-being a little faster to explore as the ordered of tactic is inscribed in the 
-shape of the tree. if the prior shapes
-
-\subsection{Node extension}
-
-To prevent tactics from looping, we add timeout of 0.02 seconds to all tactics 
-(optimized from previous paper)
-
-\begin{definition} (Productive tactic)
-During a proof search, after a tactic $t$ is applied to a node with active goal 
-$g$, 
-$t$ is called 
-productive if and only if all the following conditions are satisified:
-\begin{itemize}
-\item It does not fail. We note $l$ the list of goals returned by the tactic.
-\item It does not loop. $l$ do not contain any goals that appears in its 
-ancestor nodes.
-\item It is not a parallel step. $l$ is not a superset of a list of goals 
-produced from the same node with the same active goal.
-\end{itemize}
-
-The third point of the definition is a partial attempt to prevent confluent 
-branches. The general 
-case where two branches becomes parallel after $n$ steps is handle by a
-tactic cache which memorizes tactic effects on goals. Similarly there is also a 
-prediction cache, which memorize the tactic predictions for each goal. Those 
-two caches allow a fast exploration of confluent branches but they remain 
-separated in our search tree.
-
-\paragraph{Solving a goal and a node}
-Describe the case where a proof for a goal in a node is found.
-If one of those list where empty, the goal $g_i^0$ would 
-be proved.
-A node is solved when every of its goal are solved.
-
-\subsection{End of the proof search}
-The proof search stops when one of these 3 conditions is realized:
-\item It found a proof (i.e. the root node is solved). In this case, 
-the search returns a minimized and 
-prettified proof script (see Section~\ref{sec:proofdisplay}).
-\item It has saturated. 
-There is no more tactics to be applied to any open goal. (todo: define open 
-goal somewhere).
-
-\item It exceeded a given time limit.
-
-The second case happens very rarely (less than 10 times when evaluating 7088 
-theorems) and only at the beginning of the library when the learned dataset is 
-small. So in our experiments, we just give the number of proof found relative 
-to the number of proof attempts.
-
-\end{definition}
-
-\subsection{Evaluation}
 
@@
 \begin{align*}
-E_{none} (a_i) &= 0 \\
-E_{productive} (a_i) &= 1\\
-E_r (a_i) &= \frac{card\ Positive(r,a_i)}{r}\\
+PriorEvaluation_{none} (a_i) &= 0 \\
+PriorEvaluation_{productive} (a_i) &= 1\\
+PriorEvaluation_r (a_i) &= \frac{card\ Positive(r,a_i)}{r}\\
 \end{align*}
@@
 output to evaluate, so the result is a failure and a reward of $0$ will be 
-backpropagated.
+given for this node. The same thing happens if the node selected is already 
+proven or a descendant of a proven goal. 
+Experimentally, it is preferable not to 
+filter proven (or descendants of proven) nodes before 
+node selection to keep the estimation of the algorithm balanced.
+
+\subsection{Node selection}
+
+A similar trade-off between exploration and exploitation exists in algorithm 
+like Monte Carlo. However, Monte Carlo tree search metrics are based on many 
+simulations and do not reflect our exploration strategy that only
+
+We present here a monte carlo tree search with no roll out which is essentially 
+the same as the one use by AlphaGo Zero~\cite{} originally defined by \cite{} 
+in \cite{}.
+
+\[Value(a_i) = Exploitation(a_i) + c_{exploration} * Exploration(a_i)\]
+
+The exploration term is determined by the prior policy and the current policy 
+where as the exploitation is exactly the current evaluation build from prior of 
+evaluations of descendants of $a_i$.
+
+\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{Policy(a_i)}\]
+
+\[Policy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
+The policy is approximatively the  percentage of time a node was visited.The 
+square root skews this probability to favor even more exploration of nodes with 
+few visits.
+
+The exploitation function $Exploitation$ is the average of the evaluation of 
+all nodes descendant of $a_i$ (itself included) in the current search tree.   
+\[Exploitation{a_i} = Average_{b \in Desc(a_i))} {PriorEvaluation(b)}\]
+
+MCTS with prior policy
+This modification of the search tree can be simulated by weighting each arrow 
+by the distance shown by the diagonalized tree. To simulated the 
+diagonalization in the original tree, the arrow $t_i$ will be given weight $i$.
+Methods for exploring a tree with a large branching factor has been explored  
+in Go~\cite{alphago} and
+
+However for very large branching factor RTS Combinatorial Multi-armed Bandits
+for Real-Time Strategy Games. The method of choice for solving those 
+problems 
+is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
+to each branches
+
+In the context of changing wiehgts, the diagonalized tree has the advantage of 
+being a little faster to explore as the ordered of tactic is inscribed in the 
+shape of the tree. if the prior shapes
+
+\subsection{Node extension}
+
+To prevent tactics from looping, we add timeout of 0.02 seconds to all tactics 
+(optimized from previous paper)
+
+\begin{definition} (Productive tactic)
+During a proof search, after a tactic $t$ is applied to a node with active goal 
+$g$, 
+$t$ is called 
+productive if and only if all the following conditions are satisified:
+\begin{itemize}
+\item It does not fail. We note $l$ the list of goals returned by the tactic.
+\item It does not loop. $l$ do not contain any goals that appears in its 
+ancestor nodes.
+\item It is not a parallel step. $l$ is not a superset of a list of goals 
+produced from the same node with the same active goal.
+\end{itemize}
+
+The third point of the definition is a partial attempt to prevent confluent 
+branches. The general 
+case where two branches becomes parallel after $n$ steps is handle by a
+tactic cache which memorizes tactic effects on goals. Similarly there is also a 
+prediction cache, which memorize the tactic predictions for each goal. Those 
+two caches allow a fast exploration of confluent branches but they remain 
+separated in our search tree.
+
+\paragraph{Solving a goal and a node}
+Describe the case where a proof for a goal in a node is found.
+If one of those list where empty, the goal $g_i^0$ would 
+be proved.
+A node is solved when every of its goal are solved.
+
+\subsection{End of the proof search}
+The proof search stops when one of these 3 conditions is realized:
+\begin{itemize}
+\item It found a proof (i.e. the root node is solved). In this case, 
+the search returns a minimized and 
+prettified proof script (see Section~\ref{sec:proofdisplay}).
+\item It has saturated. 
+There is no more tactics to be applied to any open goal. \todo{define open 
+goal}
+\item It exceeded a given time limit.
+\end{itemize}
+
+The second case happens very rarely (less than 10 times when evaluating 7088 
+theorems) and only at the beginning of the library when the learned dataset is 
+small. So in our experiments, we just give the number of proof found relative 
+to the number of proof attempts.
+
+\end{definition}
+
+\subsection{Evaluation}
 

### 2018-01-20 tactictoe_jv.tex
removing old experiment

@@
 %approach.
+
+\todo{include latest results: 66.56 percent of 7123 theorems in 60 seconds and
+do a similar experiment for hh with minimization}
 
@@
 \subsection{Preselecting feature vectors and their  
-dependencies}\label{sec:dependencies}
+recursive dependencies}\label{sec:dependencies}
 We first rank feature vectors by their distance to the conjecture. But better 
@@
 problem, we reorder the list so that descendants (see 
-Definition~\ref{def:descendant}) of vectors inherits the similarity of their 
+Definition~\ref{def:recdep}) of vectors inherits the similarity of their 
 parent.
 
-\begin{definition}(Descendant of a feature vector)\label{def:descendant}
+\begin{definition}(Recursive dependencies of a feature 
+vector)\label{def:recdep}
 Let $\mathfrak{F}$ be a set of recorded feature vectors. 
-The set of all descendants $D_\infty$ of a feature vector $f$ is 
+The set of all recursive dependencies $D_\infty$ of a feature vector $f$ is 
 inductively defined by:
@@
 We augment the ordered predictions by inserting after each predicted vector 
-its descendants and removing lower ranked duplicates. This contrast with 
-theorem predictions where we only select parents. This phase is only applied 
+its descendants and removing lower ranked duplicates. 
+%This contrast with 
+%theorem predictions where we only select parents. 
+ This phase is only applied 
 during preselection, so is of no great consequence as long as we select enough 

### 2018-01-20 tactictoe_jv.tex
data from the latest experiments available for manual proof mining

@@
 %amount of time.
+\todo{include latest results: 66.56 percent of 7123 theorems in 60 seconds and
+do a similar experiment for hh with minimization}
 
@@
 %approach.
-
-\todo{include latest results: 66.56 percent of 7123 theorems in 60 seconds and
-do a similar experiment for hh with minimization}
 
@@
 
-\section{Case study}\todo{
+\section{Case study: comparison with original proofs}\todo{
   TG + RK: describe the two examples.
@@
 
+%val SURJ_INJ_INV = store_thm(
+%  "SURJ_INJ_INV",
+%  ``SURJ f s t ==> ?g. INJ g t s /\ !y. y IN t ==> (f (g y) = y)``,
+
 %SURJ_INJ_INV
@@
 
-%\begin{remark}
-%The additional difficulty produced by the necessary induction is such that the 
-%theorem could not be proven by \holyhammer.
-%\end{remark}
-
-\newcommand{\wed}{\begin{math}\wedge\end{math}}
-
-Proofs discovered by \tactictoe discovered during evaluation.
-
-for theorem \texttt{SURJ\_INJ\_INV} In 
-theory \texttt{pred\_set})
-
-\begin{example}\label{ex:casestudy}()
-
-%val SURJ_INJ_INV = store_thm(
-%  "SURJ_INJ_INV",
-%  ``SURJ f s t ==> ?g. INJ g t s /\ !y. y IN t ==> (f (g y) = y)``,
-\vspace{5mm}
-\small
-
-Goal
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-!f s t. SURJ f s t ==> ?g. INJ g t s wedge !y. y IN t ==> (f (g y) = y)
-\end{lstlisting}
-
-Original \holfour proof script (2.03 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-REWRITE_TAC [IMAGE_SURJ] THEN
-  DISCH_TAC THEN Q.EXISTS_TAC `THE o LINV_OPT f s` THEN
-  BasicProvers.VAR_EQ_TAC THEN REPEAT STRIP_TAC
-  THENL [
-  irule INJ_COMPOSE THEN Q.EXISTS_TAC `IMAGE SOME s` THEN
-    REWRITE_TAC [INJ_LINV_OPT_IMAGE] THEN REWRITE_TAC [INJ_DEF, IN_IMAGE] THEN
-    REPEAT STRIP_TAC THEN REPEAT BasicProvers.VAR_EQ_TAC THEN
-    FULL_SIMP_TAC std_ss [THE_DEF],
-  ASM_REWRITE_TAC [LINV_OPT_def, o_THM, THE_DEF] THEN
-    RULE_ASSUM_TAC (Ho_Rewrite.REWRITE_RULE
-      [IN_IMAGE', GSYM SELECT_THM, BETA_THM]) THEN ASM_REWRITE_TAC [] ]
-\end{lstlisting}
-
-\vspace{5mm}
-
-\tactictoe proof script (50.85 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-SRW_TAC [] [SURJ_DEF, INJ_DEF] THEN METIS_TAC []
-\end{lstlisting}
-
-Goal
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-!l1 l2 n x. LENGTH l1 <= n ==> 
-  (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n - LENGTH l1) l2))
-\end{lstlisting}
-
-\vspace{5mm}
-
-Original \holfour proof script (62.69 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-  rw[] THEN simp[LIST_EQ_REWRITE] THEN Q.X_GEN_TAC `z` THEN
-  simp[EL_LUPDATE] THEN rw[] THEN simp[EL_APPEND2,EL_LUPDATE] THEN
-  fs[] THEN Cases_on `z < LENGTH l1` THEN
-  fs[] THEN simp[EL_APPEND1,EL_APPEND2,EL_LUPDATE]
-\end{lstlisting}
-
-%Name: 224 LUPDATE_APPEND2
-%Original proof time:
-
-%val LUPDATE_APPEND2 = Q.store_thm("LUPDATE_APPEND2",
-%   `!l1 l2 n x.
-%      LENGTH l1 <= n ==>
-%      (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n-LENGTH l1) l2))`,
-%);
-\vspace{5mm}
-
-\tactictoe proof script (17.16 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-Induct_on `l1` THENL [SRW_TAC [] [],
-  Cases_on `n` THENL [SRW_TAC [] [],
-    FULL_SIMP_TAC (srw_ss ()) [] THEN METIS_TAC [LUPDATE_def]]]
-\end{lstlisting}
-
-If a discovered proof is reasonable short, fast, maintainable  and 
-understandable, they could 
-replace its original in the \holfour standard library. In the first case the 
-tactictoe proof is slower but could still be considered becaues it is much more 
-readable. In the second case, it is faster and more readable.
+%Name: 0 IS_GCD_UNIQUE
+%proof length: 12
+%proof time: 0.080819
+%Statistics
+%  infstep : 2539
+%  nodes   : 396
+%  maxdepth: 7
+%Time: 40.309547
+%  inferstep time: 30.15097
+%  node_find time: 0.58993
+%  node_crea time: 9.142918
+%    pred time: 2.425785
+%    thmpred time: 0.000101
+%    mc time: 4.417004
+%    inst time: 0.0
+%Proof found: STRIP_TAC THEN REWRITE_TAC [fetch "gcd" "is_gcd_def"] THEN REPEAT 
+%Cases THENL [metisTools.METIS_TAC [], REWRITE_TAC [SUC_NOT, ALL_DIVIDES_0, 
+%compute_divides] THEN metisTools.METIS_TAC [numTheory.NOT_SUC], 
+%metisTools.METIS_TAC [numTheory.NOT_SUC, DIVIDES_ANTISYM], metisLib.METIS_TAC 
+%[LESS_EQUAL_ANTISYM, (fetch "divides" "DIVIDES_LE"), prim_recTheory.LESS_0], 
+%metisTools.METIS_TAC [], (RW_TAC numLib.arith_ss) [(fetch "divides" 
+%"divides_def")], metisTools.METIS_TAC [DIVIDES_ANTISYM], metisLib.METIS_TAC 
+%[LESS_EQUAL_ANTISYM, (fetch "divides" "DIVIDES_LE"), prim_recTheory.LESS_0]]
+%Org tac number: 1
+%Original proof time: 0.0
+%Recording proof time: 0.008556
 
@@
 
-\end{example}
+Goal: \texttt{IS\_GCD\_UNIQUE} in theory \texttt{pred\_set}).
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+!a b c d. is_gcd a b c wedge is_gcd a b d ==> (c = d)
+\end{lstlisting}
+
+Original \holfour proof script (2.03 milliseconds)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+PROVE_TAC[IS_GCD,DIVIDES_ANTISYM]
+\end{lstlisting}
+
+\vspace{5mm}
+
+\tactictoe proof script
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+STRIP_TAC THEN 
+REWRITE_TAC [fetch "gcd" "is_gcd_def"] THEN 
+REPEAT Cases THENL 
+  [METIS_TAC [], 
+   REWRITE_TAC [SUC_NOT, ALL_DIVIDES_0, compute_divides] THEN 
+     METIS_TAC [NOT_SUC], 
+   METIS_TAC [NOT_SUC, DIVIDES_ANTISYM], 
+   METIS_TAC [LESS_EQUAL_ANTISYM, DIVIDES_LE, LESS_0], 
+   METIS_TAC [], 
+   RW_TAC numLib.arith_ss [divides_def], 
+   METIS_TAC [DIVIDES_ANTISYM],
+   METIS_TAC [LESS_EQUAL_ANTISYM, DIVIDES_LE, LESS_0]]
+\end{lstlisting}
+
+tacictoe is resourceful.
+
+Goal: \texttt{SURJ\_INJ\_INV} in theory \texttt{pred\_set})
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+!f s t. SURJ f s t ==> ?g. INJ g t s wedge !y. y IN t ==> (f (g y) = y)
+\end{lstlisting}
+
+Original \holfour proof script (2.03 milliseconds)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+REWRITE_TAC [IMAGE_SURJ] THEN
+  DISCH_TAC THEN Q.EXISTS_TAC `THE o LINV_OPT f s` THEN
+  BasicProvers.VAR_EQ_TAC THEN REPEAT STRIP_TAC
+  THENL [
+  irule INJ_COMPOSE THEN Q.EXISTS_TAC `IMAGE SOME s` THEN
+    REWRITE_TAC [INJ_LINV_OPT_IMAGE] THEN REWRITE_TAC [INJ_DEF, IN_IMAGE] THEN
+    REPEAT STRIP_TAC THEN REPEAT BasicProvers.VAR_EQ_TAC THEN
+    FULL_SIMP_TAC std_ss [THE_DEF],
+  ASM_REWRITE_TAC [LINV_OPT_def, o_THM, THE_DEF] THEN
+    RULE_ASSUM_TAC (Ho_Rewrite.REWRITE_RULE
+      [IN_IMAGE', GSYM SELECT_THM, BETA_THM]) THEN ASM_REWRITE_TAC [] ]
+\end{lstlisting}
+
+\vspace{5mm}
+
+\tactictoe proof script (50.85 milliseconds)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+SRW_TAC [] [SURJ_DEF, INJ_DEF] THEN METIS_TAC []
+\end{lstlisting}
+
+Goal
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+!l1 l2 n x. LENGTH l1 <= n ==> 
+  (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n - LENGTH l1) l2))
+\end{lstlisting}
+
+\vspace{5mm}
+
+Original \holfour proof script (62.69 milliseconds)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+  rw[] THEN simp[LIST_EQ_REWRITE] THEN Q.X_GEN_TAC `z` THEN
+  simp[EL_LUPDATE] THEN rw[] THEN simp[EL_APPEND2,EL_LUPDATE] THEN
+  fs[] THEN Cases_on `z < LENGTH l1` THEN
+  fs[] THEN simp[EL_APPEND1,EL_APPEND2,EL_LUPDATE]
+\end{lstlisting}
+
+%Name: 224 LUPDATE_APPEND2
+%Original proof time:
+
+%val LUPDATE_APPEND2 = Q.store_thm("LUPDATE_APPEND2",
+%   `!l1 l2 n x.
+%      LENGTH l1 <= n ==>
+%      (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n-LENGTH l1) l2))`,
+%);
+\vspace{5mm}
+
+\tactictoe proof script (17.16 milliseconds)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+Induct_on `l1` THENL [SRW_TAC [] [],
+  Cases_on `n` THENL [SRW_TAC [] [],
+    FULL_SIMP_TAC (srw_ss ()) [] THEN METIS_TAC [LUPDATE_def]]]
+\end{lstlisting}
+
+If a discovered proof is reasonable short, fast, maintainable  and 
+understandable, they could 
+replace its original in the \holfour standard library. In the first case the 
+tactictoe proof is slower but could still be considered becaues it is much more 
+readable. In the second case, it is faster and more readable.
 

### 2018-01-21 tactictoe_jv.tex
minor

@@
 Experiments on the \holfour standard 
-library show that TacticToe is able to re-prove 59\% of 7088 theorems in less 
-than 10 seconds. This is to be compared with the 29\% proved by the ATP Eprover 
-in auto-schedule mode with 128 lemmas predicted and translated by HolyHammer.
+library show that TacticToe is able to re-prove 66.6\% of 7123 theorems in less 
+than 60 seconds. 
+This is to be compared with the 42.8\% proved in the same amount of time by a 
+a barebone version of TacticToe.
+
+In the same condtion the ATP Eprover proves ..., when it is run in 
+auto-schedule mode with 128 lemmas predicted and 
+translated by HolyHammer.
+\todo{to do eprover results running}
 
@@
 %amount of time.
-\todo{include latest results: 66.56 percent of 7123 theorems in 60 seconds and
-do a similar experiment for hh with minimization}
 
@@
 Despite the best efforts of the prediction algorithms, the selected tactic may 
-not 
-solve the current goal, proceed in the wrong direction
+not solve the current goal, proceed in the wrong direction
 or even loop. For that reason, the prediction needs to be accompanied by a  
 proof search mechanism that allows for backtracking and 
-can choose which proof tree to extend next and in 
-which direction.
-Our search algorithm is an advanced tree search algorithm that relies on 
-evaluation and policy functions to guide the search.
-
-\todo{TG: add transition}
-A proof is in general not a path but a tree. This means that our search space 
-has two branching factors: 
-the choice of a tactic, and the number of goals produced by each tactic.
-The first one can be thought as a disjunction of tactics. Indeed, 
-if there exists a tactic that solves a goal then it is solved. And 
-conversely, the second branching factor is a conjunction of goals. All pending 
-goals must be proven in order to prove the parent goal. For example, in the 
-case of induction tactics the pending goals are the base case and the inductive 
-case.
-
-\todo{TG: add general overview of the algorithm and describe the plan}
-If we suppose that each tactic produces at most one goal, then the standard 
-monte carlo tree search algorithm applies. So we will first discuss how our 
-algorithm works in this case. And then explain how we extend it to support 
-tactics producing 
-multiple goals.
+can choose which proof tree to extend next and in which direction.
+Our search algorithm is a monte carlo tree search (MCTS) algorithm that relies 
+on prior evaluation and policy functions. Those priors are learned through 
+direct supervised learning in the case of the policy and via the data 
+accumulated during orthogonalization for the evaluation.
+
+So, we first give a static description of the proof tree and its component.
+Then we explain how we compute the evaluation and policy functions from learned 
+data. And finally, we show how one MCTS step changes the proof tree data 
+structures.
+
+%A proof is in general not a path but a tree. This means that our search space 
+%has two branching factors: 
+%the choice of a tactic, and the number of goals produced by each tactic.
+%The first one can be thought as a disjunction of tactics. Indeed, 
+%if there exists a tactic that solves a goal then it is solved. And 
+%conversely, the second branching factor is a conjunction of goals. All pending 
+%goals must be proven in order to prove the parent goal. For example, in the 
+%case of induction tactics the pending goals are the base case and the 
+%inductive 
+%case.
+
+%If we suppose that each tactic produces at most one goal, then the standard 
+%monte carlo tree search algorithm applies. So we will first discuss how our 
+%algorithm works in this case. And then explain how we extend it to support 
+%tactics producing 
+%multiple goals.
 
@@
 
+Definition of solved goals, open goals  and pending goals.
 Supervised learning introduction.
+
+%A proof is in general not a path but a tree. This means that our search space 
+%has two branching factors: 
+%the choice of a tactic, and the number of goals produced by each tactic.
+%The first one can be thought as a disjunction of tactics. Indeed, 
+%if there exists a tactic that solves a goal then it is solved. And 
+%conversely, the second branching factor is a conjunction of goals. All pending 
+%goals must be proven in order to prove the parent goal. For example, in the 
+%case of induction tactics the pending goals are the base case and the 
+%inductive 
+%case.
+
+\paragraph{Solving a goal and a node}
+Describe the case where a proof for a goal in a node is found.
+If one of those list where empty, the goal $g_i^0$ would 
+be proved.
+A node is solved when every of its goal are solved.
+
 \subsection{Prior tactic policy}
@@
 all nodes descendant of $a_i$ (itself included) in the current search tree.   
-\[Exploitation{a_i} = Average_{b \in Desc(a_i))} {PriorEvaluation(b)}\]
+\[Exploitation(a_i) = Average_{b \in Desc(a_i))} {PriorEvaluation(b)}\]
 
@@
 separated in our search tree.
-
-\paragraph{Solving a goal and a node}
-Describe the case where a proof for a goal in a node is found.
-If one of those list where empty, the goal $g_i^0$ would 
-be proved.
-A node is solved when every of its goal are solved.
 
@@
 
+10 seconds experiment:
+
 Until now, all our feature vectors were learned form human proofs. 
@@
 \tactictoe & 74.6 & 67.3 & 77.3 & 28.2\\
-\eprover & 50.8 & 67.6 & 63.4 & 10.8 \\
+\eprover & 50.8 & 67.6 & 63.4 & 10.8\\
 \midrule
@@
 
+60 seconds experiment + 
+adding theorems from the namespace and removing 
+additionally saved theorems + 
+recording proofs from Prove.
+
 Constant progress: 
-  From 45 percent on the same dataset in 5 seconds 
-       59 percent in december in 10 seconds
-  Probably 65-70 percent en of January in 60 seconds.
++45 percent in Septermber in 5 seconds.
++59 percent in December in 10 seconds.
++66.6 percent in January in 60 seconds.
 
@@
 \section{Case study: comparison with original proofs}\todo{
-  TG + RK: describe the two examples.
+  TG + RK: find describe examples.
   RK: After implementation and small experiment is done (in about 1 week) 
@@
 
-Original \holfour proof script (2.03 milliseconds)
+Original \holfour proof script
 \begin{lstlisting}[language=SMLSmall,frame=tb]
-PROVE_TAC[IS_GCD,DIVIDES_ANTISYM]
+PROVE_TAC[IS_GCD, DIVIDES_ANTISYM]
 \end{lstlisting}

### 2018-01-21 tactictoe_jv.tex
search: non-constructive definition of a solved goal and node

@@
 
-\subsection{A proof tree with two branching factors}
-
-A tactical proof tree is a set of nodes containing a list of goals where each 
-goal $g$ of a node $n$ is linked to a node $m$ when the list of
+\subsection{A search tree: data structures and definitions}
+
+\begin{definition}(Tactical search tree)
+A tactical search tree (or search tree) is a set of nodes containing a list of 
+goals where each goal $g$ of a node $n$ is linked to a node $m$ when the list 
+of 
+\end{definition}
 
@@
 \end{center}
-\caption{The two branching factors: choice of a goal $g_0^i$ and choice of a 
-tactic $t_j$.}
+\caption{A node of a search tree and its relation with the children of one of 
+its goal}
 \label{fig:choice}
@@
 
-Starting from the bottom node, we can construct descendant of this node by 
+Starting from the bottom parent node, we can construct children of a goal of 
+this node by 
 choosing a goal $g_i^0$ on this node and choosing tactics $t1,t2,\ldots$ to 
@@
 of goals to be proven.
+
+In each node can either be solved, open or pending. The definition of a solved 
+goal is expressed through mutual recursion with a 
+solved node and is given in ~\ref{def:solved_goal}. An open goal is goal that 
+was tried at least once by a tactic. In our implementation we have only one 
+open goals per node at once. A goal that was not explored yet is called a 
+pending goal.
+
+\begin{definition}(Solved node and solved goal)
+Let $\mathbb{T}$ be a search tree. The set of solved nodes $\mathbb{A}^*$ and 
+the set of solved goals $\mathbb{G}^*$ are the smallest set verifying the 
+following equations:
+
+\todo{should I make a constructive definition with $A^n+1$ in terms of $G^n$ 
+instead?}
+\begin{align*}
+\mathbb{A}^* &= \lbrace a \in \mathbb{A}\ |\ \forall g \in G(a).\ g \in 
+\mathbb{G}^* \rbrace\\
+\mathbb{G}^* &= \lbrace g \in \mathbb{G}\ |\ \exists t \in T(g).\ t(g) \in 
+\mathbb{A}^* \rbrace
+\end{align*}
+
+By the base case of these definition, a goal is solved if there exists a tactic 
+applied to this goal that produces no goals.
+
+\end{definition}
 
@@
 %case.
-
-\paragraph{Solving a goal and a node}
-Describe the case where a proof for a goal in a node is found.
-If one of those list where empty, the goal $g_i^0$ would 
-be proved.
-A node is solved when every of its goal are solved.
 
@@
 
-\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{Policy(a_i)}\]
-
-\[Policy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
+\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]
+
+\[CurPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
 The policy is approximatively the  percentage of time a node was visited.The 
@@
 
-The exploitation function $Exploitation$ is the average of the evaluation of 
-all nodes descendant of $a_i$ (itself included) in the current search tree.   
-\[Exploitation(a_i) = Average_{b \in Desc(a_i))} {PriorEvaluation(b)}\]
+The exploitation term is the average of the evaluation of 
+all nodes descendant of $a_i$ (itself included) in the current search tree. It 
+is also called the current evaluation of $a_i$.   
+\[Exploitation(a_i) = CurEvaluation(a_i) = 
+  Average_{b \in Desc(a_i)} {PriorEvaluation(b)}\]
 
@@
 
-tacictoe is resourceful.
+TacicToe is resourceful. Changing the shape of the feature of the theorems
+to get better predictions.
 
@@
 DISCH_TAC THEN Q.EXISTS_TAC `THE o LINV_OPT f s` THEN
-  BasicProvers.VAR_EQ_TAC THEN REPEAT STRIP_TAC
-  THENL [
-  irule INJ_COMPOSE THEN Q.EXISTS_TAC `IMAGE SOME s` THEN
+BasicProvers.VAR_EQ_TAC THEN REPEAT STRIP_TAC THENL 
+  [irule INJ_COMPOSE THEN Q.EXISTS_TAC `IMAGE SOME s` THEN
      REWRITE_TAC [INJ_LINV_OPT_IMAGE] THEN REWRITE_TAC [INJ_DEF, IN_IMAGE] THEN
@@
      RULE_ASSUM_TAC (Ho_Rewrite.REWRITE_RULE
-      [IN_IMAGE', GSYM SELECT_THM, BETA_THM]) THEN ASM_REWRITE_TAC [] ]
+       [IN_IMAGE', GSYM SELECT_THM, BETA_THM]) THEN ASM_REWRITE_TAC []]
 \end{lstlisting}
@@
 
+Make it match the new version.
 \tactictoe proof script (50.85 milliseconds)
@@
 \vspace{5mm}
-
+Make it match the new version.
 \tactictoe proof script (17.16 milliseconds)
@@
 \end{lstlisting}
-
-If a discovered proof is reasonable short, fast, maintainable  and 
-understandable, they could 
-replace its original in the \holfour standard library. In the first case the 
-tactictoe proof is slower but could still be considered becaues it is much more 
-readable. In the second case, it is faster and more readable.
+fs () is short for FULL\_SIMP\_TAC (srw\_ss ()) and rw is short for ??.
+
+If a discovered proof is short, fast, maintainable and 
+understandable, it could replace its original in the \holfour standard library.
 
@@
   \cite{conjecturing}
+  + Self-learning.  
++ MCTS was presented in manner that we only did the first step of reinforcement 
+learning. CurrentPolicy becomes PriorPolicy.
 

### 2018-01-21 tactictoe_jv.tex
search: search tree definition

@@
 (\hat{t},\hat{g}) \in D_n.\ g \in \hat{t}(\hat{g}) \rbrace  \\
-D_\infty &=_{def} \bigcup_{i \in \mathbb{N}^*} D_i\\
+D_\infty &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
 \end{align*}
@@
 
-\begin{definition}(Tactical search tree)
-A tactical search tree (or search tree) is a set of nodes containing a list of 
-goals where each goal $g$ of a node $n$ is linked to a node $m$ when the list 
-of 
+\begin{definition}(Tactical search tree)\\
+A tactical search tree is a sixtuple $(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
+that respects the following conditions:
+\begin{itemize}
+\item $\mathbb{A}$ is a set of nodes, $\mathbb{G}$ is a set of goals 
+ and $\mathbb{T}$ is a set of tactics.
+\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
+$\mathbb{A}$.
+\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$.
+\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$.
+\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
+\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
+\end{itemize}
+
+Additional constraints such as acyclicity and the fact that there is exactly 
+one root node are properties of this graph that are preserved during the 
+application of the MCTS algorithm. That is why we call it a tree although this 
+we do not specify it in this definition.
 \end{definition}
+
+We assume that any set is numbered and that we can access a nth member of set 
+$s$ by writing $s_n$. For readability in the next figure $(G(a_j))_i$ is 
+written 
+$g_j^i$.
 
@@
 \begin{definition}(Solved node and solved goal)
-Let $\mathbb{T}$ be a search tree. The set of solved nodes $\mathbb{A}^*$ and 
-the set of solved goals $\mathbb{G}^*$ are the smallest set verifying the 
-following equations:
-
-\todo{should I make a constructive definition with $A^n+1$ in terms of $G^n$ 
-instead?}
+Let $\mathbb{T}$ be a search tree. The set of solved nodes $\mathbb{A}^\infty$ 
+and 
+the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:
+
 \begin{align*}
-\mathbb{A}^* &= \lbrace a \in \mathbb{A}\ |\ \forall g \in G(a).\ g \in 
-\mathbb{G}^* \rbrace\\
-\mathbb{G}^* &= \lbrace g \in \mathbb{G}\ |\ \exists t \in T(g).\ t(g) \in 
-\mathbb{A}^* \rbrace
+\mathbb{A}^{0} &=_{def} 
+\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\ 
+\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
+\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
+\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\ 
+\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
+\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
+\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
+\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n\\
+\mathbb{G}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
 \end{align*}
@@
 \subsection{Prior tactic policy}
-
 The estimation through the distance produced by the tactic selectors are hard 
@@
 Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
-distance to a goal $g$. Let $c_policy$ be a constant heuristic that estimates 
+distance to a goal $g$. Let $c_{policy}$ be a constant heuristic that estimates 
 how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
@@
 
-\[PriorPolicy(t_{i}) = (1 - c_policy)^{i} * c_policy\]
+\[PriorPolicy(t_{i}) = (1 - c_{policy})^{i} * c_{policy}\]
 

### 2018-01-21 tactictoe_jv.tex
moving related works before the conclusion

@@
 \sml function. So, our recording is more complete and fully automatized.
-\item Existing tactic orthogonalization on recorded data is made stronger by 
-also including the internal ATP \metis in the orthogonalization 
-competition. Because we manage a database of theorems for argument 
-prediction, we apply here a pruning step to remove easy theorems.
+%\item Existing tactic orthogonalization on recorded data is made stronger by 
+%also including the internal ATP \metis in the orthogonalization 
+%competition. Because we manage a database of theorems for argument 
+%prediction, we apply here a pruning step to remove easy theorems.
 \item Two types of arguments (list of theorems and terms) can be predicted 
@@
 through supervised learning.
-\item The internal ATP \metis used during our proof search is complemented by 
-asynchronous calls to \eprover.
+
 \item First evaluation of \tactictoe and \eprover success rates relative to 
@@
 applied on the generated proof.
+\item The internal ATP \metis used during proof search can be complemented 
+by asynchronous calls to \eprover.
 \end{itemize}
-
-\section{Related Work}\todo{TG reorganise + other motivation?}
-There are several essential components of our work that are comparable to 
-previous approaches: tactic-level proof recording, tactic 
-selection through machine learning techniques and automatic tactic-based proof 
-search. Our work is also related to previous approaches that use machine 
-learning to select premises for the ATP systems and guide ATP proof search 
-internally.
-
-For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15} 
-can transform a packed tactical proof into a series of interactive tactic 
-calls. Its principal application 
-was so far refactoring the library and teaching common proof techniques to new 
-ITP users. In our work, the splitting of a proof into a sequence of tactics is 
-essential for the
-tactic recording procedure, used to train our tactic prediction module.
-
-The system \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14} 
-groups related proofs thanks to its clustering 
-algorithms. It allows \coq users to inspire themselves from similar proofs and 
-notice 
-duplicated proofs. Our predictions comes from a much more detailed description 
-of the open goal.
-However, we simply create a single label for each tactic call whereas each of 
-its
-arguments is treated independently in \textsf{ML4PG}. 
-Our choice is motivated by the k-NN algorithm already used in
-\holyhammer for the selection of theorems.
-%, which can be easly adapted to the selection of tactics but would be improved 
-%by a 
-%version that can learn to predict tactics and their arguments independently.
-
-\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to 
-generate
-proof scripts from previous \coq proof examples.
-Its strength lies in its ability to produce likely sequences 
-of tactics for solving domain specific goals. It operates by creating a model 
-for common sequences of tactics for a specific library.
-This means that in order to propose the following tactic, only the previously 
-called tactics
-are considered.
-%intuition tells us that it should not be the main factor
-Our algorithm, on the other hand, relies mainly on the characteristics of the 
-current goal 
-to decide
-which tactics to apply next. In this way, our learning mechanism has to 
-rediscover why each 
-tactic was applied for the current subgoals. It may lack some useful bias for 
-common sequences 
-of tactics, but is more reactive to subtle changes. Indeed, it can be trained 
-on a large library and only tactics relevant to the current subgoal will be 
-selected. 
-Concerning the proof search, \textsf{SEPIA}'s %simple 
-breadth-first search is replaced by an A*-algorithm which allows for heuristic 
-guidance in 
-the exploration of the search tree.
-Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
-theorems) of the 
-\coq library demonstrating that it globally outperforms individual \coq 
-tactics. In contrast, we demonstrate the competitiveness of our system against 
-the successful general-purpose hammers on the \holfour standard library (7902 
-theorems).
-
-Machine learning has also been used to advise the best library lemmas for new 
-ITP goals.
-This can be done either in an interactive way, when the user completes the 
-proof based on the recommended lemmas, as in the \textsc{Mizar Proof 
-Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma 
-selection is handed over to the ATP component of a \emph{hammer} 
-system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.
-
-Internal learning-based selection of tactical steps inside an ITP is analogous 
-to internal learning-based selection of clausal steps inside ATPs such as 
-\textsc{MaLeCoP}~\cite{malecop} and \textsc{FEMaLeCoP}~\cite{femalecop}. These 
-systems
-use the naive Bayes classifier to  select clauses for the extension steps in
-tableaux proof search based on many previous proofs. Satallax~\cite{Brown2012a} 
-can guide its
-search internally~\cite{mllax} using a command classifier, which can estimate 
-the priority of the 11 kinds of
-commands in the priority queue based on positive and negative examples.
-There are several essential components of our work that are comparable to 
-previous approaches: tactic-level proof recording, tactic 
-selection through machine learning techniques and automatic tactic-based proof 
-search. Our work is also related to previous approaches that use machine 
-learning to select premises for the ATP systems and guide ATP proof search 
-internally.
-
-For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15} 
-can transform a packed tactical proof into a series of interactive tactic 
-calls. Its principal application 
-was so far refactoring the library and teaching common proof techniques to new 
-ITP users. In our work, the splitting of a proof into a sequence of tactics is 
-essential for the
-tactic recording procedure, used to train our tactic prediction mechanism.
-
-The system \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14} 
-groups related proofs thanks to its clustering 
-algorithms. It allows \coq users to inspire themselves from similar proofs and 
-notice 
-duplicated proofs. Our predictions comes from a much more detailed description 
-of the open goal.
-However, we simply create a single label for each tactic call whereas each of 
-its
-arguments is treated independently in \textsf{ML4PG}. 
-Our choice is motivated by the k-NN algorithm already used in
-\holyhammer for the selection of theorems.
-%, which can be easly adapted to the selection of tactics but would be improved 
-%by a 
-%version that can learn to predict tactics and their arguments independently.
-
-\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to 
-generate
-proof scripts from previous \coq proof examples.
-Its strength lies in its ability to produce likely sequences 
-of tactics for solving domain specific goals. It operates by creating a model 
-for common sequences of tactics for a specific library.
-This means that in order to propose the following tactic, only the previously 
-called tactics
-are considered.
-%intuition tells us that it should not be the main factor
-Our algorithm, on the other hand, relies mainly on the characteristics of the 
-current goal 
-to decide
-which tactics to apply next. In this way, our learning mechanism has to 
-rediscover why each 
-tactic was applied for the current subgoals. It may lack some useful bias for 
-common sequences 
-of tactics, but is more reactive to subtle changes. Indeed, it can be trained 
-on a large library and only tactics relevant to the current subgoal will be 
-selected. 
-Concerning the proof search, \textsf{SEPIA}'s %simple 
-breadth-first search is replaced by MCTS which allows for heuristic 
-guidance in 
-the exploration of the search tree.
-Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
-theorems) of the 
-\coq library demonstrating that it globally outperforms individual \coq 
-tactics. In contrast, we demonstrate the competitiveness of our system against 
-the successful general-purpose hammers on the \holfour standard library.
-A recent improvement on this system~\cite{} includes the possibility to replace 
-theorems in tactics by ones belonging to the same cluster, independently of the 
-current goal. We have the opposite approach, we forget about the original 
-theorems in the tactic creating an more general abstract tactic
-and only use the goal to guide predictions. Finding a way to combine those two 
-approaches could be the key to stronger theorem predictions.
-
-A similar effort for stronger automation is performed in \isabelle. Nagashima 
-developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the user to 
-program with regular tactics and combine them with tools such as 
-\sledgehammer~\cite{sledgehammer10} or Nitpick~\cite{Nitpick10}. 
-Selection of tactics from recorded proofs has yet to be implemented and now it 
-is up to come up with a good meta-tactic.
-
-%In \holfour, 
-%a tactic is a just a \sml function of a special type and \holyhammer can 
-%easily 
-%be modified to be a tactic. So this meta-level is not 
-%necessary, all programming can occur at the tactic level and even \tactictoe  
-%is a tactic.
-
-%* Theory Exploration (Moa Johanson+, ITP 2017)  I think that is more related 
-%to conjeturing and do not think it fits here.
-
-Machine learning has also been used to advise the best library lemmas for new 
-ITP goals.
-This can be done either in an interactive way, when the user completes the 
-proof based on the recommended lemmas, as in the \textsc{Mizar Proof 
-Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma 
-selection is handed over to the ATP component of a \emph{hammer} 
-system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.
-
-Internal learning-based selection of tactical steps inside an ITP is analogous 
-to internal learning-based selection of clausal steps inside ATPs such as 
-\textsc{MaLeCoP}~\cite{malecop} and \textsc{FEMaLeCoP}~\cite{femalecop}. These 
-systems
-use the naive Bayes classifier to  select clauses for the extension steps in
-tableaux proof search based on many previous proofs. Satallax~\cite{Brown2012a} 
-can guide its
-search internally~\cite{mllax} using a command classifier, which can estimate 
-the priority of the 11 kinds of
-commands in the priority queue based on positive and negative examples.
 
@@
 \begin{definition}(Tactical search tree)\\
-A tactical search tree is a sixtuple $(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
+A tactical search tree $\mathfrak{T}$ is a sixtuple 
+$(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
 that respects the following conditions:
@@
 this node by 
-choosing a goal $g_i^0$ on this node and choosing tactics $t1,t2,\ldots$ to 
+choosing a goal $g_i^0$ on this node and choosing tactics $t_1,t_2,\ldots$ to 
 apply to this goal. Here the result produces new nodes each containing a list 
@@
 \begin{definition}(Solved node and solved goal)
-Let $\mathbb{T}$ be a search tree. The set of solved nodes $\mathbb{A}^\infty$ 
+Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
+$\mathbb{A}^\infty$ 
 and 
@@
 
-Since there is a bijection between tactics of the tree search 
-and the node they 
-produced, any function define on tactics of the tree search can be map to a 
-function defined on nodes.
+%Since the support of A gives us a bijection between tactics of the tree search 
+%and the node they 
+%produced, any function define on tactics of the tree search can be map to a 
+%function defined on nodes.
 
@@
    \tactictoe + \metis & 182 \\
-   \tactictoe + asynchronous \eprover & 182 \\
-%x tactictoe + async holyhammer 10 & 183 \\
-   \tactictoe + \metis + asynchronous \eprover & 190 \\
+   \tactictoe + \eprover & 182 \\
+   %x tactictoe + async holyhammer (10 thread) & 183 \\
+   \tactictoe + \metis + \eprover & 190 \\
 
@@
 
+Slight error in this experiments were eprover predictions have been made from 
+the preselected 500 theorems instead of the full library. But even with the 
+added overhead, we already see 
+the benefit of using stronger external prover than \metis.
+
 \subsection{Full-scale experiment}
@@
 Best parameters + comparison with holyhammer.
+asynchronous eprover is not used  for fairer comparison with eprover.
 
@@
 
+\section{Related Work}\todo{TG reorganise + other motivation?}
+There are several essential components of our work that are comparable to 
+previous approaches: tactic-level proof recording, tactic 
+selection through machine learning techniques and automatic tactic-based proof 
+search. Our work is also related to previous approaches that use machine 
+learning to select premises for the ATP systems and guide ATP proof search 
+internally.
+
+For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15} 
+can transform a packed tactical proof into a series of interactive tactic 
+calls. Its principal application 
+was so far refactoring the library and teaching common proof techniques to new 
+ITP users. In our work, the splitting of a proof into a sequence of tactics is 
+essential for the
+tactic recording procedure, used to train our tactic prediction module.
+
+The system 
+\textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
+
+groups related proofs thanks to its clustering 
+algorithms. It allows \coq users to inspire themselves from similar proofs and 
+notice 
+duplicated proofs. Our predictions comes from a much more detailed description 
+of the open goal.
+However, we simply create a single label for each tactic call whereas each of 
+its
+arguments is treated independently in \textsf{ML4PG}. 
+Our choice is motivated by the k-NN algorithm already used in
+\holyhammer for the selection of theorems.
+%, which can be easly adapted to the selection of tactics but would be improved 
+%by a 
+%version that can learn to predict tactics and their arguments independently.
+
+\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to 
+generate
+proof scripts from previous \coq proof examples.
+Its strength lies in its ability to produce likely sequences 
+of tactics for solving domain specific goals. It operates by creating a model 
+for common sequences of tactics for a specific library.
+This means that in order to propose the following tactic, only the previously 
+called tactics
+are considered.
+%intuition tells us that it should not be the main factor
+Our algorithm, on the other hand, relies mainly on the characteristics of the 
+current goal 
+to decide
+which tactics to apply next. In this way, our learning mechanism has to 
+rediscover why each 
+tactic was applied for the current subgoals. It may lack some useful bias for 
+common sequences 
+of tactics, but is more reactive to subtle changes. Indeed, it can be trained 
+on a large library and only tactics relevant to the current subgoal will be 
+selected. 
+Concerning the proof search, \textsf{SEPIA}'s %simple 
+breadth-first search is replaced by an A*-algorithm which allows for heuristic 
+guidance in 
+the exploration of the search tree.
+Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
+theorems) of the 
+\coq library demonstrating that it globally outperforms individual \coq 
+tactics. In contrast, we demonstrate the competitiveness of our system against 
+the successful general-purpose hammers on the \holfour standard library (7902 
+theorems).
+
+Machine learning has also been used to advise the best library lemmas for new 
+ITP goals.
+This can be done either in an interactive way, when the user completes the 
+proof based on the recommended lemmas, as in the \textsc{Mizar Proof 
+Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma 
+selection is handed over to the ATP component of a \emph{hammer} 
+system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.
+
+Internal learning-based selection of tactical steps inside an ITP is analogous 
+to internal learning-based selection of clausal steps inside ATPs such as 
+\textsc{MaLeCoP}~\cite{malecop} and \textsc{FEMaLeCoP}~\cite{femalecop}. These 
+systems
+use the naive Bayes classifier to  select clauses for the extension steps in
+tableaux proof search based on many previous proofs. Satallax~\cite{Brown2012a} 
+can guide its
+search internally~\cite{mllax} using a command classifier, which can estimate 
+the priority of the 11 kinds of
+commands in the priority queue based on positive and negative examples.
+There are several essential components of our work that are comparable to 
+previous approaches: tactic-level proof recording, tactic 
+selection through machine learning techniques and automatic tactic-based proof 
+search. Our work is also related to previous approaches that use machine 
+learning to select premises for the ATP systems and guide ATP proof search 
+internally.
+
+For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15} 
+can transform a packed tactical proof into a series of interactive tactic 
+calls. Its principal application 
+was so far refactoring the library and teaching common proof techniques to new 
+ITP users. In our work, the splitting of a proof into a sequence of tactics is 
+essential for the
+tactic recording procedure, used to train our tactic prediction mechanism.
+
+The system 
+\textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
+
+groups related proofs thanks to its clustering 
+algorithms. It allows \coq users to inspire themselves from similar proofs and 
+notice 
+duplicated proofs. Our predictions comes from a much more detailed description 
+of the open goal.
+However, we simply create a single label for each tactic call whereas each of 
+its
+arguments is treated independently in \textsf{ML4PG}. 
+Our choice is motivated by the k-NN algorithm already used in
+\holyhammer for the selection of theorems.
+%, which can be easly adapted to the selection of tactics but would be improved 
+%by a 
+%version that can learn to predict tactics and their arguments independently.
+
+\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to 
+generate
+proof scripts from previous \coq proof examples.
+Its strength lies in its ability to produce likely sequences 
+of tactics for solving domain specific goals. It operates by creating a model 
+for common sequences of tactics for a specific library.
+This means that in order to propose the following tactic, only the previously 
+called tactics
+are considered.
+%intuition tells us that it should not be the main factor
+Our algorithm, on the other hand, relies mainly on the characteristics of the 
+current goal 
+to decide
+which tactics to apply next. In this way, our learning mechanism has to 
+rediscover why each 
+tactic was applied for the current subgoals. It may lack some useful bias for 
+common sequences 
+of tactics, but is more reactive to subtle changes. Indeed, it can be trained 
+on a large library and only tactics relevant to the current subgoal will be 
+selected. 
+Concerning the proof search, \textsf{SEPIA}'s %simple 
+breadth-first search is replaced by MCTS which allows for heuristic 
+guidance in 
+the exploration of the search tree.
+Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
+theorems) of the 
+\coq library demonstrating that it globally outperforms individual \coq 
+tactics. In contrast, we demonstrate the competitiveness of our system against 
+the successful general-purpose hammers on the \holfour standard library.
+A recent improvement on this system~\cite{} includes the possibility to replace 
+theorems in tactics by ones belonging to the same cluster, independently of the 
+current goal. We have the opposite approach, we forget about the original 
+theorems in the tactic creating an more general abstract tactic
+and only use the goal to guide predictions. Finding a way to combine those two 
+approaches could be the key to stronger theorem predictions.
+
+A similar effort for stronger automation is performed in \isabelle. Nagashima 
+developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the 
+user to 
+program with regular tactics and combine them with tools such as 
+\sledgehammer~\cite{sledgehammer10} or Nitpick~\cite{Nitpick10}. 
+Selection of tactics from recorded proofs has yet to be implemented and now it 
+is up to come up with a good meta-tactic.
+
+%In \holfour, 
+%a tactic is a just a \sml function of a special type and \holyhammer can 
+%easily 
+%be modified to be a tactic. So this meta-level is not 
+%necessary, all programming can occur at the tactic level and even \tactictoe  
+%is a tactic.
+
+%* Theory Exploration (Moa Johanson+, ITP 2017)  I think that is more related 
+%to conjeturing and do not think it fits here.
+
+Machine learning has also been used to advise the best library lemmas for new 
+ITP goals.
+This can be done either in an interactive way, when the user completes the 
+proof based on the recommended lemmas, as in the \textsc{Mizar Proof 
+Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma 
+selection is handed over to the ATP component of a \emph{hammer} 
+system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.
+
+Internal learning-based selection of tactical steps inside an ITP is analogous 
+to internal learning-based selection of clausal steps inside ATPs such as 
+\textsc{MaLeCoP}~\cite{malecop} and \textsc{FEMaLeCoP}~\cite{femalecop}. These 
+systems
+use the naive Bayes classifier to  select clauses for the extension steps in
+tableaux proof search based on many previous proofs. Satallax~\cite{Brown2012a} 
+can guide its
+search internally~\cite{mllax} using a command classifier, which can estimate 
+the priority of the 11 kinds of
+commands in the priority queue based on positive and negative examples.
+
 \section{Conclusion}\label{sec:concl}\todo{TG + JU or CK review}

### 2018-01-22 tactictoe_jv.tex
proof search: wip

@@
 \DeclareMathOperator*{\argmax}{\arg\!\max}
+\DeclareMathOperator*{\average}{Average}
 
@@
 Experiments on the \holfour standard 
-library show that TacticToe is able to re-prove 66.6\% of 7123 theorems in less 
+library show that TacticToe is able to re-prove 66.2\% of all 7164 theorems in 
+less 
 than 60 seconds. 
-This is to be compared with the 42.8\% proved in the same amount of time by a 
-a barebone version of TacticToe.
-
-In the same condtion the ATP Eprover proves ..., when it is run in 
-auto-schedule mode with 128 lemmas predicted and 
+This is to be compared with the 42.9\% proved in the same amount of time by 
+Eprover in auto-schedule mode with 128 lemmas predicted and 
 translated by HolyHammer.
-\todo{to do eprover results running}
+\todo{run experiments including proof scripts containing let
+     run experiments without prove on 1/10th of the library.}
 
@@
 can choose which proof tree to extend next and in which direction.
-Our search algorithm is a monte carlo tree search (MCTS) algorithm that relies 
+Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies 
 on prior evaluation and policy functions. Those priors are learned through 
 direct supervised learning in the case of the policy and via the data 
-accumulated during orthogonalization for the evaluation.
+accumulated during orthogonalization for the evaluation. This MCTS do not rely 
+on roll outs. The 
+policy and evaluation are estimated by the simple supervised k-nn algorithm 
+here instead of neural networks trained through reinforcement learning in 
+AlphaGo Zero~\cite{}.
+
+%However for very large branching factor RTS Combinatorial 
+%Multi-armed Bandits
+%for Real-Time Strategy Games. The method of choice for solving those 
+%problems 
+%is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
+%to each branches
 
@@
 Then we explain how we compute the evaluation and policy functions from learned 
-data. And finally, we show how one MCTS step changes the proof tree data 
-structures.
+data. And finally, we show how the two MCTS step, node selection and node 
+extension changes the proof tree data structures.
 
@@
 
-We assume that any set is numbered and that we can access a nth member of set 
+Unless specified otherwise, we assume that any set is numbered arbitrarily and 
+that we can access a nth member of set 
 $s$ by writing $s_n$. For readability in the next figure $(G(a_j))_i$ is 
-written 
-$g_j^i$.
+written $g_j^i$.
 
@@
 \end{center}
-\caption{A node of a search tree and its relation with the children of one of 
-its goal}
+\caption{A node of a search tree and the children of one of 
+its goal.}
 \label{fig:choice}
@@
 applied to this goal that produces no goals.
-
 \end{definition}
-
-%Since the support of A gives us a bijection between tactics of the tree search 
-%and the node they 
-%produced, any function define on tactics of the tree search can be map to a 
-%function defined on nodes.
-
-Definition of solved goals, open goals  and pending goals.
-Supervised learning introduction.
 
@@
 how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
-tactics $t_{i+1},\ldots,t_500$. We make the strong assumption that $c_policy$ 
+tactics $t_{i+1},\ldots,t_{500}$. We make the strong assumption that $c_policy$ 
 is constant across the search. We can now calculate the prior policy by:
@@
 guidance to our deterministic node selection algorithm over many simulations.
+
+Since the support of A induces a bijection between tactics of the tree 
+search and the node they 
+produced, any function define on tactics of the tree search can be map to a 
+function defined on nodes. 
+Therefore, we define $PriorPolicy(a)=_def PriorPolicy(t)$, if $\exists g 
+A(g,t) = a$.
 
@@
 Experimentally, it is preferable not to 
-filter proven (or descendants of proven) nodes before 
+filter solved (or descendants of solved) nodes before 
 node selection to keep the estimation of the algorithm balanced.
 
+\subsection{Start of the search}
+The input of the algorithm is a goal $g$ that we want to prove.
+Therefore the search tree starts with only one node containing the singleton 
+$g$ and an ordered list of predicted tactics for the goal $g$.
+
 \subsection{Node selection}
 
-A similar trade-off between exploration and exploitation exists in algorithm 
-like Monte Carlo. However, Monte Carlo tree search metrics are based on many 
-simulations and do not reflect our exploration strategy that only
-
-We present here a monte carlo tree search with no roll out which is essentially 
-the same as the one use by AlphaGo Zero~\cite{} originally defined by \cite{} 
-in \cite{}.
+%A similar trade-off between exploration and exploitation exists in algorithm 
+%like Monte Carlo. However, Monte Carlo tree search metrics are based on many 
+%simulations and do not reflect our exploration strategy that only
+
+Through node extension steps the tree grows and  the number of paths to explore 
+increases. To decide which node to extend next, we proceed through the node 
+selection algorithm. Starting from the root node, we repeatedly choses the 
+children with the highest value. To allow the algorithm to select nodes that 
+are not leafs, We also assign a value to the possibility of stopping the 
+descent through the tree at a particular node and trying the next tactic in the 
+predictions. This value has an exploitation equal to 0 since it was never 
+explored.
+
+The value of a node with respect to a search tree, prior policies and prior 
+evaluation is determined by the 
+following computation:
 
@@
 evaluations of descendants of $a_i$.
-
-\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]
-
-\[CurPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
+\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurrentPolicy(a_i)}\]
+
+\[CurrentPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
 The policy is approximatively the  percentage of time a node was visited.The 
@@
 is also called the current evaluation of $a_i$.   
-\[Exploitation(a_i) = CurEvaluation(a_i) = 
-  Average_{b \in Desc(a_i)} {PriorEvaluation(b)}\]
-
-MCTS with prior policy
-This modification of the search tree can be simulated by weighting each arrow 
-by the distance shown by the diagonalized tree. To simulated the 
-diagonalization in the original tree, the arrow $t_i$ will be given weight $i$.
-Methods for exploring a tree with a large branching factor has been explored  
-in Go~\cite{alphago} and
-
-However for very large branching factor RTS Combinatorial Multi-armed Bandits
-for Real-Time Strategy Games. The method of choice for solving those 
-problems 
-is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
-to each branches
-
-In the context of changing wiehgts, the diagonalized tree has the advantage of 
-being a little faster to explore as the ordered of tactic is inscribed in the 
-shape of the tree. if the prior shapes
+\[Exploitation(a_i) = CurrentEvaluation(a_i) = 
+  \average_{b \in Descendant(a_i)} {PriorEvaluation(b)}\]
 
 \subsection{Node extension}
+Once a node is selected, we applied the best not-yet-applied tactics to
+the open goal this node. If the tactic is productive then the we create a new 
+node and calculate and save its prior evaluation. Otherwise, we create a 
+non-accessible node that holds a value of 0 to notice the value function that 
+node selection was a failure. This dummy node is only there to update the 
+result of the Current evaluation of other nodes and is not considered in the 
+calculation of exploration term and cannot be selected.
+
+Maybe should move this defnition in static description.
 
@@
 During a proof search, after a tactic $t$ is applied to a node with active goal 
-$g$, 
-$t$ is called 
+$g$, $t$ is called 
 productive if and only if all the following conditions are satisified:
@@
 ancestor nodes.
-\item It is not a parallel step. $l$ is not a superset of a list of goals 
-produced from the same node with the same active goal.
+\item It is not a parallel step. $l$ is not equal to or a superset of a list of 
+goals produced from the same node with the same active goal.
 \end{itemize}
@@
 
-\subsection{End of the proof search}
+\subsection{End of the search}
 The proof search stops when one of these 3 conditions is realized:

### 2018-01-22 tactictoe_jv.tex
record: simplification

@@
 
+\subsection{The problem}
+
 \todo{JU or CK review}
@@
 
-\subsection{Minimal \sml Code Trees}
+%\subsection{Minimal \sml Code Trees}
+%
+%
+%Using the standard types $string$ and $list$ and their type constructors,
+%we define a minimal code tree by the following AST:
+%\begin{align*}
+%M &=_{def} string \\
+%I &=_{def} string\ \vert\ Module(M,I)\\
+%D &=_{def} Open(M)\ \vert\  Val(I,E) \\
+%E &=_{def} Code(I\ list)\ \vert\   Let (D,E)
+%\end{align*}
+%
+%In this AST, the constructors $Let$, $Val$ and $Open$ correspond to
+%the reserved \sml tokens \texttt{let}, \texttt{val} and \texttt{open} 
+%respectively. $Module(M,I)$ refers to the \sml token \texttt{M.I} where 
+%\texttt{M} is a module (also called structure in \sml) and \texttt{I} is an 
+%identifier. $D$ is a declaration, $E$ an expression. Inside a 
+%declaration $val(I,E)$, $E$ is a definition for $I$.
+%To construct such an AST for a theory file, we rely on a custom parser. It 
+%transform a source file into a 
+%minimal code tree taking into account the scope of each constructor.
+%An additional reason to refrain from the use of the interpreter in our case, 
+%is that
+%the \polyml compiler needs to execute the code preceding an expression 
+%before parsing it, which is significantly slower than a custom parser.
+
+\subsection{Globalization}
+
 \holfour proofs are organized in theory files.
@@
 that correspond to more involved proofs.
-Functions that correspond to backward proofs, called tactics, are combined with each other with the help of 
+Functions that correspond to backward proofs, called tactics, are combined with 
+each other with the help of 
 tacticals (functions that takes
-tactics as arguments). Such a combination of tactics will be called a \emph{proof script} when it
+tactics as arguments). Such a combination of tactics will be called a 
+\emph{proof script} when it
 proves a theorem stated in a theory.
 
-In order to apply machine learning to the existing proof scripts, we will need to
+In order to apply machine learning to the existing proof scripts, we will need 
+to
 parse and modify them, which involves parsing and modification of \sml code.
@@
 As some \holfour theories are built in separate sessions, it is not possible
-to transfer \sml values directly, but instead tactic code needs to transferred instead.
+to transfer \sml values directly, but instead tactic code needs to transferred 
+instead.
 To do this, the transferred information will be saved externally.
 
-Using the standard types $string$ and $list$ and their type constructors,
-we define a minimal code tree by the following AST:
-\begin{align*}
-M &=_{def} string \\
-I &=_{def} string\ \vert\ Module(M,I)\\
-D &=_{def} Open(M)\ \vert\  Val(I,E) \\
-E &=_{def} Code(I\ list)\ \vert\   Let (D,E)
-\end{align*}
-
-In this AST, the constructors $Let$, $Val$ and $Open$ correspond to
-the reserved \sml tokens \texttt{let}, \texttt{val} and \texttt{open} 
-respectively. $Module(M,I)$ refers to the \sml token \texttt{M.I} where 
-\texttt{M} is a module (also called structure in \sml) and \texttt{I} is an 
-identifier. $D$ is a declaration, $E$ an expression. Inside a 
-declaration $val(I,E)$, $E$ is a definition for $I$.
-To construct such an AST for a theory file, we rely on a custom parser. It transform a source file into a 
-minimal code tree taking into account the scope of each constructor.
-An additional reason to refrain from the use of the interpreter in our case, is that
-the \polyml compiler needs to execute the code preceding an expression 
-before parsing it, which is significantly slower than a custom parser.
-
-\subsection{Globalization}
+%\subsubsection {Unfolding locally declared values}
 To transfer \sml values as code across theories in a conservative manner, we 
@@
 
-For a more formal explanation, we define a target AST build upon the original 
-one, where each identifer $I$ is replace by its globalized expression $E'$:
-
-\begin{align*}
-M' &=_{def} string \\
-I' &=_{def} string\ \vert\ Module(M',I')\\
-D' & =_{def} Open(M')\ \vert\  Val (I',E') \\
-E' &=_{def} Code(E'\ list)\ \vert\ Let (D',E') 
-\end{align*}
-
-Then, we express two mutually recursive functions $f_d$ and 
-$f_e$ that globalize the original representation. The 
-function $f_d$ modifies declaration and $f_e$ modifies expressions. Since a 
-\sml program is a list of declarations, we will be able transform it by 
-iteratively applying $f_d$. The stack 
-$\mathbb{S}$ keep track of each previous declaration: the identifier that is 
-declared and their globalized definition. To save a declaration to 
-the stack, we use the $push$ 
-function. Opening a module has the effect of pushing all identifiers with 
-their associated prefixed version to the stack. The scope of an identifier is 
-automatically handled by construction of the AST. To find the globalized 
-version of an identifier $i$, we use the $lookup$ function on a stack $s$, 
-which returns the globalization associated with $i$ appearing highest in $s$. 
-We then build $f_e$ and $f_d$ in the following way:
-\vspace{5mm}
-
-\begin{math}
-\begin{array}{lcl}
-f_d:& \mathbb{S} \times D &\rightarrow\ \ \mathbb{S} \times D'\\
-& (s, Open(m)) &\mapsto\ \ (push(m,s),\ Open(m)) \\
-& (s, Val(i,e)) &\mapsto\ \ (push ((i,\hat{e}),s),\ Val(i, \hat{e})) 
-\mbox{ where } \hat{e} = \pi_2^2 (f_e (s,e))
-\end{array}
-\end{math}
-
-\vspace{5mm}
-
-\begin{math}
-\begin{array}{lcl}
-f_e:& \mathbb{S} \times E &\rightarrow\ \ \mathbb{S} \times E'\\
-& (s, List(l))  &\mapsto\ \ (s,\ lookup(s,l))\\
-& (s, Let (d,e)) &\mapsto\ \ (\hat{s}, f_e (\hat{s}, e))
-\mbox{ where } \hat{s} = \pi_1^2 (f_d (s,d))
-\end{array}
-\end{math}
-
-\vspace{5mm}
+%For a more formal explanation, we define a target AST build upon the original 
+%one, where each identifer $I$ is replace by its globalized expression $E'$:
+%
+%\begin{align*}
+%M' &=_{def} string \\
+%I' &=_{def} string\ \vert\ Module(M',I')\\
+%D' & =_{def} Open(M')\ \vert\  Val (I',E') \\
+%E' &=_{def} Code(E'\ list)\ \vert\ Let (D',E') 
+%\end{align*}
+%
+%Then, we express two mutually recursive functions $f_d$ and 
+%$f_e$ that globalize the original representation. The 
+%function $f_d$ modifies declaration and $f_e$ modifies expressions. Since a 
+%\sml program is a list of declarations, we will be able transform it by 
+%iteratively applying $f_d$. The stack 
+%$\mathbb{S}$ keep track of each previous declaration: the identifier that is 
+%declared and their globalized definition. To save a declaration to 
+%the stack, we use the $push$ 
+%function. Opening a module has the effect of pushing all identifiers with 
+%their associated prefixed version to the stack. The scope of an identifier is 
+%automatically handled by construction of the AST. To find the globalized 
+%version of an identifier $i$, we use the $lookup$ function on a stack $s$, 
+%which returns the globalization associated with $i$ appearing highest in $s$. 
+%We then build $f_e$ and $f_d$ in the following way:
+%\vspace{5mm}
+%
+%\begin{math}
+%\begin{array}{lcl}
+%f_d:& \mathbb{S} \times D &\rightarrow\ \ \mathbb{S} \times D'\\
+%& (s, Open(m)) &\mapsto\ \ (push(m,s),\ Open(m)) \\
+%& (s, Val(i,e)) &\mapsto\ \ (push ((i,\hat{e}),s),\ Val(i, \hat{e})) 
+%\mbox{ where } \hat{e} = \pi_2^2 (f_e (s,e))
+%\end{array}
+%\end{math}
+%
+%\vspace{5mm}
+%
+%\begin{math}
+%\begin{array}{lcl}
+%f_e:& \mathbb{S} \times E &\rightarrow\ \ \mathbb{S} \times E'\\
+%& (s, List(l))  &\mapsto\ \ (s,\ lookup(s,l))\\
+%& (s, Let (d,e)) &\mapsto\ \ (\hat{s}, f_e (\hat{s}, e))
+%\mbox{ where } \hat{s} = \pi_1^2 (f_d (s,d))
+%\end{array}
+%\end{math}
+%
+%\vspace{5mm}
 
@@
 
-Thanks to the simplicity of the minimal AST, we are able to present our 
-recursive unfolding algorithm in a concise and formal manner. But for practical 
-purpose, the minimal AST is not sufficient. So, we extend the scope of our 
-algorithm by enumerating common \sml constructions and explain 
-independently how we globalize them. We first start by mentioning additional 
-details about module and value declarations.
-
-\paragraph{Modules}
-Declaring a module is made through the keyword 
-\texttt{open}. We can fetch 
-the values declared in a module \texttt{m} by first erasing all elements from 
-\texttt{PolyML.NameSpace} and then calling \texttt{open m}. Additionally, it 
-gives us the information of whether a value is a constructor or an exception. 
-This will prove to be particularly useful to handle pattern matching 
-construction correctly. During the opening of a module, nested modules can be 
-declared and push to the stack. In this way, we can globalize them by prefixing 
-them by their parent modules when they appear in the program.
-
-\paragraph{Values}
-The standard ML language allows to declare multiple values in a single $let$ 
-construction. In effect, this list of declarations is equivalent to nested 
-declarations.
-
-Because we do not want to unfold proofs into other proofs, we 
-make an exception 
-for values of type theorem and do not replace them by their definition.
-Instead, we fetch theorems from the 
-database by looking for its formula in the database of theorems.
-Because this database is reconstructed from files saved on disk, a call to 
-the function \texttt{DB.fetch} works across theories.
-
-\paragraph{Functions}
-To globalize a function identifier, we include 
-its declaration in the globalized expression. 
-Given a declaration \texttt{fun f v = e}, the globalized 
+%Thanks to the simplicity of the minimal AST, we are able to present our 
+%recursive unfolding algorithm in a concise and formal manner. But for 
+%practical 
+%purpose, the minimal AST is not sufficient.
+
+We present the tactical proof recording mechanism in details here because it 
+can help 
+another researcher to produce a tactical export in other interactive theorem 
+provers.
+We now enumerate common \sml constructions and explain 
+independently how we globalize them. We use a stack to keep track of previous 
+identifier declaration and their globalization.
+
+%We first start by mentioning additional 
+%details about module and value declarations.
+
+\subsection{Storing Declaration}
+
+\paragraph{Reserved tokens}
+
+\paragraph{Module}
+
+\paragraph{Global identifier}
+
+\paragraph{Local identifier}
+
+To globalize a identifier, we include 
+the code of its declaration in the globalized expression. 
+Given a declaration \texttt{fun f v = e} or \texttt{val f = e}, the globalized 
 expression for the identifier \texttt{f} is \texttt{let fun f v = e' in f 
-end}, where \texttt{e'} is the globalized version of \texttt{e}. The function 
-\texttt{f} and its argument \texttt{v} are binders. So, we prevent our 
-algorithm from trying to globalize them during the construction of \texttt{e'}.
-This approach is slightly more verbose than using an 
-anonymous function construction. But it is prettier as 
-we regroup declarations at the top of the proof expression before returning it
-to the user (see Section~\ref{sec:proofdisplay}).
-
-\paragraph{Pattern matching}
+end} respectively \texttt{let val f = e' in f end}
+The expression \texttt{e'} is the globalized version of \texttt{e}. That is to 
+say, it is the expression \texttt{e} where each of its tokens has been replaced 
+its globalized expressions as constructed in Section~\ref{}.
+
+In the case of value, we can simplify the globalized expression further and 
+replace it by \texttt{e'}.
+In the case of a function, \texttt{f} and its argument \texttt{v} are binders. 
+So, we prevent their globalization during the construction of \texttt{e'}.
+%This approach is slightly more verbose than using an 
+%anonymous function construction. But it is prettier as 
+%we regroup declarations at the top of the proof expression before returning it
+%to the user (see Section~\ref{sec:proofdisplay}).
+
+\paragraph{Bound variables}
 A pattern is an \sml expression that 
@@
 \texttt{p}, the globalized expression \texttt{e'} for \texttt{x} becomes 
-\texttt{let val p = e' in x end}.
+\texttt{let val p = e' in x 
+end}.
 
@@
 is not prefixed by a module. The infixity applies only when an identifier 
-exactly match $s$. Therefore, an identifier \texttt{x} could have been treated 
+exactly match $s$.
+
+\subsection{Globalizing Expression}
+
+\paragraph{Reserved tokens}
+
+\paragraph{Modules}
+Declaring a module is made through the keyword 
+\texttt{open}. We can fetch 
+the values declared in a module \texttt{m} by first erasing all elements from 
+\texttt{PolyML.NameSpace} and then calling \texttt{open m}. Additionally, it 
+gives us the information of whether a value is a constructor or an exception. 
+This will prove to be particularly useful to handle pattern matching 
+construction correctly. During the opening of a module, nested modules can be 
+declared and push to the stack. In this way, we can globalize them by prefixing 
+them by their parent modules when they appear in the program.
+
+\paragraph{Global identifier}
+Global values and functions are prefixed by the modules and submodules where 
+they were declared. This information obtained by looking up the first 
+identifier that corresponds to the values at the top of the stack.
+
+\paragraph{Local identifier}
+Each local identifier is replaced by its globalized expression.
+
+We do not unfold local and global values of type theorem. Indeed, we do not 
+want to 
+unfold proofs into other proofs. Instead, we fetch theorems by looking for its 
+formula in the global database of 
+theorems. If it does not exists, we consider the tactic to be local and not 
+transferable to other theories.
+
+\paragraph{Bound variables}
+binding variables are not globalized
+
+\paragraph{Infixity}
+ Therefore, an identifier \texttt{x} could have been treated 
 as infix in an original expression \texttt{E} but when 
@@
 \begin{lstlisting}
-fun a infixlX_open f = fn x => f (a,x)  
-fun g infixlX_close b = g b
-
-fun f infixrX_close b = fn x => f (x,b)
-fun a infixrX_open g = g a
+fun a infixlX_open f = fn x ==> f (a,x)      fun g infixlX_close b = g b
+fun a infixrX_open g = g a      fun f infixrX_close b = fn x ==> f (x,b) 
 \end{lstlisting}
 \end{definition}
-
-\begin{proof}(Does anyone want a proof that this functions are correctly 
-reproducing the original function?)
-\end{proof}
 
@@
 $g$ and an ordered list of predicted tactics for the goal $g$.
+Then, the MCTS algorithm alternates node selection steps and node extension 
+steps which grow the search tree.
 
@@
 \subsection{Node extension}
-Once a node is selected, we applied the best not-yet-applied tactics to
-the open goal this node. If the tactic is productive then the we create a new 
+Once a node is selected, we applied the best not-yet-applied tactics according 
+to the prediction order to the open goal this node. If the tactic is productive 
+then the we create a new 
 node and calculate and save its prior evaluation. Otherwise, we create a 

### 2018-01-23 tactictoe_jv.tex
related works: wip

@@
 Experiments on the \holfour standard 
-library show that TacticToe is able to re-prove 66.2\% of all 7164 theorems in 
-less 
-than 60 seconds. 
-This is to be compared with the 42.9\% proved in the same amount of time by 
+library show that TacticToe is able to re-prove 66.0\% of 
+all theorems proven by a tactical script in the standard library.
+This is to be compared with the 34.5\% proved in the same amount of time by 
 Eprover in auto-schedule mode with 128 lemmas predicted and 
 translated by HolyHammer.
-\todo{run experiments including proof scripts containing let
-     run experiments without prove on 1/10th of the library.}
 
@@
 \paragraph{Bound variables}
-binding variables are not globalized
+Binding variables are not globalized
 
@@
 \begin{definition}(Code of infix guards)
-\begin{lstlisting}
+\begin{lstlisting}[language=SMLSmall,frame=tb]
 fun a infixlX_open f = fn x ==> f (a,x)      fun g infixlX_close b = g b
@@
 \begin{example} (Globalization)
-\begin{lstlisting}
+\begin{lstlisting}[language=SMLSmall,frame=tb]
 Tactic.STRIP_TAC 
@@
 \begin{example}(Proof script transformation)
-\begin{alltt}
-original declaration: val id = store_thm (name, term, p)
-modified declaration: val id = store_thm (name, term, Split "p'")
-\end{alltt}
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+original file: val id = store_thm (name, term, p)
+modified file: val id = store_thm (name, term, Split "p'")
+\end{lstlisting}
 \end{example}
@@
 Most tacticals are infix, and we do not want to consider combination of 
-tactics to be a tactic unit. For our wrapper \texttt{Split}, a tactic unit is a 
+tactics to be a tactic unit. For \texttt{Split}, a tactic unit is a 
 subterm of 
@@
 \begin{example} (Wrap)
-\begin{alltt}\small
+\begin{lstlisting}[language=SMLSmall,frame=tb]
 R "Tactic.STRIP_TAC"
   infixl0_open boolLib.THEN infixl0_close
-R "Prim_rec.INDUCT_THEN (DB.fetch \bq{list}\bq{INDUCT}) Tactic.ASSUME_TAC"
+R "Prim_rec.INDUCT_THEN (DB.fetch \"list\" \"INDUCT\") Tactic.ASSUME_TAC"
   infixl0_open boolLib.THEN infixl0_close
 R "Rewrite.ASM_REWRITE_TAC 
-     [DB.fetch \bq{list} \bq{MAP}, DB.fetch \bq{list} \bq{APPEND}]"
-\end{alltt}
+     [DB.fetch \"list\" \"MAP\", DB.fetch \"list\" \"APPEND\"]"
+\end{lstlisting}
 \end{example}
@@
 \begin{example}\label{ex:record} (Code of the recording function)
-\begin{lstlisting}
+\begin{lstlisting}[language=SMLSmall,frame=tb]
 fun R tactic_string goal = 
-  (save (tactic_string, goal); (read_tactic tactic\_string) goal)
+  (save (tactic_string, goal); tactic_of_sml tactic_string goal)
 \end{lstlisting}
@@
 The time $time$ taken by $tactic$ is not currently used by our 
-predictor. The list of goals $goal list$ is the output produced by $tactic$ on 
+predictor. The list of goals $goal list$ is the output of $tactic$ on 
 $goal$. And the features $features$ of the goal $goal$ are extracted and hashed 
@@
 To simplify the description of our machine learning algorithms in future 
-sections, we consider a feature vector to be a pair $(tactic,goal)$ and only 
+sections, we consider a tactic feature vector to be a pair $(tactic,goal)$ and 
+only 
 explicitly rely on the 
@@
 
-All top-level theorems in the standard library are considered except:
-\begin{itemize}
-\item the 619 calls to store\_thm that contains a let binding, these are 
-typically 
-harder for \tactictoe as it may not have seen the particular local declaration 
-before and therefore would have to find an alternative proofs. We also do not 
-learn from these proofs because they were explictly local.
-\item the 1175 calls to save\_thm. These are typically mostly trivial and easy 
+7000+ top-level theorems that were created by . 
+In this evaluation, we do not include the 250 theorems created by
+
+All top-level theorems created by a call to 
+\texttt{store\_thm}, 
+\texttt{Q.store\_thm} or \texttt{maybe\_thm} are considered with the exception 
+of 440 calls that contains a let binding. They are typically harder to re-prove.
+We also ignore the 1175 \texttt{save\_thm} calls. Most of them are easily 
+provable theorems using instantiations or conjuncts of previous 
 theorems.
-The remaining 7088 top-level theorems are evaluated in the full scale 
-evaluation and the level of difficulty of these problem is assessed by a 
-comparison with \eprover.
-\end{itemize}
+
+%\begin{itemize}
+%%\item the , these are 
+%%typically 
+%%harder for \tactictoe as it may not have seen the particular local 
+%%declaration 
+%%before and therefore would have to find an alternative proofs. We also do not 
+%%learn from these proofs because they were explictly local.
+%%\item the 1175 calls to save\_thm. These are typically mostly trivial and 
+%%easy 
+%%theorems.
+%%The remaining 7088 top-level theorems are evaluated in the full scale 
+%%evaluation and the level of difficulty of these problem is assessed by a 
+%%comparison with \eprover.
+%\end{itemize}
 
@@
 \midrule
-  & \eprover (auto-schedule + 128 premises) & 86 & 2069 (29.19\%)\\ 
+  & \eprover & 86 & 2069 (29.19\%)\\ 
 %  & \tactictoe (without special care for metis) & 154 & 3868 (\\
-  & \tactictoe (best parameters) & 187 & 4186 (59.06\%) \\
+  & \tactictoe & 187 & 4186 (59.06\%) \\
 %  & \tactictoe (+ self-learning) & 187 (187) &  4201 (59.27\%) \\
@@
 
-\section{Related Work}\todo{TG reorganise + other motivation?}
+\section{Related Work}
 There are several essential components of our work that are comparable to 
@@
 essential for the
-tactic recording procedure, used to train our tactic prediction module.
+tactic recording procedure, used to train our tactic prediction mechanism.
 
@@
 \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
-
 groups related proofs thanks to its clustering 
@@
 Concerning the proof search, \textsf{SEPIA}'s %simple 
-breadth-first search is replaced by an A*-algorithm which allows for heuristic 
-guidance in 
-the exploration of the search tree.
-Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
-theorems) of the 
+breadth-first search is replaced by MCTS which allows for supervised learning
+guidance in the exploration of the search tree.
+Finally, \textsf{SEPIA} was evaluated on three chosen parts of the 
 \coq library demonstrating that it globally outperforms individual \coq 
 tactics. In contrast, we demonstrate the competitiveness of our system against 
-the successful general-purpose hammers on the \holfour standard library (7902 
-theorems).
-
-Machine learning has also been used to advise the best library lemmas for new 
-ITP goals.
-This can be done either in an interactive way, when the user completes the 
-proof based on the recommended lemmas, as in the \textsc{Mizar Proof 
-Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma 
-selection is handed over to the ATP component of a \emph{hammer} 
-system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.
-
-Internal learning-based selection of tactical steps inside an ITP is analogous 
-to internal learning-based selection of clausal steps inside ATPs such as 
-\textsc{MaLeCoP}~\cite{malecop} and \textsc{FEMaLeCoP}~\cite{femalecop}. These 
-systems
-use the naive Bayes classifier to  select clauses for the extension steps in
-tableaux proof search based on many previous proofs. Satallax~\cite{Brown2012a} 
-can guide its
-search internally~\cite{mllax} using a command classifier, which can estimate 
-the priority of the 11 kinds of
-commands in the priority queue based on positive and negative examples.
-There are several essential components of our work that are comparable to 
-previous approaches: tactic-level proof recording, tactic 
-selection through machine learning techniques and automatic tactic-based proof 
-search. Our work is also related to previous approaches that use machine 
-learning to select premises for the ATP systems and guide ATP proof search 
-internally.
-
-For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15} 
-can transform a packed tactical proof into a series of interactive tactic 
-calls. Its principal application 
-was so far refactoring the library and teaching common proof techniques to new 
-ITP users. In our work, the splitting of a proof into a sequence of tactics is 
-essential for the
-tactic recording procedure, used to train our tactic prediction mechanism.
-
-The system 
-\textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
-
-groups related proofs thanks to its clustering 
-algorithms. It allows \coq users to inspire themselves from similar proofs and 
-notice 
-duplicated proofs. Our predictions comes from a much more detailed description 
-of the open goal.
-However, we simply create a single label for each tactic call whereas each of 
-its
-arguments is treated independently in \textsf{ML4PG}. 
-Our choice is motivated by the k-NN algorithm already used in
-\holyhammer for the selection of theorems.
-%, which can be easly adapted to the selection of tactics but would be improved 
-%by a 
-%version that can learn to predict tactics and their arguments independently.
-
-\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to 
-generate
-proof scripts from previous \coq proof examples.
-Its strength lies in its ability to produce likely sequences 
-of tactics for solving domain specific goals. It operates by creating a model 
-for common sequences of tactics for a specific library.
-This means that in order to propose the following tactic, only the previously 
-called tactics
-are considered.
-%intuition tells us that it should not be the main factor
-Our algorithm, on the other hand, relies mainly on the characteristics of the 
-current goal 
-to decide
-which tactics to apply next. In this way, our learning mechanism has to 
-rediscover why each 
-tactic was applied for the current subgoals. It may lack some useful bias for 
-common sequences 
-of tactics, but is more reactive to subtle changes. Indeed, it can be trained 
-on a large library and only tactics relevant to the current subgoal will be 
-selected. 
-Concerning the proof search, \textsf{SEPIA}'s %simple 
-breadth-first search is replaced by MCTS which allows for heuristic 
-guidance in 
-the exploration of the search tree.
-Finally, \textsf{SEPIA} was evaluated on three chosen parts (totaling 2382 
-theorems) of the 
-\coq library demonstrating that it globally outperforms individual \coq 
-tactics. In contrast, we demonstrate the competitiveness of our system against 
-the successful general-purpose hammers on the \holfour standard library.
-A recent improvement on this system~\cite{} includes the possibility to replace 
-theorems in tactics by ones belonging to the same cluster, independently of the 
-current goal. We have the opposite approach, we forget about the original 
-theorems in the tactic creating an more general abstract tactic
-and only use the goal to guide predictions. Finding a way to combine those two 
-approaches could be the key to stronger theorem predictions.
+the ATP \eprover on the \holfour standard library.
+A recent improvement on this system~\cite{}\todo{Cezary: could not remember the 
+reference that you gave me here i think it was CICM 2017} includes the 
+possibility to 
+substitute theorems appearing in tactics by ones belonging to the same cluster. 
+In this work, we are also able to predict theorems appearing as arguments of 
+tactics but we rely only on information from the current goal. Combining the 
+two approaches by predicting from the goal and from the original tactic 
+arguments is more difficult but is likely to produce more suitable lemmas.
 
@@
 \sledgehammer~\cite{sledgehammer10} or Nitpick~\cite{Nitpick10}. 
-Selection of tactics from recorded proofs has yet to be implemented and now it 
-is up to come up with a good meta-tactic.
-
-%In \holfour, 
-%a tactic is a just a \sml function of a special type and \holyhammer can 
-%easily 
-%be modified to be a tactic. So this meta-level is not 
-%necessary, all programming can occur at the tactic level and even \tactictoe  
-%is a tactic.
-
-%* Theory Exploration (Moa Johanson+, ITP 2017)  I think that is more related 
-%to conjeturing and do not think it fits here.
+Selection of tactics from data collected in recorded proofs is in development 
+but for now it is up to the user to come up with a good meta-tactic.
 

### 2018-01-24 tactictoe_jv.tex
related works: wip
updating tactictoe data with the 60 seconds version

@@
 
-\subsection{The problem}
-
-\todo{JU or CK review}
+\subsection{The problem}\todo{}
+
+Two ways of representing a formulas.
+
+Defintion of a tactic in \holfour. In this research, we will take the definition
+that a tactic takes a goal and returns a list of goals.
+
 \subsection{Contributions}
@@
 \item term:\\
+Predict the closest term in a tactic that appears in the goal.
 
 \item theorem:\\
+Predict the theorem the closest to the current goal. Theorem dependencies are 
+added during preselection.
 
 \item list of goals:\\
-
+Used for Monte Carlo Tree Search evaluation. Predict the closest list of goals
+to the list of goal produced by a tactic.
 \end{itemize}
@@
 
-%However for very large branching factor RTS Combinatorial 
-%Multi-armed Bandits
-%for Real-Time Strategy Games. The method of choice for solving those 
-%problems 
-%is monte carlo tree search~\cite{mcts} which is a way to assign probabilities 
-%to each branches
-
 So, we first give a static description of the proof tree and its component.
@@
 extension changes the proof tree data structures.
-
-%A proof is in general not a path but a tree. This means that our search space 
-%has two branching factors: 
-%the choice of a tactic, and the number of goals produced by each tactic.
-%The first one can be thought as a disjunction of tactics. Indeed, 
-%if there exists a tactic that solves a goal then it is solved. And 
-%conversely, the second branching factor is a conjunction of goals. All pending 
-%goals must be proven in order to prove the parent goal. For example, in the 
-%case of induction tactics the pending goals are the base case and the 
-%inductive 
-%case.
-
-%If we suppose that each tactic produces at most one goal, then the standard 
-%monte carlo tree search algorithm applies. So we will first discuss how our 
-%algorithm works in this case. And then explain how we extend it to support 
-%tactics producing 
-%multiple goals.
-
-%Our search algorithm takes inspiration from the A*-algorithm~\cite{HartNR68} 
-%which uses 
-%a cost function and heuristics to estimate the total distance to the 
-%destination and
-%choose the shortest route.
-%The proof is not finished when the current tactic solves its goal because it 
-%often leaves new pending open goals along the current path.
-
-%\subsection{Diagonalized tree search}
-%(Move this section to Monte Carlo tree search and simplify it) 
-%(Don't mention diagonlazied tree search)
-
-%%
-%%Our first attempt was to create a breadth first search algorithm with 
-%%preferences for the best predicted tactic for each goal. In 
-%%Figure~\ref{fig:br1}, we try to prove the goal $g$ using list of tactics 
-%%$t_1,t_2,t_3 \ldots$ ordered by the predictor. Those tactics produce 
-%%respectively the goals $g_1,g_2,g_3 \ldots$. By choosing tactics
-%%for each produced goal $g_i$, we can recursively construct the whole search 
-%%tree. We observe that if we apply the breadth first search algorithm to this 
-%%tree then the predictor preferences are of little use, as we will not explore 
-%%deeper more promising branches. Another related inconvenience is that we may 
-%%not ever be able to explore the tree at depth 2 since there maybe too many 
-%%tactics to try at depth 1. One way to solve those this problem is to 
-%%diagonalized our tree by adding dummy square nodes as presented in 
-%%Figure~\ref{fig:diag_choice}. The new branching factor of this tree is 
-%%constant 
-%%equal to two. The exploration of lower ranked tactics are delayed by the 
-%%diagonalization during breadth search. And after $2^n$ breadth-first steps, 
-%%the 
-%%most promising branch has been explored up to depth $n$.
-%%
-%%\begin{figure}{}
-%%\begin{center}
-%%\begin{tikzpicture}[node distance=2cm]
-%%\node [tcircle] (0) {$g$};
-%%\node [above of=0] (m0) {};
-%%\node [tcircle, left of=m0, node distance=1cm] (2) {$g_2$};
-%%\node [tcircle, left of=2] (1) {$g_1$};
-%%\node [tcircle, right of=m0,node distance=1cm] (3) {$g_3$};
-%%\node [right of=3] (inf) {};
-%%
-%%\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1);
-%%\draw[-to,thick] (0) to node[xshift=-10](t2){$t_2$} (2);
-%%\draw[-to,thick] (0) to node[xshift=-10](t3){$t_3$} (3);
-%%\draw[-to,dotted,thick] (0) to (inf);
-%%\end{tikzpicture}
-%%%\end{center}
-%%%\caption{Choice of a tactic}
-%%%\label{fig:choice}
-%%%\end{figure}
-%%%\begin{figure}{}
-%%%\begin{center}
-%%\begin{tikzpicture}[node distance=1cm]
-%%\node [tcircle] (0) {$g$};
-%%\node [above of=0] (m0) {};
-%%\node [tcircle, left of=m0] (1) {$g_1$};
-%%\node [tsquare, right of=m0] (s1) {};
-%%\node [above of=s1] (m1) {};
-%%\node [tcircle, left of=m1] (2) {$g_2$};
-%%\node [tsquare, right of=m1] (s2) {};
-%%\node [above of=s2] (m2) {};
-%%\node [tcircle, left of=m2] (3) {$g_3$};
-%%\node [right of=m2] (s3) {};
-%%
-%%\draw[-to,thick] (0) to node[xshift=-5,yshift=-5](t1){$t_1$} (1);
-%%\draw[-to,thick] (0) to (s1);
-%%\draw[-to,thick] (s1) to node[xshift=-5,yshift=-5](t1){$t_2$} (2);
-%%\draw[-to,thick] (s1) to (s2);
-%%\draw[-to,thick] (s2) to node[xshift=-5,yshift=-5](t1){$t_3$} (3);
-%%\draw[-to,dotted,thick] (s2) to (s3);
-%%\end{tikzpicture}
-%%\end{center}
-%%\caption{Transformation of a search tree into a diagonalized search tree,
-%%caused   by a change in the relation between a node and its 
-%%children.}
-%%\label{fig:diag_choice}
-%%\end{figure}
-%%
-%%
-%%In the particular case where tactics produces at most one goal, a proof is 
-%%found in the search tree when a tactic $t$ does not produce any new goal. The 
-%%proof of the root goal can be reconstructed by extracting the path leading 
-%%to the final tactic $t$.
-%%
-%%
-%%
-%%\subsection{General case}
-%%We now tackle the more general case where tactics $t_i$ may produces multiple 
-%%goals $g_{i,1}$ $g_{i,2}$ $g_{i,n}$ from a parent goal $g_p$.
-%%
-%%All created node are kept in memory for later inspection but some of the 
-%%nodes 
-%%are deactivated which means they are removed from the proof search.
-%%
-%%Definition (deactivated node):
-%%  A solved node or a node that has a solved ancestor.
-%%
-%%Then at first, the breadth first search operates normally with only the first 
-%%goal $g_{i,1}$ is present in the search tree. 
-%%And as soon as a proof is found for the goal $g_{i,1}$, the subtree with root 
-%%$g_{i,1}$ is erased and replaced by a node containing $g_{i,2}$. The process 
-%%continues with node with the lowest depth being treated first.
-%%The process stops if the root goal is solved, the time limit is exceeded or 
-%%there is no more tactic to be applied to an active node.
+Intuitively, MCTS offers a good trade-off between exploitation and exploration. 
+It means that the algorithm will exploit deeper more promises branches but 
+leaving enough time for exploring less likely alternatives.
 
@@
 
-\begin{definition}(Tactical search tree)\\
-A tactical search tree $\mathfrak{T}$ is a sixtuple 
+\begin{definition}(search tree)\\
+A search tree $\mathfrak{T}$ is a sixtuple 
 $(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
@@
 goal is expressed through mutual recursion with a 
-solved node and is given in ~\ref{def:solved_goal}. An open goal is goal that 
+solved node and is given in Definition~\ref{def:solved_goal}. An open goal is 
+goal that 
 was tried at least once by a tactic. In our implementation we have only one 
@@
 Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
-$\mathbb{A}^\infty$ 
-and 
+$\mathbb{A}^\infty$ and 
 the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:
@@
 
-%A proof is in general not a path but a tree. This means that our search space 
-%has two branching factors: 
-%the choice of a tactic, and the number of goals produced by each tactic.
-%The first one can be thought as a disjunction of tactics. Indeed, 
-%if there exists a tactic that solves a goal then it is solved. And 
-%conversely, the second branching factor is a conjunction of goals. All pending 
-%goals must be proven in order to prove the parent goal. For example, in the 
-%case of induction tactics the pending goals are the base case and the 
-%inductive 
-%case.
-
 \subsection{Prior tactic policy}
@@
 function defined on nodes. 
-Therefore, we define $PriorPolicy(a)=_def PriorPolicy(t)$, if $\exists g 
-A(g,t) = a$.
+Therefore, we define $PriorPolicy(a)$ as the $PriorPolicy(t)$ where $t$ is the 
+tactic that produced $a$.
 
@@
 \end{definition}
-
-\subsection{Evaluation}
 
@@
 
-Mention the conflicts.
-Theorem orthogonalization is not turn on during instantiation of tacitcs except 
-for the metis tactic.
-Best parameters + comparison with holyhammer.
-asynchronous eprover is not used  for fairer comparison with eprover.
-
-10 seconds experiment:
-
+%Mention the conflicts.
+%Theorem orthogonalization is not turn on during instantiation of tacitcs 
+%except 
+%for the metis tactic.
+%Best parameters + comparison with holyhammer.
+
+\eprover is not called by \tactictoe for fair comparison.
+
+\begin{table}[h!]
+\centering\ra{1.3}
+\small
+\begin{tabular}{lll}
+\toprule
+  & & 7164 theorems \\
+\midrule
+  & \eprover   & 2472 (34.5\%)\\ 
+  & \tactictoe & 4741 (66.2\%)\\
+  & combined   & 4946 (69.0\%)\\
+\bottomrule
+\end{tabular}
+\caption{Experiments with a timeout of 60 seconds \label{tab:_param}}
+\end{table}
+
+Table~\ref{theories} compares the success rates of re-proving for different
+\holfour theories. \tactictoe outperforms \eprover on every 
+considered theory.
+\eprover is more suited to deal with dense theories such as 
+\texttt{real} or \texttt{complex} where a lot of related theorems are available 
+and most proofs are usually completed by rewriting tactics.
+
+\begin{table}[]
+\centering
+\setlength{\tabcolsep}{3mm}
+\begin{tabular}{@{}ccccc@{}}
+\toprule
+\phantom{ab} & {arith} & {real} & {compl} & {meas} \\
+\midrule
+\tactictoe & 81.2 & 74.0 & 79.6 & 31.3\\
+\eprover & 59.9 & 72.0 & 67.1 & 12.8\\
+\midrule
+\phantom{abc} & {proba} & {list} & {sort} & {f\_map} \\
+\midrule
+\tactictoe & 45.8 & 79.5 & 65.3 & 82.0 \\
+\eprover & 24.1 & 26.5 & 15.8 & 24.7 \\
+\bottomrule
+\end{tabular}
+\caption{\label{theories}Percentage (\%) of re-proved theorems in the theories 
+\texttt{arithmetic}, \texttt{real}, \texttt{complex}, \texttt{measure},  
+\texttt{probability}, \texttt{list}, \texttt{sorting} and \texttt{finite\_map}. 
+}
+\end{table}
+
+\subsection{Slight improvements}
+The final results could be improved by a few percent by including the 
+following additional techniques that are already available.
+
+\paragraph{Asynchronous calls to Eprover inside TacticToe}
+The technique already discussed in Section~ would most likely improve the 
+results by 2-3 percent.
+
+\paragraph{Self-learning}
 Until now, all our feature vectors were learned form human proofs. 
@@
 scripts~\cite{DBLP:conf/cade/Urban07}. It is also the first step toward 
-self-reinforcement learning.
-
-\begin{table}[h!]
-\centering\ra{1.3}
-\small
-\begin{tabular}{llll}
-\toprule
-  & & Small(273) & Big(7088) \\
-\midrule
-  & \eprover & 86 & 2069 (29.19\%)\\ 
-%  & \tactictoe (without special care for metis) & 154 & 3868 (\\
-  & \tactictoe & 187 & 4186 (59.06\%) \\
-%  & \tactictoe (+ self-learning) & 187 (187) &  4201 (59.27\%) \\
-\bottomrule
-\end{tabular}
-\caption{Experiments with a timeout of 10 seconds \label{tab:_param}}
-\end{table}
-
-It is learning to use \metis however it cannot automatically learn to give it 
-more time making it still a bit weaker than the adhoc startegy.
-
-Table~\ref{theories} compares the success rates of re-proving for different
-\holfour theories. \tactictoe(SH) outperforms \tactictoe(NH) on every 
-considered theory.
-Even if a lot weaker due to the missing premise selection component, 
-\tactictoe(NH)
-is hugely better than \holyhammer(blistr) in 4 theories: \texttt{measure}, 
-\texttt{list}, \texttt{sorting} and \texttt{finite\_map}. The main reason is 
-that 
-those theories are equipped with specialized tactics, performing complex 
-transformation such as structural induction and \tactictoe(NH) can reuse them. 
-Conversely, \holyhammer(blistr) is more suited to deal with dense theories such 
-as 
-\texttt{real} or \texttt{complex} where a lot of related theorems are available 
-and 
-most proofs are usually completed by rewriting tactics.
-
-\begin{table}[]
+self-reinforcement learning. Typically this technique improves the results by 
+less than a percent.
+
+\subsection{Complexity of the proofs}
+
+\pgfplotscreateplotcyclelist{my black}{
+solid, mark repeat=100, mark phase=0, black!100\\
+dashed, mark repeat=100, mark phase=0, black!100\\
+}
+
+Estimate the difficulty of the proof by the number of tactic units.
+
+\begin{figure}[h]
 \centering           
-\setlength{\tabcolsep}{3mm}
-\begin{tabular}{@{}ccccc@{}}
-\toprule
-\phantom{ab} & {arith} & {real} & {compl} 
-& {meas} \\
-\midrule
-\tactictoe & 74.6 & 67.3 & 77.3 & 28.2\\
-\eprover & 50.8 & 67.6 & 63.4 & 10.8\\
-\midrule
-\phantom{abc}  & {proba} & {list} & {sort} & {f\_map} \\
-\midrule
-\tactictoe & 37.3 & 71.4 & 58.4 & 73.6 \\
-\eprover & 19.3 & 20.7 & 15.8 & 20.9 \\
-\bottomrule
-\end{tabular}
-\caption{\label{theories}Percentage (\%) of re-proved theorems in the theories 
-\texttt{arithmetic}, \texttt{real}, \texttt{complex}, \texttt{measure},  
-\texttt{probability}, \texttt{list}, \texttt{sorting} and \texttt{finite\_map}. 
-}
-\end{table}
-
-\subsection{Time and space complexity}
+\begin{tikzpicture}[scale=1]
+\begin{axis}[
+  legend style={anchor=north east, at={(0.9,0.9)}},
+  width=0.5*\textwidth,
+  height=0.5*\textwidth,
+  ymin=0, ymax=2000,
+  xmin=0, xmax=10,
+  xtick={},
+  ytick={},
+  cycle list name=my black]
+\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
+\addplot table[x=length, y=proofs] {data/original_proof_length};
+\end{axis}
+\end{tikzpicture}     
+\begin{tikzpicture}[scale=1]
+\begin{axis}[
+  legend style={anchor=north east, at={(0.9,0.9)}},
+  width=0.5*\textwidth,
+  height=0.5*\textwidth,
+  ymin=0, ymax=120,
+  xmin=10, xmax=50,
+  xtick={},
+  ytick={},
+  cycle list name=my black]
+\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
+\addplot table[x=length, y=proofs] {data/original_proof_length};
+\legend{\tactictoe,original}
+\end{axis}
+\end{tikzpicture}
+\caption{Distribution of the proof length in tactic unit of evaluated theorems.}
+\end{figure}
 
@@
 we can see that the slope of the curve is far for being flat in Fig~..
-
-\pgfplotscreateplotcyclelist{my black}{
-solid, mark repeat=100, mark phase=0, black!100\\
-dashed, mark repeat=100, mark phase=0, black!100\\
-}
 
@@
   width=\textwidth,
-  height=0.7*\textwidth,xmin=0, xmax=10,
-  ymin=0, ymax=4500,
+  height=0.7*\textwidth,xmin=0, xmax=60,
+  ymin=0, ymax=4800,
   xtick={},
@@
   cycle list name=my black]
-\addplot table[x=time, y=solved] {data/tt_ptime};
-\addplot table[x=time, y=solved] {data/e_ptime};
+\addplot table[x=time, y=solved] {data/tactictoe_time};
+\addplot table[x=time, y=solved] {data/eprover_time};
 \legend{\tactictoe,\eprover}
@@
 \end{figure}
+\eprover uses at least two strategies where as \tactictoe only uses one set 
+of parameters.
 
@@
   width=\textwidth,
-  height=0.7*\textwidth,xmin=0, xmax=10,
+  height=0.7*\textwidth,xmin=0, xmax=20,
   ymin=0, ymax=100,
@@
   cycle list name=my black]
-\addplot table[x=size, y=solved] {data/tt_percent};
-\addplot table[x=size, y=solved] {data/e_percent};
+\addplot table[x=oplen, y=solved] {data/tactictoe_by_oplen};
+\addplot table[x=oplen, y=solved] {data/eprover_by_oplen};
 \legend{\tactictoe,\eprover}
@@
 
-%\begin{table*}[t]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{llllll}
-%\toprule
-% ID & \multicolumn{2}{c}{nodes}   & \multicolumn{2}{c}{proof size} & 
-% \multicolumn{1}{c}{time}\\ \cmidrule(lr){2-3} \cmidrule(lr){4-5} 
-% \cmidrule(lr){6-6}
-%     & average & max & average & max & average \\
-%\midrule
-% \tactictoe & 94.66 & 421 & 3.34 & 39 & 0.66 \\
-%\bottomrule
-%\end{tabular}
-% \caption{\label{tab:stats} old search statistics}
-%\end{table*}
-
-60 seconds experiment + 
-adding theorems from the namespace and removing 
-additionally saved theorems + 
-recording proofs from Prove.
-
-Constant progress: 
-+45 percent in Septermber in 5 seconds.
-+59 percent in December in 10 seconds.
-+66.6 percent in January in 60 seconds.
-
-The progress cannot be attributed only to the time increase but to testing and 
-debugging of parameters.
+%60 seconds experiment + 
+%adding theorems from the namespace and removing 
+%additionally saved theorems + 
+%recording proofs from Prove.
+
+% 
+%The progress cannot be attributed only to the time increase but to testing and 
+%debugging of parameters.
 

### 2018-01-24 tactictoe_jv.tex
scaling distribution

@@
  ortho + theorem list & 160\\
- ortho + metis & 162\\
+ ortho + \metis & 162\\
  %ortho + metis + thmortho & 155\\
  %ortho + metis + stacortho & 160\\
- ortho + metis + theorem list & 175\\ 
+ ortho + \metis + theorem list & 175\\ 
 \bottomrule
@@
   height=0.5*\textwidth,
-  ymin=0, ymax=2000,
+  ymin=0, ymax=3000,
   xmin=0, xmax=10,
@@
 \addplot table[x=length, y=proofs] {data/original_proof_length};
-\legend{\tactictoe,original}
+\legend{scaled \tactictoe,original}
 \end{axis}
 \end{tikzpicture}
-\caption{Distribution of the proof length in tactic unit of evaluated theorems.}
+\caption{Number of original proofs with a number $x$ of tactic units. 
+         Since there is about two-third of the original proofs that have a 
+         respective \tactictoe proofs, we scale the number of 
+         \tactictoe proofs by a factor of $\frac{1}{0.662}$.. 
+\tactictoe 
+ proofs.}
 \end{figure}
-
-Analysing failed proof attempt can give us an idea of the behaviour of the 
-algorithm. In a proof search of 10 seconds, in average 500 (todo: real 
-estimation) tactics are applied.
-The productive tactics create an average of 100 (todo: real estimation) nodes. 
-This means that in this settings 1 out of 5 tactics are productive.
 

### 2018-01-24 tactictoe_jv.tex
scaling distribution

@@
 \addplot table[x=length, y=proofs] {data/original_proof_length};
-\legend{scaled \tactictoe,original}
+\legend{\tactictoe proofs,original proofs}
 \end{axis}
 \end{tikzpicture}
-\caption{Number of original proofs with a number $x$ of tactic units. 
-         Since there is about two-third of the original proofs that have a 
-         respective \tactictoe proofs, we scale the number of 
-         \tactictoe proofs by a factor of $\frac{1}{0.662}$.. 
-\tactictoe 
- proofs.}
+\caption{Number of original and \tactictoe proofs with a number $x$ of tactic 
+units. For comparing of the distribution, the number of \tactictoe proofs was 
+scaled by the inverse of \tactictoe's success rate $\frac{1}{0.662}$.}
 \end{figure}
+
+There are more in proportion \tactictoe small proofs than original human proofs.
 

### 2018-01-25 tactictoe_jv.tex
Some

@@
 %
-\vspace{100mm}\tableofcontents
-
-\vspace{100mm}
+
+\tableofcontents
+
 \todo{TG : comparison with hammers and advantages for the user and over hammers}
@@
 
+How a user works with an ITP and how this gives rise to a machine learning problem.
+
 Two ways of representing a formulas.
 
-Defintion of a tactic in \holfour. In this research, we will take the definition
+Definition of a tactic in \holfour. In this research, we will take the definition
 that a tactic takes a goal and returns a list of goals.
@@
 \subsection{Contributions}
-This project is an extension of the work described in \cite{tgckju-lpar17}. We 
-explain here the additional contributions and their effect across our system:
-
+This work is an extension of the work described in \cite{tgckju-lpar17}.
+There we proposed the idea of ... and achieved ...
+The additional contributions and their effect on our system are:
+
+\todoi{add prediction of 4 kinds of things}
 \begin{itemize}
@@
 \section{Proof Recording}\label{sec:recording}
-\todo{MN + RK: comment technical and hard-to read presentation of the unfolding
+\todoi{MN + RK: comment technical and hard-to read presentation of the unfolding
 mechanism and replace by a high level abstraction. Alternatively, make the 
@@
 original theory file containing proof scripts into an
-abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
+abstract syntax tree (AST).\todoi{reason: reading the proof as if you were a 
 human ==> better for the machine learning algorithm (learn the right thing)}
@@
 \section{Proof Search}\label{sec:proofsearch}
-\todo
+\todoi
 {
@@
 
-\section{ATP Integration}\todo{TG + JU + CK: rework and review}
+\subsection{ATP Integration}\todo{TG + JU + CK: rework and review}
 General-purpose proof automation mechanisms which combine proof translation to
@@
 
+\todoi{Example prettified proof}
+
 %\paragraph{Step by step exploration.}
@@
 
-\section{Case study: comparison with original proofs}\todo{
+\section{Comparison with Human Proofs}\todo{
   TG + RK: find describe examples.
@@
   + Orthogonalization of theorems~\cite{ckju-jsc15}. 
-  + Try harder and smarted on term predictions. \cite{latest version of sepia}. 
+  + Try harder and smarted on term predictions. \cite{latest-version-of-sepia}. 
   \cite{conjecturing}

### 2018-01-26 tactictoe_jv.tex
prediction: wip

--- 2018-01-24 tactictoe_jv.tex
@@
 This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths. 
-Experiments on the \holfour standard 
-library show that TacticToe is able to re-prove 66.0\% of 
-all theorems proven by a tactical script in the standard library.
-This is to be compared with the 34.5\% proved in the same amount of time by 
-Eprover in auto-schedule mode with 128 lemmas predicted and 
-translated by HolyHammer.
+TacticToe proves 66.2\% of 7164 theorems of the standard library where as 
+Eprover solves 34.5\%. The success rate rises to 69.0\% by combining the 
+result of both provers.
 
@@
 
-\subsection{The problem}\todo{}
-
-Two ways of representing a formulas.
-
-Defintion of a tactic in \holfour. In this research, we will take the definition
-that a tactic takes a goal and returns a list of goals.
+\subsection{The problem}\todo{Maybe should }
+
+Definitions of common concepts: goals, tactics, theorems, terms , etc
 
@@
 
-\section{Proof Recording}\label{sec:recording}
-\todo{MN + RK: comment technical and hard-to read presentation of the unfolding
-mechanism and replace by a high level abstraction. Alternatively, make the 
-current version more accessible.}
-
-Recording proofs in an LCF-style proof assistant can be done at different levels.
-In \holfour all existing approaches relied on modifying the kernel. This was used
-either to export the primitive inference steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
-or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable for our
-purpose of learning proving strategies at the intermediate tactic level. We therefore
-discuss recording proofs in an LCF-style proof assistant, with the focus on \holfour
-in this section.
-
-% Because our aim is to 
-% learn how to prove from existing examples, we record the proofs at 
-% an intermediate tactic level closest to the one presented to 
-% a user. Since human developers build almost all their proofs from tactics, 
-% we believe this is an appropriate level to apply machine learning on.
-
-%More information is contained in tactical scripts than 
-%in top-level dependencies on which was based machine-learning guidance in 
-%interactive theorem proving so far. Learning from atomic inferences would 
-%require better predictions and abstractions of rules to handle the larger 
-%search space.
-
-%\subsection{Recording Tactic Calls}
-%We first discuss parsing the proofs and identifying 
-%individual tactic calls. We next show how tactics calls are recorded together 
-%with their corresponding goals using a modified proof script.
-%Finally, the recorded data is organized as feature vectors with tactics as 
-%labels and characteristics of their associated goals as features. 
-%These feature vectors constitute the training set for our selection algorithm.
-
-Rather than to rely on the underlying programming language, we parse the
-original theory file containing proof scripts into an
-abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
-human ==> better for the machine learning algorithm (learn the right thing)}
-
- The AST is then transformed into a globalized AST
-where every top-level value is made independent from
-the context. \todo{some reason why this is needed? easy transfer between 
-theories} We 
-split the proofs into a series of tactic units and wrap them 
-with a recording function.
-By replaying the proofs produced by the globalized AST, the wrapper can
-record the globalized tactic string with its input goal as training examples in
-a database.
-The combined effect of these steps on a \holfour proof will be illustrated on Example~\ref{ex:running}.
-
-\begin{example}\label{ex:running}(Running example)
-\small
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-open boolLib Tactic Prim_rec Rewrite
-ldots
-val LIST_INDUCT_TAC = INDUCT_THEN list_INDUCT ASSUME_TAC
-ldots
-val MAP_APPEND = store_thm("MAP_APPEND",
-  ``!(f:'a->'b) l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f l2)``,
-  STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND])
-\end{lstlisting}
-\end{example}
-
-%\subsection{Minimal \sml Code Trees}
-%
-%
-%Using the standard types $string$ and $list$ and their type constructors,
-%we define a minimal code tree by the following AST:
-%\begin{align*}
-%M &=_{def} string \\
-%I &=_{def} string\ \vert\ Module(M,I)\\
-%D &=_{def} Open(M)\ \vert\  Val(I,E) \\
-%E &=_{def} Code(I\ list)\ \vert\   Let (D,E)
-%\end{align*}
-%
-%In this AST, the constructors $Let$, $Val$ and $Open$ correspond to
-%the reserved \sml tokens \texttt{let}, \texttt{val} and \texttt{open} 
-%respectively. $Module(M,I)$ refers to the \sml token \texttt{M.I} where 
-%\texttt{M} is a module (also called structure in \sml) and \texttt{I} is an 
-%identifier. $D$ is a declaration, $E$ an expression. Inside a 
-%declaration $val(I,E)$, $E$ is a definition for $I$.
-%To construct such an AST for a theory file, we rely on a custom parser. It 
-%transform a source file into a 
-%minimal code tree taking into account the scope of each constructor.
-%An additional reason to refrain from the use of the interpreter in our case, 
-%is that
-%the \polyml compiler needs to execute the code preceding an expression 
-%before parsing it, which is significantly slower than a custom parser.
-
-\subsection{Globalization}
-
-\holfour proofs are organized in theory files.
-Each theory file contains definitions and statements of theorems together with
-their proofs. Those theories and the contained proofs
-are written
-in the functional programming language \sml. To automate
-proof search strategies a user can implement programming language functions
-that correspond to more involved proofs.
-Functions that correspond to backward proofs, called tactics, are combined with 
-each other with the help of 
-tacticals (functions that takes
-tactics as arguments). Such a combination of tactics will be called a 
-\emph{proof script} when it
-proves a theorem stated in a theory.
-
-In order to apply machine learning to the existing proof scripts, we will need 
-to
-parse and modify them, which involves parsing and modification of \sml code.
-Acquiring access to the string representation of a tactic (code for the tactic) 
-allows us generating
-proof scripts that can be inspected by a human.
-As some \holfour theories are built in separate sessions, it is not possible
-to transfer \sml values directly, but instead tactic code needs to transferred 
-instead.
-To do this, the transferred information will be saved externally.
-
-%\subsubsection {Unfolding locally declared values}
-To transfer \sml values as code across theories in a conservative manner, we 
-need to 
-globalize \sml identifiers. To this end, we replace each identifier by some 
-\sml code containing all the information to obtain the value of this identifier 
-without knowing the context. The result is a standalone \sml code executable in 
-any \holfour theories which will be interpreted in the same way in the absence 
-of side effects. Two techniques are helpful to produce such code. The first one 
-is to prefix global identifiers by the module (and sub-modules) in which they 
-were declared. And the second one is to replace local identifiers 
-by their definitions. Prefixing and 
-unfolding are applied recursively in 
-the definitions.
-
-%For a more formal explanation, we define a target AST build upon the original 
-%one, where each identifer $I$ is replace by its globalized expression $E'$:
-%
-%\begin{align*}
-%M' &=_{def} string \\
-%I' &=_{def} string\ \vert\ Module(M',I')\\
-%D' & =_{def} Open(M')\ \vert\  Val (I',E') \\
-%E' &=_{def} Code(E'\ list)\ \vert\ Let (D',E') 
-%\end{align*}
-%
-%Then, we express two mutually recursive functions $f_d$ and 
-%$f_e$ that globalize the original representation. The 
-%function $f_d$ modifies declaration and $f_e$ modifies expressions. Since a 
-%\sml program is a list of declarations, we will be able transform it by 
-%iteratively applying $f_d$. The stack 
-%$\mathbb{S}$ keep track of each previous declaration: the identifier that is 
-%declared and their globalized definition. To save a declaration to 
-%the stack, we use the $push$ 
-%function. Opening a module has the effect of pushing all identifiers with 
-%their associated prefixed version to the stack. The scope of an identifier is 
-%automatically handled by construction of the AST. To find the globalized 
-%version of an identifier $i$, we use the $lookup$ function on a stack $s$, 
-%which returns the globalization associated with $i$ appearing highest in $s$. 
-%We then build $f_e$ and $f_d$ in the following way:
-%\vspace{5mm}
-%
-%\begin{math}
-%\begin{array}{lcl}
-%f_d:& \mathbb{S} \times D &\rightarrow\ \ \mathbb{S} \times D'\\
-%& (s, Open(m)) &\mapsto\ \ (push(m,s),\ Open(m)) \\
-%& (s, Val(i,e)) &\mapsto\ \ (push ((i,\hat{e}),s),\ Val(i, \hat{e})) 
-%\mbox{ where } \hat{e} = \pi_2^2 (f_e (s,e))
-%\end{array}
-%\end{math}
-%
-%\vspace{5mm}
-%
-%\begin{math}
-%\begin{array}{lcl}
-%f_e:& \mathbb{S} \times E &\rightarrow\ \ \mathbb{S} \times E'\\
-%& (s, List(l))  &\mapsto\ \ (s,\ lookup(s,l))\\
-%& (s, Let (d,e)) &\mapsto\ \ (\hat{s}, f_e (\hat{s}, e))
-%\mbox{ where } \hat{s} = \pi_1^2 (f_d (s,d))
-%\end{array}
-%\end{math}
-%
-%\vspace{5mm}
-
-%\begin{example} Example
-%\small
-%\begin{alltt}
-%val MAP_APPEND = store_thm ("MAP_APPEND", 
-%  --`!(f:'a->'b).!l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f 
-%l2)`--,
-%    STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]);
-%\end{alltt}
-%\end{example}  
-%
-%\begin{example} Running example (extracted proof)
-%\small
-%\begin{alltt}
-%STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
-%\end{alltt}
-%\end{example}
-
-%The goal of our program is to transform the original program is to associate to
-%each expr token.
-
-%\subsubsection{Structures}
-%Structures:
-%
-%Globalizing functions:
-%examples:
-%
-%let val x = 5;  end;
-%substructres
-
-%Thanks to the simplicity of the minimal AST, we are able to present our 
-%recursive unfolding algorithm in a concise and formal manner. But for 
-%practical 
-%purpose, the minimal AST is not sufficient.
-
-We present the tactical proof recording mechanism in details here because it 
-can help 
-another researcher to produce a tactical export in other interactive theorem 
-provers.
-We now enumerate common \sml constructions and explain 
-independently how we globalize them. We use a stack to keep track of previous 
-identifier declaration and their globalization.
-
-%We first start by mentioning additional 
-%details about module and value declarations.
-
-\subsection{Storing Declaration}
-
-\paragraph{Reserved tokens}
-
-\paragraph{Module}
-
-\paragraph{Global identifier}
-
-\paragraph{Local identifier}
-
-To globalize a identifier, we include 
-the code of its declaration in the globalized expression. 
-Given a declaration \texttt{fun f v = e} or \texttt{val f = e}, the globalized 
-expression for the identifier \texttt{f} is \texttt{let fun f v = e' in f 
-end} respectively \texttt{let val f = e' in f end}
-The expression \texttt{e'} is the globalized version of \texttt{e}. That is to 
-say, it is the expression \texttt{e} where each of its tokens has been replaced 
-its globalized expressions as constructed in Section~\ref{}.
-
-In the case of value, we can simplify the globalized expression further and 
-replace it by \texttt{e'}.
-In the case of a function, \texttt{f} and its argument \texttt{v} are binders. 
-So, we prevent their globalization during the construction of \texttt{e'}.
-%This approach is slightly more verbose than using an 
-%anonymous function construction. But it is prettier as 
-%we regroup declarations at the top of the proof expression before returning it
-%to the user (see Section~\ref{sec:proofdisplay}).
-
-\paragraph{Bound variables}
-A pattern is an \sml expression that 
-contains \sml 
-constructors (or exceptions in the case of \texttt{handle}) and binding 
-variables. Our algorithm can recognize cases where pattern matching appears and 
-globalize the expression appropriately.  Let \texttt{p} be a pattern 
-and 
-\texttt{e} be abd expression, the following constructions can 
-occur: \texttt{val p = e}, \texttt{case \_ of p => e}, \texttt{fun f p = e}, 
-\texttt{fn p => e}, \texttt{handle p => e}. A binding identifiers in \texttt{p}
-is a token that is not a constructor (or exception). To produce the globalized 
-version of each construction, we globalize constructors (or exception) in the 
-pattern \texttt{p} and we protect binding variables in \texttt{e} to prevent 
-their modification. When a value identifier \texttt{x} is declared in a pattern 
-\texttt{p}, the globalized expression \texttt{e'} for \texttt{x} becomes 
-\texttt{let val p = e' in x 
-end}.
-
-\paragraph{Infixity}                                                            
-In \sml, it is possible to declare infixity rules for a string $s$ that 
-is not prefixed by a module. The infixity applies only when an identifier 
-exactly match $s$.
-
-\subsection{Globalizing Expression}
-
-\paragraph{Reserved tokens}
-
-\paragraph{Modules}
-Declaring a module is made through the keyword 
-\texttt{open}. We can fetch 
-the values declared in a module \texttt{m} by first erasing all elements from 
-\texttt{PolyML.NameSpace} and then calling \texttt{open m}. Additionally, it 
-gives us the information of whether a value is a constructor or an exception. 
-This will prove to be particularly useful to handle pattern matching 
-construction correctly. During the opening of a module, nested modules can be 
-declared and push to the stack. In this way, we can globalize them by prefixing 
-them by their parent modules when they appear in the program.
-
-\paragraph{Global identifier}
-Global values and functions are prefixed by the modules and submodules where 
-they were declared. This information obtained by looking up the first 
-identifier that corresponds to the values at the top of the stack.
-
-\paragraph{Local identifier}
-Each local identifier is replaced by its globalized expression.
-
-We do not unfold local and global values of type theorem. Indeed, we do not 
-want to 
-unfold proofs into other proofs. Instead, we fetch theorems by looking for its 
-formula in the global database of 
-theorems. If it does not exists, we consider the tactic to be local and not 
-transferable to other theories.
-
-\paragraph{Bound variables}
-Binding variables are not globalized
-
-\paragraph{Infixity}
- Therefore, an identifier \texttt{x} could have been treated 
-as infix in an original expression \texttt{E} but when 
-we substitute it locally by its 
-globalized version \texttt{e'}, \texttt{e'} is now considered with prefix 
-infixity in \texttt{E} which makes \texttt{E} an invalid expression. Therefore, 
-because we 
-want to avoid global changes to \texttt{E}, we encapsulate \texttt{e'} by two 
-newly 
-defined 
-infix guards to reproduce 
-the original behaviour.
-In total there are 20 possible infix settings \texttt{x} can have in \sml.
-Those infix guards are defined to have the 
-same 
-precedence level $\texttt{P}\in[0,9]$ and same associativity 
-$\texttt{A}\in[l,r]$ (right or left) as \texttt{x}.
-Definitions for the infix guards depend on their associativity:
-
-\begin{definition}(Code of infix guards)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-fun a infixlX_open f = fn x ==> f (a,x)      fun g infixlX_close b = g b
-fun a infixrX_open g = g a      fun f infixrX_close b = fn x ==> f (x,b) 
-\end{lstlisting}
-\end{definition}
-
-The previously decribed transformation affects each identifier in a theory 
-file. However, when we 
-create the recording theory file, only proof scripts are replaced by their 
-globalized version and the rest of the code is left unchanged. The effect on 
-the proof script contained in the running example is shown in 
-Example~\ref{ex:global}.
-
-\begin{example} (Globalization)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-Tactic.STRIP_TAC 
-  infixl0_open boolLib.THEN infixl0_close
-Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC
-  infixl0_open boolLib.THEN infixl0_close
-Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
-\end{lstlisting}
-\end{example}
-
-%\sml expression delimiters ``([,]){}='' are left 
-%unchanged by the globalization process.
-%
-% pattern matching
-%By distinguishing which values are constructors in our Pattern matching in 
-%\sml 
-%is allowed inside a value 
-%
-%
-%infix operators
-%Our goal is to create a globalized version of each generalizable token.
-%
-%In this description we are omitting handling of "fun", "case of | =>", "fn =>"
-%The only thing to consider is that var appearing as arguments of a function.
-%
-%infix constructor: 
-%  Infix constructors cannot be replaced by their globalized version without 
-%  confusing the parser.
-%
-%  
-%
-%  We can globalize an infix operartor locally in the following way;
-%
-%Original code: some\_code1 infix some\_code2
-%
-%the infix code does not carry the information on how to parse it.
-%
-%            infix\_open (infix\_code) infix\_close.
-%
-%where infix\_open and infix\_closed are declared to be the same arity with the 
-%same associactivity rule as infix.
-%
-%Depending if the operator is left associative or right associative we give two 
-%different definitions for infix\_open and infix\_close.
-%
-%
-%
-%constructions with bounded variables: fun (case of) (
-%
-%proof scripts (arguments of prove, store\_thm, maybe\_thm)
-%
-%watch list:
-%
-%stateful tactics
-
-\subsection{Recording tactic calls}
-
-We now restrict the definition of a proof script to being the third 
-argument of a call to \texttt{store\_thm}. And we precisely define the 
-modification applied to the theory file. A modified theory file is identical to 
-the original theory file except every proof scripts \texttt{p} is
-replaced by \texttt{Split "p'"} where \texttt{Split} is the splitting function 
-to be defined and \texttt{"p'"} is the string representation of the globalized 
-version of the proof script 
-\texttt{p}. 
-.
-\begin{example}(Proof script transformation)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-original file: val id = store_thm (name, term, p)
-modified file: val id = store_thm (name, term, Split "p'")
-\end{lstlisting}
-\end{example}
-
-The intended effect of the splitting function is to detect and wrap \sml 
-expressions of type tactics. Its aim is to is encapsulate each tactic 
-unit $t$ within a recording function $R$. This function $R$ is designed to 
-observe what goals $t$ receives at run time. Since there can be maybe nested 
-level of tactics, (i.e 
-tactics that contain tactics as subexpressions), we choose one that is
-close to what a human understand to be a tactic unit. 
-Most tacticals are infix, and we do not want to consider combination of 
-tactics to be a tactic unit. For \texttt{Split}, a tactic unit is a 
-subterm of 
-type tactics that do not contain a infix operator as its top symbol. Thanks to
-globalized infixity, we just have to verify that the expression 
-is of type tactic and that an infix guard is not a top symbol of the expression.
-To determine that, we run the \polyml compiler on \texttt{"p'"} to give us a 
-detailed code tree of the expression. Each tactic unit is then substituted 
-by a wrapped version to produce the final proof script.
-
-\begin{example} (Wrap)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-R "Tactic.STRIP_TAC"
-  infixl0_open boolLib.THEN infixl0_close
-R "Prim_rec.INDUCT_THEN (DB.fetch \"list\" \"INDUCT\") Tactic.ASSUME_TAC"
-  infixl0_open boolLib.THEN infixl0_close
-R "Rewrite.ASM_REWRITE_TAC 
-     [DB.fetch \"list\" \"MAP\", DB.fetch \"list\" \"APPEND\"]"
-\end{lstlisting}
-\end{example}
-
-During its execution, the recording function \texttt{R} (as coded in 
-Definition~\ref{ex:record}) 
-saves the tactic 
-string and the goal, verifies that the tactic string as the same effect as the 
-original tactic and apply it to continue the proof. To perform this task, we 
-convert the tactic string back into a tactic via the function 
-\texttt{read\_tactic}. This function will be used whenever we want to execute a 
-tactic string. So in the next sections, we will not make a difference between a 
-tactic and its code and call both of them tactics.
-
-\begin{example}\label{ex:record} (Code of the recording function)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-fun R tactic_string goal = 
-  (save (tactic_string, goal); tactic_of_sml tactic_string goal)
-\end{lstlisting}
-\end{example}
-
-\subsection{Storing Training Data}
-We describe here the format preferred to store and load the 
-data recorded by the function \texttt{save}. 
-During saving, we add fields to the input pair $(tactic,goal)$ to record 
-precomputed information in order to make the prediction faster. The final 
-feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
-The time $time$ taken by $tactic$ is not currently used by our 
-predictor. The list of goals $goal list$ is the output of $tactic$ on 
-$goal$. And the features $features$ of the goal $goal$ are extracted and hashed 
-as described in Section~\ref{sec:features}.
-
-\begin{remark}
-To simplify the description of our machine learning algorithms in future 
-sections, we consider a tactic feature vector to be a pair $(tactic,goal)$ and 
-only 
-explicitly rely on the 
-other fields when necessary.
-\end{remark}
-
-We now discuss the type of file we want to store our feature vectors on and
-how we plan to write and read \sml values. We start by taking 
-inspiration from an existing implementation in \holfour.
-
-To avoid the computational cost of re-proving every theorems at the begining of 
-each session, theorem values in \holfour are stored.
-They are encoded them in to a string by calling \texttt{Term.write\_raw}.
-Conversely, the function \texttt{Term.read\_raw} transforms the string back 
-into the original term.
-To have theorems binded to sml identifiers, it is necessary to create a 
-\sml file where those binders are declared. And for simplicity sake, all theory 
-data was fed into this \sml file. However, it is
-more efficient to store theory data into a text file and make a custom 
-parser which can be run faster than the \polyml compiler.
-Thanks to this insight, we have 
-modified the way in which \holfour theories are saved to include only \sml 
-bindings in a \sml file. All other theory data is now stored separately 
-into a text file. Michael Norrish and Ramana Kumar contributed to integrated 
-these changes in the current distribution of \holfour. After the changes, 
-\holfour theories load three times faster in average.
-
-Our \tactictoe data management system do not need to create new \sml
-bindings for the user. So we can store all \sml values into a text file which 
-can be loaded rapidly.
-Storing time, tactic string and features is easy since they are respectively 
-floats, strings and integers. And because a goal is composed of terms, we can 
-re-use the \holfour theory data reading and writing functions 
-for the two other fields.
-
-%The previous step demonstrated how to extract training data from proofs. We 
-%now 
-%need to store this data on disk so that this training data is re-usable in 
-%other theories. We gathered information about a global string representation 
-%of 
-%a tactic, the time taken to apply this tactic and the input goal and output 
-%list of goals. For printing the tactic string and the time, we just make use 
-%of 
-%printing functions for those types  and 
-%\texttt{Real.toString}. 
-%However for the terms that forms the goals, it is not advisable to use the 
-%function 
-%\texttt{term\_to\_string} as it may not contain all the information necessary 
-%to 
-%reconstruct the same term. Instead we re-use the functions for storing terms 
-%of 
-%theorems in theories \texttt{Term.write\_raw} and and \texttt{Term.read\_raw}, 
-%constructing sharing tables for types and terms.
-%Contrary to \holfour theories that are saved into a \sml file, our data is 
-%stored as text files. This allow us a much faster loading time, since we are 
-%never calling the PolyML compiler. 
-%Already exists in \holfour for recording theorems.
-%Mention better export for HOL4.
-%
-%Side effects:
-%By minimizing the size of the sml files, we manage to multiply by 3 the speed 
-%at which theories are loaded in \holfour. 
-%Mention Ramana Kumar and Michael Norrish.
-
-Finally, we mention some limitations that comes with 
-the need to record tactics as a string compared to being able to store a \sml 
-value. Access to the name of the tactic enhances machine 
-learning ability and user interaction. Yet, because some tactics 
-depends on the state of \holfour, for example the state of the global 
-simpset, a tactic is re-parsed before each proof attempt to be up-to-date.
-As a result, a tactic may have different effect on goals than suggested by the 
-database which could compromise the prediction accuracy. In practice, this 
-should rarely happen. Indeed, most tactics are stateless and an update on a 
-stateful tactic is usually improving it.
-
-%By recording also the sml value it would also possible to eliminate those 
-%drawbacks as long as you stay in the same session. This is typically the case 
-%for the \hollight theorem prover but not in \holfour where every theory is 
-%build in a different session loading previous theorems from disks.
-
 \section{Prediction}\label{s:prediction}\todo{TG: complete term theorem and 
@@
 The learning-based selection of relevant lemmas significantly improves the 
-automation for hammers~\cite{BlanchetteGKKU16}. Therefore we propose to adapt 
-one of the strongest hammer lemma selection 
+automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we adapt 
+one of the strongest lemma selection 
 methods: the modified distance-weighted \emph{$k$ nearest-neighbour} 
-classifier~\cite{ckju-pxtp13,DudaniS76}.
-In \tactictoe, we predict four different objects: theorems, terms, tactics and 
-list of goals. We derive from tactic selection a prior policy for 
+classifier~\cite{ckju-pxtp13,DudaniS76} to predict four different objects: 
+terms, goals, theorems and list of goals.
+
+Before going into the details of the machine learning algorithm, we introduce
+specificities associated with the prediction of each object: the data from 
+which the feature vectors are build and the different use cases for the 
+prediction.
+
+We derive from tactic selection a prior policy for 
 our search algorithm. Lists of goals (products of tactics) evaluate the success 
@@
 \begin{itemize}
-\item tactic:\\
+\item term:\\
+Predict the closest term in a tactic that appears in the goal.
+
+\item :\\
+Predict the theorem the closest to the current goal. Theorem dependencies are 
+added during preselection.
+
+\item goal:\\
 Given a goal $g$, the classifier selects a list of previously solved goals $l$
@@
 estimate the similarity between goals with mathematically relevant features.
-\item term:\\
-Predict the closest term in a tactic that appears in the goal.
-
-\item theorem:\\
-Predict the theorem the closest to the current goal. Theorem dependencies are 
-added during preselection.
 
@@
 
-\subsection{Features for Terms and Goals}\label{sec:features}
-
-We start by extracting the syntactic features that have been successfully used 
-in premise selection from terms:
+\subsection{Feature Extraction}\label{sec:features}
+The four objects considered are different representation of mathematical 
+formulas. Therefore, we start by describing which features we extract from the 
+simplest representation terms. From there, we extend the extraction mechanism 
+to goals theorems and lists of goals. Duplicated features are always removed 
+so each object has an associated set of features.
+
+Here are the different kind of feature we extract from terms:
 \begin{itemize}
@@
 \item subterms with all variables replaced by a single place holder $V$.
-\item names of the variables
+\item names of the variables.
 \end{itemize}
 
-We found that the names of variables present in a term are particularly 
-important for predicting tactics such as 
-case splitting on a variable 
-(\texttt{Cases\_on var}) or instantiation of a variable 
-(\texttt{SPEC\_TAC var term}).
-
-When a formula is represented as a goal, we also add a tag to each 
-feature to remember whether it was extracted from the assumptions or the 
-conclusion of a goal.
+Since goals and theorems are represented by a couple of a list of terms 
+(assumptions) and a term (conclusion). We compute their features by extracting 
+the features of all terms describing them. Additionally, we distinguish between 
+features of the assumptions and features of the conclusion by adding a 
+different tag to elements of each set. For a list of goals, we simply construct 
+the union of the features of each goal 
+in the list.
+
+%We found that the names of variables present in a term are particularly 
+%important for predicting tactics such as 
+%case splitting on a variable 
+%(\texttt{Cases\_on var}) or instantiation of a variable 
+%(\texttt{SPEC\_TAC var term}). 
 %We additionally extract the following features:
@@
 
+Width did include failing tactics in this experiments. $c_{exploration}$ = 1.
+$E_{none}$.
+
 \begin{table}[ht]
@@
 \small
-\begin{tabular}{llllll}
+\begin{tabular}{ll}
 \toprule
- $c_{policy}$ & Width & Ortho & Eval & $c_{exploration}$ & Solved \\
+ $c_{policy}$ & Solved \\
 \midrule
- 2/3 & All & No & $E_{none}$ & 1 & 120\\
- 1/2 & All & No & $E_{none}$ & 1 & 125\\
-% 1/2 & All & Yes & $E_{none}$ & 1 & 130 \\
- 1/3 & All & No & $E_{none}$ & 1 & 117\\
+ $0.66$ & 120\\
+ $0.5$ & 125\\
+ $0.33$ & 117\\
 % 1/2 & Productive & No & $E_{none}$ & 1 & 120\\
@@
 \end{tabular}
-\caption{\label{tab:cost_param} on a local machine}
+\caption{\label{tab:cost_param} 
+%on a local machine
+}
 \end{table}
@@
 %function.
+
+This experiments are performed with orthogonalization on and policy coefficient 
+of $0.5$.
 
@@
 \small
-\begin{tabular}{llllll}
+\begin{tabular}{lll}
 \toprule
-  $c_{policy}$ & Width & Ortho & Eval & $c_{exploration}$ & Solved \\
+ Evaluation & $c_{exploration}$ & Solved \\
 \midrule
- 1/2 & All & No & $E_{none}$ & 1 & 122\\
- 1/2 & All & Yes & $E_{none}$ & 1 & 128\\
- 1/2 & Productive & Yes & $E_{none}$ & 1 & 128\\
- 1/2 & Productive & Yes & $E_{productive}$ & 1 & 134\\ 
- &  & &  & 4 & 130\\
- 1/2 & Productive & Yes & $E_1$ & 1 & 127\\ 
- & & & & 4 & 126\\ 
- 1/2 & Productive & Yes & $E_{10}$ & 1 & 136\\ 
- & & & & 4 & 137\\ 
+% 1/2 & All & No & $E_{none}$ & 1 & 122\\
+% 1/2 & All & Yes & $E_{none}$ & 1 & 128\\
+ $E_{none}$ & 1 & 128\\
+ $E_{productive}$ & 1 & 134\\ 
+ & 4 & 130\\
+ $E_1$ & 1 & 127\\ 
+ & 4 & 126\\ 
+ $E_{10}$ & 1 & 136\\ 
+ & 4 & 137\\ 
 \bottomrule
 \end{tabular}
-\caption{\label{tab:learning} on the server}
+\caption{\label{tab:learning} 
+%on the server
+}
 \end{table}
@@
 its timeout or that the human proof is too deep to be rediscovered (more than 
-10 steps long).
+10 steps long).In all those experiments orthogonalization is performed with 
+radius 20.
 
@@
 Here, we test the benefit of different kind of tactic abstraction.
-In all those experiments orthogonalization is performed with radius 20 but not 
-orthogonalization with respect to metis.
-
 All experiments in this table are using orthogonalization.
+
 \begin{table}[t]
@@
 
-\subsection{Slight improvements}
+\subsection{Slight improvements}\todo{maybe remove this subsection}
 The final results could be improved by a few percent by including the 
 following additional techniques that are already available.
-
-\paragraph{Asynchronous calls to Eprover inside TacticToe}
 The technique already discussed in Section~ would most likely improve the 
-results by 2-3 percent.
-
-\paragraph{Self-learning}
+results by 2-3 percent. Reducing the exploration coefficient from 4 to 2 would 
+also probably give a 3 percent increase in the number of proven theorems as 
+shown by a 1 in 10 evaluation on the full library.
+
 Until now, all our feature vectors were learned form human proofs. 
@@
 discovered proofs than from the human proof 
-scripts~\cite{DBLP:conf/cade/Urban07}. It is also the first step toward 
-self-reinforcement learning. Typically this technique improves the results by 
+scripts~\cite{DBLP:conf/cade/Urban07}. 
+Self-learning improves the results by 
 less than a percent.
+
+%It is also the first step toward self-reinforcement learning.
 
@@
 \caption{Number of original and \tactictoe proofs with a number $x$ of tactic 
-units. For comparing of the distribution, the number of \tactictoe proofs was 
+units. For easier comparison of the distributions, the number of \tactictoe 
+proofs was 
 scaled by the inverse of \tactictoe's success rate $\frac{1}{0.662}$.}
@@
 \caption{Percentage of problem solved with respect to the length of the 
-original 
-proof}
+original proof until length 20.}
 \end{figure}
@@
 %debugging of parameters.
+
+\section{Proof Recording}\label{sec:recording}
+\todo{MN + RK: comment technical and hard-to read presentation of the unfolding
+mechanism and replace by a high level abstraction. Alternatively, make the 
+current version more accessible.}
+
+Recording proofs in an LCF-style proof assistant can be done at different 
+levels.
+In \holfour all existing approaches relied on modifying the kernel. This was 
+used
+either to export the primitive inference 
+steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
+or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable 
+for our
+purpose of learning proving strategies at the intermediate tactic level. We 
+therefore
+discuss recording proofs in an LCF-style proof assistant, with the focus on 
+\holfour
+in this section.
+
+% Because our aim is to 
+% learn how to prove from existing examples, we record the proofs at 
+% an intermediate tactic level closest to the one presented to 
+% a user. Since human developers build almost all their proofs from tactics, 
+% we believe this is an appropriate level to apply machine learning on.
+
+%More information is contained in tactical scripts than 
+%in top-level dependencies on which was based machine-learning guidance in 
+%interactive theorem proving so far. Learning from atomic inferences would 
+%require better predictions and abstractions of rules to handle the larger 
+%search space.
+
+%\subsection{Recording Tactic Calls}
+%We first discuss parsing the proofs and identifying 
+%individual tactic calls. We next show how tactics calls are recorded together 
+%with their corresponding goals using a modified proof script.
+%Finally, the recorded data is organized as feature vectors with tactics as 
+%labels and characteristics of their associated goals as features. 
+%These feature vectors constitute the training set for our selection algorithm.
+
+Rather than to rely on the underlying programming language, we parse the
+original theory file containing proof scripts into an
+abstract syntax tree (AST).\todo{reason: reading the proof as if you were a 
+human ==> better for the machine learning algorithm (learn the right thing)}
+
+ The AST is then transformed into a globalized AST
+where every top-level value is made independent from
+the context. \todo{some reason why this is needed? easy transfer between 
+theories} We 
+split the proofs into a series of tactic units and wrap them 
+with a recording function.
+By replaying the proofs produced by the globalized AST, the wrapper can
+record the globalized tactic string with its input goal as training examples in
+a database.
+The combined effect of these steps on a \holfour proof will be illustrated on 
+Example~\ref{ex:running}.
+
+\begin{example}\label{ex:running}(Running example)
+\small
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+open boolLib Tactic Prim_rec Rewrite
+ldots
+val LIST_INDUCT_TAC = INDUCT_THEN list_INDUCT ASSUME_TAC
+ldots
+val MAP_APPEND = store_thm("MAP_APPEND",
+  ``!(f:'a->'b) l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f l2)``,
+  STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND])
+\end{lstlisting}
+\end{example}
+
+%\subsection{Minimal \sml Code Trees}
+%
+%
+%Using the standard types $string$ and $list$ and their type constructors,
+%we define a minimal code tree by the following AST:
+%\begin{align*}
+%M &=_{def} string \\
+%I &=_{def} string\ \vert\ Module(M,I)\\
+%D &=_{def} Open(M)\ \vert\  Val(I,E) \\
+%E &=_{def} Code(I\ list)\ \vert\   Let (D,E)
+%\end{align*}
+%
+%In this AST, the constructors $Let$, $Val$ and $Open$ correspond to
+%the reserved \sml tokens \texttt{let}, \texttt{val} and \texttt{open} 
+%respectively. $Module(M,I)$ refers to the \sml token \texttt{M.I} where 
+%\texttt{M} is a module (also called structure in \sml) and \texttt{I} is an 
+%identifier. $D$ is a declaration, $E$ an expression. Inside a 
+%declaration $val(I,E)$, $E$ is a definition for $I$.
+%To construct such an AST for a theory file, we rely on a custom parser. It 
+%transform a source file into a 
+%minimal code tree taking into account the scope of each constructor.
+%An additional reason to refrain from the use of the interpreter in our case, 
+%is that
+%the \polyml compiler needs to execute the code preceding an expression 
+%before parsing it, which is significantly slower than a custom parser.
+
+\subsection{Globalization}
+
+\holfour proofs are organized in theory files.
+Each theory file contains definitions and statements of theorems together with
+their proofs. Those theories and the contained proofs
+are written
+in the functional programming language \sml. To automate
+proof search strategies a user can implement programming language functions
+that correspond to more involved proofs.
+Functions that correspond to backward proofs, called tactics, are combined with 
+each other with the help of 
+tacticals (functions that takes
+tactics as arguments). Such a combination of tactics will be called a 
+\emph{proof script} when it
+proves a theorem stated in a theory.
+
+In order to apply machine learning to the existing proof scripts, we will need 
+to
+parse and modify them, which involves parsing and modification of \sml code.
+Acquiring access to the string representation of a tactic (code for the tactic) 
+allows us generating
+proof scripts that can be inspected by a human.
+As some \holfour theories are built in separate sessions, it is not possible
+to transfer \sml values directly, but instead tactic code needs to transferred 
+instead.
+To do this, the transferred information will be saved externally.
+
+%\subsubsection {Unfolding locally declared values}
+To transfer \sml values as code across theories in a conservative manner, we 
+need to 
+globalize \sml identifiers. To this end, we replace each identifier by some 
+\sml code containing all the information to obtain the value of this identifier 
+without knowing the context. The result is a standalone \sml code executable in 
+any \holfour theories which will be interpreted in the same way in the absence 
+of side effects. Two techniques are helpful to produce such code. The first one 
+is to prefix global identifiers by the module (and sub-modules) in which they 
+were declared. And the second one is to replace local identifiers 
+by their definitions. Prefixing and 
+unfolding are applied recursively in 
+the definitions.
+
+%For a more formal explanation, we define a target AST build upon the original 
+%one, where each identifer $I$ is replace by its globalized expression $E'$:
+%
+%\begin{align*}
+%M' &=_{def} string \\
+%I' &=_{def} string\ \vert\ Module(M',I')\\
+%D' & =_{def} Open(M')\ \vert\  Val (I',E') \\
+%E' &=_{def} Code(E'\ list)\ \vert\ Let (D',E') 
+%\end{align*}
+%
+%Then, we express two mutually recursive functions $f_d$ and 
+%$f_e$ that globalize the original representation. The 
+%function $f_d$ modifies declaration and $f_e$ modifies expressions. Since a 
+%\sml program is a list of declarations, we will be able transform it by 
+%iteratively applying $f_d$. The stack 
+%$\mathbb{S}$ keep track of each previous declaration: the identifier that is 
+%declared and their globalized definition. To save a declaration to 
+%the stack, we use the $push$ 
+%function. Opening a module has the effect of pushing all identifiers with 
+%their associated prefixed version to the stack. The scope of an identifier is 
+%automatically handled by construction of the AST. To find the globalized 
+%version of an identifier $i$, we use the $lookup$ function on a stack $s$, 
+%which returns the globalization associated with $i$ appearing highest in $s$. 
+%We then build $f_e$ and $f_d$ in the following way:
+%\vspace{5mm}
+%
+%\begin{math}
+%\begin{array}{lcl}
+%f_d:& \mathbb{S} \times D &\rightarrow\ \ \mathbb{S} \times D'\\
+%& (s, Open(m)) &\mapsto\ \ (push(m,s),\ Open(m)) \\
+%& (s, Val(i,e)) &\mapsto\ \ (push ((i,\hat{e}),s),\ Val(i, \hat{e})) 
+%\mbox{ where } \hat{e} = \pi_2^2 (f_e (s,e))
+%\end{array}
+%\end{math}
+%
+%\vspace{5mm}
+%
+%\begin{math}
+%\begin{array}{lcl}
+%f_e:& \mathbb{S} \times E &\rightarrow\ \ \mathbb{S} \times E'\\
+%& (s, List(l))  &\mapsto\ \ (s,\ lookup(s,l))\\
+%& (s, Let (d,e)) &\mapsto\ \ (\hat{s}, f_e (\hat{s}, e))
+%\mbox{ where } \hat{s} = \pi_1^2 (f_d (s,d))
+%\end{array}
+%\end{math}
+%
+%\vspace{5mm}
+
+%\begin{example} Example
+%\small
+%\begin{alltt}
+%val MAP_APPEND = store_thm ("MAP_APPEND", 
+%  --`!(f:'a->'b).!l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f 
+%l2)`--,
+%    STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]);
+%\end{alltt}
+%\end{example}  
+%
+%\begin{example} Running example (extracted proof)
+%\small
+%\begin{alltt}
+%STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
+%\end{alltt}
+%\end{example}
+
+%The goal of our program is to transform the original program is to associate to
+%each expr token.
+
+%\subsubsection{Structures}
+%Structures:
+%
+%Globalizing functions:
+%examples:
+%
+%let val x = 5;  end;
+%substructres
+
+%Thanks to the simplicity of the minimal AST, we are able to present our 
+%recursive unfolding algorithm in a concise and formal manner. But for 
+%practical 
+%purpose, the minimal AST is not sufficient.
+
+We present the tactical proof recording mechanism in details here because it 
+can help 
+another researcher to produce a tactical export in other interactive theorem 
+provers.
+We now enumerate common \sml constructions and explain 
+independently how we globalize them. We use a stack to keep track of previous 
+identifier declaration and their globalization.
+
+%We first start by mentioning additional 
+%details about module and value declarations.
+
+\subsection{Storing Declaration}
+
+\paragraph{Reserved tokens}
+
+\paragraph{Module}
+
+\paragraph{Global identifier}
+
+\paragraph{Local identifier}
+
+To globalize a identifier, we include 
+the code of its declaration in the globalized expression. 
+Given a declaration \texttt{fun f v = e} or \texttt{val f = e}, the globalized 
+expression for the identifier \texttt{f} is \texttt{let fun f v = e' in f 
+end} respectively \texttt{let val f = e' in f end}
+The expression \texttt{e'} is the globalized version of \texttt{e}. That is to 
+say, it is the expression \texttt{e} where each of its tokens has been replaced 
+its globalized expressions as constructed in Section~\ref{}.
+
+In the case of value, we can simplify the globalized expression further and 
+replace it by \texttt{e'}.
+In the case of a function, \texttt{f} and its argument \texttt{v} are binders. 
+So, we prevent their globalization during the construction of \texttt{e'}.
+%This approach is slightly more verbose than using an 
+%anonymous function construction. But it is prettier as 
+%we regroup declarations at the top of the proof expression before returning it
+%to the user (see Section~\ref{sec:proofdisplay}).
+
+\paragraph{Bound variables}
+A pattern is an \sml expression that 
+contains \sml 
+constructors (or exceptions in the case of \texttt{handle}) and binding 
+variables. Our algorithm can recognize cases where pattern matching appears and 
+globalize the expression appropriately.  Let \texttt{p} be a pattern 
+and 
+\texttt{e} be abd expression, the following constructions can 
+occur: \texttt{val p = e}, \texttt{case \_ of p => e}, \texttt{fun f p = e}, 
+\texttt{fn p => e}, \texttt{handle p => e}. A binding identifiers in \texttt{p}
+is a token that is not a constructor (or exception). To produce the globalized 
+version of each construction, we globalize constructors (or exception) in the 
+pattern \texttt{p} and we protect binding variables in \texttt{e} to prevent 
+their modification. When a value identifier \texttt{x} is declared in a pattern 
+\texttt{p}, the globalized expression \texttt{e'} for \texttt{x} becomes 
+\texttt{let val p = e' in x 
+end}.
+
+\paragraph{Infixity}                                                            
+In \sml, it is possible to declare infixity rules for a string $s$ that 
+is not prefixed by a module. The infixity applies only when an identifier 
+exactly match $s$.
+
+\subsection{Globalizing Expression}
+
+\paragraph{Reserved tokens}
+
+\paragraph{Modules}
+Declaring a module is made through the keyword 
+\texttt{open}. We can fetch 
+the values declared in a module \texttt{m} by first erasing all elements from 
+\texttt{PolyML.NameSpace} and then calling \texttt{open m}. Additionally, it 
+gives us the information of whether a value is a constructor or an exception. 
+This will prove to be particularly useful to handle pattern matching 
+construction correctly. During the opening of a module, nested modules can be 
+declared and push to the stack. In this way, we can globalize them by prefixing 
+them by their parent modules when they appear in the program.
+
+\paragraph{Global identifier}
+Global values and functions are prefixed by the modules and submodules where 
+they were declared. This information obtained by looking up the first 
+identifier that corresponds to the values at the top of the stack.
+
+\paragraph{Local identifier}
+Each local identifier is replaced by its globalized expression.
+
+We do not unfold local and global values of type theorem. Indeed, we do not 
+want to 
+unfold proofs into other proofs. Instead, we fetch theorems by looking for its 
+formula in the global database of 
+theorems. If it does not exists, we consider the tactic to be local and not 
+transferable to other theories.
+
+\paragraph{Bound variables}
+Binding variables are not globalized
+
+\paragraph{Infixity}
+ Therefore, an identifier \texttt{x} could have been treated 
+as infix in an original expression \texttt{E} but when 
+we substitute it locally by its 
+globalized version \texttt{e'}, \texttt{e'} is now considered with prefix 
+infixity in \texttt{E} which makes \texttt{E} an invalid expression. Therefore, 
+because we 
+want to avoid global changes to \texttt{E}, we encapsulate \texttt{e'} by two 
+newly 
+defined 
+infix guards to reproduce 
+the original behaviour.
+In total there are 20 possible infix settings \texttt{x} can have in \sml.
+Those infix guards are defined to have the 
+same 
+precedence level $\texttt{P}\in[0,9]$ and same associativity 
+$\texttt{A}\in[l,r]$ (right or left) as \texttt{x}.
+Definitions for the infix guards depend on their associativity:
+
+\begin{definition}(Code of infix guards)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+fun a infixlX_open f = fn x ==> f (a,x)      fun g infixlX_close b = g b
+fun a infixrX_open g = g a      fun f infixrX_close b = fn x ==> f (x,b) 
+\end{lstlisting}
+\end{definition}
+
+The previously decribed transformation affects each identifier in a theory 
+file. However, when we 
+create the recording theory file, only proof scripts are replaced by their 
+globalized version and the rest of the code is left unchanged. The effect on 
+the proof script contained in the running example is shown in 
+Example~\ref{ex:global}.
+
+\begin{example} (Globalization)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+Tactic.STRIP_TAC 
+  infixl0_open boolLib.THEN infixl0_close
+Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC
+  infixl0_open boolLib.THEN infixl0_close
+Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
+\end{lstlisting}
+\end{example}
+
+%\sml expression delimiters ``([,]){}='' are left 
+%unchanged by the globalization process.
+%
+% pattern matching
+%By distinguishing which values are constructors in our Pattern matching in 
+%\sml 
+%is allowed inside a value 
+%
+%
+%infix operators
+%Our goal is to create a globalized version of each generalizable token.
+%
+%In this description we are omitting handling of "fun", "case of | =>", "fn =>"
+%The only thing to consider is that var appearing as arguments of a function.
+%
+%infix constructor: 
+%  Infix constructors cannot be replaced by their globalized version without 
+%  confusing the parser.
+%
+%  
+%
+%  We can globalize an infix operartor locally in the following way;
+%
+%Original code: some\_code1 infix some\_code2
+%
+%the infix code does not carry the information on how to parse it.
+%
+%            infix\_open (infix\_code) infix\_close.
+%
+%where infix\_open and infix\_closed are declared to be the same arity with the 
+%same associactivity rule as infix.
+%
+%Depending if the operator is left associative or right associative we give two 
+%different definitions for infix\_open and infix\_close.
+%
+%
+%
+%constructions with bounded variables: fun (case of) (
+%
+%proof scripts (arguments of prove, store\_thm, maybe\_thm)
+%
+%watch list:
+%
+%stateful tactics
+
+\subsection{Recording tactic calls}
+
+We now restrict the definition of a proof script to being the third 
+argument of a call to \texttt{store\_thm}. And we precisely define the 
+modification applied to the theory file. A modified theory file is identical to 
+the original theory file except every proof scripts \texttt{p} is
+replaced by \texttt{Split "p'"} where \texttt{Split} is the splitting function 
+to be defined and \texttt{"p'"} is the string representation of the globalized 
+version of the proof script 
+\texttt{p}. 
+.
+\begin{example}(Proof script transformation)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+original file: val id = store_thm (name, term, p)
+modified file: val id = store_thm (name, term, Split "p'")
+\end{lstlisting}
+\end{example}
+
+The intended effect of the splitting function is to detect and wrap \sml 
+expressions of type tactics. Its aim is to is encapsulate each tactic 
+unit $t$ within a recording function $R$. This function $R$ is designed to 
+observe what goals $t$ receives at run time. Since there can be maybe nested 
+level of tactics, (i.e 
+tactics that contain tactics as subexpressions), we choose one that is
+close to what a human understand to be a tactic unit. 
+Most tacticals are infix, and we do not want to consider combination of 
+tactics to be a tactic unit. For \texttt{Split}, a tactic unit is a 
+subterm of 
+type tactics that do not contain a infix operator as its top symbol. Thanks to
+globalized infixity, we just have to verify that the expression 
+is of type tactic and that an infix guard is not a top symbol of the expression.
+To determine that, we run the \polyml compiler on \texttt{"p'"} to give us a 
+detailed code tree of the expression. Each tactic unit is then substituted 
+by a wrapped version to produce the final proof script.
+
+\begin{example} (Wrap)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+R "Tactic.STRIP_TAC"
+  infixl0_open boolLib.THEN infixl0_close
+R "Prim_rec.INDUCT_THEN (DB.fetch \"list\" \"INDUCT\") Tactic.ASSUME_TAC"
+  infixl0_open boolLib.THEN infixl0_close
+R "Rewrite.ASM_REWRITE_TAC 
+     [DB.fetch \"list\" \"MAP\", DB.fetch \"list\" \"APPEND\"]"
+\end{lstlisting}
+\end{example}
+
+During its execution, the recording function \texttt{R} (as coded in 
+Definition~\ref{ex:record}) 
+saves the tactic 
+string and the goal, verifies that the tactic string as the same effect as the 
+original tactic and apply it to continue the proof. To perform this task, we 
+convert the tactic string back into a tactic via the function 
+\texttt{read\_tactic}. This function will be used whenever we want to execute a 
+tactic string. So in the next sections, we will not make a difference between a 
+tactic and its code and call both of them tactics.
+
+\begin{example}\label{ex:record} (Code of the recording function)
+\begin{lstlisting}[language=SMLSmall,frame=tb]
+fun R tactic_string goal = 
+  (save (tactic_string, goal); tactic_of_sml tactic_string goal)
+\end{lstlisting}
+\end{example}
+
+\subsection{Storing Training Data}
+We describe here the format preferred to store and load the 
+data recorded by the function \texttt{save}. 
+During saving, we add fields to the input pair $(tactic,goal)$ to record 
+precomputed information in order to make the prediction faster. The final 
+feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
+The time $time$ taken by $tactic$ is not currently used by our 
+predictor. The list of goals $goal list$ is the output of $tactic$ on 
+$goal$. And the features $features$ of the goal $goal$ are extracted and hashed 
+as described in Section~\ref{sec:features}.
+
+\begin{remark}
+To simplify the description of our machine learning algorithms in future 
+sections, we consider a tactic feature vector to be a pair $(tactic,goal)$ and 
+only 
+explicitly rely on the 
+other fields when necessary.
+\end{remark}
+
+We now discuss the type of file we want to store our feature vectors on and
+how we plan to write and read \sml values. We start by taking 
+inspiration from an existing implementation in \holfour.
+
+To avoid the computational cost of re-proving every theorems at the begining of 
+each session, theorem values in \holfour are stored.
+They are encoded them in to a string by calling \texttt{Term.write\_raw}.
+Conversely, the function \texttt{Term.read\_raw} transforms the string back 
+into the original term.
+To have theorems binded to sml identifiers, it is necessary to create a 
+\sml file where those binders are declared. And for simplicity sake, all theory 
+data was fed into this \sml file. However, it is
+more efficient to store theory data into a text file and make a custom 
+parser which can be run faster than the \polyml compiler.
+Thanks to this insight, we have 
+modified the way in which \holfour theories are saved to include only \sml 
+bindings in a \sml file. All other theory data is now stored separately 
+into a text file. Michael Norrish and Ramana Kumar contributed to integrated 
+these changes in the current distribution of \holfour. After the changes, 
+\holfour theories load three times faster in average.
+
+Our \tactictoe data management system do not need to create new \sml
+bindings for the user. So we can store all \sml values into a text file which 
+can be loaded rapidly.
+Storing time, tactic string and features is easy since they are respectively 
+floats, strings and integers. And because a goal is composed of terms, we can 
+re-use the \holfour theory data reading and writing functions 
+for the two other fields.
+
+%The previous step demonstrated how to extract training data from proofs. We 
+%now 
+%need to store this data on disk so that this training data is re-usable in 
+%other theories. We gathered information about a global string representation 
+%of 
+%a tactic, the time taken to apply this tactic and the input goal and output 
+%list of goals. For printing the tactic string and the time, we just make use 
+%of 
+%printing functions for those types  and 
+%\texttt{Real.toString}. 
+%However for the terms that forms the goals, it is not advisable to use the 
+%function 
+%\texttt{term\_to\_string} as it may not contain all the information necessary 
+%to 
+%reconstruct the same term. Instead we re-use the functions for storing terms 
+%of 
+%theorems in theories \texttt{Term.write\_raw} and and \texttt{Term.read\_raw}, 
+%constructing sharing tables for types and terms.
+%Contrary to \holfour theories that are saved into a \sml file, our data is 
+%stored as text files. This allow us a much faster loading time, since we are 
+%never calling the PolyML compiler. 
+%Already exists in \holfour for recording theorems.
+%Mention better export for HOL4.
+%
+%Side effects:
+%By minimizing the size of the sml files, we manage to multiply by 3 the speed 
+%at which theories are loaded in \holfour. 
+%Mention Ramana Kumar and Michael Norrish.
+
+Finally, we mention some limitations that comes with 
+the need to record tactics as a string compared to being able to store a \sml 
+value. Access to the name of the tactic enhances machine 
+learning ability and user interaction. Yet, because some tactics 
+depends on the state of \holfour, for example the state of the global 
+simpset, a tactic is re-parsed before each proof attempt to be up-to-date.
+As a result, a tactic may have different effect on goals than suggested by the 
+database which could compromise the prediction accuracy. In practice, this 
+should rarely happen. Indeed, most tactics are stateless and an update on a 
+stateful tactic is usually improving it.
+
+%By recording also the sml value it would also possible to eliminate those 
+%drawbacks as long as you stay in the same session. This is typically the case 
+%for the \hollight theorem prover but not in \holfour where every theory is 
+%build in a different session loading previous theorems from disks.
 

### 2018-01-26 tactictoe_jv.tex

@@
 
-\subsection{The problem}\todo{Maybe should }
-
-Definitions of common concepts: goals, tactics, theorems, terms , etc
+\subsection{The problem}\todo{ }
+
+How a user works with an ITP and how this gives rise to a machine learning problem.
+
+Two ways of representing a formulas.
+
+Definition of a tactic in \holfour. In this research, we will take the definition
+that a tactic takes a goal and returns a list of goals.
 
 \subsection{Contributions}
-This project is an extension of the work described in \cite{tgckju-lpar17}. We 
-explain here the additional contributions and their effect across our system:
-
+This work is an extension of the work described in \cite{tgckju-lpar17}.
+There we proposed the idea of ... and achieved ...
+The additional contributions and their effect on our system are:
+
+\todoi{add prediction of 4 kinds of things}
 \begin{itemize}
@@
 \section{Proof Search}\label{sec:proofsearch}
-\todo
-{
-TG:
-describe the prior policy, the evaluation, 
-the node selection algorithm,
-the node extension (+ case of solved goal) + backpropagation.
-add description of end of the algorithm.
-}
+
 Despite the best efforts of the prediction algorithms, the selected tactic may 
@@
 %- Easy to link conjunction estimation   together.
--
-
-\section{ATP Integration}\todo{TG + JU + CK: rework and review}
+
+\subsection{ATP Integration}
 General-purpose proof automation mechanisms which combine proof translation to
@@
 
-Rather than to rely on the underlying programming language, we parse the
+Rather than relying on the underlying programming language, we parse the
 original theory file containing proof scripts into an
@@
 
-\subsection{Globalization}
-
 \holfour proofs are organized in theory files.
@@
 
+\subsection{Globalization}
 %\subsubsection {Unfolding locally declared values}
@@
 
-\subsection{Storing Declaration}
+\subsubsection{Declaration}
 
@@
 
-\subsection{Globalizing Expression}
+\subsubsection{Expression}
 
@@
 
-\subsection{Recording tactic calls}
+\subsection{Tactic wrap}
 
@@
 
-\subsection{Storing Training Data}
-We describe here the format preferred to store and load the 
+\subsection{Data Transfer}
+We describe here the format preferred to save and load the 
 data recorded by the function \texttt{save}. 

### 2018-01-26 tactictoe_jv.tex
More

@@
 timeout.
-And the strongest external provers according to past experiments \cite{hh4h4}.
+And the strongest external provers according to past experiments \cite{tgck-cpp15}.
 Where as...
@@
 This is a generalization of the simplest method used for minimizing a list of 
-theorems in ``hammers'' \cite{}. The fact that lists are generally small
+theorems in ``hammers'' \cite{hammers4qed}. The fact that lists are generally small
 make this approach feasible.
@@
 
+Proof patching: \cite{RingerYLG18}
+
 A similar effort for stronger automation is performed in \isabelle. Nagashima 

### 2018-01-26 tactictoe_jv.tex
prediction: wip

--- 2018-01-26 tactictoe_jv.tex
@@
 %goal list: (product of tactics)
-\begin{itemize}
-\item term:\\
+\paragraph{term}
 Predict the closest term in a tactic that appears in the goal.
 
-\item :\\
+\paragraph{theorem}
 Predict the theorem the closest to the current goal. Theorem dependencies are 
@@
 
-\item goal:\\
+\paragraph{tactic via goal}
 Given a goal $g$, the classifier selects a list of previously solved goals $l$
@@
 
-\item list of goals:\\
+In all proofs, we record each tactic invocation and link (associate) the 
+currently open goal with the tactic's name in our database.
+Given a new open goal $g$, the \emph{score of a tactic $T$ wrt. $g$} is defined 
+to be the score (similarity) of $T$'s 
+associated goal which is most similar to $g$.
+The idea is that tactics with high scores are more likely to solve the open 
+goal $g$, 
+since they were able to solve similar goals before.
+
+Whenever we mention that we want to predict a goal, we 
+actually intend to predict its associated tactics.
+
+\paragraph{list of goals}
 Used for Monte Carlo Tree Search evaluation. Predict the closest list of goals
 to the list of goal produced by a tactic.
-\end{itemize}
 
@@
 
-\subsection{Feature distance}\label{sec:predictions}
-
-In all proofs, we record each tactic invocation and link (associate) the 
-currently open goal with the tactic's name in our database.
-Given a new open goal $g$, the \emph{score of a tactic $T$ wrt. $g$} is defined 
-to be the score (similarity) of $T$'s 
-associated goal which is most similar to $g$.
-The idea is that tactics with high scores are more likely to solve the open 
-goal $g$, 
-since they were able to solve similar goals before.
-
-We estimate the similarity (or co-distance) between an open goal $g_o$ and a 
-previously recorded goal $g_p$ using their respective feature sets $f_o$ and 
-$f_p$.
-The similarity function $sim$ computed by the k-NN algorithm is analogous 
-to the
-one used in the premise selection task~\cite{ckju-pxtp13}. The main idea 
-is to find the features shared by the two goals and estimate the rarity of 
-those features calculated via the TF-IDF~\cite{Jones72astatistical} heuristics.
+\subsection{Similarity}\label{sec:predictions}
+
+We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$ 
+through their respective feature sets $f_1$ and $f_2$.
+%The similarity function $sim$ computed by the k-NN algorithm is analogous 
+%to the one used in the premise selection task~\cite{ckju-pxtp13}.
+The estimation is based on the frequency of the features in the intersection of 
+$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features 
+are, the more similar two goals should be. This can be realized by 
+calculating TF-IDF weights for each feature~\cite{Jones72astatistical}.
+We additionally raise the weights to the sixth power giving even more credence 
+to rare features which was found experimentally optimal in \cite{cezary_cite}.
+
+The first co-distance $sim_1$ adds a modified weight of each shared 
+features to compute the total score.
 In a second co-distance $sim_2$, we additionally take into account 
-the
-total number of 
-features to reduce the seemingly unfair advantage of big feature sets in the 
-first scoring function.
-\[sim_1 (f_o, f_p) = {\sum\nolimits_{\,f \in f_o \cap 
-f_p}{\text{tfidf}(f)^{\tau_1}}}\]
+the total number of features to reduce the seemingly unfair advantage of big 
+feature sets in the first scoring function.
+\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_o \cap 
+f_p}{\text{tfidf}(f)^{6}}}\]
 \[sim_2 (f_o, f_p) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
-f_p}{\text{tfidf}(f)^{\tau_1}}}}
+f_p}{\text{tfidf}(f)^{6}}}}
 {ln (e + card(f_o) + card(f_p))}
@@
 
-\begin{remark}
-We use $\tau_1 = 6$ because it was found optimal in previous experiments.
-\end{remark}
-
-In addition to predicting tactics, we also used these similarity function to 
-predict theorems and terms. Suprisingly, the first similarity function $sim_1$
-is best to predict tactics and theorems. It is probably because there exists 
-monotonocity heuristic where if a tactic can be effective on a large goal, it 
-will be also effective on smaller goals that contain subterms of the larger 
-goal. The same heuristic can be applied for theorems, when a large theorem 
-contains more information than a theorem which is a subformula of the larger 
-theorem. Nevertheless, we want to be able to predict variables (which are small 
-terms) during term argument prediction. So in this case, weight down big terms, 
-as implemented by the $sim_2$ function, should have a beneficial impact.
-The distance $sim_2$ is also used to evaluate list of goals for proof node 
-selection in the Monte Carlo tree search(MCTS) algorithm described in 
-Section~\ref{sec:proofsearch}.
+Making prediction for an object $o$ consists sorting a set of objects 
+by their similarity to $o$.
+\todo{mention preselection and which dataset of objects?}
+
+We use $sim_1$ for goals and theorems and $sim_2$ for terms and list of goals.
+The similarity $sim_1$ is used for predicting a tactic (via goals) because if a
+tactic is effective on large goal $g$ then it is often suitable for a goal $g'$ 
+containing sub-formulas of $g$. The same monotonicity heuristic can justify the 
+use of $sim_1$ in theorems. Indeed, in practice, a large theorem is often a 
+conjunction of small theorems from the same domain. And if a conjunction 
+contains a formula related to a problem, then the other conjuncts 
+from the same domain may also help.
 

### 2018-01-27 tactictoe_jv.tex
removing comments

@@
 
-%
-%\begin{abstract}
-%Formalization of mathematical statements in interactive theorem provers (ITPs) 
-%relies nowadays on many methods to automate the construction of a proof.
-%Some of this method are general purpose and with enough time can prove any 
-%theorems. Yet, in a formalized proof script 
-%a mathematician take advantages of a combination of specialized rules, 
-%theory-based strategy and general ones. All of them are implemented as tactics 
-%in HOL4. Automated theorem provers (ATPs) rely on parameterizable general 
-%=======
 \todo{JU review: Still feels more like an intro than an abstract. Could be 
@@
 the relevance of tactics in particular proof situations.
-% are effective from 
-% tactic-level proofs and what kind of formulas are the results of effective 
-% steps, 
 This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths. 
-TacticToe proves 66.2\% of 7164 theorems of the standard library where as 
+On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.2\% 
+of 7164 theorems of the standard library where as 
 Eprover solves 34.5\%. The success rate rises to 69.0\% by combining the 
 result of both provers.
-
-%on their first-order translations?
-
-%
-%Techniques combining machine learning with translation to automated
-%reasoning have recently become an important component of formal proof 
-%assistants. 
-%Such ``hammer'' techniques complement traditional proof
-%assistant automation as implemented by tactics and decision
-%procedures. n this paper we present a unified proof assistant
-%automation approach which attempts to automate the selection of
-%appropriate tactics and tactic-sequences combined with an optimized
-%small-scale hammering approach.
-%
-%We implement the technique as a tactic-level automation for HOL4:
-%TacticToe. It implements a modified A*-algorithm directly in HOL4 that
-%explores different tactic-level proof paths, guiding their selection
-%by learning from a large number of previous tactic-level
-%proofs. Unlike the existing hammer methods, TacticToe avoids
-%translation to FOL, working directly on the HOL level. By combining tactic 
-%prediction and premise selection, TacticToe is able to re-prove 39\% of 7902 
-%HOL4 theorems in 
-%%less than
-%5 seconds whereas the best single HOL(y)Hammer strategy solves 32\% in the 
-%same 
-%amount of time.
-
 \end{abstract}
 
-\section{Introduction}\todo{TG + JU + CK}
-%\begin{example} Proof automatically generated by \tactictoe for the user given 
-%goal
-%\small
-%\begin{center}
-%
-%
-%\begin{BVerbatim}[commandchars=\\\{\}]
-%Goal: ``\(\forall\)l. FOLDL (\(\lambda\)xs x. SNOC x xs) [] l = l``
-%Proof:
-%  SNOC_INDUCT_TAC THENL
-%  [ REWRITE_TAC [APPEND_NIL, FOLDL],
-%    ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC]
-%      THEN CONV_TAC (DEPTH_CONV BETA_CONV)
-%      THEN ASM_REWRITE_TAC [APPEND_SNOC, FOLDL_SNOC] ]
-%\end{BVerbatim}
-%
-%
-%\end{center}
-%
-%\end{example}
-%
-
+\section{Introduction}
 \tableofcontents
 
-\todo{TG : comparison with hammers and advantages for the user and over hammers}
+\todo{TG : comparison with hammers in usability}
 Many of the state-of-the-art interactive theorem provers (ITPs) such as
@@
   system used with a single strategy and a single efficient external prover.
-%Thanks to a very different proof-search procedure, \tactictoe also
-%proves many theorems whose proofs are not found by the hammering
-%approach.
-
-\subsection{The problem}\todo{ }
-
+
+\subsection{The problem}
 How a user works with an ITP and how this gives rise to a machine learning problem.
@@
 that a tactic takes a goal and returns a list of goals.
+
+Input of the prediction: goal
 
@@
 \sml function. So, our recording is more complete and fully automatized.
-%\item Existing tactic orthogonalization on recorded data is made stronger by 
-%also including the internal ATP \metis in the orthogonalization 
-%competition. Because we manage a database of theorems for argument 
-%prediction, we apply here a pruning step to remove easy theorems.
 \item Two types of arguments (list of theorems and terms) can be predicted 
@@
 
-\section{Prediction}\label{s:prediction}\todo{TG: complete term theorem and 
-list of goals specificities in terms of predictions and justify by their use 
-case}
-
+\section{Prediction}\label{s:prediction}
 The learning-based selection of relevant lemmas significantly improves the 
@@
 methods: the modified distance-weighted \emph{$k$ nearest-neighbour} 
-classifier~\cite{ckju-pxtp13,DudaniS76} to predict four different objects: 
-terms, goals, theorems and list of goals.
-
-Before going into the details of the machine learning algorithm, we introduce
-specificities associated with the prediction of each object: the data from 
-which the feature vectors are build and the different use cases for the 
-prediction.
-
-We derive from tactic selection a prior policy for 
-our search algorithm. Lists of goals (products of tactics) evaluate the success 
-of the policy and adapt it accordingly. Term and theorems predictions are 
-necessary to find the suitable arguments for abstracted tactics (see Section 
-\ref{sec:synthesis}).
-
-%Premise selection usually only prunes the initial set of formulas given to the 
-%ATPs, which then try to solve the pruned problems on their own. 
-%the goal of predicting useful lemmas is to reduce the search space,
-% of the external ATPs called by a hammer,
-
-%Mention explictly all the predictors that will be using this distance.
-%tactics:
-%theorems:
-%terms:
-%goal list: (product of tactics)
-\paragraph{term}
-Predict the closest term in a tactic that appears in the goal.
-
-\paragraph{theorem}
-Predict the theorem the closest to the current goal. Theorem dependencies are 
-added during preselection.
-
-\paragraph{tactic via goal}
-Given a goal $g$, the classifier selects a list of previously solved goals $l$
-ordered by their similarity to $g$. Since each solved goal is associated 
+classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects: 
+ goals, theorems and list of goals.
+
+We first introduce
+specificities associated with the prediction of each object: the dataset from
+which objects are selected and the different use cases for the 
+prediction. We then details how we predict objects suitable for the current 
+goal by comparing their features with the feature of the goal.
+
+\paragraph{Tactics}
+Given a goal $g$, the classifier selects a list of input goals
+ordered by their similarity to $g$. Since each input goal is associated 
 with a tactic in our database, the ordered list of goals induces a ordered  
-list of associated tactics. And tactics are tried in this order during the 
+list of associated tactics which gives us a prior policy function for  MCTS. 
+And tactics are tried in this order during the 
 proof search. Therefore, to order the tactics correctly, it is crucial to 
@@
 actually intend to predict its associated tactics.
-
-\paragraph{list of goals}
-Used for Monte Carlo Tree Search evaluation. Predict the closest list of goals
-to the list of goal produced by a tactic.
-
-%We will next discuss 
-%the extraction 
-%of features from goals and the
-%actual prediction of tactics. 
-%Both have been integrated in the \sml proof search.
+We derive from tactic selection a prior policy for 
+our search algorithm. Lists of goals (products of tactics) evaluate the success 
+of the policy and adapt it accordingly.
+
+\paragraph{Theorems}
+Because we may augment the dataset of tactics with abstracted tactics where 
+list of theorems have been erased, we need to find a way to instantiate them.
+
+This algorithm is also useful for selecting premise to ATPs. (see 
+Section~\ref{sec:atp})
+
+Given a open goal $g$, the classifier selects a list of theorems ordered by 
+their 
+similarity to $g$.
+A dataset of theorems is constructed from the \holfour database and local 
+theorem values.
+We predict theorems most similar to the goal for instantiate list of theorems 
+that appears as arguments of abstracted tactics (see Section 
+\ref{sec:synthesis}). A list of 
+predicted theorems 
+are also given as input to ATPs. The number of predictions is 16 for tactic 
+arguments and \metis, but is increased to 128 for \eprover.
+
+\paragraph{List of goals}
+The dataset of learned list of goals is compiled during orthogonalization.
+This set is separated into positive and negative examples.
+Given a list of goals outputted by a tactic,
+we compute the closest lists of goals in our dataset. The ratio of positive 
+examples from the total number of examples gives us a prior evaluation function 
+for MCTS (see Section~{}).
 
@@
 simplest representation terms. From there, we extend the extraction mechanism 
-to goals theorems and lists of goals. Duplicated features are always removed 
+to goals, theorems and lists of goals. Duplicated features are always removed 
 so each object has an associated set of features.
@@
 in the list.
-
-%We found that the names of variables present in a term are particularly 
-%important for predicting tactics such as 
-%case splitting on a variable 
-%(\texttt{Cases\_on var}) or instantiation of a variable 
-%(\texttt{SPEC\_TAC var term}). 
-%We additionally extract the following features:
-
-%\begin{itemize}
-%\item the top-level logical structure with atoms substituted by a single place 
-%holder $A$ and all its substructures.
-%\end{itemize}
-
-%Determining the presence of top-level logical 
-%operators (i.e 
-%implication) is essential to assess if a "logical" tactic should be applied. 
-%For example, 
-%the presence of an implication may lead to the application of the tactic 
-%\texttt{DISCH\_TAC} that moves the precondition to the assumptions. Top-level 
-%logical structure gives a more detailed view of the relationship between those 
-%logical components.
 
@@
 by their similarity to $o$.
-\todo{mention preselection and which dataset of objects?}
-
-We use $sim_1$ for goals and theorems and $sim_2$ for terms and list of goals.
+
+We use $sim_1$ for goals and theorems and $sim_2$ for list of goals.
 The similarity $sim_1$ is used for predicting a tactic (via goals) because if a
@@
 
-%\[tactic\_score_2 (f_o,f_p) = \frac{tactic\_score_1 (f_o,f_p)}{(1+ ln\ (1 + 
-%card\ f_o))} \]
-
-%Moreover, we would like to compare the distance of a recorded goal to 
-%different goals opened at different moment of the search. That is why we 
-%normalize the scores by dividing them by the similarity of the open goal with 
-%itself. As a result, every score will lie in the interval $[0, 1]$ where 1 is 
-%the highest degree of similarity (i.e. the shortest distance). We respectively 
-%refer to those normalized scores later as $tactic\_norm\_score_1$ and 
-%$tactic\_norm\_score_2$.
-
-\subsection{Preselecting feature vectors and their  
-recursive dependencies}\label{sec:dependencies}
+\subsection{Preselection}\label{sec:dependencies}
+
+The predictions algorithm is called many times during proof search to chose 
+suitable tactics, instantiate tactic arguments by terms and theorems and
+estimate the provability of a list of goals.
+In order to speed up the predictions, we preselect 500 features vectors of 
+tactics, theorems and list of goals.
+
 We first rank feature vectors by their distance to the conjecture. But better 
@@
 recomputed which gives a preferential order of tactics to be tried on $g$.
-
-%In practice, such preselection is necessary because our nearest neighbor 
-%algorithm  is consuming more than half of the proof search time when there were
-%more than 50,000 feature vectors to select from.
-
-%\begin{remark}
-%A preselection technique is also in place for the 
-%theorem predictor and the evaluation predictor defined later in this paper, 
-%because those predictors are also called during the proof search.
-%\end{remark}
 
@@
 
-\subsection{Argument Predictors}\label{sec:arg_pred}
-We made argument predictors for two types of arguments: list 
-of theorems and term. We choose this two types because they are the most common 
-inside tactics and because perfect synthesis of those arguments would lead to 
-the largest improvements (see Section~\ref{sec:perfect_exp}). Arguments with 
-different types 
-are currently never abstracted in tactics since we do 
-not have yet a way to predict those arguments and instantiate those tactics.
-
-\paragraph{List of Theorems}
-Predicting theorems is made through the same system as premise selection
-used for \holyhammer for general purpose automation which produces a list of 
-theorems $l$. The original list of theorem $l'$ abstracted in the tactic is 
-forgotten. For the abstracted tactic 
-to replace the original one, the assumption is that the tactic does not break 
-when adding arguments. If this condition is met, the abstracted tactic will 
-always subsumes the original tactic when $l$ is a superset of $l'$. Such 
-condition is called monotonicity of the 
-tactic in its argument. Predicting such superset may not be necessary as the  
-predicted theorems can sometimes be even more suitable than the original one 
-because they depend directly on the current goal. One of the biggest advantage 
-of predicting theorems independently is that tactics may now rely on theorems 
-that were never part of a tactic.
-
-\paragraph{Term}
-Predicting terms is harder since we do not have a list of terms to choose from.
-As an approximation for this list, we take the list of subterms $l$ contained 
-in a goal $g$. Even then choosing the right subterms require us to look at the 
-term $t$ the tactic argument was abstracted from. The term predictor then 
-selects the subterm in $l$ that is closest to $t$. In an attempt to predict 
-variables (which are terms in \holfour) at the place where they were previously 
-variable, we use the distance $sim_2$ that penalizes large subterms.
-Since we rely on the original term argument $t$, the kind of tactics produced 
-are much less general that for abstracted list of terms and thus have smaller
-coverage.
-
-%In the orthogonalization settings, the abstracted forms compete alongside the 
-%originals and are ranked according to their frequency. 
-%Freq (abstract) = max (orig,abstr). And the abstracted from is always 
-%preferred 
-%to the original form of a tactic if both had the effect that was fought upon.
-%
-%
-%Abstracted tactics then become part of the feature space and can be predicted 
-%by the tactic predictors but are slightly slower because of the added overhead
-%of predicting list of theorems for their unspecified arguments.
-%
-%We first detect if the tactics contains one or more theorem list. This 
-%detection
-%is done by looking if the first element of the list is of type thm.
-%
-%This method is difficult to generalize to any tokens?
-%
-%By forgetting what was the original argument, we lose an important piece of 
-%information on what kind of arguments a tactic accepts. This is less problem 
-%for
-%a list of theorems because predicted theorems that are not usable by the 
-%tactic 
-%are usually ignored.
-%
-%What if you would have to predict a single theorem, or term, or function?
-%
-%There are two complementary guiding methods to predict them:
-%
-%- Dependent on the goal.
-%- Similar to previous arguments for this tactic.
-%
-%For example one way to predict subterms, would be to take the subterm that is 
-%the closest to predicted subterm that appears on the goal.
-%
-%For that we take slight variation of the previously defined distance to 
-%eliminate large subterms:
-
-%\subsubsection{Estimating the difficulty of a goal}
-%A first idea is to guess what is the expected length of its shortest proof.
-%
-%From the recorded feature vectors, we can compute the length of the shortest 
-%proof for each goal $g$ and can be defined recursively by:
-%
-%length :::= mathematical stuff
-%
-%Now to compute the difficulty of a goal. We can average the difficulty of the 
-%$k$ nearest neighbor. 
-%
-%However, we believe that it is more important to try to guess if a goal is 
-%likely to be provable or not.
-%
-%We augment our data by looking at all the produced list of goals
-%that did not subsumes the original one during 
-%orthogonalization. We had the list to our list of negative examples. The list 
-%of positive examples are the goal list that are present in the feature vectors.
-%
-%
-%We try to estimate if a goal was useful by taking the k nearest neighbor of 
-%the 
-%goal and see how many appeared in a previous proofs and how many are hard 
-%negatives. The average give us a probability on how provable it is.
-%We multiply this probability by a coefficient k.
-%
-%Using this heuristic to guide our proof search, we hope to reduce the amount 
-%of 
-%time that the search spends trying to prove unprovable goals.
-%
-%Formula for the heuristic.
-
-%\begin{remark}
-%In practice, each proof node contains a list of goals to be proven and we 
-%estimate the difficulty of proving a conjunction of goals.
-%\end{remark}
+We made argument predictors for one type of arguments: list 
+of theorems. We choose this very common type because perfect instantiation of 
+those arguments would lead to the largest improvements (see 
+Section~\ref{sec:perfect_exp}). We plan to 
+extend our algorithm to term which is another frequent argument types.
+But term prediction have been so far proven to weak to positively 
+influence the proof search.
 
@@
 
-\subsection{A search tree: data structures and definitions}
+\subsection{Search Tree}
 
@@
 
-Maybe should move this defnition in static description.
+Maybe should move this definition in static description.
 
@@
 to the number of proof attempts.
-
 \end{definition}
 
-%A large difference between diagonalized search and our MCTS algorithm is in 
-%what is considered to be the width of the tree. In diagonalized search, 
-%non-productive tactics contribute to the extension of the width where as in 
-%MCTSn the width includes only the total number of children produced. The 
-%influence of non-productive on the proof exploration is recovered when we 
-%include an 
-%evaluation function that takes into account the productivity of tactics such 
-%as 
-%$E_{productive}$ or $E_r$.
-
-%Obviously, we can consider the 
-%trivial evaluation function where 
-%a goal is given reward if and only if it is proven. Yet, to make this approach 
-%effective would require to learn on a very large amount of relatively easy 
-%problems. To lessen the requirement of an very large amount of data, we 
-%propose 
-%to use our goal distance to select the $k$ closest goals that were proven 
-%before. The following formulas shows how we can combine positive examples and 
-%negatives examples to evaluate an unseen 
-%goals.
-%
-%%Hard Negative mining (cite Deep math)
-%
-%\[\mathfrak{N}_m=\lbrace g\ |\ (N(g) \wedge n(g) \geq m) \vee 
-%(P(g) \wedge p(g) >  m)\rbrace\]
-%\[\mathfrak{P}_m=\lbrace g\ |\ P(g) \wedge p(g) \leq m \rbrace\]
-%
-%
-%\[E(m) = \frac{1}{m} \times \frac{card\ \mathfrak{P}_m}
-%    {card\ \mathfrak{N}_m + card\ \mathfrak{P}_m} \]
-%
-%
-%\[E = \argmax_{m \in \mathbb{N}} E(m)\]
-%
-
-%We need
-%MCTS.
-%We use supervised learning to estimate evaluation functions.
-%
-%Positives are given the number of intermediate steps needed to finish the 
-%proof.
-%And the distance for negatives is two times the number of intermediate steps 
-%tried. If we only tried one more tactic on this goal we would solve it.
-
-%We use K-nn on them with k = 10 and calculate an the following weighted mean.
-%
-%
-%The heuristic component should represent the number of steps (or total time) 
-%estimated required to 
-%solve the goal.
-%If a goal is not provable then it's the number of time the goal was visited 
-%during the search + 1.
-
-%
-%
-%Combining with general automation:
-%metis.
-%holyhammer. (with orthogonalized theorems).
-%divide by the sum of the weight of all features.
-%
-%
-%New astar heuristic:
-%The astar algorithm was simplified and now is represented in the standard 
-%additive form instead of the multiplicative one described in our previous 
-%paper.
-%
-%We also tried to estimate the difficulty of a goal. How likely that a 
-%conjunction of goals are provable.
-%
-%Heuristic from prediction score:
-%
-%Heuristic from positives:
-%
-%Heuristic from positives + hard negatives.
-%
-%Admissibility of the heuristic:
-%  
-%Connection with UCB, UCT and Monte Carlo.
-%
-%Why is it not applicable in our case?
-%
-%1) Monte carlo requires to explore every action at least once. 
-%This could be fixed by predicting fewer tactics 10-20. But we do not have
-%yet an automatic way to tell.
-%
-%2) Monte Carlo requires a evaluation of the provability of the goal.
-%This is insight can be gain.
-%
-%\subsection{Node selection formula}
-%Monte Carlo tree search is a tree search algorithm. It explores the proof tree
-%by performing simulation from the root to a leaf of the tree by selecting each 
-%node by looking at that includes exploration and exploitation. Intuitively, 
-%the 
-%trade-off between exploitation and exploration means that the algorithm will 
-%exploit deeper more promises branches but leaving enough time for exoloring 
-%less likely alternatives.
-%
-%
-%
-%Connection with previous tree search. The prior probabilities are decreasing
-%exponential of 2s. $\sqrt (n) > 1/2^n$
-%
-%
-%
-%
-%
-%calls simulation. 
-%
-%
-%
-%\subsection{Description of the evaluation function}
-%
-%Non local heuristic for evaluation. Should consider also what is needed to be 
-%proven.
-%
-%
-%In order to evaluate how long it takes to prove a goal. We collect data from 
-%previous proof searches. Those include negative and positive examples.
-%Positives examples are associated with the minimal number of steps needed to 
-%find their proofs. Negative examples carry along the largest proof search 
-%unsuccessfully performed.
-%
-%Combining this measure in an single evaluation of the proof length is a 
-%challenging task. We use the following formulas P1 for estimating the 
-%probability of being provable in less than one step.
-%
-%Neg(i) = not provable in i steps.
-%70 percent change of not being provable in two steps.
-%50 percent chance of closing the proof in 1 step: 2
-%1 percent chance of being provable in 1 steps.
-%20 percent chance of being provable in 2 steps. 5*2 = value = 10
-%100 percent chance of being provable in 5 steps. 5
-%
-%\[P_i = Neg(i) | Knn(10) \] 
-%
-%the total estimation is given by:
-%\[E = (\sum_{i=1}^5 (1-P_i)) + 1\] 
-%
-%\subsection{Combining evaluation function on goals}
-%
-%Independence of each probabilities.
-%Full correlation: $(P1 * P2 * P3) ^{1/3}$
-%Min probabiity:
-
-%
-%\subsection{Advantage of Monte Carlo Tree Search over A*-search}
-%- Take into account all estimation along the proof and not only the last leaf.
-%- Easy to link conjunction estimation   together.
-
-\subsection{ATP Integration}
+\subsection{ATP Integration}~\label{sec:atp}
 General-purpose proof automation mechanisms which combine proof translation to
@@
 We chose to distinguish two these general purpose tactic \metis and \eprover.
-\metis is integrated in \holfour, so can be call with frequently with low 
+\metis is integrated in \holfour, so is already used by the default \tactictoe 
+but we give it a boost as we are confident it is a strong tactic. The boos 
+consits in ...
+
+can be call with frequently with low 
 timeout.
@@
 from slowing down the search.
-
-%The number of parallel threads can be capped.
-%And if tactictoe or an asynchronous thread proves a goal, all descendants of 
-%these goal are deactivated, terminating associated threads. This frees space 
-%for more pending threads, associated to newly created nodes, to start.
-
-%needs to be taken after each proof step to close proofs solved by \eprover
-%When the number of thread reaches a limit, a ``hammer'' tactic call 
-%should
-%wait for a thread to be freed but the proof search goes on and the feedback of 
-%the ``hammer'' tactic is delayed. 
-%
-%
-%It is best tactic available but we do not have any data about it since it does 
-%not appear in any proofs. So , we 
-%artificially try it on
-
-%\paragraph{Effect on orthogonalization of tactics}
-%During orthogonalization, the general purpose tactic will be the first 
-%competitor and if it manages to prove the goal it will subsume the original 
-%tactic on this goal since the empty set of goals is a subset of any set. 
-%Therefore, it will replace the original tactic in the feature vector that is 
-%competed for. This is desirable since we would like to remove alternative 
-%proofs.
-
-%\paragraph{Orthogonalization of theorems}
-%A ``hammer'' tactic can be useful to remove parallel
-%theorems. We consider a theorem to be parallel if it can be proven 
-%from previous theorems using a general tactic with a timeout equal or less 
-%than the one it is given during the proof search. Parallel theorems are marked 
-%and removed from the set of theorems $\mathfrak{T}$ from 
-%which premise selection is performed.
-%
-%\begin{remark}
-%The ``smaller hammer'' is chosen to remove parallel theorems and as the first 
-%try during orthogonalization of tactics. 
-%\end{remark}
-
-%\paragraph{Effect on the proof search}
-%First, before the proof search, we preselect 500 theorems 
-%for the whole proof search tree using a premise selection algorithm 
-%that takes into account dependencies between theorems.
-%During the proof search, when a new goal is created or a fresh pending goal is 
-%considered, the ``hammer'' tactic is always tried first because if it succeeds 
-%the node would be closed rapidly.
-
-%\paragraph{Parallelization of external calls}
-%It is possible to parallelize calls to external provers so that they happen 
-%asynchronously during the proof search. We let the prover runs \tactictoe 
-%continue its proof in parallel of a ``hammer'' tactic call.
 
@@
 \midrule
- $0.66$ & 120\\
- $0.5$ & 125\\
- $0.33$ & 117\\
+ $2/3$ & 120\\
+ $1/2$ & 125\\
+ $1/3$ & 117\\
 % 1/2 & Productive & No & $E_{none}$ & 1 & 120\\
@@
 This experiments are performed with orthogonalization on and policy coefficient 
-of $0.5$.
+of $1/2$.
 
@@
 \toprule
- Evaluation & $c_{exploration}$ & Solved \\
+ PriorEvaluation & $c_{exploration}$ & Solved \\
 \midrule
@@
 % 1/2 & All & Yes & $E_{none}$ & 1 & 128\\
- $E_{none}$ & 1 & 128\\
- $E_{productive}$ & 1 & 134\\ 
+ none & 1 & 128\\
+ productive & 1 & 134\\ 
  & 4 & 130\\
- $E_1$ & 1 & 127\\ 
+ radius = 1 & 1 & 127\\ 
  & 4 & 126\\ 
- $E_{10}$ & 1 & 136\\ 
+ radius = 10 & 1 & 136\\ 
  & 4 & 137\\ 
@@
 \end{tabular}
-\caption{\label{tab:learning} 
+\caption{\label{tab:search_parameters}}
 %on the server
-}
 \end{table}
@@
 further investigation.
-
-%\begin{table}[ht]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{ll}
-%\toprule
-% & Solved \\
-%\midrule
-% diag coeff: 0.5 & 120\\
-% diag coeff: 1   & 125\\
-% diag coeff: 1 + orthogonalization & 130 \\
-% diag coeff: 2 &  117\\
-% no evaluation & 120\\
-% trivial evaluation E1 & 127\\
-%\bottomrule
-%\end{tabular}
-%\caption{\label{tab:cost_param} on a local machine}
-%\end{table}
-%
-%\begin{table}[ht]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{ll}
-%\toprule
-% & Solved \\
-%\midrule
-% none & 122\\
-% ortho & 129\\
-% no evaluation + ortho & 128\\
-% trivial evaluation + ortho    & E1:134 E4:130\\
-% evaluation radius: 1 + ortho  & E1:127 E4:126\\
-% evaluation radius: 10 + ortho & E1:136 E4:137\\
-%\bottomrule
-%\end{tabular}
-%\caption{\label{tab:learning} on the server}
-%\end{table}
-
-We preselect around each feature vectors, one hundred feature of but this takes 
-time. What kills us is the time needed to select the neighbors among the 
-maximum of 50,000 feature vectors but maybe we can hope to 
-taking half the time of the proof search.
 
@@
 \end{tabular}
-\caption{\label{tab:featue_param} Success rates of with an artificially 
-augmented dataset of features fetched from the human proof of the tested 
-theorem}
+\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
+with tactics from the proof of the tested theorem.} 
 \end{table}
+\todo{good title for this table Additional features}
 
@@
 
-\subsubsection{Abstraction}
+\subsubsection{Orthogonalization Abstraction}
 Here, we test the benefit of different kind of tactic abstraction.
@@
 \toprule
-  & Solved \\
+ Technique & Solved \\
 \midrule
 % holyhammer & 79 \\
- none & 122\\
- ortho & 129\\
- ortho + term & 126\\    
- ortho + theorem list & 160\\
- ortho + \metis & 162\\
+  & 122\\
+ Orthogonalization & 129\\
+% ortho + term & 126\\    
+ Orthogonalization + Abstraction & 160\\
+% ortho + \metis & 162\\
  %ortho + metis + thmortho & 155\\
  %ortho + metis + stacortho & 160\\
- ortho + \metis + theorem list & 175\\ 
+% ortho + \metis + theorem list & 175\\ 
 \bottomrule
@@
 \caption{\label{tab:cfot_param} 
-Sucess rates of tactictoe with synthesis for different arguments.}
+ Effect of orthogonalization and abstraction}
 \end{table}
@@
 
-%The third experiment, presented in Table~\ref{tab:lh_param},
-%evaluates the effect of integrating the ``small hammer'' in the \tactictoe 
-%search.
-%At a first glance, the increased success rate is significant for all tested 
-%parameters. 
-%Further analysis reveals that increasing the number of premises from 8 to 16 
-%with a 
-%timeout of 0.02 seconds is detrimental. The $D_{19}$ experiment  demonstrates 
-%that 0.1 
-%seconds is a better time limit for reasoning with 16 premises. And the 
-%$D_{18}$ 
-%experiment reveals the disadvantage of unnecessarily increasing the timeout of 
-%\metis. This reduces the time available for the rest of the proof search, 
-%which 
-%makes the success rate drop.
-
+\todo{paragraph is copied}
 The best setting for \metis was to be in previous experiments to be 16
-theorems and
-
-%\begin{table*}[t]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{llll}
-%\toprule
-% ID & ``small hammer'' parameter & Solved &  $U(D_{19})$ \\
-%\midrule
-% $D_9$ & $codist_5(0.8,0.8)$ (default: no small hammer) & 211 (24.5\%) & 19 \\
-% $D_{16}$ & 8 premises + timeout 0.02 sec & 281 (32.7\%) & 21 \\ 
-% $D_{17}$ & 16 premises + timeout 0.02 sec & 270 (31.4\%) & 18 \\
-% $D_{18}$ & 8 premises + timeout 0.1 sec & 280 (32.6\%) & 11 \\
-% $D_{19}$ & 16 premises + timeout 0.1 sec & 289 (33.6\%) & 0 \\
-%\bottomrule
-%\end{tabular}
-% \caption{\label{tab:lh_param}Success rate of strategies with different 
-% parameters of
-% ``small hammer'' on the training set.}
-%\end{table*}
-
-The best strategy which does not rely on the ``small hammer'' approach $D_9$ 
+theorems an The best strategy which does not rely on the ``small hammer'' 
+approach $D_9$ 
 will be called \tactictoe(NH) (for no ``small hammer'') in the remaining part 
@@
 referred to as \tactictoe(SH) (``small hammer'').
-
-%\begin{table}[h!]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{lll}
-%\toprule
-% ID & Orthogonalization of theorems & Solved \\
-%\midrule
-% $NONE$           &                             & 118\\
-% $METIS16$        & using metis                 & 143\\
-% $METIS16_NEW$    & no theorems from boolTheory & 155\\
-% $METIS16v2NOINT$ & no internal theorems        & 159\\
-% $METIS16v2DIAG$  & diagonal exploration        & 166\\ 
-% $METIS16v2INT$   & internal theorems           & 168\\
-% $METIS16_30$     & 30 seconds timeout          & 162\\
-%\bottomrule
-%\end{tabular}
-%\caption{\label{tab:cst_param}}
-%\end{table}
 
@@
 
-\subsection{Slight improvements}\todo{maybe remove this subsection}
+\subsection{Recent improvements}\todo{maybe remove this subsection}
 The final results could be improved by a few percent by including the 
@@
 also probably give a 3 percent increase in the number of proven theorems as 
-shown by a 1 in 10 evaluation on the full library.
+shown by a recent 1 in 10 evaluation on the full library.
 
@@
 
-%\subsection{Recording Tactic Calls}
-%We first discuss parsing the proofs and identifying 
-%individual tactic calls. We next show how tactics calls are recorded together 
-%with their corresponding goals using a modified proof script.
-%Finally, the recorded data is organized as feature vectors with tactics as 
-%labels and characteristics of their associated goals as features. 
-%These feature vectors constitute the training set for our selection algorithm.
+We first discuss parsing the proofs and identifying 
+individual tactic calls. We next show how tactics calls are recorded together 
+with their corresponding goals using a modified proof script.
+Finally, the recorded data is organized as feature vectors with tactics as 
+labels and characteristics of their associated goals as features. 
+These feature vectors constitute the training set for our selection algorithm.
 
@@
 
-%For a more formal explanation, we define a target AST build upon the original 
-%one, where each identifer $I$ is replace by its globalized expression $E'$:
-%
-%\begin{align*}
-%M' &=_{def} string \\
-%I' &=_{def} string\ \vert\ Module(M',I')\\
-%D' & =_{def} Open(M')\ \vert\  Val (I',E') \\
-%E' &=_{def} Code(E'\ list)\ \vert\ Let (D',E') 
-%\end{align*}
-%
-%Then, we express two mutually recursive functions $f_d$ and 
-%$f_e$ that globalize the original representation. The 
-%function $f_d$ modifies declaration and $f_e$ modifies expressions. Since a 
-%\sml program is a list of declarations, we will be able transform it by 
-%iteratively applying $f_d$. The stack 
-%$\mathbb{S}$ keep track of each previous declaration: the identifier that is 
-%declared and their globalized definition. To save a declaration to 
-%the stack, we use the $push$ 
-%function. Opening a module has the effect of pushing all identifiers with 
-%their associated prefixed version to the stack. The scope of an identifier is 
-%automatically handled by construction of the AST. To find the globalized 
-%version of an identifier $i$, we use the $lookup$ function on a stack $s$, 
-%which returns the globalization associated with $i$ appearing highest in $s$. 
-%We then build $f_e$ and $f_d$ in the following way:
-%\vspace{5mm}
-%
-%\begin{math}
-%\begin{array}{lcl}
-%f_d:& \mathbb{S} \times D &\rightarrow\ \ \mathbb{S} \times D'\\
-%& (s, Open(m)) &\mapsto\ \ (push(m,s),\ Open(m)) \\
-%& (s, Val(i,e)) &\mapsto\ \ (push ((i,\hat{e}),s),\ Val(i, \hat{e})) 
-%\mbox{ where } \hat{e} = \pi_2^2 (f_e (s,e))
-%\end{array}
-%\end{math}
-%
-%\vspace{5mm}
-%
-%\begin{math}
-%\begin{array}{lcl}
-%f_e:& \mathbb{S} \times E &\rightarrow\ \ \mathbb{S} \times E'\\
-%& (s, List(l))  &\mapsto\ \ (s,\ lookup(s,l))\\
-%& (s, Let (d,e)) &\mapsto\ \ (\hat{s}, f_e (\hat{s}, e))
-%\mbox{ where } \hat{s} = \pi_1^2 (f_d (s,d))
-%\end{array}
-%\end{math}
-%
-%\vspace{5mm}
-
-%\begin{example} Example
-%\small
-%\begin{alltt}
-%val MAP_APPEND = store_thm ("MAP_APPEND", 
-%  --`!(f:'a->'b).!l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f 
-%l2)`--,
-%    STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]);
-%\end{alltt}
-%\end{example}  
-%
-%\begin{example} Running example (extracted proof)
-%\small
-%\begin{alltt}
-%STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
-%\end{alltt}
-%\end{example}
-
-%The goal of our program is to transform the original program is to associate to
-%each expr token.
-
-%\subsubsection{Structures}
-%Structures:
-%
-%Globalizing functions:
-%examples:
-%
-%let val x = 5;  end;
-%substructres
-
-%Thanks to the simplicity of the minimal AST, we are able to present our 
-%recursive unfolding algorithm in a concise and formal manner. But for 
-%practical 
-%purpose, the minimal AST is not sufficient.
-
 We present the tactical proof recording mechanism in details here because it 
@@
 
-%\sml expression delimiters ``([,]){}='' are left 
-%unchanged by the globalization process.
-%
-% pattern matching
-%By distinguishing which values are constructors in our Pattern matching in 
-%\sml 
-%is allowed inside a value 
-%
-%
-%infix operators
-%Our goal is to create a globalized version of each generalizable token.
-%
-%In this description we are omitting handling of "fun", "case of | =>", "fn =>"
-%The only thing to consider is that var appearing as arguments of a function.
-%
-%infix constructor: 
-%  Infix constructors cannot be replaced by their globalized version without 
-%  confusing the parser.
-%
-%  
-%
-%  We can globalize an infix operartor locally in the following way;
-%
-%Original code: some\_code1 infix some\_code2
-%
-%the infix code does not carry the information on how to parse it.
-%
-%            infix\_open (infix\_code) infix\_close.
-%
-%where infix\_open and infix\_closed are declared to be the same arity with the 
-%same associactivity rule as infix.
-%
-%Depending if the operator is left associative or right associative we give two 
-%different definitions for infix\_open and infix\_close.
-%
-%
-%
-%constructions with bounded variables: fun (case of) (
-%
-%proof scripts (arguments of prove, store\_thm, maybe\_thm)
-%
-%watch list:
-%
-%stateful tactics
-
 \subsection{Tactic wrap}
@@
 bindings in a \sml file. All other theory data is now stored separately 
-into a text file. Michael Norrish and Ramana Kumar contributed to integrated 
-these changes in the current distribution of \holfour. After the changes, 
-\holfour theories load three times faster in average.
+into a text file. After integrating the changes, \holfour theories load three 
+times faster in average.
 
@@
 for the two other fields.
-
-%The previous step demonstrated how to extract training data from proofs. We 
-%now 
-%need to store this data on disk so that this training data is re-usable in 
-%other theories. We gathered information about a global string representation 
-%of 
-%a tactic, the time taken to apply this tactic and the input goal and output 
-%list of goals. For printing the tactic string and the time, we just make use 
-%of 
-%printing functions for those types  and 
-%\texttt{Real.toString}. 
-%However for the terms that forms the goals, it is not advisable to use the 
-%function 
-%\texttt{term\_to\_string} as it may not contain all the information necessary 
-%to 
-%reconstruct the same term. Instead we re-use the functions for storing terms 
-%of 
-%theorems in theories \texttt{Term.write\_raw} and and \texttt{Term.read\_raw}, 
-%constructing sharing tables for types and terms.
-%Contrary to \holfour theories that are saved into a \sml file, our data is 
-%stored as text files. This allow us a much faster loading time, since we are 
-%never calling the PolyML compiler. 
-%Already exists in \holfour for recording theorems.
-%Mention better export for HOL4.
-%
-%Side effects:
-%By minimizing the size of the sml files, we manage to multiply by 3 the speed 
-%at which theories are loaded in \holfour. 
-%Mention Ramana Kumar and Michael Norrish.
 
@@
 stateful tactic is usually improving it.
-
-%By recording also the sml value it would also possible to eliminate those 
-%drawbacks as long as you stay in the same session. This is typically the case 
-%for the \hollight theorem prover but not in \holfour where every theory is 
-%build in a different session loading previous theorems from disks.
 
@@
 }
-%Investigating further the different qualities of \tactictoe, we study its 
-%generated proof scripts on an example in \texttt{list} theory (see 
-%Example~\ref{ex:casestudy}). 
-%The theorem to be proven states the equivalence 
-%between the fact that a number $n$ is greater than the length of a list 
-%$\mathit{ls}$ with the 
-%fact that dropping $n$ elements from this list returns an empty list.
-%
-%The human proof proceeds by induction on $n$ followed by solving both goals 
-%using rewrite steps combined with an arithmetic decision procedure. 
-%Both \tactictoe proofs (NH and SH) follow the general idea of reasoning by 
-%induction but solve 
-%the base case and the inductive case in a different way.
-%The base case only needs rewriting using the global simpset in the 
-%\tactictoe(NH)
-%proof, which is simulated by a call to \metis in the (SH) proof.
-%The inductive case should require an arithmetic decision procedure as hinted 
-%by 
-%the human 
-%proof. This is achieved by rewriting using an arithmetic simpset in the second 
-%proof. In the first proof however, a rewriting step and case splitting step 
-%were used
-%to arrive at a point where \metis calls succeed.
-%The tactic proof produced by \tactictoe(NH) often looks better than the one 
-%discovered by \tactictoe(SH) in that it does not involve \metis calls with a 
-%large numbers of premises.
-
-%In general, \tactictoe does not search for beauty, but sometimes it is 
-%discovered by accident.
-
-%The term and the conversion rules were not synthesized but simply copied from 
-%previous 
-%proofs.
-
-%Name: 218 SURJ_INJ_INV
-%Original proof time: 0.002033
-
-%val SURJ_INJ_INV = store_thm(
-%  "SURJ_INJ_INV",
-%  ``SURJ f s t ==> ?g. INJ g t s /\ !y. y IN t ==> (f (g y) = y)``,
-
-%SURJ_INJ_INV
-%proof length: 2
-%proof time: 0.050853
-%Statistics
-%  infstep : 31
-%  nodes   : 11
-%  maxdepth: 3
-%Time: 1.300789
-%  inferstep time: 1.071869
-%  node_find time: 0.018046
-%  node_crea time: 0.147604
-%    pred time: 0.04263
-%    thmpred time: 0.000005
-%    mc time: 0.071282
-%    inst time: 0.0
-%Proof found: 
-%Org tac number: 15
-%Replaying proof: 0.562829
-
-%Name: 0 IS_GCD_UNIQUE
-%proof length: 12
-%proof time: 0.080819
-%Statistics
-%  infstep : 2539
-%  nodes   : 396
-%  maxdepth: 7
-%Time: 40.309547
-%  inferstep time: 30.15097
-%  node_find time: 0.58993
-%  node_crea time: 9.142918
-%    pred time: 2.425785
-%    thmpred time: 0.000101
-%    mc time: 4.417004
-%    inst time: 0.0
-%Proof found: STRIP_TAC THEN REWRITE_TAC [fetch "gcd" "is_gcd_def"] THEN REPEAT 
-%Cases THENL [metisTools.METIS_TAC [], REWRITE_TAC [SUC_NOT, ALL_DIVIDES_0, 
-%compute_divides] THEN metisTools.METIS_TAC [numTheory.NOT_SUC], 
-%metisTools.METIS_TAC [numTheory.NOT_SUC, DIVIDES_ANTISYM], metisLib.METIS_TAC 
-%[LESS_EQUAL_ANTISYM, (fetch "divides" "DIVIDES_LE"), prim_recTheory.LESS_0], 
-%metisTools.METIS_TAC [], (RW_TAC numLib.arith_ss) [(fetch "divides" 
-%"divides_def")], metisTools.METIS_TAC [DIVIDES_ANTISYM], metisLib.METIS_TAC 
-%[LESS_EQUAL_ANTISYM, (fetch "divides" "DIVIDES_LE"), prim_recTheory.LESS_0]]
-%Org tac number: 1
-%Original proof time: 0.0
-%Recording proof time: 0.008556
-
-%
-%LUPDATE_APPEND2
-%proof length: 6
-%proof time: 0.017156
-%Statistics
-%  infstep : 70
-%  nodes   : 27
-%  maxdepth: 4
-%Time: 3.02716
-%  inferstep time: 2.12842
-%  node_find time: 0.07146
-%  node_crea time: 0.508384
-%    pred time: 0.164047
-%    thmpred time: 0.000012
-%    mc time: 0.265314
-%    inst time: 0.0
-%Proof found: 
-%Org tac number: 10
-%Original proof time: 0.0
-%Replaying proof: 0.470118
 
@@
 
-Original \holfour proof script (2.03 milliseconds)
+Original \holfour proof script (2 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 Make it match the new version.
-\tactictoe proof script (50.85 milliseconds)
+\tactictoe proof script (50 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 
-Original \holfour proof script (62.69 milliseconds)
+Original \holfour proof script (63 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 Make it match the new version.
-\tactictoe proof script (17.16 milliseconds)
+\tactictoe proof script (17 milliseconds)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 \end{lstlisting}
-fs () is short for FULL\_SIMP\_TAC (srw\_ss ()) and rw is short for ??.
+fs is a \sml binding for FULL\_SIMP\_TAC (srw\_ss ()) and rw is short for ??.
 
@@
 \coq library demonstrating that it globally outperforms individual \coq 
-tactics. In contrast, we demonstrate the competitiveness of our system against 
-the ATP \eprover on the \holfour standard library.
-A recent improvement on this system~\cite{}\todo{Cezary: could not remember the 
-reference that you gave me here i think it was CICM 2017} includes the 
-possibility to 
-substitute theorems appearing in tactics by ones belonging to the same cluster. 
-In this work, we are also able to predict theorems appearing as arguments of 
-tactics but we rely only on information from the current goal. Combining the 
-two approaches by predicting from the goal and from the original tactic 
-arguments is more difficult but is likely to produce more suitable lemmas.
+tactics. Here, we demonstrate the competitiveness of our system against 
+\eprover on the \holfour standard library.
+%A recent improvement on this system~\cite{}\todo{Cezary: could not remember 
+%the 
+%reference that you gave me here i think it was CICM 2017} includes the 
+%possibility to 
+%substitute theorems appearing in tactics by ones belonging to the same 
+%cluster. 
+%In this work, we are also able to predict theorems appearing as arguments of 
+%tactics but we rely only on information from the current goal. Combining the 
+%two approaches by predicting from the goal and from the original tactic 
+%arguments is more difficult but is likely to produce more suitable lemmas.
 
@@
 
-\section{Conclusion}\label{sec:concl}\todo{TG + JU or CK review}
-
-Future work:
-  Promising techniques that could not be evaluated. 
-  + Orthogonalization of theorems~\cite{ckju-jsc15}. 
-  + Try harder and smarted on term predictions. \cite{latest-version-of-sepia}. 
-  \cite{conjecturing}
-  + Self-learning.  
-+ MCTS was presented in manner that we only did the first step of reinforcement 
-learning. CurrentPolicy becomes PriorPolicy.
-
-More training than just orthogonalization of tactics.
+\section{Conclusion}\label{sec:concl}
 
@@
 
-\todo{is copied (modified a bit)}
+\todo{is copied and modified}
 We proposed a new proof assistant automation technique which combines 
@@
 
+%Future work:
+%  Promising techniques that could not be evaluated. 
+%  + Orthogonalization of theorems~\cite{ckju-jsc15}. 
+%  + Try harder and smarted on term predictions. 
+%  \cite{conjecturing}
+%  + Self-learning.  
+%+ MCTS was presented in manner that we only did the first step of 
+%reinforcement 
+%learning. CurrentPolicy becomes PriorPolicy.
+% + More training than just orthogonalization of tactics.
+
 %%%Improve tactic pred with term predictions. 

### 2018-01-28 tactictoe_jv.tex
predictions: almost done

@@
 
-\todo{TG : comparison with hammers in usability}
+\todo{Comparison with hammers in usability}
 Many of the state-of-the-art interactive theorem provers (ITPs) such as
@@
 
-\todo{TG: rewrite the plan}
+\todo{Rewrite the plan}
   To do this, we \emph{extract and record} tactic invocations from the ITP 
@@
 \subsection{The problem}
-How a user works with an ITP and how this gives rise to a machine learning problem.
-
+How a user works with an ITP and how this gives rise to a machine learning 
+problem.
 Two ways of representing a formulas.
-
 Definition of a tactic in \holfour. In this research, we will take the definition
 that a tactic takes a goal and returns a list of goals.
-
 Input of the prediction: goal
+If a tactic produces no goals then its input goal is solved.
 
@@
 
-\todoi{add prediction of 4 kinds of things}
+\todoi{add prediction of 3 kinds of things}
 \begin{itemize}
@@
 classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects: 
- goals, theorems and list of goals.
-
-We first introduce
-specificities associated with the prediction of each object: the dataset from
-which objects are selected and the different use cases for the 
-prediction. We then details how we predict objects suitable for the current 
-goal by comparing their features with the feature of the goal.
+ tactics, theorems and list of goals.
+
+We first introduce specificities associated with the prediction of each object. 
+In particular, we present the dataset from which objects are selected and the 
+purpose of the selection process. 
+We then explain the construction of the similarity measure by which relevant 
+objects are predicted.
 
 \paragraph{Tactics}
-Given a goal $g$, the classifier selects a list of input goals
-ordered by their similarity to $g$. Since each input goal is associated 
-with a tactic in our database, the ordered list of goals induces a ordered  
-list of associated tactics which gives us a prior policy function for  MCTS. 
-And tactics are tried in this order during the 
-proof search. Therefore, to order the tactics correctly, it is crucial to 
-estimate the similarity between goals with mathematically relevant features.
-
-In all proofs, we record each tactic invocation and link (associate) the 
-currently open goal with the tactic's name in our database.
-Given a new open goal $g$, the \emph{score of a tactic $T$ wrt. $g$} is defined 
-to be the score (similarity) of $T$'s 
-associated goal which is most similar to $g$.
-The idea is that tactics with high scores are more likely to solve the open 
-goal $g$, 
-since they were able to solve similar goals before.
-
-Whenever we mention that we want to predict a goal, we 
-actually intend to predict its associated tactics.
-We derive from tactic selection a prior policy for 
-our search algorithm. Lists of goals (products of tactics) evaluate the success 
-of the policy and adapt it accordingly.
-
-\paragraph{Theorems}
-Because we may augment the dataset of tactics with abstracted tactics where 
-list of theorems have been erased, we need to find a way to instantiate them.
-
-This algorithm is also useful for selecting premise to ATPs. (see 
-Section~\ref{sec:atp})
-
-Given a open goal $g$, the classifier selects a list of theorems ordered by 
-their 
-similarity to $g$.
-A dataset of theorems is constructed from the \holfour database and local 
-theorem values.
-We predict theorems most similar to the goal for instantiate list of theorems 
-that appears as arguments of abstracted tactics (see Section 
-\ref{sec:synthesis}). A list of 
-predicted theorems 
-are also given as input to ATPs. The number of predictions is 16 for tactic 
-arguments and \metis, but is increased to 128 for \eprover.
-
-\paragraph{List of goals}
-The dataset of learned list of goals is compiled during orthogonalization.
+To start with, we build a database of tactics consisting
+of (tactic,goal) pairs recorded from successful human proofs (see 
+Section~\ref{sec:recording}).
+Then, during proof search, we order recorded goals according to their 
+similarity with a targeted goal $g$. This induces through the recorded pairs an 
+order on tactics. The intuition is that tactics that have been successful on 
+goals similar $g$ should be tried first. 
+This predicted tactic order is translated into a prior policy for our MCTS 
+algorithm (see Section~\ref{sec:policy}).
+Additionally, tactic selection allows us to improve the quality of the database 
+of tactics during orthogonalization (see Section~\ref{sec:ortho}).
+
+\paragraph{Theorems as Argument of Tactics}
+We first collect a set of theorems from the \holfour theorem database and local 
+theorem values, for our theorem predictor to select from.
+Predicted tactics during proof search and orthogonalization may contain 
+abstracted tactics where list of theorems have been erased (see Section 
+\ref{sec:synthesis}).
+We instantiate those tactics by theorems that are syntactically the closest to 
+the targeted goal as they are more likely to help.
+The same algorithm selects suitable premises for ATPs integrated with  
+\tactictoe (see Section~\ref{sec:atp}).
+
+\paragraph{List of Goals as Output of Tactics}
+A dataset of list of goals is compiled during orthogonalization (see 
+Section~\ref{sec:evaluation}).
 This set is separated into positive and negative examples.
-Given a list of goals outputted by a tactic,
-we compute the closest lists of goals in our dataset. The ratio of positive 
-examples from the total number of examples gives us a prior evaluation function 
-for MCTS (see Section~{}).
+List of goals that subsumes the result of the original tactic on a 
+targeted goal are considered positive. During proof search, given a list of 
+goals $l$ created by a tactic, we select lists of goals that are most similar 
+to $l$.
+The ratio of positive examples in the selection gives us a prior node 
+evaluation for MCTS.
 
@@
 f_p}{\text{tfidf}(f)^{6}}}\]
-\[sim_2 (f_o, f_p) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
+\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
 f_p}{\text{tfidf}(f)^{6}}}}
-{ln (e + card(f_o) + card(f_p))}
-\]
-
-Making prediction for an object $o$ consists sorting a set of objects 
-by their similarity to $o$.
-
-We use $sim_1$ for goals and theorems and $sim_2$ for list of goals.
-The similarity $sim_1$ is used for predicting a tactic (via goals) because if a
+{ln (e + card(f_o) + card(f_p))}\]
+
+In our setting, making prediction for an object $o$ consists of sorting a set 
+of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems 
+and $sim_2$ for list of goals. The similarity $sim_1$ is used for predicting a 
+tactic via its associated goal because if a
 tactic is effective on large goal $g$ then it is often suitable for a goal $g'$ 
 containing sub-formulas of $g$. The same monotonicity heuristic can justify the 
-use of $sim_1$ in theorems. Indeed, in practice, a large theorem is often a 
+use of $sim_1$ in theorems. Indeed in practice, a large theorem is often a 
 conjunction of small theorems from the same domain. And if a conjunction 
@@
 
-The predictions algorithm is called many times during proof search to chose 
-suitable tactics, instantiate tactic arguments by terms and theorems and
-estimate the provability of a list of goals.
-In order to speed up the predictions, we preselect 500 features vectors of 
-tactics, theorems and list of goals.
-
-We first rank feature vectors by their distance to the conjecture. But better 
-ranked vectors only contains symbols from 
-the conjecture and may not be suitable for later stages of the proof where the 
-goals may have derived significantly from the conjecture. To solve this 
-problem, we reorder the list so that descendants (see 
-Definition~\ref{def:recdep}) of vectors inherits the similarity of their 
-parent.
-
-\begin{definition}(Recursive dependencies of a feature 
-vector)\label{def:recdep}
-Let $\mathfrak{F}$ be a set of recorded feature vectors. 
-The set of all recursive dependencies $D_\infty$ of a feature vector $f$ is 
+In order to speed up the predictions during the search for a proof of a 
+conjecture $c$, we preselect 500 tactic-goal pairs and theorems. 
+Preselection for list of goals is induced by the preselection of the input 
+goals in the tactic-goal pairs. Those will 
+be the only objects available to \tactictoe for proving $c$.
+
+The first idea is to select tactic-goal pairs and theorems by their distance to 
+the conjecture. However, this selection may not be adapted to later stages of 
+the proof where goals may have derived significantly from the 
+initial conjecture $c$. In order to 
+anticipate the production of diverse goals, our prediction during preselection 
+take into account dependencies between objects of the same dataset.
+The dependence relation is assymetric. Through the relation, each object has 
+multiple children and at most one parent.
+Once a dependency relation is established, we can calculate a dependency score 
+for each object which is the maximum of its similarity score and the similarity 
+score of its parent. In the end, the 500 feature vectors in each dataset with 
+the highest dependency score are preselected.
+
+In the following, we give a mathematical definition for the dependency relation 
+between labels for tactic-goal pairs and theorems.
+
+\begin{definition}(Dependencies of tactic-goal pairs)\\
+Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
+The dependencies $D_\infty$ of a tactic-goal pair $(t_0,g_0)$ is 
 inductively defined by:
@@
 \begin{align*}
-D_0 &=_{def} \lbrace f \rbrace \\
+D_0 &=_{def} \lbrace (t_0,g_0) \rbrace \\
 D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists 
@@
 
-We augment the ordered predictions by inserting after each predicted vector 
-its descendants and removing lower ranked duplicates. 
-%This contrast with 
-%theorem predictions where we only select parents. 
- This phase is only applied 
-during preselection, so is of no great consequence as long as we select enough 
-theorems and tactics.
-
-Since predicting efficiently is crucial during proof 
-search, we preselect a number of tactics before the search.
-We take the first 500 vectors of the augmented ordered list. Only members of 
-this list can be predicted during proof search. When an 
-open goal $g$ is created, the scores of those 500 preselected tactics are
-recomputed which gives a preferential order of tactics to be tried on $g$.
-
-\section{Orthogonalization}\todo{MN + RK: review + fix typos}
+\begin{definition}(Dependencies of a theorem)\\
+Let $\mathfrak{T}$ be the set of recorded theorems.
+The dependencies $D_\infty$ of a theorem $t$ are the set of theorems in 
+$\mathfrak{T}$ appearing in the proof $t$ as recorded in \cite{hh4h4}.
+Contrarily to the dependency relation in tactic-goal pairs, this relation is 
+not a transitive closure.
+\end{definition}
+
+To preselect a list of goal $l$, we consider the tactic input $g$ that was 
+at the origin of the production of this list. If and only if there exists a 
+tactic $t$ such as $(t,g)$ is in the preselected tactic-goal pairs, the 
+list of goals $l$ is preselected. This has the desirable effect that
+negative and positive examples created during one orthogonalization 
+step are chosen or discarded together. \todo{problematic forward reference?}
+
+\section{Orthogonalization}\label{sec:ortho}
 
@@
 We made argument predictors for one type of arguments: list 
-of theorems. We choose this very common type because perfect instantiation of 
-those arguments would lead to the largest improvements (see 
-Section~\ref{sec:perfect_exp}). We plan to 
-extend our algorithm to term which is another frequent argument types.
-But term prediction have been so far proven to weak to positively 
+of theorems. We choose it is the most common type and the structure of theorems 
+are complex so they contain enough information to be predicted accurately. 
+We plan to extend our algorithm to term which is another frequent argument 
+types but term synthesis is so far too weak to positively 
 influence the proof search.
@@
 \exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
-\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n\\
-\mathbb{G}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
+\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
+\mathbb{G}^\infty =_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
 \end{align*}
-
-By the base case of these definition, a goal is solved if there exists a tactic 
-applied to this goal that produces no goals.
 \end{definition}
 
-\subsection{Prior tactic policy}
-The estimation through the distance produced by the tactic selectors are hard 
+\subsection{Prior tactic policy}\label{sec:policy}
+Distances between goals produced by the tactic predictor are hard 
 to translate into a prior tactic policy, so we rely only on the order of 
-tactics to estimate the probability with which a tactic should be tried.
+tactics to estimate the probability with which each tactic should be tried.
 Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
@@
 
-\subsection{Prior node evaluation}
+\subsection{Prior node evaluation}\label{sec:evaluation}
 
@@
 influences node selection in MCTS. Essentially, we estimate how 
-interesting is the list of goals of a tactic to determine if we should continue 
-searching in this direction. Theoretically, an interesting list of goal is one 
+interesting is a list of goals produced by a tactic. This way, we can determine 
+if we should continue searching in this direction. Theoretically, an 
+interesting list of goal is one 
 that has the shortest proof, and we could measure the likelihood of being 
@@
 Deep math)}. So to reduce this number, we limit 
-ourselves to feature vectors tested during orthogonalization. We declare a list 
+ourselves to tactics tested during orthogonalization. We declare a list 
 of goals l to be positive, noted $l \in Positive$, if it is has been 
-produced by the winner of an orthogonalization competition. We extend the 
-definition of a $distance$ to a list of goals by considering the union of the 
-features of the goals. The set of the $r$ set of goals closest to a set of 
-goal $l$ is noted $Neighbour(r,l)$. We can define the set of positives and 
-negatives within a radius $r$ of a 
-list of goals $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
+produced by the winner of an orthogonalization competition. The set of the $r$ 
+goals closest to a list of goals $l$ is noted $Neighbour(r,l)$. We can 
+define the set of positives within a 
+radius $r$ of $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
 Positive \rbrace$.
 
-During our experiments, we evaluate output goals in three different manners 
-with the following functions:
+During our experiments, we evaluate output goals with the following function:
 
 \begin{align*}
-PriorEvaluation_{none} (a_i) &= 0 \\
-PriorEvaluation_{productive} (a_i) &= 1\\
 PriorEvaluation_r (a_i) &= \frac{card\ Positive(r,a_i)}{r}\\
@@
 
-For all evaluations functions, if a tactic is not productive then there is no 
-output to evaluate, so the result is a failure and a reward of $0$ will be 
-given for this node. The same thing happens if the node selected is already 
-proven or a descendant of a proven goal. 
-Experimentally, it is preferable not to 
-filter solved (or descendants of solved) nodes before 
-node selection to keep the estimation of the algorithm balanced.
+If a tactic is not productive then there is no output to evaluate, so the 
+result is a failure and a reward of $0$ will be 
+given for this node. Failure also happens if the node selected is already 
+a proven goal or a descendant of a proven goal. Experimentally, it is 
+preferable not 
+to filter solved nodes and descendants of solved nodes before node selection 
+to keep the estimations of the algorithm balanced.
 
@@
 
+mention theorem dependencies for eprover but not for metis.
+
 can be call with frequently with low 
@@
 
-\section{Experimental Evaluation}\label{s:experiments}\todo{TG: introduction + 
-consistent tables}
-
-Disclaimer:
-Conditions of the experiments may vary slightly between different tables 
-because of various factors such 
-as overhead cost of additional features or the fact that they were run on 
-different machines on a local machine or on the server.
-
-\subsection{Methodology} \todo{TG: is copied, update references of sections and
-new algorithm + mention when orthogonalization is done}
-
+\section{Experimental Evaluation}\label{s:experiments}
+\todo{Josef: give the specification} 
+Each re-proving experiment is performed on a single CPU of a server in Prague 
+with the exception of the experiments 
+relying on asynchronous \eprover calls that run on two CPUs.
+
+\subsection{Methodology} 
 The evaluation imitates the construction of the library: For each theorem only 
@@
 \holfour.
-As theorems are proved, their tactical proofs and their statements are recorded 
-and included in the 
-training examples.
+As theorems are proved, their tactical proofs and their statements are 
+recorded and included in the training examples.
 For each theorem we first attempt to run the \tactictoe search with a time 
-limit of 5 seconds,
-before processing the original proof script.
-In this way, the fairness of the
-experiments is guaranteed by construction. 
+limit of 60 seconds, before processing the original proof script.
+In this way, the fairness of the experiments is guaranteed by construction. 
 Only previously declared \sml 
@@
 
-\paragraph{Dataset}
-
-7000+ top-level theorems that were created by . 
-In this evaluation, we do not include the 250 theorems created by
-
-All top-level theorems created by a call to 
-\texttt{store\_thm}, 
-\texttt{Q.store\_thm} or \texttt{maybe\_thm} are considered with the exception 
-of 440 calls that contains a let binding. They are typically harder to re-prove.
-We also ignore the 1175 \texttt{save\_thm} calls. Most of them are easily 
-provable theorems using instantiations or conjuncts of previous 
-theorems.
-
-%\begin{itemize}
-%%\item the , these are 
-%%typically 
-%%harder for \tactictoe as it may not have seen the particular local 
-%%declaration 
-%%before and therefore would have to find an alternative proofs. We also do not 
-%%learn from these proofs because they were explictly local.
-%%\item the 1175 calls to save\_thm. These are typically mostly trivial and 
-%%easy 
-%%theorems.
-%%The remaining 7088 top-level theorems are evaluated in the full scale 
-%%evaluation and the level of difficulty of these problem is assessed by a 
-%%comparison with \eprover.
-%\end{itemize}
-
-Among these types of theorems, every tenth theorems of the first third of the 
-library is evaluated during tuning, which give us a total of 273 theorems to 
-train different parameters on.
-
-\paragraph{Strategy selection}
-Although the training process in each strategy on its own is fair, the 
-selection of the 
-best strategy in Section~\ref{sec:sel_param} should also be considered as a 
-learning 
-process. To ensure the global fairness, the final experiments in 
-Section~\ref{sec:exp_full} 
-runs the best strategy on the full dataset which is about 10 times larger. The
-performance is minimally better on this validation set.
-
-\subsection{Tuning \tactictoe}
-
-\subsubsection{Prediction parameters}\todo{TG: remove this section or mention
-that parameters were optimized in the previous version}
-The first experiment concerns the choice of the right kind of features and 
-feature scoring 
-mechanism. The results are presented in Table~\ref{tab:feature_param}.
-We observe that the higher-order features and the feature of the top logical 
-structure increase minimally the number of problems solved. The attempted 
-length penalty on the total number of features is actually slightly harmful.
-We did not turn on the top feature flag in the next experiments because they 
-are slightly more expensive too compute.
-
+\paragraph{Datasets: optimization and validation}\todo{rephrase to something 
+more 
+coherent}
+All top-level theorems from the standard library are considered with the 
+exception of 440 hard problems (containing a \texttt{let} construction in their 
+proof) and 1175 easy problems (build from \texttt{save\_thm} calls). 
+Therefore, during the full-scale experiments, we evaluate 7164 theorems.
+We use every tenth theorem of the first third of the standard library of 
+them for parameter optimization, which amounts to 273 theorems.
+Although the evaluation of each set of parameters on its own is fair, 
+the selection of the best strategy in Section~\ref{sec:tuning} should also be 
+considered as a learning process. To ensure the global fairness, the final 
+experiments in Section~\ref{sec:full_exp} runs the best strategy on the full 
+dataset which is about 30 times larger.
+
+\subsection{Tuning \tactictoe}~\label{sec:tuning}
+
+Fix parameters: not optimized here. 
+ - Features extraction method.
+ - k-nn distance.
+ - MCTS evaluation radius is set to 10. 
+ - Metis : 16 lemmas.
+ - Eprover: 64 lemmas and 5 seconds timeout 1 asynchronous call
+
+\begin{table}[ht]
+\centering\ra{1.3}
+\small
+\begin{tabular}{llc}
+\toprule
+ Technique & Parameters & Solved \\
+\midrule
+Tactic time out & 0.02s & 155 (56.8\%)\\
+                & 0.05s (default) & 155 (56.8\%)\\
+                & 0.1s & 153 (56.0\%)\\
+\midrule
+Orthogonalization & $radius = 20$ (default) & 155 (56.8\%)\\ 
+                  & $radius = 10$ & 125 (50\%)\\
+                  & $radius = 40$ & 125 (50\%)\\
+\midrule
+Abstraction       & none (default) & 125 (50\%)\\
+                  & $theorems = 8$ & 125 (50\%)\\
+                  & $theorems = 16$ & 125 (50\%)\\
+                  & $theorems = 32$ & 125 (50\%)\\
+\midrule
+MCTS Policy & $c_{policy} = 1/2$ (default) & 125 (50\%)\\
+            & $c_{policy} = 2/3$ & 125 (50\%)\\
+            & $c_{policy} = 1/3$ & 125 (50\%)\\
+\midrule
+MCTS Evaluation & none (default) & 125 (50\%) \\
+           & $c_{exploration} =\sqrt{2}$ & 125 (50\%)\\ 
+           & $c_{exploration} = 1$ & 125 (50\%)\\ 
+           & $c_{exploration} = 2$ & 125 (50\%)\\ 
+\midrule
+ATP integration & none (default) & 125 (50\%)\\
+                & \metis 0.1s & 125 (50\%)\\
+                & \metis 0.2s & 125 (50\%)\\
+                & \metis 0.4s & 125 (50\%)\\ 
+                & \eprover & 125 (50\%)\\
+                & \metis 0.1s + \eprover & 190\\
+\midrule
+All & best parameters (without \eprover) & \\ 
+\bottomrule
+\end{tabular}
+\end{table}
+
+%\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
+%
+%We investigate what are the limits of \tactictoe and what benefits additional 
+%knowledge could provide. To this end, we will modify the set of features to 
+%include features vectors that should not be known in a regular experiment. We  
+%include some vectors from the recorded human proof and perform the proof 
+%search 
+%after the recording instead of before in all other experiments. This 
+%experiment 
+%is not fair and should only be compared to 
+%experiments in the same settings.
+%
+%For clarity, we will name the set of feature vectors that can be recorded from 
+%the current human proof $\mathbb{H}$ and the set of vectors recorded from 
+%previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
+%$\pi_1^1$ the projection on the 
+%first component returning the label of a feature vector. And we define the 
+%function $T$ that takes a tactic and return the list of its tokens.
+%
+%We will progressively add bigger and bigger 
+%subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
+%search success rates. 
+%
+%Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
+%consider:
+%\begin{align*}
+%\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
+%   \rbrace \\
+%\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
+%\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
+%\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
+%\rbrace\\
+%\end{align*}
+%
+%
+%$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
+%prediction for those tactics.
+%
+%
+%\begin{table}[h]
+%\centering\ra{1.3}
+%\small
+%\begin{tabular}{lll}
+%\toprule
+% & Additional features & Solved \\
+%\midrule
+%$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
+%$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
+%$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
+%$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
+%$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
+%$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
+%% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
+%\bottomrule
+%\end{tabular}
+%\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
+%with tactics from the proof of the tested theorem.} 
+%\end{table}
+%\todo{good title for this table Additional features}
+%
+%And we observe in Table~\ref{tab:featue_param} a 
+%slight increase compared to the default \tactictoe. We then included all 
+%vectors that contains tactics that can be rebuild from tokens of
+%tactics in our database. 
+%Our search algorithm proved two proofs less that if it had perfect prediction 
+%so it means that our prediction algorithm (policy) is doing a good  job at 
+%least in this early settings where proofs are easy. 
+%Suprisingly, being able to recombine tokens of tactics from the database is 
+%barely sufficient to prove more than half of the theorems in this dataset. The 
+%Therefore, additionally including in tactics theorems and terms that did not 
+%appear in previous tactics is essential. Creating from scratch other arguments 
+%may not help as much. This is why we restricted ourselves to predicting 
+%arguments of theorem list type and the term type during tactic synthesis. 
+%Finally our proof search algorithm sometimes fails even if it knows all the 
+%tactic necessary. The two main reasons are that either a crucial tactic 
+%exceeds 
+%its timeout or that the human proof is too deep to be rediscovered (more than 
+%10 steps long).In all those experiments orthogonalization is performed with 
+%radius 20.
+%
+%\subsubsection{Orthogonalization Abstraction}
+%Here, we test the benefit of different kind of tactic abstraction.
+%All experiments in this table are using orthogonalization.
 %
@@
 %\small
-%\begin{tabular}{llll}
+%\begin{tabular}{ll}
 %\toprule
-% ID & Learning parameter & Solved & $U(D_1)$ \\
+% Technique & Solved \\
 %\midrule
-% $D_0$ & $codist_1$ (length penalty) & 172 (20.0\%) & 5 \\
-% $D_1$ & $codist_0$ (default) & 179 (20.8\%) & 0 \\
-% $D_2$ & no top features & 175 (20.3\%) & 18 \\
-% $D_3$ & no higher-order features & 178 (20.7\%) & 8 \\
+%% holyhammer & 79 \\
+%  & 122\\
+% Orthogonalization & 129\\
+%% ortho + term & 126\\    
+% Orthogonalization + Abstraction & 160\\
+%% ortho + \metis & 162\\
+% %ortho + metis + thmortho & 155\\
+% %ortho + metis + stacortho & 160\\
+%% ortho + \metis + theorem list & 175\\ 
 %\bottomrule
 %\end{tabular}
-%\caption{\label{tab:feature_param}Success rate of strategies with different 
-%learning parameters on the training set.}
+%\caption{\label{tab:cfot_param} 
+% Effect of orthogonalization and abstraction}
+%\end{table}
+%We want to know what the optimal number of predictions for tactics different 
+%from \metis since \metis has a special status in our algorithm is handled 
+%independently.
+
+%\subsubsection{ATP integration}
+%
+%
+%\todo{paragraph is copied}
+%The best setting for \metis was to be in previous experiments to be 16
+%theorems an The best strategy which does not rely on the ``small hammer'' 
+%approach $D_9$ 
+%will be called \tactictoe(NH) (for no ``small hammer'') in the remaining part 
+%of the paper, and the best strategy relying on the approach, $D_{19}$, will be 
+%referred to as \tactictoe(SH) (``small hammer'').
+%
+%\begin{table}[h!]
+%\centering\ra{1.3}
+%\small
+%\begin{tabular}{ll}
+%\toprule
+%  & Solved \\
+%\midrule
+%   \eprover &  106 \\
+%   \tactictoe &  139 \\
+%   \tactictoe + \metis & 182 \\
+%   \tactictoe + \eprover & 182 \\
+%   %x tactictoe + async holyhammer (10 thread) & 183 \\
+%   \tactictoe + \metis + \eprover & 190 \\
+%  
+%\bottomrule
+%\end{tabular}
+%\caption{Experiments with a timeout of 60 seconds \label{tab:param}}
 %\end{table}
 %
-%\subsubsection{Search parameters}
-%
-%\paragraph{tactic timeout}
-%\begin{table}[t]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{llll}
-%\toprule
-% ID & Searching parameter & Solved & $U(D_{9})$ \\
-%\midrule
-% $D_1$ & $codist_0$ (default)      & 179 (20.8\%) & 10 \\
-% $D_4$ & tactic timeout 0.004 sec & 175 (20.3\%) & 9 \\
-% $D_5$ & tactic timeout 0.1 sec   & 178 (20.7\%) & 10 \\
-%\bottomrule
-%\end{tabular}
-% \caption{\label{tab:search_param}Success rate of strategies with different 
-% search parameters on the training set.}
-%\end{table}
-
-\subsubsection{MCTS parameters: policy and evaluation}
-
-Width did include failing tactics in this experiments. $c_{exploration}$ = 1.
-$E_{none}$.
-
-\begin{table}[ht]
-\centering\ra{1.3}
-\small
-\begin{tabular}{ll}
-\toprule
- $c_{policy}$ & Solved \\
-\midrule
- $2/3$ & 120\\
- $1/2$ & 125\\
- $1/3$ & 117\\
-% 1/2 & Productive & No & $E_{none}$ & 1 & 120\\
-% 1/2 & Productive & No & $E_{productive}$ & 1 & 127\\
-\bottomrule
-\end{tabular}
-\caption{\label{tab:cost_param} 
-%on a local machine
-}
-\end{table}
-
-%This experiments were performed without the square root in the policy 
-%function.
-
-This experiments are performed with orthogonalization on and policy coefficient 
-of $1/2$.
-
-\begin{table}[ht]
-\centering\ra{1.3}
-\small
-\begin{tabular}{lll}
-\toprule
- PriorEvaluation & $c_{exploration}$ & Solved \\
-\midrule
-% 1/2 & All & No & $E_{none}$ & 1 & 122\\
-% 1/2 & All & Yes & $E_{none}$ & 1 & 128\\
- none & 1 & 128\\
- productive & 1 & 134\\ 
- & 4 & 130\\
- radius = 1 & 1 & 127\\ 
- & 4 & 126\\ 
- radius = 10 & 1 & 136\\ 
- & 4 & 137\\ 
-\bottomrule
-\end{tabular}
-\caption{\label{tab:search_parameters}}
-%on the server
-\end{table}
-
-We settled for the first parameters where an evaluation function based on 
-machine learning was beneficial. It is highly likely than better parameters for 
-the policy, radius of evaluation and exploration coefficient can be found with 
-further investigation.
-
-\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
-
-We investigate what are the limits of \tactictoe and what benefits additional 
-knowledge could provide. To this end, we will modify the set of features to 
-include features vectors that should not be known in a regular experiment. We  
-include some vectors from the recorded human proof and perform the proof search 
-after the recording instead of before in all other experiments. This experiment 
-is not fair and should only be compared to 
-experiments in the same settings.
-
-For clarity, we will name the set of feature vectors that can be recorded from 
-the current human proof $\mathbb{H}$ and the set of vectors recorded from 
-previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
-$\pi_1^1$ the projection on the 
-first component returning the label of a feature vector. And we define the 
-function $T$ that takes a tactic and return the list of its tokens.
-
-We will progressively add bigger and bigger 
-subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
-search success rates.
-
-Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we consider:
-\begin{align*}
-\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
-   \rbrace \\
-\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
-\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
-\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
-\rbrace\\
-\end{align*}
-
-$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
-prediction for those tactics.
-
-\begin{table}[h]
-\centering\ra{1.3}
-\small
-\begin{tabular}{lll}
-\toprule
- & Additional features & Solved \\
-\midrule
-$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
-$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
-$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
-$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
-$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
-$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
-% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
-\bottomrule
-\end{tabular}
-\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
-with tactics from the proof of the tested theorem.} 
-\end{table}
-\todo{good title for this table Additional features}
-
-And we observe in Table~\ref{tab:featue_param} a 
-slight increase compared to the default \tactictoe. We then included all 
-vectors that contains tactics that can be rebuild from tokens of
-tactics in our database. 
-Our search algorithm proved two proofs less that if it had perfect prediction 
-so it means that our prediction algorithm (policy) is doing a good  job at 
-least in this early settings where proofs are easy. 
-Suprisingly, being able to recombine tokens of tactics from the database is 
-barely sufficient to prove more than half of the theorems in this dataset. The 
-Therefore, additionally including in tactics theorems and terms that did not 
-appear in previous tactics is essential. Creating from scratch other arguments 
-may not help as much. This is why we restricted ourselves to predicting 
-arguments of theorem list type and the term type during tactic synthesis. 
-Finally our proof search algorithm sometimes fails even if it knows all the 
-tactic necessary. The two main reasons are that either a crucial tactic exceeds 
-its timeout or that the human proof is too deep to be rediscovered (more than 
-10 steps long).In all those experiments orthogonalization is performed with 
-radius 20.
-
-\subsubsection{Orthogonalization Abstraction}
-Here, we test the benefit of different kind of tactic abstraction.
-All experiments in this table are using orthogonalization.
-
-\begin{table}[t]
-\centering\ra{1.3}
-\small
-\begin{tabular}{ll}
-\toprule
- Technique & Solved \\
-\midrule
-% holyhammer & 79 \\
-  & 122\\
- Orthogonalization & 129\\
-% ortho + term & 126\\    
- Orthogonalization + Abstraction & 160\\
-% ortho + \metis & 162\\
- %ortho + metis + thmortho & 155\\
- %ortho + metis + stacortho & 160\\
-% ortho + \metis + theorem list & 175\\ 
-\bottomrule
-\end{tabular}
-\caption{\label{tab:cfot_param} 
- Effect of orthogonalization and abstraction}
-\end{table}
-We want to know what the optimal number of predictions for tactics different 
-from \metis since \metis has a special status in our algorithm is handled 
-independently.
-
-\subsubsection{ATP integration}
-
-\todo{paragraph is copied}
-The best setting for \metis was to be in previous experiments to be 16
-theorems an The best strategy which does not rely on the ``small hammer'' 
-approach $D_9$ 
-will be called \tactictoe(NH) (for no ``small hammer'') in the remaining part 
-of the paper, and the best strategy relying on the approach, $D_{19}$, will be 
-referred to as \tactictoe(SH) (``small hammer'').
-
-\begin{table}[h!]
-\centering\ra{1.3}
-\small
-\begin{tabular}{ll}
-\toprule
-  & Solved \\
-\midrule
-   \eprover &  106 \\
-   \tactictoe &  139 \\
-   \tactictoe + \metis & 182 \\
-   \tactictoe + \eprover & 182 \\
-   %x tactictoe + async holyhammer (10 thread) & 183 \\
-   \tactictoe + \metis + \eprover & 190 \\
-
-\bottomrule
-\end{tabular}
-\caption{Experiments with a timeout of 60 seconds \label{tab:param}}
-\end{table}
-
-Slight error in this experiments were eprover predictions have been made from 
-the preselected 500 theorems instead of the full library. But even with the 
-added overhead, we already see 
-the benefit of using stronger external prover than \metis.
-
-\subsection{Full-scale experiment}
+%Slight error in this experiments were eprover predictions have been made from 
+%the preselected 500 theorems instead of the full library. But even with the 
+%added overhead, we already see 
+%the benefit of using stronger external prover than \metis.
+
+\subsection{Full-scale experiment}~\label{sec:full_exp}
 
@@
 
-\eprover is not called by \tactictoe for fair comparison.
+The best \tactictoe parameters are chosen for the full-scale evaluation. Yet,
+\eprover calls are not included in the best \tactictoe strategy for fairer 
+comparison.
 
@@
 \small
-\begin{tabular}{lll}
+\begin{tabular}{llc}
 \toprule
-  & & 7164 theorems \\
+  & Parameters & Solved \\
 \midrule
-  & \eprover   & 2472 (34.5\%)\\ 
-  & \tactictoe & 4741 (66.2\%)\\
-  & combined   & 4946 (69.0\%)\\
+   \eprover   & 128 premises & 2472 (34.5\%)\\ 
+   \tactictoe & best parameters without \eprover & 4741 (66.2\%)\\
+   combined   &              & 4946 (69.0\%)\\
 \bottomrule
 \end{tabular}
-\caption{Experiments with a timeout of 60 seconds \label{tab:_param}}
+\caption{Evaluation on 7164 top-level theorems of the \holfour standard library 
+\label{tab:_param}}
 \end{table}

### 2018-01-28 tactictoe_jv.tex
predictions: almost done

@@
 \subsection{Similarity}\label{sec:predictions}
-
 We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$ 
@@
 We additionally raise the weights to the sixth power giving even more credence 
-to rare features which was found experimentally optimal in \cite{cezary_cite}.
+to rare features which was found experimentally optimal in \cite{}.\todo{CK: 
+citation}
 
@@
 feature sets in the first scoring function.
-\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_o \cap 
-f_p}{\text{tfidf}(f)^{6}}}\]
-\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_o \cap 
-f_p}{\text{tfidf}(f)^{6}}}}
-{ln (e + card(f_o) + card(f_p))}\]
+\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap 
+f_1}{\text{tfidf}(f)^{6}}}\]
+\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap 
+f_1}{\text{tfidf}(f)^{6}}}}
+{ln (e + card(f_0) + card(f_1))}\]
 

### 2018-01-28 tactictoe_jv.tex
predictions: almost done

@@
 $f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features 
-are, the more similar two goals should be. This can be realized by 
-calculating TF-IDF weights for each feature~\cite{Jones72astatistical}.
+are, the more similar two goals should be. The relative rarity of each feature 
+can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
 We additionally raise the weights to the sixth power giving even more credence 
@@
 Let $\mathfrak{T}$ be the set of recorded theorems.
-The dependencies $D_\infty$ of a theorem $t$ are the set of theorems in 
-$\mathfrak{T}$ appearing in the proof $t$ as recorded in \cite{hh4h4}.
+The dependencies of a theorem $t$ are the set of theorems in 
+$\mathfrak{T}$ appearing in the proof of $t$ as recorded in 
+\cite{hh4h4}.\todo{cite}
 Contrarily to the dependency relation in tactic-goal pairs, this relation is 

### 2018-01-28 tactictoe_jv.tex
predictions: wip

@@
 
-\begin{definition}(Dependencies of tactic-goal pairs)\\
+\begin{definition}(Dependencies of a tactic-goal pair)\\
 Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
@@
 \cite{hh4h4}.\todo{cite} 
-Contrarily to the dependency relation in tactic-goal pairs, this relation is 
-not a transitive closure.
 \end{definition}
+
+\begin{remark}
+We can note that the relation chosen for tactic-goal pairs is a transitive 
+closure, which is not the case for theorem dependencies.
+\end{remark}
 

### 2018-01-28 tactictoe_jv.tex
putting back: increasing TacticToe's knowledge section

@@
 application of the MCTS algorithm. That is why we call it a tree although this 
-we do not specify it in this definition.
+is not specified in this definition.
 \end{definition}
@@
 
-%\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
-%
-%We investigate what are the limits of \tactictoe and what benefits additional 
-%knowledge could provide. To this end, we will modify the set of features to 
-%include features vectors that should not be known in a regular experiment. We  
-%include some vectors from the recorded human proof and perform the proof 
-%search 
-%after the recording instead of before in all other experiments. This 
-%experiment 
-%is not fair and should only be compared to 
-%experiments in the same settings.
-%
-%For clarity, we will name the set of feature vectors that can be recorded from 
-%the current human proof $\mathbb{H}$ and the set of vectors recorded from 
-%previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
-%$\pi_1^1$ the projection on the 
-%first component returning the label of a feature vector. And we define the 
-%function $T$ that takes a tactic and return the list of its tokens.
-%
-%We will progressively add bigger and bigger 
-%subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
-%search success rates. 
-%
-%Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
-%consider:
-%\begin{align*}
-%\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
-%   \rbrace \\
-%\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
-%\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
-%\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
-%\rbrace\\
-%\end{align*}
-%
-%
-%$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
-%prediction for those tactics.
-%
-%
-%\begin{table}[h]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{lll}
-%\toprule
-% & Additional features & Solved \\
-%\midrule
-%$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
-%$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
-%$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
-%$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
-%$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
-%$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
-%% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
-%\bottomrule
-%\end{tabular}
-%\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
-%with tactics from the proof of the tested theorem.} 
-%\end{table}
-%\todo{good title for this table Additional features}
-%
-%And we observe in Table~\ref{tab:featue_param} a 
-%slight increase compared to the default \tactictoe. We then included all 
-%vectors that contains tactics that can be rebuild from tokens of
-%tactics in our database. 
-%Our search algorithm proved two proofs less that if it had perfect prediction 
-%so it means that our prediction algorithm (policy) is doing a good  job at 
-%least in this early settings where proofs are easy. 
-%Suprisingly, being able to recombine tokens of tactics from the database is 
-%barely sufficient to prove more than half of the theorems in this dataset. The 
-%Therefore, additionally including in tactics theorems and terms that did not 
-%appear in previous tactics is essential. Creating from scratch other arguments 
-%may not help as much. This is why we restricted ourselves to predicting 
-%arguments of theorem list type and the term type during tactic synthesis. 
-%Finally our proof search algorithm sometimes fails even if it knows all the 
-%tactic necessary. The two main reasons are that either a crucial tactic 
-%exceeds 
-%its timeout or that the human proof is too deep to be rediscovered (more than 
-%10 steps long).In all those experiments orthogonalization is performed with 
-%radius 20.
+\subsection{Increasing \tactictoe'sknowledge}\label{sec:perfect_exp}
+
+\todo{decide to keep this section or not}
+
+We investigate what are the limits of \tactictoe and what benefits additional 
+knowledge could provide. To this end, we will modify the set of features to 
+include features vectors that should not be known in a regular experiment. We  
+include some vectors from the recorded human proof and perform the proof 
+search 
+after the recording instead of before in all other experiments. This 
+experiment 
+is not fair and should only be compared to 
+experiments in the same settings.
+
+For clarity, we will name the set of feature vectors that can be recorded from 
+the current human proof $\mathbb{H}$ and the set of vectors recorded from 
+previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
+$\pi_1^1$ the projection on the 
+first component returning the label of a feature vector. And we define the 
+function $T$ that takes a tactic and return the list of its tokens.
+
+We will progressively add bigger and bigger 
+subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
+search success rates.
+
+Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
+consider:
+\begin{align*}
+\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
+   \rbrace \\
+\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
+\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
+\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
+\rbrace\\
+\end{align*}
+
+$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
+prediction for those tactics.
+
+\begin{table}[h]
+\centering\ra{1.3}
+\small
+\begin{tabular}{lll}
+\toprule
+ & Additional features & Solved \\
+\midrule
+$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
+$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
+$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
+$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
+$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
+$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
+% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
+\bottomrule
+\end{tabular}
+\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
+with tactics from the proof of the tested theorem. The experiments was run with 
+a time out of 5sec on an older version of \tactictoe, as can be seen in the low 
+success rate of the
+default strategy.} 
+\end{table}
+
+And we observe in Table~\ref{tab:featue_param} a 
+slight increase compared to the default \tactictoe. We then included all 
+vectors that contains tactics that can be rebuild from tokens of
+tactics in our database. 
+Our search algorithm proved two proofs less that if it had perfect prediction 
+so it means that our prediction algorithm (policy) is doing a good  job at 
+least in this early settings where proofs are easy. 
+Suprisingly, being able to recombine tokens of tactics from the database is 
+barely sufficient to prove more than half of the theorems in this dataset. The 
+Therefore, additionally including in tactics theorems and terms that did not 
+appear in previous tactics is essential. Creating from scratch other arguments 
+may not help as much. This is why we restricted ourselves to predicting 
+arguments of theorem list type and the term type during tactic synthesis. 
+Finally our proof search algorithm sometimes fails even if it knows all the 
+tactic necessary. The two main reasons are that either a crucial tactic 
+exceeds 
+its timeout or that the human proof is too deep to be rediscovered (more than 
+10 steps long).In all those experiments orthogonalization is performed with 
+radius 20.
 %

### 2018-01-28 tactictoe_jv.tex
prediction: first final version

@@
 
+\section{Preliminaries: Search Tree}\label{sec:prelim}
+
+In order to be able to describe the different components of \tactictoe, 
+we need to give to describe what a search tree is, what are its elements and 
+how they are organized relative to one other. The following description is 
+static and the evolution of the search tree toward a proof is given in 
+Section~\ref{sec:proofsearch}.
+
+\begin{definition}(search tree)\\
+A search tree $\mathfrak{T}$ is a sixtuple 
+$(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$ 
+that respects the following conditions:
+\begin{itemize}
+\item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals 
+ and $\mathbb{A}$ is a set of nodes. These sets contains only objects 
+ appearing in the search tree.
+\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$. It takes a 
+goal and return the set of tactics already applied to this goal.
+\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$. It takes a 
+node a return the set of goals in this node.
+\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
+\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
+$\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the 
+application of the tactic to the goal, if this tactic has been applied to this 
+goal.
+\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
+\end{itemize}
+
+Additional constraints such as acyclicity and the fact that there is exactly 
+one root node are properties of this graph that are preserved during the 
+application of the MCTS algorithm. That is why we call it a tree although this 
+is not specified in this definition.
+\end{definition}
+
+If no explicit order is given, we assume that any set is equipped with an 
+arbitrary order. In this way, we can access the nth member of set $s$ by 
+writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written $g_j^i$.
+
+\begin{figure}{}
+\begin{center}
+\begin{tikzpicture}[node distance=0.7cm]
+\node [tcircle] (0) {$g_0^i$};
+\node [right of=0] (0r) {.\ .\ .};
+\node [tcircle,right of=0r] (0rr) {$g_0^n$};
+\node [left of=0] (0l) {.\ .\ .};
+\node [tcircle,left of=0l] (0ll) {$g_0^1$};
+\node [draw,fit=(0rr) (0ll)] {}; 
+\node [above of=0,node distance=2cm] (2) {.\ .\ .};
+\node [left of=2, tcircle] (2l) {$g_j^1$};
+\node [tcircle,right of=2] (2r) {$g_j^p$};
+\node [draw, fit=(2l) (2r)] (2b) {};
+
+\node [tcircle, left of=2l, node distance=3cm] (1r) {$g_1^q$};
+\node [left of=1r] (1) {.\ .\ .};
+\node [tcircle, left of=1] (1l) {$g_1^1$};
+\node [draw, fit=(1l) (1r)] (1b) {};
+
+\node [right of=2r,node distance=3cm] (3l) {$\phantom{g_3^p}$};
+\node [right of=3l] (3) {\phantom{.\ .\ .}};
+\node [right of=3] (3r) {$\phantom{g_3^p}$};
+\node [fit=(3l) (3r)] (3b) {};
+
+\node [fit=(1r) (2l)] (1m) {}; 
+\node [fit=(2l) (3r)] (2m) {};
+
+\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
+\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
+\draw[-to,dotted,thick] (0) to (1m);
+\draw[-to,dotted,thick] (0) to (2m);
+\draw[-to,dotted,thick] (0) to (3b);
+\end{tikzpicture}
+\end{center}
+\caption{A node of a search tree and the children of one of 
+its goal.}
+\label{fig:choice}
+\end{figure}
+
+Starting from the bottom parent node, we can construct children of a goal of 
+this node by choosing a goal $g_i^0$ on this node and choosing tactics 
+$t_1,t_2,\ldots$ to 
+apply to this goal. Here the result produces new nodes each containing a list 
+of goals to be proven. In each node of the proof tree, a goal can either be 
+solved, open or pending (see Definition~\ref{def:solved}).
+
+\begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal, 
+pending goal)\\
+Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
+$\mathbb{A}^\infty$ and 
+the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:
+
+\begin{align*}
+\mathbb{A}^{0} &=_{def} 
+\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\ 
+\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
+\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
+\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\ 
+\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
+\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
+\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
+\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
+\mathbb{G}^\infty =_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
+\end{align*}
+
+An open goal is an unsolved goal that was tried at least once by a tactic. The 
+set of open goals can be formally defined as:  
+\[\lbrace g \in G\ |\  T(g) \neq \emptyset \wedge g \notin \mathbb{G}^\infty 
+\rbrace\]
+
+The set of pending goals is the remaining set of goals in the tree. They are 
+goal that were produced by tactics that were not explored yet. This set can 
+also be also defined by:
+\[\lbrace g \in G\ |\  T(g) = \emptyset \rbrace\]
+\end{definition}
+
 \section{Prediction}\label{s:prediction}
@@
 To start with, we build a database of tactics consisting
-of (tactic,goal) pairs recorded from successful human proofs (see 
+of tactic-goal pairs recorded from successful tactic applications in human 
+proofs (see 
 Section~\ref{sec:recording}).
 Then, during proof search, we order recorded goals according to their 
-similarity with a targeted goal $g$. This induces through the recorded pairs an 
-order on tactics. The intuition is that tactics that have been successful on 
-goals similar $g$ should be tried first. 
+similarity with a targeted goal $g$ (usually the first pending goal of a node). 
+This induces 
+through the recorded pairs an 
+order on tactics. The intuition is that tactics which have been successful on 
+goals similar to $g$ should be tried first, since they are more likely to lead 
+to a proof. 
 This predicted tactic order is translated into a prior policy for our MCTS 
 algorithm (see Section~\ref{sec:policy}).
-Additionally, tactic selection allows us to improve the quality of the database 
+Tactic selection is also used to improve the quality of the database 
 of tactics during orthogonalization (see Section~\ref{sec:ortho}).
@@
 \paragraph{Theorems as Argument of Tactics}
-We first collect a set of theorems from the \holfour theorem database and local 
-theorem values, for our theorem predictor to select from.
-Predicted tactics during proof search and orthogonalization may contain 
-abstracted tactics where list of theorems have been erased (see Section 
+We first collect a set of theorems for our theorem predictor to select from.
+It includes the \holfour theorem database and theorems from the local namespace.
+Predicted tactics during proof search and orthogonalization may include
+tactics where arguments (list of theorems) have been erased (see Section 
 \ref{sec:synthesis}).
@@
 \paragraph{List of Goals as Output of Tactics}
-A dataset of list of goals is compiled during orthogonalization (see 
+A dataset of tactic outputs is compiled during orthogonalization (see 
 Section~\ref{sec:evaluation}).
 This set is separated into positive and negative examples.
-List of goals that subsumes the result of the original tactic on a 
+List of goals that subsumes the result of an original tactic on a 
 targeted goal are considered positive. During proof search, given a list of 
-goals $l$ created by a tactic, we select lists of goals that are most similar 
-to $l$.
-The ratio of positive examples in the selection gives us a prior node 
-evaluation for MCTS.
+goals $l$ created by a tactic, we select a set of lists of goals that are most 
+similar to $l$. The ratio of positive examples in the selection gives us a 
+prior node evaluation for MCTS.
 
 \subsection{Feature Extraction}\label{sec:features}
-The four objects considered are different representation of mathematical 
-formulas. Therefore, we start by describing which features we extract from the 
-simplest representation terms. From there, we extend the extraction mechanism 
+The three objects considered are different representation of mathematical 
+formulas. Therefore, we start by describing which features we extract can 
+extract terms. From there, we extend the extraction mechanism 
 to goals, theorems and lists of goals. Duplicated features are always removed 
-so each object has an associated set of features.
+so that each object has an associated set of features.
 
@@
 features of the assumptions and features of the conclusion by adding a 
-different tag to elements of each set. For a list of goals, we simply construct 
-the union of the features of each goal 
-in the list.
+different tag to elements of each set. Features of a list of goals are
+the union of the features of each goal in the list.
 
@@
 can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
-We additionally raise the weights to the sixth power giving even more credence 
+We additionally modify the weight by raising them to the sixth power giving 
+even more credence 
 to rare features which was found experimentally optimal in \cite{}.\todo{CK: 
@@
 
-The first co-distance $sim_1$ adds a modified weight of each shared 
+The first co-distance $sim_1$ adds the weight of each shared 
 features to compute the total score.
@@
 of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems 
-and $sim_2$ for list of goals. The similarity $sim_1$ is used for predicting a 
-tactic via its associated goal because if a
-tactic is effective on large goal $g$ then it is often suitable for a goal $g'$ 
-containing sub-formulas of $g$. The same monotonicity heuristic can justify the 
-use of $sim_1$ in theorems. Indeed in practice, a large theorem is often a 
-conjunction of small theorems from the same domain. And if a conjunction 
-contains a formula related to a problem, then the other conjuncts 
-from the same domain may also help.
+and $sim_2$ for list of goals. The reason why $sim_1$ is adapted for
+predicting tactics is that tactics effective on large goal $g$ are often also  
+suitable for a goal made of sub-formulas of $g$. The same monotonicity 
+heuristic can justify the use of $sim_1$ in theorems. Indeed in practice, a 
+large theorem is often a conjunction of theorems from the same domain. 
+And if a conjunction contains a formula related to a problem, then the other 
+conjuncts from the same domain may also help to solve it.
 
@@
 conjecture $c$, we preselect 500 tactic-goal pairs and theorems. 
-Preselection for list of goals is induced by the preselection of the input 
-goals in the tactic-goal pairs. Those will 
-be the only objects available to \tactictoe for proving $c$.
+Preselection for list of goals is induced by the preselection of input 
+goals in tactic-goal pairs. Those will be the only objects available to 
+\tactictoe's search algorithm for proving $c$.
 
@@
 anticipate the production of diverse goals, our prediction during preselection
-take into account dependencies between objects of the same dataset.
+takes into account dependencies between objects of the same dataset.
 The dependence relation is assymetric. Through the relation, each object has 
@@
 $\mathfrak{T}$ appearing in the proof of $t$ as recorded in 
-\cite{hh4h4}.\todo{cite} 
+\cite{tgck-cpp15}.
 \end{definition}
@@
 \begin{remark}
-We can note that the relation chosen for tactic-goal pairs is a transitive 
+The relation chosen for tactic-goal pairs is a transitive 
 closure, which is not the case for theorem dependencies.
@@
 To preselect a list of goal $l$, we consider the tactic input $g$ that was 
-at the origin of the production of this list. If and only if there exists a 
-tactic $t$ such as $(t,g)$ is in the preselected tactic-goal pairs, the 
-list of goals $l$ is preselected. This has the desirable effect that
-negative and positive examples created during one orthogonalization 
-step are chosen or discarded together. \todo{problematic forward reference?}
+at the origin of the production of $l$. The list of goals $l$ is preselected if 
+and only if $g$ appears in the preselected tactic-goal pairs. This has the 
+desirable effect that negative and positive examples derived from the same 
+input goal are chosen or discarded together.
 
@@
 direct supervised learning in the case of the policy and via the data 
-accumulated during orthogonalization for the evaluation. This MCTS do not rely 
-on roll outs. The 
+accumulated during orthogonalization for the evaluation. Therefore, this MCTS 
+algorithm does not need to rely on roll outs for evaluation. The 
 policy and evaluation are estimated by the simple supervised k-nn algorithm 
@@
 
-So, we first give a static description of the proof tree and its component.
-Then we explain how we compute the evaluation and policy functions from learned 
-data. And finally, we show how the two MCTS step, node selection and node 
-extension changes the proof tree data structures.
-Intuitively, MCTS offers a good trade-off between exploitation and exploration. 
-It means that the algorithm will exploit deeper more promises branches but 
-leaving enough time for exploring less likely alternatives.
-
-\subsection{Search Tree}
-
-\begin{definition}(search tree)\\
-A search tree $\mathfrak{T}$ is a sixtuple 
-$(\mathbb{A},\mathbb{G},\mathbb{T},A,G,T)$ 
-that respects the following conditions:
-\begin{itemize}
-\item $\mathbb{A}$ is a set of nodes, $\mathbb{G}$ is a set of goals 
- and $\mathbb{T}$ is a set of tactics.
-\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
-$\mathbb{A}$.
-\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$.
-\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$.
-\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
-\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
-\end{itemize}
-
-Additional constraints such as acyclicity and the fact that there is exactly 
-one root node are properties of this graph that are preserved during the 
-application of the MCTS algorithm. That is why we call it a tree although this 
-is not specified in this definition.
-\end{definition}
-
-Unless specified otherwise, we assume that any set is numbered arbitrarily and 
-that we can access a nth member of set 
-$s$ by writing $s_n$. For readability in the next figure $(G(a_j))_i$ is 
-written $g_j^i$.
-
-\begin{figure}{}
-\begin{center}
-\begin{tikzpicture}[node distance=0.7cm]
-\node [tcircle] (0) {$g_0^i$};
-\node [right of=0] (0r) {.\ .\ .};
-\node [tcircle,right of=0r] (0rr) {$g_0^n$};
-\node [left of=0] (0l) {.\ .\ .};
-\node [tcircle,left of=0l] (0ll) {$g_0^1$};
-\node [draw,fit=(0rr) (0ll)] {}; 
-\node [above of=0,node distance=2cm] (2) {.\ .\ .};
-\node [left of=2, tcircle] (2l) {$g_j^1$};
-\node [tcircle,right of=2] (2r) {$g_j^p$};
-\node [draw, fit=(2l) (2r)] (2b) {};
-
-\node [tcircle, left of=2l, node distance=3cm] (1r) {$g_1^q$};
-\node [left of=1r] (1) {.\ .\ .};
-\node [tcircle, left of=1] (1l) {$g_1^1$};
-\node [draw, fit=(1l) (1r)] (1b) {};
-
-\node [right of=2r,node distance=3cm] (3l) {$\phantom{g_3^p}$};
-\node [right of=3l] (3) {\phantom{.\ .\ .}};
-\node [right of=3] (3r) {$\phantom{g_3^p}$};
-\node [fit=(3l) (3r)] (3b) {};
-
-\node [fit=(1r) (2l)] (1m) {}; 
-\node [fit=(2l) (3r)] (2m) {};
-
-\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
-\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
-\draw[-to,dotted,thick] (0) to (1m);
-\draw[-to,dotted,thick] (0) to (2m);
-\draw[-to,dotted,thick] (0) to (3b);
-\end{tikzpicture}
-\end{center}
-\caption{A node of a search tree and the children of one of 
-its goal.}
-\label{fig:choice}
-\end{figure}
-
-Starting from the bottom parent node, we can construct children of a goal of 
-this node by 
-choosing a goal $g_i^0$ on this node and choosing tactics $t_1,t_2,\ldots$ to 
-apply to this goal. Here the result produces new nodes each containing a list 
-of goals to be proven.
-
-In each node can either be solved, open or pending. The definition of a solved 
-goal is expressed through mutual recursion with a 
-solved node and is given in Definition~\ref{def:solved_goal}. An open goal is 
-goal that 
-was tried at least once by a tactic. In our implementation we have only one 
-open goals per node at once. A goal that was not explored yet is called a 
-pending goal.
-
-\begin{definition}(Solved node and solved goal)
-Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
-$\mathbb{A}^\infty$ and 
-the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:
-
-\begin{align*}
-\mathbb{A}^{0} &=_{def} 
-\lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\ 
-\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
-\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
-\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\ 
-\forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
-\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\ 
-\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
-\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
-\mathbb{G}^\infty =_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
-\end{align*}
-\end{definition}
+%In this section, we first give a static description of the proof tree and its 
+%component. 
+We first explain how we compute the evaluation and policy functions 
+from learned data. 
+And then during the application of the MCTS algorithm, we show how 
+node selection steps and node extension steps make the search tree
+(described in Section~\ref{sec:prelim}) evolve.
+A characteristic of the MCTS algorithm is that it offers a good trade-off 
+between exploitation and exploration. This means that the algorithm searches
+deeper more promises branches and leaves enough time for the exploration 
+of less likely alternatives.
 
@@
 
-\[PriorPolicy(t_{i}) = (1 - c_{policy})^{i} * c_{policy}\]
+\[PriorPolicy(t_i) = PP(i) = (1 - c_{policy})^{i} * c_{policy}\]
 
@@
 
-Through node extension steps the tree grows and  the number of paths to explore 
-increases. To decide which node to extend next, we proceed through the node 
-selection algorithm. Starting from the root node, we repeatedly choses the 
-children with the highest value. To allow the algorithm to select nodes that 
-are not leafs, We also assign a value to the possibility of stopping the 
-descent through the tree at a particular node and trying the next tactic in the 
-predictions. This value has an exploitation equal to 0 since it was never 
-explored.
-
-The value of a node with respect to a search tree, prior policies and prior 
-evaluation is determined by the 
-following computation:
-
-\[Value(a_i) = Exploitation(a_i) + c_{exploration} * Exploration(a_i)\]
+Through node extension steps the search tree grows and the number of paths to 
+explore increases. To decide which node to extend next, we proceed through the 
+node selection algorithm. Starting from the root node, we repeatedly chose the 
+children with the highest value.
+
+The value of a $a_i$ with respect to a search tree is determined by:
+\[Value(a_i) = CurrentEvaluation(a_i) + c_{exploration} * Exploration(a_i)\]
 
@@
 The exploitation term is the average of the evaluation of 
-all nodes descendant of $a_i$ (itself included) in the current search tree. It 
+all nodes descendant of $a_i$ (itself included) in the current search tree.
+This evaluation also include failures of the selection process.
+
+ It 
 is also called the current evaluation of $a_i$.   
-\[Exploitation(a_i) = CurrentEvaluation(a_i) = 
-  \average_{b \in Descendant(a_i)} {PriorEvaluation(b)}\]
+\[CurrentEvaluation(a_i) = 
+  \sum_{b \in Desc(a_i)} \frac{PriorEvaluation(b)} {(card Desc(a_i) + card 
+  (Failure(a_i))}\]
+
+where $Failure(a_i)$ is the number of run that went through $a_i$ that ended in 
+a failure.
+
+To allow the algorithm to select nodes that 
+are not leafs, we also assign a value to the possibility of stopping during the 
+descent of the tree. This value has an exploitation equal to 0 since it was 
+never explored and is given a prior policy equal to $PP(n)$ where $n$ is the 
+number of 
+tactics already tried on the open goal if there is one and $PP(0)$ if there is 
+none.
 
@@
 Once a node is selected, we applied the best not-yet-applied tactics according 
-to the prediction order to the open goal this node. If the tactic is productive 
-then the we create a new 
-node and calculate and save its prior evaluation. Otherwise, we create a 
-non-accessible node that holds a value of 0 to notice the value function that 
-node selection was a failure. This dummy node is only there to update the 
-result of the Current evaluation of other nodes and is not considered in the 
-calculation of exploration term and cannot be selected.
+to the prediction order for the open goal of this node.
+If no open goals exists we apply it to the first pending goal (according to 
+some arbitrary order). If the node is a solved node or (a descendant of a 
+solved node) then 
+If the tactic is productive then the we create a new 
+node and calculate and save its prior evaluation. Otherwise we evaluate the 
+output of the tactic to 0 and backpropagate the failure to the ancestries.
 
@@
 
-To prevent tactics from looping, we add timeout of 0.02 seconds to all tactics 
-(optimized from previous paper)
+To prevent tactics from looping, we interrupt tactics after
+0.01 to 0.1 seconds. The optimal tactic timeout varies in function of the 
+search time limit.
 
@@
 
+\subsection{Backpropagation}
+Number of failed selection and 
+the number of visits are backpropagated to the ancestors
+
+which update the current evaluation in each node.
+
 \subsection{End of the search}
@@
 power of tactics, we would like to combine the \tactictoe search with 
-stronger general purpose provers such as the one in \holyhammer for 
+stronger general purpose provers such as the ones found in \holyhammer for 
 \holfour~\cite{tgck-cpp15}.
 
-During our experiments,
-We chose to distinguish two these general purpose tactic \metis and \eprover.
-\metis is integrated in \holfour, so is already used by the default \tactictoe 
-but we give it a boost as we are confident it is a strong tactic. The boos 
-consits in ...
-
-mention theorem dependencies for eprover but not for metis.
+Such a prover already exists in \holfour and is called \metis. It is already 
+recognized by the tactic selection mechanism in \tactictoe and can have its 
+premises predicted dynamically when in an abstracted form. Nevertheless,
+we think that the performance of \tactictoe can be boosted by giving the ATP 
+\metis a special status. This arrangement consist in always predicting premises 
+for \metis, giving it a slightly higher timeout than other tactics and always 
+trying it first on each node. These modifications only induce a linear overhead 
+on the number nodes as a general purpose prover does not produce more goal 
+obligations.
+
+The integration of an external prover is also possible and 
+\eprover~\cite{eprover} is the one we experimented with .
+To this end, we created a tactic which expects a number of premises as argument 
+and calls \eprover on a version of the problem (premises + goal) translated to 
+first order. We then extract the premises that were required in the \eprover 
+proof and reconstruct it with \metis. Giving a special status to external 
+prover is really necessary as external provers calls do not appear in human 
+proof scripts. The setting in place is about the same as for \metis, the 
+difference being an even higher timeout (5 seconds) and a larger number of 
+predicted premises (128). Since those external calls to \eprover can be 
+computationally expensive, we call it in parallel and asynchronously. This 
+avoids slowing down \tactictoe's main loop. The number of asynchronous calls to 
+an external provers that can run at the same time is limited by the number of 
+cores available to the user.
 
@@
 timeout.
-And the strongest external provers according to past experiments \cite{tgck-cpp15}.
+And the strongest external provers according to past experiments .
 Where as...
-external calls to \eprover are computationally expensive (5 seconds at a 
-minimum +  128 premises), we  
+
 experiment with a small hammer tactic comprised of: a faster premise 
@@
 To integrate a general tactic that requires a higher time out such as \eprover,
-we make use of asynchronous calls via PolyML.fork to prevent the ATP 
-from slowing down the search.
 
@@
  - Metis : 16 lemmas.
- - Eprover: 64 lemmas and 5 seconds timeout 1 asynchronous call
+ - Eprover: 128 lemmas and 5 seconds timeout 1 asynchronous call
 
@@
                 & \metis 0.4s & 125 (50\%)\\ 
-                & \eprover & 125 (50\%)\\
-                & \metis 0.1s + \eprover & 190\\
+                & asynchronous \eprover & 125 (50\%)\\
+                & \metis 0.1s + asynchronous \eprover & 190\\
 \midrule
@@
 tactic necessary. The two main reasons are that either a crucial tactic 
-exceeds 
-its timeout or that the human proof is too deep to be rediscovered (more than 
-10 steps long).In all those experiments orthogonalization is performed with 
+exceeds its timeout or that the human proof is too deep to be rediscovered 
+(more than 10 steps long).In all those experiments orthogonalization is 
+performed with 
 radius 20.
@@
 
-\subsection{Recent improvements}\todo{maybe remove this subsection}
-The final results could be improved by a few percent by including the 
-following additional techniques that are already available.
-The technique already discussed in Section~ would most likely improve the 
-results by 2-3 percent. Reducing the exploration coefficient from 4 to 2 would 
-also probably give a 3 percent increase in the number of proven theorems as 
-shown by a recent 1 in 10 evaluation on the full library.
-
-Until now, all our feature vectors were learned form human proofs. 
-human proofs. We can also add to this tactics that were part of a proof found 
-by our algorithm. To prevent duplication of effort, orthogonalization of those 
-tactics is essential to have a beneficial effect (see Experiments).
-Since recording and re-proving are intertwined, the 
-additional data is available for the next proof search.
+%\subsection{Recent improvements}\todo{maybe remove this subsection}
+%The final results could be improved by a few percent by including the 
+%following additional techniques that are already available.
+%The technique already discussed in Section~ would most likely improve the 
+%results by 2-3 percent. Reducing the exploration coefficient from 4 to 2 would 
+%also probably give a 3 percent increase in the number of proven theorems as 
+%shown by a recent 1 in 10 evaluation on the full library.
+
+\paragraph{Self-learning}\todo{explain better or remove or move to future work}
+All our feature vectors have been learned form human proofs. We now can now 
+also add tactic-goal pairs that appears in final \tactictoe's proof. To prevent 
+duplication of effort, orthogonalization of those 
+tactics is essential to have a beneficial effect.
+Since recording and re-proving are intertwined, the additional data is 
+available for the next proof search.
 The hope is that the algorithm will improve faster by learning from its own 
 discovered proofs than from the human proof 
-scripts~\cite{DBLP:conf/cade/Urban07}. 
-Self-learning improves the results by 
-less than a percent.
-
-%It is also the first step toward self-reinforcement learning.
+scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that 
+self-learning improves \tactictoe's success rate by less than one percent.
 
@@
 
-%%%Improve tactic pred with term predictions. 
-%%%just use a similarity on terms.
-%%%should also take into account the things that are not similar.
-%%%Variables should lead to predicting variables. Maybe treat variables 
-%%%separately make a is_var.
-
 \paragraph{Acknowledgments}\label{sect:acks}

### 2018-01-28 tactictoe_jv.tex
search: todo backpropagation

@@
 \section{Orthogonalization}\label{sec:ortho}
-
 Different tactics may transform a single goal in the same way. Exploring such 
@@
 In our experiments, we take a minimal subsumption defined by:
-\[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is $\alpha$-equivalent to }  
- g_2\]
+\[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is }\alpha\mbox{-equivalent 
+to } g_2\]
 \end{definition}
@@
 We can naturally extend a goal subsumption to a set of goals.
-\begin{definition} (Goal set subsumption)\\
+\begin{definition} (Goal list subsumption)\\
 A goal set subsumption is an partial relation on a set of goals.
-Given two sets of goals $l_1$ and $l_2$, we can define it from goal 
+Given two list of goals $l_1$ and $l_2$, we can define it from goal 
 subsumption $\ge$ by:
@@
 database exceeds the coverage of $T$, $A$ will be selected before $T$ without 
-the 
-previously required forcing.
+the previously required forcing.
 \end{remark}
@@
 
-Despite the best efforts of the prediction algorithms, the selected tactic may 
+Despite the best efforts of the prediction algorithms, a selected tactic may 
 not solve the current goal, proceed in the wrong direction
@@
 Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies 
-on prior evaluation and policy functions. Those priors are learned through 
-direct supervised learning in the case of the policy and via the data 
+on a prior evaluation function and a prior policy function. Those priors are 
+learned through direct supervised learning for the policy and via the data 
 accumulated during orthogonalization for the evaluation. Therefore, this MCTS 
 algorithm does not need to rely on roll outs for evaluation. The 
-policy and evaluation are estimated by the simple supervised k-nn algorithm 
+policy and evaluation are estimated by the simple supervised k-NN algorithm 
 here instead of neural networks trained through reinforcement learning in 
-AlphaGo Zero~\cite{}.
+AlphaGo Zero~\cite{silver2017mastering}.
 
@@
 %component. 
-We first explain how we compute the evaluation and policy functions 
-from learned data. 
-And then during the application of the MCTS algorithm, we show how 
-node selection steps and node extension steps make the search tree
-(described in Section~\ref{sec:prelim}) evolve.
+We first explain how to compute the policy and evaluation functions that 
+influence the behaviour of the MCTS algorithm.
+Then we describe the five steps of the MCTS algorithm and their effect on 
+the search tree (described in Section~\ref{sec:prelim}): 
+initialization, node selection, node extension, backpropagation and resolution. 
 A characteristic of the MCTS algorithm is that it offers a good trade-off 
@@
 
-\subsection{Start of the search}
+\subsection{Initialization}
 The input of the algorithm is a goal $g$ that we want to prove.
@@
 
-\subsection{End of the search}
-The proof search stops when one of these 3 conditions is realized:
+\subsection{Resolution}
+The search ends when the algorithm reaches one of these 3 conditions:
 \begin{itemize}
-\item It found a proof (i.e. the root node is solved). In this case, 
+\item It finds a proof (i.e. the root node is solved). In this case, 
 the search returns a minimized and 
 prettified proof script (see Section~\ref{sec:proofdisplay}).
-\item It has saturated. 
-There is no more tactics to be applied to any open goal. \todo{define open 
-goal}
-\item It exceeded a given time limit.
+\item It saturates. There is no tactics to be applied to any open goal.
+\item It exceeds a given time limit.
 \end{itemize}
 
-The second case happens very rarely (less than 10 times when evaluating 7088 
-theorems) and only at the beginning of the library when the learned dataset is 
-small. So in our experiments, we just give the number of proof found relative 
-to the number of proof attempts.
+The second case happens very rarely (less than 10 times in the full scale 
+experiment). So our experiments report only the number of proofs found relative 
+to the number of attempted problems.
 \end{definition}
@@
 we think that the performance of \tactictoe can be boosted by giving the ATP 
-\metis a special status. This arrangement consist in always predicting premises 
-for \metis, giving it a slightly higher timeout than other tactics and always 
-trying it first on each node. These modifications only induce a linear overhead 
-on the number nodes as a general purpose prover does not produce more goal 
-obligations.
+\metis a special status. This arrangement consists in always predicting 
+premises for \metis, giving it a slightly higher timeout than other tactics and 
+always trying it first on each node. These modifications only induce a
+overhead linear on the number nodes as a general purpose prover does not create
+goal obligations.
 
@@
 difference being an even higher timeout (5 seconds) and a larger number of 
-predicted premises (128). Since those external calls to \eprover can be 
-computationally expensive, we call it in parallel and asynchronously. This 
-avoids slowing down \tactictoe's main loop. The number of asynchronous calls to 
-an external provers that can run at the same time is limited by the number of 
-cores available to the user.
-
-can be call with frequently with low 
-timeout.
-And the strongest external provers according to past experiments .
-Where as...
-
-experiment with a small hammer tactic comprised of: a faster premise 
-selection algorithm and
-a short call (0.1 seconds + 16 premises optimized by the previous paper) to the 
-internal prover \metis~\cite{metis} or a
-longer call to external prover \eprover.
-
-\paragraph{Prediction of general purpose tactics}
-Since we do not always have the data to learn on which goal this general 
-tactics are successful, we assume as a first approximation that they are the 
-strongest all around.
-
-And we try them first on each node during the proof search with the advantage 
-that if it succeeds the node is closed rapidly.
-
-This general tactics requires a list of relevant lemmas.
-
-This forced behavior could often be rediscovered from the training data for the 
-internal provers \metis by using theorem abstraction. essentially a call to 
-\metis where its 
-theorem list argument has been abstracted.
-
-\paragraph{Application of general purpose tactics}
-The slightly higher timeout of 0.1 seconds given to \metis does not slowdown 
-the proof too much but...
-To integrate a general tactic that requires a higher time out such as \eprover,
+predicted premises (128). Since calls to \eprover are
+computationally, they are run in parallel and asynchronously. This 
+avoids slowing down \tactictoe's search loop. The number of asynchronous calls 
+to an external provers that can be executed at the same time is limited by the 
+number of cores available to the user.
+
+%can be call with frequently with low 
+%timeout.
+%And the strongest external provers according to past experiments .
+%Where as...
+%
+%experiment with a small hammer tactic comprised of: a faster premise 
+%selection algorithm and
+%a short call (0.1 seconds + 16 premises optimized by the previous paper) to 
+%the 
+%internal prover \metis~\cite{metis} or a
+%longer call to external prover \eprover. 
+%
+%
+%\paragraph{Prediction of general purpose tactics}
+%Since we do not always have the data to learn on which goal this general 
+%tactics are successful, we assume as a first approximation that they are the 
+%strongest all around. 
+%
+%And we try them first on each node during the proof search with the advantage 
+%that if it succeeds the node is closed rapidly.
+%
+%This general tactics requires a list of relevant lemmas.
+%
+%This forced behavior could often be rediscovered from the training data for 
+%the 
+%internal provers \metis by using theorem abstraction. essentially a call to 
+%\metis where its 
+%theorem list argument has been abstracted.
+%
+%
+%\paragraph{Application of general purpose tactics}
+%The slightly higher timeout of 0.1 seconds given to \metis does not slowdown 
+%the proof too much but...
+%To integrate a general tactic that requires a higher time out such as 
+%\eprover,
 

### 2018-01-29 tactictoe_jv.tex
minor

@@
 \section{Preliminaries: Search Tree}\label{sec:prelim}
+\todoi{Print words correctly in math format by avoiding math format}
 
@@
 
+\begin{definition}\label{def:desc}(Children, Descendant)\\
+We now define what are the children and the descendants of a node $a$ in a 
+search tree $\mathfrak{T}$. The children of a relative to a goal $g$ in $G(a)$ 
+is the set of nodes produced by tactics applied to $g$:
+  \[Children(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]
+
+From that, we define the children of a node $a$ by:
+  \[Children(a) = \bigcup_{g \in G(a)} Children(a,g) \]
+
+And descendants of $a$ by transitive closure:
+\begin{align*}
+Descendant^{0}(a) &=_{def} \lbrace a \rbrace \\ 
+Descendant^{n+1}(a) &=_{def} \bigcup_{a' \in Descendant^{n}(a)} 
+Children(a') \\
+Descendant(a) &=_{def} \bigcup_{n \in \mathbb{N}} Descendant^n(a)\\
+\end{align*}
+
 Starting from the bottom parent node, we can construct children of a goal of 
@@
 solved, open or pending (see Definition~\ref{def:solved}). 
+\end{definition}
 
 \begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal, 
-pending goal)\\
+pending goal , active node)\\
 Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
@@
 \exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
-\mathbb{A}^\infty &=_{def} \bigcup_{i \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
-\mathbb{G}^\infty =_{def} \bigcup_{i \in \mathbb{N}} \mathbb{G}^n\\
+\mathbb{A}^\infty &=_{def} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \ 
+\mathbb{G}^\infty =_{def} \bigcup_{n \in \mathbb{N}} \mathbb{G}^n\\
 \end{align*}
@@
 \[\lbrace g \in G\ |\  T(g) = \emptyset \rbrace\]
+
+To facilitate the description  of our algorithms we make a slight adjustement 
+to the definition of an open goal. If a node $a$ does not contain any open 
+goal, the first pending goal (according 
+to some preset order) of $a$ is also considered to be an open goal. In this 
+way, each unsolved node contains a unique open goal.
 \end{definition}
+
+\todo{make definition in preliminaries of active node and descendant?}
+
+We can also judge the productivity of a tactic by its contribution to the 
+search 
+tree.
+
+\begin{definition} (Productive tactic) 
+The application of a tactic on an open goal is called productive if and only if 
+all the following conditions are satisified:
+\begin{itemize}
+\item It does not fail or timeout. To prevent tactics from looping, we 
+interrupt tactics after a short amount of time (0.05 seconds). 
+\item It does not loop. Its output does not contain any goals that appears in 
+its ancestor nodes.
+\item It is not a parallel step. Its output is not equal or a superset 
+of a list of goals produced from the same open goal.
+\end{itemize}
+
+The third point of the definition is a partial attempt at preventing confluent 
+branches. The general case where two branches join after $n$ steps is handled 
+by a tactic cache which memorizes tactic effects on goals. This
+cache allows faster re-exploration of confluent branches which remain 
+separated in the search tree.
+
+\todo{make definition in preliminaries of active node and descendant?}
 
@@
 We first explain how to compute the policy and evaluation functions that 
-influence the behaviour of the MCTS algorithm.
-Then we describe the five steps of the MCTS algorithm and their effect on 
-the search tree (described in Section~\ref{sec:prelim}): 
-initialization, node selection, node extension, backpropagation and resolution. 
+influence the behavior of the MCTS algorithm.
+We then describe initialization and resolution of the search as well as the 
+different part of a step in the search loop: node selection, node 
+extension, backpropagation.
 A characteristic of the MCTS algorithm is that it offers a good trade-off 
@@
 \subsection{Prior tactic policy}\label{sec:policy}
+
+\todoi{mention productive tactics}
 Distances between goals produced by the tactic predictor are hard 
-to translate into a prior tactic policy, so we rely only on the order of 
+to translate into a prior tactic policy, so we rely solely on the order of 
 tactics to estimate the probability with which each tactic should be tried.
-Let $t_0,\ldots,t_500$ be the list of preselected tactics ordered by their 
+Let $t_0,\ldots,t_n$ be the list of preselected tactics ordered by their 
 distance to a goal $g$. Let $c_{policy}$ be a constant heuristic that estimates 
@@
 
-\[PriorPolicy(t_i) = PP(i) = (1 - c_{policy})^{i} * c_{policy}\]
+\[PriorPolicy(a_i) = PriorPolicy(t_i) = PriorPolicy(i) = (1 - c_{policy})^{i} * 
+c_{policy}\]
+
+Widening policy for a node $a$.
+
+\[WideningPolicy(a) = \]
 
 The first term of the multiplication is the probability that none of the better 
-ranked tactic is selected.
-
-In fact, these probabilities are never used directly and only serves as a 
-guidance to our deterministic node selection algorithm over many simulations.
-
-Since the support of A induces a bijection between tactics of the tree 
-search and the node they 
-produced, any function define on tactics of the tree search can be map to a 
-function defined on nodes. 
-Therefore, we define $PriorPolicy(a)$ as the $PriorPolicy(t)$ where $t$ is the 
-tactic that produced $a$.
+ranked tactic is selected. In fact, these probabilities are never used directly 
+and only serves as a guidance to our deterministic node selection algorithm 
+over many MCTS step.
 
@@
 provable in a number of steps from previous proof search. However, this 
-approach produces a lot of negatives\todo{TG: cite Hard Negative mining (cite 
-Deep math)}. So to reduce this number, we limit 
+approach produces a lot of negatives\todo{CK: cite hard negative mining 
+or/and deep math}. So to reduce this number, we limit 
 ourselves to tactics tested during orthogonalization. We declare a list 
@@
 node selection algorithm. Starting from the root node, we repeatedly chose the 
-children with the highest value.
+children relative to the open goal with the highest value.
 
 The value of a $a_i$ with respect to a search tree is determined by:
-\[Value(a_i) = CurrentEvaluation(a_i) + c_{exploration} * Exploration(a_i)\]
+\[Value(a_i) = CurEvaluation(a_i) + c_{exploration} * Exploration(a_i)\]
 
@@
 evaluations of descendants of $a_i$.
-\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurrentPolicy(a_i)}\]
-
-\[CurrentPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
+\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]
+
+\[CurPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
 The policy is approximatively the  percentage of time a node was visited.The 
@@
 
-The exploitation term is the average of the evaluation of 
-all nodes descendant of $a_i$ (itself included) in the current search tree.
-This evaluation also include failures of the selection process.
-
- It 
-is also called the current evaluation of $a_i$.   
-\[CurrentEvaluation(a_i) = 
-  \sum_{b \in Desc(a_i)} \frac{PriorEvaluation(b)} {(card Desc(a_i) + card 
-  (Failure(a_i))}\]
-
-where $Failure(a_i)$ is the number of run that went through $a_i$ that ended in 
-a failure.
+The exploitation term $CurEvaluation(a_i)$ is the average of the evaluation of 
+all MCTS step that went through $a_i$ (itself included) in the current search 
+tree.
+
+\[CurEvaluation(a_i) = 
+  \sum_{a' \in Descendant(a_i)} \frac{PriorEvaluation(a')} {card\ 
+  Descendant(a_i) 
+  + card\ Failure(a_i)}\]
+
+where $Failure(a_i)$ is the number of MCTS steps that went through $a_i$ which 
+ended in a failure.
 
@@
 descent of the tree. This value has an exploitation equal to 0 since it was 
-never explored and is given a prior policy equal to $PP(n)$ where $n$ is the 
+never explored and is given a prior policy equal to $PriorPolicy(n)$ where $n$ 
+is the 
 number of 
-tactics already tried on the open goal if there is one and $PP(0)$ if there is 
+tactics already tried on the open goal if there is one and $PriorPolicy(0)$ if 
+there is 
 none.
@@
 \subsection{Node extension}
-Once a node is selected, we applied the best not-yet-applied tactics according 
-to the prediction order for the open goal of this node.
-If no open goals exists we apply it to the first pending goal (according to 
-some arbitrary order). If the node is a solved node or (a descendant of a 
-solved node) then 
-If the tactic is productive then the we create a new 
-node and calculate and save its prior evaluation. Otherwise we evaluate the 
-output of the tactic to 0 and backpropagate the failure to the ancestries.
-
-Maybe should move this definition in static description.
-
-To prevent tactics from looping, we interrupt tactics after
-0.01 to 0.1 seconds. The optimal tactic timeout varies in function of the 
-search time limit.
-
-\begin{definition} (Productive tactic)
-During a proof search, after a tactic $t$ is applied to a node with active goal 
-$g$, $t$ is called 
-productive if and only if all the following conditions are satisified:
-\begin{itemize}
-\item It does not fail. We note $l$ the list of goals returned by the tactic.
-\item It does not loop. $l$ do not contain any goals that appears in its 
-ancestor nodes.
-\item It is not a parallel step. $l$ is not equal to or a superset of a list of 
-goals produced from the same node with the same active goal.
-\end{itemize}
-
-The third point of the definition is a partial attempt to prevent confluent 
-branches. The general 
-case where two branches becomes parallel after $n$ steps is handle by a
-tactic cache which memorizes tactic effects on goals. Similarly there is also a 
-prediction cache, which memorize the tactic predictions for each goal. Those 
-two caches allow a fast exploration of confluent branches but they remain 
-separated in our search tree.
+Once a node is selected, we first check if it is an active part of the search 
+tree (not a solved node .
+
+If the selected node is not active, then the result of node extension is a 
+failure.
+Otherwise, we apply the best untested tactic on the open goal $g$ of this node  
+according to the prediction order for $g$.
+
+we apply the next tactic in the prediction order for 
+the open goal of this node.
+
+The result of the extension step is either a failure or a list of goals $l$.
+From $l$, we can assess if the tactic was productive. 
+If not, we also return a failure.
+Otherwise, the list of goals $l$ is used to create a new node in the search 
+tree. Give precise definition.
+
+order of predicted tactics.
+
+During the creation of node, we predict tactics for the open goals
 
 \subsection{Backpropagation}
-Number of failed selection and 
-the number of visits are backpropagated to the ancestors
-
-which update the current evaluation in each node.
+
+If node extension resulted in a failure, we increment the failure field of the 
+selected node and its parents. If node extension was successful, we 
+backpropagate the evaluation of the new nodes to their parents.
+In both cases, we increment the number of visits of all nodes traversed.
+This changes induce an update on the current evaluation and current policy for 
+each node. And the process loops back to the node selection step.
 
@@
 \item It saturates. There is no tactics to be applied to any open goal.
-\item It exceeds a given time limit.
+This happens less than 10 times in the full-scale experiment.
+\item It exceeds a given time limit set by default to 60 seconds.
 \end{itemize}
 
-The second case happens very rarely (less than 10 times in the full scale 
-experiment). So our experiments report only the number of proofs found relative 
-to the number of attempted problems.
 \end{definition}

### 2018-01-29 tactictoe_jv.tex
search: first final version

@@
 
-\begin{definition}\label{def:desc}(Children, Descendant)\\
-We now define what are the children and the descendants of a node $a$ in a 
-search tree $\mathfrak{T}$. The children of a relative to a goal $g$ in $G(a)$ 
-is the set of nodes produced by tactics applied to $g$:
-  \[Children(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]
-
-From that, we define the children of a node $a$ by:
-  \[Children(a) = \bigcup_{g \in G(a)} Children(a,g) \]
-
-And descendants of $a$ by transitive closure:
-\begin{align*}
-Descendant^{0}(a) &=_{def} \lbrace a \rbrace \\ 
-Descendant^{n+1}(a) &=_{def} \bigcup_{a' \in Descendant^{n}(a)} 
-Children(a') \\
-Descendant(a) &=_{def} \bigcup_{n \in \mathbb{N}} Descendant^n(a)\\
-\end{align*}
-
 Starting from the bottom parent node, we can construct children of a goal of 
@@
 solved, open or pending (see Definition~\ref{def:solved}).
-\end{definition}
+
+In the following, new concepts and properties are defined in the context of a 
+search tree $\mathfrak{T}$.
 
 \begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal, 
-pending goal , active node)\\
-Let $\mathfrak{T}$ be a search tree. The set of solved nodes 
-$\mathbb{A}^\infty$ and 
+pending goal)\\
+The set of solved nodes $\mathbb{A}^\infty$ and 
 the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:
@@
 
-\todo{make definition in preliminaries of active node and descendant?}
-
 We can also judge the productivity of a tactic by its contribution to the 
@@
 separated in the search tree.
-
-\todo{make definition in preliminaries of active node and descendant?}
+\end{definition}
+
+\begin{definition}\label{def:desc}(Children, descendant, active node)\\
+We now define what are the children and the descendants of a node $a$. The 
+children of a relative to a goal $g$ in $G(a)$ 
+is the set of nodes produced by tactics applied to $g$:
+  \[Children(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]
+
+If $g$ is the open goal of $a$, then we simply call children of a the 
+children of $a$ relative to $g$. We define:\\
+
+Extended children of $a$ by:
+  \[ExtChildren(a) = \bigcup_{g \in G(a)} Children(a,g) \]
+
+Descendants of $a$ by:
+\begin{align*}
+Descendant^{0}(a) &=_{def} \lbrace a \rbrace \\ 
+Descendant^{n+1}(a) &=_{def} \bigcup_{a' \in Descendant^{n}(a)} 
+ExtChildren(a') \\
+Descendant(a) &=_{def} \bigcup_{n \in \mathbb{N}} Descendant^n(a)\\
+\end{align*}
+
+Descendants of $a$ through $g$ by:
+\[Descendant(a,g) =_{def} 
+  \lbrace a \rbrace \cup \bigcup_{a' \in Children(a,g)} Descendant(a')\]
+
+An active node is by definition a node that is not a descendant of a 
+solved node. Non-active nodes are not worth exploring.
+
+\end{definition}
 
@@
 similar to $l$. The ratio of positive examples in the selection gives us a 
-prior node evaluation for MCTS.
+prior evaluation for MCTS.
 
@@
 
-\subsection{Prior tactic policy}\label{sec:policy}
-
-\todoi{mention productive tactics}
-Distances between goals produced by the tactic predictor are hard 
-to translate into a prior tactic policy, so we rely solely on the order of 
+\subsection{Prior policy}\label{sec:policy}
+Co-distances between goals produced by the tactic predictor are hard 
+to translate into a prior policy, so we rely solely on the order of 
 tactics to estimate the probability with which each tactic should be tried.
-Let $t_0,\ldots,t_n$ be the list of preselected tactics ordered by their 
-distance to a goal $g$. Let $c_{policy}$ be a constant heuristic that estimates 
-how likely a tactic $t_i$ is to prove the goal compared to the rest of the 
-tactics $t_{i+1},\ldots,t_{500}$. We make the strong assumption that $c_policy$ 
-is constant across the search. We can now calculate the prior policy by:
-
-\[PriorPolicy(a_i) = PriorPolicy(t_i) = PriorPolicy(i) = (1 - c_{policy})^{i} * 
-c_{policy}\]
-
-Widening policy for a node $a$.
-
-\[WideningPolicy(a) = \]
-
-The first term of the multiplication is the probability that none of the better 
-ranked tactic is selected. In fact, these probabilities are never used directly 
-and only serves as a guidance to our deterministic node selection algorithm 
-over many MCTS step.
-
-\subsection{Prior node evaluation}\label{sec:evaluation}
-
-We now concentrate on defining a reasonable evaluation function that also 
-influences node selection in MCTS. Essentially, we estimate how 
-interesting is a list of goals produced by a tactic. This way, we can determine 
-if we should continue searching in this direction. Theoretically, an 
-interesting list of goal is one 
-that has the shortest proof, and we could measure the likelihood of being 
+Let $\mathfrak{T}$ be a search tree after a number of MCTS steps.
+Let $p$ be a selected node on this tree and $g$ its open goal. By the 
+application of MCTS, $T(g)$ is the list of $n$ productive tactics already 
+applied to $g$. We order them by their prediction score so that $T(G)= \lbrace 
+t_0,\ldots,t_{n-1} \rbrace$. 
+Let $c_{policy}$ be a constant heuristic that estimates how likely a tactic 
+$t_i$ is to prove the goal. 
+We can now calculate of a prior policy of a child $a_i$ produced by the tactic 
+$t_i$ by:
+
+\[PriorPolicy(a_i) = (1 - c_{policy})^{i} * c_{policy}\]
+
+In order to include the possibly of trying more tactic on $g$ we define the 
+widening policy on the parent $a$ for its open goal $g$ to be:
+
+\[WideningPolicy(p) = (1 - c_{policy})^{n} * c_{policy}\]
+
+\subsection{Prior evaluation}\label{sec:evaluation}
+
+We now concentrate on the definition a reasonable evaluation function for the 
+output of tactics. Theoretically, an interesting list of goal is one 
+that has a short proof, and we could measure the likelihood of being 
 provable in a number of steps from previous proof search. However, this 
@@
 or/and deep math}. So to reduce this number, we limit 
-ourselves to tactics tested during orthogonalization. We declare a list 
-of goals l to be positive, noted $l \in Positive$, if it is has been 
-produced by the winner of an orthogonalization competition. The set of the $r$ 
-goals closest to a list of goals $l$ is noted $Neighbour(r,l)$. We can 
-define the set of positives within a 
-radius $r$ of $l$ to be $Positive(r,l) = \lbrace m \in Neighbour(r,l)\ |\ m \in 
-Positive \rbrace$.
-
-During our experiments, we evaluate output goals with the following function:
-
+ourselves to collect output of tactics tested during orthogonalization. We 
+declare a list of goals $l$ to be positive if it is has been produced by the 
+winner of an orthogonalization competition. The predictor with radius $r$ 
+\todo{define radius} produces a set of $r$ list of goals closest to $l$. The 
+number of positive example in this set is noted $Positive(r,l)$.
+During the proof search, we evaluate nodes through the list of goals they 
+contain using the following evaluation function:
 \begin{align*}
-PriorEvaluation_r (a_i) &= \frac{card\ Positive(r,a_i)}{r}\\
+PriorEvaluation_r (a_i) &= \frac{Positive(r,G(a_i))}{r}\\
 \end{align*}
-
-If a tactic is not productive then there is no output to evaluate, so the 
-result is a failure and a reward of $0$ will be 
-given for this node. Failure also happens if the node selected is already 
-a proven goal or a descendant of a proven goal. Experimentally, it is 
-preferable not 
-to filter solved nodes and descendants of solved nodes before node selection 
-to keep the estimations of the algorithm balanced.
 
@@
 Through node extension steps the search tree grows and the number of paths to 
-explore increases. To decide which node to extend next, we proceed through the 
-node selection algorithm. Starting from the root node, we repeatedly chose the 
-children relative to the open goal with the highest value.
-
-The value of a $a_i$ with respect to a search tree is determined by:
+explore increases. To decide which node to extend next, a evolving value for 
+each node is given in Definition~\ref{def:value}. The algorithm that performs 
+node selection starts from the root of the search tree. X, it repeatedly choses 
+the children with the highest values. Depending on the widening policy, it can 
+stop at any chosen node on during its descent through the tree. We now give a 
+more precise description of 
+the algorithm with the following pseudo-code:
+
+\todoi{Replace alltt environment by something prettier}
+\begin{alltt}
+ChosenNode = Root(Tree);
+while true do
+  if Children(ChosenNode) = \(\emptyset\) then break;
+  (BestChild, BestValue) = CompareChildren (ChosenNode);
+  if WideningPolicy (ChosenNode) \(\geq\) BestValue then break;
+  ChosenNode = BestChild
+end;
+return ChosenNode;
+\end{alltt}
+
+\begin{definition}\label{def:value}(Value)
+The value of the $i^th$ child $a_i$ of a parent $p$ is determined by:
 \[Value(a_i) = CurEvaluation(a_i) + c_{exploration} * Exploration(a_i)\]
@@
 
-The exploitation term $CurEvaluation(a_i)$ is the average of the evaluation of 
-all MCTS step that went through $a_i$ (itself included) in the current search 
-tree.
+The current evaluation $CurEvaluation(a_i)$ is the average evaluation of 
+all descendants of $a_i$, through its open goal $g_i$, including node extension 
+failures:
 
 \[CurEvaluation(a_i) = 
-  \sum_{a' \in Descendant(a_i)} \frac{PriorEvaluation(a')} {card\ 
-  Descendant(a_i) 
-  + card\ Failure(a_i)}\]
-
-where $Failure(a_i)$ is the number of MCTS steps that went through $a_i$ which 
-ended in a failure.
-
-To allow the algorithm to select nodes that 
-are not leafs, we also assign a value to the possibility of stopping during the 
-descent of the tree. This value has an exploitation equal to 0 since it was 
-never explored and is given a prior policy equal to $PriorPolicy(n)$ where $n$ 
-is the 
-number of 
-tactics already tried on the open goal if there is one and $PriorPolicy(0)$ if 
-there is 
-none.
+  \sum_{a' \in Descendant(a_i,g_i)} \frac{PriorEvaluation(a')} {card\ 
+  Descendant(a_i,g_i) + Failure(a_i)}\]
+
+The number $Failure(a_i)$ is the amount of failures that occurred during node 
+extension from descendants of $a_i$ .
+\end{definition}
 
 \subsection{Node extension}
-Once a node is selected, we first check if it is an active part of the search 
-tree (not a solved node .
-
-If the selected node is not active, then the result of node extension is a 
-failure.
-Otherwise, we apply the best untested tactic on the open goal $g$ of this node  
-according to the prediction order for $g$.
-
-we apply the next tactic in the prediction order for 
-the open goal of this node.
-
-The result of the extension step is either a failure or a list of goals $l$.
-From $l$, we can assess if the tactic was productive. 
-If not, we also return a failure.
-Otherwise, the list of goals $l$ is used to create a new node in the search 
-tree. Give precise definition.
-
-order of predicted tactics.
-
-During the creation of node, we predict tactics for the open goals
+Let $a$ be the result of node selection.
+If $a$ is not an active part of the search then extending $a$ is useless and 
+the algorithm reports a failure. In our experiments, filtering nodes that not 
+active before node selection is 
+not beneficial. It may be because it breaks some symmetry in the 
+$Value$ function. 
+Otherwise, it applies the best untested tactic $t$ on the open goal $g$ of this 
+node according to the prediction order for $g$. If no such tactic exists or if 
+$t$ is not productive, a failure is returned. 
+Otherwise, a new node $c$ is created that contains the list of goals $l$ 
+produced by $t$ and is added to the children of $a$.
 
 \subsection{Backpropagation}
-
-If node extension resulted in a failure, we increment the failure field of the 
-selected node and its parents. If node extension was successful, we 
-backpropagate the evaluation of the new nodes to their parents.
-In both cases, we increment the number of visits of all nodes traversed.
-This changes induce an update on the current evaluation and current policy for 
-each node. And the process loops back to the node selection step.
+During backpropagation we update the statisitics of all the nodes traversed or 
+created during this MCTS step:
+\begin{itemize}
+\item Their number of visit is incremented.
+\item If node extension failed, their failure count is incremented.
+\item If node extension succeeded, they inherit the evaluation of the created 
+child. 
+\end{itemize}
+
+These changes updates the current evaluation and the current policy of the 
+traversed nodes. 
+After completing backpropagation, the process loops back to node 
+selection.
 
@@
 
-\end{definition}
-
 \subsection{ATP Integration}~\label{sec:atp}
@@
 \metis a special status. This arrangement consists in always predicting 
-premises for \metis, giving it a slightly higher timeout than other tactics and 
-always trying it first on each node. These modifications only induce a
+premises for \metis, giving it a slightly higher timeout and 
+always trying it first on each node. These modifications only induce an
 overhead linear on the number nodes as a general purpose prover does not create
-goal obligations.
+goal obligations. \todoi{Define goal obligation in the introduction}
 
@@
 To this end, we created a tactic which expects a number of premises as argument 
-and calls \eprover on a version of the problem (premises + goal) translated to 
-first order. We then extract the premises that were required in the \eprover 
-proof and reconstruct it with \metis. Giving a special status to external 
-prover is really necessary as external provers calls do not appear in human 
+and calls \eprover on the translated first order problem. We then extract the
+premises that were required in the \eprover 
+proof and reconstruct it inside \holfour with \metis. Giving a special status 
+to external 
+prover is essential as external provers calls do not appear in human
 proof scripts. The setting in place is about the same as for \metis, the 
@@
 predicted premises (128). Since calls to \eprover are
-computationally, they are run in parallel and asynchronously. This 
+computationally expensive, they are run in parallel and asynchronously. This 
 avoids slowing down \tactictoe's search loop. The number of asynchronous calls 
@@
 number of cores available to the user.
-
-%can be call with frequently with low 
-%timeout.
-%And the strongest external provers according to past experiments .
-%Where as...
-%
-%experiment with a small hammer tactic comprised of: a faster premise 
-%selection algorithm and
-%a short call (0.1 seconds + 16 premises optimized by the previous paper) to 
-%the 
-%internal prover \metis~\cite{metis} or a
-%longer call to external prover \eprover. 
-%
-%
-%\paragraph{Prediction of general purpose tactics}
-%Since we do not always have the data to learn on which goal this general 
-%tactics are successful, we assume as a first approximation that they are the 
-%strongest all around. 
-%
-%And we try them first on each node during the proof search with the advantage 
-%that if it succeeds the node is closed rapidly.
-%
-%This general tactics requires a list of relevant lemmas.
-%
-%This forced behavior could often be rediscovered from the training data for 
-%the 
-%internal provers \metis by using theorem abstraction. essentially a call to 
-%\metis where its 
-%theorem list argument has been abstracted.
-%
-%
-%\paragraph{Application of general purpose tactics}
-%The slightly higher timeout of 0.1 seconds given to \metis does not slowdown 
-%the proof too much but...
-%To integrate a general tactic that requires a higher time out such as 
-%\eprover,
 

### 2018-01-29 tactictoe_jv.tex
minor

@@
 
-The exploration term is determined by the prior policy and the current policy 
-where as the exploitation is exactly the current evaluation build from prior of 
-evaluations of descendants of $a_i$.
+The current evaluation $CurEvaluation(a_i)$ is the average evaluation of 
+all descendants of $a_i$, through its open goal $g_i$, including node extension 
+failures:
+
+\[CurEvaluation(a_i) = 
+  \sum_{a' \in Descendant(a_i,g_i)} \frac{PriorEvaluation(a')} {card\ 
+  Descendant(a_i,g_i) + Failure(a_i)}\]
+
+The number $Failure(a_i)$ is the amount of failures that occurred during node 
+extension from descendants of $a_i$ .
+
+The exploration term is determined by the prior policy and the current policy.
 \[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]
@@
 
-The current evaluation $CurEvaluation(a_i)$ is the average evaluation of 
-all descendants of $a_i$, through its open goal $g_i$, including node extension 
-failures:
-
-\[CurEvaluation(a_i) = 
-  \sum_{a' \in Descendant(a_i,g_i)} \frac{PriorEvaluation(a')} {card\ 
-  Descendant(a_i,g_i) + Failure(a_i)}\]
-
-The number $Failure(a_i)$ is the amount of failures that occurred during node 
-extension from descendants of $a_i$ .
+The coefficient $c_{exploration}$ is experimentally determined and adjusts the 
+trade-off between exploitation and exploration.
 \end{definition}

### 2018-01-29 tactictoe_jv.tex
adapting orthogonalization and abstraction

@@
 is undesirable, as it leads to inefficiency in automated proof search.
-To solve this problem, we do not directly assign a goal to the associated 
-tactic, but organize a competition on the closest feature vectors. The number 
-of selected feature vectors is a parameter that is set for each experiment.
-The winner is the tactic appearing in the most feature vectors provided that it 
-has the 
-same effect as the original tactic. We associate this tactic with the features 
-of the targeted goal instead of the original in our feature database. As a 
-result, already successful tactics are preferred, and new tactics are 
-considered only if they provide a different contribution. We also advantage 
-tactics that have a large coverage. In the following, we give formal definition 
-of the concepts of subsumption and coverage that are required for expressing 
-the orthogonalization algorithm.
+To solve this problem, we do not directly assign an input goal $g$ to its 
+associated tactic $t$, but organize a competition between the $n$ closest 
+tactic-goal pairs, noted $P_n(g)$. 
+The winner is the tactic that subsumes the original tactic on $g$ and that
+appears in the largest number of tactic-goal pairs in the database.
+The winning tactic $w$ is then associated with $g$, producing the pair $(w,g)$ 
+that replaces the original pair $(t,g)$ in the database. As a 
+result, already successful tactics with a large coverage are preferred, and new 
+tactics are 
+considered only if they provide a different contribution. In the following, we 
+give a formal definition of the concepts of subsumption and coverage that are 
+required for expressing the orthogonalization algorithm.
 
@@
 \[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is a consequence of } g_2\]
-In our experiments, we take a minimal subsumption defined by:
+By default, we choose a minimal subsumption defined by:
 \[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is }\alpha\mbox{-equivalent 
@@
 
-We can naturally extend a goal subsumption to a set of goals.
+We can naturally extend a goal subsumption to lists of goals.
 \begin{definition} (Goal list subsumption)\\
-A goal set subsumption is an partial relation on a set of goals.
+A goal list subsumption is an partial relation on lists of goals.
 Given two list of goals $l_1$ and $l_2$, we can define it from goal 
@@
 
-We now build a subsumption on the tactic level.
-\begin{definition} (Tactic subsumption)\\  
+This allows us to define subsumption for tactics.
+\begin{definition}(Tactic subsumption)\\  
 Given two non-failing tactics $t_1$ and $t_2$ on $g$, a tactic $t_1$ subsumes a 
@@
 
-If one of the tactic is failing then the couple $(t_1,t_2)$ is not an element 
-of the 
-relation.
-Given the minimal subsumption on goals, a tactic $t_1$ subsumes another $t_2$ 
-on a goal 
-$g$ if and only if the resulting set of goals produced by $t_1$ is a subset 
-of the output of $t_2$ on $g$. 
+If one of the tactic is failing then $t_1$ and $t_2$ are not comparable through 
+this relation.
+Given the default subsumption on goals, $t_1$ subsumes $t_2$ on $g$ if and only 
+if $t_1(g)$ is a subset (modulo $\alpha$-equivalence) of $t_2(g)$. 
 \end{definition}
 
-We can now express the orthogonalization algorithm given in a single formula 
-with parameters: $P_n$ for the tactic predictor where $n$ for the number of 
-predicted
-tactics, $S$ for the tactic subsumption relation and $\textit{\mbox{Coverage}}$ 
-for the coverage of 
-a tactic. The orthogonalization process $\mathbb{O}$ is given as input a 
-recorded feature vector $(t,g)$ where $t$ is the tactic originally applied to 
-the goal $g$. The winning tactic of the competition is given by the formula:
+In the end, the winning tactic of the orthogonalization competition can be 
+expressed by the formula:
 
@@
 
-After orthogonalization, because the orthogonalized tactic is supposedly 
-better, we record the feature vector $(\mathbb{O}(t,g),g)$ instead of 
-$(t,g)$. To sum up, the winning tactic is the one with the most coverage and 
-that is able to replicate the effect of the original tactic.
-
-\section{Abstraction}\label{sec:synthesis}\todo{MN + RK: review + fix typos}
+As a consequence $(\mathbb{O}(t,g),g)$ replaces $(t,g)$ in the database of 
+tactics.
+
+\section{Abstraction}\label{sec:synthesis}
 One of the great weakness of the previous version of \tactictoe is that
@@
 unspecified arguments. To apply an abstracted tactic $A$, we first need to 
-instantiate that should placeholders $h$ inside $A$. Because it is difficult to 
+instantiate the placeholders $h$ inside $A$. Because it is difficult to 
 build a general predictor that effective on each type of argument, we manually 
@@
 usually given a goal has input and returns the best chosen argument for this 
-goal(see Section~\ref{sec:arg_pred}). All in all, an abstracted tactics has 
-dynamic arguments, it decides how to instantiate them depending on the given 
-goal.
+goal. All in all, an abstracted tactics has 
+dynamical arguments, it instantiate them differently for different goals.
+
+We made successful argument predictors for one type of arguments: list 
+of theorems. It is a very common type and the structure of theorems 
+are complex, thus they contain enough information to be predicted accurately. 
+We plan to extend our algorithm to term which is another frequent argument 
+types but term synthesis is so far too weak to positively 
+influence the proof search.
 
 \paragraph{Selection of Abstracted Tactics}
-As abstracted tactics are not used in human proofs, we need to find a way to 
+As abstracted tactics do not appear in human proof, we need to find a way to 
 predict them so that they can contribute to the \tactictoe proof search.
@@
 In fact, we can re-use the orthogonalization module to do this task for us.
-During the orthogonalization competition on goal $g$, before trying a tactic 
-$T$, we now also try to apply its abstracted version $A$. If $A$ succeeds, it 
-becomes a a label for $g$ is including in the database of tactic features and 
-thus can be predicted during a proof search.
-
-\begin{remark}
-Forcing $A$ into the competition is necessary because it has never been seen 
-before so it has no coverage. But after the coverage $c_A$ of $A$ in the 
-database exceeds the coverage of $T$, $A$ will be selected before $T$ without 
-the previously required forcing.
-\end{remark}
-
-We made argument predictors for one type of arguments: list 
-of theorems. We choose it is the most common type and the structure of theorems 
-are complex so they contain enough information to be predicted accurately. 
-We plan to extend our algorithm to term which is another frequent argument 
-types but term synthesis is so far too weak to positively 
-influence the proof search.
+We add $A$ to the competition, initially giving it the coverage of $T$. If $A$ 
+wins, it becomes a a label for $g$ is including in the database of tactic 
+features and thus can be predicted during a proof search.
+After many competitions, the coverage of $A$ may exceed the coverage of $T$. At 
+this point, the coverage of $A$ is estimated on its own and not inherited from 
+$T$ anymore.
 

### 2018-01-29 tactictoe_jv.tex
minro

@@
 for each object which is the maximum of its similarity score and the similarity 
-score of its parent. In the end, the 500 feature vectors in each dataset with 
-the highest dependency score are preselected.
+score of its parent. In the end, the 500 entries with highest dependency score 
+are preselected in each dataset.
 
@@
 \begin{definition} (Coverage)\\ 
-Let $\mathfrak{F}$ be a recorded set of feature vectors. We can define the 
+Let $\mathfrak{F}$ be the database of tactics. We can define the 
 coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times 
@@
 predict them so that they can contribute to the \tactictoe proof search.
-A straightforward idea is to duplicate each tactic during the proof search by 
-creating an abstracted version of the tactic. However, this is computationally 
-expensive when the tactic does not instantiate back to its original form, we 
-risk doing unnecessary work as both version may have a similar effect.
-Therefore, we need to decide beforehand which of the two is better.
+A straightforward idea is to try $A$ before $T$ during the proof search. 
+However, this risks doing unnecessary work as they may have a similar effect. 
+Therefore, we would like to decide 
+beforehand if one is better than the other.
 In fact, we can re-use the orthogonalization module to do this task for us.
@@
 wins, it becomes a a label for $g$ is including in the database of tactic 
-features and thus can be predicted during a proof search.
+features and thus can be predicted during proof search.
 After many competitions, the coverage of $A$ may exceed the coverage of $T$. At 
@@
 
+\todoi{remove feature vectors from the paper}
 \section{Proof Recording}\label{sec:recording}
-\todo{MN + RK: comment technical and hard-to read presentation of the unfolding
-mechanism and replace by a high level abstraction. Alternatively, make the 
-current version more accessible.}
 

### 2018-01-29 tactictoe_jv.tex
minor

@@
 \section{Proof Search}\label{sec:proofsearch}
-
 Despite the best efforts of the prediction algorithms, a selected tactic may 
@@
 
-\section{Proof Presentation}\label{sec:proofdisplay}\todo{MN + RK: add 
-motivations and transitions}
+\section{Proof Presentation}\label{sec:proofdisplay}
 
@@
 
-\todoi{Example prettified proof}
+\todoi{Add an example}
 
@@
 
-\section{Comparison with Human Proofs}\todo{
-  TG + RK: find describe examples.
-  RK: After implementation and small experiment is done (in about 1 week) 
-      Test tactictoe find new examples on new goals and manual exploration mode.
-      Give feedback.
-}
+\section{Comparison with Human Proofs}
+\todoi{Test tactictoe find new examples on new goals and manual exploration 
+mode.}
 
@@
 
-\todo{is copied and modified}
+\todoi{Is copied and modified}
 We proposed a new proof assistant automation technique which combines 

### 2018-01-29 tactictoe_jv.tex
contribs

@@
 
-\todoi{add prediction of 3 kinds of things}
 \begin{itemize}
@@
 recording steps (parsing, unfolding and replaying) are regrouped into a single 
-\sml function. So, our recording is more complete and fully automatized.
-\item Two types of arguments (list of theorems and terms) can be predicted 
-independently of the tactic. Combinatorial explosion in the number of tactics 
-is prevented through the introduction of abstracted tactics. As desired, their 
-instantiations is dependent on the target goal.
+\sml function.
 \item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
@@
 through supervised learning.
-
-\item First evaluation of \tactictoe and \eprover success rates relative to 
+\item Proof guidance required by the MCTS algorithm is given by predictors for 
+three kind of objects: tactics, theorems and list of goals.
+\item The orthogonalization process that eliminates redundant tactics is 
+formally described.
+\item A tactic abstraction method is introduced. It enable us to create 
+more general and flexible tactics where tactic arguments are dynamically 
+predicted.
+\item The internal ATP \metis is complemented 
+by asynchronous calls to \eprover to help \tactictoe during proof search.
+\todoi{More on evaluation??}
+\item Evaluation of \tactictoe success rate relative to 
 the length of the original proofs.
@@
 applied on the generated proof.
-\item The internal ATP \metis used during proof search can be complemented 
-by asynchronous calls to \eprover.
 \end{itemize}
@@
 
-In order to be able to describe the different components of \tactictoe, 
-we need to give to describe what a search tree is, what are its elements and 
-how they are organized relative to one other. The following description is 
-static and the evolution of the search tree toward a proof is given in 
+The different components of \tactictoe revolves around the concept of a 
+search tree. The following definitions are made relative to a static search 
+tree, the description of its evolution during proof search will be given in 
 Section~\ref{sec:proofsearch}.
@@
 node a return the set of goals in this node.
-\item $Support(A) = \lbrace (g,t)\ |\ t \in T(g) \rbrace$.
+
 \item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to 
@@
 application of the tactic to the goal, if this tactic has been applied to this 
-goal.
-\item if $t \in T(g)$ and $A(g,t) = a$ then $t(g) = G(a)$.
+goal. Its support $Support(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g) 
+\rbrace$.
+\item if $(g,t) \in Support (A)$ and $A(g,t) = a$, then $t(g) = G(a)$.
 \end{itemize}
@@
 one root node are properties of this graph that are preserved during the 
-application of the MCTS algorithm. That is why we call it a tree although this 
+application of the MCTS algorithm. That is why we call it a tree, although this 
 is not specified in this definition.
@@
 If no explicit order is given, we assume that any set is equipped with an 
-arbitrary order. In this way, we can access the nth member of set $s$ by 
-writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written $g_j^i$.
+arbitrary order. In this way, we can access the $n^{th}$ member of a set $s$ by 
+writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written a $g_j^i$.
 
@@
 apply to this goal. Here the result produces new nodes each containing a list 
-of goals to be proven. In each node of the proof tree, a goal can either be 
+of goals to be proven called obligations. In each node of the proof tree, 
+a goal can either be 
 solved, open or pending (see Definition~\ref{def:solved}).
 
-In the following, new concepts and properties are defined in the context of a 
+In the following, concepts and properties are defined in the context of a 
 search tree $\mathfrak{T}$.
@@
 classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects: 
- tactics, theorems and list of goals.
+ tactics, theorems and list of goals. The number of predicted objects $k$ is 
+ called the prediction radius.
 
@@
 provable in a number of steps from previous proof search. However, this 
-approach produces a lot of negatives\todo{CK: cite hard negative mining 
+approach produces a lot of negatives\todo{CK: cite millions lemma mining 
 or/and deep math}. So to reduce this number, we limit 
@@
 declare a list of goals $l$ to be positive if it is has been produced by the 
-winner of an orthogonalization competition. The predictor with radius $r$ 
-\todo{define radius} produces a set of $r$ list of goals closest to $l$. The 
+winner of an orthogonalization competition. Our predictor with radius $r$ 
+produces a set of $r$ list of goals 
+closest to $l$. The 
 number of positive example in this set is noted $Positive(r,l)$.
@@
 overhead linear on the number nodes as a general purpose prover does not create
-goal obligations. \todoi{Define goal obligation in the introduction}
+goal obligations.
 

### 2018-01-29 tactictoe_jv.tex
minor

@@
 
-\todo{Comparison with hammers in usability}
+\todoi{Comparison with hammers in usability}
 Many of the state-of-the-art interactive theorem provers (ITPs) such as
@@
 
-\todo{Rewrite the plan}
+\todoi{Rewrite the plan}
   To do this, we \emph{extract and record} tactic invocations from the ITP 
@@
 
-\todoi{remove feature vectors from the paper}
+\todoi{Remove feature vectors from the paper}
 \section{Proof Recording}\label{sec:recording}
@@
 where every top-level value is made independent from
-the context. \todo{some reason why this is needed? easy transfer between 
+the context. \todoi{some reason why this is needed? easy transfer between 
 theories} We 
@@
 The previously decribed transformation affects each identifier in a theory 
-file. However, when we 
-create the recording theory file, only proof scripts are replaced by their 
+file. However, when creating the recording theory file, only proof scripts are 
+replaced by their 
 globalized version and the rest of the code is left unchanged. The effect on 
@@
 
-\begin{example} (Globalization)
+\begin{example}\label{ex:global} (Globalization)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
@@
 We now restrict the definition of a proof script to being the third 
-argument of a call to \texttt{store\_thm}. And we precisely define the 
+argument of a call to 
+\texttt{store\_thm}. And we precisely define the 
 modification applied to the theory file. A modified theory file is identical to 
@@
 
-Motivation: 
-+ minimal effort to integrate the tactictoe results in proof scripts.
-
-Side effect:
-+ could be also used to simplify existing proof scripts.
+\todoi{Motivation: + minimal effort to integrate the tactictoe results in proof 
+scripts.}
 
@@
 understandable, it could replace its original in the \holfour standard library.
+
+\todoi{Side effect:
++ could be also used to simplify existing proof scripts. }
 

### 2018-01-30 tactictoe_jv.tex
minor

@@
 \sml function.
-\item MCTS (Monte Carlo tree search) now replaces A* in our proof search 
-algorithm. MCTS gives a more balanced feedback to guide proof search by 
-comparing subtrees instead of leafs. The policy and evaluation are learned 
+\item Monte Carlo tree search~~\cite{montecarlo} (MCTS) replaces A* as our 
+search algorithm. The MCTS algorithm gives a more balanced feedback on 
+the success of 
+each proof step by
+comparing subtrees of the search tree instead of leafs. The policy and 
+evaluation are learned 
 through supervised learning.
-\item Proof guidance required by the MCTS algorithm is given by predictors for 
+\item Proof guidance required by MCTS is given by predictors for 
 three kind of objects: tactics, theorems and list of goals.
@@
 
-Descendants of $a$ through $g$ by:
-\[Descendant(a,g) =_{def} 
-  \lbrace a \rbrace \cup \bigcup_{a' \in Children(a,g)} Descendant(a')\]
+%Descendants of $a$ through $g$ by:
+%\[Descendant(a,g) =_{def} 
+%  \lbrace a \rbrace \cup \bigcup_{a' \in Children(a,g)} Descendant(a')\]
 
@@
 The current evaluation $CurEvaluation(a_i)$ is the average evaluation of 
-all descendants of $a_i$, through its open goal $g_i$, including node extension 
+all descendants of $a_i$ including node extension 
 failures:
@@
 \[CurEvaluation(a_i) = 
-  \sum_{a' \in Descendant(a_i,g_i)} \frac{PriorEvaluation(a')} {card\ 
-  Descendant(a_i,g_i) + Failure(a_i)}\]
+  \sum_{a' \in Descendant(a_i)} \frac{PriorEvaluation(a')} {card\ 
+  Descendant(a_i) + Failure(a_i)}\]
 
@@
 
-\subsection{Increasing \tactictoe'sknowledge}\label{sec:perfect_exp}
+\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
 
@@
 with tactics from the proof of the tested theorem. The experiments was run with 
-a time out of 5sec on an older version of \tactictoe, as can be seen in the low 
+a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
+the low 
 success rate of the
@@
 radius 20.
-%
-%\subsubsection{Orthogonalization Abstraction}
-%Here, we test the benefit of different kind of tactic abstraction.
-%All experiments in this table are using orthogonalization.
-%
-%\begin{table}[t]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{ll}
-%\toprule
-% Technique & Solved \\
-%\midrule
-%% holyhammer & 79 \\
-%  & 122\\
-% Orthogonalization & 129\\
-%% ortho + term & 126\\    
-% Orthogonalization + Abstraction & 160\\
-%% ortho + \metis & 162\\
-% %ortho + metis + thmortho & 155\\
-% %ortho + metis + stacortho & 160\\
-%% ortho + \metis + theorem list & 175\\ 
-%\bottomrule
-%\end{tabular}
-%\caption{\label{tab:cfot_param} 
-% Effect of orthogonalization and abstraction}
-%\end{table}
-%We want to know what the optimal number of predictions for tactics different 
-%from \metis since \metis has a special status in our algorithm is handled 
-%independently.
-
-%\subsubsection{ATP integration}
-%
-%
-%\todo{paragraph is copied}
-%The best setting for \metis was to be in previous experiments to be 16
-%theorems an The best strategy which does not rely on the ``small hammer'' 
-%approach $D_9$ 
-%will be called \tactictoe(NH) (for no ``small hammer'') in the remaining part 
-%of the paper, and the best strategy relying on the approach, $D_{19}$, will be 
-%referred to as \tactictoe(SH) (``small hammer'').
-%
-%\begin{table}[h!]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{ll}
-%\toprule
-%  & Solved \\
-%\midrule
-%   \eprover &  106 \\
-%   \tactictoe &  139 \\
-%   \tactictoe + \metis & 182 \\
-%   \tactictoe + \eprover & 182 \\
-%   %x tactictoe + async holyhammer (10 thread) & 183 \\
-%   \tactictoe + \metis + \eprover & 190 \\
-%  
-%\bottomrule
-%\end{tabular}
-%\caption{Experiments with a timeout of 60 seconds \label{tab:param}}
-%\end{table}
-%
-%Slight error in this experiments were eprover predictions have been made from 
-%the preselected 500 theorems instead of the full library. But even with the 
-%added overhead, we already see 
-%the benefit of using stronger external prover than \metis.
 
@@
 \midrule
-   \eprover   & 128 premises & 2472 (34.5\%)\\ 
+   \eprover   & & 2472 (34.5\%)\\ 
    \tactictoe & best parameters without \eprover & 4741 (66.2\%)\\
-   combined   &              & 4946 (69.0\%)\\
+\midrule
+   Total  &              & 4946 (69.0\%)\\
 \bottomrule

### 2018-01-30 tactictoe_jv.tex
minor

@@
 
-\section{Preliminaries: Search Tree}\label{sec:prelim}
-\todoi{Print words correctly in math format by avoiding math format}
+\section{Proof Search Tree}\label{sec:prelim}
 
@@
 
-To facilitate the description  of our algorithms we make a slight adjustement 
+To facilitate the description  of our algorithms we make a slight adjustment 
 to the definition of an open goal. If a node $a$ does not contain any open 
@@
 The application of a tactic on an open goal is called productive if and only if 
-all the following conditions are satisified:
+all the following conditions are satisfied:
 \begin{itemize}
@@
 
+some text to add here.
+
 \begin{definition}\label{def:desc}(Children, descendant, active node)\\
@@
 solved node. Non-active nodes are not worth exploring.
-
 \end{definition}
@@
 
-\section{Comparison with Human Proofs}
+\section{Case Study}
 \todoi{Test tactictoe find new examples on new goals and manual exploration 
 mode.}
+In this section, we compare \tactictoe proof script with human proof script.
 

### 2018-01-30 tactictoe_jv.tex
fixing conflicting notations in Abstraction

@@
 place holder creating an tactic with an unspecified argument.
-Suppose we have a recorded tactic $T$. Since $T$ is a \sml code tree, we can 
+Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can 
 try to abstract any of the 
-\sml subterms. Let $u$ be a \sml subterm of $T$, and $h$ a placeholder $u$, we 
+\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder $u$, we 
 can substitute $u$ by $h$ to create an abstracted tactic.
-We note this intermediate tactic $T[h/u]$. By repeating the process, multiple 
+We note this intermediate tactic $t[h/u]$. By repeating the 
+process, multiple 
 arguments may be abstracted in the same tactic. Ultimately, the more 
@@
 An abstracted tactic is not immediately applicable to a goal, since it contains
-unspecified arguments. To apply an abstracted tactic $A$, we first need to 
-instantiate the placeholders $h$ inside $A$. Because it is difficult to 
+unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need 
+to 
+instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to 
 build a general predictor that effective on each type of argument, we manually 
@@
 predict them so that they can contribute to the \tactictoe proof search.
-A straightforward idea is to try $A$ before $T$ during the proof search. 
+A straightforward idea is to try $\hat{t}$ before $t$ during the proof search. 
 However, this risks doing unnecessary work as they may have a similar effect. 
@@
 In fact, we can re-use the orthogonalization module to do this task for us.
-We add $A$ to the competition, initially giving it the coverage of $T$. If $A$ 
+We add $\hat{t}$ to the competition, initially giving it the coverage of $t$. 
+If $\hat{t}$ 
 wins, it becomes a a label for $g$ is including in the database of tactic 
 features and thus can be predicted during proof search.
-After many competitions, the coverage of $A$ may exceed the coverage of $T$. At 
-this point, the coverage of $A$ is estimated on its own and not inherited from 
-$T$ anymore.
+After many competitions, the coverage of $\hat{t}$ may exceed the coverage of 
+$t$. At 
+this point, the coverage of $\hat{t}$ is estimated on its own and not inherited 
+from 
+$t$ anymore.
 
@@
 
-\paragraph{Reconstruction}
-When a proof search succeeds (there are no more pending goals at the root)
-we need to reconstruct a \holfour human-style proof.
-The saved nodes consist of a set of trees where each edge is a tactic and
-the proof tree is the one starting at the root.
-In order to obtain a single \holfour proof, we need to combine the tactics
-gathered in the trees using tacticals.
+When \tactictoe finds a proof of a goal, it returns a search tree 
+$\mathfrak{T}$ where the root node containing this goal is solved.
+In order to transform this search tree into a proof script, we need to extract 
+the tactics that contributed to the proof and combine them using tactic 
+combinators.
+\todoi{Define tactic combinators?}
 By the design of the search, a single tactic combinator, \texttt{THENL}, is 
-sufficient. It combines a tactic with a list of subsequent ones, in such a way 
-that after the parent tactic is called, for each created goal a respective 
+sufficient. It combines a tactic $t$ with a list of subsequent ones, in such a 
+way that after $t$ is called, for each created goal a respective 
 tactic from the list is called.
-The proof tree is transformed into a final single proof script
- by the following recursive function $P$ taking a
-tree node $t$ and returning a string:
-\begin{equation*}
-P(t) =
-\begin{cases}
-P(c) & \text{if $t$ is a root},\\
-tac & \text{if $t$ is a leaf},\\
-tac\ \texttt{THENL}\ [P(c_0),\ldots,P(c_n)] & \text{otherwise.}
-\end{cases}
-\end{equation*}
+Here is a formal description of the transformation $Proof$ defined by mutual 
+recursion on the goals and nodes of $\mathfrak{T}$:
+
+%\begin{equation*}
+%Proof(g) =_{def}
+%\begin{cases}
+%T(g) & \text{if $t$ is a root},\\
+%tac & \text{if $t$ is a leaf},\\
+%tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
+%\end{cases}
+%
+%
+%\begin{cases}
+%tac & \text{if $t$ is a root and $g$ the unique goal of the root},\\
+%tac & \text{if $t$ is a leaf},\\
+%tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
+%\end{cases}
+%
+%
+%
+%
+%\end{equation*}
 where $tac$ is the tactic that produced the node, $c$ is the

### 2018-01-30 tactictoe_jv.tex
minor

@@
 \rbrace$.
-\item if $(g,t) \in Support (A)$ and $A(g,t) = a$, then $t(g) = G(a)$.
+\item If $(g,t) \in Support (A)$, then $t(g) = G(A(g,t))$.
 \end{itemize}
@@
 In order to include the possibly of trying more tactic on $g$ we define the 
-widening policy on the parent $a$ for its open goal $g$ to be:
+widening policy on the parent $p$ for its open goal $g$ to be:
 
@@
 tactic from the list is called.
-Here is a formal description of the transformation $Proof$ defined by mutual 
-recursion on the goals and nodes of $\mathfrak{T}$:
-
-%\begin{equation*}
-%Proof(g) =_{def}
-%\begin{cases}
-%T(g) & \text{if $t$ is a root},\\
-%tac & \text{if $t$ is a leaf},\\
-%tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
-%\end{cases}
+Here is a formal description of the transformation defined by mutual 
+recursion on goals and nodes of $\mathfrak{T}$ by the respective function 
+$P_goal$ and $P_node$:
+
+\begin{equation*}
+P_goal(g) =_{def}
+\begin{cases}
+T(g) & \text{if A(g,T(g)) = \emptyset},\\
+T(g) & \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
+\end{cases}
 %
 %
-%\begin{cases}
-%tac & \text{if $t$ is a root and $g$ the unique goal of the root},\\
-%tac & \text{if $t$ is a leaf},\\
-%tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
-%\end{cases}
-%
-%
-%
-%
-%\end{equation*}
+Proof(a) =_def{def}
+\begin{cases}
+tac & \text{if $t$ is a root and $g$ the unique goal of the root},\\
+tac & \text{if $t$ is a leaf},\\
+tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
+\end{cases}
+\end{equation*}
+
 where $tac$ is the tactic that produced the node, $c$ is the

### 2018-01-31 tactictoe_jv.tex
recording: wip

@@
    {!}{$\forall$}1%
+   {fff}{{DB.fetch}}1%
    {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
@@
 If a tactic produces no goals then its input goal is solved.
+
+\holfour proofs are organized in theory files.
+Each theory file contains definitions and statements of theorems together with
+their proofs. Those theories and the contained proofs
+are written
+in the functional programming language \sml. To automate
+proof search strategies a user can implement programming language functions
+that correspond to more involved proofs.
+Functions that correspond to backward proofs, called tactics, are combined with 
+each other with the help of 
+tacticals (functions that takes
+tactics as arguments). Such a combination of tactics will be called a 
+\emph{proof script} when it
+proves a theorem stated in a theory.
+
+In order to apply machine learning to the existing proof scripts, we will need 
+to
+parse and modify them, which involves parsing and modification of \sml code.
+Acquiring access to the string representation of a tactic (code for the tactic) 
+allows us generating
+proof scripts that can be inspected by a human.
+As some \holfour theories are built in separate sessions, it is not possible
+to transfer \sml values directly, but instead tactic code needs to transferred 
+instead.
+To do this, the transferred information will be saved externally.
 
@@
 \midrule
-Tactic time out & 0.02s & 155 (56.8\%)\\
-                & 0.05s (default) & 155 (56.8\%)\\
-                & 0.1s & 153 (56.0\%)\\
+Tactic time out & 0.02s & 154 (56.8\%)\\
+                & 0.05s (default) & 156 (56.8\%)\\
+                & 0.1s & 154 (56.0\%)\\
 \midrule
-Orthogonalization & $radius = 20$ (default) & 155 (56.8\%)\\ 
-                  & $radius = 10$ & 125 (50\%)\\
-                  & $radius = 40$ & 125 (50\%)\\
+Orthogonalization & none & 156 \\
+                  & $radius = 10$ & 156 \\
+                  & $radius = 20$ (default) & 156 \\ 
+                  & $radius = 40$ & 156 \\
+
 \midrule
-Abstraction       & none (default) & 125 (50\%)\\
-                  & $theorems = 8$ & 125 (50\%)\\
-                  & $theorems = 16$ & 125 (50\%)\\
-                  & $theorems = 32$ & 125 (50\%)\\
+Abstraction       & none (default)  & 156 (\%)\\ 
+                  & $theorems = 8$  & 195 (\%)\\
+                  & $theorems = 16$ & 199 (72.9\%)\\
+                  & $theorems = 32$ & 195 (\%)\\
 \midrule
-MCTS Policy & $c_{policy} = 1/2$ (default) & 125 (50\%)\\
-            & $c_{policy} = 2/3$ & 125 (50\%)\\
-            & $c_{policy} = 1/3$ & 125 (50\%)\\
+MCTS Policy & $c_{policy} = 0.4$ & 149 (\%)\\
+            & $c_{policy} = 0.5$ (default) & 155 (\%)\\
+            & $c_{policy} = 0.6$ & 153 (\%)\\
 \midrule
-MCTS Evaluation & none (default) & 125 (50\%) \\
-           & $c_{exploration} =\sqrt{2}$ & 125 (50\%)\\ 
-           & $c_{exploration} = 1$ & 125 (50\%)\\ 
-           & $c_{exploration} = 2$ & 125 (50\%)\\ 
+MCTS Evaluation & none (default) & 156 (50\%) \\
+				& $c_{exploration} = 0.25$ & 153 153 (50\%)\\ 
+				& $c_{exploration} = 0.5$ & 153 155 (50\%)\\ 
+				& $c_{exploration} = 1$ & 154 (50\%)\\
+				& $c_{exploration} =\sqrt{2}$ & 154 154 (50\%)\\ 
+				& $c_{exploration} = 2$ & 152 (50\%)\\ 
 \midrule
-ATP integration & none (default) & 125 (50\%)\\
+ATP integration & none (default) & 156 (50\%)\\
                 & \metis 0.1s & 125 (50\%)\\
@@
 
-We first discuss parsing the proofs and identifying 
-individual tactic calls. We next show how tactics calls are recorded together 
-with their corresponding goals using a modified proof script.
-Finally, the recorded data is organized as feature vectors with tactics as 
-labels and characteristics of their associated goals as features. 
-These feature vectors constitute the training set for our selection algorithm.
+%We first discuss parsing proof scripts and identifying 
+%individual tactic calls. We next show how to record tactics calls are recorded 
+%with their corresponding goals using a modified proof script.
+%The recorded data which consists of tactic-goal pairs creates the raw 
+%knowledge 
+%necessary for our tactic selection algorithm.
 
 Rather than relying on the underlying programming language, we parse the
-original theory file containing proof scripts into an
-abstract syntax tree (AST).\todoi{reason: reading the proof as if you were a 
-human ==> better for the machine learning algorithm (learn the right thing)}
-
- The AST is then transformed into a globalized AST
-where every top-level value is made independent from
-the context. \todoi{some reason why this is needed? easy transfer between 
-theories} We 
-split the proofs into a series of tactic units and wrap them 
-with a recording function.
-By replaying the proofs produced by the globalized AST, the wrapper can
-record the globalized tactic string with its input goal as training examples in
-a database.
-The combined effect of these steps on a \holfour proof will be illustrated on 
-Example~\ref{ex:running}.
+original theory file containing proof scripts. This enables us to extract the 
+code of each tactic. 
+Working with the string representation of a tactic is better that working with 
+its value:
+\begin{itemize}
+\item Equality between two functions is easy to do from 
+their string representation. The direct consequence is that we can avoid 
+predicting the same tactic over and over again. 
+\item 
+The 
+results of 
+\tactictoe can be returned in a readable form to the user. In this way, the 
+user can learn from the feedback, analyze the proof and possibly improve on it.
+An other point is that \tactictoe does not need to be installed to run  
+proof scripts generated by \tactictoe. In this way, the further development of 
+\tactictoe are not affecting the robustness of \holfour.
+\item
+It is probably difficult (even impossible) to transfer values 
+between theories as they are not run in the same \holfour session. In contrast, 
+\sml code can be exported, imported.
+\item To produce a tactic value form a code is easy in \sml, which can be 
+achieved by using a reference to a tactic and update it with the function 
+\texttt{use}.
+\end{itemize}
+
+In order to transfer our tactic knowledge between theories, we want to be able 
+to re-use the code of a tactic recorded in one theory in an 
+other. We also would like to make sure that the code is interpretable and its 
+interpretation does not change in the other theory.
+
+Even when developing one theory, the context is continually changing: the open 
+modules are different, global and local identifiers are defined and the local 
+ones can become out of scope. Therefore, it is unlikely that code extracted 
+from proof scripts is interpretable in any part of \holfour.
+To solve this issue, we prefix each global identifier by the module where it 
+was declared and we recursively unfold local declaration until a local 
+identifier can be expressed by through code containing only global identifiers.
+We call this process globalization. The result is a standalone \sml code 
+executable in any \holfour theories which will be interpreted in the same way 
+in the absence 
+of side effects.
+
+Therefore, we can guarantee that the behaviour of recorded stateless tactics 
+will not change. And because an update on a stateful tactic 
+generally increases its strength, our prediction algorithm based on past 
+examples will still be reasonably effective.
+
+\subsection{Implementation}
+We describe in more details the implementation of the recording algorithm which 
+consists of 4 phases: proof script extraction, proof script globalization, 
+tactic unit wrapping and creation of tactic-goal pairs.
+
+Because of the large number of \sml constructions, we only describe the effect 
+of these steps in a running example that is representative of a typical proof 
+script. And we comment on the main steps that transforms the original proof 
+script into a modified version that is able to record the code of a tactic and 
+the goal(s) it processed during the proof. We omit more involved constructions 
+such as the globalization of infix operators.
 
@@
 \small
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 open boolLib Tactic Prim_rec Rewrite
@@
 
-%\subsection{Minimal \sml Code Trees}
-%
-%
-%Using the standard types $string$ and $list$ and their type constructors,
-%we define a minimal code tree by the following AST:
-%\begin{align*}
-%M &=_{def} string \\
-%I &=_{def} string\ \vert\ Module(M,I)\\
-%D &=_{def} Open(M)\ \vert\  Val(I,E) \\
-%E &=_{def} Code(I\ list)\ \vert\   Let (D,E)
-%\end{align*}
-%
-%In this AST, the constructors $Let$, $Val$ and $Open$ correspond to
-%the reserved \sml tokens \texttt{let}, \texttt{val} and \texttt{open} 
-%respectively. $Module(M,I)$ refers to the \sml token \texttt{M.I} where 
-%\texttt{M} is a module (also called structure in \sml) and \texttt{I} is an 
-%identifier. $D$ is a declaration, $E$ an expression. Inside a 
-%declaration $val(I,E)$, $E$ is a definition for $I$.
-%To construct such an AST for a theory file, we rely on a custom parser. It 
-%transform a source file into a 
-%minimal code tree taking into account the scope of each constructor.
-%An additional reason to refrain from the use of the interpreter in our case, 
-%is that
-%the \polyml compiler needs to execute the code preceding an expression 
-%before parsing it, which is significantly slower than a custom parser.
-
-\holfour proofs are organized in theory files.
-Each theory file contains definitions and statements of theorems together with
-their proofs. Those theories and the contained proofs
-are written
-in the functional programming language \sml. To automate
-proof search strategies a user can implement programming language functions
-that correspond to more involved proofs.
-Functions that correspond to backward proofs, called tactics, are combined with 
-each other with the help of 
-tacticals (functions that takes
-tactics as arguments). Such a combination of tactics will be called a 
-\emph{proof script} when it
-proves a theorem stated in a theory.
-
-In order to apply machine learning to the existing proof scripts, we will need 
-to
-parse and modify them, which involves parsing and modification of \sml code.
-Acquiring access to the string representation of a tactic (code for the tactic) 
-allows us generating
-proof scripts that can be inspected by a human.
-As some \holfour theories are built in separate sessions, it is not possible
-to transfer \sml values directly, but instead tactic code needs to transferred 
-instead.
-To do this, the transferred information will be saved externally.
-
-\subsection{Globalization}
-%\subsubsection {Unfolding locally declared values}
-To transfer \sml values as code across theories in a conservative manner, we 
-need to 
-globalize \sml identifiers. To this end, we replace each identifier by some 
-\sml code containing all the information to obtain the value of this identifier 
-without knowing the context. The result is a standalone \sml code executable in 
-any \holfour theories which will be interpreted in the same way in the absence 
-of side effects. Two techniques are helpful to produce such code. The first one 
-is to prefix global identifiers by the module (and sub-modules) in which they 
-were declared. And the second one is to replace local identifiers 
-by their definitions. Prefixing and 
-unfolding are applied recursively in 
-the definitions.
-
-We present the tactical proof recording mechanism in details here because it 
-can help 
-another researcher to produce a tactical export in other interactive theorem 
-provers.
-We now enumerate common \sml constructions and explain 
-independently how we globalize them. We use a stack to keep track of previous 
-identifier declaration and their globalization.
-
-%We first start by mentioning additional 
-%details about module and value declarations.
-
-\subsubsection{Declaration}
-
-\paragraph{Reserved tokens}
-
-\paragraph{Module}
-
-\paragraph{Global identifier}
-
-\paragraph{Local identifier}
-
-To globalize a identifier, we include 
-the code of its declaration in the globalized expression. 
-Given a declaration \texttt{fun f v = e} or \texttt{val f = e}, the globalized 
-expression for the identifier \texttt{f} is \texttt{let fun f v = e' in f 
-end} respectively \texttt{let val f = e' in f end}
-The expression \texttt{e'} is the globalized version of \texttt{e}. That is to 
-say, it is the expression \texttt{e} where each of its tokens has been replaced 
-its globalized expressions as constructed in Section~\ref{}.
-
-In the case of value, we can simplify the globalized expression further and 
-replace it by \texttt{e'}.
-In the case of a function, \texttt{f} and its argument \texttt{v} are binders. 
-So, we prevent their globalization during the construction of \texttt{e'}.
-%This approach is slightly more verbose than using an 
-%anonymous function construction. But it is prettier as 
-%we regroup declarations at the top of the proof expression before returning it
-%to the user (see Section~\ref{sec:proofdisplay}).
-
-\paragraph{Bound variables}
-A pattern is an \sml expression that 
-contains \sml 
-constructors (or exceptions in the case of \texttt{handle}) and binding 
-variables. Our algorithm can recognize cases where pattern matching appears and 
-globalize the expression appropriately.  Let \texttt{p} be a pattern 
-and 
-\texttt{e} be abd expression, the following constructions can 
-occur: \texttt{val p = e}, \texttt{case \_ of p => e}, \texttt{fun f p = e}, 
-\texttt{fn p => e}, \texttt{handle p => e}. A binding identifiers in \texttt{p}
-is a token that is not a constructor (or exception). To produce the globalized 
-version of each construction, we globalize constructors (or exception) in the 
-pattern \texttt{p} and we protect binding variables in \texttt{e} to prevent 
-their modification. When a value identifier \texttt{x} is declared in a pattern 
-\texttt{p}, the globalized expression \texttt{e'} for \texttt{x} becomes 
-\texttt{let val p = e' in x 
-end}.
-
-\paragraph{Infixity}                                                            
-In \sml, it is possible to declare infixity rules for a string $s$ that 
-is not prefixed by a module. The infixity applies only when an identifier 
-exactly match $s$.
-
-\subsubsection{Expression}
-
-\paragraph{Reserved tokens}
-
-\paragraph{Modules}
-Declaring a module is made through the keyword 
-\texttt{open}. We can fetch 
-the values declared in a module \texttt{m} by first erasing all elements from 
-\texttt{PolyML.NameSpace} and then calling \texttt{open m}. Additionally, it 
-gives us the information of whether a value is a constructor or an exception. 
-This will prove to be particularly useful to handle pattern matching 
-construction correctly. During the opening of a module, nested modules can be 
-declared and push to the stack. In this way, we can globalize them by prefixing 
-them by their parent modules when they appear in the program.
-
-\paragraph{Global identifier}
-Global values and functions are prefixed by the modules and submodules where 
-they were declared. This information obtained by looking up the first 
-identifier that corresponds to the values at the top of the stack.
-
-\paragraph{Local identifier}
-Each local identifier is replaced by its globalized expression.
-
-We do not unfold local and global values of type theorem. Indeed, we do not 
-want to 
-unfold proofs into other proofs. Instead, we fetch theorems by looking for its 
-formula in the global database of 
-theorems. If it does not exists, we consider the tactic to be local and not 
-transferable to other theories.
-
-\paragraph{Bound variables}
-Binding variables are not globalized
-
-\paragraph{Infixity}
- Therefore, an identifier \texttt{x} could have been treated 
-as infix in an original expression \texttt{E} but when 
-we substitute it locally by its 
-globalized version \texttt{e'}, \texttt{e'} is now considered with prefix 
-infixity in \texttt{E} which makes \texttt{E} an invalid expression. Therefore, 
-because we 
-want to avoid global changes to \texttt{E}, we encapsulate \texttt{e'} by two 
-newly 
-defined 
-infix guards to reproduce 
-the original behaviour.
-In total there are 20 possible infix settings \texttt{x} can have in \sml.
-Those infix guards are defined to have the 
-same 
-precedence level $\texttt{P}\in[0,9]$ and same associativity 
-$\texttt{A}\in[l,r]$ (right or left) as \texttt{x}.
-Definitions for the infix guards depend on their associativity:
-
-\begin{definition}(Code of infix guards)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-fun a infixlX_open f = fn x ==> f (a,x)      fun g infixlX_close b = g b
-fun a infixrX_open g = g a      fun f infixrX_close b = fn x ==> f (x,b) 
-\end{lstlisting}
-\end{definition}
-
-The previously decribed transformation affects each identifier in a theory 
-file. However, when creating the recording theory file, only proof scripts are 
-replaced by their 
-globalized version and the rest of the code is left unchanged. The effect on 
-the proof script contained in the running example is shown in 
-Example~\ref{ex:global}.
-
-\begin{example}\label{ex:global} (Globalization)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-Tactic.STRIP_TAC 
-  infixl0_open boolLib.THEN infixl0_close
-Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC
-  infixl0_open boolLib.THEN infixl0_close
-Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
+\begin{example}\label{ex:r0}(Proof script extraction)
+\small
+\begin{lstlisting}[language=SMLSmall]
+STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
 \end{lstlisting}
@@
 
-\subsection{Tactic wrap}
-
-We now restrict the definition of a proof script to being the third 
-argument of a call to 
-\texttt{store\_thm}. And we precisely define the 
-modification applied to the theory file. A modified theory file is identical to 
-the original theory file except every proof scripts \texttt{p} is
-replaced by \texttt{Split "p'"} where \texttt{Split} is the splitting function 
-to be defined and \texttt{"p'"} is the string representation of the globalized 
-version of the proof script 
-\texttt{p}. 
-.
-\begin{example}(Proof script transformation)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
-original file: val id = store_thm (name, term, p)
-modified file: val id = store_thm (name, term, Split "p'")
+\begin{example}\label{ex:global} (Globalization)
+\begin{lstlisting}[language=SMLSmall]
+Tactic.STRIP_TAC THEN
+Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC THEN
+Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
 \end{lstlisting}
@@
 globalized infixity, we just have to verify that the expression 
-is of type tactic and that an infix guard is not a top symbol of the expression.
+is of type tactic and that an infix guard is not a top symbol of the 
+expression.
 To determine that, we run the \polyml compiler on \texttt{"p'"} to give us a 
@@
 
-\begin{example} (Wrap)
+\begin{definition}(Tactic unit)
+A tactic unit is an \sml expression of type tactics that does not contain a 
+infix operator at its root.
+\end{definition}
+
+\begin{example}\label{ex:wrap} (Tactic unit wrapping)
 \begin{lstlisting}[language=SMLSmall,frame=tb]
-R "Tactic.STRIP_TAC"
-  infixl0_open boolLib.THEN infixl0_close
-R "Prim_rec.INDUCT_THEN (DB.fetch \"list\" \"INDUCT\") Tactic.ASSUME_TAC"
-  infixl0_open boolLib.THEN infixl0_close
+R "Tactic.STRIP_TAC" THEN
+R "Prim_rec.INDUCT_THEN (DB.fetch \"list\" \"INDUCT\") Tactic.ASSUME_TAC" THEN
 R "Rewrite.ASM_REWRITE_TAC 
-     [DB.fetch \"list\" \"MAP\", DB.fetch \"list\" \"APPEND\"]"
+  [fff \"list\" \"MAP\", fff \"list\" \"APPEND\"]"
 \end{lstlisting}
@@
 convert the tactic string back into a tactic via the function 
-\texttt{read\_tactic}. This function will be used whenever we want to execute a 
-tactic string. So in the next sections, we will not make a difference between a 
-tactic and its code and call both of them tactics.
+\texttt{tactic\_of\_sml}.
 
@@
 \begin{lstlisting}[language=SMLSmall,frame=tb]
-fun R tactic_string goal = 
-  (save (tactic_string, goal); tactic_of_sml tactic_string goal)
+fun R stac goal = (save (stac,goal); tactic_of_sml stac goal)
 \end{lstlisting}
@@
 
-\subsection{Data Transfer}
-We describe here the format preferred to save and load the 
-data recorded by the function \texttt{save}. 
-During saving, we add fields to the input pair $(tactic,goal)$ to record 
-precomputed information in order to make the prediction faster. The final 
-feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
-The time $time$ taken by $tactic$ is not currently used by our 
-predictor. The list of goals $goal list$ is the output of $tactic$ on 
-$goal$. And the features $features$ of the goal $goal$ are extracted and hashed 
-as described in Section~\ref{sec:features}.
-
-\begin{remark}
-To simplify the description of our machine learning algorithms in future 
-sections, we consider a tactic feature vector to be a pair $(tactic,goal)$ and 
-only 
-explicitly rely on the 
-other fields when necessary.
-\end{remark}
-
-We now discuss the type of file we want to store our feature vectors on and
-how we plan to write and read \sml values. We start by taking 
-inspiration from an existing implementation in \holfour.
-
-To avoid the computational cost of re-proving every theorems at the begining of 
-each session, theorem values in \holfour are stored.
-They are encoded them in to a string by calling \texttt{Term.write\_raw}.
-Conversely, the function \texttt{Term.read\_raw} transforms the string back 
-into the original term.
-To have theorems binded to sml identifiers, it is necessary to create a 
-\sml file where those binders are declared. And for simplicity sake, all theory 
-data was fed into this \sml file. However, it is
-more efficient to store theory data into a text file and make a custom 
-parser which can be run faster than the \polyml compiler.
-Thanks to this insight, we have 
-modified the way in which \holfour theories are saved to include only \sml 
-bindings in a \sml file. All other theory data is now stored separately 
-into a text file. After integrating the changes, \holfour theories load three 
-times faster in average.
-
-Our \tactictoe data management system do not need to create new \sml
-bindings for the user. So we can store all \sml values into a text file which 
-can be loaded rapidly.
-Storing time, tactic string and features is easy since they are respectively 
-floats, strings and integers. And because a goal is composed of terms, we can 
-re-use the \holfour theory data reading and writing functions 
-for the two other fields.
-
-Finally, we mention some limitations that comes with 
-the need to record tactics as a string compared to being able to store a \sml 
-value. Access to the name of the tactic enhances machine 
-learning ability and user interaction. Yet, because some tactics 
-depends on the state of \holfour, for example the state of the global 
-simpset, a tactic is re-parsed before each proof attempt to be up-to-date.
-As a result, a tactic may have different effect on goals than suggested by the 
-database which could compromise the prediction accuracy. In practice, this 
-should rarely happen. Indeed, most tactics are stateless and an update on a 
-stateful tactic is usually improving it.
+The function save(stac,goal) save the pair to disk increasing the number of 
+tactic-goal pairs in our database.
+
+%\subsection{Data Transfer}
+%We describe here the format preferred to save and load the 
+%data recorded by the function \texttt{save}. 
+%During saving, we add fields to the input pair $(tactic,goal)$ to record 
+%precomputed information in order to make the prediction faster. The final 
+%feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
+%The time $time$ taken by $tactic$ is not currently used by our 
+%predictor. The list of goals $goal list$ is the output of $tactic$ on 
+%$goal$. And the features $features$ of the goal $goal$ are extracted and 
+%hashed 
+%as described in Section~\ref{sec:features}. 
+%
+%\begin{remark}
+%To simplify the description of our machine learning algorithms in future 
+%sections, we consider a tactic feature vector to be a pair $(tactic,goal)$ and 
+%only 
+%explicitly rely on the 
+%other fields when necessary.
+%\end{remark}
+%
+%We now discuss the type of file we want to store our feature vectors on and
+%how we plan to write and read \sml values. We start by taking 
+%inspiration from an existing implementation in \holfour.
+%
+%To avoid the computational cost of re-proving every theorems at the begining 
+%of 
+%each session, theorem values in \holfour are stored.
+%They are encoded them in to a string by calling \texttt{Term.write\_raw}.
+%Conversely, the function \texttt{Term.read\_raw} transforms the string back 
+%into the original term.
+%To have theorems binded to sml identifiers, it is necessary to create a 
+%\sml file where those binders are declared. And for simplicity sake, all 
+%theory 
+%data was fed into this \sml file. However, it is
+%more efficient to store theory data into a text file and make a custom 
+%parser which can be run faster than the \polyml compiler.
+%Thanks to this insight, we have 
+%modified the way in which \holfour theories are saved to include only \sml 
+%bindings in a \sml file. All other theory data is now stored separately 
+%into a text file. After integrating the changes, \holfour theories load three 
+%times faster in average.
+%
+%Our \tactictoe data management system do not need to create new \sml
+%bindings for the user. So we can store all \sml values into a text file which 
+%can be loaded rapidly.
+%Storing time, tactic string and features is easy since they are respectively 
+%floats, strings and integers. And because a goal is composed of terms, we can 
+%re-use the \holfour theory data reading and writing functions 
+%for the two other fields.
+%
+%Finally, we mention some limitations that comes with 
+%the need to record tactics as a string compared to being able to store a \sml 
+%value. Access to the name of the tactic enhances machine 
+%learning ability and user interaction. Yet, because some tactics 
+%depends on the state of \holfour, for example the state of the global 
+%simpset, a tactic is re-parsed before each proof attempt to be up-to-date.
+%As a result, a tactic may have different effect on goals than suggested by the 
+%database which could compromise the prediction accuracy. In practice, this 
+%should rarely happen. Indeed, most tactics are stateless and an update on a 
+%stateful tactic is usually improving it.
 
@@
 
-When \tactictoe finds a proof of a goal, it returns a search tree 
-$\mathfrak{T}$ where the root node containing this goal is solved.
+When \tactictoe finds a proof of a goal $g_{root}$, it returns a search tree 
+$\mathfrak{T}$ where the root node containing $g_{root}$ is solved.
 In order to transform this search tree into a proof script, we need to extract 
@@
 tactic from the list is called.
-Here is a formal description of the transformation defined by mutual 
+
+Let $T_{sol}$ be a partial function that from a goal $g$ returns a tactic $t$ 
+for which $A(g,t)$ is solved in $\mathfrak{T}$.
+The proof extraction mechanism is defined by mutual 
 recursion on goals and nodes of $\mathfrak{T}$ by the respective function 
-$P_goal$ and $P_node$:
-
-\begin{equation*}
-P_goal(g) =_{def}
-\begin{cases}
-T(g) & \text{if A(g,T(g)) = \emptyset},\\
-T(g) & \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
-\end{cases}
-%
-%
-Proof(a) =_def{def}
-\begin{cases}
-tac & \text{if $t$ is a root and $g$ the unique goal of the root},\\
-tac & \text{if $t$ is a leaf},\\
-tac\ \texttt{THENL}\ [P(a_1),\ldots,t_{n-1}] & \text{otherwise.}
-\end{cases}
-\end{equation*}
-
-where $tac$ is the tactic that produced the node, $c$ is the
-only successful child of the root and $c_0, \ldots, c_n$ are the 
-children of the node produced by the successful tactic.
-
-The readability of the created proof scripts is improved, by replacing
-replacing  \texttt{THENL} by \texttt{THEN} when the list has length 1.
+$P_{goal}$ and $P_{node}$:
+
+\begin{align*}
+P_{goal}(g) &=_{def} T_{sol}(g)\ \texttt{THENL}\ P_{node}(A(g,T_{sol}(g)))\\
+P_{node}(a) &=_{def} [P_{goal}(g_1),\ldots,P_{goal}(g_n)]\ \ \ \text{with}\ 
+G(a) = g_1,\ldots,g_n\\
+\end{align*}
+
+The extracted proof script of $\mathfrak{T}$ is $P_{goal}(g_{root})$.
+We minimally improve it by subtituting \texttt{THENL} by \texttt{THEN} when the 
+list of goals is a singleton and removing \texttt{THENL []} during the 
+extraction phase.
 Further post-processing such as
-removing unnecessary tactics and theorems has been developed and
+eliminating unnecessary tactics and theorems has been developed and
 improve the user experience greatly~\cite{DBLP:conf/sefm/Adams15}.
-
-\paragraph{Minimizing length and time of the proof} 
-First a weak minimization is possible when processing the final proof. If the 
-tactic \texttt{A THEN B} appears in the proof and has the same effect as 
-\texttt{B} then we can drop \texttt{A} from the proof.
-Stronger minimization can be obtained by rerunning the prover with feature 
-vectors containing only tactics 
-from the proof. To obtain a shorter proof, the exploration coefficient should be
-increased and the policy coefficient lowered to encourage exploration in the 
-width of the proof tree. Additionally, by giving a weaker evaluation to slow 
-tactics, we can aim for faster proofs.
+\todoi{Not sure if this reference is appropriate anymore}
+
+\paragraph{Minimizing length of the proof} 
+A fast and simple minimization is applied when processing the final proof. If 
+the tactic \texttt{A THEN B} appears in the proof and has the same effect as 
+\texttt{B} then we can replace \texttt{A THEN B} by \texttt{B} in the script.
+Optionally, stronger minimization can be obtained 
+by rerunning \tactictoe with a low prior 
+policy coefficient, no prior evaluation and a database of tactics that contains 
+tactics from the discovered proof only.
 
 \paragraph{Minimizing tactic arguments}
-One of the argument that may contain unnecessary information are lists.
-Let $t$ be a tactic applied to a goal $g$ containing a list $l$ as argument. 
-And let $t'$ be the tactic $t$ where one element $e$ of $l$ as be removed. If 
-$t$ and $t'$ have the same effect on $g$ then $t'$ can replace $t$ in the final 
-proof.
-
+Let $t$ be a tactic applied to a goal $g$ containing a list $l$ (of theorems) 
+as argument. And let $t'$ be the tactic $t$ where one element $e$ of $l$ as be 
+removed. If $t$ and $t'$ have the same effect on $g$ then $t'$ can replace $t$ 
+in the final proof. This process is repeated for each element of $l$.
 This is a generalization of the simplest method used for minimizing a list of 
-theorems in ``hammers'' \cite{hammers4qed}. The fact that lists are generally small
-make this approach feasible.
+theorems in ``hammers'' \cite{hammers4qed}.
 
@@
 information to guarantees that each \sml subterm is interpreted in the way in 
-any context. Our approach here is to strip this information from the tactic and 
-test if the tactic has the same effect as the original one to guarantee that we 
-are not changing the proof. For efficiency, 
-this is performed on each tactic step separately. We strip module prefixes, 
-local declaration. If local declaration can't be 
-stripped, we try to group all local declarations under a single "let" binding 
-which improves readability and prevents duplication of this declarations.
-
-\todoi{Add an example}
-
-%\paragraph{Step by step exploration.}
-%Because a human may have better insight on which tactic to chose next, we also 
-%provides an interactive mode, where the user is presented the best $n$ 
-%predicted tactics for a goal and to be chosen from. The number $n$ is 
-%parameter 
-%that can be chosen by the user. Those tactics are curated so that only 
-%productive tactics are returned. (define productive)
+any context. Since we return the proof at a point where the \sml interpreter is 
+a specific state, we can strip unnecessary information in the form of 
+module prefixes. If possible, we group all local declarations in the 
+script under a single \texttt{let} binding at the start of the script. And we 
+replace extended terms by their quoted version.
+All in all, if a prettified tactic $t_p$ has the same effect its original 
+$t_o$, we replace $t_p$ by $t_o$ in the proof.\\
+
+The total effect on the readability of a proof script is depicted in 
+Example~\ref{ex:pretty}.
+\todo{run without prettification}
+\begin{example}\label{ex:pretty}\ \\
+Extracted proof script:
+\begin{lstlisting}[language=SMLSmall]
+\end{lstlisting}
+Minimized and prettified proof script:
+\begin{lstlisting}[language=SMLSmall]
+\end{lstlisting}
+\end{example}
 
 \section{Case Study}
-\todoi{Test tactictoe find new examples on new goals and manual exploration 
-mode.}
 In this section, we compare \tactictoe proof script with human proof script.

### 2018-01-31 tactictoe_jv.tex
recording: wip

@@
 Because of the large number of \sml constructions, we only describe the effect 
-of these steps in a running example that is representative of a typical proof 
-script. And we comment on the main steps that transforms the original proof 
-script into a modified version that is able to record the code of a tactic and 
-the goal(s) it processed during the proof. We omit more involved constructions 
-such as the globalization of infix operators.
+of these steps on a running example that contains selected parts of the theory
+of lists.
+
+%We comment on the treatment of global and local identifiers but  omit more 
+%involved constructions 
+%such as the globalization of infix operators.
 
@@
 
+The first line of this theory open modules (called structures in \sml). Each of 
+them contains a list of global identifiers which become directly accessible in 
+the rest of the theory.
+A local identifier \texttt{LIST\_INDUCT\_TAC} is declared after, which is a 
+tactic that performs induction on lists. And even further, the theorem 
+\texttt{MAP\_APPEND} is proven.
+The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
+the goal is split into a base case and a inductive case. And finally, both of 
+theses cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which 
+rewrites assumptions with the help of theorems previously declared in this 
+theory.
+
+We first parse the theory to extract proof scripts. Each of them is found in 
+the third argument of a call to \texttt{store\_thm}. The result of proof script 
+extraction for the running example in presented in Example~\ref{ex:r0}.
+
 \begin{example}\label{ex:r0}(Proof script extraction)
@@
 
-\begin{example}\label{ex:global} (Globalization)
+In the next phase, we globalize identifiers of the proof scripts. 
+Infix operators such as \texttt{THEN} need to be processed in a special way so 
+that they can keep their infixity after globalization. For simplicity, the 
+globalization of infix operators is omitted in Example~\ref{ex:r1}.
+In this example, the three main cases that can happen during the globalization
+are depicted. The first one is the globalization of identifiers declared in 
+modules. The global identifiers \texttt{STRIP\_TAC} and 
+\texttt{ASM\_REWRITE\_TAC} are prefixed by their module \texttt{Tactic}. In 
+this way, they will still be interpretable whether \texttt{Tactic} was open or 
+not. The local identifier \texttt{LIST\_INDUCT\_TAC} is replaced by its 
+definition which happens to contain two global identifiers.
+The previous paragraph describe the globalization for all identifiers except 
+local theorems.  We do not replace a local
+theorem by its definition because we want to avoid unfolding proof scripts into 
+other proof scripts.
+If the theorem is stored in the database of the current theory, we 
+can obtain the theorem value by calling \texttt{DB.fetch}. This happens for the 
+three theorems in our script. The internal database of \holfour is available 
+across \holfour developments. 
+Otherwise the globalization process fails and the local theorem identifier is 
+kept unchanged. A recorded tactic with an unchanged local theorem as argument 
+is only interpretable inside the current theory.
+
+\begin{example}\label{ex:r1} (Globalization)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{example}
+
+\todo{rewrite this paragraph}
 
@@
 
+\todo{rewrite this paragraph}
+
 During its execution, the recording function \texttt{R} (as coded in 
@@
 
-The function save(stac,goal) save the pair to disk increasing the number of 
-tactic-goal pairs in our database.
-
-%\subsection{Data Transfer}
-%We describe here the format preferred to save and load the 
-%data recorded by the function \texttt{save}. 
-%During saving, we add fields to the input pair $(tactic,goal)$ to record 
-%precomputed information in order to make the prediction faster. The final 
-%feature vector is the tuple $(tactic,time,goal,goal\_list,features)$.
-%The time $time$ taken by $tactic$ is not currently used by our 
-%predictor. The list of goals $goal list$ is the output of $tactic$ on 
-%$goal$. And the features $features$ of the goal $goal$ are extracted and 
-%hashed 
-%as described in Section~\ref{sec:features}. 
-%
-%\begin{remark}
-%To simplify the description of our machine learning algorithms in future 
-%sections, we consider a tactic feature vector to be a pair $(tactic,goal)$ and 
-%only 
-%explicitly rely on the 
-%other fields when necessary.
-%\end{remark}
-%
-%We now discuss the type of file we want to store our feature vectors on and
-%how we plan to write and read \sml values. We start by taking 
-%inspiration from an existing implementation in \holfour.
-%
-%To avoid the computational cost of re-proving every theorems at the begining 
-%of 
-%each session, theorem values in \holfour are stored.
-%They are encoded them in to a string by calling \texttt{Term.write\_raw}.
-%Conversely, the function \texttt{Term.read\_raw} transforms the string back 
-%into the original term.
-%To have theorems binded to sml identifiers, it is necessary to create a 
-%\sml file where those binders are declared. And for simplicity sake, all 
-%theory 
-%data was fed into this \sml file. However, it is
-%more efficient to store theory data into a text file and make a custom 
-%parser which can be run faster than the \polyml compiler.
-%Thanks to this insight, we have 
-%modified the way in which \holfour theories are saved to include only \sml 
-%bindings in a \sml file. All other theory data is now stored separately 
-%into a text file. After integrating the changes, \holfour theories load three 
-%times faster in average.
-%
-%Our \tactictoe data management system do not need to create new \sml
-%bindings for the user. So we can store all \sml values into a text file which 
-%can be loaded rapidly.
-%Storing time, tactic string and features is easy since they are respectively 
-%floats, strings and integers. And because a goal is composed of terms, we can 
-%re-use the \holfour theory data reading and writing functions 
-%for the two other fields.
-%
-%Finally, we mention some limitations that comes with 
-%the need to record tactics as a string compared to being able to store a \sml 
-%value. Access to the name of the tactic enhances machine 
-%learning ability and user interaction. Yet, because some tactics 
-%depends on the state of \holfour, for example the state of the global 
-%simpset, a tactic is re-parsed before each proof attempt to be up-to-date.
-%As a result, a tactic may have different effect on goals than suggested by the 
-%database which could compromise the prediction accuracy. In practice, this 
-%should rarely happen. Indeed, most tactics are stateless and an update on a 
-%stateful tactic is usually improving it.
+The function \texttt{save} writes the tactic-goal pairs to disk increasing the 
+number of entries in our database of tactics. The function 
+\texttt{tactic\_of\_sml} interprets \sml code \texttt{stac}. The tactic is then 
+applied to the input goal which replicates the original behavior.
 
 \section{Proof Presentation}\label{sec:proofdisplay}
-
-\todoi{Motivation: + minimal effort to integrate the tactictoe results in proof 
-scripts.}
 
@@
 combinators.
-\todoi{Define tactic combinators?}
+
 By the design of the search, a single tactic combinator, \texttt{THENL}, is 
@@
 improve the user experience greatly~\cite{DBLP:conf/sefm/Adams15}.
-\todoi{Not sure if this reference is appropriate anymore}
 

### 2018-02-01 tactictoe_jv.tex
recording: wip

@@
 formally described.
-\item A tactic abstraction method is introduced. It enable us to create 
+\item A tactic abstraction method is introduced. It enables us to create 
 more general and flexible tactics where tactic arguments are dynamically 
@@
 
-\paragraph{Datasets: optimization and validation}\todo{rephrase to something 
-more 
-coherent}
+\todoi{rephrase to something more coherent}
+\paragraph{Datasets: optimization and validation}
+
 All top-level theorems from the standard library are considered with the 
@@
 \item
-It is probably difficult (even impossible) to transfer values 
+It is difficult (probably impossible) to transfer \sml values 
 between theories as they are not run in the same \holfour session. In contrast, 
@@
 Infix operators such as \texttt{THEN} need to be processed in a special way so 
-that they can keep their infixity after globalization. For simplicity, the 
+that they keep their infixity status after globalization. For simplicity, 
+the 
 globalization of infix operators is omitted in Example~\ref{ex:r1}.
@@
 can obtain the theorem value by calling \texttt{DB.fetch}. This happens for the 
-three theorems in our script. The internal database of \holfour is available 
-across \holfour developments. 
+three theorems in our script. The internal database of \holfour theorems is 
+available across \holfour developments. 
 Otherwise the globalization process fails and the local theorem identifier is 
@@
 
-\todo{rewrite this paragraph}
-
-The intended effect of the splitting function is to detect and wrap \sml 
-expressions of type tactics. Its aim is to is encapsulate each tactic 
-unit $t$ within a recording function $R$. This function $R$ is designed to 
-observe what goals $t$ receives at run time. Since there can be maybe nested 
-level of tactics, (i.e 
-tactics that contain tactics as subexpressions), we choose one that is
-close to what a human understand to be a tactic unit. 
-Most tacticals are infix, and we do not want to consider combination of 
-tactics to be a tactic unit. For \texttt{Split}, a tactic unit is a 
-subterm of 
-type tactics that do not contain a infix operator as its top symbol. Thanks to
-globalized infixity, we just have to verify that the expression 
-is of type tactic and that an infix guard is not a top symbol of the 
-expression.
-To determine that, we run the \polyml compiler on \texttt{"p'"} to give us a 
-detailed code tree of the expression. Each tactic unit is then substituted 
-by a wrapped version to produce the final proof script.
+Running the globalized version of a proof script will have the exact same 
+effect as the original. But since we want to extract information from this 
+script in the form tactics and their input goals, we modify it further.
+In particular, we need to define at which level we should record the tactics of 
+the proof scripts. The simplest idea would be to record all \sml subexpressions 
+of type tactics. However, it will damage the quality of our data by 
+dramatically increasing the number of tactic associated with a single goal.
+Imagine a proof script of the form \texttt{A THEN B THEN C} then the tactics
+\texttt{A}, \texttt{A THEN B} and \texttt{A THEN B THEN C} would be valid 
+advice for something close to their common input goal. The tactic  
+\texttt{A THEN B THEN C} is likely to be specific. In contrast, we can consider 
+the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A} 
+until 
+\texttt{A} has no effect. In this situation, it is often preferable to record 
+only \texttt{REPEAT A} and any call to \texttt{A}. Otherwise during the 
+proof search, we would have to make a branches created after each call of 
+\texttt{A}. 
+To sum up, we would like to record only the most general tactics which 
+make the most progress on a goal. As a trade-off between these two objectives,
+we split the proofs into tactic units.
 
@@
 
+Because these tactic units are constructed from visual information present in 
+the proof script, they often represent what a human consider to be one step of 
+the proof.
+
+To produce the final recording script, we encapsulate each tactic unit 
+in a recording function \texttt{R} (see Example~\ref{ex:wrap}). In order to 
+record 
+tacitcs in all \holfour theories, we replace the original proof by the 
+recording script in each theory and rebuild the \holfour library.
+
 \begin{example}\label{ex:wrap} (Tactic unit wrapping)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 R "Tactic.STRIP_TAC" THEN
@@
 
-\todo{rewrite this paragraph}
-
-During its execution, the recording function \texttt{R} (as coded in 
-Definition~\ref{ex:record}) 
-saves the tactic 
-string and the goal, verifies that the tactic string as the same effect as the 
-original tactic and apply it to continue the proof. To perform this task, we 
-convert the tactic string back into a tactic via the function 
-\texttt{tactic\_of\_sml}.
+At run time the function \texttt{R} is designed to observe what input goals a
+tactic  receives without changing its output. The implementation of \texttt{R} 
+is presented in Example~\ref{ex:record}.
 
 \begin{example}\label{ex:record} (Code of the recording function)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 fun R stac goal = (save (stac,goal); tactic_of_sml stac goal)
@@
 
-The function \texttt{save} writes the tactic-goal pairs to disk increasing the 
+The function \texttt{save} writes the tactic-goal pair to disk increasing the 
 number of entries in our database of tactics. The function 
-\texttt{tactic\_of\_sml} interprets \sml code \texttt{stac}. The tactic is then 
-applied to the input goal which replicates the original behavior.
+\texttt{tactic\_of\_sml} interprets the \sml code \texttt{stac}. The tactic is 
+then applied to the input goal to replicate the original behavior.
+After all modified theories are rebuild, each call to a wrapped tactic in a
+proof script is recorded as a pair containing its globalized code and its 
+input goal.
 
@@
 Goal: \texttt{IS\_GCD\_UNIQUE} in theory \texttt{pred\_set}).
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 !a b c d. is_gcd a b c wedge is_gcd a b d ==> (c = d)
@@
 Original \holfour proof script
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 PROVE_TAC[IS_GCD, DIVIDES_ANTISYM]
@@
 \tactictoe proof script
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 STRIP_TAC THEN 
@@
 Goal: \texttt{SURJ\_INJ\_INV} in theory \texttt{pred\_set})
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 !f s t. SURJ f s t ==> ?g. INJ g t s wedge !y. y IN t ==> (f (g y) = y)
@@
 Original \holfour proof script (2 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 REWRITE_TAC [IMAGE_SURJ] THEN
@@
 \tactictoe proof script (50 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 SRW_TAC [] [SURJ_DEF, INJ_DEF] THEN METIS_TAC []
@@
 Goal
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 !l1 l2 n x. LENGTH l1 <= n ==> 
@@
 Original \holfour proof script (63 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
   rw[] THEN simp[LIST_EQ_REWRITE] THEN Q.X_GEN_TAC `z` THEN
@@
 \tactictoe proof script (17 milliseconds)
-\begin{lstlisting}[language=SMLSmall,frame=tb]
+\begin{lstlisting}[language=SMLSmall]
 Induct_on `l1` THENL [SRW_TAC [] [],

### 2018-02-01 tactictoe_jv.tex
recording: wip

@@
 
-% Because our aim is to 
-% learn how to prove from existing examples, we record the proofs at 
-% an intermediate tactic level closest to the one presented to 
-% a user. Since human developers build almost all their proofs from tactics, 
-% we believe this is an appropriate level to apply machine learning on.
-
-%More information is contained in tactical scripts than 
-%in top-level dependencies on which was based machine-learning guidance in 
-%interactive theorem proving so far. Learning from atomic inferences would 
-%require better predictions and abstractions of rules to handle the larger 
-%search space.
-
-%We first discuss parsing proof scripts and identifying 
-%individual tactic calls. We next show how to record tactics calls are recorded 
-%with their corresponding goals using a modified proof script.
-%The recorded data which consists of tactic-goal pairs creates the raw 
-%knowledge 
-%necessary for our tactic selection algorithm.
-
 Rather than relying on the underlying programming language, we parse the
@@
 to re-use the code of a tactic recorded in one theory in an 
-other. We also would like to make sure that the code is interpretable and its 
-interpretation does not change in the other theory.
-
-Even when developing one theory, the context is continually changing: the open 
-modules are different, global and local identifiers are defined and the local 
-ones can become out of scope. Therefore, it is unlikely that code extracted 
-from proof scripts is interpretable in any part of \holfour.
-To solve this issue, we prefix each global identifier by the module where it 
-was declared and we recursively unfold local declaration until a local 
-identifier can be expressed by through code containing only global identifiers.
+other. We also would like to make sure that the code is interpretable and that 
+its interpretation does not change in the other theory.
+
+Even when developing one theory, the context is continually changing:
+modules are open and local identifiers are defined. Therefore, it is unlikely 
+that code extracted from proof scripts is interpretable in any other part of 
+\holfour without any post-processing.
+To solve this issue, we recursively replace each local identifier by its 
+definition until we are able to write any expression with global identifiers 
+only. And we prefix each global identifiers by its module.
 We call this process globalization. The result is a standalone \sml code 
-executable in any \holfour theories which will be interpreted in the same way 
-in the absence 
-of side effects.
-
-Therefore, we can guarantee that the behaviour of recorded stateless tactics 
-will not change. And because an update on a stateful tactic 
+executable in any \holfour theories.  In the absence of side effects, it is 
+interpreted in the same way across theories.
+Therefore, we can guarantee that the behavior of recorded stateless tactics 
+does not change. And because an update on a stateful tactic 
 generally increases its strength, our prediction algorithm based on past 
-examples will still be reasonably effective.
+examples is still reasonably effective in this case.
 

### 2018-02-02 tactictoe_jv.tex
interactive example

@@
 If a tactic produces no goals then its input goal is solved.
+Define tactical etc maybe import graphic that explain how tactic works form the 
+presentations.
 
@@
 proves a theorem stated in a theory.
-
-In order to apply machine learning to the existing proof scripts, we will need 
-to
-parse and modify them, which involves parsing and modification of \sml code.
-Acquiring access to the string representation of a tactic (code for the tactic) 
-allows us generating
-proof scripts that can be inspected by a human.
-As some \holfour theories are built in separate sessions, it is not possible
-to transfer \sml values directly, but instead tactic code needs to transferred 
-instead.
-To do this, the transferred information will be saved externally.
 
@@
 + could be also used to simplify existing proof scripts. }
+
+Interactive usage example.
+
+\begin{lstlisting}[language=SMLSmall]
+tactictoe ([], ``countable (UNIV:num set)``);
+let fun RWTAC thms = SRW_TAC [] thms in RWTAC [(fetch "pred_set" 
+"countable_def"), (fetch "pred_set" "INJ_DEF")] THEN Q.EXISTS_TAC `\\x.x` THEN 
+METIS_TAC [] end
+\end{lstlisting}
+
+\begin{lstlisting}[language=SMLSmall]
+tactictoe ([], ``count (n+m) DIFF count n = IMAGE ($+n) (count m)``);
+SRW_TAC [ARITH_ss] [(fetch "pred_set" "EXTENSION"), EQ_IMP_THM] THEN 
+Q.EXISTS_TAC `x - n` THEN SRW_TAC [ARITH_ss] []
+\end{lstlisting}
 
@@
 Its implementation,
-\tactictoe, achieves an overall performance of 59\% theorems on 7088 theorems  
+\tactictoe, achieves an overall performance of 66.2\% theorems on 7164 
+theorems  
 of the \holfour standard library surpassing \eprover with 

### 2018-02-05 tactictoe_jv.tex
updating todo
Minor tweak in abstract

@@
 On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.2\%
-of 7164 theorems of the standard library where as 
+of the 7164 theorems in HOL4's standard library whereas
 Eprover solves 34.5\%. The success rate rises to 69.0\% by combining the
-result of both provers.
+results of both provers.
 \end{abstract}
@@
 presentations.
+
+mention proof obligations (use this word) return by a tactic.
 

### 2018-02-05 tactictoe_jv.tex
Tweak some grammar and formatting

@@
  appearing in the search tree.
-\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$. It takes a
+\item $T$ is a total function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It takes a
 goal and return the set of tactics already applied to this goal.
-\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$. It takes a
-node a return the set of goals in this node.
+\item $G$ is a total function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$. It takes a
+node and returns the set of goals in this node.
 
@@
 application of the tactic to the goal, if this tactic has been applied to this
-goal. Its support $Support(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g)
+goal. Its support $\mathit{Support}(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g)
 \rbrace$.
-\item If $(g,t) \in Support (A)$, then $t(g) = G(A(g,t))$.
+\item If $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$.
 \end{itemize}
@@
 is the set of nodes produced by tactics applied to $g$:
-  \[Children(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]
+  \[\mathit{Children}(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]
 
@@
 Extended children of $a$ by:
-  \[ExtChildren(a) = \bigcup_{g \in G(a)} Children(a,g) \]
+  \[\mathit{ExtChildren}(a) = \bigcup_{g \in G(a)} \mathit{Children}(a,g) \]
 
@@
 \begin{align*}
-Descendant^{0}(a) &=_{def} \lbrace a \rbrace \\
-Descendant^{n+1}(a) &=_{def} \bigcup_{a' \in Descendant^{n}(a)}
-ExtChildren(a') \\
-Descendant(a) &=_{def} \bigcup_{n \in \mathbb{N}} Descendant^n(a)\\
+\mathit{Descendant}^{0}(a) &=_{def} \lbrace a \rbrace \\
+\mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
+\mathit{ExtChildren}(a') \\
+\mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
 \end{align*}
@@
 closest to $l$. The
-number of positive example in this set is noted $Positive(r,l)$.
+number of positive example in this set is noted $\mathit{Positive}(r,l)$.
 During the proof search, we evaluate nodes through the list of goals they
@@
 \begin{align*}
-PriorEvaluation_r (a_i) &= \frac{Positive(r,G(a_i))}{r}\\
+\mathit{PriorEvaluation}_r (a_i) &= \frac{\mathit{Positive}(r,G(a_i))}{r}\\
 \end{align*}
@@
 The value of the $i^th$ child $a_i$ of a parent $p$ is determined by:
-\[Value(a_i) = CurEvaluation(a_i) + c_{exploration} * Exploration(a_i)\]
+\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} * \mathit{Exploration}(a_i)\]
 
@@
 
-\[CurEvaluation(a_i) =
-  \sum_{a' \in Descendant(a_i)} \frac{PriorEvaluation(a')} {card\
-  Descendant(a_i) + Failure(a_i)}\]
-
-The number $Failure(a_i)$ is the amount of failures that occurred during node
+\[\mathit{CurEvaluation}(a_i) =
+  \sum_{a' \in Descendant(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
+  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]
+
+The number $\mathit{Failure}(a_i)$ is the amount of failures that occurred during node
 extension from descendants of $a_i$ .
@@
 The exploration term is determined by the prior policy and the current policy.
-\[Exploration(a_i) = \frac{PriorPolicy(a_i)}{CurPolicy(a_i)}\]
-
-\[CurPolicy(a_i) = \frac{1 + Visit(a_i))}{\sqrt{Visit(p))}}\]
+\[\mathit{Exploration}(a_i) = \frac{\mathit{PriorPolicy}(a_i)}{\mathit{CurPolicy}(a_i)}\]
+
+\[\mathit{CurPolicy}(a_i) = \frac{1 + \mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
 The policy is approximatively the  percentage of time a node was visited.The
@@
 not beneficial. It may be because it breaks some symmetry in the
-$Value$ function.
+$\mathit{Value}$ function.
 Otherwise, it applies the best untested tactic $t$ on the open goal $g$ of this

### 2018-02-07 tactictoe_jv.tex
minor

--- 2018-02-02 tactictoe_jv.tex
@@
 
+mention proof obligations (use this word) return by a tactic.
+
 \holfour proofs are organized in theory files.
@@
 
+%sh stats.sh MNM_AE2I
+%273 203 3 67 0
+%thibault@grid02:~/big$ sh stats.sh MNM_AE2
+%273 203 3 67 0
+%thibault@grid02:~/big$ sh stats.sh MNM_AE4
+%273 201 3 69 0
+%thibault@grid02:~/big$ sh stats.sh MNM_AE4I
+%273 200 3 70 0
+%thibault@grid02:~/big$ sh stats.sh MNM_AE1
+%273 198 3 72 0
+%thibault@grid02:~/big$ sh stats.sh MNM_AE1I
+%273 200 3 70 0
+
+%thibault@grid02:~/big$ sh stats.sh MNM_ATPE
+%273 213 3 57 0
+%thibault@grid02:~/big$ sh stats.sh MNM_ATPM01
+%273 216 3 54 0
+%thibault@grid02:~/big$ sh stats.sh MNM_ATPM01E
+%273 218 3 52 0
+%thibault@grid02:~/big$ sh stats.sh MNM_ATPM02
+%273 212 3 58 0
+%thibault@grid02:~/big$ sh stats.sh MNM_ATPM04
+%273 212 3 58 0
+
+%thibault@grid02:~/big$ sh ftim.sh MNM_O40
+%127
+%9
+%8
+%2
+%2
+%148
+%thibault@grid02:~/big$ sh ftim.sh MNM_O10
+%124
+%15
+%5
+%3
+%1
+%148
+%thibault@grid02:~/big$ sh ftim.sh MNM_ON
+%108
+%16
+%6
+%3
+%6
+%thibault@grid02:~/big$ sh ftim.sh MNM_T005
+%124
+%13
+%5
+%2
+%3
+
 \begin{table}[ht]
@@
 \small
-\begin{tabular}{llc}
+\begin{tabular}{llcc}
 \toprule
- Technique & Parameters & Solved \\
+ Technique & Parameters & Solved (1s) & Solved (60s)\\
 \midrule
-Tactic time out & 0.02s & 154 (56.8\%)\\
-                & 0.05s (default) & 156 (56.8\%)\\
-                & 0.1s & 154 (56.0\%)\\
+Tactic time out & 0.02s && 154\\
+                & 0.05s (default) && 156\\
+                & 0.1s && 154\\
 \midrule
-Orthogonalization & none & 156 \\
-                  & $radius = 10$ & 156 \\
-                  & $radius = 20$ (default) & 156 \\ 
-                  & $radius = 40$ & 156 \\
-
+Orthogonalization & none & 105 & 156 \\
+                  & $radius = 10$ & 121 & 156 \\
+                  & $radius = 20$ (default) & 121 & 156 \\ 
+                  & $radius = 40$ & 124 & 156 \\
 \midrule
-Abstraction       & none (default)  & 156 (\%)\\ 
-                  & $theorems = 8$  & 195 (\%)\\
-                  & $theorems = 16$ & 199 (72.9\%)\\
-                  & $theorems = 32$ & 195 (\%)\\
+Abstraction       & none (default)  && 156\\ 
+                  & $theorems = 8$  && 195\\
+                  & $theorems = 16$ && 199\\
+                  & $theorems = 32$ && 195\\
 \midrule
-MCTS Policy & $c_{policy} = 0.4$ & 149 (\%)\\
-            & $c_{policy} = 0.5$ (default) & 155 (\%)\\
-            & $c_{policy} = 0.6$ & 153 (\%)\\
+MCTS policy & $c_{policy} = 0.4$ && 149\\
+            & $c_{policy} = 0.5$ (default) && 156\\
+            & $c_{policy} = 0.6$ && 153\\
+
 \midrule
-MCTS Evaluation & none (default) & 156 (50\%) \\
-				& $c_{exploration} = 0.25$ & 153 153 (50\%)\\ 
-				& $c_{exploration} = 0.5$ & 153 155 (50\%)\\ 
-				& $c_{exploration} = 1$ & 154 (50\%)\\
-				& $c_{exploration} =\sqrt{2}$ & 154 154 (50\%)\\ 
-				& $c_{exploration} = 2$ & 152 (50\%)\\ 
+MCTS evaluation & none (best abstraction) && 199\\
+				& $c_{exploration} = 1$ && 201\\ 
+				& $c_{exploration} = 2$ && 203\\ 
+				& $c_{exploration} = 4$ && 198\\
 \midrule
-ATP integration & none (default) & 156 (50\%)\\
-                & \metis 0.1s & 125 (50\%)\\
-                & \metis 0.2s & 125 (50\%)\\
-                & \metis 0.4s & 125 (50\%)\\ 
-                & asynchronous \eprover & 125 (50\%)\\
-                & \metis 0.1s + asynchronous \eprover & 190\\
-\midrule
-All & best parameters (without \eprover) & \\ 
+ATP integration & none (best MCTS evaluation) && 203\\
+                & \metis 0.1s && 216\\
+                & \metis 0.2s && 212\\
+                & \metis 0.4s && 212\\ 
+                & \eprover && 213\\
+                & \metis 0.1s + \eprover && 218\\
 \bottomrule
@@
 
-\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
-
-\todo{decide to keep this section or not}
-
-We investigate what are the limits of \tactictoe and what benefits additional 
-knowledge could provide. To this end, we will modify the set of features to 
-include features vectors that should not be known in a regular experiment. We  
-include some vectors from the recorded human proof and perform the proof 
-search 
-after the recording instead of before in all other experiments. This 
-experiment 
-is not fair and should only be compared to 
-experiments in the same settings.
-
-For clarity, we will name the set of feature vectors that can be recorded from 
-the current human proof $\mathbb{H}$ and the set of vectors recorded from 
-previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
-$\pi_1^1$ the projection on the 
-first component returning the label of a feature vector. And we define the 
-function $T$ that takes a tactic and return the list of its tokens.
-
-We will progressively add bigger and bigger 
-subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
-search success rates.
-
-Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
-consider:
-\begin{align*}
-\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
-   \rbrace \\
-\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
-\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
-\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
-t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
-\rbrace\\
-\end{align*}
-
-$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
-prediction for those tactics.
-
-\begin{table}[h]
-\centering\ra{1.3}
-\small
-\begin{tabular}{lll}
-\toprule
- & Additional features & Solved \\
-\midrule
-$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
-$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
-$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
-$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
-$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
-$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
-% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
-\bottomrule
-\end{tabular}
-\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
-with tactics from the proof of the tested theorem. The experiments was run with 
-a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
-the low 
-success rate of the
-default strategy.} 
-\end{table}
-
-And we observe in Table~\ref{tab:featue_param} a 
-slight increase compared to the default \tactictoe. We then included all 
-vectors that contains tactics that can be rebuild from tokens of
-tactics in our database. 
-Our search algorithm proved two proofs less that if it had perfect prediction 
-so it means that our prediction algorithm (policy) is doing a good  job at 
-least in this early settings where proofs are easy. 
-Suprisingly, being able to recombine tokens of tactics from the database is 
-barely sufficient to prove more than half of the theorems in this dataset. The 
-Therefore, additionally including in tactics theorems and terms that did not 
-appear in previous tactics is essential. Creating from scratch other arguments 
-may not help as much. This is why we restricted ourselves to predicting 
-arguments of theorem list type and the term type during tactic synthesis. 
-Finally our proof search algorithm sometimes fails even if it knows all the 
-tactic necessary. The two main reasons are that either a crucial tactic 
-exceeds its timeout or that the human proof is too deep to be rediscovered 
-(more than 10 steps long).In all those experiments orthogonalization is 
-performed with 
-radius 20.
+Say the respective percentage after 5seconds for orthogonalization.
+
+%\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
+%
+%\todo{decide to keep this section or not}
+%
+%We investigate what are the limits of \tactictoe and what benefits additional 
+%knowledge could provide. To this end, we will modify the set of features to 
+%include features vectors that should not be known in a regular experiment. We  
+%include some vectors from the recorded human proof and perform the proof 
+%search 
+%after the recording instead of before in all other experiments. This 
+%experiment 
+%is not fair and should only be compared to 
+%experiments in the same settings.
+%
+%For clarity, we will name the set of feature vectors that can be recorded from 
+%the current human proof $\mathbb{H}$ and the set of vectors recorded from 
+%previous proofs $\mathbb{P}$ (consistent with previous notation?). We note 
+%$\pi_1^1$ the projection on the 
+%first component returning the label of a feature vector. And we define the 
+%function $T$ that takes a tactic and return the list of its tokens.
+%
+%We will progressively add bigger and bigger 
+%subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof 
+%search success rates. 
+%
+%Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we 
+%consider:
+%\begin{align*}
+%\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})  
+%   \rbrace \\
+%\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
+%\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
+%\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists 
+%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term} 
+%\rbrace\\
+%\end{align*}
+%
+%
+%$\mathbb{F}_1$ is an approximation of what we could get if we had perfect 
+%prediction for those tactics.
+%
+%\begin{table}[h]
+%\centering\ra{1.3}
+%\small
+%\begin{tabular}{lll}
+%\toprule
+% & Additional features & Solved \\
+%\midrule
+%$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
+%$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
+%$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
+%$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
+%$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
+%$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
+%% $AFTERv2$       & $\mathfrak{H}$ & 245 (89.7\%)  \\
+%\bottomrule
+%\end{tabular}
+%\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
+%with tactics from the proof of the tested theorem. The experiments was run 
+%with 
+%a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
+%the low 
+%success rate of the
+%default strategy.} 
+%\end{table}
+%
+%
+%And we observe in Table~\ref{tab:featue_param} a 
+%slight increase compared to the default \tactictoe. We then included all 
+%vectors that contains tactics that can be rebuild from tokens of
+%tactics in our database. 
+%Our search algorithm proved two proofs less that if it had perfect prediction 
+%so it means that our prediction algorithm (policy) is doing a good  job at 
+%least in this early settings where proofs are easy. 
+%Suprisingly, being able to recombine tokens of tactics from the database is 
+%barely sufficient to prove more than half of the theorems in this dataset. The 
+%Therefore, additionally including in tactics theorems and terms that did not 
+%appear in previous tactics is essential. Creating from scratch other arguments 
+%may not help as much. This is why we restricted ourselves to predicting 
+%arguments of theorem list type and the term type during tactic synthesis. 
+%Finally our proof search algorithm sometimes fails even if it knows all the 
+%tactic necessary. The two main reasons are that either a crucial tactic 
+%exceeds its timeout or that the human proof is too deep to be rediscovered 
+%(more than 10 steps long).In all those experiments orthogonalization is 
+%performed with 
+%radius 20.
 
@@
 The best \tactictoe parameters are chosen for the full-scale evaluation. Yet,
-\eprover calls are not included in the best \tactictoe strategy for fairer 
-comparison.
+\eprover asynchronous calls are not included in the best \tactictoe strategy 
+for fairer comparison.
 
@@
 \small
-\begin{tabular}{llc}
+\begin{tabular}{lc}
 \toprule
-  & Parameters & Solved \\
+  & Solved (60s) \\
 \midrule
-   \eprover   & & 2472 (34.5\%)\\ 
-   \tactictoe & best parameters without \eprover & 4741 (66.2\%)\\
+   \eprover   & 2472 (34.5\%)\\ 
+   \tactictoe & 4741 (66.2\%)\\
 \midrule
-   Total  &              & 4946 (69.0\%)\\
+   Total  & 4946 (69.0\%)\\
 \bottomrule
@@
 
-\paragraph{Self-learning}\todo{explain better or remove or move to future work}
-All our feature vectors have been learned form human proofs. We now can now 
-also add tactic-goal pairs that appears in final \tactictoe's proof. To prevent 
-duplication of effort, orthogonalization of those 
-tactics is essential to have a beneficial effect.
-Since recording and re-proving are intertwined, the additional data is 
-available for the next proof search.
-The hope is that the algorithm will improve faster by learning from its own 
-discovered proofs than from the human proof 
-scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that 
-self-learning improves \tactictoe's success rate by less than one percent.
+%\paragraph{Self-learning}\todo{explain better or remove or move to future work}
+%All our feature vectors have been learned form human proofs. We now can now 
+%also add tactic-goal pairs that appears in final \tactictoe's proof. To 
+%prevent 
+%duplication of effort, orthogonalization of those 
+%tactics is essential to have a beneficial effect.
+%Since recording and re-proving are intertwined, the additional data is 
+%available for the next proof search.
+%The hope is that the algorithm will improve faster by learning from its own 
+%discovered proofs than from the human proof 
+%scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that 
+%self-learning improves \tactictoe's success rate by less than one percent.
 
@@
 \paragraph{Acknowledgments}\label{sect:acks}
-This work has been supported by the
-ERC Consolidator grant no.\ 649043 \textit{AI4REASON} and ERC starting
+We would like to thank Lasse Blaauwbroek for his insightful comments which
+contributed to improve the quality of this paper. This work has been supported 
+by the ERC Consolidator grant no.\ 649043 \textit{AI4REASON} and ERC starting
 grant no.\ 714034 \textit{SMART}.

### 2018-02-07 tactictoe_jv.tex

--- 2018-02-05 tactictoe_jv.tex
@@
 \small
-\begin{tabular}{llc}
+\begin{tabular}{llcc}
 \toprule
- Technique & Parameters & Solved \\
+ Technique & Parameters & Solved (1s) & Solved (60s)\\
 \midrule
-Tactic time out & 0.02s & 154 (56.8\%)\\
-                & 0.05s (default) & 156 (56.8\%)\\
-                & 0.1s & 154 (56.0\%)\\
+Tactic time out & 0.02s && 154\\
+                & 0.05s (default) && 156\\
+                & 0.1s && 154\\
 \midrule
-Orthogonalization & none & 156 \\
-                  & $radius = 10$ & 156 \\
-                  & $radius = 20$ (default) & 156 \\
-                  & $radius = 40$ & 156 \\
-
+Orthogonalization & none & 105 & 156 \\
+                  & $radius = 10$ & 121 & 156 \\
+                  & $radius = 20$ (default) & 121 & 156 \\ 
+                  & $radius = 40$ & 124 & 156 \\
 \midrule
-Abstraction       & none (default)  & 156 (\%)\\
-                  & $theorems = 8$  & 195 (\%)\\
-                  & $theorems = 16$ & 199 (72.9\%)\\
-                  & $theorems = 32$ & 195 (\%)\\
+Abstraction       & none (default)  && 156\\ 
+                  & $theorems = 8$  && 195\\
+                  & $theorems = 16$ && 199\\
+                  & $theorems = 32$ && 195\\
 \midrule
-MCTS Policy & $c_{policy} = 0.4$ & 149 (\%)\\
-            & $c_{policy} = 0.5$ (default) & 155 (\%)\\
-            & $c_{policy} = 0.6$ & 153 (\%)\\
+MCTS policy & $c_{policy} = 0.4$ && 149\\
+            & $c_{policy} = 0.5$ (default) && 156\\
+            & $c_{policy} = 0.6$ && 153\\
 \midrule
-MCTS Evaluation & none (default) & 156 (50\%) \\
-				& $c_{exploration} = 0.25$ & 153 153 (50\%)\\
-				& $c_{exploration} = 0.5$ & 153 155 (50\%)\\
-				& $c_{exploration} = 1$ & 154 (50\%)\\
-				& $c_{exploration} =\sqrt{2}$ & 154 154 (50\%)\\
-				& $c_{exploration} = 2$ & 152 (50\%)\\
+MCTS evaluation & none (best abstraction) && 199\\
+				& $c_{exploration} = 1$ && 201\\ 
+				& $c_{exploration} = 2$ && 203\\ 
+				& $c_{exploration} = 4$ && 198\\
 \midrule
-ATP integration & none (default) & 156 (50\%)\\
-                & \metis 0.1s & 125 (50\%)\\
-                & \metis 0.2s & 125 (50\%)\\
-                & \metis 0.4s & 125 (50\%)\\
-                & asynchronous \eprover & 125 (50\%)\\
-                & \metis 0.1s + asynchronous \eprover & 190\\
-\midrule
-All & best parameters (without \eprover) & \\
+ATP integration & none (best MCTS evaluation) && 203\\
+                & \metis 0.1s && 216\\
+                & \metis 0.2s && 212\\
+                & \metis 0.4s && 212\\ 
+                & \eprover && 213\\
+                & \metis 0.1s + \eprover && 218\\
 \bottomrule
 \end{tabular}
+\caption{\label{tab:tuning} Number of problem solved with different set of 
+parameters for \tactictoe on a training set of 273 theorems.} 
 \end{table}
@@
 \subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
-
-\todo{decide to keep this section or not}
 
@@
 \caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge 
-with tactics from the proof of the tested theorem. The experiments was run with
+with tactics from the proof of the tested theorem. The experiments was run 
+with 
 a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
-the low
-success rate of the
-default strategy.}
+the low success rate of thedefault strategy.} 
 \end{table}
@@
 The best \tactictoe parameters are chosen for the full-scale evaluation. Yet,
-\eprover calls are not included in the best \tactictoe strategy for fairer
-comparison.
+\eprover asynchronous calls are not included in the best \tactictoe strategy 
+for fairer comparison.
 
@@
 \small
-\begin{tabular}{llc}
+\begin{tabular}{lc}
 \toprule
-  & Parameters & Solved \\
+  & Solved (60s) \\
 \midrule
-   \eprover   & & 2472 (34.5\%)\\
-   \tactictoe & best parameters without \eprover & 4741 (66.2\%)\\
+   \eprover   & 2472 (34.5\%)\\ 
+   \tactictoe & 4741 (66.2\%)\\
 \midrule
-   Total  &              & 4946 (69.0\%)\\
+   Total  & 4946 (69.0\%)\\
 \bottomrule

### 2018-02-07 tactictoe_jv.tex
commenting out section 7.3 on increasing Tactictoe's knowledge (delaying decision to remove it)

--- 2018-02-07 tactictoe_jv.tex
@@
  appearing in the search tree.
-\item $T$ is a total function from $\mathbb{G}$ to $P(\mathbb{T})$. It takes a 
+\item $T$ is a total function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It takes a
 goal and return the set of tactics already applied to this goal.
-\item $G$ is a total function from $\mathbb{A}$ to $P(\mathbb{G})$. It takes a 
-node a return the set of goals in this node.
+\item $G$ is a total function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$. It takes a
+node and returns the set of goals in this node.
 
@@
 application of the tactic to the goal, if this tactic has been applied to this
-goal. Its support $Support(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g) 
+goal. Its support $\mathit{Support}(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g)
 \rbrace$.
-\item If $(g,t) \in Support (A)$, then $t(g) = G(A(g,t))$.
+\item If $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$.
 \end{itemize}
@@
 \begin{align*}
-Descendant^{0}(a) &=_{def} \lbrace a \rbrace \\ 
-Descendant^{n+1}(a) &=_{def} \bigcup_{a' \in Descendant^{n}(a)} 
-ExtChildren(a') \\
-Descendant(a) &=_{def} \bigcup_{n \in \mathbb{N}} Descendant^n(a)\\
+\mathit{Descendant}^{0}(a) &=_{def} \lbrace a \rbrace \\
+\mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
+\mathit{ExtChildren}(a') \\
+\mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
 \end{align*}
@@
 
-\[CurEvaluation(a_i) = 
-  \sum_{a' \in Descendant(a_i)} \frac{PriorEvaluation(a')} {card\ 
-  Descendant(a_i) + Failure(a_i)}\]
-
-The number $Failure(a_i)$ is the amount of failures that occurred during node 
+\[\mathit{CurEvaluation}(a_i) =
+  \sum_{a' \in Descendant(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
+  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]
+
+The number $\mathit{Failure}(a_i)$ is the amount of failures that occurred during node
 extension from descendants of $a_i$ .
@@
 not beneficial. It may be because it breaks some symmetry in the
-$Value$ function. 
+$\mathit{Value}$ function.
 Otherwise, it applies the best untested tactic $t$ on the open goal $g$ of this
@@
  - Eprover: 128 lemmas and 5 seconds timeout 1 asynchronous call
-
-%sh stats.sh MNM_AE2I
-%273 203 3 67 0
-%thibault@grid02:~/big$ sh stats.sh MNM_AE2
-%273 203 3 67 0
-%thibault@grid02:~/big$ sh stats.sh MNM_AE4
-%273 201 3 69 0
-%thibault@grid02:~/big$ sh stats.sh MNM_AE4I
-%273 200 3 70 0
-%thibault@grid02:~/big$ sh stats.sh MNM_AE1
-%273 198 3 72 0
-%thibault@grid02:~/big$ sh stats.sh MNM_AE1I
-%273 200 3 70 0
-
-%thibault@grid02:~/big$ sh stats.sh MNM_ATPE
-%273 213 3 57 0
-%thibault@grid02:~/big$ sh stats.sh MNM_ATPM01
-%273 216 3 54 0
-%thibault@grid02:~/big$ sh stats.sh MNM_ATPM01E
-%273 218 3 52 0
-%thibault@grid02:~/big$ sh stats.sh MNM_ATPM02
-%273 212 3 58 0
-%thibault@grid02:~/big$ sh stats.sh MNM_ATPM04
-%273 212 3 58 0
-
-%thibault@grid02:~/big$ sh ftim.sh MNM_O40
-%127
-%9
-%8
-%2
-%2
-%148
-%thibault@grid02:~/big$ sh ftim.sh MNM_O10
-%124
-%15
-%5
-%3
-%1
-%148
-%thibault@grid02:~/big$ sh ftim.sh MNM_ON
-%108
-%16
-%6
-%3
-%6
-%thibault@grid02:~/big$ sh ftim.sh MNM_T005
-%124
-%13
-%5
-%2
-%3
 
@@
             & $c_{policy} = 0.6$ && 153\\
-
 \midrule
@@
 \end{tabular}
+\caption{\label{tab:tuning} Number of problem solved with different set of 
+parameters for \tactictoe on a training set of 273 theorems.} 
 \end{table}
 
-Say the respective percentage after 5seconds for orthogonalization.
-
 %\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
-%
-%\todo{decide to keep this section or not}
-%
+%\todoi{decide to keep or not}
 %We investigate what are the limits of \tactictoe and what benefits additional
@@
 %a time out of 5 seconds on an older version of \tactictoe, as can be seen in 
-%the low 
-%success rate of the
-%default strategy.} 
+%the low success rate of thedefault strategy.} 
 %\end{table}
+%
 %
@@
 
-%\paragraph{Self-learning}\todo{explain better or remove or move to future work}
-%All our feature vectors have been learned form human proofs. We now can now 
-%also add tactic-goal pairs that appears in final \tactictoe's proof. To 
-%prevent 
-%duplication of effort, orthogonalization of those 
-%tactics is essential to have a beneficial effect.
-%Since recording and re-proving are intertwined, the additional data is 
-%available for the next proof search.
-%The hope is that the algorithm will improve faster by learning from its own 
-%discovered proofs than from the human proof 
-%scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that 
-%self-learning improves \tactictoe's success rate by less than one percent.
+\paragraph{Self-learning}\todo{explain better or remove or move to future work}
+All our feature vectors have been learned form human proofs. We now can now
+also add tactic-goal pairs that appears in final \tactictoe's proof. To prevent
+duplication of effort, orthogonalization of those
+tactics is essential to have a beneficial effect.
+Since recording and re-proving are intertwined, the additional data is
+available for the next proof search.
+The hope is that the algorithm will improve faster by learning from its own
+discovered proofs than from the human proof
+scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that
+self-learning improves \tactictoe's success rate by less than one percent.
 

### 2018-02-07 tactictoe_jv.tex
some minor modifications in section 2 + example of prettified script

@@
 \item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals
- and $\mathbb{A}$ is a set of nodes. These sets contains only objects
- appearing in the search tree.
+ and $\mathbb{A}$ is a set of nodes.
 \item $T$ is a total function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It takes a
@@
 $\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the
-application of the tactic to the goal, if this tactic has been applied to this
-goal. Its support $\mathit{Support}(A)$ is by definition $\lbrace (g,t)\ |\ t \in T(g)
+application of the tactic to the goal. Its support $\mathit{Support}(A)$ is by 
+definition $\lbrace (g,t) \in 
+\mathbb{G} \times \mathbb{T}\ |\ t \in T(g)
 \rbrace$.
-\item If $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$.
+\item If $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$. This is the 
+main condition. The result of the application of a tactic $t$ 
+to a goal $g$ is the list of goals contained in the node produced by $t$ from 
+$g$.
 \end{itemize}
@@
 \node [tcircle,left of=0l] (0ll) {$g_0^1$};
-\node [draw,fit=(0rr) (0ll)] {};
+\node [draw,fit=(0rr) (0ll)] (0b) {};
+\node [left of=0ll] {$a_0$};
 \node [above of=0,node distance=2cm] (2) {.\ .\ .};
@@
 \node [draw, fit=(2l) (2r)] (2b) {};
+\node [left of=2l] {$a_j$};
 
@@
 \node [draw, fit=(1l) (1r)] (1b) {};
+\node [left of=1l] {$a_1$};
 
@@
 \end{tikzpicture}
-\caption{Number of original and \tactictoe proofs with a number $x$ of tactic
+\caption{Number of original human proofs and (minimized see Section) \tactictoe 
+proofs with 
+a number $x$ of tactic
 units. For easier comparison of the distributions, the number of \tactictoe
@@
 \begin{example}\label{ex:pretty}\ \\
-Extracted proof script:
+Discovered proof script for the theorem \texttt{EVERY\_MAP}
 \begin{lstlisting}[language=SMLSmall]
+boolLib.REWRITE_TAC [DB.fetch "list" "EVERY_CONJ", DB.fetch "list" "EVERY_MEM", 
+  DB.fetch "list" "EVERY_EL", ldots ,  combinTheory.o_DEF] THEN 
+BasicProvers.Induct_on [HolKernel.QUOTE " (*#loc 1 11380*)l"] THENL
+  [BasicProvers.SRW_TAC [] [],
+   simpLib.ASM_SIMP_TAC (BasicProvers.srw_ss ()) [boolLib.DISJ_IMP_THM, 
+     DB.fetch "list" "MAP", DB.fetch "list" "CONS_11", boolLib.FORALL_AND_THM]]
+
 \end{lstlisting}
@@
 \begin{lstlisting}[language=SMLSmall]
+Induct_on `l` THENL 
+  [SRW_TAC [] [], ASM_SIMP_TAC (srw_ss ()) [DISJ_IMP_THM, FORALL_AND_THM]]
 \end{lstlisting}

### 2018-02-07 tactictoe_jv.tex
future work: wip

@@
 {columns=fullflexible,
-  keywords={THEN,THENL,val,open,store_thm,fun,fn},%
+  keywords={THEN,THENL,val,open,store_thm,fun,fn,let,in,end},%
   frame=none,
@@
 tactictoe ([], ``countable (UNIV:num set)``);
-let fun RWTAC thms = SRW_TAC [] thms in RWTAC [(fetch "pred_set"
-"countable_def"), (fetch "pred_set" "INJ_DEF")] THEN Q.EXISTS_TAC `\\x.x` THEN
-METIS_TAC [] end
+
+let fun RWTAC thms = SRW_TAC [] thms in 
+  RWTAC [fetch "pred_set""countable_def", fetch "pred_set" "INJ_DEF"] THEN 
+  Q.EXISTS_TAC `\\x.x` THEN
+  METIS_TAC []
+end
 \end{lstlisting}
@@
 tactictoe ([], ``count (n+m) DIFF count n = IMAGE ($+n) (count m)``);
-SRW_TAC [ARITH_ss] [(fetch "pred_set" "EXTENSION"), EQ_IMP_THM] THEN
-Q.EXISTS_TAC `x - n` THEN SRW_TAC [ARITH_ss] []
+
+SRW_TAC [ARITH_ss] [fetch "pred_set" "EXTENSION", EQ_IMP_THM] THEN
+Q.EXISTS_TAC `x - n` THEN 
+SRW_TAC [ARITH_ss] []
 \end{lstlisting}
@@
 There is plenty of future work in the directions opened here.
+
 To improve the quality of the predicted tactics,
-we would like to predict their arguments independently.
-To be even more precise, the relation between the
-tactic arguments and their respective goals could be used.
-Additionally, we could aim for a tighter combination with the ATP-based hammer
-systems. This would perhaps make \tactictoe slower, but it might allow
-finding proofs that are so far both beyond the ATPs and \tactictoe's
-powers. The idea of reusing high-level blocks of reasoning and
-then learning their selection could also be explored
-further in various contexts. Larger frequent blocks of (instantiated) tactics
-in ITPs as well as blocks of inference patterns in ATPs could be detected
-automatically, their
-usefulness in particular proof situations learned from the large corpora of
-ITP and ATP proofs, and reused in high-level proof search.
+we would like to predict other type of arguments independently, as it was done 
+for theorems. In this direction, the most interesting type of arguments to 
+predict next are terms as they are ubiquitous in tactics. Further along the way,
+we would like to be to able to create new tactics not only by changing their 
+arguments but also by programming it with any construction available in a
+programming language such as \sml.
+
+There is still something to be gain by having a better guidance for the search. 
+That is why se also want to experiment with stronger machine learning algorithm 
+such as deep neural networks as a model for the policy and evaluation in MCTS.
+This quality of this model could be enhanced by additional data augmentation by 
+training and testing existing and synthesized tactics.
+Another possible way to guide the proof search is by conjecturing intermediate 
+steps \cite{}, allowing to prove a difficult theorem by decomposing it into 
+multiple easy steps.
 

### 2018-02-07 tactictoe_jv.tex
minor
comment on tuning

@@
 main condition. The result of the application of a tactic $t$ 
-to a goal $g$ is the list of goals contained in the node produced by $t$ from 
+to a goal $g$ is exactly the list of goals contained in the node produced by 
+$t$ from 
 $g$.
@@
 \midrule
-MCTS evaluation & none (best abstraction) && 199\\
+MCTS evaluation & none (default + best abstraction) && 199\\
 				& $c_{exploration} = 1$ && 201\\ 
@@
 \midrule
-ATP integration & none (best MCTS evaluation) && 203\\
+ATP integration & none (default + best abstraction && 203\\
+                &  + best MCTS evaluation) &&\\
                 & \metis 0.1s && 216\\
@@
 \end{table}
+
+In Table~\ref{tab:tuning}, we run \tactictoe with different parameters. 
+The first four techniques are tested relative to the same baseline indicated in 
+the table by the tag ``default''.  This default strategy relies on a tactic 
+time out of 0.05s, an orthogonalization radius of 20 and a policy coefficent of 
+0.5. 
+The two last techniques are tested relative 
+to a baseline consisting of the best set of parameters discovered so far and 
+are marked with the improvement over the ``default'' strategy.
+In each subtable, each experiment differs from the current baseline by exactly 
+one parameter.
+
+Thanks to the larger search time available in this experiment, the timeout for 
+each tactic can be raised from the 0.02 seconds used in \tactictoe's first 
+experiment~\cite{lpar} to 0.05 
+seconds. Removing tactics with similar effect as performed by the 
+orthogonalization process is only beneficial when running \tactictoe for short 
+period of time. It seems that the alloted time allows a strategy 
+running without orthogonalization to catch up. 
+It is best to predict a list of 16 theorems to instantiate arguments of type 
+theorem list in tactics. At any rate, argument prediction for abstracted 
+tactics is the technique that have the highest impact on the success of 
+\tactictoe. Integration of ATPs being a specialization of this technique 
+contributes significantly as well. Experiments involving \eprover, run the ATP
+on one thread asynchronously with a timeout of 5 seconds and 128 premises. 
+The fact that \metis outperforms \eprover in this setting is due to the fact 
+that \metis is run with a very short timeout, allowing to close different part 
+of search tree quickly. Because \metis is weaker that \tactictoe as an ATP, it 
+is given less predictions. But even if a lemma essential for the proof of a 
+goal is not predicted, a modification on the goal performed by a tactic can 
+change its prediction to include the necessary lemma. 
+This effect minimizes the drawback of relying on a small number of predictions.
+
+The MCTS evaluation, despite relying on the largest amount of collected data, 
+does not provide a significant improvement over a strategy relying on no 
+evaluation. One one hand, our predictor learning abilities is 
+limited. On the other hand, the evaluation function was designed to estimate 
+the provability of a goal. A more accurate evaluation could be based on an 
+estimation of the length of the proof required to close a goal.
 

### 2018-02-07 tactictoe_jv.tex
Move table of contents before content

@@
 
+\tableofcontents
+\todoi{Comparison with hammers in usability}
+
 \section{Introduction}
-\tableofcontents
-
-\todoi{Comparison with hammers in usability}
+
 Many of the state-of-the-art interactive theorem provers (ITPs) such as
@@
   \isabelle~\cite{isabelle}
-  and \coq~\cite{coq-book}
-  provide high-level parameterizable tactics for constructing the
-  proofs.  Such tactics typically analyze the current goal state and
-  assumptions, apply nontrivial proof transformations, which get
-  expanded into possibly many basic kernel-level inferences or significant
-  parts of the proof term.
+  and \coq~\cite{coq-book} provide high-level parameterizable tactics for constructing proofs.
+  Such tactics typically analyze the current proof state (goal and
+  assumptions), then apply non-trivial proof transformations that get
+  expanded into many basic kernel-level inferences or large parts of a proof term.
 %\marginpar{not Coq}
-  In this
-  work we develop a tactic-level automation procedure for the \holfour ITP
-  which guides selection of the tactics by learning from previous
+  In this work we develop a tactic-level automation procedure for the \holfour ITP
+  which guides the selection of tactics by learning from previous
   proofs.  Instead of relying on translation to first-order automated
   theorem provers (ATPs) as done by the hammer
-  systems~\cite{hammers4qed,tgck-cpp15}, the technique
+  systems~\cite{hammers4qed,tgck-cpp15}, our technique
 %\marginpar{too many cites?}
 %BlanchetteGKKU16,holyhammer,mizAR40
-  directly searches for sequences of tactic applications that lead to
-  the ITP proof, thus avoiding the translation and
-  proof-reconstruction phases needed by the hammers.
+  searches directly for sequences of tactic applications that lead to
+  an ITP proof, thus avoiding the translation and proof-reconstruction phases needed by the hammers.
 
@@
   \emph{build efficient machine learning classifiers} based on such training
-  examples (Section~\ref{s:prediction}).  The learned data serves as a
-  guidance for a \emph{Monte Carlo tree search} that explores the different
+  examples (Section~\ref{s:prediction}).  The learned data serves as
+  guidance for a \emph{Monte Carlo tree search} that explores different
   proof paths (Section~\ref{sec:proofsearch}). The result, if

### 2018-02-07 tactictoe_jv.tex
Write some draft text

@@
 \subsection{The problem}
-How a user works with an ITP and how this gives rise to a machine learning
-problem.
+% How a user works with an ITP and how this gives rise to a machine learning problem.
+\todo{RK: this is just a first draft -- feel free to edit lots!}
+An ITP is a development environment for the construction of formal proofs in which it is possible, but not common, to write a proof as a sequence of applications of primitive inference rules.
+The common approach, however, is to use and devise high-level tools and automation that abstract over useful ways of producing proof pieces.
+A specific, prevalent instance of this approach is the use of \emph{tactics} for \emph{backward} or goal-directed proof.
+Here, the ITP user operates on a proof state, initially the desired conjecture, and applies tactics that transform the proof state until it is solved (or until giving up).
+Each tactic performs a sound transformation of the proof state: essentially, it promises that if the resulting proof state can be solved, then so can the starting one.
+This gives rise to a machine learning problem: can we learn a mapping from a given proof state to the next tactic (or sequence of tactics) that will productively advance the proof towards a solution?
+
 Two ways of representing a formulas.

### 2018-02-08 tactictoe_jv.tex
minor

--- 2018-02-07 tactictoe_jv.tex
@@
 \title{TacticToe: Learning the Game of Tactics}
-%\title{TacticToe: Learning the Tactic Game}
 %\title{TacticToe: Automatic tactic-level proof generation}
@@
 \authorrunning{Gauthier et al.}
-%\titlerunning{TacticToe: Learning to Reason with HOL4 Tactics}
+%\titlerunning{TacticToe: Learning the Game of Tactics}
 
@@
 The input of the algorithm is a goal $g$ that we want to prove.
-Therefore the search tree starts with only one node containing the singleton
-$g$ and an ordered list of predicted tactics for the goal $g$.
+Therefore the search tree starts with only one node the list of goals
+$[g]$ and an ordered list of predicted tactics for the goal $g$.
 Then, the MCTS algorithm alternates node selection steps and node extension
@@
 \texttt{A} has no effect. In this situation, it is often preferable to record
-only \texttt{REPEAT A} and any call to \texttt{A}. Otherwise during the
+only \texttt{REPEAT A} than calls to \texttt{A}. Otherwise during the
 proof search, we would have to make a branches created after each call of

### 2018-02-08 tactictoe_jv.tex
graphic update and minor comments

@@
 \begin{axis}[
+  legend style={anchor=south east, at={(0.9,0.1)}},
+  width=\textwidth,
+  height=0.7*\textwidth,xmin=0, xmax=60,
+  ymin=0, ymax=4800,
+  xtick={},
+  ytick={},
+  cycle list name=my black]
+\addplot table[x=time, y=solved] {data/tactictoe_time};
+\addplot table[x=time, y=solved] {data/eprover_time};
+\legend{\tactictoe,\eprover}
+\end{axis}
+\end{tikzpicture}
+\caption{Number of problem solved in less than x seconds.}
+\end{figure}
+\eprover uses at least two strategies where as \tactictoe only uses one set
+of parameters.
+
+\begin{figure}[h]
+\centering
+\begin{tikzpicture}[scale=1]
+\begin{axis}[
   legend style={anchor=north east, at={(0.9,0.9)}},
@@
 \addplot table[x=length, y=proofs] {data/original_proof_length};
-\legend{\tactictoe proofs,original proofs}
+\legend{\tactictoe proofs,human proofs}
 \end{axis}
 \end{tikzpicture}
-\caption{Number of original human proofs and (minimized see Section) \tactictoe
-proofs with
-a number $x$ of tactic
-units. For easier comparison of the distributions, the number of \tactictoe
-proofs was
-scaled by the inverse of \tactictoe's success rate $\frac{1}{0.662}$.}
+\caption{\label{fig:distrib}Number of proofs (y axis) with respect to their 
+length (x axis) in 
+number of tactic units. Comparison between the set of \tactictoe proofs and the 
+set of human proofs.}
 \end{figure}
 
-There are more in proportion \tactictoe small proofs than original human proofs.
-
-Increasing the timeout to more than 10 seconds would probably be beneficial as
-we can see that the slope of the curve is far for being flat in Fig~..
-
-\begin{figure}[h]
-\centering
-\begin{tikzpicture}[scale=1]
-\begin{axis}[
-  legend style={anchor=south east, at={(0.9,0.1)}},
-  width=\textwidth,
-  height=0.7*\textwidth,xmin=0, xmax=60,
-  ymin=0, ymax=4800,
-  xtick={},
-  ytick={},
-  cycle list name=my black]
-\addplot table[x=time, y=solved] {data/tactictoe_time};
-\addplot table[x=time, y=solved] {data/eprover_time};
-\legend{\tactictoe,\eprover}
-\end{axis}
-\end{tikzpicture}
-\caption{Number of problem solved in less than x seconds.}
-\end{figure}
-\eprover uses at least two strategies where as \tactictoe only uses one set
-of parameters.
+In order to appreciate the difficulty of the evaluation set from a human 
+perspective, the length distribution of human proofs in the 
+validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the 
+graph that most of the proofs are short. The proofs found by tactictoe follows 
+about the same 
+distribution. This means that if \tactictoe finds a proof, there is about a 50 
+percent chance that it will be shorter than what a human would come up with.
+In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to 
+see how well \tactictoe and \eprover performs on increased level of 
+difficulties. As expected, the longer the original proof is, the harder it is 
+for \tactictoe and \eprover to re-prove the theorem by their own mean. 
+The performance of \tactictoe is compelling with more than half of the theorem 
+that required a proof of length six being reproven. It is also consistently 
+better than \eprover for any proof length.
 
@@
 \end{tikzpicture}
-\caption{Percentage of problem solved with respect to the length of the
+\caption{\label{fig:percentage}Percentage of problem solved with respect to the 
+length of the
 original proof until length 20.}
@@
 
-%60 seconds experiment +
-%adding theorems from the namespace and removing
-%additionally saved theorems +
-%recording proofs from Prove.
-
-%
-%The progress cannot be attributed only to the time increase but to testing and
-%debugging of parameters.
-
-\todoi{Remove feature vectors from the paper}
 \section{Proof Recording}\label{sec:recording}
-
 Recording proofs in an LCF-style proof assistant can be done at different
@@
 in this section.
-
+<
 Rather than relying on the underlying programming language, we parse the

### 2018-02-08 tactictoe_jv.tex
Work a little more on section 1

@@
 
-Two ways of representing a formulas.
-Definition of a tactic in \holfour. In this research, we will take the definition
-that a tactic takes a goal and returns a list of goals.
-Input of the prediction: goal
-If a tactic produces no goals then its input goal is solved.
+\paragraph{Two ways of representing a formulas.}\todo{RK: Actually I don't know what two ways you mean...?}
+
+\paragraph{Theories and script files.}
+Formal developments in \holfour are organized into named theories, which are concretely implemented by \emph{script} files defining modules in the functional programming language \sml.
+Each script file corresponds to a single theory, and contains definitions of types and constants as well as statements of theorems together with their proofs.
+In practice, tactic-based proofs are written in an informally specified \emph{embedded domain-specific language}: the language consisting of pre-defined tactics and \emph{tacticals} (functions that operate on tactics).
+However, the full power of \sml is always available for users to define their own tactics or combinations thereof.
+%To automate proof search strategies a user can implement programming language functions that correspond to more involved proofs.
+%Functions that correspond to backward proofs, called tactics, are combined with each other with the help of tacticals (functions that takes
+%tactics as arguments).
+%Such a combination of tactics will be called a \emph{proof script} when it proves a theorem stated in a theory.
+
+\paragraph{Tactics in \holfour.}
+Considering concrete implementation details now, a \emph{tactic} is a function that takes a goal (or proof obligation) and returns a list of goals together with a validation function.
+\todo{flesh the below out}
+Input of the prediction: goal If a tactic produces no goals then its input goal is solved.
 Define tactical etc maybe import graphic that explain how tactic works form the
 presentations.
-
 mention proof obligations (use this word) return by a tactic.
-
-\holfour proofs are organized in theory files.
-Each theory file contains definitions and statements of theorems together with
-their proofs. Those theories and the contained proofs
-are written
-in the functional programming language \sml. To automate
-proof search strategies a user can implement programming language functions
-that correspond to more involved proofs.
-Functions that correspond to backward proofs, called tactics, are combined with
-each other with the help of
-tacticals (functions that takes
-tactics as arguments). Such a combination of tactics will be called a
-\emph{proof script} when it
-proves a theorem stated in a theory.
 

### 2018-02-08 tactictoe_jv.tex
minor

--- 2018-02-08 tactictoe_jv.tex
@@
 {columns=fullflexible,
-  keywords={THEN,THENL,val,open,store_thm,fun,fn,let,in,end},%
+  keywords={THEN,THENL,val,open,store_thm,fun,fn,let,in,end,true,
+  while,do,if,then,else,break,return,;},%
   frame=none,
@@
    {ldots}{$\ldots$}1%
+   {emptyset}{$\emptyset$}1%
    {!}{$\forall$}1%
@@
 This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths.
-On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.2\%
+On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\%
 of the 7164 theorems in HOL4's standard library whereas
@@
 \item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals
- and $\mathbb{A}$ is a set of nodes.
+ and $\mathbb{A}$ is a set of nodes representing lists of goals.
 \item $T$ is a total function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It takes a
@@
 \item $G$ is a total function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$. It takes a
-node and returns the set of goals in this node.
-
+node and returns the list of goals of this node.
 \item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to
@@
 to a goal $g$ is exactly the list of goals contained in the node produced by
-$t$ from
-$g$.
+$t$ from $g$.
 \end{itemize}
@@
 arbitrary order. In this way, we can access the $n^{th}$ member of a set $s$ by
-writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written a $g_j^i$.
+writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written as $g_j^i$.
 
@@
 
-Starting from the bottom parent node, we can construct children of a goal of
+Starting from the parent node, we can construct children of a goal of
 this node by choosing a goal $g_i^0$ on this node and choosing tactics
@@
 is the set of nodes produced by tactics applied to $g$:
-  \[\mathit{Children}(a,g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace \]
-
-If $g$ is the open goal of $a$, then we simply call children of a the
-children of $a$ relative to $g$. We define:\\
-
-Extended children of $a$ by:
-  \[\mathit{ExtChildren}(a) = \bigcup_{g \in G(a)} \mathit{Children}(a,g) \]
-
-Descendants of $a$ by:
+\[\mathit{Children}(g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace\]
+
+We define the children of a node $a$ by:
+\[\mathit{Children}(a) = \bigcup_{g \in G(a)} \mathit{Children}(g) \]
+
+And the descendants of $a$ by:
 \begin{align*}
@@
 \mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
-\mathit{ExtChildren}(a') \\
+\mathit{Children}(a') \\
 \mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
 \end{align*}
-
-%Descendants of $a$ through $g$ by:
-%\[Descendant(a,g) =_{def}
-%  \lbrace a \rbrace \cup \bigcup_{a' \in Children(a,g)} Descendant(a')\]
 
@@
 each node is given in Definition~\ref{def:value}. The algorithm that performs
-node selection starts from the root of the search tree. X, it repeatedly choses
-the children with the highest values. Depending on the widening policy, it can
-stop at any chosen node on during its descent through the tree. We now give a
-more precise description of
-the algorithm with the following pseudo-code:
-
-\todoi{Replace alltt environment by something prettier}
-\begin{alltt}
-ChosenNode = Root(Tree);
+node selection starts from the root of the search tree. From its current node,
+it choses among the children of the open goal the one with the highest value. 
+If its value is higher than the widening policy, then the selected child 
+becomes the current node, otherwise the final result of node selection is the 
+current node. The node selection algorithm can be written as the following 
+pseudo-code:
+
+\begin{lstlisting}[language=SMLSmall]
+CurrentNode = Root(Tree);
 while true do
-  if Children(ChosenNode) = \(\emptyset\) then break;
-  (BestChild, BestValue) = CompareChildren (ChosenNode);
-  if WideningPolicy (ChosenNode) \(\geq\) BestValue then break;
-  ChosenNode = BestChild
+  if Children(CurrentNode) = emptyset then break;
+  (BestChild, BestValue) = CompareChildren (CurrentNode);
+  if WideningPolicy (ChosenNode) >= BestValue then break;
+  CurrentNode = BestChild
 end;
-return ChosenNode;
-\end{alltt}
+return CurrentNode;
+\end{lstlisting}
 
@@
    \eprover   & 2472 (34.5\%)\\
-   \tactictoe & 4741 (66.2\%)\\
+   \tactictoe & 4760 (66.4\%)\\
 \midrule
@@
 \end{tikzpicture}
-\caption{\label{fig:distrib}Number of proofs (y axis) with respect to their 
-length (x axis) in 
-number of tactic units. Comparison between the set of \tactictoe proofs and the 
-set of human proofs.}
+\caption{\label{fig:distrib}Number of proof scripts with $x$ tactic units.}
 \end{figure}
@@
 Its implementation,
-\tactictoe, achieves an overall performance of 66.2\% theorems on 7164
+\tactictoe, achieves an overall performance of 66.4\% theorems on 7164
 theorems

### 2018-02-08 tactictoe_jv.tex
problem: wip

@@
 
-\paragraph{Two ways of representing a formulas.}\todo{RK: Actually I don't know what two ways you mean...?}
+\paragraph{Goals and theorems.} 
+A goal is represented as a sequent. Each sequent is composed of 
+a list of assumptions and a conclusion which are higher-order logic formulas. 
+When a goal is proven, it becomes a theorem. And if the developer gives it a 
+name then we refer to it as a top-level theorem.
 
@@
 \paragraph{Tactics in \holfour.}
-Considering concrete implementation details now, a \emph{tactic} is a function that takes a goal (or proof obligation) and returns a list of goals together with a validation function.
-\todo{flesh the below out}
-Input of the prediction: goal If a tactic produces no goals then its input goal is solved.
-Define tactical etc maybe import graphic that explain how tactic works form the
-presentations.
-mention proof obligations (use this word) return by a tactic.
+Considering concrete implementation details now, a \emph{tactic} is a function 
+that takes a goal (or proof obligation) and returns a list of goals together 
+with a validation function. \todoi{do not mention validation?}
 
@@
 
-\section{Proof Search Tree}\label{sec:prelim}
+\section{Preliminaries}\label{sec:prelim}
 
@@
 
-We can also judge the productivity of a tactic by its contribution to the
-search
-tree.
+some text to add here.
+
+\begin{definition}\label{def:desc}(Children, descendant, active node)\\
+The children of a goal $g$ is a set of nodes defined by:
+\[\mathit{Children}(g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace\]
+
+By extension, we define the children of a node $a$ by:
+\[\mathit{Children}(a) = \bigcup_{g \in G(a)} \mathit{Children}(g) \]
+
+The descendants of $a$ by:
+\begin{align*}
+\mathit{Descendant}^{0}(a) &=_{def} \lbrace a \rbrace \\
+\mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
+\mathit{Children}(a') \\
+\mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
+\end{align*}
+
+And the ancestors of $a$ by:
+\[\mathit{Ancestors}(a) =_{def} \lbrace b \in \mathbb{A} \ | \ a \in 
+Descendant(b) \rbrace\]
+
+An active node is by definition a node that is not a descendant of a
+solved node. Non-active nodes are not worth exploring.
+\end{definition}
+
+We can now judge the productivity of a tactic by its contribution to the
+search tree.
 
@@
 \item It does not loop. Its output does not contain any goals that appears in
-its ancestor nodes.
+the ancestor of the node of the current 
+goal.
 \item It is not a parallel step. Its output is not equal or a superset
@@
 separated in the search tree.
-\end{definition}
-
-some text to add here.
-
-\begin{definition}\label{def:desc}(Children, descendant, active node)\\
-We now define what are the children and the descendants of a node $a$. The
-children of a relative to a goal $g$ in $G(a)$
-is the set of nodes produced by tactics applied to $g$:
-\[\mathit{Children}(g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace\]
-
-We define the children of a node $a$ by:
-\[\mathit{Children}(a) = \bigcup_{g \in G(a)} \mathit{Children}(g) \]
-
-And the descendants of $a$ by:
-\begin{align*}
-\mathit{Descendant}^{0}(a) &=_{def} \lbrace a \rbrace \\
-\mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
-\mathit{Children}(a') \\
-\mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
-\end{align*}
-
-An active node is by definition a node that is not a descendant of a
-solved node. Non-active nodes are not worth exploring.
 \end{definition}
@@
 The three objects considered are different representation of mathematical
-formulas. Therefore, we start by describing which features we extract can
-extract terms. From there, we extend the extraction mechanism
+formulas. Therefore, we start by describing features for terms. From there, we 
+extend the extraction mechanism
 to goals, theorems and lists of goals. Duplicated features are always removed
@@
 In order to speed up the predictions during the search for a proof of a
-conjecture $c$, we preselect 500 tactic-goal pairs and theorems.
-Preselection for list of goals is induced by the preselection of input
-goals in tactic-goal pairs. Those will be the only objects available to
-\tactictoe's search algorithm for proving $c$.
+conjecture $c$, we preselect 500 tactic-goal pairs and 500 theorems and a 
+larger number of list of goals induced by the selection of tactic-goal pairs. 
+Preselected objects will 
+be the only objects available to \tactictoe's search algorithm for proving $c$.
 

### 2018-02-10 tactictoe_jv.tex
minor

@@
 
-\section{Case Study}
+\section{Case Study}\todo{comment on the case study}
 In this section, we compare \tactictoe proof script with human proof script.
@@
 
-%A first successful use of orthogonalization techniques for theorems appeared in
-%a work on lemma mining~\cite{ckju-jsc15} in \hollight. Theorems were mined
-%from
-%proofs recorded at the kernel level. Because the number of recorded lemmas was
-%in millions, page ranking techniques was used as a filter to extract
-%``useful''
-%lemmas.
-
-\todoi{Is copied and modified}
 We proposed a new proof assistant automation technique which combines
-tactic-based proof search with machine learning tactic prediction
-Its implementation,
-\tactictoe, achieves an overall performance of 66.4\% theorems on 7164
-theorems
-of the \holfour standard library surpassing \eprover with
-auto-schedule. Its
+tactic-based proof search with machine learning prediction
+Its implementation, \tactictoe, achieves an overall performance of 66.4\% 
+theorems on 7164 theorems of the \holfour standard library surpassing \eprover 
+with auto-schedule. Its
 effectiveness is especially visible on
@@
 available in today's large formal proof corpora.
-%And it seems to get us
-%closer to the way how humans learn to do theorem proving.
-
-There is plenty of future work in the directions opened here.
 
@@
 we would like to predict other type of arguments independently, as it was done
-for theorems. In this direction, the most interesting type of arguments to
+for theorems. In this direction, the most interesting arguments to
 predict next are terms as they are ubiquitous in tactics. Further along the way,
-we would like to be to able to create new tactics not only by changing their
-arguments but also by programming it with any construction available in a
-programming language such as \sml.
-
-There is still something to be gain by having a better guidance for the search.
+new tactics could be created by programming them with any construction 
+available in \sml. There is also something to be gain by 
+having a better guidance for the search.
 That is why se also want to experiment with stronger machine learning algorithm
@@
 training and testing existing and synthesized tactics.
-Another possible way to guide the proof search is by conjecturing intermediate
-steps \cite{}, allowing to prove a difficult theorem by decomposing it into
-multiple easy steps.
-
-%Future work:
-%  Promising techniques that could not be evaluated.
-%  + Orthogonalization of theorems~\cite{ckju-jsc15}.
-%  + Try harder and smarted on term predictions.
-%  \cite{conjecturing}
-%  + Self-learning.
-%+ MCTS was presented in manner that we only did the first step of
-%reinforcement
-%learning. CurrentPolicy becomes PriorPolicy.
-% + More training than just orthogonalization of tactics.
+Conjecturing suitable intermediate steps could also allow \tactictoe to solve 
+a problem which requires a long proof by decomposing it into multiple easy 
+steps.
 

### 2018-02-13 tactictoe_jv.tex
Fix warninig

@@
 even more credence
-to rare features which was found experimentally optimal in \cite{}.\todo{CK:
-citation}
+to rare features which was found experimentally optimal in \cite{ckju-pxtp13}.
 
@@
 each tactic can be raised from the 0.02 seconds used in \tactictoe's first
-experiment~\cite{lpar} to 0.05
+experiment~\cite{tgckju-lpar17} to 0.05
 seconds. Removing tactics with similar effect as performed by the

### 2018-02-13 tactictoe_jv.tex
Sec 2
Start sec 3

@@
 
-The different components of \tactictoe revolves around the concept of a
+The different components of \tactictoe revolve around the concept of a
 search tree. The following definitions are made relative to a static search
@@
 
-some text to add here.
+\todoi{some text to add here}
 
@@
 The learning-based selection of relevant lemmas significantly improves the
-automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we adapt
-one of the strongest lemma selection
-methods: the modified distance-weighted \emph{$k$ nearest-neighbour}
-classifier~\cite{ckju-pxtp13,DudaniS76} to predict three different objects:
+automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we use the
+distance-weighted \emph{$k$ nearest-neighbour} classifier~\cite{DudaniS76}
+adapted for lemma selection~\cite{ckju-pxtp13}. It allows for hundreds of
+predictions per second, while maintaining very high prediction quality~\cite{femalecop}.
+Here it is used
+to predict three different kinds of objects:
  tactics, theorems and list of goals. The number of predicted objects $k$ is
- called the prediction radius.
+ called the \emph{prediction radius}.
 
@@
 similarity with a targeted goal $g$ (usually the first pending goal of a node).
-This induces
-through the recorded pairs an
+The recorded pairs induce an
 order on tactics. The intuition is that tactics which have been successful on
@@
 
-\paragraph{Theorems as Argument of Tactics}
+\paragraph{Theorems as Arguments of Tactics}
 We first collect a set of theorems for our theorem predictor to select from.
@@
 We instantiate those tactics by theorems that are syntactically the closest to
-the targeted goal as they are more likely to help.
+the goal as they are more likely to help.
 The same algorithm selects suitable premises for ATPs integrated with
@@
 \subsection{Feature Extraction}\label{sec:features}
-The three objects considered are different representation of mathematical
+All the three kinds of considered objects to predict correspond to (different representations of)
+mathematical
 formulas. Therefore, we start by describing features for terms. From there, we

### 2018-02-14 tactictoe_jv.tex
case study: wip

@@
 \section{Case Study}\todo{comment on the case study}
-In this section, we compare \tactictoe proof script with human proof script.
-
-Goal: \texttt{IS\_GCD\_UNIQUE} in theory \texttt{pred\_set}).
+Since the proofs generated by \tactictoe are meant to be part of a \holfour 
+development, it interesting to compare their quality with the original human 
+proofs. The quality of a proof in \holfour can be measured in terms of length, 
+speed, readability and maintainability. We study these properties in three 
+examples taken from our full-scale experiment.
+
+In Example~\ref{ex:cs1},
+
+\begin{example}\label{ex:cs1} \texttt{IS\_GCD\_UNIQUE} in theory \texttt{gcd}
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-
-Original \holfour proof script
+Human proof (5 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-
-\vspace{5mm}
-
-\tactictoe proof script
+\tactictoe proof (80 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
+\end{example}
 
@@
 
-Goal: \texttt{SURJ\_INJ\_INV} in theory \texttt{pred\_set})
+\begin{example} \texttt{SURJ\_INJ\_INV} in theory \texttt{pred\_set}
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-
-Original \holfour proof script (2 milliseconds)
+Human (2 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-
-\vspace{5mm}
-
-Make it match the new version.
-\tactictoe proof script (50 milliseconds)
+\tactictoe (50 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-
-Goal
+\end{example}
+
+\begin{example} \texttt{LUPDATE\_APPEND2} in theory \texttt{rich\_list}
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-
-\vspace{5mm}
-
-Original \holfour proof script (63 milliseconds)
+Human (63 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-
-%Name: 224 LUPDATE_APPEND2
-%Original proof time:
-
-%val LUPDATE_APPEND2 = Q.store_thm("LUPDATE_APPEND2",
-%   `!l1 l2 n x.
-%      LENGTH l1 <= n ==>
-%      (LUPDATE x n (l1 ++ l2) = l1 ++ (LUPDATE x (n-LENGTH l1) l2))`,
-%);
-\vspace{5mm}
-Make it match the new version.
-\tactictoe proof script (17 milliseconds)
+\tactictoe (17 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
+\end{example}
+
 fs is a \sml binding for FULL\_SIMP\_TAC (srw\_ss ()) and rw is short for ??.
-
 If a discovered proof is short, fast, maintainable and
@@
 
-\todoi{Side effect:
-+ could be also used to simplify existing proof scripts. }
-
-Interactive usage example.
-
+To conclude, we also demonstrate the capabilities of \tactictoe on two goals 
+that do not have a proof in \holfour.
+
+The first goal, in Example~, states that the set of 
+natural numbers is countable. \tactictoe starts from the definition of 
+countability and injection then instantiate the result with the trivial 
+injection and finalize the proof using the \metis. Here,
+\tactictoe is able to find an instantiation for an existential quantifier 
+because it is a common one. If a more complicated term was needed, \tactictoe 
+would try to find an alternative proof that do not rely on an original
+instantiation. This weakness could be mitigated by the introduction of term 
+prediction in a future version of \tactictoe.
+
+\begin{example}
 \begin{lstlisting}[language=SMLSmall]
-tactictoe ([], ``countable (UNIV:num set)``);
-
+countable (UNIV:num set)
 let fun RWTAC thms = SRW_TAC [] thms in
-  RWTAC [fetch "pred_set""countable_def", fetch "pred_set" "INJ_DEF"] THEN
-  Q.EXISTS_TAC `\\x.x` THEN
-  METIS_TAC []
+  RWTAC [countable_def, INJ_DEF] THEN Q.EXISTS_TAC `\\x.x` THEN METIS_TAC []
 end
 \end{lstlisting}
-
+\end{example}
+
+The second goal, in Example~, \todoi{TG to RK: I am not sure what the count 
+constant is doing.}
+
+\begin{example}
 \begin{lstlisting}[language=SMLSmall]
-tactictoe ([], ``count (n+m) DIFF count n = IMAGE ($+n) (count m)``);
-
+count (n+m) DIFF count n = IMAGE ($+n) (count m)
 SRW_TAC [ARITH_ss] [fetch "pred_set" "EXTENSION", EQ_IMP_THM] THEN

### 2018-02-14 tactictoe_jv.tex
example wip

@@
 
-To conclude, we also demonstrate the capabilities of \tactictoe on two goals 
-that do not have a proof in \holfour.
-
-The first goal, in Example~, states that the set of 
-natural numbers is countable. \tactictoe starts from the definition of 
-countability and injection then instantiate the result with the trivial 
-injection and finalize the proof using the \metis. Here,
-\tactictoe is able to find an instantiation for an existential quantifier 
-because it is a common one. If a more complicated term was needed, \tactictoe 
-would try to find an alternative proof that do not rely on an original
-instantiation. This weakness could be mitigated by the introduction of term 
-prediction in a future version of \tactictoe.
-
-\begin{example}
-\begin{lstlisting}[language=SMLSmall]
-countable (UNIV:num set)
-let fun RWTAC thms = SRW_TAC [] thms in
-  RWTAC [countable_def, INJ_DEF] THEN Q.EXISTS_TAC `\\x.x` THEN METIS_TAC []
-end
-\end{lstlisting}
-\end{example}
-
-The second goal, in Example~, \todoi{TG to RK: I am not sure what the count 
+To conclude, we also demonstrate the capabilities of \tactictoe on a goal
+that did not have a proof in \holfour.
+
+\todoi{commented out because the theorem exists in hoL4}.
+%The first goal, in Example~, states that the set of 
+%natural numbers is countable. \tactictoe starts from the definition of 
+%countability and injection then instantiate the result with the trivial 
+%injection and finalize the proof using the \metis. Here,
+%\tactictoe is able to find an instantiation for an existential quantifier 
+%because it is a common one. If a more complicated term was needed, \tactictoe 
+%would try to find an alternative proof that do not rely on an original
+%instantiation. This weakness could be mitigated by the introduction of term 
+%prediction in a future version of \tactictoe.
+
+%\begin{example}
+%\begin{lstlisting}[language=SMLSmall]
+%countable (UNIV:num set)
+%let fun RWTAC thms = SRW_TAC [] thms in
+%  RWTAC [countable_def, INJ_DEF] THEN Q.EXISTS_TAC `\\x.x` THEN METIS_TAC []
+%end
+%\end{lstlisting}
+%\end{example}
+
+In Example~, \todoi{TG to RK: I am not sure what the count 
 constant is doing.}

### 2018-02-15 tactictoe_jv.tex
updating tactictoe results

@@
 does not provide a significant improvement over a strategy relying on no
-evaluation. One one hand, our predictor learning abilities is
-limited. On the other hand, the evaluation function was designed to estimate
-the provability of a goal. A more accurate evaluation could be based on an
-estimation of the length of the proof required to close a goal.
+evaluation. The minor improvement seen in the table is most likely due to some 
+variability in our experiments.
+The main reason is that our predictor learning abilities is
+limited and we are using the provability of lists of goals for evaluation. A 
+more accurate evaluation could be based on an estimation of the length of the 
+proof required to close a list of goals.
 

### 2018-02-15 tactictoe_jv.tex
example commentary

@@
 by asynchronous calls to \eprover to help \tactictoe during proof search.
-\todoi{More on evaluation??}
 \item Evaluation of \tactictoe success rate relative to
@@
 
-In Example~\ref{ex:cs1},
+In Example~\ref{ex:cs1}, the uniqueness of the greatest common divisor is 
+proven. The human developer recognizes that it follows easily from two theorems.
+Our prover \tactictoe is not able to find that \texttt{DIVIDES\_ANTISYM} is 
+necessary directly from the shape of the original theorems so it rewrites the 
+goal and split the goal into multiple cases until goal obligations are similar 
+to \texttt{DIVIDES\_ANTISYM} to be predicted. As expected, \tactictoe's proof 
+also takes much longer to check.
 
@@
 
-TacicToe is resourceful. Changing the shape of the feature of the theorems
-to get better predictions.
-
-\begin{example} \texttt{SURJ\_INJ\_INV} in theory \texttt{pred\_set}
+In Example~\ref{ex:cs2}, we try to prove that for any surjective function $f$ 
+from $s$ to $t$, there exists an injective function $g$ from $t$ to $s$ such 
+that $f o g$ is the identity function. The human proof is quite complicated. In 
+contrast, \tactictoe finds a proof that first expand the definition of 
+injectivity and surjectivity and then call \metis. However, the simpler script 
+does take much longer to execute due to the proof search happening inside 
+\metis.
+
+\begin{example}\label{ex:cs2} \texttt{SURJ\_INJ\_INV} in theory 
+\texttt{pred\_set}
 \begin{lstlisting}[language=SMLSmall]
@@
 
-\begin{example} \texttt{LUPDATE\_APPEND2} in theory \texttt{rich\_list}
+In Example~\ref{ex:cs3}, we prove a theorem about lists which is domain where 
+\tactictoe excels compared to ATPs. 
+Given two list $l1$ and $l2$ where the length of $l1$ is less than a natural 
+number $n$, the theorem states that performing an update on the $n^{th}$ 
+element of the concatenation of $l1$ and $l2$ is the same as
+performing an update on the $m^{th}$ element of $l2$ where $m = n - 
+\mathit{length} l1$ followed by concatenation. Again, \tactictoe's proof is 
+much more readable starting with applying induction on $l$. It solves the base 
+case of the induction by rewriting and distinguish cases where $n=0$ or $n>0$ 
+in the inductive hypothesis. It finalizes the proof using the definition of 
+$LUPDATE$. Here, \tactictoe's proof is arguably smaller, faster and easier to 
+understand and maintain. Such proofs after an expert review could replace the 
+original human proofs in the \holfour repository.
+
+\begin{example}\label{ex:cs3} \texttt{LUPDATE\_APPEND2} in theory 
+\texttt{rich\_list}
 \begin{lstlisting}[language=SMLSmall]
@@
 
-fs is a \sml binding for FULL\_SIMP\_TAC (srw\_ss ()) and rw is short for ??.
-If a discovered proof is short, fast, maintainable and
-understandable, it could replace its original in the \holfour standard library.
-
-To conclude, we also demonstrate the capabilities of \tactictoe on a goal
-that did not have a proof in \holfour.
-
-\todoi{commented out because the theorem exists in hoL4}.
-%The first goal, in Example~, states that the set of 
-%natural numbers is countable. \tactictoe starts from the definition of 
-%countability and injection then instantiate the result with the trivial 
-%injection and finalize the proof using the \metis. Here,
-%\tactictoe is able to find an instantiation for an existential quantifier 
-%because it is a common one. If a more complicated term was needed, \tactictoe 
-%would try to find an alternative proof that do not rely on an original
-%instantiation. This weakness could be mitigated by the introduction of term 
-%prediction in a future version of \tactictoe.
-
-%\begin{example}
-%\begin{lstlisting}[language=SMLSmall]
-%countable (UNIV:num set)
-%let fun RWTAC thms = SRW_TAC [] thms in
-%  RWTAC [countable_def, INJ_DEF] THEN Q.EXISTS_TAC `\\x.x` THEN METIS_TAC []
-%end
-%\end{lstlisting}
-%\end{example}
-
-In Example~, \todoi{TG to RK: I am not sure what the count 
-constant is doing.}
-
-\begin{example}
+In Example~\ref{ex:cs4}, we demonstrate the capabilities of 
+\tactictoe on a goal that did not have a proof in \holfour.
+
+\todoi{TG to RK: I am not sure what the count 
+constant is doing. Feel free to add more examples.}
+
+\begin{example}\label{ex:cs4}
 \begin{lstlisting}[language=SMLSmall]

### 2018-02-15 tactictoe_jv.tex
examples wip

@@
 \tactictoe excels compared to ATPs. 
-Given two list $l1$ and $l2$ where the length of $l1$ is less than a natural 
+Given two list $l_1$ and $l_2$ where the length of $l_1$ (noted $p$) is less 
+than a natural 
 number $n$, the theorem states that performing an update on the $n^{th}$ 
-element of the concatenation of $l1$ and $l2$ is the same as
-performing an update on the $m^{th}$ element of $l2$ where $m = n - 
-\mathit{length} l1$ followed by concatenation. Again, \tactictoe's proof is 
+element of the concatenation of $l_1$ and $l_2$ is the same as
+performing an update on the $m^{th}$ element of $l_2$ where $m = n - p$. Again, 
+\tactictoe's proof is 
 much more readable starting with applying induction on $l$. It solves the base 
@@
 understand and maintain. Such proofs after an expert review could replace the 
-original human proofs in the \holfour repository.
+their respective original human proofs in the \holfour repository.
 

### 2018-02-15 tactictoe_jv.tex
examples minor

@@
 than a natural 
-number $n$, the theorem states that performing an update on the $n^{th}$ 
+number $n$, the theorem states updating the $n^{th}$ 
 element of the concatenation of $l_1$ and $l_2$ is the same as
-performing an update on the $m^{th}$ element of $l_2$ where $m = n - p$. Again, 
+updating the $m^{th}$ element of $l_2$ where $m = n - p$. Again, 
 \tactictoe's proof is 

### 2018-02-15 tactictoe_jv.tex
recording minor

@@
 other proof scripts.
-If the theorem is stored in the database of the current theory, we
-can obtain the theorem value by calling \texttt{DB.fetch}. This happens for the
-three theorems in our script. The internal database of \holfour theorems is
-available across \holfour developments.
+If the theorem is stored in the \holfour database available across \holfour 
+developments, we can obtain the theorem value by calling \texttt{DB.fetch}.
 Otherwise the globalization process fails and the local theorem identifier is
@@
 proofs. The quality of a proof in \holfour can be measured in terms of length, 
-speed, readability and maintainability. We study these properties in three 
-examples taken from our full-scale experiment.
+, readability, maintainability and verification speed. We study these 
+properties 
+in three examples taken from our full-scale experiment. The time needed by 
+\holfour to check a proof is shown in parentheses.
 
@@
 from $s$ to $t$, there exists an injective function $g$ from $t$ to $s$ such 
-that $f o g$ is the identity function. The human proof is quite complicated. In 
-contrast, \tactictoe finds a proof that first expand the definition of 
-injectivity and surjectivity and then call \metis. However, the simpler script 
-does take much longer to execute due to the proof search happening inside 
+that $f\ o\ g$ is the identity function. The human proof is quite complicated. 
+In 
+contrast, \tactictoe finds a much smaller proof that expands the 
+definition of injectivity and surjectivity and calls \metis. However, the 
+\tactictoe's script does take much longer to check due to the proof search 
+happening inside 
 \metis.
@@
 \end{lstlisting}
-Human (2 milliseconds)
+Human proof (2 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-\tactictoe (50 milliseconds)
+\tactictoe proof (50 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 case of the induction by rewriting and distinguish cases where $n=0$ or $n>0$ 
-in the inductive hypothesis. It finalizes the proof using the definition of 
+in the inductive hypothesis. It finalizes the proof with a short call to \metis 
+using the definition of 
 $LUPDATE$. Here, \tactictoe's proof is arguably smaller, faster and easier to 
@@
 \end{lstlisting}
-Human (63 milliseconds)
+Human proof (63 milliseconds)
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-\tactictoe (17 milliseconds)
+\tactictoe proof (17 milliseconds)
 \begin{lstlisting}[language=SMLSmall]

### 2018-02-15 tactictoe_jv.tex
Edit section 1 a little

@@
 \begin{abstract}
-Formalization of mathematics in interactive theorem provers (ITPs)
-relies today on many proof automation methods.
-Some of these methods are general-purpose and with enough time they can prove any
-theorem. Many formalized proof scripts however also
-take advantage of specialized rules and theory-based strategies.
+Formalization of mathematics in interactive theorem provers (ITPs) relies today on many proof automation methods.
+Some of these methods are general-purpose and with enough time they can prove any provable conjecture.
+Many formalized proof scripts however also take advantage of specialized rules and theory-based strategies.
 In the HOL4 ITP system, practically all of them are implemented as \emph{tactics}.
@@
 strategies. Sometimes, a strategy can be specialized for only one theory as in
-SMT solving. But it is difficult for an automated theorem prover to rely on
-rules tailored for a particular case.
+SMT solving. But it is difficult for an ATP to rely on rules tailored for a particular case.
 Indeed, this would require implementing many specific rules and more
@@
 Considering concrete implementation details now, a \emph{tactic} is a function
-that takes a goal (or proof obligation) and returns a list of goals together 
-with a validation function. \todoi{do not mention validation?}
+that takes a goal (or proof obligation) and returns a list of goals (subgoals of the original goal) together
+with a validation function.
+Calling the validation function on a list of \emph{theorems}, corresponding to the list of subgoals, results in a theorem corresponding to the original goal.
+For example, if the list of subgoals is empty, then calling the validation function on the empty list yields the originally goal sequent as a theorem.
+In this way, tactics implement the plumbing of backwards, goal-directed, proof construction.
+\todo{do not mention validation? RK: it depends whether these details are helpful for understanding or not}
 
 \subsection{Contributions}
-This work is an extension of the work described in \cite{tgckju-lpar17}.
+This paper extends the work described in \cite{tgckju-lpar17}.
 There we proposed the idea of ... and achieved ...

### 2018-02-15 tactictoe_jv.tex
Tweak some of the terminology
e.g., proof script -> tactic proof; theory file -> script file.

I have not done this globally yet

@@
 
-The first line of this theory open modules (called structures in \sml). Each of
+The first line of this script file opens modules (called structures in \sml). Each of
 them contains a list of global identifiers which become directly accessible in
-the rest of the theory.
-A local identifier \texttt{LIST\_INDUCT\_TAC} is declared after, which is a
-tactic that performs induction on lists. And even further, the theorem
+the rest of the script.
+A local identifier \texttt{LIST\_INDUCT\_TAC} is declared next, which is a
+tactic that performs induction on lists. Below that, the theorem
 \texttt{MAP\_APPEND} is proven.
 The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
-the goal is split into a base case and a inductive case. And finally, both of
+the goal is split into a base case and a inductive case. Finally, both of
 theses cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
@@
 
-We first parse the theory to extract proof scripts. Each of them is found in
-the third argument of a call to \texttt{store\_thm}. The result of proof script
+We first parse the script file to extract tactic proofs. Each of them is found in
+the third argument of a call to \texttt{store\_thm}. The result of tactic proof
 extraction for the running example in presented in Example~\ref{ex:r0}.
 
-\begin{example}\label{ex:r0}(Proof script extraction)
+\begin{example}\label{ex:r0}(Tactic proof extraction)
 \small
@@
 
-In the next phase, we globalize identifiers of the proof scripts.
+In the next phase, we globalize identifiers of the tactic proofs.
 Infix operators such as \texttt{THEN} need to be processed in a special way so
@@
 local theorems.  We do not replace a local
-theorem by its definition because we want to avoid unfolding proof scripts into
-other proof scripts.
+theorem by its definition because we want to avoid unfolding tactic proofs into
+other tactic proofs.
 If the theorem is stored in the \holfour database available across \holfour
@@
 
-Running the globalized version of a proof script will have the exact same
+Running the globalized version of a tactic proof will have the exact same
 effect as the original. But since we want to extract information from this
 script in the form tactics and their input goals, we modify it further.
-In particular, we need to define at which level we should record the tactics of
-the proof scripts. The simplest idea would be to record all \sml subexpressions
-of type tactics. However, it will damage the quality of our data by
-dramatically increasing the number of tactic associated with a single goal.
-Imagine a proof script of the form \texttt{A THEN B THEN C} then the tactics
+In particular, we need to define at which level we should record the tactics in
+the tactic proof. The simplest idea would be to record all \sml subexpressions
+of type \texttt{tactic}. However, it will damage the quality of our data by
+dramatically increasing the number of tactics associated with a single goal.
+Imagine a proof script of the form \texttt{A THEN B THEN C}; then the tactics
 \texttt{A}, \texttt{A THEN B} and \texttt{A THEN B THEN C} would be valid
@@
 make the most progress on a goal. As a trade-off between these two objectives,
-we split the proofs into tactic units.
+we split proofs into tactic units.
 
 \begin{definition}(Tactic unit)
-A tactic unit is an \sml expression of type tactics that does not contain a
+A tactic unit is an \sml expression of type \texttt{tactic} that does not contain a
 infix operator at its root.
@@
 Because these tactic units are constructed from visual information present in
-the proof script, they often represent what a human consider to be one step of
+the tactic proof, they often represent what a human consider to be one step of
 the proof.
@@
 After all modified theories are rebuild, each call to a wrapped tactic in a
-proof script is recorded as a pair containing its globalized code and its
+tactic proof is recorded as a pair containing its globalized code and its
 input goal.

### 2018-02-15 tactictoe_jv.tex
Some random grammar edits

@@
 \section{Case Study}\todo{comment on the case study}
-Since the proofs generated by \tactictoe are meant to be part of a \holfour
-development, it interesting to compare their quality with the original human
-proofs. The quality of a proof in \holfour can be measured in terms of length,
-, readability, maintainability and verification speed. We study these
-properties
-in three examples taken from our full-scale experiment. The time needed by
-\holfour to check a proof is shown in parentheses.
-
-In Example~\ref{ex:cs1}, the uniqueness of the greatest common divisor is
-proven. The human developer recognizes that it follows easily from two theorems.
+Since the proofs generated by \tactictoe are meant to be part of a \holfour development, it interesting to compare their quality with the original human proofs.
+The quality of a proof in \holfour can be measured in terms of length, readability, maintainability and verification speed.
+We study these properties in three examples taken from our full-scale experiment.
+The time needed by \holfour to check a proof is shown in parentheses.
+
+Example~\ref{ex:cs1} is about proving that greatest common divisors are unique.
+The human developer recognizes that it follows easily from two theorems.
 Our prover \tactictoe is not able to find that \texttt{DIVIDES\_ANTISYM} is
 necessary directly from the shape of the original theorems so it rewrites the
-goal and split the goal into multiple cases until goal obligations are similar
+goal then splits the goal into multiple cases until the goal obligations are similar
 to \texttt{DIVIDES\_ANTISYM} to be predicted. As expected, \tactictoe's proof
@@
 that $f\ o\ g$ is the identity function. The human proof is quite complicated.
-In
-contrast, \tactictoe finds a much smaller proof that expands the
-definition of injectivity and surjectivity and calls \metis. However, the
-\tactictoe's script does take much longer to check due to the proof search
-happening inside
-\metis.
+In contrast, \tactictoe finds a much smaller proof that expands the
+definition of injectivity and surjectivity and calls \metis.
+However, the \tactictoe's proof takes much longer to check due to the proof search happening inside \metis.
 
@@
 
-In Example~\ref{ex:cs3}, we prove a theorem about lists which is domain where
-\tactictoe excels compared to ATPs.
+In Example~\ref{ex:cs3}, we prove a theorem about lists, a domain where \tactictoe excels compared to ATPs.
 Given two list $l_1$ and $l_2$ where the length of $l_1$ (noted $p$) is less

### 2018-02-15 tactictoe_jv.tex
Explain the statement of Example 10

@@
 
-In Example~\ref{ex:cs4}, we demonstrate the capabilities of
-\tactictoe on a goal that did not have a proof in \holfour.
-
-\todoi{TG to RK: I am not sure what the count
-constant is doing. Feel free to add more examples.}
+In Example~\ref{ex:cs4}, we demonstrate the capabilities of \tactictoe on a goal that was not already proved in the \holfour libraries: that the set of numbers $\{0,...,n+m-1\} \setminus \{0,...,n-1\}$ is the same as the set obtained by adding $n$ to everything in $\{0,...,m-1\}$.
 
@@
 \begin{lstlisting}[language=SMLSmall]
-count (n+m) DIFF count n = IMAGE ($+n) (count m)
+count (n+m) DIFF count n = IMAGE ((+) n) (count m)
 SRW_TAC [ARITH_ss] [fetch "pred_set" "EXTENSION", EQ_IMP_THM] THEN
@@
 \end{example}
+
+\todoi{TG to RK: Feel free to add more examples.}
 

### 2018-02-15 tactictoe_jv.tex
minor

@@
 count (n+m) DIFF count n = IMAGE ((+) n) (count m)
-SRW_TAC [ARITH_ss] [fetch "pred_set" "EXTENSION", EQ_IMP_THM] THEN
+SRW_TAC [ARITH_ss] [EXTENSION, EQ_IMP_THM] THEN
 Q.EXISTS_TAC `x - n` THEN

### 2018-02-16 tactictoe_jv.tex
commentary on the experiments

@@
 
-%Mention the conflicts.
-%Theorem orthogonalization is not turn on during instantiation of tacitcs
-%except
-%for the metis tactic.
-%Best parameters + comparison with holyhammer.
-
-The best \tactictoe parameters are chosen for the full-scale evaluation. Yet,
-\eprover asynchronous calls are not included in the best \tactictoe strategy
-for fairer comparison.
+Based on the results of parameter tuning, we now evaluate a version of 
+\tactictoe with its best parameters on an evaluation set of 7164 theorems from 
+the \holfour standard library. We compare it with the performance of \eprover.
+For reasons of fairness, \eprover asynchronous calls are not included in the 
+best \tactictoe strategy. The ATP \eprover is run in auto-schedule mode with 
+128 premises. The settings for \tactictoe are the following:
+0.05 seconds tactic timeout, an orthogonalization radius of 20, theorem list 
+abstraction with 16 predicted theorems for instantiation, a prior policy 
+coefficient of 0.5, an evaluation radius of 10, an exploration coefficient of 2 
+and priority calls to \metis with a timeout of 0.1 seconds.
 
@@
 
-Table~\ref{theories} compares the success rates of re-proving for different
-\holfour theories. \tactictoe outperforms \eprover on every
-considered theory.
-\eprover is more suited to deal with dense theories such as
-\texttt{real} or \texttt{complex} where a lot of related theorems are available
-and most proofs are usually completed by rewriting tactics.
+The results shows that \tactictoe is able to prove almost twice as much 
+theorems as \eprover. Combining the results of \tactictoe and \eprover we get a 
+69.0\% success rate which is significantly above the 5O\% success rates 
+of hammers on this type of problems~\cite{tgck-cpp15}. Moreover, \tactictoe is 
+running a single set of parameters (strategy), where as hammers and ATPs have 
+been optimized and rely on a wide range of strategies.
+
+\paragraph{Reconstruction}
+Tactic proofs produced by \tactictoe during this experiment are all verifiable 
+in \holfour. By the design of the proof search, reconstruction of \tactictoe 
+proof succeeds, unless one of the tactic modifies the state of \holfour in a 
+way that changes the behavior of a tactic used in the final proof. Such an 
+event has not yet occurred.
+More details on how a proof is extracted from the search tree is given in 
+Section~\ref{sec:proofdisplay}.
+For \eprover, reconstruction rate varies between 95 and 99 percent, depending 
+on the reconstruction method \cite{}\todo{CK}.
 
@@
 
+The Table~\ref{theories} compares the success rates of re-proving for different
+\holfour theories. \tactictoe outperforms \eprover on every
+considered theory.
+\eprover is more suited to deal with dense theories such as
+\texttt{real} or \texttt{complex} where a lot of related theorems are available
+and most proofs are usually completed by rewriting tactics. Thanks to its 
+ability to re-use custom-built tactics, \tactictoe 
+largely surpasses \eprover on theories relying on inductive terms and 
+simplifications sets such as \texttt{arithmetic}, \texttt{list} 
+and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to 
+apply induction, a task at which ATPs are known to struggle with.
+
 %\subsection{Recent improvements}\todo{maybe remove this subsection}
@@
 
-\subsection{Complexity of the proofs}
+\subsection{Complexity of the proof search}
 
@@
 }
-
-Estimate the difficulty of the proof by the number of tactic units.
 
@@
 \end{tikzpicture}
-\caption{Number of problem solved in less than x seconds.}
+\caption{\label{fig:c1} Number of problems solved in less than x seconds.}
 \end{figure}
-\eprover uses at least two strategies where as \tactictoe only uses one set
-of parameters.
+
+We first investigate how \tactictoe and \eprover scales as the search time 
+grows in Figure~\ref{fig:c1}.
+In 10 seconds, \tactictoe solves 90 percent of the problem it can 
+solve in 60 seconds. The analysis is a bit different for \eprover. Indeed,
+we can clearly deduce from the bump at 30 seconds that \eprover is using at 
+least two strategies and probably more. Strategies are a useful method to 
+fight the exponential decline in the number of newly proven theorems over 
+time. Therefore, integrating strategy scheduling in \tactictoe could be 
+something to be experimented with. Overall, we recommend using \tactictoe and 
+\eprover with a 10 seconds timeout. It is possible also to run them with a 
+longer time limit, but we advise to do so in parallel to avoid waiting time.
 
@@
 validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
-graph that most of the proofs are short. The proofs found by tactictoe follows
-about the same
+graph that most of the human proofs are short. The proofs found by tactictoe 
+follows about the same
 distribution. This means that if \tactictoe finds a proof, there is about a 50
 percent chance that it will be shorter than what a human would come up with.
-In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
-see how well \tactictoe and \eprover performs on increased level of
-difficulties. As expected, the longer the original proof is, the harder it is
-for \tactictoe and \eprover to re-prove the theorem by their own mean.
-The performance of \tactictoe is compelling with more than half of the theorem
-that required a proof of length six being reproven. It is also consistently
-better than \eprover for any proof length.
 
@@
 \end{figure}
+
+In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
+see how well \tactictoe and \eprover performs on increased level of
+difficulties. As expected, the longer the original proof is, the harder it is
+for \tactictoe and \eprover to re-prove the theorem by their own mean.
+The performance of \tactictoe is compelling with more than half of the theorem
+that required a proof of length six being reproven. It is also consistently
+better than \eprover for any proof length.
 

### 2018-02-19 tactictoe_jv.tex
intergrating Lasse's comments

@@
 For example, if the list of subgoals is empty, then calling the validation function on the empty list yields the originally goal sequent as a theorem.
-In this way, tactics implement the plumbing of backwards, goal-directed, proof construction.
-\todo{do not mention validation? RK: it depends whether these details are helpful for understanding or not}
+In this way, tactics implement the plumbing of backwards, goal-directed, proof 
+construction. 
+Because validation functions are only important to check the final proof, we 
+omit them during the description of proof search algorithm. And we note $t(g)$ 
+the lists of goals produced by the application of the tactic $t$ to the goal 
+$g$.
 
@@
 \item Proof guidance required by MCTS is given by predictors for
-three kind of objects: tactics, theorems and list of goals.
+three kinds of objects: tactics, theorems and lists of goals.
 \item The orthogonalization process that eliminates redundant tactics is
@@
 
-\section{Preliminaries}\label{sec:prelim}
-
-The different components of \tactictoe revolve around the concept of a
-search tree. The following definitions are made relative to a static search
-tree, the description of its evolution during proof search will be given in
-Section~\ref{sec:proofsearch}.
-
-\begin{definition}(search tree)\\
+\section{Search tree}\label{sec:prelim}
+
+The application of tactic to a goal produces a list of goals. Each of the 
+created goals can in turn be the input other tactics. Moreover, it is possible 
+to try multiple tactics on the same goal. In order to keep track of the 
+progress made so far from the initial goal we want to solve, we organize
+goals and tactics into a graph with list of goals as nodes and tactics as edges
+(See Definition~\ref{def:stree}).
+
+In this section, we only give a description of the graph at a given 
+moment of the search after a certain amount of tactic applications. 
+The exploration of the tree is decided by MCTS algorithm in 
+Section~\ref{sec:proofsearch}, which is guided by a tactic prediction algorithm 
+presented in Section~\ref{sec:features}.
+
+\begin{definition}\label{def:stree}(search tree)\\
 A search tree $\mathfrak{T}$ is a sixtuple
@@
  and $\mathbb{A}$ is a set of nodes representing lists of goals.
-\item $T$ is a total function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It takes a
-goal and return the set of tactics already applied to this goal.
-\item $G$ is a total function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$. It takes a
-node and returns the list of goals of this node.
+\item $T$ is a function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It 
+takes a goal and return the set of tactics already applied to this goal.
+\item $G$ is a total function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$. 
+It takes a node and returns the list of goals of this node.
 \item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to
 $\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the
-application of the tactic to the goal. Its support $\mathit{Support}(A)$ is by
-definition $\lbrace (g,t) \in
-\mathbb{G} \times \mathbb{T}\ |\ t \in T(g)
-\rbrace$.
-\item If $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$. This is the
-main condition. The result of the application of a tactic $t$
-to a goal $g$ is exactly the list of goals contained in the node produced by
+application of the tactic to the goal. Its support $\mathit{Support}(A)$ is 
+defined as $\lbrace (g,t) \in \mathbb{G} \times \mathbb{T}\ |\ t \in T(g) 
+\rbrace$. And if $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$. This 
+means that the output $t(g)$ is exactly the list 
+of goals contained in the node produced by
 $t$ from $g$.
@@
 
-Additional constraints such as acyclicity and the fact that there is exactly
-one root node are properties of this graph that are preserved during the
+Acyclicity and the fact that there is exactly
+one root node containing only the initial goal are additional constraints to 
+the search tree that are preserved during the
 application of the MCTS algorithm. That is why we call it a tree, although this
@@
 
-If no explicit order is given, we assume that any set is equipped with an
-arbitrary order. In this way, we can access the $n^{th}$ member of a set $s$ by
-writing $s_n$. In Figure~\ref{fig:choice}, $(G(a_j))_i$ is written as $g_j^i$.
+If no explicit order is given, we assume that the sets 
+$\mathbb{T},\mathbb{G},\mathbb{A}$ are equipped with an
+arbitrary order. In Figure~\ref{fig:choice}, a part of the a search tree is 
+depicted, with goals represented by circles, nodes by rectangles and tactics by 
+arrows.
 
@@
 \begin{tikzpicture}[node distance=0.7cm]
-\node [tcircle] (0) {$g_0^i$};
+\node [tcircle] (0) {$g_i$};
 \node [right of=0] (0r) {.\ .\ .};
-\node [tcircle,right of=0r] (0rr) {$g_0^n$};
+\node [tcircle,right of=0r] (0rr) {$g_n$};
 \node [left of=0] (0l) {.\ .\ .};
-\node [tcircle,left of=0l] (0ll) {$g_0^1$};
+\node [tcircle,left of=0l] (0ll) {$g_0$};
 \node [draw,fit=(0rr) (0ll)] (0b) {};
@@
 \node [above of=0,node distance=2cm] (2) {.\ .\ .};
-\node [left of=2, tcircle] (2l) {$g_j^1$};
-\node [tcircle,right of=2] (2r) {$g_j^p$};
+\node [left of=2, tcircle] (2l) {$\phantom{g_3}$};
+\node [tcircle,right of=2] (2r) {$\phantom{g_3}$};
 \node [draw, fit=(2l) (2r)] (2b) {};
@@
 
-\node [tcircle, left of=2l, node distance=3cm] (1r) {$g_1^q$};
+\node [tcircle, left of=2l, node distance=3cm] (1r) {$\phantom{g_3}$};
 \node [left of=1r] (1) {.\ .\ .};
-\node [tcircle, left of=1] (1l) {$g_1^1$};
+\node [tcircle, left of=1] (1l) {$\phantom{g_3}$};
 \node [draw, fit=(1l) (1r)] (1b) {};
@@
 
-\node [right of=2r,node distance=3cm] (3l) {$\phantom{g_3^p}$};
+\node [right of=2r,node distance=3cm] (3l) {$\phantom{g_3}$};
 \node [right of=3l] (3) {\phantom{.\ .\ .}};
-\node [right of=3] (3r) {$\phantom{g_3^p}$};
+\node [right of=3] (3r) {$\phantom{g_3}$};
 \node [fit=(3l) (3r)] (3b) {};
@@
 \draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
-\draw[-to,thick] (0) to node[xshift=-10](t2){$t_j$} (2b);
+\draw[-to,thick] (0) to node[xshift=-4](t2){$t_j$} (2b);
 \draw[-to,dotted,thick] (0) to (1m);
 \draw[-to,dotted,thick] (0) to (2m);
-\draw[-to,dotted,thick] (0) to (3b);
+\draw[-to,dotted,thick] (0) to node[xshift=15](t2){$t_m$} (3b);
 \end{tikzpicture}
 \end{center}
-\caption{A node of a search tree and the children of one of
-its goal.}
+\caption{A node of a search tree $a_0$, the list of goals $g_0 \ldots g_n$ it 
+contains and the nodes $a_1 \ldots a_m$ derived from the application of the 
+tactics $t_1 \ldots t_m$ to $g_i$.}
 \label{fig:choice}
@@
 
-Starting from the parent node, we can construct children of a goal of
-this node by choosing a goal $g_i^0$ on this node and choosing tactics
-$t_1,t_2,\ldots$ to
-apply to this goal. Here the result produces new nodes each containing a list
-of goals to be proven called obligations. In each node of the proof tree,
-a goal can either be
-solved, open or pending (see Definition~\ref{def:solved}).
-
-In the following, concepts and properties are defined in the context of a
-search tree $\mathfrak{T}$.
-
+In the following, properties are defined in the context of a
+search tree $\mathfrak{T}$. In Definition~\ref{def:solved}, we specify 
+different status for a goal: open, pending or solved. There, we define 
+what a solved goal and solved node are by mutual recursion on the number of
+steps it takes to solve a goal and a node.
 \begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal,
@@
 \mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\
-\exists a \in A(g).\ a \in \mathbb{A}^{0} \rbrace\\
+\exists t \in T(g).\ A(g,t) \in \mathbb{A}^{0} \rbrace\\
 \mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\
@@
 \mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\
-\exists a \in A(g).\ a \in \mathbb{A}^{n+1} \rbrace \\
+\exists t \in T(g).\ A(g,t) \in \mathbb{A}^{n+1} \rbrace \\
 \mathbb{A}^\infty &=_{def} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \
@@
 
-An open goal is an unsolved goal that was tried at least once by a tactic. The
-set of open goals can be formally defined as:
-\[\lbrace g \in G\ |\  T(g) \neq \emptyset \wedge g \notin \mathbb{G}^\infty
-\rbrace\]
-
-The set of pending goals is the remaining set of goals in the tree. They are
-goal that were produced by tactics that were not explored yet. This set can
-also be also defined by:
-\[\lbrace g \in G\ |\  T(g) = \emptyset \rbrace\]
-
-To facilitate the description  of our algorithms we make a slight adjustment
-to the definition of an open goal. If a node $a$ does not contain any open
-goal, the first pending goal (according
-to some preset order) of $a$ is also considered to be an open goal. In this
-way, each unsolved node contains a unique open goal.
+We call a goal (respectively node) unsolved if it is not in the set of solved  
+goal (respectively node).
+The open goal of a node is the first unsolved goal of this node according to 
+some preset order. The other unsolved nodes are called pending nodes.
+If a node is unsolved, then it contains a unique open goal.
+During MCTS exploration, we do not explore pending goals of an unsolved node 
+before its open goal is solved which justifies our terminology. In other words, 
+our search tree has the following property: if $g$ is a pending node then 
+$T(g)$ is empty.
+
 \end{definition}
 
-\todoi{some text to add here}
-
-\begin{definition}\label{def:desc}(Children, descendant, active node)\\
+In Definition~\ref{def:desc}, we define relations between goals 
+and nodes in terms of parenthood.
+
+\begin{definition}\label{def:desc}(children, descendant, ancestors)\\
 The children of a goal $g$ is a set of nodes defined by:
@@
 
-An active node is by definition a node that is not a descendant of a
-solved node. Non-active nodes are not worth exploring.
 \end{definition}
 
-We can now judge the productivity of a tactic by its contribution to the
-search tree.
-
-\begin{definition} (Productive tactic)
+In Definition~\ref{def:prod}, we judge if a tactic is productive from its 
+contribution to the search tree.
+
+\begin{definition}\label{def:prof} (Productive tactic)
 The application of a tactic on an open goal is called productive if and only if
@@
 interrupt tactics after a short amount of time (0.05 seconds).
-\item It does not loop. Its output does not contain any goals that appears in
-the ancestor of the node of the current
-goal.
-\item It is not a parallel step. Its output is not equal or a superset
-of a list of goals produced from the same open goal.
+\item It does not loop, i.e its output does not contain any goals that appears 
+in the ancestor of the node of the current goal.
+\item It is not a parallel step, i.e its output is not a superset
+of a list of goals produced from the same goal.
 \end{itemize}
@@
 predictions per second, while maintaining very high prediction quality~\cite{femalecop}.
-Here it is used
-to predict three different kinds of objects:
- tactics, theorems and list of goals. The number of predicted objects $k$ is
- called the \emph{prediction radius}.
-
-We first introduce specificities associated with the prediction of each object.
+We rely on this supervised machine learning algorithm to predict three 
+different kinds of objects:
+ tactics, theorems and lists of goals. The number of predicted objects $k$ is
+ called the \emph{prediction radius}. For a goal $g$, we note
+$\mathit{Predictions}^{\mathit{tactic}}_k (g)$ (respectively
+$\mathit{Predictions}^{\mathit{theorem}}_k (g)$
+and $\mathit{Predictions}^{\mathit{goal\_list}}_k (g)$) the $k$ tactics 
+(respectively theorems and lists of goals) selected by our prediction algorithm.
+
+We introduce specificities associated with the prediction of each object.
 In particular, we present the dataset from which objects are selected and the
-purpose of the selection process.
-We then explain the construction of the similarity measure by which relevant
-objects are predicted.
+purpose of the selection process. Details on the similarity measure backing the 
+predictions is described in Section~\ref{sec:features} and two methods for 
+improving the quality of the predictions are presented in 
+Section~\ref{sec:ortho} and Section~\ref{sec:synthesis}.
 
@@
 \paragraph{Theorems as Arguments of Tactics}
-We first collect a set of theorems for our theorem predictor to select from.
+We first collect a set of theorems for our predictor to select from.
 It includes the \holfour theorem database and theorems from the local namespace.
 Predicted tactics during proof search and orthogonalization may include
-tactics where arguments (list of theorems) have been erased (see Section
+tactics where arguments (lists of theorems) have been erased (see Section
 \ref{sec:synthesis}).
@@
 
-\paragraph{List of Goals as Output of Tactics}
-A dataset of tactic outputs is compiled during orthogonalization (see
+\paragraph{Lists of Goals as Output of Tactics}
+A dataset of tactic outputs is compiled during orthogonalization.
+This set is separated into positive and negative examples (see
 Section~\ref{sec:evaluation}).
-This set is separated into positive and negative examples.
-List of goals that subsumes the result of an original tactic on a
+List of goals that are a subset of the result of a recorded tactic on a
 targeted goal are considered positive. During proof search, given a list of
@@
 
-\subsection{Feature Extraction}\label{sec:features}
-All the three kinds of considered objects to predict correspond to (different representations of)
-mathematical
-formulas. Therefore, we start by describing features for terms. From there, we
-extend the extraction mechanism
+\section{Proof Search}\label{sec:proofsearch}
+
+Despite the best efforts of the prediction algorithms, a selected tactic may
+not solve the current goal, proceed in the wrong direction
+or even loop. For that reason, predictions need to be accompanied by a
+proof search mechanism that allows for backtrac king and
+can choose which proof tree to extend next and in which direction.
+Our search algorithm is a Monte Carlo tree search~\cite{} algorithm that relies
+on a prior evaluation function and a prior policy function. Those priors are
+learned through direct supervised learning for the policy and via the data
+accumulated during orthogonalization for the evaluation. Therefore, this MCTS
+algorithm does not need to rely on roll outs for evaluation. The
+policy and evaluation are estimated by the simple supervised k-NN algorithm
+here instead of neural networks trained through reinforcement learning in
+AlphaGo Zero~\cite{silver2017mastering}.
+A characteristic of the MCTS algorithm is that it offers a good trade-off
+between exploitation and exploration. This means that the algorithm searches
+deeper more promises branches and leaves enough time for the exploration
+of less likely alternatives.
+
+We first describe the proof search algorithm and explain later how to compute 
+the prior policy \texttt{PriorPolicy} and prior evaluation 
+\texttt{PriorEvaluation} functions.
+
+\subsection{Proof exploration}
+
+The proof search starts by creating an initial search tree. The search tree 
+then evolves by the repetitive applications of MCTS steps. A step in the main 
+loop of MCTS consists of three parts: node selection, node extension, and 
+backpropagation. And the decision to terminate the algorithm is taken during
+resolution.
+
+\paragraph{Initialization}
+The input of the algorithm is a goal $g$ (also called conjecture) that we 
+want to prove.
+Therefore the search tree starts with only one node the list of goals
+$[g]$ and an ordered list of predicted tactics for the goal $g$.
+%Then, the MCTS algorithm alternates node selection steps and node extension
+%steps which grow the search tree.
+
+%A similar trade-off between exploration and exploitation exists in algorithm
+%like Monte Carlo. However, Monte Carlo tree search metrics are based on many
+%simulations and do not reflect our exploration strategy that only
+\paragraph{Node selection}
+Through node extension steps the search tree grows and the number of paths to
+explore increases. To decide which node to extend next, a evolving value for
+each node is given in Definition~\ref{def:value}. The algorithm that performs
+node selection starts from the root of the search tree. From the current node,
+the function \texttt{CompareChildren} choses among the children of its open 
+goal the one with the highest value.
+
+If the highest children value is higher than the widening policy, then the 
+selected child
+becomes the current node, otherwise the final result of node selection is the
+current node. The node selection algorithm can be written as the following
+pseudo-code:
+
+\begin{lstlisting}[language=SMLSmall]
+CurrentNode = Root(Tree);
+while true do
+  if Children(CurrentNode) = emptyset then break;
+  (BestChild, BestValue) = CompareChildren (CurrentNode);
+  if WideningPolicy (ChosenNode) >= BestValue then break;
+  CurrentNode = BestChild
+end;
+return CurrentNode;
+\end{lstlisting}
+
+\begin{definition}\label{def:value}(Value)
+The value of the $i^th$ child $a_i$ of the open goal of a parent node $p$ 
+is determined by:
+\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} * 
+\mathit{Exploration}(a_i)\]
+
+The number $\mathit{Failure}(a_i)$ is the amount of failures that occurred 
+during node extension from descendants of $a_i$ .
+The current evaluation $CurEvaluation(a_i)$ is the average evaluation of
+all descendants of $a_i$ including node extension failures.
+
+\[\mathit{CurEvaluation}(a_i) =
+  \sum_{a' \in Descendant(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
+  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]
+
+The exploration term is determined by the prior policy and the current policy.
+The current policy is calculate from the number of times a node $x$ has been 
+traversed during node selection, noted $\mathit{Visit}(x)$.
+
+\[\mathit{Exploration}(a_i) = 
+\frac{\mathit{PriorPolicy}(a_i)}{\mathit{CurPolicy}(a_i)}\]
+
+\[\mathit{CurPolicy}(a_i) = \frac{1 + 
+\mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
+The policy is approximatively the  percentage of time a node was visited.The
+square root skews this probability to favor even more exploration of nodes with
+few visits.
+
+The coefficient $c_{exploration}$ is experimentally determined and adjusts the
+trade-off between exploitation and exploration.
+\end{definition}
+
+\paragraph{Node extension}
+Let $a$ be the result of node selection.
+If $a$ is a solved node or the descendant of a solved node, then extending $a$ 
+is useless and the algorithm reports a failure for this MCTS step. 
+%In our 
+%experiments, filtering such nodes before node selection is detrimental.
+If $a$ is not $solved$, it applies the best untested tactic $t$ on the open 
+goal $g$ of this
+node according to the prediction order for $g$. If no such tactic exists or if
+$t$ is not productive, a failure is returned.
+If node extension succeeds, a new node containing $t(g)$ is added to the 
+children of $a$.
+
+\paragraph{Backpropagation}
+During backpropagation we update the statisitics of all the nodes traversed or
+created during this MCTS step:
+\begin{itemize}
+\item Their number of visit is incremented.
+\item If node extension failed, their failure count is incremented.
+\item If node extension succeeded, they inherit the evaluation of the created
+child.
+\end{itemize}
+
+These changes updates the current evaluation and the current policy of the
+traversed nodes.
+After completing backpropagation, the process loops back to node
+selection.
+
+\paragraph{Resolution}
+The search ends when the algorithm reaches one of these 3 conditions:
+\begin{itemize}
+\item It finds a proof, i.e. the root node is solved. In this case,
+the search returns a minimized and
+prettified proof script (see Section~\ref{sec:proofdisplay}).
+\item It saturates, i.e. there is no tactics to be applied to any open goal.
+This happens less than 10 times in the full-scale experiment.
+\item It exceeds a given time limit set by default to 60 seconds.
+\end{itemize}
+
+\subsection{Supervised learning guidance}
+Our MCTS algorithm relies on the guidance of two supervised learning methods,
+the prior policy decides the order in which tactic should be tried and the 
+prior evaluation estimates if a branch is fruitful by analysing tactic 
+outputs. Both priors influence the rates at which branches of the search tree 
+are explored.
+
+\paragraph{Prior policy}\label{sec:policy}
+Co-distances between goals produced by the tactic predictor are hard
+to translate into a prior policy, so we rely solely on the induced order of
+tactics to estimate the probability with which each tactic should be tried.
+Let $\mathfrak{T}$ be a search tree after a number of MCTS steps.
+Let $p$ be a selected node in $\mathfrak{T}$ and $g$ its open goal. We order 
+the list of $n$ productive tactics already
+applied to $g$ by their similarity score so that $T(g)= \lbrace
+t_0,\ldots,t_{n-1} \rbrace$.
+Let $c_{policy}$ be a constant heuristic that estimates how likely a tactic
+$t_i$ is to prove the goal.
+We can now calculate of a prior policy of a child $a_i$ produced by the tactic
+$t_i$ by:
+\[PriorPolicy(a_i) = (1 - c_{policy})^{i} * c_{policy}\]
+
+In order to include the possibly of trying more tactics on $g$ we define the
+widening policy on the parent $p$ for its open goal $g$ to be:
+
+\[WideningPolicy(p) = (1 - c_{policy})^{n} * c_{policy}\]
+
+\paragraph{Prior evaluation}\label{sec:evaluation}
+
+We now concentrate on the definition a reasonable evaluation function for the
+output of tactics. Theoretically, an interesting list of goals is one
+that has short proofs for each of its goals, and we could estimate their 
+lengths from previous proof searches. However, this
+approach generates too much data and slows down the proof search. So, we only 
+collect output of tactics tested during orthogonalization. We
+declare a list of goals $l$ to be positive if it is has been produced by the
+winner of an orthogonalization competition. 
+The set of positive examples in 
+$\mathit{Predictions}^{\mathit{goal\_list}}_k(l)$ is noted 
+$\mathit{Positive}^{\mathit{goal\_list}}_k(l)$.
+We chose $k=10$ as a default evaluation radius in our experiments.
+And we evaluate a node $a$ through the list of goals $G(a)$ it
+contains using the prior evaluation function:
+\begin{align*}
+\mathit{PriorEvaluation}_k (a) &= 
+  \frac{card\ \mathit{Positive}^{\mathit{goal\_list}}_k(G(a)}{k}\\
+\end{align*}
+
+\subsection{ATP Integration}~\label{sec:atp}
+General-purpose proof automation mechanisms which combine proof translation to
+ATPs with machine learning (``hammers'') have become quite successful in
+enhancing the automation level in proof assistants~\cite{hammers4qed}.
+As external automated reasoning techniques sometimes outperform the combined
+power of tactics, we would like to combine the \tactictoe search with
+general purpose provers such as the ones found in \holyhammer for
+\holfour~\cite{tgck-cpp15}.
+
+Such a prover already exists in \holfour and is called \metis. It is already
+recognized by the tactic selection mechanism in \tactictoe and can have its
+premises predicted dynamically when in an abstracted form. Nevertheless,
+we think that the performance of \tactictoe can be boosted by giving the ATP
+\metis a special status. This arrangement consists of always predicting
+premises for \metis, giving it a slightly higher timeout and
+trying it first on each open goal. These modifications only induce an
+linear overhead on the number nodes as a general purpose prover does not create
+goal obligations.
+
+The integration of an external prover is also possible. And
+\eprover~\cite{eprover} is the one we experimented with.
+To this end, we created a tactic which expects a number of premises as an 
+argument
+and calls \eprover on the translated first order problem. We then extract the
+premises that were required in the \eprover
+proof and reconstruct it inside \holfour with \metis. Giving a special status
+to external
+provers is essential as external ATP calls do not appear in human
+proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and 
+a larger number of predicted premises (128 for \eprover) and also try them 
+first on open 
+goals. Since calls to ATPs are
+computationally expensive, they are run in parallel and asynchronously. This
+avoids slowing down \tactictoe's search loop. The number of asynchronous calls
+to external provers that can be executed at the same time is limited by the
+number of cores available to the user.
+
+\section{Feature Extraction}\label{sec:features}
+We explain the construction of the similarity measure by which relevant
+objects are predicted
+
+We predicts tactics through their associated goals. So, the features 
+for each kind of object can be extracted from (different representations of)
+mathematical formulas. We start by describing features for \holfour terms. From 
+there, we extend the extraction mechanism
 to goals, theorems and lists of goals. Duplicated features are always removed
@@
 \item type constructors present in the types of constants and variables,
-\item subterms with all variables replaced by a single place holder $V$.
+\item subterms with all variables replaced by a single place holder $V$,
 \item names of the variables.
 \end{itemize} 
-
-Since goals and theorems are represented by a couple of a list of terms
-(assumptions) and a term (conclusion). We compute their features by extracting
-the features of all terms describing them. Additionally, we distinguish between
+Goals (respectively theorems) are represented by a couple of a list of 
+terms (assumptions) and a term (conclusion). So, we distinguish between
 features of the assumptions and features of the conclusion by adding a
-different tag to elements of each set. Features of a list of goals are
-the union of the features of each goal in the list.
+different tag to elements of each set. The union of all those 
+features are considered the feature of a goal (respectively theorem). From 
+that, we can construct features for a list of goals by
+from the union of the features of each goal in the list.
 
@@
 through their respective feature sets $f_1$ and $f_2$.
-%The similarity function $sim$ computed by the k-NN algorithm is analogous
-%to the one used in the premise selection task~\cite{ckju-pxtp13}.
 The estimation is based on the frequency of the features in the intersection of
@@
 even more credence
-to rare features which was found experimentally optimal in \cite{ckju-pxtp13}.
-
-The first co-distance $sim_1$ adds the weight of each shared
+to rare features. This modification was found experimentally optimal in 
+\cite{ckju-pxtp13}.
+
+The first co-distance $sim_1$ sums the weight of each shared
 features to compute the total score.
-In a second co-distance $sim_2$, we additionally take into account
-the total number of features to reduce the seemingly unfair advantage of big
+In a second co-distance $sim_2$, we additionally take
+the total number of features into account, to reduce the seemingly unfair 
+advantage of big
 feature sets in the first scoring function.
 \[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap
-f_1}{\text{tfidf}(f)^{6}}}\]
+<f_1}{\text{tfidf}(f)^{6}}}\]
 \[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
 f_1}{\text{tfidf}(f)^{6}}}}
-{ln (e + card(f_0) + card(f_1))}\]
-
-In our setting, making prediction for an object $o$ consists of sorting a set
+{ln (e + \mathit{card} f_0 + \mathit{card} f_1)}\]
+
+In our setting, making predictions for an object $o$ consists of sorting a set
 of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems
-and $sim_2$ for list of goals. The reason why $sim_1$ is adapted for
-predicting tactics is that tactics effective on large goal $g$ are often also
+and $sim_2$ for lists of goals. The reason why $sim_1$ is used for
+predicting tactics is that tactics effective on a large goal $g$ are often also
 suitable for a goal made of sub-formulas of $g$. The same monotonicity
-heuristic can justify the use of $sim_1$ in theorems. Indeed in practice, a
+heuristic can justify the use of $sim_1$ for theorems. Indeed, in practice a
 large theorem is often a conjunction of theorems from the same domain.
@@
 conjecture $c$, we preselect 500 tactic-goal pairs and 500 theorems and a
-larger number of list of goals induced by the selection of tactic-goal pairs.
-Preselected objects will
-be the only objects available to \tactictoe's search algorithm for proving $c$.
-
-The first idea is to select tactic-goal pairs and theorems by their distance to
-the conjecture. However, this selection may not be adapted to later stages of
-the proof where goals may have derived significantly from the
-initial conjecture $c$. In order to
+larger number of lists of goals induced by the selection of tactic-goal pairs.
+Preselected objects are the only objects available to \tactictoe's search 
+algorithm for proving $c$.
+
+%During search, preselected tactics are reordered for 
+%each open goal creating the prior policy of our MCTS algorithm (see 
+%Section~\ref{sec:policy}).
+
+The first idea is to select tactic-goal pairs and theorems by their similarity 
+with $c$ (i.e $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
+$\mathit{Predictions}^{\mathit{theorem}}_{500} (c)$).
+However, this selection may not be adapted to later stages of
+the proof where goals may have drifted significantly from $c$. In order to
 anticipate the production of diverse goals, our prediction during preselection
-takes into account dependencies between objects of the same dataset.
-The dependence relation is assymetric. Through the relation, each object has
+takes dependencies between objects of the same dataset into account.
+This dependency relation is assymetric. Through the relation, each object has
 multiple children and at most one parent.
-Once a dependency relation is established, we can calculate a dependency score
-for each object which is the maximum of its similarity score and the similarity
-score of its parent. In the end, the 500 entries with highest dependency score
-are preselected in each dataset.
-
-In the following, we give a mathematical definition for the dependency relation
-between labels for tactic-goal pairs and theorems.
+Once a dependency relation is established we can calculate a dependency score
+for each object, which is the maximum of its similarity score 
+and the similarity score of its parent. In the end, the 500 entries with 
+highest dependency score are preselected in each dataset.
+
+We now give a mathematical definition of the dependency relation for each kind 
+of object.
 
@@
 D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
-(\hat{t},\hat{g}) \in D_n.\ g \in \hat{t}(\hat{g}) \rbrace  \\
+(t',g') \in D_n.\ g \in t'(g') \rbrace  \\
 D_\infty &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
 \end{align*}
-where $\hat{t}(\hat{g})$ denotes the resulting set of goals produced by the
-application
-of the tactic $\hat{t}$ to the goal $\hat{g}$.
 \end{definition}
@@
 Let $\mathfrak{T}$ be the set of recorded theorems.
-The dependencies of a theorem $t$ are the set of theorems in
-$\mathfrak{T}$ appearing in the proof of $t$ as recorded in
-\cite{tgck-cpp15}.
+The dependencies of a theorem $t$ are the set of top-level theorems in
+$\mathfrak{T}$ appearing in the proof of $t$. These dependencies are
+recorded by tracking the use of top-level theorems through the inference 
+kernel~\cite{tgck-cpp15}.
 \end{definition}
@@
 \begin{remark}
-The relation chosen for tactic-goal pairs is a transitive
-closure, which is not the case for theorem dependencies.
+The relation chosen for tactic-goal pairs is transitively closed, which is not 
+the case for theorem dependencies.
 \end{remark}
 
-To preselect a list of goal $l$, we consider the tactic input $g$ that was
-at the origin of the production of $l$. The list of goals $l$ is preselected if
-and only if $g$ appears in the preselected tactic-goal pairs. This has the
+To preselect a list of goal $l$ from our database of lists of goals, we 
+consider the tactic input $g$ that was at the origin of the production of $l$ 
+during orthogonalization. The list $l$ 
+is preselected if and only if $g$ appears in the 500 preselected tactic-goal 
+pairs. This has the
 desirable effect that negative and positive examples derived from the same
 input goal are chosen or discarded together.
+\todo{this paragraph was confusing for Lasse}
 
@@
 is undesirable, as it leads to inefficiency in automated proof search.
-To solve this problem, we do not directly assign an input goal $g$ to its
-associated tactic $t$, but organize a competition between the $n$ closest
-tactic-goal pairs, noted $P_n(g)$.
-The winner is the tactic that subsumes the original tactic on $g$ and that
+To solve this problem, we modify how the database of tactics is build.
+Each time a new tactic-goal pair $(t,g)$ is extracted from a tactic proof and 
+about to be recorded, we consider if there does not exists already a better 
+tactic for $g$ in our database. To this end, we organize a competition between 
+the 
+$k$ closest tactic-goal pairs $\mathit{Predictions}^{\mathit{tactic}}_k(g)$. 
+In our experiments, 
+the default orthogonalization radius is $k=20$.
+The winner is the tactic that subsumes (see Definition~\ref{def:tacsub}) the 
+original tactic on $g$ and that
 appears in the largest number of tactic-goal pairs in the database.
 The winning tactic $w$ is then associated with $g$, producing the pair $(w,g)$
-that replaces the original pair $(t,g)$ in the database. As a
-result, already successful tactics with a large coverage are preferred, and new
-tactics are
-considered only if they provide a different contribution. In the following, we
-give a formal definition of the concepts of subsumption and coverage that are
-required for expressing the orthogonalization algorithm.
+and is stored in the database instead of the original pair $(t,g)$. 
+As a result, already successful tactics with a large coverage are preferred, 
+and new tactics are considered only if they provide a different contribution. 
+We now give a formal definition of the concepts of subsumption and coverage 
+that are required for expressing the orthogonalization algorithm.
 
 \begin{definition} (Coverage)\\
-Let $\mathfrak{F}$ be the database of tactics. We can define the
+Let $\mathfrak{T}$ be the database of tactic-goal pairs. We can define the
 coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times
@@
 appears in
-$\mathfrak{F}$. Expressing this in a formula, we get:
-  \[\textit{\mbox{Coverage}}(t) =_{def} card\ \lbrace (x,g)\ |\ (x,g) \in
-  \mathfrak{F}
-  \wedge x
-  = t
-  \rbrace  \]
+$\mathfrak{T}$. Expressing this in a formula, we get:
+  \[\textit{\mbox{Coverage}}(t) =_{def} card\ \lbrace g\ |\ (t,g) \in 
+  \mathfrak{T} \rbrace \]
 Intuitively, this notion estimates how general a tactic is by counting the
@@
 \begin{definition} (Goal subsumption)\\
-A goal subsumption $S$ is a partial relation on a set of goals.\\
-Given two goals $g_1$ and $g_2$, we note $g_1 \ge g_2$ when  $(g_1,g_2) \in
-S$.\\
-Assuming we have a way to test if a goal is a consequence of another, a
-possibly useful subsumption definition would be:
-\[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is a consequence of } g_2\]
+A goal subsumption $\le$ is a partial relation on a set of goals.\\
+Assuming we have a way to estimate the number steps required to solve a goal,  
+a possibly useful subsumption definition could be:
+\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ has a shorter proof than } 
+g_2\]
 By default, we choose a minimal subsumption defined by:
-\[g_1 \ge g_2  \Leftrightarrow_{def} g_1 \mbox{ is }\alpha\mbox{-equivalent
+\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ is }\alpha\mbox{-equivalent
 to } g_2\]
@@
 A goal list subsumption is an partial relation on lists of goals.
-Given two list of goals $l_1$ and $l_2$, we can define it from goal
-subsumption $\ge$ by:
-\[l_1 \ge l_2  \Leftrightarrow_{def} \forall g_2 \in l_2\ \exists g_1 \in l_1\
-g_1 \ge g_2\]
+Given two lists of goals $l_1$ and $l_2$, we can define it from goal
+subsumption $\le$ by:
+\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_2 \in l_2.\ \exists g_1 \in l_1.\
+g_1 \le g_2\]  
 \end{definition}
@@
 This allows us to define subsumption for tactics.
-\begin{definition}(Tactic subsumption)\\
+\begin{definition}\label{def:tacsub}(Tactic subsumption)\\
 Given two non-failing tactics $t_1$ and $t_2$ on $g$, a tactic $t_1$ subsumes a
-tactic $t_2$ on a goal $g$, noted $\ge_g$, when:
-
- \[t_1 \ge_g t_2 \Leftrightarrow_{def} t_1(g) \ge t_2(g)\]
+tactic $t_2$ on a goal $g$, noted $\le_g$, when:
+
+ \[t_1 \le_g t_2 \Leftrightarrow_{def} t_1(g) \le t_2(g)\]
 
@@
 
-\[\mathbb{O}(t,g) = \argmax_{x\ \in\ P_n(g) \cup \lbrace t \rbrace} \lbrace
+\[\mathbb{O}(t,g) = \argmax_{x\ \in\ 
+\mathit{Predictions}^{\mathit{tactic}}_k(g) \cup 
+\lbrace t 
+\rbrace} \lbrace
 \textit{\mbox{Coverage}}(x)\
-|\ x \ge_g t\rbrace\]
-
-As a consequence $(\mathbb{O}(t,g),g)$ replaces $(t,g)$ in the database of
-tactics.
+|\ x \le_g t\rbrace\]
+
+All in all, a database build with the orthogonalization process contains 
+$(\mathbb{O}(t,g),g)$ instead of $(t,g)$.
 
@@
 create tactics that are more general but have the same effect as the
-ungeneralized one.
+ungeneralized one. The generalized tactics are typically slower that their 
+counterpart, but the added flexibility is worthwhile in practice. Moreover, 
+since we impose a timeout on each tactic (0.05 seconds by default), very slow 
+tactics fails and thus are not selected.
 
@@
 The first step is to abstract arguments of tactics by replacing them by a
-place holder creating an tactic with an unspecified argument.
+place holder, creating a tactic with an unspecified argument.
 Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
@@
 abstractions are performed the more general a tactic is, as many tactics
-become instances of the same abstracted tactic. Yet, it becomes harder and
-harder to predict suitable arguments.
+become instances of the same abstracted tactic. As a consequence, it becomes 
+harder and harder to predict suitable arguments. In our experiments, we create 
+one abstraction $\hat{t}$ for each tactic $t$ by abstracting all subterms of 
+type theorem list in $t$.
 
@@
 to
-instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to
+instantiate the placeholders inside $\hat{t}$. Because it is difficult to
 build a general predictor that effective on each type of argument, we manually
 design different argument predictors for each type. Those predictors are
-usually given a goal has input and returns the best chosen argument for this
-goal. All in all, an abstracted tactics has
-dynamical arguments, it instantiate them differently for different goals.
-
-We made successful argument predictors for one type of arguments: list
-of theorems. It is a very common type and the structure of theorems
-are complex, thus they contain enough information to be predicted accurately.
-We plan to extend our algorithm to term which is another frequent argument
-types but term synthesis is so far too weak to positively
-influence the proof search.
+given a goal $g$ as input and returns the best chosen argument for this
+goal.
+
+Our default algorithm relies on argument predictions for theorems with a 
+radius of 16. It uses the produced list 
+$\mathit{Predictions}^{\mathit{tactic}}_{16}(g)$ to replace all the 
+placeholders 
+in $\hat{t}$. 
+List of theorems is a very common type in \holfour proofs and theorems contain 
+enough information to be predicted accurately.
+We also have positive results with term abstraction and plan to include it soon 
+in our default algorithm.
 
 \paragraph{Selection of Abstracted Tactics}
-As abstracted tactics do not appear in human proof, we need to find a way to
+As abstracted tactics do not appear in human proofs, we need to find a way to
 predict them so that they can contribute to the \tactictoe proof search.
@@
 If $\hat{t}$
-wins, it becomes a a label for $g$ is including in the database of tactic
+wins, it is associated with $g$ and is included in the database of tactic
 features and thus can be predicted during proof search.
-After many competitions, the coverage of $\hat{t}$ may exceed the coverage of
+After many orthogonalization competitions, the coverage of $\hat{t}$ may exceed 
+the coverage of
 $t$. At
@@
 $t$ anymore.
-
-\section{Proof Search}\label{sec:proofsearch}
-Despite the best efforts of the prediction algorithms, a selected tactic may
-not solve the current goal, proceed in the wrong direction
-or even loop. For that reason, the prediction needs to be accompanied by a
-proof search mechanism that allows for backtracking and
-can choose which proof tree to extend next and in which direction.
-Our search algorithm is a Monte Carlo tree search (MCTS) algorithm that relies
-on a prior evaluation function and a prior policy function. Those priors are
-learned through direct supervised learning for the policy and via the data
-accumulated during orthogonalization for the evaluation. Therefore, this MCTS
-algorithm does not need to rely on roll outs for evaluation. The
-policy and evaluation are estimated by the simple supervised k-NN algorithm
-here instead of neural networks trained through reinforcement learning in
-AlphaGo Zero~\cite{silver2017mastering}.
-
-%In this section, we first give a static description of the proof tree and its
-%component.
-We first explain how to compute the policy and evaluation functions that
-influence the behavior of the MCTS algorithm.
-We then describe initialization and resolution of the search as well as the
-different part of a step in the search loop: node selection, node
-extension, backpropagation.
-A characteristic of the MCTS algorithm is that it offers a good trade-off
-between exploitation and exploration. This means that the algorithm searches
-deeper more promises branches and leaves enough time for the exploration
-of less likely alternatives.
-
-\subsection{Prior policy}\label{sec:policy}
-Co-distances between goals produced by the tactic predictor are hard
-to translate into a prior policy, so we rely solely on the order of
-tactics to estimate the probability with which each tactic should be tried.
-Let $\mathfrak{T}$ be a search tree after a number of MCTS steps.
-Let $p$ be a selected node on this tree and $g$ its open goal. By the
-application of MCTS, $T(g)$ is the list of $n$ productive tactics already
-applied to $g$. We order them by their prediction score so that $T(G)= \lbrace
-t_0,\ldots,t_{n-1} \rbrace$.
-Let $c_{policy}$ be a constant heuristic that estimates how likely a tactic
-$t_i$ is to prove the goal.
-We can now calculate of a prior policy of a child $a_i$ produced by the tactic
-$t_i$ by:
-
-\[PriorPolicy(a_i) = (1 - c_{policy})^{i} * c_{policy}\]
-
-In order to include the possibly of trying more tactic on $g$ we define the
-widening policy on the parent $p$ for its open goal $g$ to be:
-
-\[WideningPolicy(p) = (1 - c_{policy})^{n} * c_{policy}\]
-
-\subsection{Prior evaluation}\label{sec:evaluation}
-
-We now concentrate on the definition a reasonable evaluation function for the
-output of tactics. Theoretically, an interesting list of goal is one
-that has a short proof, and we could measure the likelihood of being
-provable in a number of steps from previous proof search. However, this
-approach produces a lot of negatives\todo{CK: cite millions lemma mining
-or/and deep math}. So to reduce this number, we limit
-ourselves to collect output of tactics tested during orthogonalization. We
-declare a list of goals $l$ to be positive if it is has been produced by the
-winner of an orthogonalization competition. Our predictor with radius $r$
-produces a set of $r$ list of goals
-closest to $l$. The
-number of positive example in this set is noted $\mathit{Positive}(r,l)$.
-During the proof search, we evaluate nodes through the list of goals they
-contain using the following evaluation function:
-\begin{align*}
-\mathit{PriorEvaluation}_r (a_i) &= \frac{\mathit{Positive}(r,G(a_i))}{r}\\
-\end{align*}
-
-\subsection{Initialization}
-The input of the algorithm is a goal $g$ that we want to prove.
-Therefore the search tree starts with only one node the list of goals
-$[g]$ and an ordered list of predicted tactics for the goal $g$.
-Then, the MCTS algorithm alternates node selection steps and node extension
-steps which grow the search tree.
-
-\subsection{Node selection}
-
-%A similar trade-off between exploration and exploitation exists in algorithm
-%like Monte Carlo. However, Monte Carlo tree search metrics are based on many
-%simulations and do not reflect our exploration strategy that only
-
-Through node extension steps the search tree grows and the number of paths to
-explore increases. To decide which node to extend next, a evolving value for
-each node is given in Definition~\ref{def:value}. The algorithm that performs
-node selection starts from the root of the search tree. From its current node,
-it choses among the children of the open goal the one with the highest value.
-If its value is higher than the widening policy, then the selected child
-becomes the current node, otherwise the final result of node selection is the
-current node. The node selection algorithm can be written as the following
-pseudo-code:
-
-\begin{lstlisting}[language=SMLSmall]
-CurrentNode = Root(Tree);
-while true do
-  if Children(CurrentNode) = emptyset then break;
-  (BestChild, BestValue) = CompareChildren (CurrentNode);
-  if WideningPolicy (ChosenNode) >= BestValue then break;
-  CurrentNode = BestChild
-end;
-return CurrentNode;
-\end{lstlisting}
-
-\begin{definition}\label{def:value}(Value)
-The value of the $i^th$ child $a_i$ of a parent $p$ is determined by:
-\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} * \mathit{Exploration}(a_i)\]
-
-The current evaluation $CurEvaluation(a_i)$ is the average evaluation of
-all descendants of $a_i$ including node extension
-failures:
-
-\[\mathit{CurEvaluation}(a_i) =
-  \sum_{a' \in Descendant(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
-  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]
-
-The number $\mathit{Failure}(a_i)$ is the amount of failures that occurred during node
-extension from descendants of $a_i$ .
-
-The exploration term is determined by the prior policy and the current policy.
-\[\mathit{Exploration}(a_i) = \frac{\mathit{PriorPolicy}(a_i)}{\mathit{CurPolicy}(a_i)}\]
-
-\[\mathit{CurPolicy}(a_i) = \frac{1 + \mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
-The policy is approximatively the  percentage of time a node was visited.The
-square root skews this probability to favor even more exploration of nodes with
-few visits.
-
-The coefficient $c_{exploration}$ is experimentally determined and adjusts the
-trade-off between exploitation and exploration.
-\end{definition}
-
-\subsection{Node extension}
-Let $a$ be the result of node selection.
-If $a$ is not an active part of the search then extending $a$ is useless and
-the algorithm reports a failure. In our experiments, filtering nodes that not
-active before node selection is
-not beneficial. It may be because it breaks some symmetry in the
-$\mathit{Value}$ function.
-Otherwise, it applies the best untested tactic $t$ on the open goal $g$ of this
-node according to the prediction order for $g$. If no such tactic exists or if
-$t$ is not productive, a failure is returned.
-Otherwise, a new node $c$ is created that contains the list of goals $l$
-produced by $t$ and is added to the children of $a$.
-
-\subsection{Backpropagation}
-During backpropagation we update the statisitics of all the nodes traversed or
-created during this MCTS step:
-\begin{itemize}
-\item Their number of visit is incremented.
-\item If node extension failed, their failure count is incremented.
-\item If node extension succeeded, they inherit the evaluation of the created
-child.
-\end{itemize}
-
-These changes updates the current evaluation and the current policy of the
-traversed nodes.
-After completing backpropagation, the process loops back to node
-selection.
-
-\subsection{Resolution}
-The search ends when the algorithm reaches one of these 3 conditions:
-\begin{itemize}
-\item It finds a proof (i.e. the root node is solved). In this case,
-the search returns a minimized and
-prettified proof script (see Section~\ref{sec:proofdisplay}).
-\item It saturates. There is no tactics to be applied to any open goal.
-This happens less than 10 times in the full-scale experiment.
-\item It exceeds a given time limit set by default to 60 seconds.
-\end{itemize}
-
-\subsection{ATP Integration}~\label{sec:atp}
-General-purpose proof automation mechanisms which combine proof translation to
-ATPs with machine learning (``hammers'') have become quite successful in
-enhancing the automation level in proof assistants~\cite{hammers4qed}.
-As external automated reasoning techniques sometimes outperform the combined
-power of tactics, we would like to combine the \tactictoe search with
-stronger general purpose provers such as the ones found in \holyhammer for
-\holfour~\cite{tgck-cpp15}.
-
-Such a prover already exists in \holfour and is called \metis. It is already
-recognized by the tactic selection mechanism in \tactictoe and can have its
-premises predicted dynamically when in an abstracted form. Nevertheless,
-we think that the performance of \tactictoe can be boosted by giving the ATP
-\metis a special status. This arrangement consists in always predicting
-premises for \metis, giving it a slightly higher timeout and
-always trying it first on each node. These modifications only induce an
-overhead linear on the number nodes as a general purpose prover does not create
-goal obligations.
-
-The integration of an external prover is also possible and
-\eprover~\cite{eprover} is the one we experimented with .
-To this end, we created a tactic which expects a number of premises as argument
-and calls \eprover on the translated first order problem. We then extract the
-premises that were required in the \eprover
-proof and reconstruct it inside \holfour with \metis. Giving a special status
-to external
-prover is essential as external provers calls do not appear in human
-proof scripts. The setting in place is about the same as for \metis, the
-difference being an even higher timeout (5 seconds) and a larger number of
-predicted premises (128). Since calls to \eprover are
-computationally expensive, they are run in parallel and asynchronously. This
-avoids slowing down \tactictoe's search loop. The number of asynchronous calls
-to an external provers that can be executed at the same time is limited by the
-number of cores available to the user.
 
@@
 notice
-duplicated proofs. Our predictions comes from a much more detailed description
-of the open goal.
-However, we simply create a single label for each tactic call whereas each of
-its
-arguments is treated independently in \textsf{ML4PG}.
-Our choice is motivated by the k-NN algorithm already used in
-\holyhammer for the selection of theorems.
+duplicated proofs. Comparatively, our predictions comes from a much more 
+detailed description of the targeted goal.
+%However, we simply create a single label for each tactic call whereas each of
+%its
+%arguments is treated independently in \textsf{ML4PG}.
+%Our choice is motivated by the k-NN algorithm already used in
+%\holyhammer for the a of theorems.
 %, which can be easly adapted to the selection of tactics but would be improved

### 2018-02-19 tactictoe_jv.tex
minor

@@
 \item The orthogonalization process that eliminates redundant tactics is 
-formally described.
-\item A tactic abstraction method is introduced. It enables us to create
+described in a precise manner.
+\item A tactic abstraction mechanism is introduced. It enables us to create
 more general and flexible tactics where tactic arguments are dynamically
@@
 by asynchronous calls to \eprover to help \tactictoe during proof search.
-\item Evaluation of \tactictoe success rate relative to
-the length of the original proofs.
-\item User interaction is enhanced by the minimization and prettification
-applied on the generated proof.
+\item First evaluation of \tactictoe success rate relative to
+the length of the original human proofs.
+\item Minimization and prettification of the generated proof improves user 
+interaction.
 \end{itemize}

### 2018-02-19 tactictoe_jv.tex
minor

@@
 \sml function.
-\item Monte Carlo tree search~~\cite{montecarlo} (MCTS) replaces A* as our
+\item Monte Carlo tree search~\cite{montecarlo} (MCTS) replaces A* as our
 search algorithm. The MCTS algorithm gives a more balanced feedback on
@@
 can choose which proof tree to extend next and in which direction.
-Our search algorithm is a Monte Carlo tree search~\cite{} algorithm that relies
+Our search algorithm is a Monte Carlo tree search~\cite{montecarlo} algorithm 
+that relies
 on a prior evaluation function and a prior policy function. Those priors are

### 2018-02-19 tactictoe_jv.tex
minor

@@
 \tableofcontents
-\todoi{Comparison with hammers in usability}
+\todoi{Comparison with hammers in usability, easier reconstruction nice proofs}
 
@@
   tactics.  The system is evaluated on a large set of theorems originating
-  from \holfour (Section~\ref{s:experiments}), and we show that the performance
-  of the single best \tactictoe strategy exceeds the performance of a hammer
-  system used with a single strategy and a single efficient external prover.
+  from \holfour (Section~\ref{s:experiments}).
 

### 2018-02-19 tactictoe_jv.tex
minor

@@
   proof paths (Section~\ref{sec:proofsearch}). The result, if
-  successful, is a certified human-level proof composed of \holfour
+  successful, is a verified human-level proof composed of \holfour
   tactics.  The system is evaluated on a large set of theorems originating
@@
 
-The application of tactic to a goal produces a list of goals. Each of the 
-created goals can in turn be the input other tactics. Moreover, it is possible 
-to try multiple tactics on the same goal. In order to keep track of the 
-progress made so far from the initial goal we want to solve, we organize
-goals and tactics into a graph with list of goals as nodes and tactics as edges
-(See Definition~\ref{def:stree}).
+The application of tactic to an initial conjecture produces a list of goals. 
+Each of the created goals can in turn be the input other tactics. Moreover, it 
+is possible to try multiple tactics on the same goal. In order to keep track of 
+the progress made so far from the initial conjecture we want to solve, we 
+organize goals and tactics into a graph with list of goals as nodes and tactics 
+as edges (See Definition~\ref{def:stree}).
 
@@
 The exploration of the tree is decided by MCTS algorithm in 
-Section~\ref{sec:proofsearch}, which is guided by a tactic prediction algorithm 
-presented in Section~\ref{sec:features}.
+Section~\ref{sec:proofsearch}, which is guided by a prediction algorithms 
+presented in Section~\ref{sec:predictions}.
 

### 2018-02-19 tactictoe_jv.tex
Minor

--- 2018-02-19 tactictoe_jv.tex
@@
 
-\begin{definition}\label{def:prof} (Productive tactic)
+\begin{definition}\label{def:prod} (Productive tactic)
 The application of a tactic on an open goal is called productive if and only if

### 2018-02-19 tactictoe_jv.tex
minor

@@
 interrupt tactics after a short amount of time (0.05 seconds).
-\item It does not loop, i.e its output does not contain any goals that appears 
+\item It does not loop, i.e. its output does not contain any goals that appears 
 in the ancestor of the node of the current goal.
-\item It is not a parallel step, i.e its output is not a superset
+\item It is not a parallel step, i.e. its output is not a superset
 of a list of goals produced from the same goal.

### 2018-02-19 tactictoe_jv.tex
example

--- 2018-02-19 tactictoe_jv.tex
@@
 %\title{TacticToe: Proof  by Tactic Selection}
-\title{TacticToe: Learning the Game of Tactics}
+\title{TacticToe: Learning to Prove with Tactics}
 %\title{TacticToe: Automatic tactic-level proof generation}
@@
 
-\todo{JU review: Still feels more like an intro than an abstract. Could be
+\todoi{JU review: Still feels more like an intro than an abstract. Could be
 shorter and punchier.}
+
+\todoi{Move part of the abstract to the introduction.}
 \begin{abstract}
@@
 \tableofcontents
-\todoi{Comparison with hammers in usability, easier reconstruction nice proofs}
-
 \section{Introduction}
-
 Many of the state-of-the-art interactive theorem provers (ITPs) such as
@@
 
-\todoi{Rewrite the plan}
   To do this, we \emph{extract and record} tactic invocations from the ITP
@@
 \paragraph{Theories and script files.}
-Formal developments in \holfour are organized into named theories, which are concretely implemented by \emph{script} files defining modules in the functional programming language \sml.
-Each script file corresponds to a single theory, and contains definitions of types and constants as well as statements of theorems together with their proofs.
-In practice, tactic-based proofs are written in an informally specified \emph{embedded domain-specific language}: the language consisting of pre-defined tactics and \emph{tacticals} (functions that operate on tactics).
-However, the full power of \sml is always available for users to define their own tactics or combinations thereof.
-%To automate proof search strategies a user can implement programming language functions that correspond to more involved proofs.
-%Functions that correspond to backward proofs, called tactics, are combined with each other with the help of tacticals (functions that takes
+Formal developments in \holfour are organized into named theories, which are 
+concretely implemented by \emph{script} files defining modules in the 
+functional programming language \sml.
+Each script file corresponds to a single theory, and contains definitions of 
+types and constants as well as statements of theorems together with their 
+proofs.
+In practice, tactic-based proofs are written in an informally specified 
+\emph{embedded domain-specific language}: the language consisting of 
+pre-defined tactics and \emph{tacticals} (functions that operate on tactics).
+However, the full power of \sml is always available for users to define their 
+own tactics or combinations thereof.
+%To automate proof search strategies a user can implement programming language 
+functions that correspond to more involved proofs.
+%Functions that correspond to backward proofs, called tactics, are combined 
+with each other with the help of tacticals (functions that takes
 %tactics as arguments).
-%Such a combination of tactics will be called a \emph{proof script} when it proves a theorem stated in a theory.
+%Such a combination of tactics will be called a \emph{proof script} when it 
+proves a theorem stated in a theory.
 
@@
 Considering concrete implementation details now, a \emph{tactic} is a function
-that takes a goal (or proof obligation) and returns a list of goals (subgoals of the original goal) together
+that takes a goal (or proof obligation) and returns a list of goals (subgoals 
+of the original goal) together
 with a validation function.
-Calling the validation function on a list of \emph{theorems}, corresponding to the list of subgoals, results in a theorem corresponding to the original goal.
-For example, if the list of subgoals is empty, then calling the validation function on the empty list yields the originally goal sequent as a theorem.
+Calling the validation function on a list of \emph{theorems}, corresponding to 
+the list of subgoals, results in a theorem corresponding to the original goal.
+For example, if the list of subgoals is empty, then calling the validation 
+function on the empty list yields the originally goal sequent as a theorem.
 In this way, tactics implement the plumbing of backwards, goal-directed, proof 
@@
 Because validation functions are only important to check the final proof, we 
-omit them during the description of proof search algorithm. And we note $t(g)$ 
-the lists of goals produced by the application of the tactic $t$ to the goal 
+omit them during the description of proof search algorithm. We note $t(g)$ 
+the lists of goals produced by the application of a tactic $t$ to a goal 
 $g$.
@@
 This paper extends the work described in \cite{tgckju-lpar17}.
-There we proposed the idea of ... and achieved ...
+There we proposed the idea of a tactic-based prover based on supervised 
+learning guidance. And we achieved a 39\% success rate on theorems of the 
+\holfour standard library by running \tactictoe with a 5 seconds timeout.
 The additional contributions and their effect on our system are:
@@
 by asynchronous calls to \eprover to help \tactictoe during proof search.
+\item Evaluation of \tactictoe with a 60 seconds timeout achieving a 66\% 
+success rate on the standard library.
 \item First evaluation of \tactictoe success rate relative to
@@
 
-\begin{definition}\label{def:prod} (Productive tactic)
+\begin{definition}\label{def:prof} (Productive tactic)
 The application of a tactic on an open goal is called productive if and only if
@@
 \begin{itemize}
-\item Equality between two functions is easy to do from
-their string representation. The direct consequence is that we can avoid
+\item Equality between two functions is easy to compute from
+their string representation and therefore we can avoid
 predicting the same tactic over and over again.
@@
 using the definition of
-$LUPDATE$. Here, \tactictoe's proof is arguably smaller, faster and easier to
+\texttt{LUPDATE}. Here, \tactictoe's proof is arguably smaller, faster and 
+easier to
 understand and maintain. Such proofs after an expert review could replace the
@@
 
-In Example~\ref{ex:cs4}, we demonstrate the capabilities of \tactictoe on a goal that was not already proved in the \holfour libraries: that the set of numbers $\{0,...,n+m-1\} \setminus \{0,...,n-1\}$ is the same as the set obtained by adding $n$ to everything in $\{0,...,m-1\}$.
+In Example~\ref{ex:cs4}, we demonstrate the capabilities of \tactictoe on a 
+goal that was not already proved in the \holfour libraries: that the set of 
+numbers $\{0,...,n+m-1\} \setminus \{0,...,n-1\}$ is the same as the set 
+obtained by adding $n$ to everything in $\{0,...,m-1\}$. In this example, 
+\tactictoe uses the simplification set \texttt{ARITH\_ss} to reduce arithmetic 
+formulas. This exemplifies an another advantage that \tactictoe has over ATPs, 
+mainly its ability to take advantage of user-defined simplification sets.
 
@@
 \end{example}
-
-\todoi{TG to RK: Feel free to add more examples.}
 
@@
 %arguments is more difficult but is likely to produce more suitable lemmas.
-
-Proof patching: \cite{RingerYLG18}
+%Proof patching: \cite{RingerYLG18}
 
@@
 \section{Conclusion}\label{sec:concl}
-
 We proposed a new proof assistant automation technique which combines
 tactic-based proof search with machine learning prediction
-Its implementation, \tactictoe, achieves an overall performance of 66.4\%
-theorems on 7164 theorems of the \holfour standard library surpassing \eprover
+Its implementation, \tactictoe, achieves an overall success rate of 66.4\%
+on 7164 theorems of the \holfour standard library, surpassing \eprover
 with auto-schedule. Its

### 2018-02-19 tactictoe_jv.tex
Minor

--- 2018-02-19 tactictoe_jv.tex
@@
 The learning-based selection of relevant lemmas significantly improves the
-automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we use the
+automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe we use the
 distance-weighted \emph{$k$ nearest-neighbour} classifier~\cite{DudaniS76}
@@
 \section{Abstraction}\label{sec:synthesis}
-One of the great weakness of the previous version of \tactictoe is that
+One of the major weakness of the previous version of \tactictoe is that
 it could not create its own tactics. Indeed, sometimes no previous tactic is

### 2018-02-19 tactictoe_jv.tex
More

@@
 \section{Abstraction}\label{sec:synthesis}
-One of the major weakness of the previous version of \tactictoe is that
+One of the major weakness of the previous version of \tactictoe was that
 it could not create its own tactics. Indeed, sometimes no previous tactic is
@@
 %improve the success rate of \tactictoe.
-We present a way to create tactics with different arguments
+In this section we present a way to create tactics with different arguments
 by abstracting them and re-instantiating them using a predictor for that kind
@@
 ungeneralized one. The generalized tactics are typically slower that their 
-counterpart, but the added flexibility is worthwhile in practice. Moreover, 
+original variants, but the added flexibility is worthwhile in practice. Moreover, 
 since we impose a timeout on each tactic (0.05 seconds by default), very slow 
-tactics fails and thus are not selected.
+tactics fail and thus are not selected.
 

### 2018-02-19 tactictoe_jv.tex
more

@@
 Because validation functions are only important to check the final proof, we 
-omit them during the description of proof search algorithm. We note $t(g)$ 
+omit them during the description of proof search algorithm. We denote $t(g)$ 
 the lists of goals produced by the application of a tactic $t$ to a goal 
@@
 \item type constructors present in the types of constants and variables,
-\item subterms with all variables replaced by a single place holder $V$,
+\item subterms with all variables replaced by a single placeholder $V$,
 \item names of the variables.
@@
 The first step is to abstract arguments of tactics by replacing them by a
-place holder, creating a tactic with an unspecified argument.
+placeholder, creating a tactic with an unspecified argument.
 Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
 try to abstract any of the
-\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder $u$, we
+\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder in $u$, we
 can substitute $u$ by $h$ to create an abstracted tactic.
-We note this intermediate tactic $t[h/u]$. By repeating the
+We denote this intermediate tactic $t[h/u]$. By repeating the
 process, multiple

### 2018-02-19 tactictoe_jv.tex
Sec 8

@@
 instantiate the placeholders inside $\hat{t}$. Because it is difficult to
-build a general predictor that effective on each type of argument, we manually
+build a general predictor effective on each type of argument, we manually
 design different argument predictors for each type. Those predictors are
-given a goal $g$ as input and returns the best chosen argument for this
+given a goal $g$ as input and return the best predicted argument for this
 goal.
@@
 in $\hat{t}$. 
-List of theorems is a very common type in \holfour proofs and theorems contain 
+The type of theorem lists is a very common type in \holfour proofs and theorems contain 
 enough information to be predicted accurately.
@@
 A straightforward idea is to try $\hat{t}$ before $t$ during the proof search.
-However, this risks doing unnecessary work as they may have a similar effect.
+However, this risks doing unnecessary work as the two may perform similar steps.
 Therefore, we would like to decide
@@
 
-The Table~\ref{theories} compares the success rates of re-proving for different
+Table~\ref{theories} compares the re-proving success rates for different
 \holfour theories. \tactictoe outperforms \eprover on every
@@
 largely surpasses \eprover on theories relying on inductive terms and 
-simplifications sets such as \texttt{arithmetic}, \texttt{list} 
+simplification sets such as \texttt{arithmetic}, \texttt{list} 
 and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to 
@@
 
-We first investigate how \tactictoe and \eprover scales as the search time 
+We first investigate how \tactictoe and \eprover scale as the search time 
 grows in Figure~\ref{fig:c1}.
-In 10 seconds, \tactictoe solves 90 percent of the problem it can 
+In 10 seconds, \tactictoe solves 90 percent of the problems it can 
 solve in 60 seconds. The analysis is a bit different for \eprover. Indeed,
@@
 longer time limit, but we advise to do so in parallel to avoid waiting time.
+\todo{its strange to ``advise'' in a paper?}
 
@@
 validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
-graph that most of the human proofs are short. The proofs found by tactictoe 
-follows about the same
+graph that most of the human proofs are short. The proofs found by \tactictoe 
+follows a similar
 distribution. This means that if \tactictoe finds a proof, there is about a 50
@@
 difficulties. As expected, the longer the original proof is, the harder it is
-for \tactictoe and \eprover to re-prove the theorem by their own mean.
-The performance of \tactictoe is compelling with more than half of the theorem
+for \tactictoe and \eprover to re-prove the theorem on their own.
+The performance of \tactictoe is compelling with more than half of the theorems
 that required a proof of length six being reproven. It is also consistently

### 2018-02-19 tactictoe_jv.tex
More

@@
 in this section.
-<
+%
 Rather than relying on the underlying programming language, we parse the
@@
 code of each tactic.
-Working with the string representation of a tactic is better that working with
+Working with the string representation of a tactic is better than working with
 its value:
@@
 \item Equality between two functions is easy to compute from
-their string representation and therefore we can avoid
-predicting the same tactic over and over again.
+their string representation, which allows us to avoid
+predicting the same tactic repeatedly.
 \item
@@
 user can learn from the feedback, analyze the proof and possibly improve on it.
-An other point is that \tactictoe does not need to be installed to run
+Furthermore \tactictoe does not need to be installed to run
 proof scripts generated by \tactictoe. In this way, the further development of
-\tactictoe are not affecting the robustness of \holfour.
+\tactictoe does not affect the robustness of \holfour.
 \item
 It is difficult (probably impossible) to transfer \sml values
-between theories as they are not run in the same \holfour session. In contrast,
-\sml code can be exported, imported.
+between theories, since they are not run in the same \holfour session. In contrast,
+\sml code can be exported and imported.
 \item To produce a tactic value form a code is easy in \sml, which can be

### 2018-02-19 tactictoe_jv.tex
more

@@
 Even when developing one theory, the context is continually changing:
-modules are open and local identifiers are defined. Therefore, it is unlikely
+modules are opened and local identifiers are defined. Therefore, it is unlikely
 that code extracted from proof scripts is interpretable in any other part of
@@
 definition until we are able to write any expression with global identifiers
-only. And we prefix each global identifiers by its module.
-We call this process globalization. The result is a standalone \sml code
-executable in any \holfour theories.  In the absence of side effects, it is
+only. Similarly, we prefix each global identifier by its module.
+We call this process \emph{globalization}. The result is a standalone \sml code
+executable in any \holfour theory.  In the absence of side effects, it is
 interpreted in the same way across theories.
@@
 \subsection{Implementation}
-We describe in more details the implementation of the recording algorithm which
+We describe in more detail our implementation of the recording algorithm. It
 consists of 4 phases: proof script extraction, proof script globalization,
-tactic unit wrapping and creation of tactic-goal pairs.
+tactic unit wrapping, and creation of tactic-goal pairs.
 
@@
 The first line of this script file opens modules (called structures in \sml). Each of
-them contains a list of global identifiers which become directly accessible in
+the modules contains a list of global identifiers which become directly accessible in
 the rest of the script.
@@
 The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
-the goal is split into a base case and a inductive case. Finally, both of
-theses cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
+the goal is split into a base case and an inductive case. Finally, both
+of these cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
 rewrites assumptions with the help of theorems previously declared in this

### 2018-02-19 tactictoe_jv.tex
some typos

@@
 or even loop. For that reason, predictions need to be accompanied by a
-proof search mechanism that allows for backtrac king and
+proof search mechanism that allows for backtracking and
 can choose which proof tree to extend next and in which direction.
@@
 node selection starts from the root of the search tree. From the current node,
-the function \texttt{CompareChildren} choses among the children of its open 
+the function \texttt{CompareChildren} chooses among the children of its open 
 goal the one with the highest value.
@@
 \paragraph{Backpropagation}
-During backpropagation we update the statisitics of all the nodes traversed or
+During backpropagation we update the statistics of all the nodes traversed or
 created during this MCTS step:
@@
 definition which happens to contain two global identifiers.
-The previous paragraph describe the globalization for all identifiers except
+The previous paragraph describes the globalization for all identifiers except
 local theorems.  We do not replace a local

### 2018-02-19 tactictoe_jv.tex
Tweaks and fix some English

@@
 \paragraph{Goals and theorems.}
-A goal is represented as a sequent. Each sequent is composed of
-a list of assumptions and a conclusion which are higher-order logic formulas.
-When a goal is proven, it becomes a theorem. And if the developer gives it a
-name then we refer to it as a top-level theorem.
+A goal is represented as a \emph{sequent}.
+A sequent is composed of a set of assumptions and a conclusion, all of which are higher-order logic formulas.
+When a goal is proven, it becomes a theorem.
+If the developer gives the theorem a name then we refer to it as a top-level theorem.
 
@@
 own tactics or combinations thereof.
-%To automate proof search strategies a user can implement programming language 
-functions that correspond to more involved proofs.
-%Functions that correspond to backward proofs, called tactics, are combined 
-with each other with the help of tacticals (functions that takes
+%To automate proof search strategies a user can implement programming language functions that correspond to more involved proofs.
+%Functions that correspond to backward proofs, called tactics, are combined with each other with the help of tacticals (functions that takes
 %tactics as arguments).
-%Such a combination of tactics will be called a \emph{proof script} when it 
-proves a theorem stated in a theory.
+%Such a combination of tactics will be called a \emph{proof script} when it proves a theorem stated in a theory.
 
@@
 Because validation functions are only important to check the final proof, we
-omit them during the description of proof search algorithm. We denote $t(g)$ 
-the lists of goals produced by the application of a tactic $t$ to a goal 
-$g$.
+omit them during the description of proof search algorithm. We denote by $t(g)$
+the list of goals produced by the application of a tactic $t$ to a goal $g$.
 
@@
 
-The application of tactic to an initial conjecture produces a list of goals. 
-Each of the created goals can in turn be the input other tactics. Moreover, it 
-is possible to try multiple tactics on the same goal. In order to keep track of 
-the progress made so far from the initial conjecture we want to solve, we 
-organize goals and tactics into a graph with list of goals as nodes and tactics 
-as edges (See Definition~\ref{def:stree}).
+The application of a tactic to an initial conjecture produces a list of goals.
+Each of the created goals can in turn be the input to other tactics.
+Moreover, it is possible to try multiple tactics on the same goal.
+In order to keep track of the progress made so far from the initial conjecture we want to solve, we organize goals and tactics into a graph with list of goals as nodes and tactics as edges (See Definition~\ref{def:stree}).
 
@@
 moment of the search after a certain amount of tactic applications.
-The exploration of the tree is decided by MCTS algorithm in 
-Section~\ref{sec:proofsearch}, which is guided by a prediction algorithms 
+Exploration of the tree is directed by the MCTS algorithm in
+Section~\ref{sec:proofsearch}, which is guided by prediction algorithms
 presented in Section~\ref{sec:predictions}.
@@
 node selection starts from the root of the search tree. From the current node,
-the function \texttt{CompareChildren} chooses among the children of its open 
+the function \texttt{CompareChildren} choses among the children of its open
 goal the one with the highest value.

### 2018-02-19 tactictoe_jv.tex
Finish sec 9

--- 2018-02-19 tactictoe_jv.tex
@@
 effect as the original. But since we want to extract information from this
-script in the form tactics and their input goals, we modify it further.
+script in the form of tactics and their input goals, we modify it further.
 In particular, we need to define at which level we should record the tactics in
@@
 advice for something close to their common input goal. The tactic
-\texttt{A THEN B THEN C} is likely to be specific. In contrast, we can consider
+\texttt{A THEN B THEN C} is likely to be too specific. In contrast, we can consider
 the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A}
 until
-\texttt{A} has no effect. In this situation, it is often preferable to record
-only \texttt{REPEAT A} than calls to \texttt{A}. Otherwise during the
-proof search, we would have to make a branches created after each call of
-\texttt{A}.
+\texttt{A} has no effect. For such calls, it is often preferable to record
+a single call to \texttt{REPEAT A} rather than multiple calls to \texttt{A}.
+Otherwise branches after each call to \texttt{A} would be necessary as part
+of proof search.
 To sum up, we would like to record only the most general tactics which
@@
 \begin{definition}(Tactic unit)
-A tactic unit is an \sml expression of type \texttt{tactic} that does not contain a
+A tactic unit is an \sml expression of type \texttt{tactic} that does not contain an
 infix operator at its root.
@@
 
-Because these tactic units are constructed from visual information present in
-the tactic proof, they often represent what a human consider to be one step of
+Because such tactic units are constructed from visual information present in
+the tactic proof, they often represent what a human user considers to be a single step of
 the proof.
@@
 record
-tacitcs in all \holfour theories, we replace the original proof by the
+tactics in all \holfour theories, we replace the original proof by the
 recording script in each theory and rebuild the \holfour library.
@@
 then applied to the input goal to replicate the original behavior.
-After all modified theories are rebuild, each call to a wrapped tactic in a
+After all modified theories are rebuilt, each call to a wrapped tactic in a
 tactic proof is recorded as a pair containing its globalized code and its

### 2018-02-19 tactictoe_jv.tex
minor
Sec 11

@@
 
-\paragraph{Minimizing length of the proof}
+\paragraph{Proof Length Minimization}
 A fast and simple minimization is applied when processing the final proof. If
@@
 
-\paragraph{Minimizing tactic arguments}
+\paragraph{Tactic Arguments Minimization}
 Let $t$ be a tactic applied to a goal $g$ containing a list $l$ (of theorems)
@@
 any context. Since we return the proof at a point where the \sml interpreter is
-a specific state, we can strip unnecessary information in the form of
+in a specific state, we can strip unnecessary information, such as
 module prefixes. If possible, we group all local declarations in the
-script under a single \texttt{let} binding at the start of the script. And we
-replace extended terms by their quoted version.
+script under a single \texttt{let} binding at the start of the script. We
+also replace extended terms by their quoted version.
 All in all, if a prettified tactic $t_p$ has the same effect its original
@@
 \section{Case Study}\todo{comment on the case study}
-Since the proofs generated by \tactictoe are meant to be part of a \holfour development, it interesting to compare their quality with the original human proofs.
+Since the proofs generated by \tactictoe are meant to be parts of a \holfour development, it interesting to compare their quality with the original human proofs.
 The quality of a proof in \holfour can be measured in terms of length, readability, maintainability and verification speed.
 We study these properties in three examples taken from our full-scale experiment.
-The time needed by \holfour to check a proof is shown in parentheses.
-
-Example~\ref{ex:cs1} is about proving that greatest common divisors are unique.
-The human developer recognizes that it follows easily from two theorems.
-Our prover \tactictoe is not able to find that \texttt{DIVIDES\_ANTISYM} is
-necessary directly from the shape of the original theorems so it rewrites the
-goal then splits the goal into multiple cases until the goal obligations are similar
-to \texttt{DIVIDES\_ANTISYM} to be predicted. As expected, \tactictoe's proof
+We list the time needed by \holfour to check a proof in parentheses.
+
+We start with Example~\ref{ex:cs1}, which proves that greatest common divisors are unique.
+The human formalizer recognized that it follows from two theorems.
+The relevance filtering of \tactictoe is not as powerful as that used in hammers
+and is therefore not able to find the \texttt{DIVIDES\_ANTISYM} property. The proof
+proceeds instead by rewriting the
+goal with the definitions then spliting the goal into multiple cases until the goal obligations are similar
+to \texttt{DIVIDES\_ANTISYM}. As expected, \tactictoe's proof
 also takes much longer to check.
@@
 element of the concatenation of $l_1$ and $l_2$ is the same as
-updating the $m^{th}$ element of $l_2$ where $m = n - p$. Again,
+updating the $m^{th}$ element of $l_2$ where $m = n - p$. Again, the
 \tactictoe's proof is
-much more readable starting with applying induction on $l$. It solves the base
-case of the induction by rewriting and distinguish cases where $n=0$ or $n>0$
-in the inductive hypothesis. It finalizes the proof with a short call to \metis
+much more readable. It starts by applying induction on $l$. It solves the base
+case of the induction by rewriting and proceeds by cases on $n=0$ or $n>0$
+in the induction hypothesis. It finalizes the proof with a short call to \metis
 using the definition of
@@
 
-In Example~\ref{ex:cs4}, we demonstrate the capabilities of \tactictoe on a
-goal that was not already proved in the \holfour libraries: that the set of
+In Example~\ref{ex:cs4}, we experiment with \tactictoe on a
+goal that does not originate from the \holfour library. The conjecture is
+that the set of
 numbers $\{0,...,n+m-1\} \setminus \{0,...,n-1\}$ is the same as the set
@@
 \tactictoe uses the simplification set \texttt{ARITH\_ss} to reduce arithmetic
-formulas. This exemplifies an another advantage that \tactictoe has over ATPs,
-mainly its ability to take advantage of user-defined simplification sets.
+formulas. This exemplifies another advantage that \tactictoe has over ATPs,
+namely its ability to take advantage of user-defined simplification sets.
 

### 2018-02-19 tactictoe_jv.tex
Related

@@
 
-The system
 \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
-groups related proofs thanks to its clustering
+groups related proofs using clustering
 algorithms. It allows \coq users to inspire themselves from similar proofs and
@@
 duplicated proofs. Comparatively, our predictions comes from a much more
-detailed description of the targeted goal.
+detailed description of the targeted goal.\todo{and can be used as an ATP}
 %However, we simply create a single label for each tactic call whereas each of
@@
 
-\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to
+\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} can
 generate
 proof scripts from previous \coq proof examples.
-Its strength lies in its ability to produce likely sequences
+Its strength lies in the ability to produce likely sequences
 of tactics for solving domain specific goals. It operates by creating a model
 for common sequences of tactics for a specific library.
-This means that in order to propose the following tactic, only the previously
+This means that in order to propose the next tactic, only the previously
 called tactics
@@
 
+\todoi{maybe the paragraph can be generally proof strategy languages including ltac and eisbach}
 A similar effort for stronger automation is performed in \isabelle. Nagashima
@@
 This can be done either in an interactive way, when the user completes the
-proof based on the recommended lemmas, as in the \textsc{Mizar Proof
-Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma
+proof based on the recommended lemmas, as in the Mizar proof advisor
+\textsc{MizAR}~\cite{Urb04-MPTP0,mizAR40}, or attempted fully automatically, where such lemma
 selection is handed over to the ATP component of a \emph{hammer}
-system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.
+system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16}.
 
@@
 We proposed a new proof assistant automation technique which combines
-tactic-based proof search with machine learning prediction
+tactic-based proof search with machine learning prediction.
 Its implementation, \tactictoe, achieves an overall success rate of 66.4\%
@@
 Thanks to the learning abilities of \tactictoe, the generated proof scripts
-usually reveal the high-level structure of the proof. %For these reasons,
+often reveal the high-level structure of the proof. %For these reasons,
 We therefore believe that predicting ITP tactics based on the current goal

### 2018-02-19 tactictoe_jv.tex
Concl

@@
 new tactics could be created by programming them with any construction
-available in \sml. There is also something to be gain by
-having a better guidance for the search.
-That is why se also want to experiment with stronger machine learning algorithm
+available in \sml.
+The proof search guidance can also be improved, for example by considering
+stronger machine learning algorithms
 such as deep neural networks as a model for the policy and evaluation in MCTS.
-This quality of this model could be enhanced by additional data augmentation by
+The quality of such models could be enhanced by
 training and testing existing and synthesized tactics.
 Conjecturing suitable intermediate steps could also allow \tactictoe to solve
-a problem which requires a long proof by decomposing it into multiple easy
+problems which require long proofs by decomposing it into multiple easier
 steps.

### 2018-02-19 tactictoe_jv.tex
minor

@@
 steps it takes to solve a goal and a node.
-\begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal,
+\begin{definition}\label{def:solved}(solved goal, solved nodes, open goal,
 pending goal)\\
@@
 We call a goal (respectively node) unsolved if it is not in the set of solved
-goal (respectively node).
+goals (respectively nodes).
 The open goal of a node is the first unsolved goal of this node according to

### 2018-02-19 tactictoe_jv.tex
minor

@@
 
-In Definition~\ref{def:prod}, we judge if a tactic is productive from its
+In Definition~\ref{def:prod}, we decide if a tactic is productive based on its
 contribution to the search tree.
 
-\begin{definition}\label{def:prod} (Productive tactic)
+\begin{definition}\label{def:prod} (productive tactic)
 The application of a tactic on an open goal is called productive if and only if
@@
 interrupt tactics after a short amount of time (0.05 seconds).
-\item It does not loop, i.e. its output does not contain any goals that appears
+\item It does not loop, i.e., its output does not contain any goals that appear
 in the ancestor of the node of the current goal.
-\item It is not a parallel step, i.e. its output is not a superset
+\item It is not a parallel step, i.e., its output is not a superset
 of a list of goals produced from the same goal.
@@
 \begin{itemize}
-\item It finds a proof, i.e. the root node is solved. In this case,
+\item It finds a proof, i.e., the root node is solved. In this case,
 the search returns a minimized and
 prettified proof script (see Section~\ref{sec:proofdisplay}).
-\item It saturates, i.e. there is no tactics to be applied to any open goal.
+\item It saturates, i.e., there is no tactics to be applied to any open goal.
 This happens less than 10 times in the full-scale experiment.
@@
 The first idea is to select tactic-goal pairs and theorems by their similarity
-with $c$ (i.e $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
+with $c$ (i.e. $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
 $\mathit{Predictions}^{\mathit{theorem}}_{500} (c)$).

### 2018-02-19 tactictoe_jv.tex
More sec 3

@@
 
-We introduce specificities associated with the prediction of each object.
+We discuss below the specifics associated with the prediction of each kind of object.
 In particular, we present the dataset from which objects are selected and the
-purpose of the selection process. Details on the similarity measure backing the
+purpose of the selection process. The similarity measure backing the
 predictions is described in Section~\ref{sec:features} and two methods for
@@
 of tactic-goal pairs recorded from successful tactic applications in human
-proofs (see
+proofs (recording will be discussed in 
 Section~\ref{sec:recording}).
-Then, during proof search, we order recorded goals according to their
-similarity with a targeted goal $g$ (usually the first pending goal of a node).
+During proof search, the recorded goals will be reordered according to their
+similarity with a target goal $g$ (usually the first pending goal of a node).
 The recorded pairs induce an
@@
 List of goals that are a subset of the result of a recorded tactic on a
-targeted goal are considered positive. During proof search, given a list of
+target goal are considered positive. During proof search, given a list of
 goals $l$ created by a tactic, we select a set of lists of goals that are most
@@
 duplicated proofs. Comparatively, our predictions comes from a much more
-detailed description of the targeted goal.\todo{and can be used as an ATP}
+detailed description of the target goal.\todo{and can be used as an ATP}
 %However, we simply create a single label for each tactic call whereas each of

### 2018-02-19 tactictoe_jv.tex
more

@@
 between exploitation and exploration. This means that the algorithm searches
-deeper more promises branches and leaves enough time for the exploration
+deeper more promising branches and leaves enough time for the exploration
 of less likely alternatives.
@@
 want to prove.
+\todo{@TG: Check sentence below}
 Therefore the search tree starts with only one node the list of goals
@@
 Through node extension steps the search tree grows and the number of paths to
-explore increases. To decide which node to extend next, a evolving value for
+explore increases. To decide which node to extend next, an evolving value for
 each node is given in Definition~\ref{def:value}. The algorithm that performs
@@
 \begin{definition}\label{def:value}(Value)
-The value of the $i^th$ child $a_i$ of the open goal of a parent node $p$
+The value of the $i^{th}$ child $a_i$ of the open goal of a parent node $p$
 is determined by:
@@
 \mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
-The policy is approximatively the  percentage of time a node was visited.The
+The policy is approximately the  percentage of time a node was visited. The
 square root skews this probability to favor even more exploration of nodes with
@@
 %exceeds its timeout or that the human proof is too deep to be rediscovered
-%(more than 10 steps long).In all those experiments orthogonalization is
+%(more than 10 steps long). In all those experiments orthogonalization is
 %performed with

### 2018-02-19 tactictoe_jv.tex
4.1 and 4.2

@@
 \begin{itemize}
-\item Their number of visit is incremented.
+\item Their visit number is incremented.
 \item If node extension failed, their failure count is incremented.
@@
 
-These changes updates the current evaluation and the current policy of the
+These changes update the current evaluation and the current policy of the
 traversed nodes.
@@
 prettified proof script (see Section~\ref{sec:proofdisplay}).
-\item It saturates, i.e., there is no tactics to be applied to any open goal.
-This happens less than 10 times in the full-scale experiment.
-\item It exceeds a given time limit set by default to 60 seconds.
+\item It saturates, i.e., there are no tactics to be applied to any open goal.
+This happens less than 10 times in the full-scale experiment. \todo{@TG: why don't we have the original tactics as options?}
+\item It exceeds a given time limit (default is set to 60 seconds).
 \end{itemize}
@@
 \subsection{Supervised learning guidance}
-Our MCTS algorithm relies on the guidance of two supervised learning methods,
-the prior policy decides the order in which tactic should be tried and the
+Our MCTS algorithm relies on the guidance of two supervised learning methods.
+The prior policy decides the order in which tactic should be tried and the
 prior evaluation estimates if a branch is fruitful by analysing tactic
@@
 $t_i$ is to prove the goal.
-We can now calculate of a prior policy of a child $a_i$ produced by the tactic
+We can now calculate the prior policy of a child $a_i$ produced by the tactic
 $t_i$ by:
@@
 
-We now concentrate on the definition a reasonable evaluation function for the
-output of tactics. Theoretically, an interesting list of goals is one
-that has short proofs for each of its goals, and we could estimate their
-lengths from previous proof searches. However, this
+We now concentrate on the definition of a reasonable evaluation function for the
+output of tactics. Theoretically, a list of goals is interesting if short proofs
+are possible each the goals in the list, and the lengths of the proofs can be
+estimated from previous proof searches. However, this
 approach generates too much data and slows down the proof search. So, we only
-collect output of tactics tested during orthogonalization. We
+collect outputs of tactics tested during orthogonalization. We
 declare a list of goals $l$ to be positive if it is has been produced by the
@@
 
-\subsection{ATP Integration}~\label{sec:atp}
+\subsection{ATP Integration}\label{sec:atp}
 General-purpose proof automation mechanisms which combine proof translation to
@@
 
-\subsection{Tuning \tactictoe}~\label{sec:tuning}
+\subsection{Tuning \tactictoe}\label{sec:tuning}
 
@@
 
-\subsection{Full-scale experiment}~\label{sec:full_exp}
+\subsection{Full-scale experiment}\label{sec:full_exp}
 

### 2018-02-19 tactictoe_jv.tex
Finish sec 4

@@
 enhancing the automation level in proof assistants~\cite{hammers4qed}.
-As external automated reasoning techniques sometimes outperform the combined
+Since external automated reasoning techniques sometimes outperform the combined
 power of tactics, we would like to combine the \tactictoe search with
@@
 
-Such a prover already exists in \holfour and is called \metis. It is already
-recognized by the tactic selection mechanism in \tactictoe and can have its
-premises predicted dynamically when in an abstracted form. Nevertheless,
+\holfour already integrates a superposition-based prover \metis for this purpose. It is already
+recognized by the tactic selection mechanism in \tactictoe and its
+premises can be predicted dynamically when in an abstracted form. Nevertheless,
 we think that the performance of \tactictoe can be boosted by giving the ATP
@@
 trying it first on each open goal. These modifications only induce an
-linear overhead on the number nodes as a general purpose prover does not create
-goal obligations.
-
-The integration of an external prover is also possible. And
-\eprover~\cite{eprover} is the one we experimented with.
-To this end, we created a tactic which expects a number of premises as an
+linear overhead on the number nodes as \metis does not create
+new goals.
+
+The integration of an external prover is also possible. For this purpose we experiment
+with 
+\eprover~\cite{eprover}.
+Namely, we developed a tactic which expects a list of premises as an
 argument
-and calls \eprover on the translated first order problem. We then extract the
-premises that were required in the \eprover
-proof and reconstruct it inside \holfour with \metis. Giving a special status
+and calls \eprover on the goal together with the premises translated to a first-order problem. If \eprover
+succeeds on the problem, the premises used in the external proof are used to
+reconstruct the proof inside \holfour with \metis. Giving a special status
 to external
@@
 selection is handed over to the ATP component of a \emph{hammer}
-system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16}.
+system, as in the hammers for \holfour~\cite{tgck-cpp15}, \hollight~\cite{holyhammer} and \isabelle~\cite{BlanchetteGKKU16}.
 

### 2018-02-19 tactictoe_jv.tex
Sec 5

@@
 \section{Feature Extraction}\label{sec:features}
+\todo{@TG: decide which of the next sentences}
 We explain the construction of the similarity measure by which relevant
@@
 
-Here are the different kind of feature we extract from terms:
+Here are the different kinds of feature we extract from terms:
 \begin{itemize}
@@
 \end{itemize}
-Goals (respectively theorems) are represented by a couple of a list of
-terms (assumptions) and a term (conclusion). So, we distinguish between
+Goals (respectively theorems) are represented by pairs consisting of a list of
+terms (assumptions) and a term (conclusion). We distinguish between
 features of the assumptions and features of the conclusion by adding a
-different tag to elements of each set. The union of all those
-features are considered the feature of a goal (respectively theorem). From
-that, we can construct features for a list of goals by
-from the union of the features of each goal in the list.
+different tag to elements of each set. The feature of a goal (respectively theorem)
+are the union of all these tagged features. From
+that, we can construct features for a list of goals by computing the union of the features of each goal in the list.
 
@@
 can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
-We additionally modify the weight by raising them to the sixth power giving
-even more credence
+We additionally raise these weighs to the sixth power giving
+even more importance
 to rare features. This modification was found experimentally optimal in
@@
 
-The first co-distance $sim_1$ sums the weight of each shared
+The first co-distance $sim_1$ sums the weight of all shared
 features to compute the total score.
@@
 $\mathit{Predictions}^{\mathit{theorem}}_{500} (c)$).
-However, this selection may not be adapted to later stages of
+However, this selection may not be adapted at later stages of
 the proof where goals may have drifted significantly from $c$. In order to
@@
 takes dependencies between objects of the same dataset into account.
-This dependency relation is assymetric. Through the relation, each object has
+This dependency relation is asymmetric. Through the relation, each object has
 multiple children and at most one parent.
@@
 for each object, which is the maximum of its similarity score
-and the similarity score of its parent. In the end, the 500 entries with
+and the similarity score of its parent. Finally, the 500 entries with
 highest dependency score are preselected in each dataset.
@@
 Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
-The dependencies $D_\infty$ of a tactic-goal pair $(t_0,g_0)$ is
+The dependencies $D_\infty$ of a tactic-goal pair $(t_0,g_0)$ are
 inductively defined by:
@@
 input goal are chosen or discarded together.
-\todo{this paragraph was confusing for Lasse}
+\todo{this paragraph was confusing for Lasse and Cezary}
 

### 2018-02-19 tactictoe_jv.tex
Sec 6

@@
 The current policy is calculate from the number of times a node $x$ has been
-traversed during node selection, noted $\mathit{Visit}(x)$.
+traversed during node selection, denoted $\mathit{Visit}(x)$.
 
@@
 The set of positive examples in
-$\mathit{Predictions}^{\mathit{goal\_list}}_k(l)$ is noted
+$\mathit{Predictions}^{\mathit{goal\_list}}_k(l)$ is denoted
 $\mathit{Positive}^{\mathit{goal\_list}}_k(l)$.
@@
 is undesirable, as it leads to inefficiency in automated proof search.
-To solve this problem, we modify how the database of tactics is build.
+To solve this problem, we modify the construction of the tactics database.
 Each time a new tactic-goal pair $(t,g)$ is extracted from a tactic proof and
-about to be recorded, we consider if there does not exists already a better
+about to be recorded, we consider if there does not already exist a better
 tactic for $g$ in our database. To this end, we organize a competition between
@@
 \begin{definition} (Goal subsumption)\\
-A goal subsumption $\le$ is a partial relation on a set of goals.\\
+A goal subsumption $\le$ is a partial relation goals.
 Assuming we have a way to estimate the number steps required to solve a goal,
-a possibly useful subsumption definition could be:
+a possibly useful subsumption definition could be: \todo{shorter or equal}
 \[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ has a shorter proof than }
@@
 \end{definition}
+\todoi{@TG: Which of the two above?}
 
@@
 Given two non-failing tactics $t_1$ and $t_2$ on $g$, a tactic $t_1$ subsumes a
-tactic $t_2$ on a goal $g$, noted $\le_g$, when:
+tactic $t_2$ on a goal $g$, denoted $\le_g$, when:
 
@@
 
-In the end, the winning tactic of the orthogonalization competition can be
+Finally, the winning tactic of the orthogonalization competition can be
 expressed by the formula:
@@
 In Example~\ref{ex:cs3}, we prove a theorem about lists, a domain where \tactictoe excels compared to ATPs.
-Given two list $l_1$ and $l_2$ where the length of $l_1$ (noted $p$) is less
+Given two list $l_1$ and $l_2$ where the length of $l_1$ (denoted $p$) is less
 than a natural

### 2018-02-20 tactictoe_jv.tex
minor

--- 2018-02-19 tactictoe_jv.tex
@@
 \paragraph{Goals and theorems.}
-A goal is represented as a \emph{sequent}.
-A sequent is composed of a set of assumptions and a conclusion, all of which are higher-order logic formulas.
-When a goal is proven, it becomes a theorem.
-If the developer gives the theorem a name then we refer to it as a top-level theorem.
+A goal is represented as a sequent. Each sequent is composed of
+a list of assumptions and a conclusion which are higher-order logic formulas.
+When a goal is proven, it becomes a theorem. And if the developer gives it a
+name then we refer to it as a top-level theorem.
 
@@
 own tactics or combinations thereof.
-%To automate proof search strategies a user can implement programming language functions that correspond to more involved proofs.
-%Functions that correspond to backward proofs, called tactics, are combined with each other with the help of tacticals (functions that takes
+%To automate proof search strategies a user can implement programming language 
+functions that correspond to more involved proofs.
+%Functions that correspond to backward proofs, called tactics, are combined 
+with each other with the help of tacticals (functions that takes
 %tactics as arguments).
-%Such a combination of tactics will be called a \emph{proof script} when it proves a theorem stated in a theory.
+%Such a combination of tactics will be called a \emph{proof script} when it 
+proves a theorem stated in a theory.
 
@@
 Because validation functions are only important to check the final proof, we 
-omit them during the description of proof search algorithm. We denote by $t(g)$
-the list of goals produced by the application of a tactic $t$ to a goal $g$.
+omit them during the description of proof search algorithm. We note $t(g)$ 
+the lists of goals produced by the application of a tactic $t$ to a goal 
+$g$.
 
@@
 
-The application of a tactic to an initial conjecture produces a list of goals.
-Each of the created goals can in turn be the input to other tactics.
-Moreover, it is possible to try multiple tactics on the same goal.
-In order to keep track of the progress made so far from the initial conjecture we want to solve, we organize goals and tactics into a graph with list of goals as nodes and tactics as edges (See Definition~\ref{def:stree}).
+The application of tactic to an initial conjecture produces a list of goals. 
+Each of the created goals can in turn be the input other tactics. Moreover, it 
+is possible to try multiple tactics on the same goal. In order to keep track of 
+the progress made so far from the initial conjecture we want to solve, we 
+organize goals and tactics into a graph with list of goals as nodes and tactics 
+as edges (See Definition~\ref{def:stree}).
 
@@
 moment of the search after a certain amount of tactic applications. 
-Exploration of the tree is directed by the MCTS algorithm in
-Section~\ref{sec:proofsearch}, which is guided by prediction algorithms
+The exploration of the tree is decided by MCTS algorithm in 
+Section~\ref{sec:proofsearch}, which is guided by a prediction algorithms 
 presented in Section~\ref{sec:predictions}.
@@
 steps it takes to solve a goal and a node.
-\begin{definition}\label{def:solved}(solved goal, solved nodes, open goal,
+\begin{definition}\label{def:solved}(Solved goal, solved nodes, open goal,
 pending goal)\\
@@
 We call a goal (respectively node) unsolved if it is not in the set of solved  
-goals (respectively nodes).
+goal (respectively node).
 The open goal of a node is the first unsolved goal of this node according to 
@@
 
-In Definition~\ref{def:prod}, we decide if a tactic is productive based on its
+In Definition~\ref{def:prod}, we judge if a tactic is productive from its 
 contribution to the search tree.
 
-\begin{definition}\label{def:prod} (productive tactic)
+\begin{definition}\label{def:prod} (Productive tactic)
 The application of a tactic on an open goal is called productive if and only if
@@
 interrupt tactics after a short amount of time (0.05 seconds).
-\item It does not loop, i.e., its output does not contain any goals that appear
+\item It does not loop, i.e. its output does not contain any goals that appears 
 in the ancestor of the node of the current goal.
-\item It is not a parallel step, i.e., its output is not a superset
+\item It is not a parallel step, i.e. its output is not a superset
 of a list of goals produced from the same goal.
@@
 The learning-based selection of relevant lemmas significantly improves the
-automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe we use the
+automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe, we use the
 distance-weighted \emph{$k$ nearest-neighbour} classifier~\cite{DudaniS76}
@@
 
-We discuss below the specifics associated with the prediction of each kind of object.
+We introduce specificities associated with the prediction of each object.
 In particular, we present the dataset from which objects are selected and the
-purpose of the selection process. The similarity measure backing the
+purpose of the selection process. Details on the similarity measure backing the 
 predictions is described in Section~\ref{sec:features} and two methods for 
@@
 of tactic-goal pairs recorded from successful tactic applications in human
-proofs (recording will be discussed in 
+proofs (see
 Section~\ref{sec:recording}).
-During proof search, the recorded goals will be reordered according to their
-similarity with a target goal $g$ (usually the first pending goal of a node).
+Then, during proof search, we order recorded goals according to their
+similarity with a targeted goal $g$ (usually the first pending goal of a node).
 The recorded pairs induce an
@@
 List of goals that are a subset of the result of a recorded tactic on a
-target goal are considered positive. During proof search, given a list of
+targeted goal are considered positive. During proof search, given a list of
 goals $l$ created by a tactic, we select a set of lists of goals that are most
@@
 or even loop. For that reason, predictions need to be accompanied by a
-proof search mechanism that allows for backtracking and
+proof search mechanism that allows for backtrac king and
 can choose which proof tree to extend next and in which direction.
@@
 between exploitation and exploration. This means that the algorithm searches
-deeper more promising branches and leaves enough time for the exploration
+deeper more promises branches and leaves enough time for the exploration
 of less likely alternatives.
@@
 want to prove.
-\todo{@TG: Check sentence below}
 Therefore the search tree starts with only one node the list of goals
@@
 Through node extension steps the search tree grows and the number of paths to
-explore increases. To decide which node to extend next, an evolving value for
+explore increases. To decide which node to extend next, a evolving value for
 each node is given in Definition~\ref{def:value}. The algorithm that performs
@@
 \begin{definition}\label{def:value}(Value)
-The value of the $i^{th}$ child $a_i$ of the open goal of a parent node $p$
+The value of the $i^th$ child $a_i$ of the open goal of a parent node $p$ 
 is determined by:
@@
 \mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
-The policy is approximately the  percentage of time a node was visited. The
+The policy is approximatively the  percentage of time a node was visited.The
 square root skews this probability to favor even more exploration of nodes with
@@
 \paragraph{Backpropagation}
-During backpropagation we update the statistics of all the nodes traversed or
+During backpropagation we update the statisitics of all the nodes traversed or
 created during this MCTS step:
 \begin{itemize}
-\item Their visit number is incremented.
+\item Their number of visit is incremented.
 \item If node extension failed, their failure count is incremented.
@@
 
-These changes update the current evaluation and the current policy of the
+These changes updates the current evaluation and the current policy of the
 traversed nodes.
@@
 \begin{itemize}
-\item It finds a proof, i.e., the root node is solved. In this case,
+\item It finds a proof, i.e. the root node is solved. In this case,
 the search returns a minimized and
 prettified proof script (see Section~\ref{sec:proofdisplay}).
-\item It saturates, i.e., there are no tactics to be applied to any open goal.
-This happens less than 10 times in the full-scale experiment. \todo{@TG: why don't we have the original tactics as options?}
-\item It exceeds a given time limit (default is set to 60 seconds).
+\item It saturates, i.e. there is no tactics to be applied to any open goal.
+This happens less than 10 times in the full-scale experiment.
+\item It exceeds a given time limit set by default to 60 seconds.
 \end{itemize}
@@
 \subsection{Supervised learning guidance}
-Our MCTS algorithm relies on the guidance of two supervised learning methods.
-The prior policy decides the order in which tactic should be tried and the
+Our MCTS algorithm relies on the guidance of two supervised learning methods,
+the prior policy decides the order in which tactic should be tried and the 
 prior evaluation estimates if a branch is fruitful by analysing tactic 
@@
 $t_i$ is to prove the goal.
-We can now calculate the prior policy of a child $a_i$ produced by the tactic
+We can now calculate of a prior policy of a child $a_i$ produced by the tactic
 $t_i$ by:
@@
 
-We now concentrate on the definition of a reasonable evaluation function for the
-output of tactics. Theoretically, a list of goals is interesting if short proofs
-are possible each the goals in the list, and the lengths of the proofs can be
-estimated from previous proof searches. However, this
+We now concentrate on the definition a reasonable evaluation function for the
+output of tactics. Theoretically, an interesting list of goals is one
+that has short proofs for each of its goals, and we could estimate their 
+lengths from previous proof searches. However, this
 approach generates too much data and slows down the proof search. So, we only 
-collect outputs of tactics tested during orthogonalization. We
+collect output of tactics tested during orthogonalization. We
 declare a list of goals $l$ to be positive if it is has been produced by the
@@
 
-\subsection{ATP Integration}\label{sec:atp}
+\subsection{ATP Integration}~\label{sec:atp}
 General-purpose proof automation mechanisms which combine proof translation to
@@
 enhancing the automation level in proof assistants~\cite{hammers4qed}.
-Since external automated reasoning techniques sometimes outperform the combined
+As external automated reasoning techniques sometimes outperform the combined
 power of tactics, we would like to combine the \tactictoe search with
@@
 
-\holfour already integrates a superposition-based prover \metis for this purpose. It is already
-recognized by the tactic selection mechanism in \tactictoe and its
-premises can be predicted dynamically when in an abstracted form. Nevertheless,
+Such a prover already exists in \holfour and is called \metis. It is already
+recognized by the tactic selection mechanism in \tactictoe and can have its
+premises predicted dynamically when in an abstracted form. Nevertheless,
 we think that the performance of \tactictoe can be boosted by giving the ATP
@@
 trying it first on each open goal. These modifications only induce an
-linear overhead on the number nodes as \metis does not create
-new goals.
-
-The integration of an external prover is also possible. For this purpose we experiment
-with 
-\eprover~\cite{eprover}.
-Namely, we developed a tactic which expects a list of premises as an
+linear overhead on the number nodes as a general purpose prover does not create
+goal obligations.
+
+The integration of an external prover is also possible. And
+\eprover~\cite{eprover} is the one we experimented with.
+To this end, we created a tactic which expects a number of premises as an 
 argument
-and calls \eprover on the goal together with the premises translated to a first-order problem. If \eprover
-succeeds on the problem, the premises used in the external proof are used to
-reconstruct the proof inside \holfour with \metis. Giving a special status
+and calls \eprover on the translated first order problem. We then extract the
+premises that were required in the \eprover
+proof and reconstruct it inside \holfour with \metis. Giving a special status
 to external
@@
 \item type constructors present in the types of constants and variables,
-\item subterms with all variables replaced by a single placeholder $V$,
+\item subterms with all variables replaced by a single place holder $V$,
 \item names of the variables.
@@
 The first idea is to select tactic-goal pairs and theorems by their similarity 
-with $c$ (i.e. $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
+with $c$ (i.e $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
 $\mathit{Predictions}^{\mathit{theorem}}_{500} (c)$).
@@
 \section{Abstraction}\label{sec:synthesis}
-One of the major weakness of the previous version of \tactictoe was that
+One of the great weakness of the previous version of \tactictoe is that
 it could not create its own tactics. Indeed, sometimes no previous tactic is
@@
 %improve the success rate of \tactictoe.
-In this section we present a way to create tactics with different arguments
+We present a way to create tactics with different arguments
 by abstracting them and re-instantiating them using a predictor for that kind
@@
 ungeneralized one. The generalized tactics are typically slower that their 
-original variants, but the added flexibility is worthwhile in practice. Moreover,
+counterpart, but the added flexibility is worthwhile in practice. Moreover, 
 since we impose a timeout on each tactic (0.05 seconds by default), very slow 
-tactics fail and thus are not selected.
+tactics fails and thus are not selected.
 
@@
 The first step is to abstract arguments of tactics by replacing them by a
-placeholder, creating a tactic with an unspecified argument.
+place holder, creating a tactic with an unspecified argument.
 Suppose we have a recorded tactic $t$. Since $t$ is a \sml code tree, we can
 try to abstract any of the
-\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder in $u$, we
+\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder $u$, we
 can substitute $u$ by $h$ to create an abstracted tactic.
-We denote this intermediate tactic $t[h/u]$. By repeating the
+We note this intermediate tactic $t[h/u]$. By repeating the
 process, multiple
@@
 instantiate the placeholders inside $\hat{t}$. Because it is difficult to
-build a general predictor effective on each type of argument, we manually
+build a general predictor that effective on each type of argument, we manually
 design different argument predictors for each type. Those predictors are
-given a goal $g$ as input and return the best predicted argument for this
+given a goal $g$ as input and returns the best chosen argument for this
 goal.
@@
 in $\hat{t}$. 
-The type of theorem lists is a very common type in \holfour proofs and theorems contain
+List of theorems is a very common type in \holfour proofs and theorems contain 
 enough information to be predicted accurately.
@@
 A straightforward idea is to try $\hat{t}$ before $t$ during the proof search.
-However, this risks doing unnecessary work as the two may perform similar steps.
+However, this risks doing unnecessary work as they may have a similar effect.
 Therefore, we would like to decide
@@
 
-\subsection{Tuning \tactictoe}\label{sec:tuning}
+\subsection{Tuning \tactictoe}~\label{sec:tuning}
 
@@
 %exceeds its timeout or that the human proof is too deep to be rediscovered
-%(more than 10 steps long). In all those experiments orthogonalization is
+%(more than 10 steps long).In all those experiments orthogonalization is
 %performed with
@@
 
-\subsection{Full-scale experiment}\label{sec:full_exp}
+\subsection{Full-scale experiment}~\label{sec:full_exp}
 
@@
 
-Table~\ref{theories} compares the re-proving success rates for different
+The Table~\ref{theories} compares the success rates of re-proving for different
 \holfour theories. \tactictoe outperforms \eprover on every
@@
 largely surpasses \eprover on theories relying on inductive terms and 
-simplification sets such as \texttt{arithmetic}, \texttt{list}
+simplifications sets such as \texttt{arithmetic}, \texttt{list} 
 and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to 
@@
 
-We first investigate how \tactictoe and \eprover scale as the search time
+We first investigate how \tactictoe and \eprover scales as the search time 
 grows in Figure~\ref{fig:c1}.
-In 10 seconds, \tactictoe solves 90 percent of the problems it can
+In 10 seconds, \tactictoe solves 90 percent of the problem it can 
 solve in 60 seconds. The analysis is a bit different for \eprover. Indeed,
@@
 longer time limit, but we advise to do so in parallel to avoid waiting time.
-\todo{its strange to ``advise'' in a paper?}
 
@@
 validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
-graph that most of the human proofs are short. The proofs found by \tactictoe
-follows a similar
+graph that most of the human proofs are short. The proofs found by tactictoe 
+follows about the same
 distribution. This means that if \tactictoe finds a proof, there is about a 50
@@
 difficulties. As expected, the longer the original proof is, the harder it is
-for \tactictoe and \eprover to re-prove the theorem on their own.
-The performance of \tactictoe is compelling with more than half of the theorems
+for \tactictoe and \eprover to re-prove the theorem by their own mean.
+The performance of \tactictoe is compelling with more than half of the theorem
 that required a proof of length six being reproven. It is also consistently
@@
 in this section.
-%
+<
 Rather than relying on the underlying programming language, we parse the
@@
 code of each tactic.
-Working with the string representation of a tactic is better than working with
+Working with the string representation of a tactic is better that working with
 its value:
@@
 \item Equality between two functions is easy to compute from
-their string representation, which allows us to avoid
-predicting the same tactic repeatedly.
+their string representation and therefore we can avoid
+predicting the same tactic over and over again.
 \item
@@
 user can learn from the feedback, analyze the proof and possibly improve on it.
-Furthermore \tactictoe does not need to be installed to run
+An other point is that \tactictoe does not need to be installed to run
 proof scripts generated by \tactictoe. In this way, the further development of
-\tactictoe does not affect the robustness of \holfour.
+\tactictoe are not affecting the robustness of \holfour.
 \item
 It is difficult (probably impossible) to transfer \sml values
-between theories, since they are not run in the same \holfour session. In contrast,
-\sml code can be exported and imported.
+between theories as they are not run in the same \holfour session. In contrast,
+\sml code can be exported, imported.
 \item To produce a tactic value form a code is easy in \sml, which can be
@@
 Even when developing one theory, the context is continually changing:
-modules are opened and local identifiers are defined. Therefore, it is unlikely
+modules are open and local identifiers are defined. Therefore, it is unlikely
 that code extracted from proof scripts is interpretable in any other part of
@@
 definition until we are able to write any expression with global identifiers
-only. Similarly, we prefix each global identifier by its module.
-We call this process \emph{globalization}. The result is a standalone \sml code
-executable in any \holfour theory.  In the absence of side effects, it is
+only. And we prefix each global identifiers by its module.
+We call this process globalization. The result is a standalone \sml code
+executable in any \holfour theories.  In the absence of side effects, it is
 interpreted in the same way across theories.
@@
 \subsection{Implementation}
-We describe in more detail our implementation of the recording algorithm. It
+We describe in more details the implementation of the recording algorithm which
 consists of 4 phases: proof script extraction, proof script globalization,
-tactic unit wrapping, and creation of tactic-goal pairs.
+tactic unit wrapping and creation of tactic-goal pairs.
 
@@
 The first line of this script file opens modules (called structures in \sml). Each of
-the modules contains a list of global identifiers which become directly accessible in
+them contains a list of global identifiers which become directly accessible in
 the rest of the script.
@@
 The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
-the goal is split into a base case and an inductive case. Finally, both
-of these cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
+the goal is split into a base case and a inductive case. Finally, both of
+theses cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
 rewrites assumptions with the help of theorems previously declared in this
@@
 definition which happens to contain two global identifiers.
-The previous paragraph describes the globalization for all identifiers except
+The previous paragraph describe the globalization for all identifiers except
 local theorems.  We do not replace a local
@@
 effect as the original. But since we want to extract information from this
-script in the form of tactics and their input goals, we modify it further.
+script in the form tactics and their input goals, we modify it further.
 In particular, we need to define at which level we should record the tactics in
@@
 advice for something close to their common input goal. The tactic
-\texttt{A THEN B THEN C} is likely to be too specific. In contrast, we can consider
+\texttt{A THEN B THEN C} is likely to be specific. In contrast, we can consider
 the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A}
 until
-\texttt{A} has no effect. For such calls, it is often preferable to record
-a single call to \texttt{REPEAT A} rather than multiple calls to \texttt{A}.
-Otherwise branches after each call to \texttt{A} would be necessary as part
-of proof search.
+\texttt{A} has no effect. In this situation, it is often preferable to record
+only \texttt{REPEAT A} than calls to \texttt{A}. Otherwise during the
+proof search, we would have to make a branches created after each call of
+\texttt{A}.
 To sum up, we would like to record only the most general tactics which
@@
 \begin{definition}(Tactic unit)
-A tactic unit is an \sml expression of type \texttt{tactic} that does not contain an
+A tactic unit is an \sml expression of type \texttt{tactic} that does not contain a
 infix operator at its root.
@@
 
-Because such tactic units are constructed from visual information present in
-the tactic proof, they often represent what a human user considers to be a single step of
+Because these tactic units are constructed from visual information present in
+the tactic proof, they often represent what a human consider to be one step of
 the proof.
@@
 record
-tactics in all \holfour theories, we replace the original proof by the
+tacitcs in all \holfour theories, we replace the original proof by the
 recording script in each theory and rebuild the \holfour library.
@@
 then applied to the input goal to replicate the original behavior.
-After all modified theories are rebuilt, each call to a wrapped tactic in a
+After all modified theories are rebuild, each call to a wrapped tactic in a
 tactic proof is recorded as a pair containing its globalized code and its
@@
 
-When \tactictoe finds a proof of a goal $g_{root}$, it returns a search tree
-$\mathfrak{T}$ where the root node containing $g_{root}$ is solved.
+When \tactictoe finds a proof of an initial goal, it returns a search tree
+$\mathfrak{T}$ where the root node containing the initial goal is solved.
 In order to transform this search tree into a proof script, we need to extract
@@
 eliminating unnecessary tactics and theorems has been developed and
-improve the user experience greatly~\cite{DBLP:conf/sefm/Adams15}.
-
-\paragraph{Proof Length Minimization}
+improves the user experience greatly~\cite{DBLP:conf/sefm/Adams15}.
+
+\paragraph{Minimizing length of the proof}
 A fast and simple minimization is applied when processing the final proof. If
@@
 
-\paragraph{Tactic Arguments Minimization}
+\paragraph{Minimizing tactic arguments}
 Let $t$ be a tactic applied to a goal $g$ containing a list $l$ (of theorems)
@@
 any context. Since we return the proof at a point where the \sml interpreter is
-in a specific state, we can strip unnecessary information, such as
+a specific state, we can strip unnecessary information in the form of
 module prefixes. If possible, we group all local declarations in the
-script under a single \texttt{let} binding at the start of the script. We
-also replace extended terms by their quoted version.
+script under a single \texttt{let} binding at the start of the script. And we
+replace extended terms by their quoted version.
 All in all, if a prettified tactic $t_p$ has the same effect its original
@@
 \section{Case Study}\todo{comment on the case study}
-Since the proofs generated by \tactictoe are meant to be parts of a \holfour development, it interesting to compare their quality with the original human proofs.
+Since the proofs generated by \tactictoe are meant to be part of a \holfour development, it interesting to compare their quality with the original human proofs.
 The quality of a proof in \holfour can be measured in terms of length, readability, maintainability and verification speed.
 We study these properties in three examples taken from our full-scale experiment.
-We list the time needed by \holfour to check a proof in parentheses.
-
-We start with Example~\ref{ex:cs1}, which proves that greatest common divisors are unique.
-The human formalizer recognized that it follows from two theorems.
-The relevance filtering of \tactictoe is not as powerful as that used in hammers
-and is therefore not able to find the \texttt{DIVIDES\_ANTISYM} property. The proof
-proceeds instead by rewriting the
-goal with the definitions then spliting the goal into multiple cases until the goal obligations are similar
-to \texttt{DIVIDES\_ANTISYM}. As expected, \tactictoe's proof
+The time needed by \holfour to check a proof is shown in parentheses.
+
+Example~\ref{ex:cs1} is about proving that greatest common divisors are unique.
+The human developer recognizes that it follows easily from two theorems.
+Our prover \tactictoe is not able to find that \texttt{DIVIDES\_ANTISYM} is
+necessary directly from the shape of the original theorems so it rewrites the
+goal then splits the goal into multiple cases until the goal obligations are similar
+to \texttt{DIVIDES\_ANTISYM} to be predicted. As expected, \tactictoe's proof
 also takes much longer to check.
@@
 element of the concatenation of $l_1$ and $l_2$ is the same as
-updating the $m^{th}$ element of $l_2$ where $m = n - p$. Again, the
+updating the $m^{th}$ element of $l_2$ where $m = n - p$. Again,
 \tactictoe's proof is
-much more readable. It starts by applying induction on $l$. It solves the base
-case of the induction by rewriting and proceeds by cases on $n=0$ or $n>0$
-in the induction hypothesis. It finalizes the proof with a short call to \metis
+much more readable starting with applying induction on $l$. It solves the base
+case of the induction by rewriting and distinguish cases where $n=0$ or $n>0$
+in the inductive hypothesis. It finalizes the proof with a short call to \metis
 using the definition of
@@
 
-In Example~\ref{ex:cs4}, we experiment with \tactictoe on a
-goal that does not originate from the \holfour library. The conjecture is
-that the set of
+In Example~\ref{ex:cs4}, we demonstrate the capabilities of \tactictoe on a 
+goal that was not already proved in the \holfour libraries: that the set of 
 numbers $\{0,...,n+m-1\} \setminus \{0,...,n-1\}$ is the same as the set 
@@
 \tactictoe uses the simplification set \texttt{ARITH\_ss} to reduce arithmetic 
-formulas. This exemplifies another advantage that \tactictoe has over ATPs,
-namely its ability to take advantage of user-defined simplification sets.
+formulas. This exemplifies an another advantage that \tactictoe has over ATPs, 
+mainly its ability to take advantage of user-defined simplification sets.
 
@@
 
+The system
 \textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
-groups related proofs using clustering
+groups related proofs thanks to its clustering
 algorithms. It allows \coq users to inspire themselves from similar proofs and
@@
 duplicated proofs. Comparatively, our predictions comes from a much more 
-detailed description of the target goal.\todo{and can be used as an ATP}
+detailed description of the targeted goal.
 %However, we simply create a single label for each tactic call whereas each of
@@
 
-\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} can
+\textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} is a powerful system able to
 generate
 proof scripts from previous \coq proof examples.
-Its strength lies in the ability to produce likely sequences
+Its strength lies in its ability to produce likely sequences
 of tactics for solving domain specific goals. It operates by creating a model
 for common sequences of tactics for a specific library.
-This means that in order to propose the next tactic, only the previously
+This means that in order to propose the following tactic, only the previously
 called tactics
@@
 
-\todoi{maybe the paragraph can be generally proof strategy languages including ltac and eisbach}
 A similar effort for stronger automation is performed in \isabelle. Nagashima
@@
 This can be done either in an interactive way, when the user completes the
-proof based on the recommended lemmas, as in the Mizar proof advisor
-\textsc{MizAR}~\cite{Urb04-MPTP0,mizAR40}, or attempted fully automatically, where such lemma
+proof based on the recommended lemmas, as in the \textsc{Mizar Proof
+Advisor}~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma
 selection is handed over to the ATP component of a \emph{hammer}
-system, as in the hammers for \holfour~\cite{tgck-cpp15}, \hollight~\cite{holyhammer} and \isabelle~\cite{BlanchetteGKKU16}.
+system~\cite{hammers4qed,tgck-cpp15,holyhammer,BlanchetteGKKU16,mizAR40}.
 
@@
 We proposed a new proof assistant automation technique which combines
-tactic-based proof search with machine learning prediction.
+tactic-based proof search with machine learning prediction
 Its implementation, \tactictoe, achieves an overall success rate of 66.4\%
@@
 Thanks to the learning abilities of \tactictoe, the generated proof scripts
-often reveal the high-level structure of the proof. %For these reasons,
+usually reveal the high-level structure of the proof. %For these reasons,
 We therefore believe that predicting ITP tactics based on the current goal
@@
 new tactics could be created by programming them with any construction
-available in \sml.
-The proof search guidance can also be improved, for example by considering
-stronger machine learning algorithms
+available in \sml. There is also something to be gain by
+having a better guidance for the search.
+That is why se also want to experiment with stronger machine learning algorithm
 such as deep neural networks as a model for the policy and evaluation in MCTS.
-The quality of such models could be enhanced by
+This quality of this model could be enhanced by additional data augmentation by
 training and testing existing and synthesized tactics.
 Conjecturing suitable intermediate steps could also allow \tactictoe to solve
-problems which require long proofs by decomposing it into multiple easier
+a problem which requires a long proof by decomposing it into multiple easy
 steps.

### 2018-02-20 tactictoe_jv.tex

--- 2018-02-19 tactictoe_jv.tex
@@
 eliminating unnecessary tactics and theorems has been developed and
-improve the user experience greatly~\cite{DBLP:conf/sefm/Adams15}.
+improves the user experience greatly~\cite{DBLP:conf/sefm/Adams15}.
 

### 2018-02-20 tactictoe_jv.tex
minor fix in diff

@@
 Because validation functions are only important to check the final proof, we
-omit them during the description of proof search algorithm. We denote by $t(g)$
-the list of goals produced by the application of a tactic $t$ to a goal $g$.
+omit them during the description of the proof search algorithm. We denote by 
+$t(g)$ the list of goals produced by the application of a tactic $t$ to a goal 
+$g$.
 
@@
 The exploration term is determined by the prior policy and the current policy.
-The current policy is calculate from the number of times a node $x$ has been
+The current policy is calculated from the number of times a node $x$ has been
 traversed during node selection, denoted $\mathit{Visit}(x)$.

### 2018-02-20 tactictoe_jv.tex
removing early results for terms

@@
 enough information to be predicted accurately.
-We also have positive results with term abstraction and plan to include it soon
-in our default algorithm.
+
+%We have positive results 
+%with term abstraction but they are not evaluated in 
+%Section~\ref{sec:evaluation}.
 
@@
 
-%\subsection{Recent improvements}\todo{maybe remove this subsection}
-%The final results could be improved by a few percent by including the
-%following additional techniques that are already available.
-%The technique already discussed in Section~ would most likely improve the
-%results by 2-3 percent. Reducing the exploration coefficient from 4 to 2 would
-%also probably give a 3 percent increase in the number of proven theorems as
-%shown by a recent 1 in 10 evaluation on the full library.
-
-\paragraph{Self-learning}\todo{explain better or remove or move to future work}
+\paragraph{Reinforcement learning}
 All our feature vectors have been learned form human proofs. We now can now
@@
 tactics is essential to have a beneficial effect.
-Since recording and re-proving are intertwined, the additional data is
-available for the next proof search.
+Since recording and re-proving are intertwined during evaluation, the 
+additional data is available for the next proof search.
 The hope is that the algorithm will improve faster by learning from its own
@@
 scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that
-self-learning improves \tactictoe's success rate by less than one percent.
+this one shot reinforcement learning method increases \tactictoe's success rate 
+by half a percent.
 
 \subsection{Complexity of the proof search}
-
 \pgfplotscreateplotcyclelist{my black}{
 solid, mark repeat=100, mark phase=0, black!100\\
-dashed, mark repeat=100, mark phase=0, black!100\\
-}
+dashed, mark repeat=100, mark phase=0, black!100\\}
 

### 2018-02-20 tactictoe_jv.tex
removing todos

@@
 Assuming we have a way to estimate the number steps required to solve a goal,
-a possibly useful subsumption definition could be: \todo{shorter or equal}
-\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ has a shorter proof than }
+a possibly useful subsumption definition could be:
+\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ has a proof with a length 
+shorter 
+or equal to the proof of }
 g_2\]
@@
 \end{definition}
-\todoi{@TG: Which of the two above?}
-
-We can naturally extend a goal subsumption to lists of goals.
+
+We can naturally extend any goal subsumption to a subsumption on
+lists of goals.
 \begin{definition} (Goal list subsumption)\\
@@
 \section{Experimental Evaluation}\label{s:experiments}
-\todo{Josef: give the specification}
-Each re-proving experiment is performed on a single CPU of a server in Prague
-with the exception of the experiments
-relying on asynchronous \eprover calls that run on two CPUs.
+The execution of \tactictoe's main loop in each re-proving experiment is 
+performed on a single CPU. An additional CPU is needed for experiments relying 
+on asynchronous \eprover calls.
 
@@
 For each theorem we first attempt to run the \tactictoe search with a time
-limit of 60 seconds, before processing the original proof script.
-In this way, the fairness of the experiments is guaranteed by construction.
-Only previously declared \sml
-variables (essentially tactics, theorems and simpsets) are accessible.
-And for each theorem to be re-proven \tactictoe is only trained on previous
-proofs.
-
-\todoi{rephrase to something more coherent}
+limit of 60 seconds before processing the original proof script.
+In this way, the fairness of the experiments is guaranteed by construction:
+only previously declared \sml variables (essentially tactics, theorems and 
+simpsets) are accessible to \tactictoe.
+
 \paragraph{Datasets: optimization and validation}
-
 All top-level theorems from the standard library are considered with the
@@
 Therefore, during the full-scale experiments, we evaluate 7164 theorems.
-We use every tenth theorem of the first third of the standard library of
-them for parameter optimization, which amounts to 273 theorems.
+We use every tenth theorem of the first third of the standard library for 
+parameter optimization, which amounts to 273 theorems.
+
 Although the evaluation of each set of parameters on its own is fair,
@@
 Section~\ref{sec:proofdisplay}.
-For \eprover, reconstruction rate varies between 95 and 99 percent, depending
-on the reconstruction method \cite{}\todo{CK}.
+Comparatively, we achieve a reconstruction rate of 95 percent for \eprover,
+by calling \metis for 2.0 seconds with the set of theorems necessary for 
+in \eprover's proof as argument.
 
@@
 discovered proofs than from the human proof
-scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments indicate that
+scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments show that
 this one shot reinforcement learning method increases \tactictoe's success rate 
-by half a percent.
+by less than a percent.
 
@@
 time. Therefore, integrating strategy scheduling in \tactictoe could be
-something to be experimented with. Overall, we recommend using \tactictoe and
-\eprover with a 10 seconds timeout. It is possible also to run them with a
-longer time limit, but we advise to do so in parallel to avoid waiting time.
-\todo{its strange to ``advise'' in a paper?}
+something to be experimented with.
+
+%Overall, a good compromise is to run \tactictoe and \eprover with a 10 seconds 
+%timeout. It is posso run them with a
+%longer time limit, but we do so in parallel to avoid waiting time.
+%\todo{its strange to ``advise'' in a paper?}
 
@@
 
-The total effect on the readability of a proof script is depicted in
+The total effect of these minimization and prettification techniques 
+on the readability of a proof script is depicted in
 Example~\ref{ex:pretty}.
-\todo{run without prettification}
-\begin{example}\label{ex:pretty}\ \\
-Discovered proof script for the theorem \texttt{EVERY\_MAP}
+
+\begin{example}(Theorem \texttt{EVERY\_MAP})\label{ex:pretty}\\\
+Proof script before minimization and prettification:
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-Minimized and prettified proof script:
+Proof script after minimization and prettification:
 \begin{lstlisting}[language=SMLSmall]
@@
 
-\section{Case Study}\todo{comment on the case study}
+\section{Case Study}
 Since the proofs generated by \tactictoe are meant to be parts of a \holfour development, it interesting to compare their quality with the original human proofs.
@@
 algorithms. It allows \coq users to inspire themselves from similar proofs and
-notice
-duplicated proofs. Comparatively, our predictions comes from a much more
-detailed description of the target goal.\todo{and can be used as an ATP}
+notice duplicated proofs. Comparatively, our predictions comes from a much more
+detailed description of the target goal. \tactictoe can also organize the 
+predictions to produce verifiable proofs and is not restricted to user 
+advice.
 %However, we simply create a single label for each tactic call whereas each of

### 2018-02-20 tactictoe_jv.tex
3 more spotted grammar

@@
 In this way, the fairness of the experiments is guaranteed by construction:
-only previously declared \sml variables (essentially tactics, theorems and 
+only previously declared \sml values (essentially tactics, theorems and 
 simpsets) are accessible to \tactictoe.
@@
 Comparatively, we achieve a reconstruction rate of 95 percent for \eprover,
-by calling \metis for 2.0 seconds with the set of theorems necessary for 
+by calling \metis for 2.0 seconds with the set of theorems used
 in \eprover's proof as argument.
@@
 algorithms. It allows \coq users to inspire themselves from similar proofs and
-notice duplicated proofs. Comparatively, our predictions comes from a much more
+notice duplicated proofs. Comparatively, our predictions come from a much more
 detailed description of the target goal. \tactictoe can also organize the 

### 2018-02-20 tactictoe_jv.tex
moving abstract to introduction

--- 2018-02-20 tactictoe_jv.tex
@@
 \begin{abstract}
-Formalization of mathematics in interactive theorem provers (ITPs) relies today on many proof automation methods.
-Some of these methods are general-purpose and with enough time they can prove any provable conjecture.
-Many formalized proof scripts however also take advantage of specialized rules and theory-based strategies.
-In the HOL4 ITP system, practically all of them are implemented as \emph{tactics}.
-
-Automated theorem provers (ATPs) rely on parameterizable general
-strategies. Sometimes, a strategy can be specialized for only one theory as in
-SMT solving. But it is difficult for an ATP to rely on rules tailored for a particular case.
-Indeed, this would require implementing many specific rules and more
-importantly figuring out in which cases they can be applied.
-
-We bridge this gap by implementing our tactical prover TacticToe on top
-of the HOL4 ITP.
-TacticToe first learns from many proof scripts
-the relevance of tactics in particular proof situations.
-This learned knowledge is then used in a Monte Carlo tree search algorithm that explores the most promising tactic-level proof paths.
-On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\%
-of the 7164 theorems in HOL4's standard library whereas
+
+We implement a tactical prover TacticToe on top of the HOL4 interactive theorem 
+prover. TacticToe learns from many human proofs which mathematical method is 
+useful in particular proof situations.
+This knowledge is then used in a Monte Carlo tree search algorithm that 
+explores the most promising tactic-level proof paths.
+On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of 
+7164 theorems in HOL4's standard library whereas
 Eprover solves 34.5\%. The success rate rises to 69.0\% by combining the
-results of both provers.
+results of \tactictoe and \eprover.
 \end{abstract}
@@
 \tableofcontents
+
 \section{Introduction}
@@
   and \coq~\cite{coq-book} provide high-level parameterizable tactics for constructing proofs.
-  Such tactics typically analyze the current proof state (goal and
-  assumptions), then apply non-trivial proof transformations that get
-  expanded into many basic kernel-level inferences or large parts of a proof term.
-%\marginpar{not Coq}
-  In this work we develop a tactic-level automation procedure for the \holfour ITP
-  which guides the selection of tactics by learning from previous
-  proofs.  Instead of relying on translation to first-order automated
-  theorem provers (ATPs) as done by the hammer
-  systems~\cite{hammers4qed,tgck-cpp15}, our technique
-%\marginpar{too many cites?}
-%BlanchetteGKKU16,holyhammer,mizAR40
-  searches directly for sequences of tactic applications that lead to
-  an ITP proof, thus avoiding the translation and proof-reconstruction phases needed by the hammers.
-
-  To do this, we \emph{extract and record} tactic invocations from the ITP
+  Such tactics analyze the current proof state (goal and
+  assumptions) and apply non-trivial proof transformations.
+  Formalized proofs take advantage of different levels of automation: 
+  specialized rules, theory-based strategies and general purpose startegies. 
+  Thanks to the progress in proof 
+  automation, developers can delegate more and more complicated proof 
+  obligations to general purpose strategies. Those are implemented by automated 
+  theorem provers(ATPs) such as \eprover~\cite{eprover}. The communications 
+  between 
+  an ITP and ATPs is made possible by a ``hammer''  
+  system~\cite{hammers4qed,tgck-cpp15}. It acts as an interface by performing 
+  premise selection, translation and proof reconstruction.
+  Yet, ATPs are far from foolproof and more precise user-guidance, achieved 
+  by applying a particular sequence of specialized rules, is almost always 
+  necessary to develop a mathematical theory.
+
+  In this work, we develop in the ITP \holfour a procedure to select suitable 
+  tactics depending on the current proof state by learning 
+  from previous proofs. Backed by this machine-learned guidance, our prover 
+  \tactictoe relies on Monte 
+  Carlo tree search~\cite{montecarlo} to find sequences of tactic applications 
+  leading to an ITP proof.
+  To this end, we \emph{extract and record} tactic invocations from the ITP
   proofs (Section~\ref{sec:recording}) and
@@
   successful, is a verified human-level proof composed of \holfour
-  tactics.  The system is evaluated on a large set of theorems originating
+  tactics (Section~\ref{sec:proofdisplay}). The system is evaluated on a large 
+  set of theorems originating
   from \holfour (Section~\ref{s:experiments}).
@@
 \paragraph{Goals and theorems.}
-A goal is represented as a \emph{sequent}.
+A goal (or proof state) is represented as a \emph{sequent}.
 A sequent is composed of a set of assumptions and a conclusion, all of which are higher-order logic formulas.
@@
 \sml function.
-\item Monte Carlo tree search~\cite{montecarlo} (MCTS) replaces A* as our
+\item Monte Carlo tree search (MCTS) replaces A* as our
 search algorithm. The MCTS algorithm gives a more balanced feedback on
@@
 the length of the original human proofs.
-\item Minimization and prettification of the generated proof improves user
+\item Minimization and prettification of the generated proof  facilitates user
 interaction.

### 2018-02-20 tactictoe_jv.tex
abstract

@@
 
-\todoi{JU review: Still feels more like an intro than an abstract. Could be
-shorter and punchier.}
-
-\todoi{Move part of the abstract to the introduction.}
 \begin{abstract}
-
 We implement a tactical prover TacticToe on top of the HOL4 interactive theorem 
-prover. TacticToe learns from many human proofs which mathematical method is 
-useful in particular proof situations.
+prover. TacticToe learns from human proofs which mathematical method is 
+useful in a particular proof situation.
 This knowledge is then used in a Monte Carlo tree search algorithm that 
@@
 Eprover solves 34.5\%. The success rate rises to 69.0\% by combining the
-results of \tactictoe and \eprover.
+results of TacticToe and Eprover.
 \end{abstract}

### 2018-02-20 tactictoe_jv.tex
minor
introduction of evaluation

@@
 
-\section{Search tree}\label{sec:prelim}
-
+\section{Search Tree}\label{sec:prelim}
 The application of a tactic to an initial conjecture produces a list of goals.
@@
 Moreover, it is possible to try multiple tactics on the same goal.
-In order to keep track of the progress made so far from the initial conjecture we want to solve, we organize goals and tactics into a graph with list of goals as nodes and tactics as edges (See Definition~\ref{def:stree}).
+In order to keep track of the progress made so far from the initial conjecture 
+we want to solve, we organize goals and tactics into a graph with list of goals 
+as nodes and tactics as edges (See Definition~\ref{def:stree}).
 
@@
 The input of the algorithm is a goal $g$ (also called conjecture) that we
-want to prove.
-\todo{@TG: Check sentence below}
-Therefore the search tree starts with only one node the list of goals
-$[g]$ and an ordered list of predicted tactics for the goal $g$.
+want to prove. Therefore the search tree starts with only one node containing 
+the list of goals
+$[g]$.
 %Then, the MCTS algorithm alternates node selection steps and node extension
@@
 \item It saturates, i.e., there are no tactics to be applied to any open goal.
-This happens less than 10 times in the full-scale experiment. \todo{@TG: why don't we have the original tactics as options?}
+This occurs less than 10 times in the full-scale experiment.
+This happens only at the beginning of \holfour library, where the training data 
+has very few tactics. 
 \item It exceeds a given time limit (default is set to 60 seconds).
@@
 \section{Feature Extraction}\label{sec:features}
-\todo{@TG: decide which of the next sentences}
-We explain the construction of the similarity measure by which relevant
-objects are predicted
-
-We predicts tactics through their associated goals. So, the features
+%We explain the construction of the similarity measure by which relevant
+%objects are predicted
+
+We predict tactics through their associated goals. So, the features
 for each kind of object can be extracted from (different representations of)
@@
 
-To preselect a list of goal $l$ from our database of lists of goals, we
+We do preselection also for lists of goals.
+To preselect a list of goals $l$ from our database of lists of goals, we
 consider the tactic input $g$ that was at the origin of the production of $l$
-during orthogonalization. The list $l$
-is preselected if and only if $g$ appears in the 500 preselected tactic-goal
-pairs. This has the
-desirable effect that negative and positive examples derived from the same
-input goal are chosen or discarded together.
-\todo{this paragraph was confusing for Lasse and Cezary}
+during orthogonalization. The list $l$ is preselected if and only if $g$ 
+appears in the 500 preselected tactic-goal pairs. 
+%This has the
+%desirable effect that negative and positive examples derived from the same
+%input goal are chosen or discarded together.
 
@@
 
-Fix parameters: not optimized here.
- - Features extraction method.
- - k-nn distance.
- - MCTS evaluation radius is set to 10.
- - Metis : 16 lemmas.
- - Eprover: 128 lemmas and 5 seconds timeout 1 asynchronous call
+We optimize \tactictoe by tuning parameters of six different techniques:
+the timeout of tactics, orthogonalization, abstraction, MCTS policy, MCTS 
+evaluation and ATP integration. During training, optional techniques such as 
+orthogonalization may also be turned off completely.
+\tactictoe also includes other important techniques which are run with their 
+default parameters: features extraction, prediction 
+algorithms, number of premises for ATPs and MCTS evaluation radius (set to 
+10).
+By choosing a set of parameters, we create a strategy for 
+\tactictoe that is evaluated on our training set of 273 theorems. And the 
+strategy with the highest number of proven theorems is selected for 
+a full scale evaluation.
 
@@
 
-In Table~\ref{tab:tuning}, we run \tactictoe with different parameters.
+In Table~\ref{tab:tuning}, the success rate of the different \tactictoe  
+strategy is presented.
 The first four techniques are tested relative to the same baseline indicated in
@@
 
-\textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
-groups related proofs using clustering
-algorithms. It allows \coq users to inspire themselves from similar proofs and
-notice duplicated proofs. Comparatively, our predictions come from a much more
-detailed description of the target goal. \tactictoe can also organize the 
-predictions to produce verifiable proofs and is not restricted to user 
-advice.
-%However, we simply create a single label for each tactic call whereas each of
-%its
-%arguments is treated independently in \textsf{ML4PG}.
-%Our choice is motivated by the k-NN algorithm already used in
-%\holyhammer for the a of theorems.
-%, which can be easly adapted to the selection of tactics but would be improved
-%by a
-%version that can learn to predict tactics and their arguments independently.
-
 \textsf{SEPIA}~\cite{DBLP:conf/cade/GransdenWR15} can
@@
 \eprover on the \holfour standard library.
+
+\textsf{ML4PG}~\cite{DBLP:journals/corr/abs-1212-3618,DBLP:journals/mics/HerasK14}
+groups related proofs using clustering
+algorithms. It allows \coq users to inspire themselves from similar proofs and
+notice duplicated proofs. Comparatively, our predictions come from a much more
+detailed description of the target goal. \tactictoe can also organize the 
+predictions to produce verifiable proofs and is not restricted to user 
+advice.
+%However, we simply create a single label for each tactic call whereas each of
+%its
+%arguments is treated independently in \textsf{ML4PG}.
+%Our choice is motivated by the k-NN algorithm already used in
+%\holyhammer for the a of theorems.
+%, which can be easly adapted to the selection of tactics but would be improved
+%by a
+%version that can learn to predict tactics and their arguments independently.
+
 %A recent improvement on this system~\cite{}\todo{Cezary: could not remember
@@
 %arguments is more difficult but is likely to produce more suitable lemmas.
-%Proof patching: \cite{RingerYLG18}
+%
+Proof patching: \cite{RingerYLG18}
 
@@
 program with regular tactics and combine them with tools such as
-\sledgehammer~\cite{sledgehammer10} or Nitpick~\cite{Nitpick10}.
+\sledgehammer.
 Selection of tactics from data collected in recorded proofs is in development

### 2018-02-20 tactictoe_jv.tex
Reread 8.2

@@
 \tactictoe also includes other important techniques which are run with their 
-default parameters: features extraction, prediction 
-algorithms, number of premises for ATPs and MCTS evaluation radius (set to 
+default parameters: feature extraction, prediction 
+algorithms, number of premises for ATPs and MCTS evaluation radius (set by default to
 10).
@@
 Thanks to the larger search time available in this experiment, the timeout for
-each tactic can be raised from the 0.02 seconds used in \tactictoe's first
-experiment~\cite{tgckju-lpar17} to 0.05
+each tactic can be increased from the 0.02 seconds used in \tactictoe's initial
+experiments~\cite{tgckju-lpar17} to 0.05
 seconds. Removing tactics with similar effect as performed by the
@@
 contributes significantly as well. Experiments involving \eprover, run the ATP
-on one thread asynchronously with a timeout of 5 seconds and 128 premises.
+using a separate proces asynchronously with a timeout of 5 seconds and 128 premises.
 The fact that \metis outperforms \eprover in this setting is due to the fact
@@
 of search tree quickly. Because \metis is weaker that \tactictoe as an ATP, it
-is given less predictions. But even if a lemma essential for the proof of a
+is given less predictions. But even if an essential lemma for the proof of a
 goal is not predicted, a modification on the goal performed by a tactic can
@@
 
-The results shows that \tactictoe is able to prove almost twice as much
+The results shows that \tactictoe is able to prove almost twice as many
 theorems as \eprover. Combining the results of \tactictoe and \eprover we get a
-69.0\% success rate which is significantly above the 5O\% success rates
+69.0\% success rate which is significantly above the 50\% success rates
 of hammers on this type of problems~\cite{tgck-cpp15}. Moreover, \tactictoe is
@@
 proof succeeds, unless one of the tactic modifies the state of \holfour in a
-way that changes the behavior of a tactic used in the final proof. Such an
-event has not yet occurred.
+way that changes the behavior of a tactic used in the final proof. This has not
+occurred a single time in our experiments.
 More details on how a proof is extracted from the search tree is given in

### 2018-02-20 tactictoe_jv.tex
minor
Minor grammar tweaks

@@
   and \coq~\cite{coq-book} provide high-level parameterizable tactics for constructing proofs.
-  Such tactics analyze the current proof state (goal and
+  Tactics analyze the current proof state (goal and
   assumptions) and apply non-trivial proof transformations.
@@
   specialized rules, theory-based strategies and general purpose startegies.
-  Thanks to the progress in proof 
+  Thanks to progress in proof
   automation, developers can delegate more and more complicated proof
   obligations to general purpose strategies. Those are implemented by automated
-  theorem provers(ATPs) such as \eprover~\cite{eprover}. The communications 
+  theorem provers (ATPs) such as \eprover~\cite{eprover}. Communication
   between
@@
   from previous proofs. Backed by this machine-learned guidance, our prover
-  \tactictoe relies on Monte 
-  Carlo tree search~\cite{montecarlo} to find sequences of tactic applications 
+  \tactictoe relies on \emph{Monte
+  Carlo tree search}~\cite{montecarlo} to find sequences of tactic applications
   leading to an ITP proof.
@@
   examples (Section~\ref{s:prediction}).  The learned data serves as
-  guidance for a \emph{Monte Carlo tree search} that explores different
+  guidance for the Monte Carlo tree search that explores different
   proof paths (Section~\ref{sec:proofsearch}). The result, if
   successful, is a verified human-level proof composed of \holfour
-  tactics (Section~\ref{sec:proofdisplay}). The system is evaluated on a large 
+  tactics (Section~\ref{sec:proofdisplay}). We evalute the system on a large
   set of theorems originating
@@
 \tactictoe that is evaluated on our training set of 273 theorems. And the
-strategy with the highest number of proven theorems is selected for 
+strategy with the highest number of re-proven theorems is selected for
 a full scale evaluation.

### 2018-02-22 tactictoe_jv.tex
definition of search tree + preselection section

--- 2018-02-20 tactictoe_jv.tex
@@
 \item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to
-$\mathbb{A}$. It takes a goal and a tactic and returns the node produced by the
-application of the tactic to the goal. Its support $\mathit{Support}(A)$ is
-defined as $\lbrace (g,t) \in \mathbb{G} \times \mathbb{T}\ |\ t \in T(g)
-\rbrace$. And if $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$. This
-means that the output $t(g)$ is exactly the list
-of goals contained in the node produced by
-$t$ from $g$.
+$\mathbb{A}$. Its support $\mathit{Support}(A)$ is 
+$\lbrace (g,t) \in \mathbb{G} \times \mathbb{T}\ |\ t \in T(g)
+\rbrace$. And if $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$. In 
+other words, the output $t(g)$ is exactly the list
+of goals contained in the node produced by $A$ from $(g,t)$.
+\item It is acyclic, i.e. a node cannot be a descendant (see 
+Definition~\ref{def:desc}) of itself.
+\item There is exactly one node with no parents. This root node contains 
+exactly one goal which is the initial goal (also called conjecture).
 \end{itemize}
 
-Acyclicity and the fact that there is exactly
-one root node containing only the initial goal are additional constraints to
-the search tree that are preserved during the
-application of the MCTS algorithm. That is why we call it a tree, although this
-is not specified in this definition.
 \end{definition}
@@
 predictions is described in Section~\ref{sec:features} and two methods for
-improving the quality of the predictions are presented in
-Section~\ref{sec:ortho} and Section~\ref{sec:synthesis}.
+improving the speed and quality of the predictions are presented in
+Section~\ref{sec:presel},\ref{sec:ortho} and \ref{sec:synthesis}.
 
@@
 
-\subsection{Preselection}\label{sec:dependencies}
+\section{Preselection}\label{sec:presel}
 
@@
 \tactictoe that is evaluated on our training set of 273 theorems. And the 
-strategy with the highest number of proven theorems is selected for 
+strategy with the highest number of re-proven theorems is selected for 
 a full scale evaluation.
@@
 
-%\subsection{Increasing \tactictoe's knowledge}\label{sec:perfect_exp}
-%\todoi{decide to keep or not}
-%We investigate what are the limits of \tactictoe and what benefits additional
-%knowledge could provide. To this end, we will modify the set of features to
-%include features vectors that should not be known in a regular experiment. We
-%include some vectors from the recorded human proof and perform the proof
-%search
-%after the recording instead of before in all other experiments. This
-%experiment
-%is not fair and should only be compared to
-%experiments in the same settings.
-%
-%For clarity, we will name the set of feature vectors that can be recorded from
-%the current human proof $\mathbb{H}$ and the set of vectors recorded from
-%previous proofs $\mathbb{P}$ (consistent with previous notation?). We note
-%$\pi_1^1$ the projection on the
-%first component returning the label of a feature vector. And we define the
-%function $T$ that takes a tactic and return the list of its tokens.
-%
-%We will progressively add bigger and bigger
-%subset of $\mathbb{H}$ to $\mathbb{P}$ and observe the effect on the proof
-%search success rates.
-%
-%Here are the list of increasing non-trivial subsets of $\mathbb{H}$ we
-%consider:
-%\begin{align*}
-%\mathbb{F}_1 &= \lbrace (t,g) \in \mathbb{H}\ |\ t \in \pi_1^1(\mathbb{P})
-%   \rbrace \\
-%\mathbb{F}_2 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists
-%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \rbrace\\
-%\mathbb{F}_3 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists
-%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem} \rbrace\\
-%\mathbb{F}_4 &= \lbrace (t,g) \in \mathbb{H}\ |\ \forall x \in T(t).\ \exists
-%t'\in \pi_1^1(\mathbb{P}).\ x \in T(t') \vee x\ \mbox{is a theorem or a term}
-%\rbrace\\
-%\end{align*}
-%
-%
-%$\mathbb{F}_1$ is an approximation of what we could get if we had perfect
-%prediction for those tactics.
-%
-%\begin{table}[h]
-%\centering\ra{1.3}
-%\small
-%\begin{tabular}{lll}
-%\toprule
-% & Additional features & Solved \\
-%\midrule
-%$\emptyset$   & none & 119 (43.6\%) \\ %NONEv2
-%$\mathbb{F}_1$& tactic prediction & 121 (44.3\%) \\ %AFTERTACv2
-%$\mathbb{F}_2$& token recombination  & 137 (50.2\%) \\ %AFTERTOKENv2
-%$\mathbb{F}_3$& theorem prediction  & 207 (75.8\%) \\ %AFTERTHMTHMv2
-%$\mathbb{F}_4$& term prediction & 223 (81.7\%) \\ %AFTERALLv2
-%$\mathbb{H}$  & all & 239 (87.5\%) \\ %AFTERSMALLv2
-%% $AFTERv2$   & $\mathfrak{H}$ & 245 (89.7\%)  \\
-%\bottomrule
-%\end{tabular}
-%\caption{\label{tab:featue_param} Effect of augmenting \tactictoe's knowledge
-%with tactics from the proof of the tested theorem. The experiments was run
-%with
-%a time out of 5 seconds on an older version of \tactictoe, as can be seen in
-%the low success rate of thedefault strategy.}
-%\end{table}
-%
-%
-%
-%And we observe in Table~\ref{tab:featue_param} a
-%slight increase compared to the default \tactictoe. We then included all
-%vectors that contains tactics that can be rebuild from tokens of
-%tactics in our database.
-%Our search algorithm proved two proofs less that if it had perfect prediction
-%so it means that our prediction algorithm (policy) is doing a good  job at
-%least in this early settings where proofs are easy.
-%Suprisingly, being able to recombine tokens of tactics from the database is
-%barely sufficient to prove more than half of the theorems in this dataset. The
-%Therefore, additionally including in tactics theorems and terms that did not
-%appear in previous tactics is essential. Creating from scratch other arguments
-%may not help as much. This is why we restricted ourselves to predicting
-%arguments of theorem list type and the term type during tactic synthesis.
-%Finally our proof search algorithm sometimes fails even if it knows all the
-%tactic necessary. The two main reasons are that either a crucial tactic
-%exceeds its timeout or that the human proof is too deep to be rediscovered
-%(more than 10 steps long). In all those experiments orthogonalization is
-%performed with
-%radius 20.
-
 \subsection{Full-scale experiment}\label{sec:full_exp}

### 2018-02-22 tactictoe_jv.tex
fix typo
minor

@@
   successful, is a verified human-level proof composed of \holfour
-  tactics (Section~\ref{sec:proofdisplay}). We evalute the system on a large
+  tactics (Section~\ref{sec:proofdisplay}). We evaluate the system on a large
   set of theorems originating
@@
 of goals contained in the node produced by $A$ from $(g,t)$.
-\item It is acyclic, i.e. a node cannot be a descendant (see 
+\item It is acyclic, i.e. a node cannot be a strict descendant (see 
 Definition~\ref{def:desc}) of itself.

### 2018-03-05 tactictoe_jv.tex
replacing support by domain
removing domain
search tree definition
todo

@@
 
+\todoi{graphic explaining how each part relate to each other}
+
 \section{Search Tree}\label{sec:prelim}
@@
 takes a goal and return the set of tactics already applied to this goal.
-\item $G$ is a total function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$.
+\item $G$ is a function from $\mathbb{A}$ to $\mathcal{P}(\mathbb{G})$.
 It takes a node and returns the list of goals of this node.
-\item $A$ is a partial function from $\mathbb{G} \times \mathbb{T}$ to
-$\mathbb{A}$. Its support $\mathit{Support}(A)$ is 
-$\lbrace (g,t) \in \mathbb{G} \times \mathbb{T}\ |\ t \in T(g)
-\rbrace$. And if $(g,t) \in \mathit{Support}(A)$, then $t(g) = G(A(g,t))$. In 
-other words, the output $t(g)$ is exactly the list
-of goals contained in the node produced by $A$ from $(g,t)$.
+\item $A$ is a function that takes a pair $(g,t)\in \mathbb{G} \times 
+\mathbb{T}$ such as $t \in T(g)$ and 
+returns a node $A(g,t)$ such as $t(g) = G(A(g,t))$. In other words, the output 
+$t(g)$ is exactly the list of goals contained in $A(g,t)$.
 \item It is acyclic, i.e. a node cannot be a strict descendant (see 
@@
 \paragraph{Acknowledgments}\label{sect:acks}
-We would like to thank Lasse Blaauwbroek for his insightful comments which
+We would like to thank Lasse Blaauwbroek and Yutaka Nagashima for their 
+insightful comments which
 contributed to improve the quality of this paper. This work has been supported

### 2018-03-14 tactictoe_jv.tex
spellcheck

@@
   Formalized proofs take advantage of different levels of automation:
-  specialized rules, theory-based strategies and general purpose startegies.
+  specialized rules, theory-based strategies and general purpose strategies.
   Thanks to progress in proof
@@
 node selection starts from the root of the search tree. From the current node,
-the function \texttt{CompareChildren} choses among the children of its open
+the function \texttt{CompareChildren} chooses among the children of its open
 goal the one with the highest value.
@@
 the table by the tag ``default''.  This default strategy relies on a tactic
-time out of 0.05s, an orthogonalization radius of 20 and a policy coefficent of
+time out of 0.05s, an orthogonalization radius of 20 and a policy coefficient of
 0.5.
@@
 orthogonalization process is only beneficial when running \tactictoe for short
-period of time. It seems that the alloted time allows a strategy
+period of time. It seems that the allotted time allows a strategy
 running without orthogonalization to catch up.
@@
 contributes significantly as well. Experiments involving \eprover, run the ATP
-using a separate proces asynchronously with a timeout of 5 seconds and 128 premises.
+using a separate process asynchronously with a timeout of 5 seconds and 128 premises.
 The fact that \metis outperforms \eprover in this setting is due to the fact
@@
 The extracted proof script of $\mathfrak{T}$ is $P_{goal}(g_{root})$.
-We minimally improve it by subtituting \texttt{THENL} by \texttt{THEN} when the
+We minimally improve it by substituting \texttt{THENL} by \texttt{THEN} when the
 list of goals is a singleton and removing \texttt{THENL []} during the
@@
 proceeds instead by rewriting the
-goal with the definitions then spliting the goal into multiple cases until the goal obligations are similar
+goal with the definitions then splitting the goal into multiple cases until the goal obligations are similar
 to \texttt{DIVIDES\_ANTISYM}. As expected, \tactictoe's proof

### 2018-03-14 tactictoe_jv.tex
some fixes, in particular Def 10 seemed in need of flipping

@@
 \begin{definition} (Coverage)\\
-Let $\mathfrak{T}$ be the database of tactic-goal pairs. We can define the
+Let $\mathfrak{T}$ be the database of tactic-goal pairs. We define the
 coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times
@@
 \begin{definition} (Goal subsumption)\\
-A goal subsumption $\le$ is a partial relation goals.
-Assuming we have a way to estimate the number steps required to solve a goal,
-a possibly useful subsumption definition could be:
+A goal subsumption $\le$ is a partial relation on goals.
+Assuming we have a way to estimate the number of steps required to solve a goal,
+a % possibly 
+general and useful subsumption definition is for example:
+% could be:
 \[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ has a proof with a length
@@
 g_2\]
-By default, we choose a minimal subsumption defined by:
+By default, we however choose for efficiency reasons a minimal subsumption defined by:
 \[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ is }\alpha\mbox{-equivalent
@@
 \begin{definition} (Goal list subsumption)\\
-A goal list subsumption is an partial relation on lists of goals.
-Given two lists of goals $l_1$ and $l_2$, we can define it from goal
+A goal list subsumption is a partial relation on lists of goals.
+Given two lists of goals $l_1$ and $l_2$, we define it from the goal
 subsumption $\le$ by:
-\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_2 \in l_2.\ \exists g_1 \in l_1.\
+%\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_2 \in l_2.\ \exists g_1 \in l_1.\
+%g_1 \le g_2\]
+\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_1 \in l_1.\ \exists g_2 \in l_2.\
 g_1 \le g_2\]
@@
 
-If one of the tactic is failing then $t_1$ and $t_2$ are not comparable through
+If one of the tactics is failing then $t_1$ and $t_2$ are not comparable through
 this relation.
@@
 
-All in all, a database build with the orthogonalization process contains
+% All in all, 
+A database built with the orthogonalization process hus contains
 $(\mathbb{O}(t,g),g)$ instead of $(t,g)$.
@@
 \section{Abstraction}\label{sec:synthesis}
-One of the major weakness of the previous version of \tactictoe was that
+One of the major weaknesses of the previous version of \tactictoe was that
 it could not create its own tactics. Indeed, sometimes no previous tactic is
@@
 create tactics that are more general but have the same effect as the
-ungeneralized one. The generalized tactics are typically slower that their
+% ungeneralized 
+original one. The generalized tactics are typically slower that their
 original variants, but the added flexibility is worthwhile in practice. Moreover,
@@
 in \holfour. By the design of the proof search, reconstruction of \tactictoe
-proof succeeds, unless one of the tactic modifies the state of \holfour in a
+proof succeeds, unless one of the tactics modifies the state of \holfour in a
 way that changes the behavior of a tactic used in the final proof. This has not
@@
 All our feature vectors have been learned form human proofs. We now can now
-also add tactic-goal pairs that appears in final \tactictoe's proof. To prevent
+also add tactic-goal pairs that appears in \tactictoe's final proof. To prevent
 duplication of effort, orthogonalization of those
@@
 graph that most of the human proofs are short. The proofs found by \tactictoe
-follows a similar
-distribution. This means that if \tactictoe finds a proof, there is about a 50
+follow a similar
+distribution. % This means that 
+If \tactictoe finds a proof, there is about a 50
 percent chance that it will be shorter than what a human would come up with.
@@
 In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
-see how well \tactictoe and \eprover performs on increased level of
-difficulties. As expected, the longer the original proof is, the harder it is
+see how well \tactictoe and \eprover cope with increased levels of
+difficulty. As expected, the longer the original proof is, the harder it is
 for \tactictoe and \eprover to re-prove the theorem on their own.
@@
 \sml code can be exported and imported.
-\item To produce a tactic value form a code is easy in \sml, which can be
-achieved by using a reference to a tactic and update it with the function
+\item To produce a tactic value from a code is easy in \sml, which can be
+achieved by using a reference to a tactic and updating it with the function
 \texttt{use}.
@@
 In order to transfer our tactic knowledge between theories, we want to be able
-to re-use the code of a tactic recorded in one theory in an
-other. We also would like to make sure that the code is interpretable and that
+to re-use the code of a tactic recorded in one theory in another. 
+We also would like to make sure that the code is interpretable and that
 its interpretation does not change in the other theory.

### 2018-03-14 tactictoe_jv.tex
ack

@@
 contributed to improve the quality of this paper. This work has been supported
-by the ERC Consolidator grant no.\ 649043 \textit{AI4REASON} and ERC starting
-grant no.\ 714034 \textit{SMART}.
+by the ERC Consolidator grant no.\ 649043 \textit{AI4REASON}, the ERC starting
+grant no.\ 714034 \textit{SMART}, the Czech 
+project AI \& Reasoning CZ.02.1.01/0.0/0.0/15\_003/0000466 and the 
+European Regional Development Fund.
 

### 2018-03-19 tactictoe_jv.tex
Plan and diagram in the introduction

--- 2018-03-14 tactictoe_jv.tex
@@
 \usetikzlibrary{trees,positioning,fit}
-\tikzstyle{block} = [rectangle, draw,% fill=blue!20,
-    text width=5em, text centered, rounded corners, minimum height=4em]
+\tikzstyle{block} = 
+  [rectangle, draw, 
+   minimum width=9em, text centered, rounded corners, minimum height=3.5em]
 \tikzstyle{line}=[draw]
-\tikzstyle{cloud} = [draw, ellipse,%fill=red!20,
-    node distance=3cm,
-    minimum height=2em]
+\tikzstyle{cloud} = 
+  [draw, text centered, ellipse, minimum height=3.5em, minimum width=9em]
 
@@
   Formalized proofs take advantage of different levels of automation:
-  specialized rules, theory-based strategies and general purpose strategies.
+  specialized rules, theory-based strategies and general purpose startegies.
   Thanks to progress in proof
@@
   leading to an ITP proof.
-  To this end, we \emph{extract and record} tactic invocations from the ITP
-  proofs (Section~\ref{sec:recording}) and
-  \emph{build efficient machine learning classifiers} based on such training
-  examples (Section~\ref{s:prediction}).  The learned data serves as
-  guidance for the Monte Carlo tree search that explores different
-  proof paths (Section~\ref{sec:proofsearch}). The result, if
-  successful, is a verified human-level proof composed of \holfour
-  tactics (Section~\ref{sec:proofdisplay}). We evaluate the system on a large
-  set of theorems originating
-  from \holfour (Section~\ref{s:experiments}).
 
 \subsection{The problem}
-% How a user works with an ITP and how this gives rise to a machine learning problem.
-\todo{RK: this is just a first draft -- feel free to edit lots!}
 An ITP is a development environment for the construction of formal proofs in which it is possible, but not common, to write a proof as a sequence of applications of primitive inference rules.
@@
 
+\subsection{Plan}
+
+\begin{figure}[h]
+\begin{center}
+\begin{tikzpicture}[node distance = 1.9cm]
+  \node [] (aa) {\textbf{Learning}};
+  \node [cloud, below of=aa, node distance = 1cm] (a) {Formal library};
+  \node [block, below of=a] (b) {Proof recording};
+  \node [cloud, below of=b] (c) {Knowledge base}; 
+  \node [block, below of=c] (e) {Training loop};
+  \node [block, below of=e] (d) {Tactic predictors};
+
+  \draw [-to,black,thick] (a) to (b);
+  \draw [-to,black,thick] (b.south) to (c.north);
+  \draw [-to,black,thick] (c) to [bend left]  (e);
+  \draw [-to,black,thick] (e) to [bend left]  (c);
+  \draw [-to,black,thick] (c.east) to [out=330,in=30] (d);
+
+  \node [right of=aa, node distance=6.2cm] (0) {\textbf{Proving}};
+  \node [cloud, below of=0, node distance = 1cm] (1) {Conjecture};
+  \node [cloud, below of=1, node distance = 1.7cm] (2) {Search tree};
+  \node [block, below of=2, node distance=1.3cm] (3) {Tactic policy};
+  \node [above of=2, xshift=-16mm, node distance = 0.9cm] (00) {MCTS proof 
+  search};
+  \node [block, below of=3, node distance = 1.3cm] (4) {Tactic evaluation};
+  \node [block, fit=(2) (3) (4)] (5) {};
+  \node [block, below of=4, node distance = 1.7cm] (6) {Proof minimization};
+  \node [cloud, below of=6, node distance = 1.6cm] (7) {Proof};
+  \draw [-to,black,thick] (1) to (5);
+  \draw [-to,black,thick] (5) to (6);
+  \draw [-to,black,thick] (6) to (7);
+  \draw [-to,black,thick] (d) to [out=0, in=180] (3);
+  \draw [-to,black,thick] (d) to [out=0, in=180] (4);
+
+\end{tikzpicture}
+\end{center}
+\label{fig:components}
+\caption{Relation between modules of the learning and proving tool chains.
+Functions are represented by rectangles and objects by ellipses.}
+\end{figure}
+
+The different components depicted in Fig~\ref{fig:components} give a 
+high-level overview of the techniques necessary to generate 
+tactic-level proofs by learning from recorded knowledge. From a formal library 
+of ITP proofs, we extract and record tactic invocations 
+in Section~\ref{sec:recording}. Based on such training examples, we
+  build efficient machine learning classifiers in Section~\ref{s:prediction}. 
+  At the basis of those predictions are 
+  similarity measures computed from syntactic features of mathematical formulas 
+  in Section~\ref{sec:features}.
+  Modification of the knowledge base performed by orthogonalization 
+  in Section~\ref{sec:ortho} and abstraction in Section~\ref{sec:synthesis}
+  improves the quality of the databases.
+  And the preselection method in Section~\ref{sec:presel} increases the speed 
+  of the prediction algorithm.
+  The learned predictors relying on this knowledge serves as
+  guidance for the proof search algorithm in Section~\ref{sec:proofsearch} 
+  which explores different proof paths in the proof tree in 
+  Section~\ref{sec:prelim}. 
+  The result, if
+  successful, is a verified human-level proof composed of \holfour
+  tactics minimized in Section~\ref{sec:proofdisplay}. And in 
+  Section~\ref{s:experiments}, we evaluate the system 
+  on a large set of theorems originating from \holfour.
+  And generated tactic-level proofs are analysed in 
+  Section~\ref{sec:case_study}.
+
+The section order does not follow the implementation. In this way, 
+we can give a description of the proof search algorithm sooner. This is 
+beneficial for the reader as it shows the essential for our work before 
+technical details on the prediction and recording algorithms.
+
 \section{Search Tree}\label{sec:prelim}
@@
 node selection starts from the root of the search tree. From the current node,
-the function \texttt{CompareChildren} chooses among the children of its open
+the function \texttt{CompareChildren} choses among the children of its open
 goal the one with the highest value.
@@
 
-\section{Feature Extraction}\label{sec:features}
+\section{Feature Extraction and Similarity Measure}\label{sec:features}
 %We explain the construction of the similarity measure by which relevant
@@
 
-\subsection{Similarity}\label{sec:predictions}
 We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$
@@
 the table by the tag ``default''.  This default strategy relies on a tactic
-time out of 0.05s, an orthogonalization radius of 20 and a policy coefficient of
+time out of 0.05s, an orthogonalization radius of 20 and a policy coefficent of
 0.5.
@@
 orthogonalization process is only beneficial when running \tactictoe for short
-period of time. It seems that the allotted time allows a strategy
+period of time. It seems that the alloted time allows a strategy
 running without orthogonalization to catch up.
@@
 contributes significantly as well. Experiments involving \eprover, run the ATP
-using a separate process asynchronously with a timeout of 5 seconds and 128 premises.
+using a separate proces asynchronously with a timeout of 5 seconds and 128 premises.
 The fact that \metis outperforms \eprover in this setting is due to the fact
@@
 The extracted proof script of $\mathfrak{T}$ is $P_{goal}(g_{root})$.
-We minimally improve it by substituting \texttt{THENL} by \texttt{THEN} when the
+We minimally improve it by subtituting \texttt{THENL} by \texttt{THEN} when the
 list of goals is a singleton and removing \texttt{THENL []} during the
@@
 
-\section{Case Study}
+\section{Case Study}\label{sec:case_study}
 Since the proofs generated by \tactictoe are meant to be parts of a \holfour development, it interesting to compare their quality with the original human proofs.
@@
 proceeds instead by rewriting the
-goal with the definitions then splitting the goal into multiple cases until the goal obligations are similar
+goal with the definitions then spliting the goal into multiple cases until the goal obligations are similar
 to \texttt{DIVIDES\_ANTISYM}. As expected, \tactictoe's proof

### 2018-03-19 tactictoe_jv.tex
Plan and diagram in the introduction

@@
 the length of the original human proofs.
-\item Minimization and prettification of the generated proof  facilitates user
+\item Minimization and embellishment of the generated proof  facilitates user
 interaction.
 \end{itemize}
-
-\todoi{graphic explaining how each part relate to each other}
 
@@
   At the basis of those predictions are 
-  similarity measures computed from syntactic features of mathematical formulas 
-  in Section~\ref{sec:features}.
+  similarity measures computed in Section~\ref{sec:features} from syntactic 
+  features of mathematical formulas .
   Modification of the knowledge base performed by orthogonalization 
@@
   improves the quality of the databases.
-  And the preselection method in Section~\ref{sec:presel} increases the speed 
+  And the preselection method described in Section~\ref{sec:presel} increases 
+  the speed 
   of the prediction algorithm.
   The learned predictors relying on this knowledge serves as
-  guidance for the proof search algorithm in Section~\ref{sec:proofsearch} 
-  which explores different proof paths in the proof tree in 
+  guidance for the proof search algorithm presented in 
+  Section~\ref{sec:proofsearch} 
+  which explores different proof paths in the proof tree defined in 
   Section~\ref{sec:prelim}. 
@@
 
-\section{Proof Presentation}\label{sec:proofdisplay}
+\section{Minimization and Embellishment}\label{sec:proofdisplay}
 
@@
 
-\paragraph{Prettification}
-Without prettification, the returned tactic is barely readable as it contains
+\paragraph{Embellishment}
+Without embellishment, the returned tactic is barely readable as it contains
 information to guarantees that each \sml subterm is interpreted in the way in
@@
 
-The total effect of these minimization and prettification techniques
+The total effect of these minimization and embellishment techniques
 on the readability of a proof script is depicted in
@@
 \begin{example}(Theorem \texttt{EVERY\_MAP})\label{ex:pretty}\\\
-Proof script before minimization and prettification:
+Before minimization and embellishment:
 \begin{lstlisting}[language=SMLSmall]
@@
 \end{lstlisting}
-Proof script after minimization and prettification:
+After minimization and embellishment:
 \begin{lstlisting}[language=SMLSmall]

### 2018-03-19 tactictoe_jv.tex
Abstract

@@
 We implement a tactical prover TacticToe on top of the HOL4 interactive theorem
-prover. TacticToe learns from human proofs which mathematical method is
-useful in a particular proof situation.
-This knowledge is then used in a Monte Carlo tree search algorithm that
-explores the most promising tactic-level proof paths.
+prover. TacticToe learns from human proofs which mathematical technique is
+the most suitable in each proof situation.
+This knowledge is then used by a Monte Carlo tree search algorithm to
+explore the most promising tactic-level proof paths.
 On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of

### 2018-03-20 tactictoe_jv.tex
integrating Josef's comments wip

@@
 \def\vampire{\textsf{Vampire}\xspace}
-\def\eprover{\textsf{E-prover}\xspace}
+\def\eprover{\textsf{Eprover}\xspace}
 \def\zthree{\textsf{z3}\xspace}
@@
 7164 theorems in HOL4's standard library whereas
-Eprover solves 34.5\%. The success rate rises to 69.0\% by combining the
-results of TacticToe and Eprover.
+Eprover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by 
+combining the results of TacticToe and Eprover.
 \end{abstract}
-
-\tableofcontents
 
@@
 A goal (or proof state) is represented as a \emph{sequent}.
-A sequent is composed of a set of assumptions and a conclusion, all of which are higher-order logic formulas.
+A sequent is composed of a set of assumptions and a conclusion, all of which 
+are higher-order logic formulas~\cite{gordon01hol}.
 When a goal is proven, it becomes a theorem.
-If the developer gives the theorem a name then we refer to it as a top-level theorem.
+If the developer gives the theorem a name then we refer to it as a 
+\textit{top-level} theorem.
 
@@
 of the original goal) together
-with a validation function.
-Calling the validation function on a list of \emph{theorems}, corresponding to
+with a \textit{validation function}.
+Calling the validation function on a list of \emph{theorems}, 
+corresponding to
 the list of subgoals, results in a theorem corresponding to the original goal.
-For example, if the list of subgoals is empty, then calling the validation
+For example, if the list of subgoals is empty, then calling the 
+validation
 function on the empty list yields the originally goal sequent as a theorem.
@@
 construction.
-Because validation functions are only important to check the final proof, we
+Because validation functions are only important to check the final 
+proof, we
 omit them during the description of the proof search algorithm. We denote by
@@
 \holfour standard library by running \tactictoe with a 5 seconds timeout.
-The additional contributions and their effect on our system are:
+The contributions of this extension and their effect on our system are:
 
 \begin{itemize}
-\item Proof recording at the tactic level is made more precise. Pattern
+\item Proof recording at the tactic level is made more precise in 
+Section~\ref{sec:recording}. Pattern
 matching constructions, opening of submodules are supported. The different
@@
 \item Monte Carlo tree search (MCTS) replaces A* as our
-search algorithm. The MCTS algorithm gives a more balanced feedback on
+search algorithm in Section~\ref{sec:proofsearch}. The MCTS algorithm gives a 
+more balanced feedback on
 the success of
@@
 through supervised learning.
-\item Proof guidance required by MCTS is given by predictors for
+\item Proof guidance required by MCTS is given by predictors presented in 
+Section~\ref{s:prediction} for
 three kinds of objects: tactics, theorems and lists of goals.
 \item The orthogonalization process that eliminates redundant tactics is
-described in a precise manner.
-\item A tactic abstraction mechanism is introduced. It enables us to create
+described in a precise manner in Section~\ref{sec:ortho}.
+\item A tactic abstraction mechanism is introduced in 
+Section~\ref{sec:synthesis}. It enables us to create
 more general and flexible tactics where tactic arguments are dynamically
@@
 \item The internal ATP \metis is complemented
-by asynchronous calls to \eprover to help \tactictoe during proof search.
+by asynchronous calls to \eprover in Section~\ref{sec:atp} to help \tactictoe 
+during proof search.
 \item Evaluation of \tactictoe with a 60 seconds timeout achieving a 66\%
-success rate on the standard library.
-\item First evaluation of \tactictoe success rate relative to
-the length of the original human proofs.
-\item Minimization and embellishment of the generated proof  facilitates user
-interaction.
+success rate on the standard library. And detailed comparisons between 
+\tactictoe and \eprover on the different type of problems are reported in 
+Section~\ref{s:experiments}.
+\item Minimization and embellishment of the discovered proof facilitates user
+interaction in Section~\ref{sec:proofdisplay}.
 \end{itemize}
@@
 \end{center}
-\label{fig:components}
-\caption{Relation between modules of the learning and proving tool chains.
-Functions are represented by rectangles and objects by ellipses.}
+\caption{\label{fig:components}Relation between modules of the learning and 
+proving tool chains. Functions are represented by rectangles and objects by 
+ellipses.}
 \end{figure}
@@
   At the basis of those predictions are 
-  similarity measures computed in Section~\ref{sec:features} from syntactic 
-  features of mathematical formulas .
+  similarity measures computed from syntactic features of mathematical formulas 
+  .
   Modification of the knowledge base performed by orthogonalization 
@@
   on a large set of theorems originating from \holfour.
-  And generated tactic-level proofs are analysed in 
+  The generated proofs are compared with human proofs in 
   Section~\ref{sec:case_study}.
@@
 The section order does not follow the implementation. In this way, 
-we can give a description of the proof search algorithm sooner. This is 
+we can give a description of the proof search algorithm sooner. This order is 
 beneficial for the reader as it shows the essential for our work before 
@@
 In order to keep track of the progress made so far from the initial conjecture
-we want to solve, we organize goals and tactics into a graph with list of goals
+we want to solve, we organize goals and tactics into a graph with lists of goals
 as nodes and tactics as edges (See Definition~\ref{def:stree}).
@@
 moment of the search after a certain amount of tactic applications.
-Exploration of the tree is directed by the MCTS algorithm in
+Construction of the tree is done by the MCTS algorithm in
 Section~\ref{sec:proofsearch}, which is guided by prediction algorithms
-presented in Section~\ref{sec:predictions}.
+presented in Section~\ref{s:prediction}.
 
@@
 \item $\mathbb{T}$ is a set of tactics, $\mathbb{G}$ is a set of goals
- and $\mathbb{A}$ is a set of nodes representing lists of goals.
+ and $\mathbb{A}$ is a set of nodes representing lists of goals. All objects 
+ are tagged with their position in the tree (see Figure~\ref{fig:choice}).
 \item $T$ is a function from $\mathbb{G}$ to $\mathcal{P}(\mathbb{T})$. It
@@
 $\mathbb{T},\mathbb{G},\mathbb{A}$ are equipped with an
-arbitrary order. In Figure~\ref{fig:choice}, a part of the a search tree is
-depicted, with goals represented by circles, nodes by rectangles and tactics by
-arrows.
+arbitrary \textit{total order}. In Figure~\ref{fig:choice}, a part of the a 
+search tree is depicted, with goals represented by circles, nodes by rectangles 
+and tactics by arrows.
 
@@
 \end{center}
-\caption{A node of a search tree $a_0$, the list of goals $g_0 \ldots g_n$ it
+\caption{\label{fig:choice}A node of a search tree $a_0$, the list of goals 
+$g_0 \ldots g_n$ it
 contains and the nodes $a_1 \ldots a_m$ derived from the application of the
 tactics $t_1 \ldots t_m$ to $g_i$.}
-\label{fig:choice}
 \end{figure}
@@
 pending goal)\\
-The set of solved nodes $\mathbb{A}^\infty$ and
-the set of solved goals $\mathbb{G}^\infty$ are defined inductively by:
+The set of \textit{solved nodes} $\mathbb{A}^*$ and
+the set of \textit{solved goals} $\mathbb{G}^*$ are defined inductively by:
 
@@
 \exists t \in T(g).\ A(g,t) \in \mathbb{A}^{n+1} \rbrace \\
-\mathbb{A}^\infty &=_{def} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \
-\mathbb{G}^\infty =_{def} \bigcup_{n \in \mathbb{N}} \mathbb{G}^n\\
+\mathbb{A}^* &=_{def} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \
+\mathbb{G}^* =_{def} \bigcup_{n \in \mathbb{N}} \mathbb{G}^n\\
 \end{align*}
 
-We call a goal (respectively node) unsolved if it is not in the set of solved
+We call a goal (respectively node) \textit{unsolved} if it is not in the set of 
+solved
 goals (respectively nodes).
-The open goal of a node is the first unsolved goal of this node according to
+The \textit{open goal} of a node is the first unsolved goal of this node 
+according to
 some preset order. The other unsolved nodes are called pending nodes.
@@
 \begin{definition}\label{def:desc}(children, descendant, ancestors)\\
-The children of a goal $g$ is a set of nodes defined by:
+The \textit{children of a goal} $g$ is a set of nodes defined by:
 \[\mathit{Children}(g) = \lbrace A(g,t)\in \mathbb{A}\ |\ t \in T(g) \rbrace\]
 
-By extension, we define the children of a node $a$ by:
+By extension, we define the \textit{children of a node} $a$ by:
 \[\mathit{Children}(a) = \bigcup_{g \in G(a)} \mathit{Children}(g) \]
 
-The descendants of $a$ by:
+The \textit{descendants of a node} $a$ by:
 \begin{align*}
@@
 
-And the ancestors of $a$ by:
+And the \textit{ancestors of a node} $a$ by:
 \[\mathit{Ancestors}(a) =_{def} \lbrace b \in \mathbb{A} \ | \ a \in
 Descendant(b) \rbrace\]
-
 \end{definition}
@@
 \begin{definition}\label{def:prod} (productive tactic)
-The application of a tactic on an open goal is called productive if and only if
+The application of a tactic $t$ on an open goal $g$ is called 
+\textit{productive} if and only if
 all the following conditions are satisfied:
@@
 in the ancestor of the node of the current goal.
-\item It is not a parallel step, i.e., its output is not a superset
-of a list of goals produced from the same goal.
+\item It is not a parallel step, i.e., there does not exist $t'$ in $T(g)$ 
+such as $t'(g) \subseteq t(g)$.
 \end{itemize}
@@
  tactics, theorems and lists of goals. The number of predicted objects $k$ is
- called the \emph{prediction radius}. For a goal $g$, we note
-$\mathit{Predictions}^{\mathit{tactic}}_k (g)$ (respectively
+ called the \emph{prediction radius}. For a goal $g$, we denote
+by $\mathit{Predictions}^{\mathit{tactic}}_k (g)$ (respectively
 $\mathit{Predictions}^{\mathit{theorem}}_k (g)$
@@
 similarity with a target goal $g$ (usually the first pending goal of a node).
-The recorded pairs induce an
+The recorded pairs and their similarity to $g$ induce an
 order on tactics. The intuition is that tactics which have been successful on
@@
 \paragraph{Lists of Goals as Output of Tactics}
-A dataset of tactic outputs is compiled during orthogonalization.
-This set is separated into positive and negative examples (see
+A dataset of tactic outputs is compiled during orthogonalization (see 
+Section~\ref{sec:ortho}).
+Some elements of this set will be considered positive examples (see 
 Section~\ref{sec:evaluation}).
-List of goals that are a subset of the result of a recorded tactic on a
-target goal are considered positive. During proof search, given a list of
+Lists of goals that are a subset or equal to the result of a recorded tactic on 
+a target goal are considered positive because if a set of goals is solvable 
+then a subset of it is too. During proof search, given a list of
 goals $l$ created by a tactic, we select a set of lists of goals that are most
@@
 prior evaluation for MCTS.
+
+\subsection{Feature Extraction and Similarity Measures}\label{sec:features}
+
+We predict tactics through their associated goals. So, the features
+for each kind of object can be extracted from (different representations of)
+mathematical formulas. We start by describing features for \holfour terms. From
+there, we extend the extraction mechanism
+to goals, theorems and lists of goals. Duplicated features are always removed
+so that each object has an associated set of features.
+
+Here are the different kinds of feature we extract from terms:
+\begin{itemize}
+\item names of constants, including the logical operators,
+\item type constructors present in the types of constants and variables,
+\item subterms with all variables replaced by a single placeholder $V$,
+\item names of the variables.
+\end{itemize}
+Goals (respectively theorems) are represented by pairs consisting of a list of
+terms (assumptions) and a term (conclusion). We distinguish between
+features of the assumptions and features of the conclusion by adding a
+different tag to elements of each set. The feature of a goal (respectively 
+theorem)
+are the union of all these tagged features. From
+that, we can construct features for a list of goals by computing the union of 
+the features of each goal in the list.
+
+We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$
+through their respective feature sets $f_1$ and $f_2$.
+The estimation is based on the frequency of the features in the intersection of
+$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features
+are, the more similar two goals should be. The relative rarity of each feature
+can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
+We additionally raise these weighs to the sixth power giving
+even more importance
+to rare features. This modification was found experimentally optimal in
+\cite{ckju-pxtp13}.
+
+The first co-distance $sim_1$ sums the weight of all shared
+features to compute the total score.
+In a second co-distance $sim_2$, we additionally take
+the total number of features into account, to reduce the seemingly unfair
+advantage of big
+feature sets in the first scoring function.
+\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap
+<f_1}{\text{tfidf}(f)^{6}}}\]
+\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
+f_1}{\text{tfidf}(f)^{6}}}}
+{ln (e + \mathit{card} f_0 + \mathit{card} f_1)}\]
+
+In our setting, making predictions for an object $o$ consists of sorting a set
+of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems
+and $sim_2$ for lists of goals. The reason why $sim_1$ is used for
+predicting tactics is that tactics effective on a large goal $g$ are often also
+suitable for a goal made of sub-formulas of $g$. The same monotonicity
+heuristic can justify the use of $sim_1$ for theorems. Indeed, in practice a
+large theorem is often a conjunction of theorems from the same domain.
+And if a conjunction contains a formula related to a problem, then the other
+conjuncts from the same domain may also help to solve it.
 
@@
 policy and evaluation are estimated by the simple supervised k-NN algorithm
-here instead of neural networks trained through reinforcement learning in
-AlphaGo Zero~\cite{silver2017mastering}.
+here instead of neural networks trained through reinforcement learning
+as e.g. in AlphaGo Zero~\cite{silver2017mastering}.
 A characteristic of the MCTS algorithm is that it offers a good trade-off
 between exploitation and exploration. This means that the algorithm searches
-deeper more promising branches and leaves enough time for the exploration
+deeper the more promising branches and leaves enough time for the exploration
 of less likely alternatives.
@@
 
-If the highest children value is higher than the widening policy, then the
+If the highest children value is higher than the \textit{widening policy} (see 
+Section~\ref{sec:policy}), then the
 selected child
@@
 
-The number $\mathit{Failure}(a_i)$ is the amount of failures that occurred
+The number $\mathit{Failure}(a_i)$ is the number of failures that occurred
 during node extension from descendants of $a_i$ .
@@
 \[\mathit{CurPolicy}(a_i) = \frac{1 +
-\mathit{Visit}(a_i))}{\sqrt{\mathit{Visit}(p))}}\]
+\mathit{Visit}(a_i)}{\sqrt{\mathit{Visit}(p)}}\]
 The policy is approximately the  percentage of time a node was visited. The
@@
 number of cores available to the user.
-
-\section{Feature Extraction and Similarity Measure}\label{sec:features}
-%We explain the construction of the similarity measure by which relevant
-%objects are predicted
-
-We predict tactics through their associated goals. So, the features
-for each kind of object can be extracted from (different representations of)
-mathematical formulas. We start by describing features for \holfour terms. From
-there, we extend the extraction mechanism
-to goals, theorems and lists of goals. Duplicated features are always removed
-so that each object has an associated set of features.
-
-Here are the different kinds of feature we extract from terms:
-\begin{itemize}
-\item names of constants, including the logical operators,
-\item type constructors present in the types of constants and variables,
-\item subterms with all variables replaced by a single placeholder $V$,
-\item names of the variables.
-\end{itemize}
-Goals (respectively theorems) are represented by pairs consisting of a list of
-terms (assumptions) and a term (conclusion). We distinguish between
-features of the assumptions and features of the conclusion by adding a
-different tag to elements of each set. The feature of a goal (respectively theorem)
-are the union of all these tagged features. From
-that, we can construct features for a list of goals by computing the union of the features of each goal in the list.
-
-We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$
-through their respective feature sets $f_1$ and $f_2$.
-The estimation is based on the frequency of the features in the intersection of
-$f_1$ and $f_2$. A good rule of thumb is that the rarer the shared features
-are, the more similar two goals should be. The relative rarity of each feature
-can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
-We additionally raise these weighs to the sixth power giving
-even more importance
-to rare features. This modification was found experimentally optimal in
-\cite{ckju-pxtp13}.
-
-The first co-distance $sim_1$ sums the weight of all shared
-features to compute the total score.
-In a second co-distance $sim_2$, we additionally take
-the total number of features into account, to reduce the seemingly unfair
-advantage of big
-feature sets in the first scoring function.
-\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap
-<f_1}{\text{tfidf}(f)^{6}}}\]
-\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
-f_1}{\text{tfidf}(f)^{6}}}}
-{ln (e + \mathit{card} f_0 + \mathit{card} f_1)}\]
-
-In our setting, making predictions for an object $o$ consists of sorting a set
-of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems
-and $sim_2$ for lists of goals. The reason why $sim_1$ is used for
-predicting tactics is that tactics effective on a large goal $g$ are often also
-suitable for a goal made of sub-formulas of $g$. The same monotonicity
-heuristic can justify the use of $sim_1$ for theorems. Indeed, in practice a
-large theorem is often a conjunction of theorems from the same domain.
-And if a conjunction contains a formula related to a problem, then the other
-conjuncts from the same domain may also help to solve it.
 
@@
 \end{tabular}
-\caption{Evaluation on 7164 top-level theorems of the \holfour standard library
-\label{tab:_param}}
+\caption{\label{tab:_param} Evaluation on 7164 top-level theorems of the 
+\holfour standard library
+}
 \end{table}
@@
 \end{tabular}
-\caption{\label{theories}Percentage (\%) of re-proved theorems in the theories
+\caption{\label{theories} Percentage (\%) of re-proved theorems in the theories
 \texttt{arithmetic}, \texttt{real}, \texttt{complex}, \texttt{measure},

### 2018-03-20 tactictoe_jv.tex
integrating Josef's comments

@@
 \def\vampire{\textsf{Vampire}\xspace}
-\def\eprover{\textsf{Eprover}\xspace}
+\def\eprover{\textsf{E~prover}\xspace}
 \def\zthree{\textsf{z3}\xspace}
@@
 Eprover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by 
-combining the results of TacticToe and Eprover.
+combining the results of TacticToe and E~prover.
 \end{abstract}
@@
 \subsection{The problem}
-An ITP is a development environment for the construction of formal proofs in which it is possible, but not common, to write a proof as a sequence of applications of primitive inference rules.
-The common approach, however, is to use and devise high-level tools and automation that abstract over useful ways of producing proof pieces.
-A specific, prevalent instance of this approach is the use of \emph{tactics} for \emph{backward} or goal-directed proof.
-Here, the ITP user operates on a proof state, initially the desired conjecture, and applies tactics that transform the proof state until it is solved (or until giving up).
-Each tactic performs a sound transformation of the proof state: essentially, it promises that if the resulting proof state can be solved, then so can the starting one.
-This gives rise to a machine learning problem: can we learn a mapping from a given proof state to the next tactic (or sequence of tactics) that will productively advance the proof towards a solution?
+An ITP is a development environment for the construction of formal proofs in 
+which it is possible, but not common, to write a proof as a sequence of 
+applications of primitive inference rules.
+The common approach, however, is to use and devise high-level tools and 
+automation that abstract over useful ways of producing proof pieces.
+A specific, prevalent instance of this approach is the use of \emph{tactics} 
+for \emph{backward} or goal-directed proof.
+Here, the ITP user operates on a proof state, initially the desired conjecture, 
+and applies tactics that transform the proof state until it is solved (or until 
+giving up).
+Each tactic performs a sound transformation of the proof state: essentially, it 
+promises that if the resulting proof state can be solved, then so can the 
+starting one.
+This gives rise to a machine learning problem: can we learn a mapping from a 
+given proof state to the next tactic (or sequence of tactics) that will 
+productively advance the proof towards a solution?
 
@@
 Section~\ref{sec:evaluation}).
-Lists of goals that are a subset or equal to the result of a recorded tactic on 
-a target goal are considered positive because if a set of goals is solvable 
-then a subset of it is too. During proof search, given a list of
+During proof search, given a list of
 goals $l$ created by a tactic, we select a set of lists of goals that are most
@@
 can be estimated by calculating its TF-IDF weight~\cite{Jones72astatistical}.
-We additionally raise these weighs to the sixth power giving
+We additionally raise these weights to the sixth power giving
 even more importance
 to rare features. This modification was found experimentally optimal in
-\cite{ckju-pxtp13}.
+\cite{ckju-pxtp13}. \todoi {Cezary: Josef says sixth power is not in this 
+citation}.
 
@@
 \[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap 
-<f_1}{\text{tfidf}(f)^{6}}}\]
+f_1}{\text{tfidf}(f)^{6}}}\]
 \[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
@@
 This occurs less than 10 times in the full-scale experiment.
-This happens only at the beginning of \holfour library, where the training data
+This happens only at the beginning of the \holfour library, where the training 
+data
 has very few tactics.
@@
 We now concentrate on the definition of a reasonable evaluation function for the
-output of tactics. Theoretically, a list of goals is interesting if short proofs
-are possible each the goals in the list, and the lengths of the proofs can be
-estimated from previous proof searches. However, this
+output of tactics. Theoretically, a list of goals is interesting if we can find 
+a short proof for it. We could estimate the length of a proof from previous 
+proof searches. However, this
 approach generates too much data and slows down the proof search. So, we only
-collect outputs of tactics tested during orthogonalization. We
-declare a list of goals $l$ to be positive if it is has been produced by the
-winner of an orthogonalization competition.
+collect outputs of tactics tested during orthogonalization (see 
+Section~\ref{sec:ortho}). We
+declare a list of goals $l$ to be \textit{positive} if it is has been produced 
+by the winner of an orthogonalization competition.
 The set of positive examples in
@@
 
-\holfour already integrates a superposition-based prover \metis for this purpose. It is already
+\holfour already integrates a superposition-based prover \metis~\cite{metis} 
+for this purpose. It is already
 recognized by the tactic selection mechanism in \tactictoe and its
@@
 goals. Since calls to ATPs are
-computationally expensive, they are run in parallel and asynchronously. This
+computationally expensive, they are run in parallel and asynchronously. To 
+update the MCTS values, the 
+result of each thread is back-propagated in the tree after completion. This
 avoids slowing down \tactictoe's search loop. The number of asynchronous calls
@@
 Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
-The dependencies $D_\infty$ of a tactic-goal pair $(t_0,g_0)$ are
+The dependencies $D^*$ of a tactic-goal pair $(t_0,g_0)$ are
 inductively defined by:
@@
 (t',g') \in D_n.\ g \in t'(g') \rbrace  \\
-D_\infty &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
+D^* &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
 \end{align*}
@@
 \begin{definition}(Dependencies of a theorem)\\
-Let $\mathfrak{T}$ be the set of recorded theorems.
+Let $\mathfrak{R}$ be the set of recorded theorems.
 The dependencies of a theorem $t$ are the set of top-level theorems in
-$\mathfrak{T}$ appearing in the proof of $t$. These dependencies are
+$\mathfrak{R}$ appearing in the proof of $t$. These dependencies are
 recorded by tracking the use of top-level theorems through the inference
@@
 this relation.
-Given the default subsumption on goals, $t_1$ subsumes $t_2$ on $g$ if and only
-if $t_1(g)$ is a subset (modulo $\alpha$-equivalence) of $t_2(g)$.
+%Given the default subsumption on goals, $t_1$ subsumes $t_2$ on $g$ if and only
+%if $t_1(g)$ is a subset (modulo $\alpha$-equivalence) of $t_2(g)$.
 \end{definition}
@@
 % All in all, 
-A database built with the orthogonalization process hus contains
+A database built with the orthogonalization process thus contains
 $(\mathbb{O}(t,g),g)$ instead of $(t,g)$.
@@
 try to abstract any of the
-\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a placeholder in $u$, we
-can substitute $u$ by $h$ to create an abstracted tactic.
+\sml subterms. Let $u$ be a \sml subterm of $t$, and $h$ a variable (playing 
+the role of a placeholder), we
+can replace $u$ by $h$ to create an abstracted tactic.
 We denote this intermediate tactic $t[h/u]$. By repeating the
@@
 to
-instantiate the placeholders inside $\hat{t}$. Because it is difficult to
+instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to
 build a general predictor effective on each type of argument, we manually
@@
 considered as a learning process. To ensure the global fairness, the final
-experiments in Section~\ref{sec:full_exp} runs the best strategy on the full
+experiment in Section~\ref{sec:full_exp} runs the best strategy on the full
 dataset which is about 30 times larger.
@@
 way that changes the behavior of a tactic used in the final proof. This has not
-occurred a single time in our experiments.
+occurred a single time in our experiments. And a tactical proof generated by 
+\tactictoe during the final experiment takes on average 0.37 seconds to replay.
 More details on how a proof is extracted from the search tree is given in
@@
 All our feature vectors have been learned form human proofs. We now can now
-also add tactic-goal pairs that appears in \tactictoe's final proof. To prevent
+also add tactic-goal pairs that appears in the last proof of \tactictoe. To 
+prevent
 duplication of effort, orthogonalization of those
@@
 follow a similar
-distribution. % This means that 
+distribution.
 If \tactictoe finds a proof, there is about a 50
@@
 
-In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
-see how well \tactictoe and \eprover cope with increased levels of
-difficulty. As expected, the longer the original proof is, the harder it is
+In Figure~\ref{fig:percentage}, we regroup problems by the length of their 
+human proof to measure how well \tactictoe and \eprover cope with increasing 
+levels 
+of difficulty. As expected, the longer the original proof is, the harder it is
 for \tactictoe and \eprover to re-prove the theorem on their own.
@@
 Therefore, we can guarantee that the behavior of recorded stateless tactics
-does not change. And because an update on a stateful tactic
-generally increases its strength, our prediction algorithm based on past
-examples is still reasonably effective in this case.
+does not change. Even with some stateful tactics, the prediction algorithm is  
+effective because most updates on the state of \holfour increase the 
+strength of stateful tactics.
 
@@
 local theorems.  We do not replace a local
-theorem by its definition because we want to avoid unfolding tactic proofs into
-other tactic proofs.
+theorem by its tactic proof (which its \sml definition) because we want to 
+avoid unfolding proofs inside other proofs.
 If the theorem is stored in the \holfour database available across \holfour

### 2018-03-21 tactictoe_jv.tex
Yutaka's comments

@@
   assumptions) and apply non-trivial proof transformations.
-  Formalized proofs take advantage of different levels of automation:
+  Formalized proofs take advantage of different levels of automation which are 
+  in increasing order of generality:
   specialized rules, theory-based strategies and general purpose strategies.
@@
   premise selection, translation and proof reconstruction.
-  Yet, ATPs are far from foolproof and more precise user-guidance, achieved
+  Yet, ATPs are not flawless and more precise user-guidance, achieved
   by applying a particular sequence of specialized rules, is almost always
@@
   from previous proofs. Backed by this machine-learned guidance, our prover
-  \tactictoe relies on \emph{Monte
-  Carlo tree search}~\cite{montecarlo} to find sequences of tactic applications
+  \tactictoe executes a \emph{Monte
+  Carlo tree search}~\cite{montecarlo} algorithm to find sequences of tactic 
+  applications
   leading to an ITP proof.
@@
 An ITP is a development environment for the construction of formal proofs in 
-which it is possible, but not common, to write a proof as a sequence of 
-applications of primitive inference rules.
+which it is possible to write a proof as a sequence of 
+applications of primitive inference rules. This approach is not preferred 
+by developers because proving a valuable theorem requires in general more than 
+thousand primitive inferences.
 The common approach, however, is to use and devise high-level tools and 
@@
 Here, the ITP user operates on a proof state, initially the desired conjecture, 
-and applies tactics that transform the proof state until it is solved (or until 
-giving up).
+and applies tactics that transform the proof state until it is solved.
 Each tactic performs a sound transformation of the proof state: essentially, it 
 promises that if the resulting proof state can be solved, then so can the 
-starting one.
-This gives rise to a machine learning problem: can we learn a mapping from a 
-given proof state to the next tactic (or sequence of tactics) that will 
-productively advance the proof towards a solution?
+starting one. This gives rise to a machine learning problem: can we learn a 
+mapping from a given proof state to the next tactic (or sequence of tactics) 
+that will productively advance the proof towards a solution?
 
@@
 When a goal is proven, it becomes a theorem.
-If the developer gives the theorem a name then we refer to it as a 
+And when the developer has given the theorem a name, we refer to it as a 
 \textit{top-level} theorem.
@@
 Formal developments in \holfour are organized into named theories, which are
-concretely implemented by \emph{script} files defining modules in the
+implemented by \emph{script} files defining modules in the
 functional programming language \sml.
@@
 \paragraph{Tactics in \holfour.}
-Considering concrete implementation details now, a \emph{tactic} is a function
+A \emph{tactic} is a function
 that takes a goal (or proof obligation) and returns a list of goals (subgoals
@@
 function on the empty list yields the originally goal sequent as a theorem.
-In this way, tactics implement the plumbing of backwards, goal-directed, proof
-construction.
-Because validation functions are only important to check the final 
+In this way, tactics implement the plumbing of goal-directed proof
+construction, transforming the conjecture by reasoning in a backward manner.
+Since validation functions are only important to check the final 
 proof, we
@@
 \subsection{Contributions}
-This paper extends the work described in \cite{tgckju-lpar17}.
+This paper extends the work described in our previous work~\cite{tgckju-lpar17}.
 There we proposed the idea of a tactic-based prover based on supervised
@@
 \holfour standard library by running \tactictoe with a 5 seconds timeout.
-The contributions of this extension and their effect on our system are:
+The contributions of this paper and their effect on our system are:
 
@@
 Section~\ref{sec:recording}. Pattern
-matching constructions, opening of submodules are supported. The different
-recording steps (parsing, unfolding and replaying) are regrouped into a single
-\sml function.
+matching constructions, opening of submodules are supported. 
+%The different
+%recording steps (parsing, unfolding and replaying) are regrouped into a single
+%\sml function.
 \item Monte Carlo tree search (MCTS) replaces A* as our
@@
 through supervised learning.
-\item Proof guidance required by MCTS is given by predictors presented in 
-Section~\ref{s:prediction} for
-three kinds of objects: tactics, theorems and lists of goals.
+\item Proof guidance required by MCTS is given by prediction algorithms 
+through supervised learning (see 
+Section~\ref{s:prediction}).
+These predictors are implemented for three kind of objects: tactics, 
+theorems and lists of goals.
 \item The orthogonalization process that eliminates redundant tactics is
@@
 during proof search.
-\item Evaluation of \tactictoe with a 60 seconds timeout achieving a 66\%
-success rate on the standard library. And detailed comparisons between 
-\tactictoe and \eprover on the different type of problems are reported in 
+\item Evaluation of \tactictoe with a 60 seconds timeout achieves a 66\%
+success rate on the standard library. And comparisons between 
+\tactictoe and \eprover on different type of problems are reported in 
 Section~\ref{s:experiments}.
@@
 Construction of the tree is done by the MCTS algorithm in
-Section~\ref{sec:proofsearch}, which is guided by prediction algorithms
-presented in Section~\ref{s:prediction}.
+Section~\ref{sec:proofsearch}. The MCTS algorithm is guided by prediction 
+algorithms presented in Section~\ref{s:prediction}.
 
 \begin{definition}\label{def:stree}(search tree)\\
-A search tree $\mathfrak{T}$ is a sixtuple
+A search tree $\mathfrak{T}$ is a sextuple
 $(\mathbb{T},\mathbb{G},\mathbb{A},T,G,A)$
@@
 arbitrary \textit{total order}. In Figure~\ref{fig:choice}, a part of the a 
-search tree is depicted, with goals represented by circles, nodes by rectangles 
-and tactics by arrows.
+search tree is depicted. Goals, nodes and tactics are represented 
+respectively by circles, rectangles and arrows.
 
@@
 
-We call a goal (respectively node) \textit{unsolved} if it is not in the set of 
-solved
-goals (respectively nodes).
+We call a goal \textit{unsolved} if it is not in the set 
+of solved goals. Similarly, we call a node \textit{unsolved} if it does not 
+belong to the set of solved nodes.
 The \textit{open goal} of a node is the first unsolved goal of this node 
 according to
-some preset order. The other unsolved nodes are called pending nodes.
+some preset order. The other unsolved nodes are called \textit{pending nodes}.
 If a node is unsolved, then it contains a unique open goal.
@@
 in the ancestor of the node of the current goal.
-\item It is not a parallel step, i.e., there does not exist $t'$ in $T(g)$ 
+\item It is not a redundant step, i.e., there does not exist $t'$ in $T(g)$ 
 such as $t'(g) \subseteq t(g)$.
@@
 The learning-based selection of relevant lemmas significantly improves the
-automation for hammers~\cite{BlanchetteGKKU16}. In \tactictoe we use the
+automation in hammer systems~\cite{BlanchetteGKKU16}. In \tactictoe we use the
 distance-weighted \emph{$k$ nearest-neighbour} classifier~\cite{DudaniS76}
@@
 predictions per second, while maintaining very high prediction quality~\cite{femalecop}.
-We rely on this supervised machine learning algorithm to predict three
-different kinds of objects:
- tactics, theorems and lists of goals. The number of predicted objects $k$ is
+We rely on this supervised machine learning algorithm to predict objects of 
+three kinds: tactics, theorems and lists of goals. The number of predicted 
+objects $k$ is
  called the \emph{prediction radius}. For a goal $g$, we denote
@@
 The recorded pairs and their similarity to $g$ induce an
-order on tactics. The intuition is that tactics which have been successful on
-goals similar to $g$ should be tried first, since they are more likely to lead
+order on tactics. Intuitively, tactics which have been successful on
+goals similar to $g$ should be tried first. Indeed, they are more likely to lead
 to a proof.
-This predicted tactic order is translated into a prior policy for our MCTS
+This predicted tactic order is then transformed into a prior policy of our MCTS
 algorithm (see Section~\ref{sec:policy}).
@@
 \paragraph{Lists of Goals as Output of Tactics}
-A dataset of tactic outputs is compiled during orthogonalization (see 
+We compile a dataset of tactic outputs during orthogonalization (see 
 Section~\ref{sec:ortho}).
@@
 We predict tactics through their associated goals. So, the features
-for each kind of object can be extracted from (different representations of)
+for each kind of object can be extracted from different representations of
 mathematical formulas. We start by describing features for \holfour terms. From
@@
 
-Here are the different kinds of feature we extract from terms:
+Here are the kinds of features we extract from terms:
 \begin{itemize}
@@
 
-We estimate the similarity (or co-distance) between two objects $o_1$ and $o_2$
+We estimate the similarity between two objects $o_1$ and $o_2$
 through their respective feature sets $f_1$ and $f_2$.
@@
 even more importance
-to rare features. This modification was found experimentally optimal in
-\cite{ckju-pxtp13}. \todoi {Cezary: Josef says sixth power is not in this 
+to rare features. The second and third authors found this modification 
+experimentally 
+optimal~\cite{ckju-pxtp13}. \todoi {Cezary: Josef says sixth 
+power is not 
+in this 
 citation}.
 
-The first co-distance $sim_1$ sums the weight of all shared
+The first similarity measure $sim_1$ sums the weight of all shared
 features to compute the total score.
-In a second co-distance $sim_2$, we additionally take
+In a second similarity measure $sim_2$, we additionally take
 the total number of features into account, to reduce the seemingly unfair
@@
 
-Despite the best efforts of the prediction algorithms, a selected tactic may
-not solve the current goal, proceed in the wrong direction
-or even loop. For that reason, predictions need to be accompanied by a
+Our prediction algorithms are not always accurate. 
+Therefore, a selected tactic may
+fail, proceed in the wrong direction or even loop. 
+For that reason, predictions need to be accompanied by a
 proof search mechanism that allows for backtracking and
@@
 algorithm does not need to rely on roll outs for evaluation. The
-policy and evaluation are estimated by the simple supervised k-NN algorithm
-here instead of neural networks trained through reinforcement learning
-as e.g. in AlphaGo Zero~\cite{silver2017mastering}.
+policy and evaluation are estimated by the simple supervised k-NN algorithm.
 A characteristic of the MCTS algorithm is that it offers a good trade-off
@@
 
+\begin{remark}
+Neural networks trained through reinforcement learning can be 
+very effective for approximating the policy and evaluation as demonstrated in 
+.e.g. AlphaGo Zero~\cite{silver2017mastering}. 
+But training neural networks is 
+computationally expensive and has not yet been 
+proven effective in our context. That is why we chose a simpler machine 
+learning model for our project.
+\end{remark}
+
 We first describe the proof search algorithm and explain later how to compute
@@
 
-The proof search starts by creating an initial search tree. The search tree
-then evolves by the repetitive applications of MCTS steps. A step in the main
-loop of MCTS consists of three parts: node selection, node extension, and
-backpropagation. And the decision to terminate the algorithm is taken during
-resolution.
+The proof search starts by creating a search tree during initialization. 
+The search tree then evolves by the repetitive applications of MCTS steps. A 
+step in the main loop of MCTS consists of three parts: node selection, node 
+extension, and backpropagation. And the decision to stop the loop is taken 
+by the resolution module.
 
@@
 Through node extension steps the search tree grows and the number of paths to
-explore increases. To decide which node to extend next, an evolving value for
-each node is given in Definition~\ref{def:value}. The algorithm that performs
+explore increases. To decide which node to extend next, the MCTS algorithm
+computes for each node a value (see Definition~\ref{def:value}) that changes
+after each MCTS step. 
+The algorithm that performs
 node selection starts from the root of the search tree. From the current node,
-the function \texttt{CompareChildren} chooses among the children of its open
-goal the one with the highest value.
+the function \texttt{CompareChildren} chooses the child node with the highest 
+value among the children of its open goal.
 
@@
 becomes the current node, otherwise the final result of node selection is the
-current node. The node selection algorithm can be written as the following
-pseudo-code:
+current node. The following pseudo-code illustrates our node selection 
+algorithm:
 
@@
 
-\begin{definition}\label{def:value}(Value)
+\begin{definition}\label{def:value}(Value)\\
 The value of the $i^{th}$ child $a_i$ of the open goal of a parent node $p$
@@
 
-The number $\mathit{Failure}(a_i)$ is the number of failures that occurred
-during node extension from descendants of $a_i$ .
 The current evaluation $CurEvaluation(a_i)$ is the average evaluation of
@@
 
+where the number $\mathit{Failure}(a_i)$ is the number of failures that occurred
+during node extension from descendants of $a_i$ .
+
 The exploration term is determined by the prior policy and the current policy.
@@
 \mathit{Visit}(a_i)}{\sqrt{\mathit{Visit}(p)}}\]
-The policy is approximately the  percentage of time a node was visited. The
-square root skews this probability to favor even more exploration of nodes with
-few visits.
-
-The coefficient $c_{exploration}$ is experimentally determined and adjusts the
-trade-off between exploitation and exploration.
+The policy can be seen as a skewed percentage of the number of times a node was 
+visited. The square root favors exploration of nodes with
+few visits. The coefficient $c_{exploration}$ is experimentally determined and 
+adjusts the trade-off between exploitation and exploration.
 \end{definition}
@@
 If $a$ is a solved node or the descendant of a solved node, then extending $a$
-is useless and the algorithm reports a failure for this MCTS step.
-%In our
-%experiments, filtering such nodes before node selection is detrimental.
-If $a$ is not $solved$, it applies the best untested tactic $t$ on the open
+would not be productive and the algorithm reports a failure for this MCTS step.
+If $a$ is not solved, it applies the best untested tactic $t$ on the open
 goal $g$ of this
-node according to the prediction order for $g$. If no such tactic exists or if
-$t$ is not productive, a failure is returned.
+node according to the prediction order for $g$. 
+If no such tactic exists or if $t$ is not productive, the algorithm returns a 
+failure.
 If node extension succeeds, a new node containing $t(g)$ is added to the
@@
 \item It saturates, i.e., there are no tactics to be applied to any open goal.
-This occurs less than 10 times in the full-scale experiment.
-This happens only at the beginning of the \holfour library, where the training 
+This occurs less than 0.1 percent of the time in the full-scale experiment.
+And this happens only at the beginning of the \holfour library, where the 
+training 
 data
@@
 Our MCTS algorithm relies on the guidance of two supervised learning methods.
-The prior policy decides the order in which tactic should be tried and the
-prior evaluation estimates if a branch is fruitful by analysing tactic
+The prior policy estimates the probability of success of a tactic $t$ on a 
+goal $g$ and associate it with the potential children that would be created by 
+the 
+application of $t$ to $g$. The prior evaluation judges if a branch is fruitful 
+by analyzing tactic
 outputs. Both priors influence the rates at which branches of the search tree
@@
 \paragraph{Prior policy}\label{sec:policy}
-Co-distances between goals produced by the tactic predictor are hard
-to translate into a prior policy, so we rely solely on the induced order of
-tactics to estimate the probability with which each tactic should be tried.
+Similarity between goals produced by the tactic predictor are hard
+to translate into a prior policy, so we rely solely on the prediction order
+to estimate the probability with which each tactic should be tried.
 Let $\mathfrak{T}$ be a search tree after a number of MCTS steps.
@@
 the list of $n$ productive tactics already
-applied to $g$ by their similarity score so that $T(g)= \lbrace
-t_0,\ldots,t_{n-1} \rbrace$.
-Let $c_{policy}$ be a constant heuristic that estimates how likely a tactic
-$t_i$ is to prove the goal.
-We can now calculate the prior policy of a child $a_i$ produced by the tactic
+applied to $g$ by their similarity score. We note the resulting list 
+$t_0,\ldots,t_{n-1}$.
+Let $c_{policy}$ be a constant heuristic optimized during our experiments.
+We calculate the prior policy of a child $a_i$ produced by the tactic
 $t_i$ by:
@@
 We now concentrate on the definition of a reasonable evaluation function for the
-output of tactics. Theoretically, a list of goals is interesting if we can find 
-a short proof for it. We could estimate the length of a proof from previous 
-proof searches. However, this
-approach generates too much data and slows down the proof search. So, we only
+output of tactics. Intuitively, a list of goals is worth exploring further if 
+we believe there exists a short proof for it. We could estimate the likelihood 
+of finding such a proof from 
+previous proof searches. However, extracting lists of goals from all proof 
+attempts creates too much data which slows down the prediction algorithm. 
+So, we only
 collect outputs of tactics tested during orthogonalization (see 
 Section~\ref{sec:ortho}). We
-declare a list of goals $l$ to be \textit{positive} if it is has been produced 
+declare a list of goals $l$ to be \textit{positive} if it has been produced 
 by the winner of an orthogonalization competition.
@@
 Since external automated reasoning techniques sometimes outperform the combined
-power of tactics, we would like to combine the \tactictoe search with
+power of tactics, we combine the \tactictoe search with
 general purpose provers such as the ones found in \holyhammer for
@@
 for this purpose. It is already
-recognized by the tactic selection mechanism in \tactictoe and its
-premises can be predicted dynamically when in an abstracted form. Nevertheless,
+recognized by the tactic selection mechanism in \tactictoe and thanks to 
+abstraction, its
+premises can be predicted dynamically. Nevertheless,
 we think that the performance of \tactictoe can be boosted by giving the ATP
-\metis a special status. This arrangement consists of always predicting
+\metis a special status among tactics. This arrangement consists of always 
+predicting
 premises for \metis, giving it a slightly higher timeout and
-trying it first on each open goal. These modifications only induce an
-linear overhead on the number nodes as \metis does not create
-new goals.
-
-The integration of an external prover is also possible. For this purpose we experiment
-with
-\eprover~\cite{eprover}.
-Namely, we developed a tactic which expects a list of premises as an
-argument
-and calls \eprover on the goal together with the premises translated to a first-order problem. If \eprover
+trying it first on each open goal. Since \metis does not create
+new goals, these modifications only induce an overhead that increases linearly 
+with the number of nodes.
+
+In the following, we present the implementation of asynchronous calls to 
+\eprover~\cite{eprover} during \tactictoe's proof search. 
+Other external ATPs can be integrated with \tactictoe in a similar manner.
+
+First, we develop a tactic which expects a list of premises as an
+argument and calls \eprover on the goal together with the premises translated 
+to a first-order problem. If \eprover
 succeeds on the problem, the premises used in the external proof are used to
 reconstruct the proof inside \holfour with \metis. Giving a special status
-to external
-provers is essential as external ATP calls do not appear in human
-proof scripts. For external ATPs, we use an even higher timeout (5 seconds) and
-a larger number of predicted premises (128 for \eprover) and also try them
-first on open
-goals. Since calls to ATPs are
+to \eprover is essential as \eprover calls do not appear in human
+proof scripts. Here, we use an even higher timeout (5 seconds) and
+a larger number of predicted premises (128). We also try \eprover
+first every time a new goal is created. Since calls to \eprover are
 computationally expensive, they are run in parallel and asynchronously. To 
@@
 avoids slowing down \tactictoe's search loop. The number of asynchronous calls
-to external provers that can be executed at the same time is limited by the
+to \eprover that can be executed at the same time is limited by the
 number of cores available to the user.
@@
 \end{definition}
-
-\begin{remark}
-The relation chosen for tactic-goal pairs is transitively closed, which is not
-the case for theorem dependencies.
-\end{remark}
 
@@
 Proof patching: \cite{RingerYLG18}
-
-\todoi{maybe the paragraph can be generally proof strategy languages including ltac and eisbach}
+\todoi{Cezary; comment on proof patching or remove the citation.}
+
+\todoi{Cezary; add ltac and eisbach or remove PSL if you think it is not worth 
+mentionning.}
 A similar effort for stronger automation is performed in \isabelle. Nagashima

### 2018-03-21 tactictoe_jv.tex
section 1.3

@@
   build efficient machine learning classifiers in Section~\ref{s:prediction}.
-  At the basis of those predictions are 
-  similarity measures computed from syntactic features of mathematical formulas 
-  .
-  Modification of the knowledge base performed by orthogonalization 
-  in Section~\ref{sec:ortho} and abstraction in Section~\ref{sec:synthesis}
-  improves the quality of the databases.
+  Modification of the knowledge base, shown as the training loop in 
+  Fig~\ref{fig:components} and
+  achieved by orthogonalization in Section~\ref{sec:ortho} and completed by 
+  an abstraction technique in Section~\ref{sec:synthesis},
+  improves the quality of the predictions.
   And the preselection method described in Section~\ref{sec:presel} increases 
-  the speed 
-  of the prediction algorithm.
-  The learned predictors relying on this knowledge serves as
-  guidance for the proof search algorithm presented in 
-  Section~\ref{sec:proofsearch} 
-  which explores different proof paths in the proof tree defined in 
-  Section~\ref{sec:prelim}. 
+  the speed of the prediction algorithm.
+  The learned predictors relying on this knowledge guide the proof search 
+  algorithm presented in Section~\ref{sec:proofsearch}.
+  The search algorithm explores different proof paths in the proof tree which 
+  is defined in Section~\ref{sec:prelim}. 
   The result, if
@@
 
-The section order does not follow the implementation. In this way, 
-we can give a description of the proof search algorithm sooner. This order is 
-beneficial for the reader as it shows the essential for our work before 
+The section order does not follow the implementation to give a description of 
+the proof search algorithm sooner. This order is 
+beneficial for the reader as it shows the essential of our work before 
 technical details on the prediction and recording algorithms.

### 2018-03-21 tactictoe_jv.tex
minor

@@
 
-The different components depicted in Fig~\ref{fig:components} give a 
+The different components depicted in Figure~\ref{fig:components} give a 
 high-level overview of the techniques necessary to generate 
@@
   build efficient machine learning classifiers in Section~\ref{s:prediction}.
-  Modification of the knowledge base, shown as the training loop in 
-  Fig~\ref{fig:components} and
+  Modification of the knowledge base (shown as the training loop in 
+  Figure~\ref{fig:components})
   achieved by orthogonalization in Section~\ref{sec:ortho} and completed by 
-  an abstraction technique in Section~\ref{sec:synthesis},
+  an abstraction technique in Section~\ref{sec:synthesis}
   improves the quality of the predictions.

### 2018-03-21 tactictoe_jv.tex
Fix Cezary's todos

@@
 even more importance
-to rare features. The second and third authors found this modification 
-experimentally 
-optimal~\cite{ckju-pxtp13}. \todoi {Cezary: Josef says sixth 
-power is not 
-in this 
-citation}.
+to rare features~\cite{BlanchetteGKKU16}.
 
@@
 %
-Proof patching: \cite{RingerYLG18}
-\todoi{Cezary; comment on proof patching or remove the citation.}
-
-\todoi{Cezary; add ltac and eisbach or remove PSL if you think it is not worth 
-mentionning.}
-A similar effort for stronger automation is performed in \isabelle. Nagashima
-developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the
-user to
-program with regular tactics and combine them with tools such as
-\sledgehammer.
-Selection of tactics from data collected in recorded proofs is in development
-but for now it is up to the user to come up with a good meta-tactic.
+Proof Patching \cite{RingerYLG18} is an approach that attempts to learn the
+proof changes necessary for a modification of a prover library or system. The
+approach is hard to apply for new proofs. On the other had, \tactictoe could also
+try to learn from different versions of the prover library.
+
+% \todoi{Cezary; add ltac and eisbach or remove PSL if you think it is not worth 
+% mentionning.}
+% A similar effort for stronger automation is performed in \isabelle. Nagashima
+% developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the
+% user to
+% program with regular tactics and combine them with tools such as
+% \sledgehammer.
+% Selection of tactics from data collected in recorded proofs is in development
+% but for now it is up to the user to come up with a good meta-tactic.
 

### 2018-03-21 tactictoe_jv.tex
typo
minor

@@
 %
-Proof Patching \cite{RingerYLG18} is an approach that attempts to learn the
+Proof patching \cite{RingerYLG18} is an approach that attempts to learn the
 proof changes necessary for a modification of a prover library or system. The
-approach is hard to apply for new proofs. On the other had, \tactictoe could also
-try to learn from different versions of the prover library.
+approach is hard to apply for new proofs. Following this idea, \tactictoe 
+could also try to learn from different versions of the prover library.
 

### 2018-03-21 tactictoe_jv.tex
Make minor edits in Section 1

@@
 applications of primitive inference rules. This approach is not preferred
-by developers because proving a valuable theorem requires in general more than 
-thousand primitive inferences.
-The common approach, however, is to use and devise high-level tools and 
+by developers because proving a valuable theorem requires often many
+thousands of primitive inferences.
+The preferred approach is to use and devise high-level tools and
 automation that abstract over useful ways of producing proof pieces.
@@
 promises that if the resulting proof state can be solved, then so can the
-starting one. This gives rise to a machine learning problem: can we learn a 
+initial one. This gives rise to a machine learning problem: can we learn a
 mapping from a given proof state to the next tactic (or sequence of tactics)
@@
 validation
-function on the empty list yields the originally goal sequent as a theorem.
+function on the empty list yields the original goal sequent as a theorem.
 In this way, tactics implement the plumbing of goal-directed proof
@@
 \subsection{Contributions}
-This paper extends the work described in our previous work~\cite{tgckju-lpar17}.
-There we proposed the idea of a tactic-based prover based on supervised
-learning guidance. And we achieved a 39\% success rate on theorems of the
-\holfour standard library by running \tactictoe with a 5 seconds timeout.
+This paper extends work described previously~\cite{tgckju-lpar17}, in which
+we proposed the idea of a tactic-based prover based on supervised
+learning guidance. We achieved a 39\% success rate on theorems of the
+\holfour standard library by running \tactictoe with a 5 second timeout.
 The contributions of this paper and their effect on our system are:
@@
 \begin{itemize}
-\item Proof recording at the tactic level is made more precise in 
-Section~\ref{sec:recording}. Pattern
-matching constructions, opening of submodules are supported. 
+\item Proof recording at the tactic level is made more precise
+  (Section~\ref{sec:recording}). We now support pattern
+matching constructions and opening SML modules.
 %The different
@@
 \item Monte Carlo tree search (MCTS) replaces A* as our
-search algorithm in Section~\ref{sec:proofsearch}. The MCTS algorithm gives a 
+  search algorithm (Section~\ref{sec:proofsearch}). The MCTS algorithm gives
 more balanced feedback on
@@
 each proof step by
-comparing subtrees of the search tree instead of leafs. The policy and
+comparing subtrees of the search tree instead of leaves. The policy and
 evaluation are learned
@@
 \item Proof guidance required by MCTS is given by prediction algorithms
-through supervised learning (see 
-Section~\ref{s:prediction}).
+through supervised learning (Section~\ref{s:prediction}).
 These predictors are implemented for three kind of objects: tactics,
 theorems and lists of goals.
-\item The orthogonalization process that eliminates redundant tactics is
-described in a precise manner in Section~\ref{sec:ortho}.
-\item A tactic abstraction mechanism is introduced in 
-Section~\ref{sec:synthesis}. It enables us to create
-more general and flexible tactics where tactic arguments are dynamically
-predicted.
+\item We introduce an orthogonalization process that eliminates redundant tactics (Section~\ref{sec:ortho}).
+\item We introduce a tactic abstraction mechanism
+  (Section~\ref{sec:synthesis}), which enables us to create
+more general and flexible tactics by dynamically
+predicting tactic arguments.
 \item The internal ATP \metis is complemented
-by asynchronous calls to \eprover in Section~\ref{sec:atp} to help \tactictoe 
+  by asynchronous calls to \eprover (Section~\ref{sec:atp}) to help \tactictoe
 during proof search.
 \item Evaluation of \tactictoe with a 60 seconds timeout achieves a 66\%
-success rate on the standard library. And comparisons between 
+success rate on the standard library. Comparisons between
 \tactictoe and \eprover on different type of problems are reported in
@@
 \item Minimization and embellishment of the discovered proof facilitates user
-interaction in Section~\ref{sec:proofdisplay}.
+  interaction (Section~\ref{sec:proofdisplay}).
 \end{itemize}
@@
 \caption{\label{fig:components}Relation between modules of the learning and
-proving tool chains. Functions are represented by rectangles and objects by 
+proving toolchains. Functions are represented by rectangles and objects by
 ellipses.}
@@
 of ITP proofs, we extract and record tactic invocations
-in Section~\ref{sec:recording}. Based on such training examples, we
-  build efficient machine learning classifiers in Section~\ref{s:prediction}.
-  Modification of the knowledge base (shown as the training loop in 
+(Section~\ref{sec:recording}). Based on such training examples, we
+  build efficient machine learning classifiers (Section~\ref{s:prediction}).
+  We then modify the knowledge base (shown as the training loop in
   Figure~\ref{fig:components})
-  achieved by orthogonalization in Section~\ref{sec:ortho} and completed by 
-  an abstraction technique in Section~\ref{sec:synthesis}
-  improves the quality of the predictions.
-  And the preselection method described in Section~\ref{sec:presel} increases 
+  using orthogonalization (Section~\ref{sec:ortho}) and
+  an abstraction technique (Section~\ref{sec:synthesis})
+  to improve the quality of the predictions.
+  Additionally, the preselection method described in Section~\ref{sec:presel} increases
   the speed of the prediction algorithm.
@@
   algorithm presented in Section~\ref{sec:proofsearch}.
-  The search algorithm explores different proof paths in the proof tree which 
+  The search algorithm explores different proof paths in the proof tree, which
   is defined in Section~\ref{sec:prelim}.
@@
   successful, is a verified human-level proof composed of \holfour
-  tactics minimized in Section~\ref{sec:proofdisplay}. And in 
+  tactics, minimized according to Section~\ref{sec:proofdisplay}.
+  In
   Section~\ref{s:experiments}, we evaluate the system
@@
 
-The section order does not follow the implementation to give a description of 
-the proof search algorithm sooner. This order is 
-beneficial for the reader as it shows the essential of our work before 
-technical details on the prediction and recording algorithms.
+The section order does not follow the implementation, because we have prioritized giving a description of
+the proof search algorithm early on. We hope this order
+benefits the reader by showing the essence of our work before
+the technical details of the prediction and recording algorithms.
 

### 2018-03-23 tactictoe_jv.tex
minor

@@
 On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of
-7164 theorems in HOL4's standard library whereas
+7164 theorems in the HOL4's standard library whereas
 Eprover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by
@@
 so that each object has an associated set of features.
-
 Here are the kinds of features we extract from terms:

### 2018-03-23 tactictoe_jv.tex
Make some edits

@@
 Moreover, it is possible to try multiple tactics on the same goal.
-In order to keep track of the progress made so far from the initial conjecture
-we want to solve, we organize goals and tactics into a graph with lists of goals
+In order to keep track of progress made from the initial conjecture, we organize goals and tactics into a graph with lists of goals
 as nodes and tactics as edges (See Definition~\ref{def:stree}).
@@
 In this section, we only give a description of the graph at a given
-moment of the search after a certain amount of tactic applications.
+moment of the search, after some number of tactic applications.
 Construction of the tree is done by the MCTS algorithm in
@@
 \mathbb{T}$ such as $t \in T(g)$ and
-returns a node $A(g,t)$ such as $t(g) = G(A(g,t))$. In other words, the output
+returns a node $A(g,t)$ such that $t(g) = G(A(g,t))$. In other words, the output
 $t(g)$ is exactly the list of goals contained in $A(g,t)$.
-\item It is acyclic, i.e. a node cannot be a strict descendant (see
+\item It is acyclic, i.e., a node cannot be a strict descendant (see
 Definition~\ref{def:desc}) of itself.
 \item There is exactly one node with no parents. This root node contains
-exactly one goal which is the initial goal (also called conjecture).
+exactly one goal which is the initial goal (conjecture).
 \end{itemize}
@@
 $\mathbb{T},\mathbb{G},\mathbb{A}$ are equipped with an
-arbitrary \textit{total order}. In Figure~\ref{fig:choice}, a part of the a
-search tree is depicted. Goals, nodes and tactics are represented
+arbitrary \textit{total order}. Figure~\ref{fig:choice} depicts part of a
+search tree. Goals, nodes and tactics are represented
 respectively by circles, rectangles and arrows.
@@
 search tree $\mathfrak{T}$. In Definition~\ref{def:solved}, we specify
-different status for a goal: open, pending or solved. There, we define
+different states for a goal: open, pending or solved. There, we define
 what a solved goal and solved node are by mutual recursion on the number of
@@
 \begin{align*}
-\mathbb{A}^{0} &=_{def}
+\mathbb{A}^{0} &=_{\mathrm{def}}
 \lbrace a \in \mathbb{A}\ |\ G(a) = \emptyset \rbrace \\
-\mathbb{G}^{0} &=_{def} \lbrace g \in \mathbb{G}\ |\
+\mathbb{G}^{0} &=_{\mathrm{def}} \lbrace g \in \mathbb{G}\ |\
 \exists t \in T(g).\ A(g,t) \in \mathbb{A}^{0} \rbrace\\
-\mathbb{A}^{n+1} &=_{def} \lbrace a \in \mathbb{A}\ |\
+\mathbb{A}^{n+1} &=_{\mathrm{def}} \lbrace a \in \mathbb{A}\ |\
 \forall g \in G(a).\ g \in \mathbb{G}^{n} \rbrace\\
-\mathbb{G}^{n+1} &=_{def} \lbrace g \in \mathbb{G}\ |\
+\mathbb{G}^{n+1} &=_{\mathrm{def}} \lbrace g \in \mathbb{G}\ |\
 \exists t \in T(g).\ A(g,t) \in \mathbb{A}^{n+1} \rbrace \\
-\mathbb{A}^* &=_{def} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \
-\mathbb{G}^* =_{def} \bigcup_{n \in \mathbb{N}} \mathbb{G}^n\\
+\mathbb{A}^* &=_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathbb{A}^n \ \ \ \ \
+\mathbb{G}^* =_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathbb{G}^n\\
 \end{align*}
@@
 \begin{align*}
-\mathit{Descendant}^{0}(a) &=_{def} \lbrace a \rbrace \\
-\mathit{Descendant}^{n+1}(a) &=_{def} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
+\mathit{Descendant}^{0}(a) &=_{\mathrm{def}} \lbrace a \rbrace \\
+\mathit{Descendant}^{n+1}(a) &=_{\mathrm{def}} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
 \mathit{Children}(a') \\
-\mathit{Descendant}(a) &=_{def} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
+\mathit{Descendant}(a) &=_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
 \end{align*}
@@
 And the \textit{ancestors of a node} $a$ by:
-\[\mathit{Ancestors}(a) =_{def} \lbrace b \in \mathbb{A} \ | \ a \in
+\[\mathit{Ancestors}(a) =_{\mathrm{def}} \lbrace b \in \mathbb{A} \ | \ a \in
 Descendant(b) \rbrace\]
@@
 is determined by:
-\[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{exploration} *
+  \[\mathit{Value}(a_i) = \mathit{CurEvaluation}(a_i) + c_{\mathit{exploration}} *
 \mathit{Exploration}(a_i)\]
@@
 \[\mathit{CurEvaluation}(a_i) =
-  \sum_{a' \in Descendant(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
+  \sum_{a' \in \mathit{Descendant}(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
   \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]
@@
 \begin{align*}
-D_0 &=_{def} \lbrace (t_0,g_0) \rbrace \\
-D_{n+1} &=_{def} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
+D_0 &=_{\mathrm{def}} \lbrace (t_0,g_0) \rbrace \\
+D_{n+1} &=_{\mathrm{def}} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
 (t',g') \in D_n.\ g \in t'(g') \rbrace  \\
-D^* &=_{def} \bigcup_{i \in \mathbb{N}} D_i\\
+D^* &=_{\mathrm{def}} \bigcup_{i \in \mathbb{N}} D_i\\
 \end{align*}
@@
 $\mathfrak{T}$. Expressing this in a formula, we get:
-  \[\textit{\mbox{Coverage}}(t) =_{def} card\ \lbrace g\ |\ (t,g) \in
+\[\textit{\mbox{Coverage}}(t) =_{\mathrm{def}} card\ \lbrace g\ |\ (t,g) \in
   \mathfrak{T} \rbrace \]
@@
 % could be:
-\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ has a proof with a length
+\[g_1 \le g_2  \Leftrightarrow_{\mathrm{def}} g_1 \mbox{ has a proof with a length
 shorter
@@
 By default, we however choose for efficiency reasons a minimal subsumption defined by:
-\[g_1 \le g_2  \Leftrightarrow_{def} g_1 \mbox{ is }\alpha\mbox{-equivalent
+\[g_1 \le g_2  \Leftrightarrow_{\mathrm{def}} g_1 \mbox{ is }\alpha\mbox{-equivalent
 to } g_2\]
@@
 subsumption $\le$ by:
-%\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_2 \in l_2.\ \exists g_1 \in l_1.\
+%\[l_1 \le l_2  \Leftrightarrow_{\mathrm{def}} \forall g_2 \in l_2.\ \exists g_1 \in l_1.\
 %g_1 \le g_2\]
-\[l_1 \le l_2  \Leftrightarrow_{def} \forall g_1 \in l_1.\ \exists g_2 \in l_2.\
+\[l_1 \le l_2  \Leftrightarrow_{\mathrm{def}} \forall g_1 \in l_1.\ \exists g_2 \in l_2.\
 g_1 \le g_2\]
@@
 
- \[t_1 \le_g t_2 \Leftrightarrow_{def} t_1(g) \le t_2(g)\]
+\[t_1 \le_g t_2 \Leftrightarrow_{\mathrm{def}} t_1(g) \le t_2(g)\]
 
@@
 \begin{align*}
-P_{goal}(g) &=_{def} T_{sol}(g)\ \texttt{THENL}\ P_{node}(A(g,T_{sol}(g)))\\
-P_{node}(a) &=_{def} [P_{goal}(g_1),\ldots,P_{goal}(g_n)]\ \ \ \text{with}\
+P_{goal}(g) &=_{\mathrm{def}} T_{sol}(g)\ \texttt{THENL}\ P_{node}(A(g,T_{sol}(g)))\\
+P_{node}(a) &=_{\mathrm{def}} [P_{goal}(g_1),\ldots,P_{goal}(g_n)]\ \ \ \text{with}\
 G(a) = g_1,\ldots,g_n\\

### 2018-03-23 tactictoe_jv.tex
Some more edits

@@
 \[\mathit{CurEvaluation}(a_i) =
-  \sum_{a' \in \mathit{Descendant}(a_i)} \frac{\mathit{PriorEvaluation}(a')} {card\
-  \mathit{Descendant}(a_i) + \mathit{Failure}(a_i)}\]
+  \sum_{a' \in \mathit{Descendant}(a_i)} \frac{\mathit{PriorEvaluation}(a')}
+  {|\mathit{Descendant}(a_i)| + \mathit{Failure}(a_i)}\]
 
@@
 visited. The square root favors exploration of nodes with
-few visits. The coefficient $c_{exploration}$ is experimentally determined and
+  few visits. The coefficient $c_{\mathit{exploration}}$ is experimentally determined and
 adjusts the trade-off between exploitation and exploration.
@@
 has very few tactics.
-\item It exceeds a given time limit (default is set to 60 seconds).
+\item It exceeds a given time limit (60 seconds by default).
 \end{itemize}
@@
 $t_0,\ldots,t_{n-1}$.
-Let $c_{policy}$ be a constant heuristic optimized during our experiments.
+Let $c_{\mathit{policy}}$ be a constant heuristic optimized during our experiments.
 We calculate the prior policy of a child $a_i$ produced by the tactic
 $t_i$ by:
-\[PriorPolicy(a_i) = (1 - c_{policy})^{i} * c_{policy}\]
+\[\mathit{PriorPolicy}(a_i) = (1 - c_{\mathit{policy}})^{i} * c_{\mathit{policy}}\]
 
@@
 
-\[WideningPolicy(p) = (1 - c_{policy})^{n} * c_{policy}\]
+\[\mathit{WideningPolicy}(p) = (1 - c_{\mathit{policy}})^{n} * c_{\mathit{policy}}\]
 
@@
 \mathit{PriorEvaluation}_k (a) &=
-  \frac{card\ \mathit{Positive}^{\mathit{goal\_list}}_k(G(a)}{k}\\
+  \frac{|\mathit{Positive}^{\mathit{goal\_list}}_k(G(a))|}{k}\\
 \end{align*}

### 2018-03-23 tactictoe_jv.tex
Fix some minor issues
Change tactic-goal to goal-tactic

@@
 \begin{align*}
-\mathit{Descendant}^{0}(a) &=_{\mathrm{def}} \lbrace a \rbrace \\
-\mathit{Descendant}^{n+1}(a) &=_{\mathrm{def}} \bigcup_{a' \in \mathit{Descendant}^{n}(a)}
+\mathit{Descendants}^{0}(a) &=_{\mathrm{def}} \lbrace a \rbrace \\
+\mathit{Descendants}^{n+1}(a) &=_{\mathrm{def}} \bigcup_{a' \in \mathit{Descendants}^{n}(a)}
 \mathit{Children}(a') \\
-\mathit{Descendant}(a) &=_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathit{Descendant}^n(a)\\
+\mathit{Descendants}(a) &=_{\mathrm{def}} \bigcup_{n \in \mathbb{N}} \mathit{Descendants}^n(a)\\
 \end{align*}
@@
 \[\mathit{Ancestors}(a) =_{\mathrm{def}} \lbrace b \in \mathbb{A} \ | \ a \in
-Descendant(b) \rbrace\]
+Descendants(b) \rbrace\]
 \end{definition}
@@
 To start with, we build a database of tactics consisting
-of tactic-goal pairs recorded from successful tactic applications in human
+of goal-tactic pairs recorded from successful tactic applications in human
 proofs (recording will be discussed in
@@
 During proof search, the recorded goals will be reordered according to their
-similarity with a target goal $g$ (usually the first pending goal of a node).
+similarity with a target goal $g$ (usually the open goal of a node).
 The recorded pairs and their similarity to $g$ induce an
@@
 \[\mathit{CurEvaluation}(a_i) =
-  \sum_{a' \in \mathit{Descendant}(a_i)} \frac{\mathit{PriorEvaluation}(a')}
-  {|\mathit{Descendant}(a_i)| + \mathit{Failure}(a_i)}\]
+  \sum_{a' \in \mathit{Descendants}(a_i)} \frac{\mathit{PriorEvaluation}(a')}
+  {|\mathit{Descendants}(a_i)| + \mathit{Failure}(a_i)}\]
 
@@
 In order to speed up the predictions during the search for a proof of a
-conjecture $c$, we preselect 500 tactic-goal pairs and 500 theorems and a
-larger number of lists of goals induced by the selection of tactic-goal pairs.
+conjecture $c$, we preselect 500 goal-tactic pairs and 500 theorems and a
+larger number of lists of goals induced by the selection of goal-tactic pairs.
 Preselected objects are the only objects available to \tactictoe's search
@@
 
-The first idea is to select tactic-goal pairs and theorems by their similarity
+The first idea is to select goal-tactic pairs and theorems by their similarity
 with $c$ (i.e. $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
@@
 
-\begin{definition}(Dependencies of a tactic-goal pair)\\
-Let $\mathfrak{F}$ be the set of recorded tactic-goal pairs.
-The dependencies $D^*$ of a tactic-goal pair $(t_0,g_0)$ are
+\begin{definition}(Dependencies of a goal-tactic pair)\\
+Let $\mathfrak{F}$ be the set of recorded goal-tactic pairs.
+The dependencies $D^*$ of a goal-tactic pair $(t_0,g_0)$ are
 inductively defined by:
@@
 during orthogonalization. The list $l$ is preselected if and only if $g$
-appears in the 500 preselected tactic-goal pairs.
+appears in the 500 preselected goal-tactic pairs.
 %This has the
@@
 To solve this problem, we modify the construction of the tactics database.
-Each time a new tactic-goal pair $(t,g)$ is extracted from a tactic proof and
+Each time a new goal-tactic pair $(t,g)$ is extracted from a tactic proof and
 about to be recorded, we consider if there does not already exist a better
@@
 the
-$k$ closest tactic-goal pairs $\mathit{Predictions}^{\mathit{tactic}}_k(g)$.
+$k$ closest goal-tactic pairs $\mathit{Predictions}^{\mathit{tactic}}_k(g)$.
 In our experiments,
@@
 original tactic on $g$ and that
-appears in the largest number of tactic-goal pairs in the database.
+appears in the largest number of goal-tactic pairs in the database.
 The winning tactic $w$ is then associated with $g$, producing the pair $(w,g)$
@@
 \begin{definition} (Coverage)\\
-Let $\mathfrak{T}$ be the database of tactic-goal pairs. We define the
+Let $\mathfrak{T}$ be the database of goal-tactic pairs. We define the
 coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times
@@
 All our feature vectors have been learned form human proofs. We now can now
-also add tactic-goal pairs that appears in the last proof of \tactictoe. To
+also add goal-tactic pairs that appears in the last proof of \tactictoe. To
 prevent
@@
 consists of 4 phases: proof script extraction, proof script globalization,
-tactic unit wrapping, and creation of tactic-goal pairs.
+tactic unit wrapping, and creation of goal-tactic pairs.
 
@@
 
-The function \texttt{save} writes the tactic-goal pair to disk increasing the
+The function \texttt{save} writes the goal-tactic pair to disk increasing the
 number of entries in our database of tactics. The function

### 2018-03-23 tactictoe_jv.tex
More formatting edits

@@
 
-The first similarity measure $sim_1$ sums the weight of all shared
+The first similarity measure $\mathit{sim}_1$ sums the weight of all shared
 features to compute the total score.
-In a second similarity measure $sim_2$, we additionally take
+In a second similarity measure $\mathit{sim}_2$, we additionally take
 the total number of features into account, to reduce the seemingly unfair
@@
 feature sets in the first scoring function.
-\[sim_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap
-f_1}{\text{tfidf}(f)^{6}}}\]
-\[sim_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
-f_1}{\text{tfidf}(f)^{6}}}}
-{ln (e + \mathit{card} f_0 + \mathit{card} f_1)}\]
+\[\mathit{sim}_1 (o_1, o_2) = {\sum\nolimits_{\,f \in f_0 \cap
+f_1}{\mathrm{tf\_idf}(f)^{6}}}\]
+\[\mathit{sim}_2 (o_1, o_2) = \frac{{\sum\nolimits_{\,f \in f_0 \cap
+f_1}{\mathrm{tf\_idf}(f)^{6}}}}
+{ln (e + |f_0| + |f_1|)}\]
 
 In our setting, making predictions for an object $o$ consists of sorting a set
-of objects by their similarity to $o$. We use $sim_1$ for tactics and theorems
-and $sim_2$ for lists of goals. The reason why $sim_1$ is used for
+of objects by their similarity to $o$. We use $\mathit{sim}_1$ for tactics and theorems
+and $\mathit{sim}_2$ for lists of goals. The reason why $\mathit{sim}_1$ is used for
 predicting tactics is that tactics effective on a large goal $g$ are often also
 suitable for a goal made of sub-formulas of $g$. The same monotonicity
-heuristic can justify the use of $sim_1$ for theorems. Indeed, in practice a
+heuristic can justify the use of $\mathit{sim}_1$ for theorems. Indeed, in practice a
 large theorem is often a conjunction of theorems from the same domain.
@@
 $\mathfrak{T}$. Expressing this in a formula, we get:
-\[\textit{\mbox{Coverage}}(t) =_{\mathrm{def}} card\ \lbrace g\ |\ (t,g) \in
-  \mathfrak{T} \rbrace \]
+\[\textit{\mbox{Coverage}}(t) =_{\mathrm{def}} |\lbrace g\ |\ (t,g) \in
+  \mathfrak{T} \rbrace| \]
 Intuitively, this notion estimates how general a tactic is by counting the

### 2018-03-23 tactictoe_jv.tex
Change terminology: proof scripts -> tactic proofs

@@
 the search returns a minimized and
-prettified proof script (see Section~\ref{sec:proofdisplay}).
+prettified tactic proof (see Section~\ref{sec:proofdisplay}).
 \item It saturates, i.e., there are no tactics to be applied to any open goal.
@@
 reconstruct the proof inside \holfour with \metis. Giving a special status
-to \eprover is essential as \eprover calls do not appear in human
-proof scripts. Here, we use an even higher timeout (5 seconds) and
+to \eprover is essential as \eprover calls do not appear in human-written
+tactic proofs. Here, we use an even higher timeout (5 seconds) and
 a larger number of predicted premises (128). We also try \eprover
@@
 For each theorem we first attempt to run the \tactictoe search with a time
-limit of 60 seconds before processing the original proof script.
+limit of 60 seconds before processing the original tactic proof.
 In this way, the fairness of the experiments is guaranteed by construction:
@@
 \midrule
-MCTS policy & $c_{policy} = 0.4$ && 149\\
-            & $c_{policy} = 0.5$ (default) && 156\\
-            & $c_{policy} = 0.6$ && 153\\
+  MCTS policy & $c_{\mathit{policy}} = 0.4$ && 149\\
+            & $c_{\mathit{policy}} = 0.5$ (default) && 156\\
+            & $c_{\mathit{policy}} = 0.6$ && 153\\
 \midrule
 MCTS evaluation & none (default + best abstraction) && 199\\
-				& $c_{exploration} = 1$ && 201\\
-				& $c_{exploration} = 2$ && 203\\
-				& $c_{exploration} = 4$ && 198\\
+        & $c_{\mathit{exploration}} = 1$ && 201\\
+        & $c_{\mathit{exploration}} = 2$ && 203\\
+        & $c_{\mathit{exploration}} = 4$ && 198\\
 \midrule
@@
 The hope is that the algorithm will improve faster by learning from its own
-discovered proofs than from the human proof
-scripts~\cite{DBLP:conf/cade/Urban07}. Side experiments show that
+discovered proofs than from the human-written proofs~\cite{DBLP:conf/cade/Urban07}. Side experiments show that
 this one shot reinforcement learning method increases \tactictoe's success rate
@@
 \end{tikzpicture}
-\caption{\label{fig:distrib}Number of proof scripts with $x$ tactic units.}
+\caption{\label{fig:distrib}Number of tactic proofs with $x$ tactic units.}
 \end{figure}
@@
 Rather than relying on the underlying programming language, we parse the
-original theory file containing proof scripts. This enables us to extract the
+original script file containing tactic proofs. This enables us to extract the
 code of each tactic.
@@
 Furthermore \tactictoe does not need to be installed to run
-proof scripts generated by \tactictoe. In this way, the further development of
+tactic proofs generated by \tactictoe. In this way, the further development of
 \tactictoe does not affect the robustness of \holfour.
@@
 modules are opened and local identifiers are defined. Therefore, it is unlikely
-that code extracted from proof scripts is interpretable in any other part of
+that code extracted from a tactic proof is interpretable in any other part of
 \holfour without any post-processing.
@@
 We describe in more detail our implementation of the recording algorithm. It
-consists of 4 phases: proof script extraction, proof script globalization,
+consists of 4 phases: tactic proof extraction, tactic proof globalization,
 tactic unit wrapping, and creation of goal-tactic pairs.
@@
 effect as the original. But since we want to extract information from this
-script in the form of tactics and their input goals, we modify it further.
+proof in the form of tactics and their input goals, we modify it further.
 In particular, we need to define at which level we should record the tactics in
@@
 dramatically increasing the number of tactics associated with a single goal.
-Imagine a proof script of the form \texttt{A THEN B THEN C}; then the tactics
+Imagine a tactic proof of the form \texttt{A THEN B THEN C}; then the tactics
 \texttt{A}, \texttt{A THEN B} and \texttt{A THEN B THEN C} would be valid
@@
 
-To produce the final recording script, we encapsulate each tactic unit
+To produce the final recording tactic proof, we encapsulate each tactic unit
 in a recording function \texttt{R} (see Example~\ref{ex:wrap}). In order to
@@
 tactics in all \holfour theories, we replace the original proof by the
-recording script in each theory and rebuild the \holfour library.
+recording proof in each theory and rebuild the \holfour library.
 
@@
 $\mathfrak{T}$ where the root node containing the initial goal is solved.
-In order to transform this search tree into a proof script, we need to extract
+In order to transform this search tree into a tactic proof, we need to extract
 the tactics that contributed to the proof and combine them using tactic
@@
 
-The extracted proof script of $\mathfrak{T}$ is $P_{goal}(g_{root})$.
+The extracted tactic proof of $\mathfrak{T}$ is $P_{goal}(g_{root})$.
 We minimally improve it by substituting \texttt{THENL} by \texttt{THEN} when the
@@
 the tactic \texttt{A THEN B} appears in the proof and has the same effect as
-\texttt{B} then we can replace \texttt{A THEN B} by \texttt{B} in the script.
+\texttt{B} then we can replace \texttt{A THEN B} by \texttt{B} in the proof.
 Optionally, stronger minimization can be obtained
@@
 module prefixes. If possible, we group all local declarations in the
-script under a single \texttt{let} binding at the start of the script. We
+proof under a single \texttt{let} binding at the start of the proof. We
 also replace extended terms by their quoted version.
@@
 The total effect of these minimization and embellishment techniques
-on the readability of a proof script is depicted in
+on the readability of a tactic proof is depicted in
 Example~\ref{ex:pretty}.
@@
 and custom built simplification sets.
-Thanks to the learning abilities of \tactictoe, the generated proof scripts
+Thanks to the learning abilities of \tactictoe, the generated tactic proofs
 often reveal the high-level structure of the proof. %For these reasons,

### 2018-03-27 tactictoe_jv.tex
minor line breaks
ldots in figure

@@
 \lstset{language=SML}
+
 \begin{document}
@@
 
-The section order does not follow the implementation, because we have prioritized giving a description of
+The section order does not follow the implementation, because we have 
+prioritized giving a description of
 the proof search algorithm early on. We hope this order
@@
 \node [tcircle] (0) {$g_i$};
-\node [right of=0] (0r) {.\ .\ .};
+\node [right of=0] (0r) {...};
 \node [tcircle,right of=0r] (0rr) {$g_n$};
-\node [left of=0] (0l) {.\ .\ .};
+\node [left of=0] (0l) {...};
 \node [tcircle,left of=0l] (0ll) {$g_0$};
@@
 \node [left of=0ll] {$a_0$};
-\node [above of=0,node distance=2cm] (2) {.\ .\ .};
+\node [above of=0,node distance=2cm] (2) {...};
 \node [left of=2, tcircle] (2l) {$\phantom{g_3}$};
@@
 \node [tcircle, left of=2l, node distance=3cm] (1r) {$\phantom{g_3}$};
-\node [left of=1r] (1) {.\ .\ .};
+\node [left of=1r] (1) {...};
 \node [tcircle, left of=1] (1l) {$\phantom{g_3}$};
@@
 \node [right of=2r,node distance=3cm] (3l) {$\phantom{g_3}$};
-\node [right of=3l] (3) {\phantom{.\ .\ .}};
+\node [right of=3l] (3) {\phantom{...}};
 \node [right of=3] (3r) {$\phantom{g_3}$};
@@
 It is difficult (probably impossible) to transfer \sml values
-between theories, since they are not run in the same \holfour session. In contrast,
+between theories, since they are not run in the same \holfour session. In 
+contrast,
 \sml code can be exported and imported.
@@
 
-The first line of this script file opens modules (called structures in \sml). Each of
-the modules contains a list of global identifiers which become directly accessible in
+The first line of this script file opens modules (called structures in \sml). 
+Each of
+the modules contains a list of global identifiers which become directly 
+accessible in
 the rest of the script.
@@
 
-We first parse the script file to extract tactic proofs. Each of them is found in
+We first parse the script file to extract tactic proofs. Each of them is found 
+in
 the third argument of a call to \texttt{store\_thm}. The result of tactic proof
@@
 advice for something close to their common input goal. The tactic
-\texttt{A THEN B THEN C} is likely to be too specific. In contrast, we can consider
+\texttt{A THEN B THEN C} is likely to be too specific. In contrast, we can 
+consider
 the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A}
@@
 \begin{definition}(Tactic unit)
-A tactic unit is an \sml expression of type \texttt{tactic} that does not contain an
+A tactic unit is an \sml expression of type \texttt{tactic} that does not 
+contain an
 infix operator at its root.
@@
 Because such tactic units are constructed from visual information present in
-the tactic proof, they often represent what a human user considers to be a single step of
+the tactic proof, they often represent what a human user considers to be a 
+single step of
 the proof.
@@
 \section{Minimization and Embellishment}\label{sec:proofdisplay}
-
 When \tactictoe finds a proof of an initial goal, it returns a search tree

### 2018-03-27 tactictoe_jv.tex
minor

@@
 does not provide a significant improvement over a strategy relying on no
-evaluation. The minor improvement seen in the table is most likely due to some
-variability in our experiments.
-The main reason is that our predictor learning abilities is
-limited and we are using the provability of lists of goals for evaluation. A
-more accurate evaluation could be based on an estimation of the length of the
-proof required to close a list of goals.
+evaluation. The main reason is that our predictor learning abilities is
+limited and we are using an estimate of the provability of lists of goals for 
+evaluation. A more accurate evaluation could be based on an estimation of the 
+length of the proof required to close a list of goals.
 

### 2018-03-28 presentation/presentation.tex
presentation

\documentclass{beamer}
\usepackage{etex}
\usepackage{booktabs}
\usepackage{xspace}

\usepackage{ifthen}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}

\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{headline}{}
\setbeamertemplate{footline}
{\hfill\insertframenumber{} /\inserttotalframenumber\hspace*{1ex}}

% Define my settings
\definecolor{mygreen}{cmyk}{0.82,0.11,1,0.25}
\setbeamercolor{structure}{fg=blue}
%packages indispensables

%packages utiles
\usepackage{alltt} %program code
\usepackage{amssymb} %lettres mathématiques
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{bussproofs} %derivation
\usepackage{hyperref} %to write path.
\usepackage{color} % colouring text
\usepackage{tabularx} % table
\usepackage{graphicx}
\usepackage{enumerate}

\def\cakeml{\textsf{CakeML}\xspace}
\def\holfour{\textsf{HOL4}\xspace}
\def\isabellehol{\textsf{Isabelle/HOL}\xspace}
\def\hollight{\textsf{HOL Light}\xspace}
\def\coq{\textsf{Coq}\xspace}
\def\ocaml{\textsf{OCaml}\xspace}
\def\sml{\textsf{SML}\xspace}
\def\holyhammer{\textsf{HOL(y)Hammer}\xspace}
\def\vampire{\textsf{Vampire}\xspace}
\def\eprover{\textsf{E~prover}\xspace}
\def\zthree{\textsf{Z3}\xspace}
\def\metis{\textsf{Metis}\xspace}
\def\o{\scriptsize{$\emptyset$}}
\def\tactictoe{\textsf{TacticToe}\xspace}
% GRAPHICS
\usepackage{graphicx} % margin definition
\usepackage{pgfplots} % graphics
\usepackage{tikz} % graphics
% graphs
\usetikzlibrary{arrows,shapes,calc}
\usetikzlibrary{trees,positioning,fit}
\tikzstyle{arrow}=[draw,-to,thick]
\tikzstyle{bluearrow}=[draw,-to,thick,blue]
\tikzstyle{tcircle} = [circle, draw, fill=white]
\tikzstyle{bcircle} = [circle, draw, blue, fill=blue]
\tikzstyle{gcircle} = [circle, draw, mygreen, fill=mygreen]

\tikzstyle{block} =
  [rectangle, draw,
   minimum width=9em, text centered, rounded corners, minimum height=2.0em]
\tikzstyle{line}=[draw]
\tikzstyle{cloud} =
  [draw, text centered, ellipse, minimum height=2.0em, minimum width=9em]

\title{Learning to Prove with Tactics} 
\author{Thibault Gauthier, 
Cezary Kaliszyk, Josef Urban, Ramana~Kumar, Michael Norrish}

\usepackage[final]{listings}
%\usepackage[scaled=.95]{couriers}
\lstdefinelanguage{SMLSmall}%
{columns=fixed,
  keywords={THEN,THENL,val,open,store_thm,fun,fn,let,in,end,true,
  while,do,if,then,else,break,return,;},%
  frame=none,
  sensitive=true,
  keywordstyle=\fontfamily{lmss}\scriptsize\selectfont,%
  basicstyle=\fontfamily{pcr}\small\selectfont,%
  stringstyle=\tt,
  morestring=[b]",
  literate=
   {=}{{\tt\raisebox{-.15mm}{=}}}1%
   {[}{{\tt\raisebox{-.15mm}{[}}}1%
   {]}{{\tt\raisebox{-.15mm}{]}}}1%
   {->}{{$\rightarrow$}}1%
   {wedge}{{$\wedge$}}1%
   {vee}{{$\vee$}}1%
   {==>}{{$\Rightarrow$}}1%
   {<==>}{{$\Leftrightarrow$}}1%
   {vee}{{$\vee$}}1%
   {<=}{{$\leq$}}1%
   {>=}{{$\geq$}}1%
   {'a}{{$\alpha$}}1%
   {'b}{{$\beta$}}1%
   {ldots}{$\ldots$}1%
   {emptyset}{$\emptyset$}1%
   {!}{$\forall$}1%
   {fff}{{DB.fetch}}1%
   {``}{\hspace{-.5mm}`\hspace{-1mm}`}1%
  }
\lstset{language=SMLSmall}
\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Problem}

Can we formally prove mathematical formulas automatically?

\end{frame}

\begin{frame}{Our solution}

\begin{center}
\begin{tikzpicture}[node distance = 1.5cm]
  \node [] (aa) {\textbf{Supervised learning}};
  \only<1> 
  {\node [cloud, below of=aa, node distance = 1cm] (a) {Formal 
  library};}
  \only<2> 
  {\node [cloud, below of=aa, node distance = 1cm, blue] (a) {Formal 
  library};}

  \node [block, below of=a] (b) {Proof recording};
  \node [cloud, below of=b] (c) {Knowledge base};
  \node [block, below of=c] (e) {Training};
  \node [block, below of=e] (d) {Predictors};

  \draw [-to,black,thick] (a) to (b);
  \draw [-to,black,thick] (b.south) to (c.north);
  \draw [-to,black,thick] (c) to [bend left]  (e);
  \draw [-to,black,thick] (e) to [bend left]  (c);
  \draw [-to,black,thick] (c.east) to [out=330,in=30] (d);

  \node [right of=aa, node distance=6.2cm] (0) {\textbf{Proving}};
  \node [cloud, below of=0, node distance = 1cm] (1) {Conjecture};
  \node [cloud, below of=1, node distance = 1.3cm] (2) {Search tree};
  \node [block, below of=2, node distance=1.1cm] (3) {Tactic policy};
  \node [above of=2, xshift=-16mm, node distance = 0.7cm] (00) {Proof
  search};
  \node [block, below of=3, node distance = 1.3cm] (4) {Tactic evaluation};
  \node [block, fit=(2) (3) (4)] (5) {};
  \node [block, below of=4, node distance = 1.3cm] (6) {Proof minimization};
  \node [cloud, below of=6, node distance = 1.3cm] (7) {Proof};
  \draw [-to,black,thick] (1) to (5);
  \draw [-to,black,thick] (5) to (6);
  \draw [-to,black,thick] (6) to (7);
  \draw [-to,black,thick] (d) to [out=0, in=180] (3);
  \draw [-to,black,thick] (d) to [out=0, in=180] (4);
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}{Formal library: reasoning with inference rules}

 \begin{center}
 	\begin{tikzpicture}[node distance=3cm]
 	\node [gcircle] (1) {};
 	\node [gcircle,right of=1] (2) {};
 	\node [gcircle,right of=2] (3) {};
    \node[right of=2,node distance=1.5cm] (m23) {};
    \uncover<2-3>{\node [tcircle, below of=m23,node distance=2cm] (4) {};
                  \draw[-to,thick] (3) to [out=270,in=90] (4);
                  \draw[-to,thick] (2) to [out=270,in=90] (4);
                 }
    \node [bcircle, below of=2,node distance=4cm] (5) {};
    \uncover<3-3>{
    	\draw[-to,thick] (1) to [out=270,in=90] (5);
    	\draw[-to,thick] (4) to [out=270,in=90] (5);
    }

 	\end{tikzpicture}
 \end{center}

\begin{tikzpicture} \node[gcircle] {}; \end{tikzpicture} axiom \\
\begin{tikzpicture} \node[bcircle] {}; \end{tikzpicture} conjecture \\
$\rightarrow$ rule\\
\begin{tikzpicture} \node[tcircle] {}; \end{tikzpicture} lemma

\end{frame}

\begin{frame}{Formal library: reasoning with tactics}
 \begin{center}
 	\begin{tikzpicture}[node distance=3cm]
 	\node [gcircle] (1) {};
 	\node [gcircle,right of=1] (2) {};
 	\node [gcircle,right of=2] (3) {};
 	\node[right of=2,node distance=1.5cm] (m23) {};
 	\uncover<2-3>{\node [tcircle, below of=m23,node distance=2cm] (4) {};}
 	\uncover<3-3>{
      \draw[-to,thick] (4) to [in=270,out=90] (3);
 	  \draw[-to,thick] (4) to [in=270,out=90] (2);
 	}
 	\node [bcircle, below of=2,node distance=4cm] (5) {};
 	\uncover<2-3>{
 		\draw[-to,thick] (5) to [in=270,out=90] (1);
 		\draw[-to,thick] (5) to [in=270,out=90] (4);
 	}
 	\end{tikzpicture}
  \end{center}

\vspace{5mm}

\begin{tikzpicture} \node[gcircle] {}; \end{tikzpicture} axiom\\
\begin{tikzpicture} \node[bcircle] {}; \end{tikzpicture} conjecture\\
$\rightarrow$ tactic\\
\begin{tikzpicture} \node[tcircle] {}; \end{tikzpicture} goal\\

\end{frame}

\begin{frame}{Formal library: tactics}

\texttt{REWRITE\_TAC}

\vspace{5mm}

\texttt{INDUCT\_TAC}

\vspace{5mm}

\texttt{METIS\_TAC}

\end{frame}

\begin{frame}{Formal library: composing tactics}
\centering
\texttt{THENL} tactical composes the effect of tactics.\\
\vspace{5mm}

\begin{tikzpicture}[node distance=3cm]
\uncover<3->{\node [tcircle] (1) {};}
\uncover<3->{\node [right of=1] (2) {};}

\node [right of=1,node distance=1.5cm] (m12) {};
\uncover<5->{\node [tcircle,right of=2,node distance=2cm] (l1) {};}

\uncover<2->{
\node [tcircle, below of=1,node distance=2cm] (b1) {};
\node [tcircle, below of=2,node distance=2cm] (b2) {};
}
\uncover<3->{
\draw[-to,thick] (b1) to node [xshift=10mm,yshift=-2mm] {\scriptsize 
\texttt{REWRITE\_TAC}} (1);
\draw[-to,thick] (b2) to node [xshift=-8mm,yshift=2mm] {\scriptsize 
\texttt{METIS\_TAC}} (2);
}
\node [bcircle, below of=m12,node distance=4cm] (cj1) {};
\uncover<4->{\node [bcircle, below of=l1,node distance=4cm] (cj2) {};}
\uncover<5->{
\draw[-to,thick] (cj2) to (l1);
\node [below of=l1,node distance=2cm,yshift=3mm,xshift=20mm] () 
{\scriptsize \texttt{INDUCT\_TAC}};
\node [below of=l1,node distance=2cm,yshift=0mm,xshift=20mm] () 
{\scriptsize \texttt{THENL}};
\node [below of=l1,node distance=2cm,yshift=-3mm,xshift=20mm] () 
{\scriptsize \texttt{[REWRITE\_TAC,METIS\_TAC]}};
}

\uncover<2->{
\node [below of=m12,node distance=2.5cm] () {\scriptsize \texttt{INDUCT\_TAC}};
\draw[-to,thick] (cj1) to [out=90,in=270] (b1);
\draw[-to,thick] (cj1) to [out=90,in=270] (b2);
}
\end{tikzpicture}
\end{frame}

\begin{frame}{}
\centering \Large Demo
\end{frame}

\begin{frame}{Plan}
\begin{center}
\begin{tikzpicture}[node distance = 1.5cm]
  \node [] (aa) {\textbf{Supervised learning}};
  \node [cloud, below of=aa, node distance = 1cm] (a) {Formal library};
  \node [block, blue, below of=a] (b) {Proof recording};
  \node [cloud, below of=b] (c) {Knowledge base};
  \node [block, below of=c] (e) {Training};
  \node [block, below of=e] (d) {Predictors};

  \draw [-to,black,thick] (a) to (b);
  \draw [-to,black,thick] (b.south) to (c.north);
  \draw [-to,black,thick] (c) to [bend left]  (e);
  \draw [-to,black,thick] (e) to [bend left]  (c);
  \draw [-to,black,thick] (c.east) to [out=330,in=30] (d);

  \node [right of=aa, node distance=6.2cm] (0) {\textbf{Proving}};
  \node [cloud, below of=0, node distance = 1cm] (1) {Conjecture};
  \node [cloud, below of=1, node distance = 1.3cm] (2) {Search tree};
  \node [block, below of=2, node distance=1.1cm] (3) {Tactic policy};
  \node [above of=2, xshift=-16mm, node distance = 0.7cm] (00) {Proof
  search};
  \node [block, below of=3, node distance = 1.3cm] (4) {Tactic 
  evaluation};
  \node [block, fit=(2) (3) (4)] (5) {};
  \node [block, below of=4, node distance = 1.3cm] (6) {Proof minimization};
  \node [cloud, below of=6, node distance = 1.3cm] (7) {Proof};
  \draw [-to,black,thick] (1) to (5);
  \draw [-to,black,thick] (5) to (6);
  \draw [-to,black,thick] (6) to (7);
  \draw [-to,black,thick] (d) to [out=0, in=180] (3);
  \draw [-to,black,thick] (d) to [out=0, in=180] (4);
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}[fragile]{Recording}
Original proof:
\begin{lstlisting}[language=SMLSmall]
INDUCT_TAC THENL [REWRITE_TAC, METIS_TAC]
\end{lstlisting}

Modified proof:
\begin{lstlisting}[language=SMLSmall]
(R numLib.INDUCT_TAC) THENL 
  [R boolLib.REWRITE_TAC, R metisLib.METIS_TAC]
\end{lstlisting}

Database of tactics:
\begin{lstlisting}[language=SMLSmall]
R (f n) (f (SUC n)) ==> transitive R: INDUCT_TAC
n * m <= n * p ==> (n = 0) vee m <= p   : REWRITE_TAC
INJ f U(:num) s ==> INFINITE s      : METIS_TAC
ldots
\end{lstlisting}
\end{frame}

\begin{frame}{Plan}
\begin{center}
\begin{tikzpicture}[node distance = 1.5cm]
  \node [] (aa) {\textbf{Supervised learning}};
  \node [cloud, below of=aa, node distance = 1cm] (a) {Formal library};
  \node [block, below of=a] (b) {Proof recording};
  \node [cloud, below of=b] (c) {Knowledge base};
  \node [block, below of=c] (e) {Training};
  \node [block, blue, below of=e] (d) {Predictors};

  \draw [-to,black,thick] (a) to (b);
  \draw [-to,black,thick] (b.south) to (c.north);
  \draw [-to,black,thick] (c) to [bend left]  (e);
  \draw [-to,black,thick] (e) to [bend left]  (c);
  \draw [-to,black,thick] (c.east) to [out=330,in=30] (d);

  \node [right of=aa, node distance=6.2cm] (0) {\textbf{Proving}};
  \node [cloud, below of=0, node distance = 1cm] (1) {Conjecture};
  \node [cloud, below of=1, node distance = 1.3cm] (2) {Search tree};
  \node [block, below of=2, blue, node distance=1.1cm] (3) {Tactic policy};
  \node [above of=2, xshift=-16mm, node distance = 0.7cm] (00) {Proof
  search};
  \node [block, below of=3, blue, node distance = 1.3cm] (4) {Tactic 
  evaluation};
  \node [block, fit=(2) (3) (4)] (5) {};
  \node [block, below of=4, node distance = 1.3cm] (6) {Proof minimization};
  \node [cloud, below of=6, node distance = 1.3cm] (7) {Proof};
  \draw [-to,black,thick] (1) to (5);
  \draw [-to,black,thick] (5) to (6);
  \draw [-to,black,thick] (6) to (7);
  \draw [-to,black,thick] (d) to [out=0, in=180] (3);
  \draw [-to,black,thick] (d) to [out=0, in=180] (4);
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}{Prediction algorithm}

Algorithm:\\ \ \  
  Nearest neighbor weighted by TF-IDF heuristics

\vspace{5mm}

Effect:\\ \ \
  Order goals from the database according to\\
  their distance to a target goal.

\vspace{5mm}

Remark:\\ \ \ 
  This is algorithm performs premise selection.\\ \ \
  How do we adapt it to predict tactics?

\end{frame}

\begin{frame}[fragile]{Policy: choosing a tactic}

Database of tactics is a map from goals to tactics.
\begin{lstlisting}[language=SMLSmall]
R (f n) (f (SUC n)) ==> transitive R: INDUCT_TAC
n * m <= n * p ==> (n = 0) vee m <= p   : REWRITE_TAC
INJ f U(:num) s ==> INFINITE s      : METIS_TAC
ldots
\end{lstlisting}
An order on goals induces an order on tactics.

Goal appearing during proof search:
\begin{lstlisting}[language=SMLSmall]
LENGTH (MAP f l) = LENGTH l
\end{lstlisting}

Policy for the new goal:
\begin{lstlisting}[language=SMLSmall]
Rank Tactic      Policy
1    REWRITE_TAC 0.5 
2    METIS_TAC   0.25
ldots 
4    INDUCT_TAC  0.0625
ldots
\end{lstlisting}
\end{frame}

\begin{frame}{Value function: provability of a list of goals}
Database of lists of goals:\\
\begin{itemize}
\item Positive examples: appears in human proofs.
\item Negative examples: produced during \tactictoe search but do not appear in 
the final proof.
\end{itemize}

Value function:\\
\ \ Percentage of positives in the 10 closest lists of goals of a target list 
of goals.

\vspace{5mm}

Future work:\\
\ \ Estimate the number of steps needed to prove a list of goals.

\end{frame}

\begin{frame}{Plan}
\begin{center}
\begin{tikzpicture}[node distance = 1.5cm]
  \node [] (aa) {\textbf{Supervised learning}};
  \node [cloud, below of=aa, node distance = 1cm] (a) {Formal library};
  \node [block, below of=a] (b) {Proof recording};
  \node [cloud, below of=b] (c) {Knowledge base};
  \node [block, below of=c, blue] (e) {Training};
  \node [block, below of=e] (d) {Predictors};

  \draw [-to,black,thick] (a) to (b);
  \draw [-to,black,thick] (b.south) to (c.north);
  \draw [-to,black,thick] (c) to [bend left]  (e);
  \draw [-to,black,thick] (e) to [bend left]  (c);
  \draw [-to,black,thick] (c.east) to [out=330,in=30] (d);

  \node [right of=aa, node distance=6.2cm] (0) {\textbf{Proving}};
  \node [cloud, below of=0, node distance = 1cm] (1) {Conjecture};
  \node [cloud, below of=1, node distance = 1.3cm] (2) {Search tree};
  \node [block, below of=2, node distance=1.1cm] (3) {Tactic policy};
  \node [above of=2, xshift=-16mm, node distance = 0.7cm] (00) {Proof
  search};
  \node [block, below of=3, node distance = 1.3cm] (4) {Tactic 
  evaluation};
  \node [block, fit=(2) (3) (4)] (5) {};
  \node [block, below of=4, node distance = 1.3cm] (6) {Proof minimization};
  \node [cloud, below of=6, node distance = 1.3cm] (7) {Proof};
  \draw [-to,black,thick] (1) to (5);
  \draw [-to,black,thick] (5) to (6);
  \draw [-to,black,thick] (6) to (7);
  \draw [-to,black,thick] (d) to [out=0, in=180] (3);
  \draw [-to,black,thick] (d) to [out=0, in=180] (4);
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}{Training}

Improve recorded data to create better predictions during search.

\end{frame}

\begin{frame}[fragile]{Training: orthogonalization}

Issue: Many tactics are doing the same job on a goal $g$.

\vspace{5mm}

Solution: Competition for $g$ where the most popular tactic wins.
\end{frame}

\begin{frame}[fragile]{Training: orthogonalization}

Recorded goal-tactic pair:
\begin{lstlisting}[language=SMLSmall]
LENGTH (MAP f l) = LENGTH l: INDUCT_TAC
\end{lstlisting}

Competition:
\begin{lstlisting}[language=SMLSmall]
            Progress  Coverage 
INDUCT_TAC  Yes       136
REWRITE_TAC No        2567
METIS_TAC   Yes       694
\end{lstlisting}

Added to the database:
\begin{lstlisting}[language=SMLSmall]
LENGTH (MAP f l) = LENGTH l: METIS_TAC
\end{lstlisting}
\end{frame}

\begin{frame}{Training: abstraction}
Issue: Some theorems are never used inside tactics.

\vspace{5mm}

Solution: Abstract all lists of theorems in a tactic\\ \ \ 
and instantiate them depending on the target goal.
\end{frame}

\begin{frame}[fragile]{Training: abstraction}

Abstraction algorithm:
\begin{lstlisting}[language=SMLSmall]
Original     : REWRITE_TAC [T1,T2]
Abstraction  : REWRITE_TAC X
Instantiation: REWRITE_TAC [T67, T1, T43, ldots]
\end{lstlisting}

\vspace{5mm}

Question: Dow we keep the original or the abstraction ?\\

\vspace{5mm}

Answer: Let them compete during orthogonalization.\\

\end{frame}

\begin{frame}{Training: preselection}

Issue: Predictions are too slow during proof search.

\vspace{5mm}

Solution: Preselect 1000 suitable tactics using dependencies.

\vspace{5mm}

Dependency: Appear in the same proof.

\end{frame}

\begin{frame}{Plan}
\begin{center}
\begin{tikzpicture}[node distance = 1.5cm]
  \node [] (aa) {\textbf{Supervised learning}};
  \node [cloud, below of=aa, node distance = 1cm] (a) {Formal library};
  \node [block, below of=a] (b) {Proof recording};
  \node [cloud, below of=b] (c) {Knowledge base};
  \node [block, below of=c] (e) {Training};
  \node [block, below of=e] (d) {Predictors};

  \draw [-to,black,thick] (a) to (b);
  \draw [-to,black,thick] (b.south) to (c.north);
  \draw [-to,black,thick] (c) to [bend left]  (e);
  \draw [-to,black,thick] (e) to [bend left]  (c);
  \draw [-to,black,thick] (c.east) to [out=330,in=30] (d);

  \node [right of=aa, node distance=6.2cm] (0) {\textbf{Proving}};
  \node [cloud, below of=0, node distance = 1cm] (1) {Conjecture};
  \node [cloud, below of=1, blue, node distance = 1.3cm] (2) {Search tree};
  \node [block, below of=2, blue, node distance=1.1cm] (3) {Tactic policy};
  \node [above of=2, xshift=-16mm, blue, node distance = 0.7cm] (00) {Proof
  search};
  \node [block, below of=3, blue, node distance = 1.3cm] (4) {Tactic 
  evaluation};
  \node [block, blue, fit=(2) (3) (4)] (5) {};
  \node [block, below of=4, node distance = 1.3cm] (6) {Proof minimization};
  \node [cloud, below of=6, node distance = 1.3cm] (7) {Proof};
  \draw [-to,black,thick] (1) to (5);
  \draw [-to,black,thick] (5) to (6);
  \draw [-to,black,thick] (6) to (7);
  \draw [-to,black,thick] (d) to [out=0, in=180] (3);
  \draw [-to,black,thick] (d) to [out=0, in=180] (4);
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}{Proof search: search tree}
\centering
\begin{tikzpicture}[node distance=0.7cm]
\node [tcircle] (0) {$g_i$};
\node [right of=0] (0r) {...};
\node [tcircle,right of=0r] (0rr) {$g_n$};
\node [left of=0] (0l) {...};
\node [tcircle,left of=0l] (0ll) {$g_0$};
\node [draw,fit=(0rr) (0ll)] (0b) {};
\node [left of=0ll] {$a_0$};
\node [above of=0,node distance=3cm] (2) {...};
\node [left of=2, tcircle] (2l) {$\phantom{g_3}$};
\node [tcircle,right of=2] (2r) {$\phantom{g_3}$};
\node [draw, fit=(2l) (2r)] (2b) {};
\node [left of=2l] {$a_j$};

\node [tcircle, left of=2l, node distance=2.4cm] (1r) {$\phantom{g_3}$};
\node [left of=1r] (1) {...};
\node [tcircle, left of=1] (1l) {$\phantom{g_3}$};
\node [draw, fit=(1l) (1r)] (1b) {};
\node [left of=1l] {$a_1$};

\node [right of=2r,node distance=2.4cm] (3l) {$\phantom{g_3}$};
\node [right of=3l] (3) {\phantom{...}};
\node [right of=3] (3r) {$\phantom{g_3}$};
\node [fit=(3l) (3r)] (3b) {};

\node [fit=(1r) (2l)] (1m) {};
\node [fit=(2l) (3r)] (2m) {};

\draw[-to,thick] (0) to node[xshift=-10](t1){$t_1$} (1b);
\draw[-to,thick] (0) to node[xshift=-4](t2){$t_j$} (2b);
\draw[-to,dotted,thick] (0) to (1m);
\draw[-to,dotted,thick] (0) to (2m);
\draw[-to,dotted,thick] (0) to node[xshift=15](t2){$t_m$} (3b);
\end{tikzpicture}
\end{frame}

\begin{frame}{Proof search: advanced tree search}
\begin{center}
\begin{tikzpicture}[node distance=3cm]

\node [tcircle] (0) {\only<1>{}\only<2>{0.7}\only<3>{0.5}\only<4->{0.6}};
\node [tcircle, above of=0] (1) {\only<4->{0.8}};
\node [tcircle, left of=1] (2) {\only<2>{0.7}\only<3->{0.5}};
\node [tcircle, right of=1] (3) {};

\draw[-to,thick] (0) to node[xshift=15]{0.25} (1);
\draw[-to,thick] (0) to node[xshift=15]{0.5} (2);
\draw[-to,thick] (0) to node[xshift=20]{0.125} (3);

\uncover<2->{
\node [tcircle, above of=2] (4) {};
\node [tcircle, left of=4] (5) {\only<3->{0.3}};
\node [tcircle, right of=4] (6) {};

\draw[-to,thick] (2) to node[xshift=15]{0.25} (4);
\draw[-to,thick] (2) to node[xshift=15]{0.5} (5);
\draw[-to,thick] (2) to node[xshift=20]{0.125} (6);
}
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}{Re-proving: results}
Evaluation is "fair". (not totally if you ask Freek)\\
Only previous proofs are available for training.

\begin{center}
\begin{tabular}{lcc}
\toprule
  & 7164 \holfour theorems (60s) & 3329 \cakeml theorems (15s)\\
\midrule
   \eprover   & 2472 (34.5\%) & \phantom{0000 } 20.0\%?\\
   \tactictoe & 4760 (66.4\%) & 1161 (34.9\%)\\
\midrule
   Total  & 4946 (69.0\%) & \\
\bottomrule
\end{tabular}
\end{center}
\end{frame}

\begin{frame}{Re-proving: number of proofs found in less that x seconds}
\pgfplotscreateplotcyclelist{my black}{
solid, mark repeat=100, mark phase=0, black!100\\
dashed, mark repeat=100, mark phase=0, black!100\\}

\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=south east, at={(0.9,0.1)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=60,
  ymin=0, ymax=4800,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=time, y=solved] {data/tactictoe_time};
\addplot table[x=time, y=solved] {data/eprover_time};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\end{frame}

\begin{frame}{Re-proving: number of human proofs of size x}
\pgfplotscreateplotcyclelist{my black}{
solid, mark repeat=100, mark phase=0, red!100\\
solid, mark repeat=100, mark phase=0, blue!100\\}

add cakeml
\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=\textwidth,
  height=0.7*\textwidth,
  ymin=0, ymax=2000,
  xmin=0, xmax=20,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
\legend{\holfour}
\end{axis}
\end{tikzpicture}
\end{frame}

\begin{frame}{Re-proving: percentage of proof of size x solved}
\pgfplotscreateplotcyclelist{my black}{
solid, mark repeat=100, mark phase=0, black!100\\
dashed, mark repeat=100, mark phase=0, black!100\\}

\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=north east, at={(0.9,0.9)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=20,
  ymin=0, ymax=100,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=oplen, y=solved] {data/tactictoe_by_oplen};
\addplot table[x=oplen, y=solved] {data/eprover_by_oplen};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\end{frame}

\begin{frame}{Plan}
\begin{center}
\begin{tikzpicture}[node distance = 1.5cm]
  \node [] (aa) {\textbf{Supervised learning}};
  \node [cloud, below of=aa, node distance = 1cm] (a) {Formal library};
  \node [block, below of=a] (b) {Proof recording};
  \node [cloud, below of=b] (c) {Knowledge base};
  \node [block, below of=c] (e) {Training};
  \node [block, below of=e] (d) {Predictors};

  \draw [-to,black,thick] (a) to (b);
  \draw [-to,black,thick] (b.south) to (c.north);
  \draw [-to,black,thick] (c) to [bend left]  (e);
  \draw [-to,black,thick] (e) to [bend left]  (c);
  \draw [-to,black,thick] (c.east) to [out=330,in=30] (d);

  \node [right of=aa, node distance=6.2cm] (0) {\textbf{Proving}};
  \node [cloud, below of=0, node distance = 1cm] (1) {Conjecture};
  \node [cloud, below of=1, node distance = 1.3cm] (2) {Search tree};
  \node [block, below of=2, node distance=1.1cm] (3) {Tactic policy};
  \node [above of=2, xshift=-16mm, node distance = 0.7cm] (00) {Proof
  search};
  \node [block, below of=3, node distance = 1.3cm] (4) {Tactic 
  evaluation};
  \node [block, fit=(2) (3) (4)] (5) {};
  \node [block, below of=4, blue, node distance = 1.3cm] (6) {Proof 
  minimization};
  \node [cloud, below of=6, blue, node distance = 1.3cm] (7) {Proof};
  \draw [-to,black,thick] (1) to (5);
  \draw [-to,black,thick] (5) to (6);
  \draw [-to,black,thick] (6) to (7);
  \draw [-to,black,thick] (d) to [out=0, in=180] (3);
  \draw [-to,black,thick] (d) to [out=0, in=180] (4);
\end{tikzpicture}
\end{center}
\end{frame}

\begin{frame}[fragile]{Minimization and embellishment}

Raw proof:
\begin{lstlisting}[language=SMLSmall]
boolLib.REWRITE_TAC [DB.fetch "list" "EVERY_CONJ",... ] 
  THEN
BasicProvers.Induct_on [HolKernel.QUOTE "l"] 
  THENL
  [BasicProvers.SRW_TAC [] [],
   simpLib.ASM_SIMP_TAC (BasicProvers.srw_ss ()) 
   [boolLib.DISJ_IMP_THM, DB.fetch "list" "MAP", 
    DB.fetch "list" "CONS_11", boolLib.FORALL_AND_THM]]
\end{lstlisting}

\vspace{5mm}
Processed proof:
\begin{lstlisting}[language=SMLSmall]
Induct_on `l` THENL
  [SRW_TAC [] [], 
   ASM_SIMP_TAC (srw_ss ()) 
   [DISJ_IMP_THM, FORALL_AND_THM]]
\end{lstlisting}

\end{frame}

\begin{frame}{Conclusion}
Summary: TacticToe learns from human proofs to solve new goals.

\vspace{5mm}

Advantages over ATPs (\eprover) for ITP (\holfour) users:
\begin{itemize}
\item Includes domain specific automation found in the ITP.
\item Generated proofs are human-level proofs.\\
\item No reconstruction needed.\\
\end{itemize}

\end{frame}

\begin{frame}{Future work}
Enlarge the action space:\\ \ \
  parameter synthesis, sequence of tactics, forward proofs.

\vspace{5mm}

Train tactics by evaluating input/output pairs.

\vspace{5mm}

Conjecture intermediate lemmas.
\vspace{5mm}
\end{frame}

\end{document}

### 2018-03-28 presentation/presentation.tex
minor

@@
 \toprule
-  & 7164 \holfour theorems (60s) & 3329 \cakeml theorems (15s)\\
+  & \holfour: 7164, 60s & \cakeml: 3329, 15s\\
 \midrule
-   \eprover   & 2472 (34.5\%) & \phantom{0000 } 20.0\%?\\
+   \eprover   & 2472 (34.5\%) & \\
    \tactictoe & 4760 (66.4\%) & 1161 (34.9\%)\\
@@
 
-add cakeml
-\centering
 \begin{tikzpicture}[scale=1]
@@
   legend style={anchor=north east, at={(0.9,0.9)}},
-  width=\textwidth,
+  width=0.3*\textwidth,
   height=0.7*\textwidth,
   ymin=0, ymax=2000,
-  xmin=0, xmax=20,
+  xmin=0, xmax=15,
   xtick={},
@@
   cycle list name=my black]
-\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
-\legend{\holfour}
+\addplot table[x=length, y=proofs] {data/original_proof_length};
+\addplot table[x=length, y=proofs] {data/cakeml_original_proof_length};
+\end{axis}
+\end{tikzpicture}
+\begin{tikzpicture}[scale=1]
+\begin{axis}[
+  legend style={anchor=north east, at={(0.9,0.9)}},
+  width=0.6*\textwidth,
+  height=0.7*\textwidth,
+  ymin=0, ymax=100,
+  xmin=15, xmax=50,
+  xtick={},
+  ytick={},
+  cycle list name=my black]
+\addplot table[x=length, y=proofs] {data/original_proof_length};
+\addplot table[x=length, y=proofs] {data/cakeml_original_proof_length};
+\legend{\holfour,\cakeml}
 \end{axis}

### 2018-03-28 presentation/presentation.tex
reconstruction

@@
 
-Goal appearing during proof search:
+\vspace{10mm}
+
+New goal appearing during proof search:
 \begin{lstlisting}[language=SMLSmall]
@@
 \begin{frame}{Re-proving: results}
-Evaluation is "fair". (not totally if you ask Freek)\\
-Only previous proofs are available for training.
+Evaluation is "fair". (not totally if you ask Freek)\\\ \ 
+Only previous proofs are available for training.\\\ \
+\tactictoe does not call external provers.\\
 
@@
 
-\begin{frame}{Re-proving: number of proofs found in less that x seconds}
+\begin{frame}{Re-proving: proofs of size x}
+\pgfplotscreateplotcyclelist{my black}{
+solid, red!100\\
+solid, blue!100\\
+dashed, red!100\\
+dashed, blue!100\\
+}
+\begin{tikzpicture}[scale=1]
+\begin{axis}[
+  legend style={anchor=north east, at={(0.9,0.9)}},
+  width=0.3*\textwidth,
+  height=0.7*\textwidth,
+  ymin=0, ymax=2000,
+  xmin=0, xmax=15,
+  xtick={},
+  ytick={},
+  cycle list name=my black]
+\addplot table[x=length, y=proofs] {data/original_proof_length};
+\addplot table[x=length, y=proofs] {data/cakeml_original_proof_length};
+\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
+\addplot table[x=length, y=proofs] {data/cakeml_proof_length};
+\end{axis}
+\end{tikzpicture}
+\begin{tikzpicture}[scale=1]
+\begin{axis}[
+  legend style={anchor=north east, at={(0.9,0.9)}},
+  width=0.7*\textwidth,
+  height=0.7*\textwidth,
+  ymin=0, ymax=100,
+  xmin=15, xmax=50,
+  xtick={},
+  ytick={},
+  cycle list name=my black]
+\addplot table[x=length, y=proofs] {data/original_proof_length};
+\addplot table[x=length, y=proofs] {data/cakeml_original_proof_length};
+\addplot table[x=length, y=proofs] {data/tactictoe_proof_length};
+\addplot table[x=length, y=proofs] {data/cakeml_proof_length};
+\legend{\holfour,\cakeml,\tactictoe in \holfour,\tactictoe in \cakeml}
+\end{axis}
+\end{tikzpicture}
+\end{frame}
+
+\begin{frame}{Re-proving: \holfour proofs found in less than x 
+seconds}
 \pgfplotscreateplotcyclelist{my black}{
@@
 
-\begin{frame}{Re-proving: number of human proofs of size x}
-\pgfplotscreateplotcyclelist{my black}{
-solid, mark repeat=100, mark phase=0, red!100\\
-solid, mark repeat=100, mark phase=0, blue!100\\}
-
-\begin{tikzpicture}[scale=1]
-\begin{axis}[
-  legend style={anchor=north east, at={(0.9,0.9)}},
-  width=0.3*\textwidth,
-  height=0.7*\textwidth,
-  ymin=0, ymax=2000,
-  xmin=0, xmax=15,
-  xtick={},
-  ytick={},
-  cycle list name=my black]
-\addplot table[x=length, y=proofs] {data/original_proof_length};
-\addplot table[x=length, y=proofs] {data/cakeml_original_proof_length};
-\end{axis}
-\end{tikzpicture}
-\begin{tikzpicture}[scale=1]
-\begin{axis}[
-  legend style={anchor=north east, at={(0.9,0.9)}},
-  width=0.6*\textwidth,
-  height=0.7*\textwidth,
-  ymin=0, ymax=100,
-  xmin=15, xmax=50,
-  xtick={},
-  ytick={},
-  cycle list name=my black]
-\addplot table[x=length, y=proofs] {data/original_proof_length};
-\addplot table[x=length, y=proofs] {data/cakeml_original_proof_length};
-\legend{\holfour,\cakeml}
-\end{axis}
-\end{tikzpicture}
-\end{frame}
-
-\begin{frame}{Re-proving: percentage of proof of size x solved}
+\begin{frame}{Re-proving: percentage of solved \holfour proof of size x }
 \pgfplotscreateplotcyclelist{my black}{
@@
 \item Generated proofs are human-level proofs.\\
-\item No reconstruction needed.\\
+\item No translation or reconstruction needed.\\
 \end{itemize}

### 2018-03-31 tactictoe_jv.tex
presentation
tactictoe-jv: saving orthogonalization
Start editing the paper again

@@
 \begin{abstract}
-We implement a tactical prover TacticToe on top of the HOL4 interactive theorem
+We implement a automated tactical prover TacticToe on top of the HOL4 interactive theorem
 prover. TacticToe learns from human proofs which mathematical technique is
-the most suitable in each proof situation.
-This knowledge is then used by a Monte Carlo tree search algorithm to
-explore the most promising tactic-level proof paths.
+suitable in each proof situation.
+This knowledge is then used in a Monte Carlo tree search algorithm to
+explore promising tactic-level proof paths.
 On a single CPU, with a time limit of 60 seconds, TacticToe proves 66.4\% of
-7164 theorems in the HOL4's standard library whereas
-Eprover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by
+the 7164 theorems in HOL4's standard library, whereas
+E~prover with auto-schedule solves 34.5\%. The success rate rises to 69.0\% by
 combining the results of TacticToe and E~prover.
@@
 \begin{itemize}
-\item Proof recording at the tactic level is made more precise
-  (Section~\ref{sec:recording}). We now support pattern
-matching constructions and opening SML modules.
 %The different
@@
 comparing subtrees of the search tree instead of leaves. The policy and
-evaluation are learned
+value function are learned
 through supervised learning.
@@
 through supervised learning (Section~\ref{s:prediction}).
-These predictors are implemented for three kind of objects: tactics,
-theorems and lists of goals.
+These predictors are implemented for three kinds of objects: tactics,
+theorems, and lists of goals.
 \item We introduce an orthogonalization process that eliminates redundant tactics (Section~\ref{sec:ortho}).
@@
 Section~\ref{s:experiments}.
+\item Proof recording at the tactic level is made more precise
+  (Section~\ref{sec:recording}). We now support pattern
+matching constructions and opening SML modules.
 \item Minimization and embellishment of the discovered proof facilitates user
@@
 period of time. It seems that the allotted time allows a strategy
-running without orthogonalization to catch up.
+running without orthogonalization to catch up. Yet, side experiments show
+that orthogonalization becomes relevant again when we add additional
+tactics through abstraction.
 It is best to predict a list of 16 theorems to instantiate arguments of type
@@
 \centering\ra{1.3}
-\small
 \begin{tabular}{lc}

### 2018-03-31 tactictoe_jv.tex
Rewrite the plan

@@
 The different components depicted in Figure~\ref{fig:components} give a
-high-level overview of the techniques necessary to generate
-tactic-level proofs by learning from recorded knowledge. From a formal library
-of ITP proofs, we extract and record tactic invocations
-(Section~\ref{sec:recording}). Based on such training examples, we
-  build efficient machine learning classifiers (Section~\ref{s:prediction}).
-  We then modify the knowledge base (shown as the training loop in
-  Figure~\ref{fig:components})
-  using orthogonalization (Section~\ref{sec:ortho}) and
-  an abstraction technique (Section~\ref{sec:synthesis})
-  to improve the quality of the predictions.
-  Additionally, the preselection method described in Section~\ref{sec:presel} increases
-  the speed of the prediction algorithm.
-  The learned predictors relying on this knowledge guide the proof search
-  algorithm presented in Section~\ref{sec:proofsearch}.
-  The search algorithm explores different proof paths in the proof tree, which
-  is defined in Section~\ref{sec:prelim}.
-  The result, if
-  successful, is a verified human-level proof composed of \holfour
-  tactics, minimized according to Section~\ref{sec:proofdisplay}.
-  In
-  Section~\ref{s:experiments}, we evaluate the system
-  on a large set of theorems originating from \holfour.
-  The generated proofs are compared with human proofs in
-  Section~\ref{sec:case_study}.
-
-The section order does not follow the implementation, because we have
-prioritized giving a description of
-the proof search algorithm early on. We hope this order
-benefits the reader by showing the essence of our work before
-the technical details of the prediction and recording algorithms.
+high-level overview of TacticToe and our approach to generating tactic-level proofs by learning from recorded knowledge.
+On one side, we have the learning aspect of TacticToe whose purpose is to produce a high quality function to predict a tactic given a goal state.
+Tactic prediction uses a knowledge base that is gleaned, in the first place, from human-written proofs.
+On the other side, we have the proving aspect of TacticToe, which uses tactic prediction to guide a proof search algorithm in the context of a given conjecture.
+
+This paper is organized firstly around the proving aspect:
+We define the proof search tree (Section~\ref{sec:prelim}), explain the essence of our approach to learning to predict tactics (Section~\ref{s:prediction}), and then present the prediction-guided proof search algorithm (Section~\ref{sec:proofsearch}).
+We also describe our preselection method (Section~\ref{sec:presel}) for speeding up prediction during search.
+Afterwards, we delve into some details of the learning aspect of TacticToe:
+We describe approaches to improving the knowledge base that supports prediction (orthogonalization in Section~\ref{sec:ortho} and abstraction in Section~\ref{sec:synthesis}), and describe how we build the knowledge base in the first place by extracting and recording tactic invocations from a library of human-written proofs
+(Section~\ref{sec:recording}).
+
+In Section~\ref{s:experiments}, we present an evaluation of TacticToe on a large set of theorems originating from \holfour.
+Finally, we present a feature of TacticToe that makes it more suitable for use by human provers, namely, minimization and prettification (Section~\ref{sec:proofdisplay}) and compare TacticToe's generated proofs to human proofs (Section~\ref{sec:case_study}).
 
@@
 purpose of the selection process. The similarity measure backing the
-predictions is described in Section~\ref{sec:features} and two methods for
+predictions is described in Section~\ref{sec:features} and three methods for
 improving the speed and quality of the predictions are presented in
-Section~\ref{sec:presel},\ref{sec:ortho} and \ref{sec:synthesis}.
+Sections~\ref{sec:presel}, \ref{sec:ortho} and \ref{sec:synthesis}.
 

### 2018-03-31 tactictoe_jv.tex
switching section 8 and 9
adding psl,ltac,eisbach references

@@
 \def\mizAR{\textsf{MizAR}\xspace}
+
+\def\ltac{\textsf{Ltac}\xspace}
 
@@
 during proof search.
+\item Proof recording at the tactic level is made more precise
+  (Section~\ref{sec:recording}). We now support pattern
+matching constructions and opening SML modules.
 \item Evaluation of \tactictoe with a 60 seconds timeout achieves a 66\%
@@
 Section~\ref{s:experiments}.
-\item Proof recording at the tactic level is made more precise
-  (Section~\ref{sec:recording}). We now support pattern
-matching constructions and opening SML modules.
+
 \item Minimization and embellishment of the discovered proof facilitates user
@@
 $t$ anymore.
+
+\section{Proof Recording}\label{sec:recording}
+Recording proofs in an LCF-style proof assistant can be done at different
+levels.
+In \holfour all existing approaches relied on modifying the kernel. This was
+used
+either to export the primitive inference
+steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
+or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable
+for our
+purpose of learning proving strategies at the intermediate tactic level. We
+therefore
+discuss recording proofs in an LCF-style proof assistant, with the focus on
+\holfour
+in this section.
+%
+Rather than relying on the underlying programming language, we parse the
+original script file containing tactic proofs. This enables us to extract the
+code of each tactic.
+Working with the string representation of a tactic is better than working with
+its value:
+\begin{itemize}
+\item Equality between two functions is easy to compute from
+their string representation, which allows us to avoid
+predicting the same tactic repeatedly.
+\item
+The
+results of
+\tactictoe can be returned in a readable form to the user. In this way, the
+user can learn from the feedback, analyze the proof and possibly improve on it.
+Furthermore \tactictoe does not need to be installed to run
+tactic proofs generated by \tactictoe. In this way, the further development of
+\tactictoe does not affect the robustness of \holfour.
+\item
+It is difficult (probably impossible) to transfer \sml values
+between theories, since they are not run in the same \holfour session. In
+contrast,
+\sml code can be exported and imported.
+\item To produce a tactic value from a code is easy in \sml, which can be
+achieved by using a reference to a tactic and updating it with the function
+\texttt{use}.
+\end{itemize}
+
+In order to transfer our tactic knowledge between theories, we want to be able
+to re-use the code of a tactic recorded in one theory in another.
+We also would like to make sure that the code is interpretable and that
+its interpretation does not change in the other theory.
+
+Even when developing one theory, the context is continually changing:
+modules are opened and local identifiers are defined. Therefore, it is unlikely
+that code extracted from a tactic proof is interpretable in any other part of
+\holfour without any post-processing.
+To solve this issue, we recursively replace each local identifier by its
+definition until we are able to write any expression with global identifiers
+only. Similarly, we prefix each global identifier by its module.
+We call this process \emph{globalization}. The result is a standalone \sml code
+executable in any \holfour theory.  In the absence of side effects, it is
+interpreted in the same way across theories.
+Therefore, we can guarantee that the behavior of recorded stateless tactics
+does not change. Even with some stateful tactics, the prediction algorithm is
+effective because most updates on the state of \holfour increase the
+strength of stateful tactics.
+
+\subsection{Implementation}
+We describe in more detail our implementation of the recording algorithm. It
+consists of 4 phases: tactic proof extraction, tactic proof globalization,
+tactic unit wrapping, and creation of goal-tactic pairs.
+
+Because of the large number of \sml constructions, we only describe the effect
+of these steps on a running example that contains selected parts of the theory
+of lists.
+
+%We comment on the treatment of global and local identifiers but  omit more
+%involved constructions
+%such as the globalization of infix operators.
+
+\begin{example}\label{ex:running}(Running example)
+\small
+\begin{lstlisting}[language=SMLSmall]
+open boolLib Tactic Prim_rec Rewrite
+ldots
+val LIST_INDUCT_TAC = INDUCT_THEN list_INDUCT ASSUME_TAC
+ldots
+val MAP_APPEND = store_thm("MAP_APPEND",
+  ``!(f:'a->'b) l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f l2)``,
+  STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND])
+\end{lstlisting}
+\end{example}
+
+The first line of this script file opens modules (called structures in \sml).
+Each of
+the modules contains a list of global identifiers which become directly
+accessible in
+the rest of the script.
+A local identifier \texttt{LIST\_INDUCT\_TAC} is declared next, which is a
+tactic that performs induction on lists. Below that, the theorem
+\texttt{MAP\_APPEND} is proven.
+The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
+the goal is split into a base case and an inductive case. Finally, both
+of these cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
+rewrites assumptions with the help of theorems previously declared in this
+theory.
+
+We first parse the script file to extract tactic proofs. Each of them is found
+in
+the third argument of a call to \texttt{store\_thm}. The result of tactic proof
+extraction for the running example in presented in Example~\ref{ex:r0}.
+
+\begin{example}\label{ex:r0}(Tactic proof extraction)
+\small
+\begin{lstlisting}[language=SMLSmall]
+STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
+\end{lstlisting}
+\end{example}
+
+In the next phase, we globalize identifiers of the tactic proofs.
+Infix operators such as \texttt{THEN} need to be processed in a special way so
+that they keep their infixity status after globalization. For simplicity,
+the
+globalization of infix operators is omitted in Example~\ref{ex:r1}.
+In this example, the three main cases that can happen during the globalization
+are depicted. The first one is the globalization of identifiers declared in
+modules. The global identifiers \texttt{STRIP\_TAC} and
+\texttt{ASM\_REWRITE\_TAC} are prefixed by their module \texttt{Tactic}. In
+this way, they will still be interpretable whether \texttt{Tactic} was open or
+not. The local identifier \texttt{LIST\_INDUCT\_TAC} is replaced by its
+definition which happens to contain two global identifiers.
+The previous paragraph describes the globalization for all identifiers except
+local theorems.  We do not replace a local
+theorem by its tactic proof (which its \sml definition) because we want to
+avoid unfolding proofs inside other proofs.
+If the theorem is stored in the \holfour database available across \holfour
+developments, we can obtain the theorem value by calling \texttt{DB.fetch}.
+Otherwise the globalization process fails and the local theorem identifier is
+kept unchanged. A recorded tactic with an unchanged local theorem as argument
+is only interpretable inside the current theory.
+
+\begin{example}\label{ex:r1} (Globalization)
+\begin{lstlisting}[language=SMLSmall]
+Tactic.STRIP_TAC THEN
+Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC THEN
+Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
+\end{lstlisting}
+\end{example}
+
+Running the globalized version of a tactic proof will have the exact same
+effect as the original. But since we want to extract information from this
+proof in the form of tactics and their input goals, we modify it further.
+In particular, we need to define at which level we should record the tactics in
+the tactic proof. The simplest idea would be to record all \sml subexpressions
+of type \texttt{tactic}. However, it will damage the quality of our data by
+dramatically increasing the number of tactics associated with a single goal.
+Imagine a tactic proof of the form \texttt{A THEN B THEN C}; then the tactics
+\texttt{A}, \texttt{A THEN B} and \texttt{A THEN B THEN C} would be valid
+advice for something close to their common input goal. The tactic
+\texttt{A THEN B THEN C} is likely to be too specific. In contrast, we can
+consider
+the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A}
+until
+\texttt{A} has no effect. For such calls, it is often preferable to record
+a single call to \texttt{REPEAT A} rather than multiple calls to \texttt{A}.
+Otherwise branches after each call to \texttt{A} would be necessary as part
+of proof search.
+To sum up, we would like to record only the most general tactics which
+make the most progress on a goal. As a trade-off between these two objectives,
+we split proofs into tactic units.
+
+\begin{definition}(Tactic unit)
+A tactic unit is an \sml expression of type \texttt{tactic} that does not
+contain an
+infix operator at its root.
+\end{definition}
+
+Because such tactic units are constructed from visual information present in
+the tactic proof, they often represent what a human user considers to be a
+single step of
+the proof.
+
+To produce the final recording tactic proof, we encapsulate each tactic unit
+in a recording function \texttt{R} (see Example~\ref{ex:wrap}). In order to
+record
+tactics in all \holfour theories, we replace the original proof by the
+recording proof in each theory and rebuild the \holfour library.
+
+\begin{example}\label{ex:wrap} (Tactic unit wrapping)
+\begin{lstlisting}[language=SMLSmall]
+R "Tactic.STRIP_TAC" THEN
+R "Prim_rec.INDUCT_THEN (DB.fetch \"list\" \"INDUCT\") Tactic.ASSUME_TAC" THEN
+R "Rewrite.ASM_REWRITE_TAC
+  [fff \"list\" \"MAP\", fff \"list\" \"APPEND\"]"
+\end{lstlisting}
+\end{example}
+
+At run time the function \texttt{R} is designed to observe what input goals a
+tactic  receives without changing its output. The implementation of \texttt{R}
+is presented in Example~\ref{ex:record}.
+
+\begin{example}\label{ex:record} (Code of the recording function)
+\begin{lstlisting}[language=SMLSmall]
+fun R stac goal = (save (stac,goal); tactic_of_sml stac goal)
+\end{lstlisting}
+\end{example}
+
+The function \texttt{save} writes the goal-tactic pair to disk increasing the
+number of entries in our database of tactics. The function
+\texttt{tactic\_of\_sml} interprets the \sml code \texttt{stac}. The tactic is
+then applied to the input goal to replicate the original behavior.
+After all modified theories are rebuilt, each call to a wrapped tactic in a
+tactic proof is recorded as a pair containing its globalized code and its
+input goal.
 
@@
 The performance of \tactictoe is compelling with more than half of the theorems
-that required a proof of length six being reproven. It is also consistently
+that required a proof of length six being re-proven. It is also consistently
 better than \eprover for any proof length.
-
-\section{Proof Recording}\label{sec:recording}
-Recording proofs in an LCF-style proof assistant can be done at different
-levels.
-In \holfour all existing approaches relied on modifying the kernel. This was
-used
-either to export the primitive inference
-steps~\cite{Wong95recordingand,DBLP:conf/itp/KumarH12}
-or to record dependencies of theorems~\cite{tgck-cpp15}. This was not suitable
-for our
-purpose of learning proving strategies at the intermediate tactic level. We
-therefore
-discuss recording proofs in an LCF-style proof assistant, with the focus on
-\holfour
-in this section.
-%
-Rather than relying on the underlying programming language, we parse the
-original script file containing tactic proofs. This enables us to extract the
-code of each tactic.
-Working with the string representation of a tactic is better than working with
-its value:
-\begin{itemize}
-\item Equality between two functions is easy to compute from
-their string representation, which allows us to avoid
-predicting the same tactic repeatedly.
-\item
-The
-results of
-\tactictoe can be returned in a readable form to the user. In this way, the
-user can learn from the feedback, analyze the proof and possibly improve on it.
-Furthermore \tactictoe does not need to be installed to run
-tactic proofs generated by \tactictoe. In this way, the further development of
-\tactictoe does not affect the robustness of \holfour.
-\item
-It is difficult (probably impossible) to transfer \sml values
-between theories, since they are not run in the same \holfour session. In
-contrast,
-\sml code can be exported and imported.
-\item To produce a tactic value from a code is easy in \sml, which can be
-achieved by using a reference to a tactic and updating it with the function
-\texttt{use}.
-\end{itemize}
-
-In order to transfer our tactic knowledge between theories, we want to be able
-to re-use the code of a tactic recorded in one theory in another.
-We also would like to make sure that the code is interpretable and that
-its interpretation does not change in the other theory.
-
-Even when developing one theory, the context is continually changing:
-modules are opened and local identifiers are defined. Therefore, it is unlikely
-that code extracted from a tactic proof is interpretable in any other part of
-\holfour without any post-processing.
-To solve this issue, we recursively replace each local identifier by its
-definition until we are able to write any expression with global identifiers
-only. Similarly, we prefix each global identifier by its module.
-We call this process \emph{globalization}. The result is a standalone \sml code
-executable in any \holfour theory.  In the absence of side effects, it is
-interpreted in the same way across theories.
-Therefore, we can guarantee that the behavior of recorded stateless tactics
-does not change. Even with some stateful tactics, the prediction algorithm is
-effective because most updates on the state of \holfour increase the
-strength of stateful tactics.
-
-\subsection{Implementation}
-We describe in more detail our implementation of the recording algorithm. It
-consists of 4 phases: tactic proof extraction, tactic proof globalization,
-tactic unit wrapping, and creation of goal-tactic pairs.
-
-Because of the large number of \sml constructions, we only describe the effect
-of these steps on a running example that contains selected parts of the theory
-of lists.
-
-%We comment on the treatment of global and local identifiers but  omit more
-%involved constructions
-%such as the globalization of infix operators.
-
-\begin{example}\label{ex:running}(Running example)
-\small
-\begin{lstlisting}[language=SMLSmall]
-open boolLib Tactic Prim_rec Rewrite
-ldots
-val LIST_INDUCT_TAC = INDUCT_THEN list_INDUCT ASSUME_TAC
-ldots
-val MAP_APPEND = store_thm("MAP_APPEND",
-  ``!(f:'a->'b) l1 l2. MAP f (APPEND l1 l2) = APPEND (MAP f l1) (MAP f l2)``,
-  STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND])
-\end{lstlisting}
-\end{example}
-
-The first line of this script file opens modules (called structures in \sml).
-Each of
-the modules contains a list of global identifiers which become directly
-accessible in
-the rest of the script.
-A local identifier \texttt{LIST\_INDUCT\_TAC} is declared next, which is a
-tactic that performs induction on lists. Below that, the theorem
-\texttt{MAP\_APPEND} is proven.
-The global tactic \texttt{STRIP\_TAC} first removes universal quantifiers. Then,
-the goal is split into a base case and an inductive case. Finally, both
-of these cases are solved by \texttt{ASM\_REWRITE\_TAC [MAP, APPEND]}, which
-rewrites assumptions with the help of theorems previously declared in this
-theory.
-
-We first parse the script file to extract tactic proofs. Each of them is found
-in
-the third argument of a call to \texttt{store\_thm}. The result of tactic proof
-extraction for the running example in presented in Example~\ref{ex:r0}.
-
-\begin{example}\label{ex:r0}(Tactic proof extraction)
-\small
-\begin{lstlisting}[language=SMLSmall]
-STRIP_TAC THEN LIST_INDUCT_TAC THEN ASM_REWRITE_TAC [MAP, APPEND]
-\end{lstlisting}
-\end{example}
-
-In the next phase, we globalize identifiers of the tactic proofs.
-Infix operators such as \texttt{THEN} need to be processed in a special way so
-that they keep their infixity status after globalization. For simplicity,
-the
-globalization of infix operators is omitted in Example~\ref{ex:r1}.
-In this example, the three main cases that can happen during the globalization
-are depicted. The first one is the globalization of identifiers declared in
-modules. The global identifiers \texttt{STRIP\_TAC} and
-\texttt{ASM\_REWRITE\_TAC} are prefixed by their module \texttt{Tactic}. In
-this way, they will still be interpretable whether \texttt{Tactic} was open or
-not. The local identifier \texttt{LIST\_INDUCT\_TAC} is replaced by its
-definition which happens to contain two global identifiers.
-The previous paragraph describes the globalization for all identifiers except
-local theorems.  We do not replace a local
-theorem by its tactic proof (which its \sml definition) because we want to
-avoid unfolding proofs inside other proofs.
-If the theorem is stored in the \holfour database available across \holfour
-developments, we can obtain the theorem value by calling \texttt{DB.fetch}.
-Otherwise the globalization process fails and the local theorem identifier is
-kept unchanged. A recorded tactic with an unchanged local theorem as argument
-is only interpretable inside the current theory.
-
-\begin{example}\label{ex:r1} (Globalization)
-\begin{lstlisting}[language=SMLSmall]
-Tactic.STRIP_TAC THEN
-Prim_rec.INDUCT_THEN (DB.fetch "list" "list_INDUCT") Tactic.ASSUME_TAC THEN
-Rewrite.ASM_REWRITE_TAC [DB.fetch "list" "MAP", DB.fetch "list" "APPEND"]
-\end{lstlisting}
-\end{example}
-
-Running the globalized version of a tactic proof will have the exact same
-effect as the original. But since we want to extract information from this
-proof in the form of tactics and their input goals, we modify it further.
-In particular, we need to define at which level we should record the tactics in
-the tactic proof. The simplest idea would be to record all \sml subexpressions
-of type \texttt{tactic}. However, it will damage the quality of our data by
-dramatically increasing the number of tactics associated with a single goal.
-Imagine a tactic proof of the form \texttt{A THEN B THEN C}; then the tactics
-\texttt{A}, \texttt{A THEN B} and \texttt{A THEN B THEN C} would be valid
-advice for something close to their common input goal. The tactic
-\texttt{A THEN B THEN C} is likely to be too specific. In contrast, we can
-consider
-the tactic \texttt{REPEAT A} that is repetitively calling the tactic \texttt{A}
-until
-\texttt{A} has no effect. For such calls, it is often preferable to record
-a single call to \texttt{REPEAT A} rather than multiple calls to \texttt{A}.
-Otherwise branches after each call to \texttt{A} would be necessary as part
-of proof search.
-To sum up, we would like to record only the most general tactics which
-make the most progress on a goal. As a trade-off between these two objectives,
-we split proofs into tactic units.
-
-\begin{definition}(Tactic unit)
-A tactic unit is an \sml expression of type \texttt{tactic} that does not
-contain an
-infix operator at its root.
-\end{definition}
-
-Because such tactic units are constructed from visual information present in
-the tactic proof, they often represent what a human user considers to be a
-single step of
-the proof.
-
-To produce the final recording tactic proof, we encapsulate each tactic unit
-in a recording function \texttt{R} (see Example~\ref{ex:wrap}). In order to
-record
-tactics in all \holfour theories, we replace the original proof by the
-recording proof in each theory and rebuild the \holfour library.
-
-\begin{example}\label{ex:wrap} (Tactic unit wrapping)
-\begin{lstlisting}[language=SMLSmall]
-R "Tactic.STRIP_TAC" THEN
-R "Prim_rec.INDUCT_THEN (DB.fetch \"list\" \"INDUCT\") Tactic.ASSUME_TAC" THEN
-R "Rewrite.ASM_REWRITE_TAC
-  [fff \"list\" \"MAP\", fff \"list\" \"APPEND\"]"
-\end{lstlisting}
-\end{example}
-
-At run time the function \texttt{R} is designed to observe what input goals a
-tactic  receives without changing its output. The implementation of \texttt{R}
-is presented in Example~\ref{ex:record}.
-
-\begin{example}\label{ex:record} (Code of the recording function)
-\begin{lstlisting}[language=SMLSmall]
-fun R stac goal = (save (stac,goal); tactic_of_sml stac goal)
-\end{lstlisting}
-\end{example}
-
-The function \texttt{save} writes the goal-tactic pair to disk increasing the
-number of entries in our database of tactics. The function
-\texttt{tactic\_of\_sml} interprets the \sml code \texttt{stac}. The tactic is
-then applied to the input goal to replicate the original behavior.
-After all modified theories are rebuilt, each call to a wrapped tactic in a
-tactic proof is recorded as a pair containing its globalized code and its
-input goal.
 
@@
 
-For \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15}
+In \hollight, the Tactician tool~\cite{DBLP:conf/sefm/Adams15}
 can transform a packed tactical proof into a series of interactive tactic
@@
 
-% \todoi{Cezary; add ltac and eisbach or remove PSL if you think it is not worth
-% mentionning.}
-% A similar effort for stronger automation is performed in \isabelle. Nagashima
-% developed a proof strategy language PSL~\cite{NagashimaK17psl} which allows the
-% user to
-% program with regular tactics and combine them with tools such as
-% \sledgehammer.
-% Selection of tactics from data collected in recorded proofs is in development
-% but for now it is up to the user to come up with a good meta-tactic.
+It is important to consider the strength of the language used for building 
+proofs and proof procedures when designing a system such as \tactictoe.
+In \isabelle, the language {\textsf{Eisbach}\xspace}~\cite{eisbach} was created 
+on top of the standard tactic language by giving access to constructors of the 
+tactic language to the end-user. The language 
+{\textsf{PSL}\xspace}~\cite{NagashimaK17psl} was also developed. It can combine 
+regular tactics and tools such as \sledgehammer.
+In \coq, the \ltac~\cite{ltac} meta language was designed to enrich the 
+tactic combinators of the \coq language. It adds functionalities such as 
+recursors and matching operators for terms and proof contexts.
+In \holfour, the strength of \sml as a proof strategy 
+language has been demonstrated by the implementation of complex proof 
+procedures such as \metis.
+
+%Comparatively, in \holfour, new tactics can be written by end-users in the 
+%implementation 
+%language \sml. Therefore, the design of a meta language was 
+%not necessary as every proof method is a \sml function. 
+%Using this language, the authors 
+%created a default search strategy ``try-hard'' that includes calls to 
+%\sledgehammer, and that outperforms \sledgehammer.
 
@@
 selection is handed over to the ATP component of a \emph{hammer}
-system, as in the hammers for \holfour~\cite{tgck-cpp15}, \hollight~\cite{holyhammer} and \isabelle~\cite{BlanchetteGKKU16}.
+system, as in the hammers for \holfour~\cite{tgck-cpp15}, 
+\hollight~\cite{holyhammer} and \isabelle~\cite{BlanchetteGKKU16}.
 

### 2018-08-14 review_response.tex
minor comments from 1st reviewer

\documentclass[]{scrartcl}
\usepackage{csquotes}
\MakeOuterQuote{"}

%opening
\title{Response for SCSS-JSC submission 2}
\subtitle{Aligning Concepts across Proof Assistant
Libraries}
\author{Thibault Gauthier \and Cezary Kaliszyk}

\usepackage{blindtext} 
\usepackage{framed}
%\usepackage[parfill]{parskip}

\setlength{\parskip}{1.5mm}
\setlength{\parindent}{0mm}

\begin{document}

\maketitle

We thank the reviewers for the valuable comments and questions. We have 
attempted to correct all
the issues suggested in the reviews and we give further answers to the 
questions raised in this response letter.

\section*{Reviewer 1}

\begin{leftbar}
My main concern about the paper is related to the usability by people that are 
not the authors of the paper. In several places of the paper, the authors 
indicate that some steps of the alignment process are carried out manually, but 
this is not completely clear; so, the authors should clarify this point in 
order to know which steps are conducted manually and which steps are automated. 
Moreover, the output results are huge plain text file, and exploring them is 
really time-consuming (as indicated previously, the authors have already done 
this, but it might be difficult to the same by other users in a different 
context). In addition, the output results must be manually analysed by the user.

Related to the previous point, the authors provide the final data obtained in 
their
experiments; however, they do not provide the programs that they have used. 
Hence, it is not possible to reproduce the authors' results or use their 
algorithms in different libraries and contexts.
\end{leftbar}

We have now provided the source code of all the experiments performed together 
with instructions how to rerun them
to ease the repeatability and reusability of our project.

When it comes to the section on human advice, the purpose is to allow the user 
to intervene in the decision process,
which could allow the algorithms to continue with more precise matches. Indeed, 
such a hybrid method is very
time consuming and was not evaluated in its full extent.
Except for the logical mappings, all presented matches were automatically 
derived.
We have now explained this in more detail in Section 5.4.

\begin{leftbar}Another issue that arose me when reading the paper were some of 
the applications of this work: translating results and conjecturing new 
theorems. Since the alignment of concepts is based on the quantity (and 
quality) of shared properties, how many results do you need to prove in order 
to discover the alignment of concepts? This is probably one drawback of your 
work: the need of big libraries (it is probably not a problem using the already 
developed libraries, but it might be a problem when creating a new library).
\end{leftbar}

Indeed, the proposed approach would not work well with small proof libraries.
However, given the bridges we constructed between all the considered libraries, 
a relatively small development may
map to specific structures in other libraries and allow the import of related 
theorems.

\begin{leftbar}Another point is whether you have noticed that some matchings 
are missed.
I mean, two concepts that were intended to represent the same mathematical 
object, but that are not matched. I guess that you do not have an automatic way 
of discovering those missing matches.
\end{leftbar}

The only proposed method of discovering such missing matches that our algorithm 
supports is through transitive matches.
We have now added some comments about it in section 4.6.

\begin{leftbar}It is a bit strange that you have not included the alignments in 
the same ITP, have you tried that experiment? Have you discovered something 
interesting? As an idea for the future, you could create some kind of alert 
system that inform the user about the matching of concepts in the libraries of 
the ITP to avoid duplications.
\end{leftbar}

We have tried experiments within a single ITP library. For curated libraries 
where no exact duplicates
occur (like the 6 libraries we consider) the alignments found are often 
concepts that are dual. These
include union and intersection of sets in Mizar, or addition and substraction 
in HOL libraries or
closely related concepts like lipschitzian and continuous.

In a work related to your question (presented at CICM Work in Progress last 
year), we tried to find the most likely substitutions (which mappings
should be applied together) and create a large number of conjectured formulas. 
Only 5 percent of them can be reproved. However they have not been too useful 
for current automation methods. This is now also shortly discussed in the paper.

\begin{leftbar}I wonder if you could define some threshold to differentiate the 
optimal matches, approximate matches and singular matches.
\end{leftbar}

Our current method is statistical, therefore the thresholds are estimates 
learned from previous library alignments.
We could also try to prove equivalence (modulo other mappings) of two similar 
concept but this would most of the times fail even for optimal matches because 
of minor differences in the definitions.

\begin{leftbar}
From my point of view, a graph presenting the workflow that is followed in your 
work to
align concepts might be useful for the reader.
\end{leftbar}
We included a graph together with clarifications to the overview of the 
procedure in Section 1.5.

\section*{Reviewer 2}

\subsection*{COMMENTS ABOUT THE RELEVANCE OF THE PROBLEM}

\begin{leftbar}
Automatically generating alignments is an important practical precondition to 
tackle the harder task of transferring knowledge between mathematical libraries 
at the shallow level, i.e. by translating a constant to the corresponding 
constant in the other library (avoiding to obtain two related but not 
identified notions). The problem seems therefore of primary importance.

However, the authors tackle the problem of aligning existing libraries from 
scratch. This needs to be done only once and forall and the automatically found 
result can be throughly corrected and improved by humans. After it is done, the 
most significant scenario, becomes the one where an author is starting a new 
formalization and he needs to pull in results from other libraries on the same 
constants he is interested in. This scenario is rather different for a series 
of reasons: 1) at the beginning of the new development there are usually only a 
small number of theorems proved and this makes it harder to find correct 
alignments; 2) the algorithm to automatically find alignments could be 
restricted to a subset of the system libraries that are already known to be 
potentially interesting.

I would like to hear from the authors what are their expectations on the 
behaviour of their algorithm on the new scenario. Maybe they could also run 
some benchmarks to receive a first impression, e.g. by showing how the matches 
improve when re-running the algorithm after adding one by one the lemmas in one 
theory.
\end{leftbar}

The proposed scenario is indeed quite different, and like the reviewer 
suggests, it would not work very well because of the first reason.
The only way to cope with it would be to automatically explore the theory 
before the alignment algorithm would be strong enough to discover new theorems 
by itself. As for the second reason why the proposed scenario is different, the 
restriction could be automatically found by a first pass of the alignment
algorithm.

Unfortunately for a number of the considered libraries no theory exploration 
mechanisms exist so far. Developing such mechanisms would in particular require 
powerful general purpose automation, which also has been limited in some of the 
systems. We completely agree that combining the presented work with theory 
exploration would be a very interesting line of work, but it would also be a 
large endeavor and has not been considered so far.

\subsection*{COMMENTS ABOUT THE TECHNICAL SOLUTION}
\begin{leftbar}All the choices the authors took seem reasonable. Of course, it 
is possible to think about totally different techniques to automatically 
identify alignments, for example borrowing other tools from AI and machine 
learning. However, being the first to address the problem at this scale, I 
believe they found a great compromise between simpliciy and efficiency. I will 
add later specific minor comments.

However, one remark I want to do now is that the kind of alignments the authors 
want to find should be directly related to the normalization they apply. For 
example, imagine that the authors want to align binary real division from HOL 
to ternary real division in some Coq libraries. Normalization should somehow 
``normalize away'' the proof argument in order to hope to match properties up 
to alpha-conversion. I can easily come up with other similar examples where 
aligning the library of a constructive, predicative type theory with that of an 
higher order logic requires to be more aggressive in the generalization process 
that produces properties. In previsous papers co-authored by the same authors 
certain classes of alignments were identified. I would have liked to see an 
explanation of normalization that is linked directly to those kind of 
alignments.
\end{leftbar}

Indeed, the manually constructed logical mappings are a weakness of our 
procedure.
However we have been able to find the logical mappings rather quickly, and they 
do not need to be extended as proof libraries grow.
It is true that for an algorithm to discover the logical mappings, the actual 
mappings would need to be more flexible.
In this respect, a possible extension of the algorithm could be to allow 
different parts of the term tree to be conceptualized.
However, we expect that the additional flexibility would reduce precision. We 
have included some discussion of this when considering
subterm conceptualizations.

\subsection*{COMMENTS ABOUT THE EVALUATION METHODOLOGY}

\begin{leftbar}
The evaluation of the technique is probably the weakest part of the paper and 
the one that suggests further future improvements. In particular, the algorithm 
only uses positive pre-existing knowledge in order to identify the alignments 
and complex heuristics to try to single out promising alignments and 
dependencies between the alignments. Finally, the responsibility of checking 
the quality of the alignments is totally left to the user.

A significant improvement would be to try to use automatic techniques to verify 
the quality of the alignments based on the generation of new knowledge, both 
positive and negative. For example, suppose that the system detect a possible 
alignment between concept A and B because A and B share a certain number of 
properties with a good score. In order to evaluate the alignment, the system 
could take for example all the properties on A that are not present for B and 
try to conjecture them to hold for B as well. Once the conjecture is done, 
automatic proof checkers and counterexample generators could be used to prove 
or disprove the conjecture. Even if they fail, the conjectures could often 
still be tested on a certain number of randomly generated tests. For example, 
it should not be hard to automatically understand that the alignment between 
natural and integer numbers is not optimal.
\end{leftbar}

This is a very good idea. There are three reasons why we have not been able to 
do this.
First, we do not have powerful enough general-purpose automation for all the 
provers considered.
Second, it is difficult to evaluate a match on its own. It makes little sense 
to map addition to multiplication but not 0 and 1 at the same time. Therefore 
we would need to consider and evaluate structural alignments consisting of a 
coherent set of matches.
Finally, given the number of matches evaluating each match using many  
automated reasoning experiments would be would be prohibitively time-consuming.

The closest to what the reviewer suggests and is currently possible has been 
considered by us in the CICM Work in Progress paper:
focusing on certain structural alignments one can evaluate the newly generated 
conjectures in some of the provers. The results
have been however much weaker than with other applications presented here. This 
is discussed in Section 1.5 and Section 5.

\begin{leftbar}
Coming to non optimal alignments, I would like to understand if the score 
computed by the algorithms is informative or not on the usefulness of the 
alignment. For example, aligning natural numbers with lists is quite useful 
because a natural number is a degenerate case of list (a list of elements of 
the unit type) and every theorem over lists induces a (possibly trivialized) 
theorem on natural numbers. Another example is aligning natural numbers with 
integers: even if the alignment is non optimal, the alignment becomes optimal 
(or almost so) if restricted to subsets of the two types (natural vs positive 
integers). Most other alignments, instead, are really likely to be completely 
useless.
\end{leftbar}

The scores are representative of the number and quality of the property shared 
by the two constants. So for a particular
pair of provers, the actual value of the score is indeed informative about the 
quality of the alignment, and the best
evidence of this are the actual correct alignments being found despite 
thousands of concepts present.

\begin{leftbar}
Did you ask a mathematician expert in the field what he thinks about the 
conjecture?
\end{leftbar}
We did not ask any expert mathematicians in the domain, instead relying of our 
own limited knowledge of the field to make those conjectures about the 
dynamical system.
We tried to prove those conjectures but failed a bit short so we did some 
empirical evaluation at dimension 2 confirming our intuition that the fix point 
we reach is the only stable fix point.

\begin{leftbar}
Minor comments
\end{leftbar}
We added examples describing the normalization, typing levels and logical 
mappings.
We also gave the patterns representing the most frequent properties, and 
commented on the effect of the coherence constraints on the conev
\end{document}

### 2018-08-14 tactictoe_jv.tex
minor
TODO
minor comments from 1st reviewer

@@
 \begin{abstract}
-We implement a automated tactical prover TacticToe on top of the HOL4 interactive theorem
+We implement an automated tactical prover TacticToe on top of the HOL4 
+interactive theorem
 prover. TacticToe learns from human proofs which mathematical technique is
@@
 For example, if the list of subgoals is empty, then calling the
-validation
-function on the empty list yields the original goal sequent as a theorem.
+validation function on the empty list yields the original goal as a theorem.
 In this way, tactics implement the plumbing of goal-directed proof
@@
 Since validation functions are only important to check the final
-proof, we
-omit them during the description of the proof search algorithm. We denote by
-$t(g)$ the list of goals produced by the application of a tactic $t$ to a goal
-$g$.
+proof, we omit them during the description of the proof search algorithm. We 
+denote by $t(g)$ the list of goals produced by the application of a tactic $t$ 
+to a goal $g$.
 
@@
 more balanced feedback on
-the success of
-each proof step by
-comparing subtrees of the search tree instead of leaves. The policy and
+the success of each proof step. The policy and
 value function are learned
@@
 according to
-some preset order. The other unsolved nodes are called \textit{pending nodes}.
+some preset order. The other unsolved goals are called \textit{pending goals}.
 If a node is unsolved, then it contains a unique open goal.
@@
 before its open goal is solved which justifies our terminology. In other words,
-our search tree has the following property: if $g$ is a pending node then
+our search tree has the following property: if $g$ is a pending goal then
 $T(g)$ is empty.
@@
 \item It does not loop, i.e., its output does not contain any goals that appear
-in the ancestor of the node of the current goal.
+in the ancestors of the node of the current goal.
 \item It is not a redundant step, i.e., there does not exist $t'$ in $T(g)$
@@
 \paragraph{Tactics}
-To start with, we build a database of tactics consisting
-of goal-tactic pairs recorded from successful tactic applications in human
+To start with, we build a database of tactics
+$\mathbb{D}_{\mathit{tactic}}$ consisting
+of tactic-goal pairs recorded from successful tactic applications in human
 proofs (recording will be discussed in
@@
 \paragraph{Theorems as Arguments of Tactics}
-We first collect a set of theorems for our predictor to select from.
+We first collect a database of theorems $\mathbb{D}_{\mathit{theorem}}$ for our 
+predictor to select from.
 It includes the \holfour theorem database and theorems from the local namespace.
@@
 \paragraph{Lists of Goals as Output of Tactics}
-We compile a dataset of tactic outputs during orthogonalization (see
+We compile a dataset of tactic outputs $\mathbb{D}_{\mathit{goallist}}$ during 
+orthogonalization (see
 Section~\ref{sec:ortho}).
@@
 Our search algorithm is a Monte Carlo tree search~\cite{montecarlo} algorithm
-that relies
-on a prior evaluation function and a prior policy function. Those priors are
-learned through direct supervised learning for the policy and via the data
-accumulated during orthogonalization for the evaluation. Therefore, this MCTS
-algorithm does not need to rely on roll outs for evaluation. The
+that relies on a prior evaluation function and a prior policy function. Those 
+priors are learned through direct supervised learning for the policy and via 
+the data accumulated during orthogonalization for the evaluation. The
 policy and evaluation are estimated by the simple supervised k-NN algorithm.
@@
 \paragraph{Prior policy}\label{sec:policy}
-Similarity between goals produced by the tactic predictor are hard
-to translate into a prior policy, so we rely solely on the prediction order
-to estimate the probability with which each tactic should be tried.
 Let $\mathfrak{T}$ be a search tree after a number of MCTS steps.
-Let $p$ be a selected node in $\mathfrak{T}$ and $g$ its open goal. We order
-the list of $n$ productive tactics already
-applied to $g$ by their similarity score. We note the resulting list
+Let $n$ be the number of predictions returned by the prior policy (500 in our 
+experiments).
+Let $p$ be a selected parent node in $\mathfrak{T}$ and $g$ its open goal. We 
+order the list of $n$ productive tactics already
+applied to $g$ by their similarity score. Indeed, a tactic is likely to 
+be useful to solve $g$ if it has been applied on a goal similar to 
+$g$ in a previous proofs. 
+We note the resulting list
 $t_0,\ldots,t_{n-1}$.
@@
 
-\[\mathit{WideningPolicy}(p) = (1 - c_{\mathit{policy}})^{n} * c_{\mathit{policy}}\]
+\[\mathit{WideningPolicy}(p) = (1 - c_{\mathit{policy}})^{n} * 
+c_{\mathit{policy}}\]
 
@@
 In order to speed up the predictions during the search for a proof of a
-conjecture $c$, we preselect 500 goal-tactic pairs and 500 theorems and a
-larger number of lists of goals induced by the selection of goal-tactic pairs.
+conjecture $c$, we preselect 500 tactic-goal pairs and 500 theorems and a
+larger number of lists of goals induced by the selection of tactic-goal pairs.
 Preselected objects are the only objects available to \tactictoe's search
@@
 
-The first idea is to select goal-tactic pairs and theorems by their similarity
+The first idea is to select tactic-goal pairs and theorems by their similarity
 with $c$ (i.e. $\mathit{Predictions}^{\mathit{tactic}}_{500} (c)$ and
@@
 multiple children and at most one parent.
-Once a dependency relation is established we can calculate a dependency score
+Once a dependency relation is established (see Definition~\ref{def:tacdep} and 
+Definition~\ref{def:thmdep}), we can calculate a dependency score
 for each object, which is the maximum of its similarity score
@@
 
-We now give a mathematical definition of the dependency relation for each kind
-of object.
-
-\begin{definition}(Dependencies of a goal-tactic pair)\\
-Let $\mathfrak{F}$ be the set of recorded goal-tactic pairs.
-The dependencies $D^*$ of a goal-tactic pair $(t_0,g_0)$ are
+\begin{definition}\label{def:tacdep}(Dependencies of a tactic-goal pair)\\
+Let $\mathbb{D}_{\mathit{tactic}}$ be the database of tactic-goal pairs.
+The dependencies $D^*$ of a tactic-goal pair $(t_0,g_0)$ are
 inductively defined by:
@@
 D_0 &=_{\mathrm{def}} \lbrace (t_0,g_0) \rbrace \\
-D_{n+1} &=_{\mathrm{def}} D_n \cup \lbrace (t,g)\in \mathfrak{F}\  |\ \exists
+D_{n+1} &=_{\mathrm{def}} D_n \cup \lbrace (t,g)\in 
+\mathbb{D}_{\mathit{tactic}}\  |\ \exists
 (t',g') \in D_n.\ g \in t'(g') \rbrace  \\
@@
 
-\begin{definition}(Dependencies of a theorem)\\
-Let $\mathfrak{R}$ be the set of recorded theorems.
+\begin{definition}\label{def:thmdep}(Dependencies of a theorem)\\
 The dependencies of a theorem $t$ are the set of top-level theorems in
-$\mathfrak{R}$ appearing in the proof of $t$. These dependencies are
-recorded by tracking the use of top-level theorems through the inference
+the database of theorems $\mathbb{D}_{\mathit{theorem}}$ appearing in the proof 
+of $t$. These 
+dependencies are recorded by tracking the use of top-level theorems through the 
+inferences of the
 kernel~\cite{tgck-cpp15}.
@@
 We do preselection also for lists of goals.
-To preselect a list of goals $l$ from our database of lists of goals, we
+To preselect a list of goals $l$ from our database of lists of goals 
+$\mathbb{D}_{\mathit{goallist}}$, we
 consider the tactic input $g$ that was at the origin of the production of $l$
 during orthogonalization. The list $l$ is preselected if and only if $g$
-appears in the 500 preselected goal-tactic pairs.
+appears in the 500 preselected tactic-goal pairs.
 %This has the
@@
 To solve this problem, we modify the construction of the tactics database.
-Each time a new goal-tactic pair $(t,g)$ is extracted from a tactic proof and
+Each time a new tactic-goal pair $(t,g)$ is extracted from a tactic proof and
 about to be recorded, we consider if there does not already exist a better
@@
 the
-$k$ closest goal-tactic pairs $\mathit{Predictions}^{\mathit{tactic}}_k(g)$.
+$k$ closest tactic-goal pairs $\mathit{Predictions}^{\mathit{tactic}}_k(g)$.
 In our experiments,
@@
 original tactic on $g$ and that
-appears in the largest number of goal-tactic pairs in the database.
+appears in the largest number of tactic-goal pairs in the database.
 The winning tactic $w$ is then associated with $g$, producing the pair $(w,g)$
@@
 \begin{definition} (Coverage)\\
-Let $\mathfrak{T}$ be the database of goal-tactic pairs. We define the
+Let $\mathfrak{T}$ be the database of tactic-goal pairs. We define the
 coverage $\textit{\mbox{Coverage}}(t)$ of a tactic $t$ by the number of times
@@
 g_2\]
+In this example, the definition of the 
+length of a proof is not specified, different definitions may produce different 
+subsumptions.
 By default, we however choose for efficiency reasons a minimal subsumption defined by:
@@
 \section{Abstraction}\label{sec:synthesis}
-One of the major weaknesses of the previous version of \tactictoe was that
+One of the major weaknesses of the previous version of 
+\tactictoe~\cite{tgckju-lpar17} was that
 it could not create its own tactics. Indeed, sometimes no previous tactic is
@@
 % ungeneralized
-original one. The generalized tactics are typically slower that their
+original one. The generalized tactics are typically slower than their
 original variants, but the added flexibility is worthwhile in practice. Moreover,
@@
 consists of 4 phases: tactic proof extraction, tactic proof globalization,
-tactic unit wrapping, and creation of goal-tactic pairs.
+tactic unit wrapping, and creation of tactic-goal pairs.
 
@@
 
-The function \texttt{save} writes the goal-tactic pair to disk increasing the
+The function \texttt{save} writes the tactic-goal pair to disk increasing the
 number of entries in our database of tactics. The function
@@
 In Table~\ref{tab:tuning}, the success rate of the different \tactictoe
-strategy is presented.
+strategies is presented.
 The first four techniques are tested relative to the same baseline indicated in
@@
 theorem list in tactics. At any rate, argument prediction for abstracted
-tactics is the technique that have the highest impact on the success of
+tactics is the technique that has the highest impact on the success of
 \tactictoe. Integration of ATPs being a specialization of this technique
@@
 that \metis is run with a very short timeout, allowing to close different part
-of search tree quickly. Because \metis is weaker that \tactictoe as an ATP, it
+of search tree quickly. Because \metis is weaker than \tactictoe as an ATP, it
 is given less predictions. But even if an essential lemma for the proof of a
@@
 of hammers on this type of problems~\cite{tgck-cpp15}. Moreover, \tactictoe is
-running a single set of parameters (strategy), where as hammers and ATPs have
+running a single set of parameters (strategy), whereas hammers and ATPs have
 been optimized and rely on a wide range of strategies.
@@
 \paragraph{Reinforcement learning}
-All our feature vectors have been learned form human proofs. We now can now
-also add goal-tactic pairs that appears in the last proof of \tactictoe. To
+All our feature vectors have been learned form human proofs. We can now
+also add tactic-goal pairs that appears in the last proof of \tactictoe. To
 prevent
@@
 \section{Case Study}\label{sec:case_study}
-Since the proofs generated by \tactictoe are meant to be parts of a \holfour development, it interesting to compare their quality with the original human proofs.
+Since the proofs generated by \tactictoe are meant to be parts of a \holfour 
+development, it is interesting to compare their quality with the original human 
+proofs.
 The quality of a proof in \holfour can be measured in terms of length, readability, maintainability and verification speed.
@@
 It is important to consider the strength of the language used for building 
-proofs and proof procedures when designing a system such as \tactictoe.
+proofs and proof procedures when designing a system such as \tactictoe. Indeed, 
+such system will be re-using those procedures.
 In \isabelle, the language {\textsf{Eisbach}\xspace}~\cite{eisbach} was created 

### 2018-08-14 review_response.tex
minor + review response selected issues

@@
 %opening
-\title{Response for SCSS-JSC submission 2}
-\subtitle{Aligning Concepts across Proof Assistant
-Libraries}
-\author{Thibault Gauthier \and Cezary Kaliszyk}
+\title{Response for JAR submission}
+\subtitle{TacticToe: Learning to Prove with Tactics}
+\author{\mbox{Thibault Gauthier} \and \mbox{Cezary Kaliszyk} \and \mbox{Josef
+		Urban} \and \mbox{Ramana Kumar} \and \mbox{Michael Norrish}}
 
@@
 We thank the reviewers for the valuable comments and questions. We have 
-attempted to correct all
-the issues suggested in the reviews and we give further answers to the 
-questions raised in this response letter.
+attempted to correct all the issues suggested in the reviews and we give 
+further answers to the questions raised in this response letter.
 
@@
 \begin{leftbar}
-My main concern about the paper is related to the usability by people that are 
-not the authors of the paper. In several places of the paper, the authors 
-indicate that some steps of the alignment process are carried out manually, but 
-this is not completely clear; so, the authors should clarify this point in 
-order to know which steps are conducted manually and which steps are automated. 
-Moreover, the output results are huge plain text file, and exploring them is 
-really time-consuming (as indicated previously, the authors have already done 
-this, but it might be difficult to the same by other users in a different 
-context). In addition, the output results must be manually analysed by the user.
-
-Related to the previous point, the authors provide the final data obtained in 
-their
-experiments; however, they do not provide the programs that they have used. 
-Hence, it is not possible to reproduce the authors' results or use their 
-algorithms in different libraries and contexts.
+No data is provided to help other researchers reproduce this experiment.  I 
+consider this is a flaw of the paper.  I think the paper should be improved in 
+the direction of making the experiment reproducible by other researchers, as 
+much as possible. The paper provides numerical evaluation of its results, 
+mostly in the form of 
+percentage of success under a given timeout.   I would also be interesting to 
+know how much it takes to re-run the whole process of learning and then using 
+for the complete database.
 \end{leftbar}
-
-We have now provided the source code of all the experiments performed together 
-with instructions how to rerun them
-to ease the repeatability and reusability of our project.
-
-When it comes to the section on human advice, the purpose is to allow the user 
-to intervene in the decision process,
-which could allow the algorithms to continue with more precise matches. Indeed, 
-such a hybrid method is very
-time consuming and was not evaluated in its full extent.
-Except for the logical mappings, all presented matches were automatically 
-derived.
-We have now explained this in more detail in Section 5.4.
-
-\begin{leftbar}Another issue that arose me when reading the paper were some of 
-the applications of this work: translating results and conjecturing new 
-theorems. Since the alignment of concepts is based on the quantity (and 
-quality) of shared properties, how many results do you need to prove in order 
-to discover the alignment of concepts? This is probably one drawback of your 
-work: the need of big libraries (it is probably not a problem using the already 
-developed libraries, but it might be a problem when creating a new library).
-\end{leftbar}
-
-Indeed, the proposed approach would not work well with small proof libraries.
-However, given the bridges we constructed between all the considered libraries, 
-a relatively small development may
-map to specific structures in other libraries and allow the import of related 
-theorems.
-
-\begin{leftbar}Another point is whether you have noticed that some matchings 
-are missed.
-I mean, two concepts that were intended to represent the same mathematical 
-object, but that are not matched. I guess that you do not have an automatic way 
-of discovering those missing matches.
-\end{leftbar}
-
-The only proposed method of discovering such missing matches that our algorithm 
-supports is through transitive matches.
-We have now added some comments about it in section 4.6.
-
-\begin{leftbar}It is a bit strange that you have not included the alignments in 
-the same ITP, have you tried that experiment? Have you discovered something 
-interesting? As an idea for the future, you could create some kind of alert 
-system that inform the user about the matching of concepts in the libraries of 
-the ITP to avoid duplications.
-\end{leftbar}
-
-We have tried experiments within a single ITP library. For curated libraries 
-where no exact duplicates
-occur (like the 6 libraries we consider) the alignments found are often 
-concepts that are dual. These
-include union and intersection of sets in Mizar, or addition and substraction 
-in HOL libraries or
-closely related concepts like lipschitzian and continuous.
-
-In a work related to your question (presented at CICM Work in Progress last 
-year), we tried to find the most likely substitutions (which mappings
-should be applied together) and create a large number of conjectured formulas. 
-Only 5 percent of them can be reproved. However they have not been too useful 
-for current automation methods. This is now also shortly discussed in the paper.
-
-\begin{leftbar}I wonder if you could define some threshold to differentiate the 
-optimal matches, approximate matches and singular matches.
-\end{leftbar}
-
-Our current method is statistical, therefore the thresholds are estimates 
-learned from previous library alignments.
-We could also try to prove equivalence (modulo other mappings) of two similar 
-concept but this would most of the times fail even for optimal matches because 
-of minor differences in the definitions.
-
-\begin{leftbar}
-From my point of view, a graph presenting the workflow that is followed in your 
-work to
-align concepts might be useful for the reader.
-\end{leftbar}
-We included a graph together with clarifications to the overview of the 
-procedure in Section 1.5.
 
@@
 
-\subsection*{COMMENTS ABOUT THE RELEVANCE OF THE PROBLEM}
+\begin{leftbar}
+The only reservation I have with the paper is that it is not clear to me 
+whether and how the proof tool is able to generate two commonly-used tactics. 
+Firstly,
+generating witnesses to existential goals using "EXISTS_TAC tm". If these 
+tactics are out of scope for the proof tool, then are all goals containing 
+existential
+quantifiers impossible for the tool to prove? Secondly, reducing a goal using 
+an implication theorem using "MATCH_MP_TAC th". Often such implication theorems 
+are
+highly specific to a single theory or even theorem, so how can tactics like 
+this be effectively learned? In my opinion the paper would be improved by a 
+brief
+discussion on how these tactics could be learned, either in the present or a 
+future system.
+\end{leftbar}
+
+A paragraph about predicting arrguments for these commonly used tactics has 
+been added. more later
+
+\section*{Reviewer 3}
 
 \begin{leftbar}
-Automatically generating alignments is an important practical precondition to 
-tackle the harder task of transferring knowledge between mathematical libraries 
-at the shallow level, i.e. by translating a constant to the corresponding 
-constant in the other library (avoiding to obtain two related but not 
-identified notions). The problem seems therefore of primary importance.
+The TacticToe system seems to particularly shine on inductive proofs over
+recursive data structures, as the authors themselves note. It seems
+therefore that the main place it wins over standard automated provers is
+that it can recognize when and how to perform proof by induction. This
+suggests that it may be less effective in other domains (e.g. where one has
+to synthesize non-obvious witness terms) but also suggests to me that the
+authors should also be comparing with the extensive literature on the
+automation of inductive proof, e.g. Boyer-Moore, Bundy and "Proof
+Planning", "inductionless induction" in rewriting, and the concrete
+realizations in HOL family theorem provers by Boulton and Papapanagiotou.
 
-However, the authors tackle the problem of aligning existing libraries from 
-scratch. This needs to be done only once and forall and the automatically found 
-result can be throughly corrected and improved by humans. After it is done, the 
-most significant scenario, becomes the one where an author is starting a new 
-formalization and he needs to pull in results from other libraries on the same 
-constants he is interested in. This scenario is rather different for a series 
-of reasons: 1) at the beginning of the new development there are usually only a 
-small number of theorems proved and this makes it harder to find correct 
-alignments; 2) the algorithm to automatically find alignments could be 
-restricted to a subset of the system libraries that are already known to be 
-potentially interesting.
-
-I would like to hear from the authors what are their expectations on the 
-behaviour of their algorithm on the new scenario. Maybe they could also run 
-some benchmarks to receive a first impression, e.g. by showing how the matches 
-improve when re-running the algorithm after adding one by one the lemmas in one 
-theory.
+The authors might also say a bit more about how much of this is specific to
+HOL4, and how much would change if one wanted to implement something
+similar in another prover either with a similar (HOL Light, Isabelle/HOL)
+or different (ACL2, Coq, PVS) logic. While the authors give quite an
+explicit discussion of the method, I did not see a link to the actual code.
+Is this available?
 \end{leftbar}
 
-The proposed scenario is indeed quite different, and like the reviewer 
-suggests, it would not work very well because of the first reason.
-The only way to cope with it would be to automatically explore the theory 
-before the alignment algorithm would be strong enough to discover new theorems 
-by itself. As for the second reason why the proposed scenario is different, the 
-restriction could be automatically found by a first pass of the alignment
-algorithm.
+More details for the re
 
-Unfortunately for a number of the considered libraries no theory exploration 
-mechanisms exist so far. Developing such mechanisms would in particular require 
-powerful general purpose automation, which also has been limited in some of the 
-systems. We completely agree that combining the presented work with theory 
-exploration would be a very interesting line of work, but it would also be a 
-large endeavor and has not been considered so far.
-
-\subsection*{COMMENTS ABOUT THE TECHNICAL SOLUTION}
-\begin{leftbar}All the choices the authors took seem reasonable. Of course, it 
-is possible to think about totally different techniques to automatically 
-identify alignments, for example borrowing other tools from AI and machine 
-learning. However, being the first to address the problem at this scale, I 
-believe they found a great compromise between simpliciy and efficiency. I will 
-add later specific minor comments.
-
-However, one remark I want to do now is that the kind of alignments the authors 
-want to find should be directly related to the normalization they apply. For 
-example, imagine that the authors want to align binary real division from HOL 
-to ternary real division in some Coq libraries. Normalization should somehow 
-``normalize away'' the proof argument in order to hope to match properties up 
-to alpha-conversion. I can easily come up with other similar examples where 
-aligning the library of a constructive, predicative type theory with that of an 
-higher order logic requires to be more aggressive in the generalization process 
-that produces properties. In previsous papers co-authored by the same authors 
-certain classes of alignments were identified. I would have liked to see an 
-explanation of normalization that is linked directly to those kind of 
-alignments.
-\end{leftbar}
-
-Indeed, the manually constructed logical mappings are a weakness of our 
-procedure.
-However we have been able to find the logical mappings rather quickly, and they 
-do not need to be extended as proof libraries grow.
-It is true that for an algorithm to discover the logical mappings, the actual 
-mappings would need to be more flexible.
-In this respect, a possible extension of the algorithm could be to allow 
-different parts of the term tree to be conceptualized.
-However, we expect that the additional flexibility would reduce precision. We 
-have included some discussion of this when considering
-subterm conceptualizations.
-
-\subsection*{COMMENTS ABOUT THE EVALUATION METHODOLOGY}
-
-\begin{leftbar}
-The evaluation of the technique is probably the weakest part of the paper and 
-the one that suggests further future improvements. In particular, the algorithm 
-only uses positive pre-existing knowledge in order to identify the alignments 
-and complex heuristics to try to single out promising alignments and 
-dependencies between the alignments. Finally, the responsibility of checking 
-the quality of the alignments is totally left to the user.
-
-A significant improvement would be to try to use automatic techniques to verify 
-the quality of the alignments based on the generation of new knowledge, both 
-positive and negative. For example, suppose that the system detect a possible 
-alignment between concept A and B because A and B share a certain number of 
-properties with a good score. In order to evaluate the alignment, the system 
-could take for example all the properties on A that are not present for B and 
-try to conjecture them to hold for B as well. Once the conjecture is done, 
-automatic proof checkers and counterexample generators could be used to prove 
-or disprove the conjecture. Even if they fail, the conjectures could often 
-still be tested on a certain number of randomly generated tests. For example, 
-it should not be hard to automatically understand that the alignment between 
-natural and integer numbers is not optimal.
-\end{leftbar}
-
-This is a very good idea. There are three reasons why we have not been able to 
-do this.
-First, we do not have powerful enough general-purpose automation for all the 
-provers considered.
-Second, it is difficult to evaluate a match on its own. It makes little sense 
-to map addition to multiplication but not 0 and 1 at the same time. Therefore 
-we would need to consider and evaluate structural alignments consisting of a 
-coherent set of matches.
-Finally, given the number of matches evaluating each match using many  
-automated reasoning experiments would be would be prohibitively time-consuming.
-
-The closest to what the reviewer suggests and is currently possible has been 
-considered by us in the CICM Work in Progress paper:
-focusing on certain structural alignments one can evaluate the newly generated 
-conjectures in some of the provers. The results
-have been however much weaker than with other applications presented here. This 
-is discussed in Section 1.5 and Section 5.
-
-\begin{leftbar}
-Coming to non optimal alignments, I would like to understand if the score 
-computed by the algorithms is informative or not on the usefulness of the 
-alignment. For example, aligning natural numbers with lists is quite useful 
-because a natural number is a degenerate case of list (a list of elements of 
-the unit type) and every theorem over lists induces a (possibly trivialized) 
-theorem on natural numbers. Another example is aligning natural numbers with 
-integers: even if the alignment is non optimal, the alignment becomes optimal 
-(or almost so) if restricted to subsets of the two types (natural vs positive 
-integers). Most other alignments, instead, are really likely to be completely 
-useless.
-\end{leftbar}
-
-The scores are representative of the number and quality of the property shared 
-by the two constants. So for a particular
-pair of provers, the actual value of the score is indeed informative about the 
-quality of the alignment, and the best
-evidence of this are the actual correct alignments being found despite 
-thousands of concepts present.
-
-\begin{leftbar}
-Did you ask a mathematician expert in the field what he thinks about the 
-conjecture?
-\end{leftbar}
-We did not ask any expert mathematicians in the domain, instead relying of our 
-own limited knowledge of the field to make those conjectures about the 
-dynamical system.
-We tried to prove those conjectures but failed a bit short so we did some 
-empirical evaluation at dimension 2 confirming our intuition that the fix point 
-we reach is the only stable fix point.
-
-\begin{leftbar}
-Minor comments
-\end{leftbar}
-We added examples describing the normalization, typing levels and logical 
-mappings.
-We also gave the patterns representing the most frequent properties, and 
-commented on the effect of the coherence constraints on the conev
 \end{document}

### 2018-08-14 tactictoe_jv.tex
minor + review response selected issues

@@
 the data accumulated during orthogonalization for the evaluation. The
-policy and evaluation are estimated by the simple supervised k-NN algorithm.
-A characteristic of the MCTS algorithm is that it offers a good trade-off
-between exploitation and exploration. This means that the algorithm searches
-deeper the more promising branches and leaves enough time for the exploration
-of less likely alternatives.
+prior policy and evaluation are estimated by the simple supervised k-NN 
+algorithm. We call them priors because these function are known a priori before 
+starting the proof search. In contrast, we define later in this section a 
+current policy and a current evaluation which are continuously updated by the 
+MCTS algorithm during proof search through a feedback mechanism. A 
+characteristic of the MCTS 
+algorithm is that it offers a good 
+trade-off between exploitation and exploration. This means that the algorithm 
+searches deeper the more promising branches and leaves enough time for the 
+exploration of less likely alternatives.
 
@@
 The prior policy estimates the probability of success of a tactic $t$ on a
-goal $g$ and associate it with the potential children that would be created by
+goal $g$ and associates it with the potential children that would be created by
 the
@@
 seconds. Removing tactics with similar effect as performed by the
-orthogonalization process is only beneficial when running \tactictoe for short
+orthogonalization process is only beneficial when running \tactictoe for a short
 period of time. It seems that the allotted time allows a strategy
@@
 Without embellishment, the returned tactic is barely readable as it contains
-information to guarantees that each \sml subterm is interpreted in the way in
-any context. Since we return the proof at a point where the \sml interpreter is
+information to guarantee that each \sml subterm is interpreted in the same way 
+in any context. Since we return the proof at a point where the \sml interpreter 
+is
 in a specific state, we can strip unnecessary information, such as
@@
 procedures such as \metis. 
+% TODO: Add a sentence about try-hard
 

### 2018-08-15 review_response.tex
Reproducibility paragraph

@@
 whether and how the proof tool is able to generate two commonly-used tactics. 
-Firstly,
-generating witnesses to existential goals using "EXISTS_TAC tm". If these 
+Firstly, generating witnesses to existential goals using \texttt{EXISTS\_TAC 
+tm}. If these 
 tactics are out of scope for the proof tool, then are all goals containing 
@@
 quantifiers impossible for the tool to prove? Secondly, reducing a goal using 
-an implication theorem using "MATCH_MP_TAC th". Often such implication theorems 
+an implication theorem using \texttt{MATCH\_MP\_TAC thm}. Often such 
+implication 
+theorems 
 are

### 2018-08-15 tactictoe_jv.tex
Reproducibility paragraph

@@
 
+\paragraph{Reproducibility}
+The \tactictoe project is part of the \holfour distribution 
+which can be downloaded at \url{https://github.com/HOL-Theorem-Prover/HOL}.
+The \tactictoe modules can be found in the directory
+\url{HOL/src/tactictoe}. The README explains how to launch the recording 
+process (which includes learning techniques such as orthogonalization),
+use \tactictoe interactively and run a re-proving experiments on a theory. The 
+recording process takes less than eight hours. An upper bound for 
+the time necessary for the final experiment can be given by the formula:
+\[\mathit{number\ of\ theorems} \times \mathit{timeout} = 7164 \times 60\ 
+\mathit{seconds} = 
+4.975\ \mathit{days}\] which translate in practice
+to about $2.5\ \mathit{days}$ because successful searches stop before reaching 
+the timeout. And finally, by running this experiment on 20 cores, we cut down 
+the actual total time requirement to a few hours.
+
 \subsection{Tuning \tactictoe}\label{sec:tuning}
-
 We optimize \tactictoe by tuning parameters of six different techniques:

### 2018-08-15 review_response.tex
Other types of arguments: wip

@@
 
+A paragraph was added in the methodology subsection where we talk about the 
+reproducibility of our results. We mention that TacticToe is part of the HOL4 
+(development) distribution and give the location of its directory.
+The time it would take to re-do the final experiment independently is about 3 
+days on a single CPU.
+
 \section*{Reviewer 2}
@@
 an implication theorem using \texttt{MATCH\_MP\_TAC thm}. Often such 
-implication 
-theorems 
+implication theorems 
 are
@@
 
-A paragraph about predicting arrguments for these commonly used tactics has 
-been added. more later
+The is able to predict this tactics with arguments fixed. We also tested a way 
+to predict the arguments independently taking account that our predictor is 
+hardly accurate by being able to try multiple arguments.
 

### 2018-08-15 tactictoe_jv.tex
Other types of arguments: wip

@@
 
-\paragraph{Abstraction of Tactic Arguments}
+\subsection{Abstraction of Tactic Arguments}
 The first step is to abstract arguments of tactics by replacing them by a
@@
 
-\paragraph{Instantiation of an Abstracted Tactic}
+\subsection{Instantiation of an Abstracted Tactic}
 An abstracted tactic is not immediately applicable to a goal, since it contains
@@
 given a goal $g$ as input and return the best predicted argument for this
-goal.
+goal. We discuss here specific algorithm for instantiating three common types 
+of arguments: theorem list, theorem, term.
+
+\paragraph{Theorem List}
+A typical example for
 
@@
 in $\hat{t}$.
-The type of theorem lists is a very common type in \holfour proofs and theorems contain
-enough information to be predicted accurately.
+The type of theorem lists is a very common type in \holfour proofs and theorems 
+contain enough information to be predicted accurately.
+We are helped by the fact that the order of theorems rarely matters in this 
+type of tactics and
+
+\paragraph{Theorem}
+
+\paragraph{Term}
+
+We re-use the scheme describe previously for changing the prediction of one 
+argument into the prediction of a list of arguments.
+The only things that changes is that we have to design a special predictor for 
+terms. In this context we think that the predictions should come from the list 
+of subterms present in the goal as a first approximation. So all algorithm will 
+try
+
+Term synthesis and instantiation is vast problem for higher-order theorems 
+provers such as Satallax and inductive programming Hipsec.
+
+Side experiments with independent term argument predictions 
+shows that trying to instantiate the tactic with subterms from the goal improve 
+the success rate by another 4 percent. 
+Nevertheless, in the evaluation presented in this paper, the 
+abstraction/instantiation 
+method is evaluated only for list of 
+theorems, tactics with other type of arguments were predicted with their 
+original arguments.
 
@@
 
-\paragraph{Selection of Abstracted Tactics}
+\subsection{Selection of Abstracted Tactics}
 As abstracted tactics do not appear in human proofs, we need to find a way to
@@
 which can be downloaded at \url{https://github.com/HOL-Theorem-Prover/HOL}.
-The \tactictoe modules can be found in the directory
-\url{HOL/src/tactictoe}. The README explains how to launch the recording 
+\tactictoe modules can be found in the directory
+\url{HOL/src/tactictoe}. The \url{README} file explains how to launch the 
+recording 
 process (which includes learning techniques such as orthogonalization),

### 2018-08-17 tactictoe_jv.tex
theorem list

@@
 become instances of the same abstracted tactic. As a consequence, it becomes
-harder and harder to predict suitable arguments. In our experiments, we create
-one abstraction $\hat{t}$ for each tactic $t$ by abstracting all subterms of
-type theorem list in $t$.
+harder and harder to predict suitable arguments.
 
@@
 unspecified arguments. To apply an abstracted tactic $\hat{t}$, we first need
-to
-instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to
+to instantiate the placeholders $h$ inside $\hat{t}$. Because it is difficult to
 build a general predictor effective on each type of argument, we manually
@@
 goal. We discuss here specific algorithm for instantiating three common types 
-of arguments: theorem list, theorem, term.
+of arguments: theorem list, theorem, term. Creating predictors for other types 
+is harder as they lack the structure that is present in mathematical 
+formulas.
 
 \paragraph{Theorem List}
-A typical example for
-
-Our default algorithm relies on argument predictions for theorems with a
-radius of 16. It uses the produced list
-$\mathit{Predictions}^{\mathit{tactic}}_{16}(g)$ to replace all the
-placeholders
-in $\hat{t}$.
-The type of theorem lists is a very common type in \holfour proofs and theorems 
-contain enough information to be predicted accurately.
-We are helped by the fact that the order of theorems rarely matters in this 
-type of tactics and
+The type of theorem lists is the most commonly used type in \holfour proofs.
+Here are two examples of such constructions with placeholder 
+\texttt{[thm\_list]}: \texttt{REWRITE\_TAC [thm\_list]},  \texttt{METIS\_TAC 
+[thm\_list]}.
+
+In general, the list of arguments in such tactics respect the two following 
+property:
+\begin{itemize}
+\item Additional theorems do not cause the tactic to fail and useless theorems 
+are ignored. For \texttt{REWRITE\_TAC}, loops are generally avoided because 
+most equations appears with only one orientation in the database of theorems.
+\item The order of theorems in the list does not influence the final result and 
+thus we can consider the list to be a set in our algorithm.
+\end{itemize}
+
+Since our predictor is not perfectly accurate and the presented tactics are 
+robust, we select a list of theorems usually longer that the original one to 
+instantiate these tactics.
+Our default predictor for list of theorems selects 16 theorems from the 
+database of theorems $\mathbb{D}_{\mathit{theorem}}$. It uses the produced list
+$\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= [thm1,\ldots,thm16]$ to 
+replace the
+placeholders in $\hat{t}$. The instantiation of the first abstracted is thus
+\texttt{REWRITE\_TAC [thm1,...,thm16]}.
 
 \paragraph{Theorem}
+Instantiating a single theorem is much harder than a list of theorems. Indeed,
+the predictions needs to be very precise for the tactic to be effective.
+For example, the tactic MPTAC only works if the theorem is an instance
+of the precondition. The best option would be to have a train a theorem 
+predictor that can adapt its predictions to the given tactic. However, this 
+would require a stronger machine learning model than nearest neighbor algorithm.
+Therefore, in order to mitigate the gap in the quality of the predictions, we 
+propose to create a tactic that tries instantiations until one succeeds. (i.e 
+by succeeding we mean produces a list of goal that does not contain the 
+original goal)
 
 \paragraph{Term}
-
 We re-use the scheme describe previously for changing the prediction of one 
@@
 
-Term synthesis and instantiation is vast problem for higher-order theorems 
-provers such as Satallax and inductive programming Hipsec.
-
-Side experiments with independent term argument predictions 
-shows that trying to instantiate the tactic with subterms from the goal improve 
-the success rate by another 4 percent. 
-Nevertheless, in the evaluation presented in this paper, the 
-abstraction/instantiation 
-method is evaluated only for list of 
-theorems, tactics with other type of arguments were predicted with their 
-original arguments.
-
-%We have positive results
-%with term abstraction but they are not evaluated in
-%Section~\ref{sec:evaluation}.
+Side experiments with term argument predictions 
+show that trying to instantiate the tactic with subterms 
+from the goal improve the success rate by another 4 percent. 
+However, in the evaluation presented in this paper, the 
+abstraction/instantiation method is evaluated only for the type of theorem list.
+Moreover, the current argument predictions are only affected by the given goal, 
+and we believe it would be beneficial to have more flexible predictors that can 
+adapt to their predictions to the abstracted tactic.
 

### 2018-08-17 tactictoe_jv.tex
theorem

@@
 \item The order of theorems in the list does not influence the final result and 
-thus we can consider the list to be a set in our algorithm.
+thus we can consider the list to be a set in our algorithm. This is 
+particularly true for the first-order theorem prover \texttt{METIS\_TAC}.
 \end{itemize}
@@
 replace the
-placeholders in $\hat{t}$. The instantiation of the first abstracted is thus
-\texttt{REWRITE\_TAC [thm1,...,thm16]}.
+placeholders in $\hat{t}$. Thus, the instantiation of the first construction 
+would be \texttt{REWRITE\_TAC [thm1,...,thm16]}.
 
 \paragraph{Theorem}
-Instantiating a single theorem is much harder than a list of theorems. Indeed,
-the predictions needs to be very precise for the tactic to be effective.
-For example, the tactic MPTAC only works if the theorem is an instance
-of the precondition. The best option would be to have a train a theorem 
-predictor that can adapt its predictions to the given tactic. However, this 
-would require a stronger machine learning model than nearest neighbor algorithm.
-Therefore, in order to mitigate the gap in the quality of the predictions, we 
-propose to create a tactic that tries instantiations until one succeeds. (i.e 
-by succeeding we mean produces a list of goal that does not contain the 
-original goal)
+To instantiate a single theorem correctly is much harder than finding a 
+superset of theorems. The following examples with placeholder \texttt{thm} 
+illustrate this point:
+  \texttt{MATCH\_MP\_TAC thm}, \texttt{INDUCT\_THEN thm ASSUME\_TAC}
+
+Here, the two examples accepts only very specific kind of theorems: the theorem 
+argument should match the precondition of the goal for \texttt{MATCH\_MP\_TAC} 
+and an induction principle should be given to \texttt{INDUCT\_THEN}.
+The likelihood of finding such theorem is low, thus given the list of 
+predictions $\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= 
+[thm1,\ldots,thm16]$, we create 16 instantiated tactics and try them in the 
+order given by the predictor, and we return the result of the first successful 
+tactic and produces new goals. For readers familiar with \holfour tacticals, we 
+give the single new tactic that can be created using this procedure on the 
+first example:  
+  \texttt{CHANGED\_TAC (MATCH\_MP\_TAC thm1) ORELSE ... ORELSE CHANGED\_TAC 
+  (MATCH\_MP\_TAC thm16)}.
 
@@
 abstraction/instantiation method is evaluated only for the type of theorem list.
-Moreover, the current argument predictions are only affected by the given goal, 
-and we believe it would be beneficial to have more flexible predictors that can 
-adapt to their predictions to the abstracted tactic.
+Moreover, the current argument predictions are only influenced by the input 
+goal, 
+and we believe that in the future it would be beneficial to have more flexible 
+predictors that can adapt their predictions to each abstracted tactic.
 

### 2018-08-20 tactictoe_jv.tex
Term instantiation

@@
 predictions $\mathit{Predictions}^{\mathit{theorem}}_{16}(g)= 
-[thm1,\ldots,thm16]$, we create 16 instantiated tactics and try them in the 
+[thm1,\ldots,thm16]$, we create 16 instantiations and try them in the 
 order given by the predictor, and we return the result of the first successful 
-tactic and produces new goals. For readers familiar with \holfour tacticals, we 
-give the single new tactic that can be created using this procedure on the 
-first example:  
-  \texttt{CHANGED\_TAC (MATCH\_MP\_TAC thm1) ORELSE ... ORELSE CHANGED\_TAC 
-  (MATCH\_MP\_TAC thm16)}.
+tactic and produces new goals. With a modified version of 
+the tactical \texttt{ORELSE} that fails if the tactic 
+did not have any effect, we
+can write for the first example a single new tactic encoding this procedure: 
+  \[\texttt{MATCH\_MP\_TAC thm1 ORELSE ... ORELSE MATCH\_MP\_TAC thm16}\]
 
 \paragraph{Term}
-We re-use the scheme describe previously for changing the prediction of one 
-argument into the prediction of a list of arguments.
-The only things that changes is that we have to design a special predictor for 
-terms. In this context we think that the predictions should come from the list 
-of subterms present in the goal as a first approximation. So all algorithm will 
-try
-
-Side experiments with term argument predictions 
-show that trying to instantiate the tactic with subterms 
-from the goal improve the success rate by another 4 percent. 
-However, in the evaluation presented in this paper, the 
-abstraction/instantiation method is evaluated only for the type of theorem list.
-Moreover, the current argument predictions are only influenced by the input 
-goal, 
-and we believe that in the future it would be beneficial to have more flexible 
-predictors that can adapt their predictions to each abstracted tactic.
+Some proofs requires to instantiate a tactic argument by a term. Here are two 
+abstracted 
+tactics that uses term 
+arguments with placeholders \texttt{term} and \texttt{thm}: 
+  \[\texttt{EXISTS\_TAC term},\ \texttt{SPEC\_THEN term STRUCT\_CASES\_TAC 
+  thm}\]
+Finding a witness to instantiate an existential formula is difficult because it 
+is likely that this witness may not have been proposed before and there is no 
+general rule on how to generate it. Instantiating 
+the term argument in the second tactic is easier because the expected term is a 
+variable of the goal. That is why we think that the first step toward term 
+prediction is to select subterms of the goal ordered by their similarity with 
+the term argument appearing in the original tactic. 
+Side experiments using a tactic that tries the 16 closest subterms 
+$[term1,\ldots,term16]$ to the goal with the procedure described in the 
+previous paragraph 
+shows that this simple term predictor improves the success rate of 
+\tactictoe by another 4 percent. The tactic derived from the abstraction
+\texttt{EXISTS\_TAC term} in this experiment is:
+   \[\texttt{EXISTS\_TAC term1 ORELSE ... ORELSE EXISTS\_TAC term16}\]
+
+In the future, with stronger machine learning models, we should be able to 
+select the right term from larger sets of terms including:
+\begin{itemize}
+ \item subterms of the goal (already implemented),
+ \item terms appearing as arguments of the abstracted tactic,
+ \item tactic sub-expressions of type term and their subterms,
+ \item subterms of theorems from the database $\mathbb{D}_{\mathit{theorem}}$,
+ \item terms created by term synthesis techniques.
+\end{itemize}
+
+\begin{remark} (Limitations)
+In the experiments presented in Section~\ref{sec:evaluation}, the 
+abstraction/instantiation method is evaluated only for the type ``theorem 
+list''.
+Another issue is that the argument predictions are only influenced by the input 
+goal. But we believe that it would be beneficial to have
+more flexible predictors that can adapt their selection to the abstracted 
+tactic they procure instantiations for.
+\end{remark}
 

### 2018-08-20 tactictoe_jv.tex
Term instantiation

@@
 did not have any effect, we
-can write for the first example a single new tactic encoding this procedure: 
+can write for the first example a single new tactic that implements this 
+procedure: 
   \[\texttt{MATCH\_MP\_TAC thm1 ORELSE ... ORELSE MATCH\_MP\_TAC thm16}\]
@@
   thm}\]
-Finding a witness to instantiate an existential formula is difficult because it 
-is likely that this witness may not have been proposed before and there is no 
+Finding a witness to instantiate an existential formula is challenging because 
+it is likely that this witness may not have been proposed before and there is 
+no 
 general rule on how to generate it. Instantiating 
 the term argument in the second tactic is easier because the expected term is a 
-variable of the goal. That is why we think that the first step toward term 
-prediction is to select subterms of the goal ordered by their similarity with 
-the term argument appearing in the original tactic. 
-Side experiments using a tactic that tries the 16 closest subterms 
-$[term1,\ldots,term16]$ to the goal with the procedure described in the 
-previous paragraph 
-shows that this simple term predictor improves the success rate of 
-\tactictoe by another 4 percent. The tactic derived from the abstraction
-\texttt{EXISTS\_TAC term} in this experiment is:
+variable of the goal. That is why we think that the first step 
+toward term argument prediction is to select subterms of the goal ordered by 
+their similarity with the term argument appearing in the original tactic. 
+Side experiments, using a tactic that tries the 16 closest subterms 
+$[term1,\ldots,term16]$ to the goal with the trial and error procedure 
+described in the previous paragraph, shows that this simple term predictor 
+improves the success 
+rate of 
+\tactictoe by another four percent. With the modified 
+\texttt{ORELSE} tactical, the tactic derived from the abstraction
+\texttt{EXISTS\_TAC term} in this experiment can be written as:
    \[\texttt{EXISTS\_TAC term1 ORELSE ... ORELSE EXISTS\_TAC term16}\]

### 2018-08-20 tactictoe_jv.tex
Proof plan draft

@@
 advice.
+
+A closely related approach is proof planning that was firstly describe in 
+~\cite{}. Proof planning is a meta approach to tactics constructing more and 
+more general tactics by sharing common proof patterns between different tactics.
+Comparatively, our approach tries to automate the construction of proof plans 
+by letting MCTS policy decide which tactics is the best in each situation. Our 
+approach can be thought as one general proof plan only guided by previous human 
+proofs. 
+From another angle, if we think of proof plans as ``domain'' specific tactics 
+our approach does not try to synthetize proof plan automatically but just 
+re-use them if they already exists.
+
 %However, we simply create a single label for each tactic call whereas each of
@@
 procedures such as \metis.
-% TODO: Add a sentence about try-hard
-
-%Comparatively, in \holfour, new tactics can be written by end-users in the 
-%implementation 
-%language \sml. Therefore, the design of a meta language was 
-%not necessary as every proof method is a \sml function. 
-%Using this language, the authors 
-%created a default search strategy ``try-hard'' that includes calls to 
-%\sledgehammer, and that outperforms \sledgehammer.
 

### 2018-08-20 tactictoe_jv.tex
draft: proof planning related work

@@
 A closely related approach is proof planning that was firstly describe in 
-~\cite{}. Proof planning is a meta approach to tactics constructing more and 
+~\cite{proofplan}. ~\cite{oyster}
+The philosophy of proof planning is to create (automatically or not) a few 
+tactics can prove a wide range of theorems. In this respect, if you consider 
+MCTS to be a proof plan then we have produced one general tactic that learns 
+from examples and has a large success rate.
+But proof planning is also concerned with tactic synthesis where new proof 
+plans (tactics in \holfour) may be discovered automatically from examples.
+From this point of view, \tactictoe does not look at proof patterns to infer 
+new tactics and creates new tactics only by abstracting arguments. Therefore, 
+it would be interesting to see how including tactic synthesis or even 
+human-designed proof plans (such as rippling~\cite{rippling}) in the database 
+of tactics $\mathbb{D}_{\mathit{tactic}}$.
+Proof planning is a meta approach to tactics constructing more and 
 more general tactics by sharing common proof patterns between different tactics.
@@
 proofs. 
-From another angle, if we think of proof plans as ``domain'' specific tactics 
+From another angle, if we think of proof plans as clever more general specific 
+tactics (like inductionless induction\cite{})
 our approach does not try to synthetize proof plan automatically but just 

### 2018-08-20 tactictoe_jv.tex
portability start
portability wip

@@
 the actual total time requirement to a few hours.
+
+\paragraph{Portability}
+Although \tactictoe is designed for \holfour, the ideas presented could be 
+ported to another ITP. We highlight here the major modules of \tactictoe  some 
+of the hurdles that developers might encounter during their re-implementation 
+for another ITP:
+\begin{itemize}
+	\item A recording mechanism that can extract context-independent tactics 
+	and the goal they were trying to solve. In our implementation, few tactics 
+	are stateful and they usually become more powerful over time.
+	\item Efficient tactics:
+	\item Efficient predictors
+	\item Efficient search: 
+	 the MCTS algorithm should be re-implemented in the language of the 
+	ITP, because better integration improves the efficiency of the algorithm.
+
+\end{itemize}
 

### 2018-08-21 tactictoe_jv.tex
proof plan citations

@@
 are considered.
-%intuition tells us that it should not be the main factor
 Our algorithm, on the other hand, relies mainly on the characteristics of the
@@
 A closely related approach is proof planning that was firstly describe in 
-~\cite{proofplan}. ~\cite{oyster}
+~\cite{DBLP:conf/cade/Bundy88}. 
+Proof planners have been implemented to facilitate interactive theorem proving 
+development in Oyster~\cite{DBLP:conf/cade/BundyHHS90},  
+Omega~\cite{DBLP:conf/cade/BenzmullerCFFHKKKMMSSS97} and 
+\isabelle~\cite{DBLP:conf/cade/DixonF03}.
 The philosophy of proof planning is to create (automatically or not) a few 
@@
 it would be interesting to see how including tactic synthesis or even 
-human-designed proof plans (such as rippling~\cite{rippling}) in the database 
+human-designed proof plans in the database 
 of tactics $\mathbb{D}_{\mathit{tactic}}$.
@@
 proofs. 
-From another angle, if we think of proof plans as clever more general specific 
-tactics (like inductionless induction\cite{})
 our approach does not try to synthetize proof plan automatically but just 
@@
 
-%However, we simply create a single label for each tactic call whereas each of
-%its
-%arguments is treated independently in \textsf{ML4PG}.
-%Our choice is motivated by the k-NN algorithm already used in
-%\holyhammer for the a of theorems.
-%, which can be easly adapted to the selection of tactics but would be improved
-%by a
-%version that can learn to predict tactics and their arguments independently.
-
-%A recent improvement on this system~\cite{}\todo{Cezary: could not remember
-%the
-%reference that you gave me here i think it was CICM 2017} includes the
-%possibility to
-%substitute theorems appearing in tactics by ones belonging to the same
-%cluster.
-%In this work, we are also able to predict theorems appearing as arguments of
-%tactics but we rely only on information from the current goal. Combining the
-%two approaches by predicting from the goal and from the original tactic
-%arguments is more difficult but is likely to produce more suitable lemmas.
-%
 Proof patching \cite{RingerYLG18} is an approach that attempts to learn the

### 2018-08-21 tactictoe_jv.tex
proof planning related work

@@
 A closely related approach is proof planning that was firstly describe in 
-~\cite{DBLP:conf/cade/Bundy88}. 
-Proof planners have been implemented to facilitate interactive theorem proving 
+~\cite{DBLP:conf/cade/Bundy88}. Proof planners have been implemented to 
+facilitate interactive theorem proving 
 development in Oyster~\cite{DBLP:conf/cade/BundyHHS90},  
@@
 \isabelle~\cite{DBLP:conf/cade/DixonF03}.
-The philosophy of proof planning is to create (automatically or not) a few 
-tactics can prove a wide range of theorems. In this respect, if you consider 
-MCTS to be a proof plan then we have produced one general tactic that learns 
-from examples and has a large success rate.
-But proof planning is also concerned with tactic synthesis where new proof 
-plans (tactics in \holfour) may be discovered automatically from examples.
-From this point of view, \tactictoe does not look at proof patterns to infer 
-new tactics and creates new tactics only by abstracting arguments. Therefore, 
-it would be interesting to see how including tactic synthesis or even 
-human-designed proof plans in the database 
-of tactics $\mathbb{D}_{\mathit{tactic}}$.
-Proof planning is a meta approach to tactics constructing more and 
-more general tactics by sharing common proof patterns between different tactics.
-Comparatively, our approach tries to automate the construction of proof plans 
-by letting MCTS policy decide which tactics is the best in each situation. Our 
-approach can be thought as one general proof plan only guided by previous human 
-proofs. 
-our approach does not try to synthetize proof plan automatically but just 
-re-use them if they already exists.
+A proof plan can be seen as a program that given a goal 
+returns the appropriate tactic for this goal. From this point of view, 
+\tactictoe is a general proof plan that learns to apply the right tactic from 
+examples and possess a backtracking mechanism given by the MCTS algorithm.
+The philosophy of proof planning is also to create a few 
+tactics can prove a wide range of theorems. In this respect, if we omit the 
+abstraction technique, \tactictoe is not synthesizing any tactic but simply 
+selecting them out of large tactic set. In the end, proof planning methodology 
+may become necessary for improving \tactictoe as it provides a way to generate 
+tactic applicable to a 
+large range of formulas. And more general tactics are usually easier to predict 
+accurately by machine learning models as they are associated with more 
+(successful) examples in our database.
 

### 2018-08-22 tactictoe_jv.tex
portability

@@
 \paragraph{Portability}
-Although \tactictoe is designed for \holfour, the ideas presented could be 
-ported to another ITP. We highlight here the major modules of \tactictoe  some 
-of the hurdles that developers might encounter during their re-implementation 
-for another ITP:
-\begin{itemize}
-	\item A recording mechanism that can extract context-independent tactics 
-	and the goal they were trying to solve. In our implementation, few tactics 
-	are stateful and they usually become more powerful over time.
-	\item Efficient tactics:
-	\item Efficient predictors
-	\item Efficient search: 
-	 the MCTS algorithm should be re-implemented in the language of the 
-	ITP, because better integration improves the efficiency of the algorithm.
-
-\end{itemize}
+Although \tactictoe is designed for \holfour, the \tactictoe modules 
+implemented in this paper could be ported to another ITP. In order to 
+facilitate this work, we give
+ideas on how to adapt the major components of \tactictoe in the context of 
+another ITP and warn about some of the issues that could arise during this work.
+First, the recording mechanism requires the ability to parse and manipulate 
+proof scripts and include the possibility to define and recognize what a tactic 
+is. In \holfour, \sml is our proof scripts language, so this was achieved by 
+parsing the \sml programming language. In other ITPs, this task may be 
+facilitated by the fact that proof scripts are written with a specifically 
+designed language so the ability to modify syntax trees of proof scripts may 
+already be available. Another issue that can negatively impact the quality of 
+the recorded data is the fact that the effect of the tactics may depend on the 
+state of the ITP (i.e. flags), which rarely happens in \holfour. If the 
+behavior of a tactic drastically changes under different context, then the part 
+of the context that affects it should become part of the tactic.
+Porting the other modules (abstraction, orthogonalization, predictors and MCTS 
+search) should be quite straightforward. Provided that application of 
+tactics and the computation of predictions is fast enough (in the order of 
+10 milliseconds each), the MCTS algorithm should explore enough of the tactic 
+space to be a valuable tool for theorem proving.
 

### 2018-08-22 review_response.tex
theorem list
review_response

--- 2018-08-15 review_response.tex
@@
 
+A paragraph on ``Reproducibility'' was added to the Subsection ``Methodology''. 
+There, 
+we talk about the 
+reproducibility of our results with information about the availability of our 
+source code (part of the HOL4 standard distribution) and the 
+time it takes to reproduce the results (learning and evaluation).
+
 \section*{Reviewer 2}
@@
 
-A paragraph about predicting arrguments for these commonly used tactics has 
-been added. more later
+The current version of TacticToe cannot synthetize non-obvious 
+witness, unless it leverages some clever 
+tactics to perform this task. In practice, the number of theorems containing 
+an existential quantifiers is significant (about 10 percent) but they are 
+specialized tactics that can solve problems containing existentials (e.g. 
+linear arithmetic). It is true that as TacticToe improves this weakness 
+may become more palatable and methods coming from proof planning, inductive 
+theorem proving or even machine learning in general, could be integrated.
+Possible solutions of the problems of predicting exact arguments are now 
+proposed in Section ``Abstraction''. And we give examples for the two most 
+common type of arguments appearing in HOL4 tactics (terms and theorems).
 
@@
 
-More details for the re
-
+A paragraph on Proof Planning was added in the Section ``Related Work'' 
+comparing aspects of proof planning to our work: proof automation, tactic 
+generalization and tactic synthesis. Term synthesis limitations are
+addressed in our reply to Reviewer 2.
+The TacticToe framework is mostly dependent on the proof script language of the 
+prover and not its logical foundation. Indeed, tactics can be seen as an 
+interface between the user and the logical foundations. We now added a 
+paragraph on ``Portability'' in the Subsection ``Methodology'', where we 
+discuss how to port our work to a different ITP. The source code is part of the 
+standard HOL4 distribution and a link is provided in the same subsection.
 \end{document}

### 2018-08-22 tactictoe_jv.tex
portability
minor

@@
 ideas on how to adapt the major components of \tactictoe in the context of 
-another ITP and warn about some of the issues that could arise during this work.
+another ITP and warn about some of the issues that could arise.
 First, the recording mechanism requires the ability to parse and manipulate 
@@
 the recorded data is the fact that the effect of the tactics may depend on the 
-state of the ITP (i.e. flags), which rarely happens in \holfour. If the 
-behavior of a tactic drastically changes under different context, then the part 
-of the context that affects it should become part of the tactic.
+state of the ITP (i.e. flags), which rarely happens in \holfour.
 Porting the other modules (abstraction, orthogonalization, predictors and MCTS 

### 2018-08-24 tactictoe_jv.tex
Minor grammar fixes

@@
 
-Here, the two examples accepts only very specific kind of theorems: the theorem 
+Here, the two examples accept only very specific kind of theorems: the theorem 
 argument should match the precondition of the goal for \texttt{MATCH\_MP\_TAC} 
@@
 \paragraph{Term}
-Some proofs requires to instantiate a tactic argument by a term. Here are two 
+Some proofs require to instantiate a tactic argument by a term. Here are two 
 abstracted 
-tactics that uses term 
+tactics that use term 
 arguments with placeholders \texttt{term} and \texttt{thm}: 
@@
 their similarity with the term argument appearing in the original tactic. 
-Side experiments, using a tactic that tries the 16 closest subterms 
+Additional experiments, using a tactic that tries the 16 closest subterms 
 $[term1,\ldots,term16]$ to the goal with the trial and error procedure 
-described in the previous paragraph, shows that this simple term predictor 
+described in the previous paragraph, show that this simple term predictor 
 improves the success 

### 2018-08-24 review_response.tex
review_response
Minor grammar fixes

@@
 The current version of TacticToe cannot synthetize non-obvious 
-witness, unless it leverages some clever 
+witnesses, unless it leverages some clever 
 tactics to perform this task. In practice, the number of theorems containing 
-an existential quantifiers is significant (about 10 percent) but they are 
+an existential quantifiers is significant (about 10 percent) but there are 
 specialized tactics that can solve problems containing existentials (e.g. 
-linear arithmetic). It is true that as TacticToe improves this weakness 
-may become more palatable and methods coming from proof planning, inductive 
-theorem proving or even machine learning in general, could be integrated.
+linear arithmetic). It is true that as TacticToe improves, this weakness 
+might diminish. Methods coming from proof planning, inductive 
+theorem proving or even machine learning in general could be integrated.
 Possible solutions of the problems of predicting exact arguments are now 
@@
 prover and not its logical foundation. Indeed, tactics can be seen as an 
-interface between the user and the logical foundations. We now added a 
+interface between the user and the logical foundations. We added a 
 paragraph on ``Portability'' in the Subsection ``Methodology'', where we 

### 2018-08-24 tactictoe_jv.tex
more in proof planning

@@
 Another issue is that the argument predictions are only influenced by the input 
-goal. But we believe that it would be beneficial to have
+goal. % But 
+We believe that it would be beneficial to have
 more flexible predictors that can adapt their selection to the abstracted 
@@
 
-A closely related approach is proof planning that was firstly describe in 
+A closely related approach is proof planning that was first described in 
 ~\cite{DBLP:conf/cade/Bundy88}. Proof planners have been implemented to 
@@
 The philosophy of proof planning is also to create a few 
-tactics can prove a wide range of theorems. In this respect, if we omit the 
-abstraction technique, \tactictoe is not synthesizing any tactic but simply 
-selecting them out of large tactic set. In the end, proof planning methodology 
-may become necessary for improving \tactictoe as it provides a way to generate 
-tactic applicable to a 
-large range of formulas. And more general tactics are usually easier to predict 
+tactics that can prove a wide range of theorems. % In this respect,  if we omit 
+Apart from the 
+abstraction technique, \tactictoe is not yet synthesizing any tactics but simply 
+selecting them out of a large tactic set. This may become necessary in the future, 
+using a variety of statistical/symbolic learning methods.
+% In the end, proof planning methodology 
+% may become necessary for improving \tactictoe as it provides a way to generate 
+% tactic applicable to a 
+% large range of formulas. 
+More general tactics are usually easier to predict 
 accurately by machine learning models as they are associated with more 

### 2018-08-24 tactictoe_jv.tex
more

@@
 Proof patching \cite{RingerYLG18} is an approach that attempts to learn the
-proof changes necessary for a modification of a prover library or system. The
-approach is hard to apply for new proofs. Following this idea, \tactictoe
-could also try to learn from different versions of the prover library.
+proof changes necessary for a modification of a prover library or system. This
+approach is however hard to use to create completely new proofs. Following this idea, \tactictoe
+could try to learn from different versions of the prover library.
 

### 2018-09-03 tactictoe_jv.tex
JU tweaks in reproducibility

@@
 \paragraph{Reproducibility}
-The \tactictoe project is part of the \holfour distribution 
+The \tactictoe project is a part of the \holfour distribution 
 which can be downloaded at \url{https://github.com/HOL-Theorem-Prover/HOL}.
-\tactictoe modules can be found in the directory
+The \tactictoe modules can be found in the directory
 \url{HOL/src/tactictoe}. The \url{README} file explains how to launch the 
@@
 process (which includes learning techniques such as orthogonalization),
-use \tactictoe interactively and run a re-proving experiments on a theory. The 
-recording process takes less than eight hours. An upper bound for 
+use \tactictoe interactively and run re-proving experiments on a theory. The 
+recording process takes less than eight hours for the \holfour standard library. An upper bound for 
 the time necessary for the final experiment can be given by the formula:
@@
 \mathit{seconds} = 
-4.975\ \mathit{days}\] which translate in practice
-to about $2.5\ \mathit{days}$ because successful searches stop before reaching 
-the timeout. And finally, by running this experiment on 20 cores, we cut down 
-the actual total time requirement to a few hours.
+4.975\ \mathit{days}\] 
+This in practice translates 
+to about $2.5\ \mathit{days}$, because successful searches stop before reaching 
+the timeout. Finally, by running this experiment on 20 cores, we reduce
+the actual wall clock time to several hours.
 

### 2018-09-03 tactictoe_jv.tex
minor tweaks

@@
 no 
-general rule on how to generate it. Instantiating 
+general rule to generate it. Instantiating 
 the term argument in the second tactic is easier because the expected term is a 
@@
 proof scripts from previous \coq proof examples.
-Its strength lies in the ability to produce likely sequences
+Its strength lies in the ability to produce probable sequences
 of tactics for solving domain specific goals. It operates by creating a model
@@
 \tactictoe is a general proof plan that learns to apply the right tactic from 
-examples and possess a backtracking mechanism given by the MCTS algorithm.
+examples and has a backtracking mechanism given by the MCTS algorithm.
 The philosophy of proof planning is also to create a few 
@@
 This can be done either in an interactive way, when the user completes the
-proof based on the recommended lemmas, as in the Mizar proof advisor
-\textsc{MizAR}~\cite{Urb04-MPTP0,mizAR40}, or attempted fully automatically, where such lemma
+proof based on the recommended lemmas, as in the Mizar Proof Advisor~\cite{Urb04-MPTP0}, or attempted fully automatically, where such lemma
 selection is handed over to the ATP component of a \emph{hammer}
 system, as in the hammers for \holfour~\cite{tgck-cpp15}, 
-\hollight~\cite{holyhammer} and \isabelle~\cite{BlanchetteGKKU16}.
+\hollight~\cite{holyhammer}, \isabelle~\cite{BlanchetteGKKU16} and \mizar~\cite{UrbanRS13,mizAR40}.
 
